### 2018-07-24 thesis/paper/paper.tex
Added paper and required files

\documentclass[hyperref,bachelorofscience]{cgvpub}
%weitere Optionen zum Ergänzen (in eckigen Klammern):
% 
% bibnum	numerische Literaturschlüssel
% final 	für Abgabe	
% lof			Abbildungsverzeichis
% lot			Tabellenverzeichnis
% noproblem	keine Aufgabenstellung
% notoc			kein Inhaltsverzeichnis
% twoside		zweiseitig
\author{Roland Raytracer}
\title{Titeltext}
\birthday{1. Januar 1234}
\placeofbirth{Dresden}
\matno{123456}
\betreuer{Dr. B. Spline}
\bibfiles{literatur}
\problem{Text der Aufgabenstellung...}
\copyrighterklaerung{Hier soll jeder Autor die von ihm eingeholten
Zustimmungen der Copyright-Besitzer angeben bzw. die in Web Press
Rooms angegebenen generellen Konditionen seiner Text- und
Bild"ubernahmen zitieren.}
\acknowledgments{Die Danksagung...}
\abstracten{abstract text english}
\abstractde{ Zusammenfassung Text Deutsch}
\usepackage{rotating}

\begin{document}
\chapter{Introduction/Preliminaries}
Motivation
\cite{test}
\section{Wikidata}
Wikidata is the knowledge base sister project of Wikipedia. It is a public, open database system that, at its core, stores statements about specific items. The basic unit of information is a triple subject-predicate-object stating an item (subject) has a property (predicate) which is either another item or a value of some kind (object). 
\chapter{Property Constraints}
To improve the data quality, help editors avoid common mistakes and clarify the usage of properties, Wikidata allows constraints to be defined on properties. They are however not firm rules and execptions can and should be made if necessary. A property constraint says something about the way a property may or may not be used. This can mean simple restrictions on the statement a property is part of, e.g. only certain values are allowed or more complex requirements on connected parts of the data graph, e.g. the value of a statement with this property should have certain statements. On the following page you can find table \ref{tab_basestats} which list the existing property constraints and a short description as well as classifying information which will be explained shortly. Afterwards the constraints will be explained in detail and either a basic approach for their modeling or reasoning why it is not possible or sensible to model it given.\\

To apply a constraint to a property a statement <property> <P2302> <constraint> is added to the properties item page, where <property> is the property that should be constrained, <P2302> (property constraint) denotes that the property has a constraint and <constraint> is the specific constraint. Note that a constrained statement from hereon denotes a statement whose predicate is a property that has a specific constraint. A simple example for this would be distinct values (Q21502410) stating that values of the property must be unique across all constrained statements. For this case no additional information regarding the constraint is necessary and violation could be found by surveying all constrained statements and comparing their values.\\
Constraints can be further specified by using qualifiers on the constraint statement. These qualifiers are either optional or mandatory and may be applied once or multiple times. For example, the none of constraint (Q52558054) states that values for a specific property may not be in a set of disallowed values. These are specified as qualifiers using item of property constraint (P2305) on the constraint statement and require at least one such statement.\\
More comlex dependencies are also possible. A constraint can depend on all constrained statements of a specific item (single value constraint [Q19474404]), on all statements of the value of a constrained statement (value requires statement constraint [Q21510864]) or in the most complex case on all statements using the subclassOf-predicate (type constraint [Q21503250]).
\newpage
\begin{table}[t]
\caption{property constraints and basic information}\label{tab_basestats}
\begin{tabular}{rcl}

\# of properties & constraint id & constraint name \\
\hline
4265 & Q21503247 & Item requires statement \\
3021 & Q21502404 & Format \\
2909 & Q19474404 & Single value \\
2843 & Q21502410 & Unique value \\
2613 & Q21503250 & Type \\
1410 & Q53869507 & Scope \\
781 & Q21502838 & Conflicts with \\
707 & Q21510865 & Value type \\
343 & Q21514353 & Allowed units \\
303 & Q21510851 & Allowed qualifiers \\
244 & Q21510864 & Value requires statement \\
203 & Q21510856 & Mandatory qualifier \\
188 & Q21510860 & Range \\
119 & Q21510855 & Inverse \\
116 & Q21510859 & One of \\
106 & Q52848401 & Integer \\
65 & Q51723761 & No bounds \\
53 & Q21510852 & Commons link \\
33 & Q21510862 & Symmetric \\
27 & Q52004125 & Allowed entity types \\
17 & Q52558054 & None of \\
13 & Q21510857 & Multi-value \\
10 & Q54554025 & Link \\
6 & Q21510854 & Difference within range \\
3 & Q52060874 & Single best value \\
1 & Q25796498 &  \\
1 & Q52712340 &  \\

\end{tabular}
\end{table}
\newpage

\chapter{Rules}
\chapter{Evaluation/Comparison}
\end{document}

### 2018-07-24 thesis/paper/paper.tex
Added a small piece on datalog

@@
 The basic unit of information is a triple subject-predicate-object stating an item (subject) has a property (predicate) which is either another item or a value of some kind (object).
+\section{Datalog}
+Datalog is a logic programming language.
+It is fully delarative and uses simple rules to derive additional knowledge from given facts.
+The most basic unit of information in datalog is a constant, which is a given [string]?
+expressing a fix point.
+In the context of Wikidata any entity or property, e.g.
+the universe (Q1) or head of government (P6), could be represented as a constant.
+Together with variables constants form terms.
+To form rules and define facts it is necessary to express their properties and relations.
+This is achieved using predicates
 \chapter{Property Constraints}

### 2018-07-25 thesis/paper/paper.tex
Expanded content

@@
 \abstractde{ Zusammenfassung Text Deutsch}
+
 \usepackage{rotating}
+\usepackage{float}
+\restylefloat{table}
+\usepackage{makecell}
+\renewcommand{\cellalign}{tl}
+\usepackage{tabularx}
 
@@
 The basic unit of information is a triple subject-predicate-object stating an item (subject) has a property (predicate) which is either another item or a value of some kind (object).
+These store the relations of entities with each other or with specific data.
+An entity is either an item, which represents a topic, concept or real world entity, or a property, which denotes a relation an item has to another or with a value.
+A value is one of several datatypes noted in table \ref{tab_datatypes}.
+\\
+
+\begin{table}
+\caption{wikidata datatypes}\label{tab_datatypes}
+\begin{tabularx}{\textwidth}{lllX}
+Datatype & Fields & Description \\
+\hline
+string & string (string) & simple string literal \\
+globecoordinate & \makecell{latitude (float) \\ longitude (float) \\ precision (float) \\ globe(URI)} & \makecell{coordinates on a given celestial body \\ or geographic standard} \\
+quantity & \makecell {amount (decimal) \\ upperBound (decimal) \\ lowerBound (decimal) \\ unit (URI or 1)} & quantity of a specified unit (or none) \\ with uncertainty \\
+time & \makecell {time (string) \\ calendarmodel (URI) \\ precision (integer)} & \makecell {time as a string according to calendarmodel \\ (julian or gregorian),\\ with precision mapping to orders of magnitude \\ (eg.
+10 years, months)} \\
+
+\end{tabularx}
+\end{table}
+
+While this is simliar to RDF and Wikidata has in fact been exported as RDF \cite{EGKMV2014}, the data structure is more complex than simple triples.
+It allows, in essence, to store additional information about statements.
+This is done via a statement id, which is assigend to every triple storing information about an item.
 \section{Datalog}
@@
 expressing a fix point.
-In the context of Wikidata any entity or property, e.g.
+In the context of Wikidata any item or property, e.g.
 the universe (Q1) or head of government (P6), could be represented as a constant.
-Together with variables constants form terms.
-To form rules and define facts it is necessary to express their properties and relations.
-This is achieved using predicates
+\section{Terminology}
+\begin{table}[H]
+\caption{shorthands}
+\begin{tabular}{ll}
+Shorthand & Description \\
+\hline
+constrained statement & statement with the constrained property as predicate \\
+item statement & statement on the same item as the constrained statement \\
+constrained item & item with a constrained statement
+\end{tabular}
+\end{table}
 \chapter{Property Constraints}
+
+%\begin{table}[t]
+%\caption{property constraints and basic information}\label{tab_basestats}
+%\begin{tabular}{rcl}
+
+%\# of properties & constraint id & constraint name \\
+%\hline
+%4265 & Q21503247 & Item requires statement \\
+%3021 & Q21502404 & Format \\
+%2909 & Q19474404 & Single value \\
+%2843 & Q21502410 & Unique value \\
+%2613 & Q21503250 & Type \\
+%1410 & Q53869507 & Scope \\
+%781 & Q21502838 & Conflicts with \\
+%707 & Q21510865 & Value type \\
+%343 & Q21514353 & Allowed units \\
+%303 & Q21510851 & Allowed qualifiers \\
+%244 & Q21510864 & Value requires statement \\
+%203 & Q21510856 & Mandatory qualifier \\
+%188 & Q21510860 & Range \\
+%119 & Q21510855 & Inverse \\
+%116 & Q21510859 & One of \\
+%106 & Q52848401 & Integer \\
+%65 & Q51723761 & No bounds \\
+%53 & Q21510852 & Commons link \\
+%33 & Q21510862 & Symmetric \\
+%27 & Q52004125 & Allowed entity types \\
+%17 & Q52558054 & None of \\
+%13 & Q21510857 & Multi-value \\
+%10 & Q54554025 & Link \\
+%6 & Q21510854 & Difference within range \\
+%3 & Q52060874 & Single best value \\
+%1 & Q25796498 &  \\
+%1 & Q52712340 &  \\
+%
+%\end{tabular}
+%\end{table}
 To improve the data quality, help editors avoid common mistakes and clarify the usage of properties, Wikidata allows constraints to be defined on properties.
@@
 A constraint can depend on all constrained statements of a specific item (single value constraint [Q19474404]), on all statements of the value of a constrained statement (value requires statement constraint [Q21510864]) or in the most complex case on all statements using the subclassOf-predicate (type constraint [Q21503250]).
+
+\section{Wikidata constraints explained}
+This section will explain all interesting, meaning translatable or partially translatable, currently used property constraints on Wikidata.
+They will be presented rougly in descending order of the number of properties with this constraint, although constraints with very similar function will be explained directly after one another.
+Also, all constraints that were considered untranslatable are listed at the and, each with a short explanation as to why they were placed in that group.\\
+
+Every constraint will start with a table showing basic properties of the constraint.
+This includes the requirement, relevant qualifiers, possible violations of the constraint and the data the constraint depends on.
+The requirement states which part of the data graph in relation to the constrained statement are restricted or need to exist.
+The qualifiers list their cardinality, meaning if they are required and if they may be used multiple times, and a short description.
+The violations list the conflicting (part of a) statement in relation to the constrained statement.
+The dependecies list their relation to the constrained statement.
+Table \ref{tab_example_constraint} shows the formatting as an example.
+
+\begin{table}[H]
+\caption{example constraint (Q31415926)}\label{tab_example_constraint}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Requirement & requirement \\
+\hline
+Qualifiers & \makecell{example qualifier (P1234) -- 1..* \\ description \\ another qualifier (P4321) -- 0..1 \\ different description} \\
+\hline
+Violation & \makecell{one way to violate the constraint \\ another violation possibility} \\
+\hline
+Dependency & statements, qualifiers or references in relation to the constrained statement\\
+\hline
+Rule features & features needed (missing if no features needed) \\
+\hline
+\end{tabularx}
+\end{table}
+
 \newpage
-\begin{table}[t]
-\caption{property constraints and basic information}\label{tab_basestats}
-\begin{tabular}{rcl}
+\subsection{Item requires statement (Q21503247)}\label{subsec_item_requires_statement}
+The item requires statement constraint requires an additional statement on an item with a statement using the constrained property, possibly also using a specific value from the list of allowed values.
+\begin{table}[H]
+\caption{item requires statement constraint (Q21503247)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Requirement & item statement \\
+\hline
+Qualifiers & \makecell{required property (P2306) -- 1 \\ property the item statement must have \\ allowed value (P2305) -- 0..* \\ values the item statement may have} \\
+\hline
+Violation & \makecell{no item statement with required property \\ all item statements with required property have no allowed value} \\
+\hline
+Dependency &  all statements of a constrained item\\
+\hline
+Rule features & negation \\
+\hline
+\end{tabularx}
+\end{table}
 
-\# of properties & constraint id & constraint name \\
+An item violates the constraint if it a) has a statement using constrained property as predicate, b1) has no statement using the required property as predicate, or (if allowed values are specified) b2) has only statement using required property as predicate that do not use one of the allowed values as value.\\
+While condition a) can easily be tested by a tripleEDB-atom with the constrained property, condition b1) and b2) pose the question if an item does not have a statement.
+This is adressed in section [?].
+
+\subsection{Value requires statement (Q21510864)}
+The value requires statement constraint requires an additional statement on the value of a statement using the constrained property, possibly also using a specific value from the list of allowed values.
+Since the constraint works very similar to item requires statement (see \ref{subsec_item_requires_statement}) --- requiring a statement on the value of the constrained statement instead of on the item --- it is not elaborated on.
+
+\subsection{Single value (Q19474404)}
+The single value constraint constrains the values a property may take on a single item.
+A property marked with the single value constraint can have properties marked as separators, which define exceptions under specific circumstances.
+\begin{table}[H]
+\caption{single value constraint (Q19474404)}
+\begin{tabularx}{\textwidth}{ ll X}
 \hline
-4265 & Q21503247 & Item requires statement \\
-3021 & Q21502404 & Format \\
-2909 & Q19474404 & Single value \\
-2843 & Q21502410 & Unique value \\
-2613 & Q21503250 & Type \\
-1410 & Q53869507 & Scope \\
-781 & Q21502838 & Conflicts with \\
-707 & Q21510865 & Value type \\
-343 & Q21514353 & Allowed units \\
-303 & Q21510851 & Allowed qualifiers \\
-244 & Q21510864 & Value requires statement \\
-203 & Q21510856 & Mandatory qualifier \\
-188 & Q21510860 & Range \\
-119 & Q21510855 & Inverse \\
-116 & Q21510859 & One of \\
-106 & Q52848401 & Integer \\
-65 & Q51723761 & No bounds \\
-53 & Q21510852 & Commons link \\
-33 & Q21510862 & Symmetric \\
-27 & Q52004125 & Allowed entity types \\
-17 & Q52558054 & None of \\
-13 & Q21510857 & Multi-value \\
-10 & Q54554025 & Link \\
-6 & Q21510854 & Difference within range \\
-3 & Q52060874 & Single best value \\
-1 & Q25796498 &  \\
-1 & Q52712340 &  \\
-
-\end{tabular}
+Requirement & constrained statement value \\
+\hline
+Qualifiers & \makecell{required property (P2306) -- 1 \\ property the item statement must have \\ allowed value (P2305) -- 0..* \\ values the item statement may have} \\
+\hline
+Violation & \makecell{no item statement with required property \\ all item statements with required property have no allowed value} \\
+\hline
+Dependency &  all statements of a constrained item\\
+\hline
+Rule features & negation \\
+\hline
+\end{tabularx}
 \end{table}
-\newpage
 

### 2018-07-25 thesis/paper/paper.tex
Added basic modelling and more constraints

@@
 
-\begin{table}
+\begin{table}[H]
 \caption{wikidata datatypes}\label{tab_datatypes}
@@
 the universe (Q1) or head of government (P6), could be represented as a constant.
+\section{Representation}
+To model the basic represenation of triples, qualifiers and references three predicates are introduced.\\
+tripleEDB (STATEMENT, ITEM, PROPERTY, VALUE) represents a single statement consisting of the statement ID, the item ID, the property ID and the value as either and item ID or a string-representation of the data value.\\
+qualifierEDB (STATEMENT, PROPERTY, VALUE) and referenceEDB (STATEMENT, PROPERTY, VALUE) represent the qualifiers and references on a given statement ID with their respective property id and the value as usual.\\
+
+Additional predicates modelling different preprocessing steps are explained in chapter \ref{cha_rules}.
 \section{Terminology}
@@
 On the following page you can find table \ref{tab_basestats} which list the existing property constraints and a short description as well as classifying information which will be explained shortly.
-Afterwards the constraints will be explained in detail and either a basic approach for their modeling or reasoning why it is not possible or sensible to model it given.\\
+Afterwards the constraints will be explained in detail and either a basic approach for their modelling or reasoning why it is not possible or sensible to model it given.\\
 
@@
 \hline
-Qualifiers & \makecell{required property (P2306) -- 1 \\ property the item statement must have \\ allowed value (P2305) -- 0..* \\ values the item statement may have} \\
+Qualifiers & \makecell{separator (P4155) -- 0..* \\ qualifiers that must be different for exception} \\
 \hline
-Violation & \makecell{no item statement with required property \\ all item statements with required property have no allowed value} \\
+Violation & \makecell{two constrained statements with same item and value \\ (exception: statements with separators as qualifiers with different values)} \\
 \hline
-Dependency &  all statements of a constrained item\\
+Dependency & \makecell{without separators: all constrained statements on one item \\ with separators: additionally all qualifiers with separators on those statements }\\
 \hline
-Rule features & negation \\
+Rule features & inequality, partially negation \\
 \hline
@@
 
-\chapter{Rules}
+If no separators are specified, two triple form a violation of this constraint if they a) belong to the same item, b) have the same constrained property, and c) have the same value.
+If separators are specified, a violation as above can be ignored if the statements each have at least one qualifier with the same separator as predicate and different values.\\
+In the case without separators the modelling idea would be to find two triples tripleEDB with the same item and constrained property and test if their id is unequal.
+If separators were specified a violation would be the two triples and all their separator qualifiers either equal or nonexistent.
+
+\subsection{Distinct values (Q21502410)}
+The distinct values constraint forbids the existence of two constrained statements with the same value.
+\begin{table}[H]
+\caption{distinct values constraint (Q21502410)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Requirement & constrained statement value \\
+\hline
+Violation & two constrained statements with same value \\
+\hline
+Dependency & all constrained statements\\
+\hline
+Rule features & inequality \\
+\hline
+\end{tabularx}
+\end{table}
+To model the distinct values constraint it is necessary to find triples tripleEDB
+
+\chapter{Rules}\label{cha_rules}
 \chapter{Evaluation/Comparison}

### 2018-07-25 thesis/paper/paper.tex
Added inverse and symmetric

@@
 
-\subsection{Value requires statement (Q21510864)}
-The value requires statement constraint requires an additional statement on the value of a statement using the constrained property, possibly also using a specific value from the list of allowed values.
+\subsection{Value requires statement (Q21510864)}\label{subsec_value_requires_statement}
+The value requires statement constraint requires an additional statement on the value of constrained statement, possibly also using a specific value from the list of allowed values.
 Since the constraint works very similar to item requires statement (see \ref{subsec_item_requires_statement}) --- requiring a statement on the value of the constrained statement instead of on the item --- it is not elaborated on.
+
+\subsection{Inverse (Q21510855)}\label{subsec_inverse}
+The inverse constraint requires an additional statement on the value of a constrained statement with a specific property and the constrained item as value.
+It is very similar to value requires statement (see \ref{subsec_value_requires_statement}), although it does not specify allowed values, since the required statement has one specific value: the constrained item.
+While this means it could not be modelled exactly as a value requires statement constraint, it would only require very small rule changes and will thus not be elaborated on.
+
+\subsection{Symmetric (Q21510862)}
+The symmetric constraint requires an additional statement on the value of a constrained statement with the constrained property and the constrained item as value.
+It is thus equal to an inverse constraint (see \ref{subsec_inverse}) with the same property as constrained property and required property and will not be elaborated on.
 
@@
 \end{table}
-To model the distinct values constraint it is necessary to find triples tripleEDB
+To model the distinct values constraint it is necessary to find triples tripleEDB with the constrained property and the same value whose statement id is unequal.
 

### 2018-07-25 thesis/paper/paper.tex
Added type

@@
 \hline
-Violation & two constrained statements with same value \\
+Violations & two constrained statements with same value \\
 \hline
@@
 
+\subsection{Type (Q21503250)}
+The type constraint requires the existence of either one of a specific statement (instanceOf <class>) or a chain of statements using subclassOf to connect directly (subclassOf* <class>)
+or via instanceOf (instanceOf/subclassOf* <class>) for all classes.
+\begin{table}[H]
+\caption{type constraint (Q21503250)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Requirement & chain of statements \\
+\hline
+Qualifiers & \makecell{relation (P2309) -- 1 \\ must be instance of (Q21503252), subclass of (Q21514624), \\ or instance or subclass of (Q30208840); \\ specifies the relation the constrained item must have to the classes\\
+class (P308) -- 1..* \\ entities that must be reached using the relation} \\
+\hline
+Violation & constrained item without relation to a class \\
+\hline
+Dependency & \makecell{statements of constrained items using instanceOf (P31) or subclassOf(P279) \\ all statements using subclassOf (P279) }\\
+\hline
+Rule features & inequality, negation \\
+\hline
+\end{tabularx}
+\end{table}
+
+If the relation is instanceOf, the constraint can be evaluated like multiple item requires statement constraints (see \ref{subsec_item_requires_statement}) by setting the required property to instanceOf (P31) and, for each created item requires statement constraint, the allowed value to one of the classes.
+\\
+For the relations involving subclasses all transivite subclass relationships would be concluded by setting a subclass relationship if a third item can be found that is subclass of one and subclassed by the other item.
+This would be extended with one step of instanceOf.
+A violation would then be any constrained item that is not subclass of any of the classes.
+
 \chapter{Rules}\label{cha_rules}

### 2018-07-25 thesis/paper/paper.tex
Added value type

@@
 This section will explain all interesting, meaning translatable or partially translatable, currently used property constraints on Wikidata.
-They will be presented rougly in descending order of the number of properties with this constraint, although constraints with very similar function will be explained directly after one another.
+They will be presented rougly in descending order of the number of properties with this constraint, although constraints with similar function will be explained directly after one another.
 Also, all constraints that were considered untranslatable are listed at the and, each with a short explanation as to why they were placed in that group.\\
@@
 
-\subsection{Type (Q21503250)}
+\subsection{Type (Q21503250)}\label{subsec_type}
 The type constraint requires the existence of either one of a specific statement (instanceOf <class>) or a chain of statements using subclassOf to connect directly (subclassOf* <class>)
-or via instanceOf (instanceOf/subclassOf* <class>) for all classes.
+or via instanceOf (instanceOf/subclassOf* <class>) for all classes on the constrained item.
 \begin{table}[H]
@@
 
+\subsection{Value type (Q21510865)}
+The value type constraint works like the type constraint (see \ref{subsec_type}) except the statement chains are required on the value of the constrained statement instead of the constrained item.
+It will thus not be expanded upon.
+
 \chapter{Rules}\label{cha_rules}

### 2018-07-25 thesis/paper/paper.tex
Added property scope

@@
 constrained statement & statement with the constrained property as predicate \\
+constrained qualifier & qualifier with the constrained property as predicate \\
+constrained reference & reference with the constrained property as predicate \\
+constrained item & item with a constrained statement \\
 item statement & statement on the same item as the constrained statement \\
-constrained item & item with a constrained statement
 \end{tabular}
@@
 
+\subsection{Property scope (Q53869507)}
+The property scope constraint specifies if a property may be used in a statement, qualifier or reference as predicate.
+\begin{table}[H]
+\caption{property scope constraint (Q53869507)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Requirement & constrained statement \\
+\hline
+Qualifiers & \makecell{property scope (P5314) -- 1..3 \\ must be as main value (Q54828448), as qualifier (Q54828449) or as reference (Q54828450) \\ specifies allowed use in statements, qualifiers or references} \\
+\hline
+Violation & constrained statement in a position that is not allowed \\
+\hline
+Dependency & constrained statement, qualifier or reference \\
+\hline
+\end{tabularx}
+\end{table}
+
+Since only three posibilities exist, this is equivalent to forbidding all other possibilities and the rules are simple mapping all of tripleEDB, qualifierEDB and referenceEDB with the constrained property that are not allowed.
+
 \chapter{Rules}\label{cha_rules}

### 2018-07-25 thesis/paper/paper.tex
Added conflicts-with

@@
 \hline
-Dependency &  all statements of a constrained item\\
+Dependency &  all item statements\\
 \hline
@@
 
+\subsection{Conflicts-with (Q21502838)}
+The conflicts with constraint forbids either certain statements or statements with a specific property on constrained items.
+\begin{table}[H]
+\caption{Conflicts-with constraint (Q21502838)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Requirement & item statement \\
+\hline
+Qualifiers & \makecell{conflicting property (P2306) -- 1 \\ property no item statement may have (with conflicting values if specified) \\
+conflicting values (P2305) -- 0..* \\ values that may not appear in item statements with the conflicting property} \\
+\hline
+Violation & \makecell{item statement with the conflicting property \\ item statement with the conflicting property and one of the conflicting values} \\
+\hline
+Dependency & all item statements \\
+\hline
+\end{tabularx}
+\end{table}
+The rule modelling can simply be done by choosing all tripleEDB with the constrained property and then all tripleEDB on the same item with the conflicting property, specifying the values as well if necessary.
+
 \chapter{Rules}\label{cha_rules}

### 2018-07-26 thesis/paper/paper.tex
Added wikidata preliminary

@@
 Motivation
-\cite{test}
+\newpage
 \section{Wikidata}
@@
 An entity is either an item, which represents a topic, concept or real world entity, or a property, which denotes a relation an item has to another or with a value.
-A value is one of several datatypes noted in table \ref{tab_datatypes}.
+They have numeric identifiers that are prefixed with Q and P respectively.
+A value is one of several datatypes noted in table \ref{tab_datatypes}, although it is also possible to use none or some as values for statements that have no value (akin to negation) or have a currently unknown value.
+Every property determines the datatype it accepts as an object.
 \\
@@
 globecoordinate & \makecell{latitude (float) \\ longitude (float) \\ precision (float) \\ globe(URI)} & \makecell{coordinates on a given celestial body \\ or geographic standard} \\
-quantity & \makecell {amount (decimal) \\ upperBound (decimal) \\ lowerBound (decimal) \\ unit (URI or 1)} & quantity of a specified unit (or none) \\ with uncertainty \\
+quantity & \makecell {amount (decimal) \\ upperBound (decimal) \\ lowerBound (decimal) \\ unit (URI or 1)} & \makecell{quantity of a specified unit (or none) \\ with uncertainty} \\
 time & \makecell {time (string) \\ calendarmodel (URI) \\ precision (integer)} & \makecell {time as a string according to calendarmodel \\ (julian or gregorian),\\ with precision mapping to orders of magnitude \\ (eg.
@@
 While this is simliar to RDF and Wikidata has in fact been exported as RDF \cite{EGKMV2014}, the data structure is more complex than simple triples.
-It allows, in essence, to store additional information about statements.
-This is done via a statement id, which is assigend to every triple storing information about an item.
+It allows, to store additional information about statements.
+This is done via a statement ID, which is assigend to every triple storing information about an item.
+Every statement can be references by qualifiers and references which contain a predicate and an object, but have no item as subject but instead the statement ID.
+Qualifiers provide contextual information for statements.
+The most common context is the timeframe in which the statement was accurate, e.g.
+stating the start and end times of a head of state, but other information is possible, e.g.
+determination method for measurements or the elevation above sea level for the highest point.
+References are information that supports the claim by citing the source and possibly additional information like when it was retrieved or when the source was published.
+
+Other information regarding items like labels and descriptions in different languages as well as aliases and site links to other Wikimedia projects are also part of the data model but are not subject to constraints and are thus not explored further.
 \section{Datalog}

### 2018-07-26 thesis/paper/paper.tex
Added allowed units

@@
 
+\subsection{Allowed units (Q21514353)}
+The allowed units constraint specifies that a constrained statement may only have a unit that is in a set of specific units.
+
+\begin{table}[H]
+\caption{Conflicts-with constraint (Q21502838)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Requirement & constrained statement unit \\
+\hline
+Qualifiers & \makecell{allowed unit (P2305) -- 1..* \\ units that the statement may have (or none for dimensionless units)} \\
+\hline
+Violation & \makecell{constrained statement with a unit unequal to all allowed units} \\
+\hline
+Dependency & constrained statement and its unit \\
+\hline
+Rule features & inequality \\
+\hline
+\end{tabularx}
+\end{table}
+
+Modelling the constraint requires a predicate associating the value or the statement with its unit and inequality between every unit and all allowed units.
+Then a violation can be found by finding all tripleEDB with the constrained property and a unit unequal to all allowed units.
+
 \chapter{Rules}\label{cha_rules}

### 2018-07-26 thesis/paper/paper.tex
Added allowed qualifiers

@@
 \begin{table}[H]
-\caption{Conflicts-with constraint (Q21502838)}
+\caption{Allowed units constraint (Q21514353)}
 \begin{tabularx}{\textwidth}{ ll X}
@@
 
+\subsection{Allowed qualifiers (Q21510851)}
+The allowed qualifiers constraint specifies that only a limited set of properties may be used as qualifiers on a constrained statement.
+
+\begin{table}[H]
+\caption{Allowed qualifiers constraint (Q21510851)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Requirement & qualifier on constrained statement \\
+\hline
+Qualifiers & \makecell{	allowed qualifier (P2306) -- 1..* \\ qualifiers that the statement may have} \\
+\hline
+Violation & \makecell{constrained statement with a qualifier unequal to all allowed qualifiers} \\
+\hline
+Dependency & constrained statement and its qualifiers \\
+\hline
+Rule features & inequality \\
+\hline
+\end{tabularx}
+\end{table}
+
+To represent a violation of this constraint in datalog inequality between all specified allowed qualifiers and the properties of the qualifiers on constrained statements is required.
+A tripleEDB with the constrained property and a qualifierEDB on that statement with a property different from all allowed qualifiers would violate the constraint.
+
 \chapter{Rules}\label{cha_rules}

### 2018-07-26 thesis/paper/paper.tex
Added mandatory qualifier

@@
 
+\subsection{Mandatory qualifier (Q21510856)}
+The mandatory qualifier constraint requires that a constrained statement has a qualifier with a specific property.
+
+\begin{table}[H]
+\caption{Mandatory qualifier constraint (Q21510856)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Requirement & qualifier on constrained statement \\
+\hline
+Qualifiers & \makecell{required property (P2306) -- 1 \\ qualifier that the statement must have} \\
+\hline
+Violation & \makecell{constrained statement without a qualifier with the required property} \\
+\hline
+Dependency & constrained statement and its qualifiers \\
+\hline
+Rule features & negation \\
+\hline
+\end{tabularx}
+\end{table}
+
+To find violations of this constraint one would find a tripleEDB with the constrained property, but where no qualifierEDB for the statement and with the required property exists.
+
 \chapter{Rules}\label{cha_rules}

### 2018-07-26 thesis/paper/paper.tex
Added one-of

@@
 
+\subsection{One-of (Q21510859)}
+The one-of constraint limits the values a property may use to a set of given values.
+
+\begin{table}[H]
+\caption{One-of constraint (Q21510859)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Requirement & constrained statement \\
+\hline
+Qualifiers & \makecell{allowed values (P2305) -- 1..* \\ values that may be used in the constrained statement} \\
+\hline
+Violation & \makecell{constrained statement with a value unequal to all allowed values} \\
+\hline
+Dependency & constrained statement \\
+\hline
+Rule features & inequality \\
+\hline
+\end{tabularx}
+\end{table}
+
+Modelling the one-of constraint requires a rule finding a tripleEDB with the constrained property whose value is unequal to all allowed values, meaning the inequality between each value of constrained statements and all allowed values must be given.
+
 \chapter{Rules}\label{cha_rules}

### 2018-07-26 thesis/paper/paper.tex
Added allowed entity types

@@
 
+\subsection{Allowed entity types (Q52004125)}
+The allowed entity types constraint limits if a property may be used on an item, a property or a lexeme, although lexemes are not yet deployed.
+
+\begin{table}[H]
+\caption{Allowed entity types constraint (Q52004125)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Requirement & constrained statement \\
+\hline
+Qualifiers & \makecell{allowed entity type (P2305) -- 1..3 \\ must be wikibase item (Q29934200), wikibase property (Q29934218) \\ or lexeme (Q51885771); \\ specifies the type the property may be used on} \\
+\hline
+Violation & \makecell{constrained statement on an entity of a type unequal to all allowed types} \\
+\hline
+Dependency & constrained statement and entity type of the subject \\
+\hline
+\end{tabularx}
+\end{table}
+
+Since there are only three options (and only two in use), violations can be found without inequality by finding the tripleEDB with the constrained property whose subject is of the type that is not allowed, meaning either wikibase item or wikibase property.
+
 \chapter{Rules}\label{cha_rules}

### 2018-07-26 thesis/paper/paper.tex
Added none of

@@
 
+\subsection{None of (Q52558054)}
+The none of constraint restricts the values that the property may refer to by listing a set of values as forbidden.
+
+\begin{table}[H]
+\caption{None of constraint (Q52558054)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Requirement & constrained statement \\
+\hline
+Qualifiers & \makecell{forbidden value (P2305) -- 1..* \\ values that may not appear in a constrained property} \\
+\hline
+Violation & \makecell{constrained statement with a forbidden value} \\
+\hline
+Dependency & constrained statement \\
+\hline
+\end{tabularx}
+\end{table}
+
+Modelling this constraint is very straightforward, since each forbidden value forms a property-value pair with the constrained statement that may not exists.
+Thus a violations is a tripleEDB with the constrained statement and a forbidden value.
+
 \chapter{Rules}\label{cha_rules}

### 2018-07-26 thesis/paper/paper.tex
Added multi-value

@@
 
+\subsection{Multi-value (Q21510857)}
+The multi-value constraint specifies that the constrained property should appear in multiple statements of an item or none.
+
+\begin{table}[H]
+\caption{Multi-value constraint (Q21510857)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Requirement & item statement \\
+\hline
+Violation & \makecell{constrained item with exactly one constrained statement} \\
+\hline
+Dependency & all item statements \\
+\hline
+Rule features & count \\
+\hline
+\end{tabularx}
+\end{table}
+
+To model this constraint a count of statements using the constrained property per item would be needed.
+Then a violation would be any constrained item with a count of one for the constrained property.
+Since there is no need for arbitrarily large counts, but only for the difference between one and two, this could be solved explicitly.
+
 \chapter{Rules}\label{cha_rules}

### 2018-07-26 thesis/paper/paper.tex
Added single best value

@@
 
+\subsection{Single best value (Q52060874)}
+The single best value constraint requires that only one constrained statement per item has the rank preferred.
+
+\begin{table}[H]
+\caption{Single best value constraint (Q52060874)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Requirement & item statement rank \\
+\hline
+Violation & \makecell{items with two constrained statements with rank preferred} \\
+\hline
+Dependency & all item statements and their rank \\
+\hline
+Rule features & count \\
+\hline
+\end{tabularx}
+\end{table}
+
+Modelling the constraint means a count of all constrained statements with rank preferred is necessary, with a violation constituted if an item has a count of two or higher.
+Since it is only needed to differentiate between one and many, the rules could be written explicitly.
+
 \chapter{Rules}\label{cha_rules}

### 2018-07-26 thesis/paper/paper.tex
Added one-of qualifier value

@@
 
+\subsection{One-of qualifier value (Q52712340)}
+The one-of qualifier value constraint restricts which values may be used on a qualifier with a given property on a constrained statement.
+
+\begin{table}[H]
+\caption{One-of qualifier value constraint (Q52712340)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Requirement & constrained statement qualifier value \\
+\hline
+Qualifiers: & \makecell{qualifier property (P2306) -- 1 \\ property of the restricted qualifier\\
+qualifier values (P2305) -- 1..* \\ values the qualifiers may take}\\
+\hline
+Violation & \makecell{constrained statement with a qualifier with the qualifier property and a value unequal to all qualifier values} \\
+\hline
+Dependency & constrained item and qualifiers with the qualifier property \\
+\hline
+Rule features & inequality \\
+\hline
+\end{tabularx}
+\end{table}
+
+The constraint can be modelled if the inequality between a value of a qualifier property with all qualifier values is given.
+The violations could then be found by taking a tripleEDB with the constrained property and a qualifier with the qualifier property on the former statement whose value is unequal to all qualifier values.
+
 \chapter{Rules}\label{cha_rules}

### 2018-07-26 thesis/paper/paper.tex
Added reasoning for unconsidered constraints

@@
 
+\section{Reasoning for unconsidered constraints}
+Multiple constraints have not been described above for different reasons.
+The most prominent of these is the format constraint (Q21502404) which specifies a regular expressions the values of the constrained property must adhere to.
+Evaluating these expressions would be theoretically possible but \\
+a) primarily results in a large amount of preprocessing and\\
+b) does not solve the question of negation, since violations would be values that do not conform to their required regular expression. \\
+
+The integer constraint (Q52848401), which requires values of a constrained property be integers (notably a quantity without decimal places), could easily be seen as a specific format constraint.
+While precomputing if a value is an integer or not could be done, it would mean that no actually relevant rules would be needed for the evaluation in datalog.
+
+Another problem that could not be sensibly solved is the comparison of values, needed for range (Q21510860), difference within range (Q21510854) and contemporary constraint (Q25796498).
+Range restricts the values of a constrained property within a certain range, difference within range specifies that two item statements should have values within a certain range and contemporary constraint specifies that subject and object of a constrained statement have to have coexisted at some point in time.
+While it would be possible to extract all values that might need to be compared and precompute the comparison, it would make most of the work precomputing and leave little to be done in datalog.\\
+
+Link constraint (Q54554025), requiring that constraint statements must have one or more references, requires a negation that cannot be concluded except by explicitly precomputing it, which makes the evaluation unnecessary.
+\\
+
+And finally commons link constraint (Q21510852) requires that values for a property should be valid links to to Wikimedia Commons.
+This validity cannot be testes from within datalog since there is no way to query the web for information.
+
 \chapter{Rules}\label{cha_rules}

### 2018-07-28 thesis/paper/paper.tex
Continued Datalog and Rules and fixed typos

@@
 % twoside		zweiseitig
-\author{Roland Raytracer}
-\title{Titeltext}
-\birthday{1. Januar 1234}
+\author{Adrian Bielefeldt}
+\title{Tit}le
+\birthday{15. October 1994}
 \placeofbirth{Dresden}
-\matno{123456}
-\betreuer{Dr. B. Spline}
+\matno{3694323}
+\betreuer{Prof. Dr. Markus Krötzsch}
 \bibfiles{literatur}
-\problem{Text der Aufgabenstellung...}
-\copyrighterklaerung{Hier soll jeder Autor die von ihm eingeholten
-Zustimmungen der Copyright-Besitzer angeben bzw. die in Web Press
-Rooms angegebenen generellen Konditionen seiner Text- und
-Bild"ubernahmen zitieren.}
-\acknowledgments{Die Danksagung...}
-\abstracten{abstract text english}
-\abstractde{ Zusammenfassung Text Deutsch}
+\problem{problem}
+\copyrighterklaerung{copyright stuff}
+\acknowledgments{Acknowledgements}
+\abstracten{abstract text English}
 
@@
 \usepackage{tabularx}
+\usepackage{todonotes}
 
@@
 They have numeric identifiers that are prefixed with Q and P respectively.
-A value is one of several datatypes noted in table \ref{tab_datatypes}, although it is also possible to use none or some as values for statements that have no value (akin to negation) or have a currently unknown value.
-Every property determines the datatype it accepts as an object.
+A value is one of several data types noted in table \ref{tab_datatypes}, although it is also possible to use none or some as values for statements that have no value (akin to negation) or have a currently unknown value.
+Every property determines the data type it accepts as an object.
 \\
@@
 \begin{table}[H]
-\caption{wikidata datatypes}\label{tab_datatypes}
+\caption{wikidata data types}\label{tab_datatypes}
 \begin{tabularx}{\textwidth}{lllX}
-Datatype & Fields & Description \\
+Data type & Fields & Description \\
 \hline
@@
 quantity & \makecell {amount (decimal) \\ upperBound (decimal) \\ lowerBound (decimal) \\ unit (URI or 1)} & \makecell{quantity of a specified unit (or none) \\ with uncertainty} \\
-time & \makecell {time (string) \\ calendarmodel (URI) \\ precision (integer)} & \makecell {time as a string according to calendarmodel \\ (julian or gregorian),\\ with precision mapping to orders of magnitude \\ (eg.
+time & \makecell {time (string) \\ calendarmodel (URI) \\ precision (integer)} & \makecell {time as a string according to calendarmodel \\ (julian or gregorian),\\ with precision mapping to orders of magnitude \\ (e.g..
 10 years, months)} \\
@@
 
-While this is simliar to RDF and Wikidata has in fact been exported as RDF \cite{EGKMV2014}, the data structure is more complex than simple triples.
+While this is similar to RDF and Wikidata has in fact been exported as RDF \cite{EGKMV2014}, the data structure is more complex than simple triples.
 It allows, to store additional information about statements.
-This is done via a statement ID, which is assigend to every triple storing information about an item.
+This is done via a statement ID, which is assigned to every triple storing information about an item.
 Every statement can be references by qualifiers and references which contain a predicate and an object, but have no item as subject but instead the statement ID.
 Qualifiers provide contextual information for statements.
-The most common context is the timeframe in which the statement was accurate, e.g.
+The most common context is the time frame in which the statement was accurate, e.g.
 stating the start and end times of a head of state, but other information is possible, e.g.
@@
 Datalog is a logic programming language.
-It is fully delarative and uses simple rules to derive additional knowledge from given facts.
-The most basic unit of information in datalog is a constant, which is a given [string]?
-expressing a fix point.
+It is fully declarative and uses simple rules to derive additional knowledge from given facts.
+The most basic unit of information in datalog is a \emph{constant}, which will be denoted as \(someConstant\).
 In the context of Wikidata any item or property, e.g.
-the universe (Q1) or head of government (P6), could be represented as a constant.
+Earth (Q2) or head of government (P6), could be represented as a constant.
+Relations in datalog are expressed as \emph{predicates} with a specified arity \(n\), written as predicate\([n]\).
+To stay close to Wikidata in our examples let us take the predicate statement\([3]\) as an example.
+This would be used to express the \emph{subject}-\emph{predicate}-\emph{object} relation with each position representing one part of the triple.\\
+Additionally, datalog also knows \emph{variables}, which will be capitalized like \(X\).
+Constants and variables are collectively known as \emph{terms} and are, together with predicates, used to form \emph{atoms}.
+An atom has the form predicate\((t_1, \ldots , t_n)\) with \(n\) as the arity of the predicate and \(t_1,\ldots,t_n\) terms.\\
+Atoms are the building blocks of \emph{rules}, which are used to derive information in datalog.
+Given atoms \(C\) and \(B_n\), a rule takes the form \(C \leftarrow P_1 \wedge \ldots \wedge P_m\), with \(C\) being the conclusion and \(B_1 \wedge \ldots \wedge B_m\) the premise.
+If the premise does not contain any atoms such a rule is called a \emph{fact}, or a \emph{ground rule} if it additionally only uses constants.
+A finite set of rules forms a \emph{program}.\\
+
+Within the rule predicates we can differentiate between two different schemas: the \emph{extensional schema} and the \emph{intensional schema}, called EDB and IDB \todo{Shouldn't this be called \\EDP/IDP?} (extensional/intensional database predicate).
+An EDB may only occur in the premise of a rule, meaning it can never be derived.
+All others (predicates that appear in the conclusion at least once) are IDB.
 \section{Representation}
-To model the basic represenation of triples, qualifiers and references three predicates are introduced.\\
-tripleEDB (STATEMENT, ITEM, PROPERTY, VALUE) represents a single statement consisting of the statement ID, the item ID, the property ID and the value as either and item ID or a string-representation of the data value.\\
-qualifierEDB (STATEMENT, PROPERTY, VALUE) and referenceEDB (STATEMENT, PROPERTY, VALUE) represent the qualifiers and references on a given statement ID with their respective property id and the value as usual.\\
+To model the basic representation of triples, qualifiers and references three predicates are introduced.\\
+tripleEDB\([4]\) represents a single statement consisting of the statement ID, the item ID, the property ID and the value as either and item ID or a string-representation of the data value.\\
+qualifierEDB\([3]\) and referenceEDB\([3]\) represent the qualifiers and references on a given statement ID with their respective property id and the value as usual.\\
 
-Additional predicates modelling different preprocessing steps are explained in chapter \ref{cha_rules}.
+Additional predicates modelling different pre-processing steps are explained in chapter \ref{cha_rules}.
 \section{Terminology}
 \begin{table}[H]
-\caption{shorthands}
+\caption{abbreviations}
 \begin{tabular}{ll}
@@
 To improve the data quality, help editors avoid common mistakes and clarify the usage of properties, Wikidata allows constraints to be defined on properties.
-They are however not firm rules and execptions can and should be made if necessary.
+They are however not firm rules and exceptions can and should be made if necessary.
 A property constraint says something about the way a property may or may not be used.
@@
 To apply a constraint to a property a statement <property> <P2302> <constraint> is added to the properties item page, where <property> is the property that should be constrained, <P2302> (property constraint) denotes that the property has a constraint and <constraint> is the specific constraint.
-Note that a constrained statement from hereon denotes a statement whose predicate is a property that has a specific constraint.
+Note that a constrained statement means a statement whose predicate is a property that has a specific constraint.
 A simple example for this would be distinct values (Q21502410) stating that values of the property must be unique across all constrained statements.
@@
 These are specified as qualifiers using item of property constraint (P2305) on the constraint statement and require at least one such statement.\\
-More comlex dependencies are also possible.
+More complex dependencies are also possible.
 A constraint can depend on all constrained statements of a specific item (single value constraint [Q19474404]), on all statements of the value of a constrained statement (value requires statement constraint [Q21510864]) or in the most complex case on all statements using the subclassOf-predicate (type constraint [Q21503250]).
@@
 This section will explain all interesting, meaning translatable or partially translatable, currently used property constraints on Wikidata.
-They will be presented rougly in descending order of the number of properties with this constraint, although constraints with similar function will be explained directly after one another.
+They will be presented roughly in descending order of the number of properties with this constraint, although constraints with similar function will be explained directly after one another.
 Also, all constraints that were considered untranslatable are listed at the and, each with a short explanation as to why they were placed in that group.\\
@@
 The violations list the conflicting (part of a) statement in relation to the constrained statement.
-The dependecies list their relation to the constrained statement.
+The dependencies list their relation to the constrained statement.
 Table \ref{tab_example_constraint} shows the formatting as an example.
@@
 While condition a) can easily be tested by a tripleEDB-atom with the constrained property, condition b1) and b2) pose the question if an item does not have a statement.
-This is adressed in section [?].
+This is addressed in section [?].
 
@@
 In the case without separators the modelling idea would be to find two triples tripleEDB with the same item and constrained property and test if their id is unequal.
-If separators were specified a violation would be the two triples and all their separator qualifiers either equal or nonexistent.
+If separators were specified a violation would be the two triples and all their separator qualifiers either equal or non-existent.
 
@@
 Qualifiers & \makecell{relation (P2309) -- 1 \\ must be instance of (Q21503252), subclass of (Q21514624), \\ or instance or subclass of (Q30208840); \\ specifies the relation the constrained item must have to the classes\\
-class (P308) -- 1..* \\ entities that must be reached using the relation} \\
+class (P2308) -- 1..* \\ entities that must be reached using the relation} \\
 \hline
@@
 \\
-For the relations involving subclasses all transivite subclass relationships would be concluded by setting a subclass relationship if a third item can be found that is subclass of one and subclassed by the other item.
+For the relations involving subclasses all transitive subclass relationships would be concluded by setting a subclass relationship if a third item can be found that is subclass of one and subclassed by the other item.
 This would be extended with one step of instanceOf.
@@
 
-Since only three posibilities exist, this is equivalent to forbidding all other possibilities and the rules are simple mapping all of tripleEDB, qualifierEDB and referenceEDB with the constrained property that are not allowed.
+Since only three possibilities exist, this is equivalent to forbidding all other possibilities and the rules are simple mapping all of tripleEDB, qualifierEDB and referenceEDB with the constrained property that are not allowed.
 
@@
 \hline
-Rule features & count \\
+Rule features & inequality \\
 \hline
@@
 While precomputing if a value is an integer or not could be done, it would mean that no actually relevant rules would be needed for the evaluation in datalog.
+\\
 
@@
 \\
+A similar situation is the no bounds constraint (Q51723761), which requires that quantity values have neither upper nor lower bound.
+While it would be possible to precompute if a statement has no upper or lower bound, this would shift basically all evaluation logic into precomputing.
 
@@
 \chapter{Rules}\label{cha_rules}
+The previous chapter has show that most constraints cannot be modelled without at least some amount of preprocessing.
+Only Scope, Conflicts with, Allowed entity types and None of can be modelled directly.
+This is due to the fact that Conflicts with and None of exclude a limited and specified set of statements from validity and these cases can thus be directly takes as violations.
+Scope and Allowed entity types work very similar, since there are only three respectively two possible values and violations are simply the options that have not been allowed.
+The rules for these constraints are described in section \ref{sec_no_preprocessing}.\\
+
+A problem that occurs multiple times is the need to determine inequality between constants.
+Distinct value, Allowed units, Allowed qualifiers, One of, Single best value and One-of qualifier value each require a rule to determine if a value, unit or property is not equal to another.
+To this end a certain amount of preprocessing is necessary, most notably the extraction of all constant pairs whose inequality has to be determined for finding the violations.
+To avoid putting and undue amount of processing outside of datalog this extraction will not check more than that the constant is in the right position relative to the constrained statement, e.g.
+all values on statements constrained by Distinct value or all qualifiers on statements constrained by Allowed qualifiers.
+From this set inequality will be generated, different versions of which will be introduced in section \ref{sec_inequality} and evaluated with regards to their performance in section \ref{sec_inequality_evaluation}.\\
+
+Negation as a general problem cannot be solved in datalog.
+One could pre-process every necessary fact to be negated and add this to the ground facts, but this is not the purpose of this work.
+However, not all constraints require a general negation.
+Item requires statement, Value requires statement, Inverse and Symmetric are all violated if a statement does not exist on a specific item.
+This problem can be solved with limited amounts of preprocessing by leveraging a preprocessed statement order per item and 'iterating' through the statements until the required conditions are met.\\
+A similar approach can be taken to find items with exactly one statement matching certain criteria , required for 	Multi-value.
+The idea is to ask the question if a second statement exists, then evaluating it as above.
+Both methods will be explained in section \ref{sec_statement_nonexistence}.
+
+\section{Task}
+Each property constraint will be modelled and evaluated as one set of rules per property per constraint, meaning each violating statement will be of a form similar to
+\begin{verbatim}
+violation(ID, ITEM, propertyConstant, VALUE) :-
+    <rules determining the violation>
+\end{verbatim}
+where \texttt{propertyConstant} denotes the constrained property. While it would be possible to load all data at once and restrict the property constant with a separate predicate akin to
+\begin{verbatim}
+violation(ID, ITEM, PROPERTY, VALUE) :-
+    constrained(PROPERTY, constraintConstant),
+    <rules determining the violation>
+\end{verbatim}
+this would increase the amount of facts loaded significantly while almost not reducing preprocessing.\\
+
+The task for each constraint is thus: Given the constraint and a constrained property, potentially with qualifiers, which rule finds all violations of the rule in a given set of triples, qualifiers and references if applicable?
+The necessary preprocessing and filtering will be listed and the amount of rules necessary to model the constrained depending on the number of qualifiers given.
+
+\section{No preprocessing}\label{sec_no_preprocessing}
+
+\section{Inequality}\label{sec_inequality}
+
+\section{Statement nonexistence}\label{sec_statement_nonexistence}
+
 \chapter{Evaluation/Comparison}
+
+\section{Inequality evaluation}\label{sec_inequality_evaluation}
 \end{document}

### 2018-07-28 thesis/paper/paper.tex
FIX: wrong bracket
Added example to datalog

@@
 \author{Adrian Bielefeldt}
-\title{Tit}le
+\title{Title}
 \birthday{15. October 1994}
@@
 It is fully declarative and uses simple rules to derive additional knowledge from given facts.
-The most basic unit of information in datalog is a \emph{constant}, which will be denoted as \(someConstant\).
+The most basic unit of information in datalog is a \emph{constant}, which will be denoted as someConstant.
 In the context of Wikidata any item or property, e.g.
@@
 To stay close to Wikidata in our examples let us take the predicate statement\([3]\) as an example.
-This would be used to express the \emph{subject}-\emph{predicate}-\emph{object} relation with each position representing one part of the triple.\\
+This would be used to express the \emph{item}-\emph{property}-\emph{value} relation with each position representing one part of the triple.\\
 Additionally, datalog also knows \emph{variables}, which will be capitalized like \(X\).
@@
 All others (predicates that appear in the conclusion at least once) are IDB.
+
+\todo[inline]{Explain the relation to FO-logic?}
+
+For the following example taken from Wikidata descriptive names will be used instead of the IDs to increase readability.
+The individual entities and properties are introduced with their ID beforehand so the example can be retraced.
+The blue whales (Q42196) parent taxon (P171) is balaenoptera (Q133320), which in turn has the parent taxon rorqual (Q232829).
+The predicates used are tripleEDB\([3]\) as the input and isInTaxonIDB\([2]\), which will be derived.
+The basic facts are:\\
+\[\text{tripleEDB}(\text{blue whale}, \text{parent taxon}, \text{balaenoptera})\]
+\[\text{tripleEDB}(\text{balaenoptera}, \text{parent taxon}, \text{rorqual})\]
+From this we can draw conclusions. The first step is to represent the already existing facts as IDB:
+\[\text{isInTaxonIDB}(X, Y) \leftarrow \text{tripleEDB}(X, \text{parent taxon}, Y)\]
+Finally the transitive closure of this relation can be derived with
+\[\text{isInTaxonIDB}(X, Y) \leftarrow \text{isInTaxonIDB}(X, Z) \wedge \text{isInTaxonIDB}(Z, Y)\]
+
+\todo[inline]{Probably a good place to explain existential and universal quantifiers.}
 \section{Representation}

### 2018-07-28 thesis/paper/paper.tex
Added table and description

@@
 \usepackage{todonotes}
+\usepackage{adjustbox}
 
@@
 
-%\begin{table}[t]
-%\caption{property constraints and basic information}\label{tab_basestats}
-%\begin{tabular}{rcl}
-
-%\# of properties & constraint id & constraint name \\
-%\hline
-%4265 & Q21503247 & Item requires statement \\
-%3021 & Q21502404 & Format \\
-%2909 & Q19474404 & Single value \\
-%2843 & Q21502410 & Unique value \\
-%2613 & Q21503250 & Type \\
-%1410 & Q53869507 & Scope \\
-%781 & Q21502838 & Conflicts with \\
-%707 & Q21510865 & Value type \\
-%343 & Q21514353 & Allowed units \\
-%303 & Q21510851 & Allowed qualifiers \\
-%244 & Q21510864 & Value requires statement \\
-%203 & Q21510856 & Mandatory qualifier \\
-%188 & Q21510860 & Range \\
-%119 & Q21510855 & Inverse \\
-%116 & Q21510859 & One of \\
-%106 & Q52848401 & Integer \\
-%65 & Q51723761 & No bounds \\
-%53 & Q21510852 & Commons link \\
-%33 & Q21510862 & Symmetric \\
-%27 & Q52004125 & Allowed entity types \\
-%17 & Q52558054 & None of \\
-%13 & Q21510857 & Multi-value \\
-%10 & Q54554025 & Link \\
-%6 & Q21510854 & Difference within range \\
-%3 & Q52060874 & Single best value \\
-%1 & Q25796498 &  \\
-%1 & Q52712340 &  \\
-%
-%\end{tabular}
-%\end{table}
 To improve the data quality, help editors avoid common mistakes and clarify the usage of properties, Wikidata allows constraints to be defined on properties.
@@
 the value of a statement with this property should have certain statements.
-On the following page you can find table \ref{tab_basestats} which list the existing property constraints and a short description as well as classifying information which will be explained shortly.
+Further down you can find table \ref{tab_property_constraints}, which list the existing property constraints.
 Afterwards the constraints will be explained in detail and either a basic approach for their modelling or reasoning why it is not possible or sensible to model it given.\\
@@
 A constraint can depend on all constrained statements of a specific item (single value constraint [Q19474404]), on all statements of the value of a constrained statement (value requires statement constraint [Q21510864]) or in the most complex case on all statements using the subclassOf-predicate (type constraint [Q21503250]).
+
+Table \ref{tab_property_constraints} shows all currently used property constrains in Wikidata.
+It is sorted by difficulty, which is an estimate of the necessary pre-processing and rule complexity based on required features.
+These required features are conditions that cannot natively be expressed in datalog and would require some kind of pre-processing to be made available.
+Difficulty 0 does not need any pre-processing except for loading the data into datalog as ground rules.
+1 requires inequality, which can be easily solved with brute force but will be improved in chapter \ref{cha_rules}.
+2 needs negation in a form that can be solved without pre-processing everything and 3 in a form that would require all logic as part of pre-processing.
+Other features needed for category three are regular expressions, which could theoretically be translated but would require large amounts of pre-processing, value comparison, general negation and finally web requests to test if a link is valid.
+
+\begin{table}[H]
+\caption{property constraints}\label{tab_property_constraints}
+\begin{tabularx}{\textheight}{rrllllX}
+Difficulty & \makecell{\# constrained\\ properties} & Identifier & Name & Required Features \\
+\hline
+0 & 1410 & Q53869507 & Scope &  \\
+0 & 781 & Q21502838 & Conflicts with &  \\
+0 & 27 & Q52004125 & Allowed entity types &  \\
+0 & 17 & Q52558054 & None of &  \\
+1 & 2843 & Q21502410 & Distinct value & Inequality \\
+1 & 343 & Q21514353 & Allowed units & Inequality \\
+1 & 303 & Q21510851 & Allowed qualifiers & Inequality \\
+1 & 116 & Q21510859 & One of & Inequality \\
+1 & 3 & Q52060874 & Single best value & Inequality \\
+1 & 1 & Q52712340 & One-of qualifier value & Inequality \\
+2 & 4265 & Q21503247 & Item requires statement & Negation (statement non-existence) \\
+2 & 244 & Q21510864 & Value requires statement & Negation (statement non-existence) \\
+2 & 119 & Q21510855 & Inverse & Negation (statement non-existence) \\
+2 & 33 & Q21510862 & Symmetric & Negation (statement non-existence) \\
+2 & 13 & Q21510857 & Multi-value & Negation, Count (up to two) \\
+2 & 2909 & Q19474404 & Single value & Inequality, partially negation (qualifier non-existence) \\
+2 & 203 & Q21510856 & Mandatory qualifier & Negation (qualifier non-existence) \\
+3 & 2613 & Q21503250 & Type & Negation (derived fact negation) \\
+3 & 707 & Q21510865 & Value type & Negation (derived fact negation) \\
+3 & 3021 & Q21502404 & Format & Regular expressions \\
+3 & 188 & Q21510860 & Range & Value comparison \\
+3 & 106 & Q52848401 & Integer & Regular expressions \\
+3 & 65 & Q51723761 & No bounds & Negation \\
+3 & 53 & Q21510852 & Commons link & Web request \\
+3 & 10 & Q54554025 & Link & Negation \\
+3 & 6 & Q21510854 & Difference within range & Value comparison \\
+3 & 1 & Q25796498 & Contemporary constraint & Negation \\
+\end{tabularx}
+\end{table}
 

### 2018-07-28 thesis/paper/paper.tex
Sorted constraints in ch2 by difficulty
Switched verbatim to equation

@@
-\documentclass[hyperref,bachelorofscience]{cgvpub}
+\documentclass[hyperref,bachelorofscience,fleqn]{cgvpub}
 %weitere Optionen zum Ergänzen (in eckigen Klammern):
@@
 
-\newpage
-\subsection{Item requires statement (Q21503247)}\label{subsec_item_requires_statement}
-The item requires statement constraint requires an additional statement on an item with a statement using the constrained property, possibly also using a specific value from the list of allowed values.
-\begin{table}[H]
-\caption{item requires statement constraint (Q21503247)}
-\begin{tabularx}{\textwidth}{ ll X}
-\hline
-Requirement & item statement \\
-\hline
-Qualifiers & \makecell{required property (P2306) -- 1 \\ property the item statement must have \\ allowed value (P2305) -- 0..* \\ values the item statement may have} \\
-\hline
-Violation & \makecell{no item statement with required property \\ all item statements with required property have no allowed value} \\
-\hline
-Dependency &  all item statements\\
-\hline
-Rule features & negation \\
-\hline
-\end{tabularx}
-\end{table}
-
-An item violates the constraint if it a) has a statement using constrained property as predicate, b1) has no statement using the required property as predicate, or (if allowed values are specified) b2) has only statement using required property as predicate that do not use one of the allowed values as value.\\
-While condition a) can easily be tested by a tripleEDB-atom with the constrained property, condition b1) and b2) pose the question if an item does not have a statement.
-This is addressed in section [?].
-
-\subsection{Value requires statement (Q21510864)}\label{subsec_value_requires_statement}
-The value requires statement constraint requires an additional statement on the value of constrained statement, possibly also using a specific value from the list of allowed values.
-Since the constraint works very similar to item requires statement (see \ref{subsec_item_requires_statement}) --- requiring a statement on the value of the constrained statement instead of on the item --- it is not elaborated on.
-
-\subsection{Inverse (Q21510855)}\label{subsec_inverse}
-The inverse constraint requires an additional statement on the value of a constrained statement with a specific property and the constrained item as value.
-It is very similar to value requires statement (see \ref{subsec_value_requires_statement}), although it does not specify allowed values, since the required statement has one specific value: the constrained item.
-While this means it could not be modelled exactly as a value requires statement constraint, it would only require very small rule changes and will thus not be elaborated on.
-
-\subsection{Symmetric (Q21510862)}
-The symmetric constraint requires an additional statement on the value of a constrained statement with the constrained property and the constrained item as value.
-It is thus equal to an inverse constraint (see \ref{subsec_inverse}) with the same property as constrained property and required property and will not be elaborated on.
-
-\subsection{Single value (Q19474404)}
-The single value constraint constrains the values a property may take on a single item.
-A property marked with the single value constraint can have properties marked as separators, which define exceptions under specific circumstances.
-\begin{table}[H]
-\caption{single value constraint (Q19474404)}
-\begin{tabularx}{\textwidth}{ ll X}
-\hline
-Requirement & constrained statement value \\
-\hline
-Qualifiers & \makecell{separator (P4155) -- 0..* \\ qualifiers that must be different for exception} \\
-\hline
-Violation & \makecell{two constrained statements with same item and value \\ (exception: statements with separators as qualifiers with different values)} \\
-\hline
-Dependency & \makecell{without separators: all constrained statements on one item \\ with separators: additionally all qualifiers with separators on those statements }\\
-\hline
-Rule features & inequality, partially negation \\
-\hline
-\end{tabularx}
-\end{table}
-
-If no separators are specified, two triple form a violation of this constraint if they a) belong to the same item, b) have the same constrained property, and c) have the same value.
-If separators are specified, a violation as above can be ignored if the statements each have at least one qualifier with the same separator as predicate and different values.\\
-In the case without separators the modelling idea would be to find two triples tripleEDB with the same item and constrained property and test if their id is unequal.
-If separators were specified a violation would be the two triples and all their separator qualifiers either equal or non-existent.
-
-\subsection{Distinct values (Q21502410)}
-The distinct values constraint forbids the existence of two constrained statements with the same value.
-\begin{table}[H]
-\caption{distinct values constraint (Q21502410)}
-\begin{tabularx}{\textwidth}{ ll X}
-\hline
-Requirement & constrained statement value \\
-\hline
-Violations & two constrained statements with same value \\
-\hline
-Dependency & all constrained statements\\
-\hline
-Rule features & inequality \\
-\hline
-\end{tabularx}
-\end{table}
-To model the distinct values constraint it is necessary to find triples tripleEDB with the constrained property and the same value whose statement id is unequal.
-
-\subsection{Type (Q21503250)}\label{subsec_type}
-The type constraint requires the existence of either one of a specific statement (instanceOf <class>) or a chain of statements using subclassOf to connect directly (subclassOf* <class>)
-or via instanceOf (instanceOf/subclassOf* <class>) for all classes on the constrained item.
-\begin{table}[H]
-\caption{type constraint (Q21503250)}
-\begin{tabularx}{\textwidth}{ ll X}
-\hline
-Requirement & chain of statements \\
-\hline
-Qualifiers & \makecell{relation (P2309) -- 1 \\ must be instance of (Q21503252), subclass of (Q21514624), \\ or instance or subclass of (Q30208840); \\ specifies the relation the constrained item must have to the classes\\
-class (P2308) -- 1..* \\ entities that must be reached using the relation} \\
-\hline
-Violation & constrained item without relation to a class \\
-\hline
-Dependency & \makecell{statements of constrained items using instanceOf (P31) or subclassOf(P279) \\ all statements using subclassOf (P279) }\\
-\hline
-Rule features & inequality, negation \\
-\hline
-\end{tabularx}
-\end{table}
-
-If the relation is instanceOf, the constraint can be evaluated like multiple item requires statement constraints (see \ref{subsec_item_requires_statement}) by setting the required property to instanceOf (P31) and, for each created item requires statement constraint, the allowed value to one of the classes.
-\\
-For the relations involving subclasses all transitive subclass relationships would be concluded by setting a subclass relationship if a third item can be found that is subclass of one and subclassed by the other item.
-This would be extended with one step of instanceOf.
-A violation would then be any constrained item that is not subclass of any of the classes.
-
-\subsection{Value type (Q21510865)}
-The value type constraint works like the type constraint (see \ref{subsec_type}) except the statement chains are required on the value of the constrained statement instead of the constrained item.
-It will thus not be expanded upon.
-
 \subsection{Property scope (Q53869507)}
@@
 The rule modelling can simply be done by choosing all tripleEDB with the constrained property and then all tripleEDB on the same item with the conflicting property, specifying the values as well if necessary.
+
+\subsection{Allowed entity types (Q52004125)}
+The allowed entity types constraint limits if a property may be used on an item, a property or a lexeme, although lexemes are not yet deployed.
+
+\begin{table}[H]
+\caption{Allowed entity types constraint (Q52004125)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Requirement & constrained statement \\
+\hline
+Qualifiers & \makecell{allowed entity type (P2305) -- 1..3 \\ must be wikibase item (Q29934200), wikibase property (Q29934218) \\ or lexeme (Q51885771); \\ specifies the type the property may be used on} \\
+\hline
+Violation & \makecell{constrained statement on an entity of a type unequal to all allowed types} \\
+\hline
+Dependency & constrained statement and entity type of the subject \\
+\hline
+\end{tabularx}
+\end{table}
+
+Since there are only three options (and only two in use), violations can be found without inequality by finding the tripleEDB with the constrained property whose subject is of the type that is not allowed, meaning either wikibase item or wikibase property.
+
+\subsection{None of (Q52558054)}
+The none of constraint restricts the values that the property may refer to by listing a set of values as forbidden.
+
+\begin{table}[H]
+\caption{None of constraint (Q52558054)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Requirement & constrained statement \\
+\hline
+Qualifiers & \makecell{forbidden value (P2305) -- 1..* \\ values that may not appear in a constrained property} \\
+\hline
+Violation & \makecell{constrained statement with a forbidden value} \\
+\hline
+Dependency & constrained statement \\
+\hline
+\end{tabularx}
+\end{table}
+
+Modelling this constraint is very straightforward, since each forbidden value forms a property-value pair with the constrained statement that may not exists.
+Thus a violations is a tripleEDB with the constrained statement and a forbidden value.
+
+\subsection{Distinct values (Q21502410)}
+The distinct values constraint forbids the existence of two constrained statements with the same value.
+\begin{table}[H]
+\caption{distinct values constraint (Q21502410)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Requirement & constrained statement value \\
+\hline
+Violations & two constrained statements with same value \\
+\hline
+Dependency & all constrained statements\\
+\hline
+Rule features & inequality \\
+\hline
+\end{tabularx}
+\end{table}
+To model the distinct values constraint it is necessary to find triples tripleEDB with the constrained property and the same value whose statement id is unequal.
 
@@
 
-\subsection{Mandatory qualifier (Q21510856)}
-The mandatory qualifier constraint requires that a constrained statement has a qualifier with a specific property.
-
-\begin{table}[H]
-\caption{Mandatory qualifier constraint (Q21510856)}
-\begin{tabularx}{\textwidth}{ ll X}
-\hline
-Requirement & qualifier on constrained statement \\
-\hline
-Qualifiers & \makecell{required property (P2306) -- 1 \\ qualifier that the statement must have} \\
-\hline
-Violation & \makecell{constrained statement without a qualifier with the required property} \\
-\hline
-Dependency & constrained statement and its qualifiers \\
-\hline
-Rule features & negation \\
-\hline
-\end{tabularx}
-\end{table}
-
-To find violations of this constraint one would find a tripleEDB with the constrained property, but where no qualifierEDB for the statement and with the required property exists.
-
 \subsection{One-of (Q21510859)}
@@
 Modelling the one-of constraint requires a rule finding a tripleEDB with the constrained property whose value is unequal to all allowed values, meaning the inequality between each value of constrained statements and all allowed values must be given.
-
-\subsection{Allowed entity types (Q52004125)}
-The allowed entity types constraint limits if a property may be used on an item, a property or a lexeme, although lexemes are not yet deployed.
-
-\begin{table}[H]
-\caption{Allowed entity types constraint (Q52004125)}
-\begin{tabularx}{\textwidth}{ ll X}
-\hline
-Requirement & constrained statement \\
-\hline
-Qualifiers & \makecell{allowed entity type (P2305) -- 1..3 \\ must be wikibase item (Q29934200), wikibase property (Q29934218) \\ or lexeme (Q51885771); \\ specifies the type the property may be used on} \\
-\hline
-Violation & \makecell{constrained statement on an entity of a type unequal to all allowed types} \\
-\hline
-Dependency & constrained statement and entity type of the subject \\
-\hline
-\end{tabularx}
-\end{table}
-
-Since there are only three options (and only two in use), violations can be found without inequality by finding the tripleEDB with the constrained property whose subject is of the type that is not allowed, meaning either wikibase item or wikibase property.
-
-\subsection{None of (Q52558054)}
-The none of constraint restricts the values that the property may refer to by listing a set of values as forbidden.
-
-\begin{table}[H]
-\caption{None of constraint (Q52558054)}
-\begin{tabularx}{\textwidth}{ ll X}
-\hline
-Requirement & constrained statement \\
-\hline
-Qualifiers & \makecell{forbidden value (P2305) -- 1..* \\ values that may not appear in a constrained property} \\
-\hline
-Violation & \makecell{constrained statement with a forbidden value} \\
-\hline
-Dependency & constrained statement \\
-\hline
-\end{tabularx}
-\end{table}
-
-Modelling this constraint is very straightforward, since each forbidden value forms a property-value pair with the constrained statement that may not exists.
-Thus a violations is a tripleEDB with the constrained statement and a forbidden value.
-
-\subsection{Multi-value (Q21510857)}
-The multi-value constraint specifies that the constrained property should appear in multiple statements of an item or none.
-
-\begin{table}[H]
-\caption{Multi-value constraint (Q21510857)}
-\begin{tabularx}{\textwidth}{ ll X}
-\hline
-Requirement & item statement \\
-\hline
-Violation & \makecell{constrained item with exactly one constrained statement} \\
-\hline
-Dependency & all item statements \\
-\hline
-Rule features & count \\
-\hline
-\end{tabularx}
-\end{table}
-
-To model this constraint a count of statements using the constrained property per item would be needed.
-Then a violation would be any constrained item with a count of one for the constrained property.
-Since there is no need for arbitrarily large counts, but only for the difference between one and two, this could be solved explicitly.
 
@@
 
+\subsection{Item requires statement (Q21503247)}\label{subsec_item_requires_statement}
+The item requires statement constraint requires an additional statement on an item with a statement using the constrained property, possibly also using a specific value from the list of allowed values.
+\begin{table}[H]
+\caption{item requires statement constraint (Q21503247)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Requirement & item statement \\
+\hline
+Qualifiers & \makecell{required property (P2306) -- 1 \\ property the item statement must have \\ allowed value (P2305) -- 0..* \\ values the item statement may have} \\
+\hline
+Violation & \makecell{no item statement with required property \\ all item statements with required property have no allowed value} \\
+\hline
+Dependency &  all item statements\\
+\hline
+Rule features & negation \\
+\hline
+\end{tabularx}
+\end{table}
+
+An item violates the constraint if it a) has a statement using constrained property as predicate, b1) has no statement using the required property as predicate, or (if allowed values are specified) b2) has only statement using required property as predicate that do not use one of the allowed values as value.\\
+While condition a) can easily be tested by a tripleEDB-atom with the constrained property, condition b1) and b2) pose the question if an item does not have a statement.
+This is addressed in section [?].
+
+\subsection{Value requires statement (Q21510864)}\label{subsec_value_requires_statement}
+The value requires statement constraint requires an additional statement on the value of constrained statement, possibly also using a specific value from the list of allowed values.
+Since the constraint works very similar to item requires statement (see \ref{subsec_item_requires_statement}) --- requiring a statement on the value of the constrained statement instead of on the item --- it is not elaborated on.
+
+\subsection{Inverse (Q21510855)}\label{subsec_inverse}
+The inverse constraint requires an additional statement on the value of a constrained statement with a specific property and the constrained item as value.
+It is very similar to value requires statement (see \ref{subsec_value_requires_statement}), although it does not specify allowed values, since the required statement has one specific value: the constrained item.
+While this means it could not be modelled exactly as a value requires statement constraint, it would only require very small rule changes and will thus not be elaborated on.
+
+\subsection{Symmetric (Q21510862)}
+The symmetric constraint requires an additional statement on the value of a constrained statement with the constrained property and the constrained item as value.
+It is thus equal to an inverse constraint (see \ref{subsec_inverse}) with the same property as constrained property and required property and will not be elaborated on.
+
+\subsection{Multi-value (Q21510857)}
+The multi-value constraint specifies that the constrained property should appear in multiple statements of an item or none.
+
+\begin{table}[H]
+\caption{Multi-value constraint (Q21510857)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Requirement & item statement \\
+\hline
+Violation & \makecell{constrained item with exactly one constrained statement} \\
+\hline
+Dependency & all item statements \\
+\hline
+Rule features & count \\
+\hline
+\end{tabularx}
+\end{table}
+
+To model this constraint a count of statements using the constrained property per item would be needed.
+Then a violation would be any constrained item with a count of one for the constrained property.
+Since there is no need for arbitrarily large counts, but only for the difference between one and two, this could be solved explicitly.
+
+\subsection{Single value (Q19474404)}
+The single value constraint constrains the values a property may take on a single item.
+A property marked with the single value constraint can have properties marked as separators, which define exceptions under specific circumstances.
+\begin{table}[H]
+\caption{single value constraint (Q19474404)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Requirement & constrained statement value \\
+\hline
+Qualifiers & \makecell{separator (P4155) -- 0..* \\ qualifiers that must be different for exception} \\
+\hline
+Violation & \makecell{two constrained statements with same item and value \\ (exception: statements with separators as qualifiers with different values)} \\
+\hline
+Dependency & \makecell{without separators: all constrained statements on one item \\ with separators: additionally all qualifiers with separators on those statements }\\
+\hline
+Rule features & inequality, partially negation \\
+\hline
+\end{tabularx}
+\end{table}
+
+If no separators are specified, two triple form a violation of this constraint if they a) belong to the same item, b) have the same constrained property, and c) have the same value.
+If separators are specified, a violation as above can be ignored if the statements each have at least one qualifier with the same separator as predicate and different values.\\
+In the case without separators the modelling idea would be to find two triples tripleEDB with the same item and constrained property and test if their id is unequal.
+If separators were specified a violation would be the two triples and all their separator qualifiers either equal or non-existent.
+
+\subsection{Mandatory qualifier (Q21510856)}
+The mandatory qualifier constraint requires that a constrained statement has a qualifier with a specific property.
+
+\begin{table}[H]
+\caption{Mandatory qualifier constraint (Q21510856)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Requirement & qualifier on constrained statement \\
+\hline
+Qualifiers & \makecell{required property (P2306) -- 1 \\ qualifier that the statement must have} \\
+\hline
+Violation & \makecell{constrained statement without a qualifier with the required property} \\
+\hline
+Dependency & constrained statement and its qualifiers \\
+\hline
+Rule features & negation \\
+\hline
+\end{tabularx}
+\end{table}
+
+To find violations of this constraint one would find a tripleEDB with the constrained property, but where no qualifierEDB for the statement and with the required property exists.
+
+\subsection{Type (Q21503250)}\label{subsec_type}
+The type constraint requires the existence of either one of a specific statement (instanceOf <class>) or a chain of statements using subclassOf to connect directly (subclassOf* <class>)
+or via instanceOf (instanceOf/subclassOf* <class>) for all classes on the constrained item.
+\begin{table}[H]
+\caption{type constraint (Q21503250)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Requirement & chain of statements \\
+\hline
+Qualifiers & \makecell{relation (P2309) -- 1 \\ must be instance of (Q21503252), subclass of (Q21514624), \\ or instance or subclass of (Q30208840); \\ specifies the relation the constrained item must have to the classes\\
+class (P2308) -- 1..* \\ entities that must be reached using the relation} \\
+\hline
+Violation & constrained item without relation to a class \\
+\hline
+Dependency & \makecell{statements of constrained items using instanceOf (P31) or subclassOf(P279) \\ all statements using subclassOf (P279) }\\
+\hline
+Rule features & inequality, negation \\
+\hline
+\end{tabularx}
+\end{table}
+
+If the relation is instanceOf, the constraint can be evaluated like multiple item requires statement constraints (see \ref{subsec_item_requires_statement}) by setting the required property to instanceOf (P31) and, for each created item requires statement constraint, the allowed value to one of the classes.
+\\
+For the relations involving subclasses all transitive subclass relationships would be concluded by setting a subclass relationship if a third item can be found that is subclass of one and subclassed by the other item.
+This would be extended with one step of instanceOf.
+A violation would then be any constrained item that is not subclass of any of the classes.
+
+\subsection{Value type (Q21510865)}
+The value type constraint works like the type constraint (see \ref{subsec_type}) except the statement chains are required on the value of the constrained statement instead of the constrained item.
+It will thus not be expanded upon.
+
 \section{Reasoning for unconsidered constraints}
@@
 This problem can be solved with limited amounts of preprocessing by leveraging a preprocessed statement order per item and 'iterating' through the statements until the required conditions are met.\\
-A similar approach can be taken to find items with exactly one statement matching certain criteria , required for 	Multi-value.
+A similar approach can be taken to find items with exactly one statement matching certain criteria, required for 	Multi-value.
 The idea is to ask the question if a second statement exists, then evaluating it as above.
@@
 Each property constraint will be modelled and evaluated as one set of rules per property per constraint, meaning each violating statement will be of a form similar to
-\begin{verbatim}
-violation(ID, ITEM, propertyConstant, VALUE) :-
-    <rules determining the violation>
-\end{verbatim}
-where \texttt{propertyConstant} denotes the constrained property. While it would be possible to load all data at once and restrict the property constant with a separate predicate akin to
-\begin{verbatim}
-violation(ID, ITEM, PROPERTY, VALUE) :-
-    constrained(PROPERTY, constraintConstant),
-    <rules determining the violation>
-\end{verbatim}
+\begin{equation}
+\begin{split}
+\text{violation}(X, Y, \text{propertyConstant}, Z) \leftarrow \text{atoms determining the violation}
+\end{split}
+\end{equation}
+where propertyConstant denotes the constrained property. While it would be possible to load all data at once and restrict the property constant with a separate predicate akin to
+\begin{equation}
+\begin{split}
+\text{violation}&(X, Y, \text{PROPERTY}, Z) \leftarrow \\
+&\text{constrained}(\text{PROPERTY}, \text{constrainedConstant}) \wedge\\
+&\text{rules determining the violation}
+\end{split}
+\end{equation}
 this would increase the amount of facts loaded significantly while almost not reducing preprocessing.\\
@@
 
-\section{No preprocessing}\label{sec_no_preprocessing}
+\section{No additional preprocessing}\label{sec_no_preprocessing}
 

### 2018-07-29 thesis/paper/paper.tex
Started rules

@@
 10 years, months)} \\
-
 \end{tabularx}
@@
 
-\section{No additional preprocessing}\label{sec_no_preprocessing}
+\subsection{General approach}
+The goal outlined above is to identify violations of constrained properties, meaning a triple, qualifier, or reference with the constrained property will always be part of the rule body.
+With the exception of Property scope all constraints work on triples.
+The atom that will thus always be included is \(\text{tripleEDB}(X, Y, \text{propertyConstant}, Z)\).\\
+
+Every constraint violation depends on some constant in relation to the constrained statement.
+These requirements have been roughly identified in the previous chapter and will now be specified as rules.
+This step can include expanding the pre-processing to load more data such as units for quantities.\\
+
+The second step is the most complex: Depending on the constrained statement and the requirements, has a violation occurred?
+For the easiest category of constraints this step does not entail any additional work since the existence of the dependent structure is itself a violation.
+However, most constraints require at least evaluation of inequality between allowed values and existing values (\ref{sec_inequality}) or the non-existence of certain facts (\ref{sec_statement_nonexistence}).
+Both processes will be explained in their respective sections.
+
+The final step of translation is not necessary to identify violations per se, but provides additional information as to where exactly a violation has occurred.
+\todo[inline]{lets see how much we can actually do here}
+
+\section{No additional pre-processing}\label{sec_no_preprocessing}
+Four constraints do not require any additional features and can be easily translated, since they are violated if specific statements exists.
+A violation of these statements is thus always a triple, qualifier, or reference and violations will be expressed as the result of a query atom
+\[\text{violation\_triple}(S, I, P, V)\]
+\[\text{violation\_qualifier}(S, P, V)\]
+\[\text{violation\_reference}(S, P, V)\]
+where \(S\) specifies the statement ID, \(I\) the item, \(P\) the property, and \(V\) the value. The term propertyConstant will be used for the constant of the constrained property.
+
+\subsection{Property scope (Q53869507)}
+The property scope constraint requires no context information to be violated; the triple, qualifier, or reference with the constrained property is enough. The dependent atoms are thus:
+\[\text{tripleEDB}(S, I, \text{propertyConstant}, V)\]
+\[\text{qualifierEDB}(S, \text{propertyConstant}, V)\]
+\[\text{referenceEDB}(S, \text{propertyConstant}, V)\]
+depending on if the property scope qualifier is not set to as main value, as qualifier, or as reference respectively.\\
+
+A violation for a property scope constraint is
+\[\text{violation\_triple}(S, I, \text{propertyConstant}, V) \leftarrow \text{tripleEDB}(S, I, \text{propertyConstant}, V)\]
+\[\text{violation\_qualifier}(S, \text{propertyConstant}, V) \leftarrow \text{qualifierEDB}(S, \text{propertyConstant}, V)\]
+\[\text{violation\_reference}(S, \text{propertyConstant}, V) \leftarrow \text{referenceEDB}(S, \text{propertyConstant}, V)\]
+depending on if as main value, as qualifier, or as reference was not set in the qualifiers.\\
+
+There is no further information that can be materialized regarding this violations.
+
+\subsection{Conflicts-with (Q21502838)}
+The conflicts-with constraint is violated if another statement on the same item either has a conflicting property or a conflicting property and a conflicting value depending on specified qualifiers.
+It depends on a conjunction (\(O\) denoting another statement ID, \(C\) another value)
+\begin{equation*}
+\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \text{tripleEDB}(O, I, \text{conflictingPropertyConstant}, C)
+\end{equation*}
+\begin{equation*}
+\begin{split}
+\text{tripleEDB}&(S, I, \text{propertyConstant}, V) \wedge \\
+&\text{tripleEDB}(O, I, \text{conflictingPropertyConstant}, \text{conflictingValueConstant})
+\end{split}
+\end{equation*}
+
+A violation for a conflicts-with constraint is thus either
+\begin{equation*}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow\\
+&\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge\\
+&\text{tripleEDB}(O, I, \text{conflictingPropertyConstant}, C)
+\end{split}
+\end{equation*}
+for constraints with no conflicting values specified and one
+\begin{equation*}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow\\
+&\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge\\
+&\text{tripleEDB}(O, I, \text{conflictingPropertyConstant}, \text{conflictingValueConstant})
+\end{split}
+\end{equation*}
+for each conflicting value if specified.\\
 
 \section{Inequality}\label{sec_inequality}
+All rules except the four covered in the previous section require inequality in some way or another.
 

### 2018-07-30 thesis/paper/paper.tex
FIX: Typo
Added allowed entity types

@@
 \subsection{Property scope (Q53869507)}
-The property scope constraint requires no context information to be violated; the triple, qualifier, or reference with the constrained property is enough. The dependent atoms are thus:
-\[\text{tripleEDB}(S, I, \text{propertyConstant}, V)\]
-\[\text{qualifierEDB}(S, \text{propertyConstant}, V)\]
-\[\text{referenceEDB}(S, \text{propertyConstant}, V)\]
-depending on if the property scope qualifier is not set to as main value, as qualifier, or as reference respectively.\\
-
+The property scope constraint requires no context information to be violated; the triple, qualifier, or reference with the constrained property is enough.
 A violation for a property scope constraint is
@@
 The conflicts-with constraint is violated if another statement on the same item either has a conflicting property or a conflicting property and a conflicting value depending on specified qualifiers.
-It depends on a conjunction (\(O\) denoting another statement ID, \(C\) another value)
-\begin{equation*}
-\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \text{tripleEDB}(O, I, \text{conflictingPropertyConstant}, C)
-\end{equation*}
-\begin{equation*}
-\begin{split}
-\text{tripleEDB}&(S, I, \text{propertyConstant}, V) \wedge \\
-&\text{tripleEDB}(O, I, \text{conflictingPropertyConstant}, \text{conflictingValueConstant})
-\end{split}
-\end{equation*}
-
-A violation for a conflicts-with constraint is thus either
+It depends on a conjunction (\(O\) denoting another statement ID, \(C\) another value) and a violation for a conflicts-with constraint is thus either
 \begin{equation*}
@@
 \end{equation*}
-for constraints with no conflicting values specified and one
+for constraints with no conflicting values specified or one
 \begin{equation*}
@@
 
-\section{Inequality}\label{sec_inequality}
-All rules except the four covered in the previous section require inequality in some way or another.
+\subsection{Allowed entity types (Q52004125)}
+The allowed entity types constraint is violated by a statement with the wrong entity type.
+This means that during extraction additional data has to be readied.
+To encode this two new predicates are introduced, item\([1]\) and property\([1]\).
+All items and properties with a constrained statement will then be sorted into one of these two categories.
+Constraints regarding the lexeme entity type can be ignored since they are currently not deployed.
+The violations the depend on a simple conjunction and follow directly as
+\begin{equation*}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+&\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \text{item}(I)
+\end{split}
+\end{equation*}
+\begin{equation*}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+&\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \text{property}(I)
+\end{split}
+\end{equation*}
 

### 2018-07-30 thesis/paper/paper.tex
Added non of

@@
 
+\subsection{None of (Q52558054)}
+The none of constraint simply forbids certain values for properties, thus effectively prohibiting certain triples, qualifiers and references.
+The violations, depending on nothing but specific database entries, are
+\begin{equation*}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
+&\text{tripleEDB}(S, I, \text{propertyConstant}, \text{forbiddenValueConstant})
+\end{split}
+\end{equation*}
+\begin{equation*}
+\begin{split}
+\text{violation\_qualifier}&(S, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
+&\text{qualifierEDB}(S, \text{propertyConstant}, \text{forbiddenValueConstant})
+\end{split}
+\end{equation*}
+\begin{equation*}
+\begin{split}
+\text{violation\_reference}&(S, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
+&\text{referenceEDB}(S, \text{propertyConstant}, \text{forbiddenValueConstant})
+\end{split}
+\end{equation*}
+
 \section{Statement nonexistence}\label{sec_statement_nonexistence}

### 2018-07-30 thesis/paper/paper.tex
Added inequality introduction

@@
 
+\section{Inequality}\label{sec_inequality}
+All rules except the four covered in the previous section require inequality in some way or another.
+This means it is necessary to extract additional information during pre-processing, namely the constants for whom inequality needs to be stated.
+Each constraint will list the possible unequal entities in relation to the constrained statement.
+Multiple ways of implementing inequality while keeping the number of rules small will be discussed.
+Afterwards all constraints that can be translated using inequality will be explained.
+
+\subsection{Approaches}
+In every approach it will be necessary to load some basic inequalities into the database.
+To this end the predicate unequalEDB\([2]\) is introduced.
+Since the goal is to keep the number of applications of this predicate small, an IDB predicate unequal\([2]\) has to be introduced.
+This leads to the linear rule
+\begin{equation*}
+\text{unequal}(X, Y) \leftarrow \text{unequalEDB}(X, Y)
+\end{equation*}
+
+Given a set of entities \(E\), the task is to generate rules so that \(\forall x, y \in E, x \ne y: \text{unequal}(x, y)\).
+In doing so, dependant on the number of entities \(n = |E|\), the sum of the number of introduced EDB facts \(f(n)\), the number of introduced rules \(r(n)\), and the number of materialised facts \(m(n)\) should be kept as small as possible.
+The effects will be evaluated in section \ref{sec_inequality_evaluation}.
+
+\todo[inline]{How are the three numbers weighted? Or do I determine that in evaluation?}
+
 \section{Statement nonexistence}\label{sec_statement_nonexistence}

### 2018-07-30 thesis/paper/paper.tex
Added naive inequality

@@
 \todo[inline]{How are the three numbers weighted? Or do I determine that in evaluation?}
+\subsubsection{Naive}
+The naive approach would be to to simply add all possible combinations:
+\begin{equation*}
+\forall x, y \in E, x \ne y \text{ add } \text{unequalEDB}(x, y)
+\end{equation*}.
+This can be easily improved using the inverse and adding the following rule:
+\begin{equation}\label{equ_inverse}
+\text{unequal}(X, Y) \leftarrow \text{unequal}(Y, X)
+\end{equation}
+The it is sufficient to add each combinations once if disregarding their order:
+\begin{equation*}
+\forall x, y \in E, x < y \text{ add } \text{unequalEDB}(x, y)
+\end{equation*}.
+
+The naive approach with inverse would result in \(f(n) = \binom{n}{2}\) EDB facts, \(r(n) = 1\) additional rule and \(m(n) = \binom{n}{2}\) materialized facts since equation \ref{equ_inverse} would be applied exactly once to every EDB fact.
 

### 2018-07-30 thesis/paper/paper.tex
Added future sections
n > 1 added
Added encoding

@@
 Given a set of entities \(E\), the task is to generate rules so that \(\forall x, y \in E, x \ne y: \text{unequal}(x, y)\).
-In doing so, dependant on the number of entities \(n = |E|\), the sum of the number of introduced EDB facts \(f(n)\), the number of introduced rules \(r(n)\), and the number of materialised facts \(m(n)\) should be kept as small as possible.
+In doing so, dependent on the number of entities \(n = |E|\), the sum of the number of introduced EDB facts \(f(n)\) and the number of introduced rules \(r(n)\) should be kept as small as possible.
+Note that \(n > 1\) is presumed.
 The effects will be evaluated in section \ref{sec_inequality_evaluation}.
@@
 \todo[inline]{How are the three numbers weighted? Or do I determine that in evaluation?}
-\subsubsection{Naive}
+\todo[inline]{Is the number of materialised facts even relevant? Maybe look at rule applications.}
+%, and the number of materialised facts \(m(n)\)
+\subsubsection{Naive}\label{subsubsec_naive}
 The naive approach would be to to simply add all possible combinations:
@@
 
-The naive approach with inverse would result in \(f(n) = \binom{n}{2}\) EDB facts, \(r(n) = 1\) additional rule and \(m(n) = \binom{n}{2}\) materialized facts since equation \ref{equ_inverse} would be applied exactly once to every EDB fact.
+The naive approach with inverse would result in \(f(n) = \binom{n}{2}\) EDB facts and \(r(n) = 1\) additional rule.
+%and \(m(n) = \binom{n}{2}\) materialized facts since equation \ref{equ_inverse} would be applied exactly once to every EDB fact.
+
+\subsubsection{Encoding}
+One option to determine inequality would be to utilise that all entities are strings of single characters.
+The idea is to determine the set of characters used in all entities \(C\), set them as unequal as outlined above in section \ref{subsubsec_naive} and associate each entity with their characters.\\
+
+An entity \(e \in E\) is always a string and can be written as \(c_0 \ldots c_{l - 1}\) with \(l = len(e)\) being the length of the string.
+Denote the set \(C = \{c | \exists e \in E: c \in e\}\)\todo{\(c \in e\) does not seem correct} as the set of characters occurring in entities.
+Then \(l_{\max} = \max\limits_{\forall e \in E}(len(e))\) is the length of the longest entity string.
+Based on this a group of predicates letter\_iEDB[2] is introduced where \(i \in \{0 \ldots l_{\max}-1\}\) and letter\_iEBD denotes the ith letter of the entity string.\\
+
+For each entity string \(e = c_0 \ldots c_{l - 1}\) each predicate letter\_iEDB is created as \(\text{letter\_iEDB}(e, f(e, i))\), where
+\begin{equation*}
+f(e, i) = 
+\begin{cases}
+      c_i & i < len(e) \\
+      \text{NONE} & i \geq len(e)
+\end{cases}
+\end{equation*}
+
+Now that inequality between all constitution characters and the link between entities and their characters are introduced, the inequality between two entities can be derived by creating the following rules for all predicates letter\_i:
+\begin{equation*}
+\text{unequal}(X, Y) \leftarrow \text{letter\_i}(X, A) \wedge \text{letter\_i}(Y, B) \wedge \text{unequal}(A, B)
+\end{equation*}
+
+With this approach the statistics depend on the number of distinct characters \(n_c = |C|\).
+The number of introduced EDB facts is \(f(n, n_c) = \binom{n_c}{2} + n*l_{\max}\) and the number of introduced rules \(r(n, n_c) = 1 + l_{\max}\).
+The effectiveness of this approach thus depends on the ratio between \(n\) and \(n_c\).
+
+\subsubsection{Demand-driven materialization}
 

### 2018-07-30 thesis/paper/paper.tex
Started demand-driven and added distinct values

@@
 
-\subsubsection{Demand-driven materialization}
+\subsubsection{Demand-driven materialisation}
+The approaches outlined above will always result in all inequalities between the members of \(E\) being materialised.
+However, the purpose of determining inequality is to assist in finding violations of specific constraints.
+This means that only some inequalities -- the ones that actually appear in the dataset -- have to be evaluated.
+The rules can then be limited by introducing a new predicate req\_inequality\([2]\) which will only be derived if a condition occurs in which the inequality must be evaluated.\\
+
+\subsection{Distinct values (Q21502410)}
+The distinct values constraints requires two distinct constrained statements with the same value. Inequality needs to be evaluated if two statements with the same value have been found.
+\begin{equation*}
+\text{req\_inequality}(S, O) \leftarrow \text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \text{tripleEDB}(O, X, \text{propertyConstant}, V)
+\end{equation*}
+This premise in addition with inequality between the statement IDs can then be used to identify all violations.
+\begin{equation*}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+&\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \\
+&\text{tripleEDB}(O, X, \text{propertyConstant}, V) \wedge \\
+&\text{unequal}(S, O)
+\end{split}
+\end{equation*}
+While it is possible to set this constraint on qualifiers or references, this is currently not done and is thus not evaluated here.
 

### 2018-07-30 thesis/paper/paper.tex
Added allowed units

@@
 
+\subsection{Allowed units (Q21514353)}
+The allowed units constraint needs to determine inequality between all allowed units and each unit on a constrained statement.
+To this end the unit of a triple, qualifier or reference needs to be extracted during pre-processing.
+Three predicates unitTriple\([2]\), unitQualifier\([4]\) and unitReference\([4]\) are introduced.
+While each triple has a unique identifier and only need to associate this with the unit, qualifiers and references need to be identified completely to assign the correct unit.\\
+
+Inequality between units needs to be determined if a constrained triple, qualifier or reference exists. For every allowed unit of the constraint the following rules are introduced
+\begin{equation*}
+\begin{split}
+\text{req\_inequality}&(X, \text{allowedUnitConstant}) \leftarrow \\
+&\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \\
+&\text{unitTriple}(S, X)
+\end{split}
+\end{equation*}
+\begin{equation*}
+\begin{split}
+\text{req\_inequality}&(X, \text{allowedUnitConstant}) \leftarrow \\
+&\text{qualifierEDB}(S, \text{propertyConstant}, V) \wedge \\
+&\text{unitQualifier}(S, \text{propertyConstant}, V, X)
+\end{split}
+\end{equation*}
+\begin{equation*}
+\begin{split}
+\text{req\_inequality}&(X, \text{allowedUnitConstant}) \leftarrow \\
+&\text{referenceEDB}(S, \text{propertyConstant}, V) \wedge \\
+&\text{unitReference}(S, \text{propertyConstant}, V, X)
+\end{split}
+\end{equation*}
+
+The violations can then be found using the same premises with inequality added.
+\todo[inline]{Kind of ridiculous to write basically the same premise repeatedly. Should probably write the general approach in demand-driven and just write the normal inequality here}
+
 \section{Statement nonexistence}\label{sec_statement_nonexistence}

### 2018-07-31 thesis/paper/paper.tex
More demand-driven

@@
 
-\subsubsection{Encoding}
+\subsubsection{Encoding}\label{subsubsec_encoding}
 One option to determine inequality would be to utilise that all entities are strings of single characters.
@@
 The rules can then be limited by introducing a new predicate req\_inequality\([2]\) which will only be derived if a condition occurs in which the inequality must be evaluated.\\
+
+To save space and avoid redundancy the following rules will not explicitly state the inequality requirement predicate.
+Instead, assume that for every premise with the form \(B_1 \wedge \ldots\wedge B_m \wedge I\) where \(B_1 ..
+B_m\) are arbitrary atoms, \(I\) is an atom of the form \(\text{unequal}(t_1, t_2)\) with \(t_1,t_2\) terms and all of \(t_1,t_2\) that are variables appear in at least one of \(B_1 \ldots B_m\), the rule
+\begin{equation*}
+\text{req\_inequality}(t_1, t_2)
+ \leftarrow B_1 \wedge \ldots \wedge B_m
+\end{equation*} is introduced.\\
+
+The rules outlined in section \ref{subsubsec_encoding} are changed to utilise this new requirement.
+\begin{equation*}
+\text{unequal}(X, Y) \leftarrow \text{req\_inequality}(X, Y) \wedge \text{letter\_i}(X, A) \wedge \text{letter\_i}(Y, B) \wedge \text{unequal}(A, B)
+\end{equation*}
+
+\todo{could use this to restrict the character inequality too but do we gain much?}
 

### 2018-07-31 thesis/paper/paper.tex
Modded distinct values

@@
 Then a violation can be found by finding all tripleEDB with the constrained property and a unit unequal to all allowed units.
+In theory the constraint can also be applied to qualifiers and references by setting the respective value on a constraint scope (P4680) qualifier.
+However, there is currently no such case and will therefore not be regarded.
 
@@
 \subsection{Distinct values (Q21502410)}
-The distinct values constraints requires two distinct constrained statements with the same value. Inequality needs to be evaluated if two statements with the same value have been found.
-\begin{equation*}
-\text{req\_inequality}(S, O) \leftarrow \text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \text{tripleEDB}(O, X, \text{propertyConstant}, V)
-\end{equation*}
-This premise in addition with inequality between the statement IDs can then be used to identify all violations.
+The distinct values constraints requires two distinct constrained statements with the same value. 
+While finding constrained statements with the same value is straightforward, violations only occur if the statements are actually different, i.e. have an unequal statement id.
 \begin{equation*}

### 2018-07-31 thesis/paper/paper.tex
Moved text to correct section

@@
 To model the distinct values constraint it is necessary to find triples tripleEDB with the constrained property and the same value whose statement id is unequal.
+In theory the constraint can also be applied to qualifiers and references by setting the respective value on a constraint scope (P4680) qualifier.
+However, there is currently no such case and will therefore not be regarded.
 
@@
 Modelling the constraint requires a predicate associating the value or the statement with its unit and inequality between every unit and all allowed units.
-Then a violation can be found by finding all tripleEDB with the constrained property and a unit unequal to all allowed units.
-In theory the constraint can also be applied to qualifiers and references by setting the respective value on a constraint scope (P4680) qualifier.
-However, there is currently no such case and will therefore not be regarded.
+Then a violation can be found by finding all tripleEDB, qualifierEDB and referenceEDB with the constrained property and a unit unequal to all allowed units.
+Theoretically some of these can be excluded by using the constraint scope (P4680) qualifier on the constraint but there is no current case where this is done.
 
@@
 The violations can then be found using the same premises with inequality added.
-\todo[inline]{Kind of ridiculous to write basically the same premise repeatedly. Should probably write the general approach in demand-driven and just write the normal inequality here}
 

### 2018-07-31 thesis/paper/paper.tex
Modded allowed units

@@
 To this end the unit of a triple, qualifier or reference needs to be extracted during pre-processing.
-Three predicates unitTriple\([2]\), unitQualifier\([4]\) and unitReference\([4]\) are introduced.
-While each triple has a unique identifier and only need to associate this with the unit, qualifiers and references need to be identified completely to assign the correct unit.\\
+Quantities are rendered as string during pre-processing, specifying their value, bounds and unit.
+The new predicate unit\([2]\) will associate a value with its unit.\\
 
-Inequality between units needs to be determined if a constrained triple, qualifier or reference exists. For every allowed unit of the constraint the following rules are introduced
+To determine if a violation occurred the unit to the constrained statement needs to be found. This is done via the unit-predicate as below
+\begin{equation*}
+\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \text{unit}(V, U)
+\end{equation*}
+with qualifiers and references working on the same lines.
+
+The full rules require multiple unequal-atoms for all allowed units.
+Given a set \(A\) of constants, in this case allowed units, \(\text{unequal}(\{A\}, t)\) is short for the conjunction of the set \(\{\text{unequal}(a, t) | \forall a \in A\}\), with \(t\) being an arbitrary term.
+
+Violation are thus
 \begin{equation*}
 \begin{split}
-\text{req\_inequality}&(X, \text{allowedUnitConstant}) \leftarrow \\
+\text{violation\_triple}&(S, I \text{propertyConstant}, V) \leftarrow \\
 &\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \\
-&\text{unitTriple}(S, X)
+&\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
 \end{split}
@@
 \begin{split}
-\text{req\_inequality}&(X, \text{allowedUnitConstant}) \leftarrow \\
+\text{violation\_qualifier}&(S, I \text{propertyConstant}, V) \leftarrow \\
 &\text{qualifierEDB}(S, \text{propertyConstant}, V) \wedge \\
-&\text{unitQualifier}(S, \text{propertyConstant}, V, X)
+&\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
 \end{split}
@@
 \begin{split}
-\text{req\_inequality}&(X, \text{allowedUnitConstant}) \leftarrow \\
+\text{violation\_reference}&(S, I \text{propertyConstant}, V) \leftarrow \\
 &\text{referenceEDB}(S, \text{propertyConstant}, V) \wedge \\
-&\text{unitReference}(S, \text{propertyConstant}, V, X)
+&\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
 \end{split}
 \end{equation*}
-
-The violations can then be found using the same premises with inequality added.
 

### 2018-07-31 thesis/paper/paper.tex
Modded Datalog

@@
 Datalog is a logic programming language.
-It is fully declarative and uses simple rules to derive additional knowledge from given facts.
+It is fully declarative and uses existential rules to derive additional knowledge from given facts.
 The most basic unit of information in datalog is a \emph{constant}, which will be denoted as someConstant.
@@
 An atom has the form predicate\((t_1, \ldots , t_n)\) with \(n\) as the arity of the predicate and \(t_1,\ldots,t_n\) terms.\\
-Atoms are the building blocks of \emph{rules}, which are used to derive information in datalog.
-Given atoms \(C\) and \(B_n\), a rule takes the form \(C \leftarrow P_1 \wedge \ldots \wedge P_m\), with \(C\) being the conclusion and \(B_1 \wedge \ldots \wedge B_m\) the premise.
-If the premise does not contain any atoms such a rule is called a \emph{fact}, or a \emph{ground rule} if it additionally only uses constants.
-A finite set of rules forms a \emph{program}.\\
+Atoms are the building blocks of \emph{existential rules}, which are used to derive information in datalog.
+A rule is premise that implies a conclusion, and both are conjunctions of atoms.
+Denote \(P_n\) as the premise atoms, \(C_n\) as the conclusion atoms.
+The variables used are divided into three disjunct sets \(V_p\), \(V_c\), and \(V_r\) which are the variables occurring only in premise atoms, only in conclusion atoms, and occurring in both respectively.
+The general from of an existential rule is (with \(k \in \mathbb{N}^+\), \(l \in \mathbb{N}\))
+\begin{equation*}
+\forall p \in V_p, r \in V_r : \exists c \in V_c: C_1 \ldots C_k \leftarrow P_1 \ldots P_l
+\end{equation*}
+
+\(V_c\) are called existential variables.
+If \(V_c = \emptyset\) the rule is called a \emph{datalog rule} and if \(l = 0\) is called a \emph{fact}.
+Since \(\forall p \in V_p, r \in V_r\) can be taken as given for all rules, it will be omitted in this work.
 
@@
 All others (predicates that appear in the conclusion at least once) are IDB.
-
-\todo[inline]{Explain the relation to FO-logic?}
 
@@
 
-\todo[inline]{Probably a good place to explain existential and universal quantifiers.}
+\todo[inline]{Should I explain the chase? And in what detail...}
 \section{Representation}

### 2018-07-31 thesis/paper/paper.tex
Added allowed qualifiers

@@
 
-\subsection{Allowed units (Q21514353)}
+\subsection{Allowed units (Q21514353)}\label{subsec_allowed_units}
 The allowed units constraint needs to determine inequality between all allowed units and each unit on a constrained statement.
-To this end the unit of a triple, qualifier or reference needs to be extracted during pre-processing.
+To this end the units of a triple, qualifier or reference needs to be extracted during pre-processing and their inequality established.
 Quantities are rendered as string during pre-processing, specifying their value, bounds and unit.
@@
 
+\subsection{Allowed qualifiers (Q21510851)}
+The allowed qualifiers constraint is violated if a qualifier on a constrained statement exists that is different from all allowed qualifiers.
+During pre-processing all qualifiers on constrained statements will be extracted and the inequality of their properties and the allowed qualifiers established.
+We will use the notation for set inequality as introduced in section \ref{subsec_allowed_units}.
+With \(A\) the set of allowed qualifiers the rule is:
+\begin{equation*}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+&\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \\
+&\text{qualifierEDB}(S, Q, O) \wedge \text{unequal}(\{A\}, Q)
+\end{split}
+\end{equation*}
+
 \section{Statement nonexistence}\label{sec_statement_nonexistence}

### 2018-07-31 thesis/paper/paper.tex
Added one-of

@@
 
+\subsection{One-of (Q21510859)}
+The one-of constraint fits in the pattern established by the previous two constraints and disallows certain property-value combinations.
+With \(A\) the set of allowed values the rules for triple, qualifier and reference are respectively:
+\begin{equation*}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+&\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
+\end{split}
+\end{equation*}
+\begin{equation*}
+\begin{split}
+\text{violation\_qualifier}&(S, \text{propertyConstant}, V) \leftarrow \\
+&\text{qualifierEDB}(S, \text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
+\end{split}
+\end{equation*}
+\begin{equation*}
+\begin{split}
+\text{violation\_reference}&(S, \text{propertyConstant}, V) \leftarrow \\
+&\text{referenceEDB}(S, \text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
+\end{split}
+\end{equation*}
+
 \section{Statement nonexistence}\label{sec_statement_nonexistence}

### 2018-07-31 thesis/paper/paper.tex
Started single best value

@@
 The one-of constraint fits in the pattern established by the previous two constraints and disallows certain property-value combinations.
+Inequality is pre-processed between all values of constrained statements.
 With \(A\) the set of allowed values the rules for triple, qualifier and reference are respectively:
@@
 
+\subsection{Single best value (Q52060874)}
+The single best value constraint limits the number of constrained statements with rank preferred on the same item to one.
+This means during pre-processing the rank information per triple must be extracted and attached to the statement with the new predicate rank\([2]\) which links the statement ID and the rank.
+Qualifiers and references are not affected since they do not have ranks.
+Inequality is established between all constrained statement IDs.
+For the simple case without separators the rule is:
+\todo[inline]{Well, this isn't done.}
+
 \chapter{Evaluation/Comparison}

### 2018-07-31 thesis/paper/paper.tex
Moved single best value to correct position
Added One-of qualifier value

@@
 
-\subsection{Single best value (Q52060874)}
-The single best value constraint requires that only one constrained statement per item has the rank preferred.
-
-\begin{table}[H]
-\caption{Single best value constraint (Q52060874)}
-\begin{tabularx}{\textwidth}{ ll X}
-\hline
-Requirement & item statement rank \\
-\hline
-Violation & \makecell{items with two constrained statements with rank preferred} \\
-\hline
-Dependency & all item statements and their rank \\
-\hline
-Rule features & inequality \\
-\hline
-\end{tabularx}
-\end{table}
-
-Modelling the constraint means a count of all constrained statements with rank preferred is necessary, with a violation constituted if an item has a count of two or higher.
-Since it is only needed to differentiate between one and many, the rules could be written explicitly.
-
 \subsection{One-of qualifier value (Q52712340)}
@@
 To find violations of this constraint one would find a tripleEDB with the constrained property, but where no qualifierEDB for the statement and with the required property exists.
+
+\subsection{Single best value (Q52060874)}
+The single best value constraint requires that only one constrained statement per item has the rank preferred.
+
+\begin{table}[H]
+\caption{Single best value constraint (Q52060874)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Requirement & item statement rank \\
+\hline
+Violation & \makecell{items with two constrained statements with rank preferred} \\
+\hline
+Dependency & all item statements and their rank \\
+\hline
+Rule features & inequality \\
+\hline
+\end{tabularx}
+\end{table}
+
+Modelling the constraint means a count of all constrained statements with rank preferred is necessary, with a violation constituted if an item has a count of two or higher.
+Since it is only needed to differentiate between one and many, the rules could be written explicitly.
 
@@
 
+\subsection{One-of qualifier value (Q52712340)}
+The one-of qualifier value constraint is violated if a constrained statement has a qualifier with a value unequal to all allowed values.
+Inequality needs to be pre-processed between all values of qualifiers on constrained statements.
+With the set of allowed Values \(A\) the rules are
+\begin{equation*}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+&\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \\
+&\text{qualifierEDB}(S, \text{qualifierProperty}, O) \wedge \text{unequal}(\{A\}, O)
+\end{split}
+\end{equation*}
+
 \section{Statement nonexistence}\label{sec_statement_nonexistence}

### 2018-07-31 thesis/paper/paper.tex
Added statement non-existence

@@
 The idea is to ask the question if a second statement exists, then evaluating it as above.
-Both methods will be explained in section \ref{sec_statement_nonexistence}.
+Both methods will be explained in section \ref{sec_statement_non-existence}.
 
@@
 For the easiest category of constraints this step does not entail any additional work since the existence of the dependent structure is itself a violation.
-However, most constraints require at least evaluation of inequality between allowed values and existing values (\ref{sec_inequality}) or the non-existence of certain facts (\ref{sec_statement_nonexistence}).
+However, most constraints require at least evaluation of inequality between allowed values and existing values (\ref{sec_inequality}) or the non-existence of certain facts (\ref{sec_statement_non-existence}).
 Both processes will be explained in their respective sections.
@@
 
-\section{Statement nonexistence}\label{sec_statement_nonexistence}
+\section{Statement non-existence}\label{sec_statement_non-existence}
+Multiple statements are violated if specific statements do not exist in relation to constrained statements.
+It is not possible to simply negate said statements and pre-processing every required statement would defeat the purpose of implementing them in datalog.
+However, non-existence can be derived utilising the order of statements.
+Three predicates are introduced: first\([2]\), last\([2]\), and next\([2]\), which denote the first and last statement of an item and the following statement to another using their statement IDs.\\
+
+If qualifier non-existence should be tested the predicates have to be adapted, since they do not have a unique identifier.
+Their order has to be established using the entire set of statement ID, property and value.
+For this the predicates first\_qualifier\([3]\) and last\_qualifier\([3]\) taking the statement ID, the property and the value and next\_qualifier\([6]\) with ID, property and value for the current and next qualifier are introduced.\\
+
+Using this pre-processed data the approach is to apply a require-predicate to each statement or qualifier.
+They will be introduced in the respective constraint sections since they may need to carry different information.
+For the introduction the predicate require\([1]\) simply denotes that the statement requires something.
+The qualifier variant works analogous.\\
+
+The pattern consists of two rules, both utilising the same conjunction \(t = \text{tripleEDB}(S, I, P, V) \wedge C_1 ..
+C_m\) with arbitrary atoms \(C_n\) which is used to test if the triple does not fulfil the requirement.
+Then rule one
+\begin{equation*}
+\text{require}(S) \leftarrow \text{first}(S) \wedge t
+\end{equation*}
+states that the statement requires the condition and
+\begin{equation*}
+\text{require}(O) \leftarrow \text{next}(S, O) \wedge \text{require}(S) \wedge t
+\end{equation*}
+propagates this down the statement chain. If the last statement of an item or the last qualifier of a statement has require the violation occurred.\\
+
+In the following constraints only the conjunction \(t\) and adjustments to this pattern will be described. The require-propagation is regarded as given.
 

### 2018-07-31 thesis/paper/paper.tex
Added item requires statement

@@
 
+\subsection{Item requires statement (Q21503247)}
+The item requires statement constraint requires the pre-processing of all item statements.
+Additionally, inequality between all properties of item statements and the required property as well as between all values and the allowed values has to be established.
+It occurs in two varieties, depending on if allowed values have been specified or not.
+The predicate is defined as require\([2]\) denoting the statement and the required property.
+Without separators the conjunction is \(t = \text{tripleEDB}(S, I, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)\).\\
+
+If a set of allowed values \(A\) has been specified, a second set of require-rules have to be introduced with \(t = \text{tripleEDB}(S, I, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)\).
+Note that the first set of rules is still required, since the preceding statements do not need to have the required property.\\
+
+In both cases violations can be found with
+\begin{equation*}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+&\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \\
+&\text{tripleEDB}(O, I, P, X) \wedge \text{last}(O, I) \wedge \\
+&\text{require}(O, \text{requiredPropertyConstant})
+\end{split}
+\end{equation*}
+
 \subsection{Single best value (Q52060874)}

### 2018-07-31 thesis/paper/paper.tex
Added value requires statement

@@
 constrained item & item with a constrained statement \\
-item statement & statement on the same item as the constrained statement \\
+item statement & statement on the same item of a constrained statement \\
+value statement & statement on the value of a constrained statement
 \end{tabular}
@@
 
+\subsection{Value requires statement (Q21510864)}
+The value requires statement constraint is syntactically very similar to the previous item requires statement constraint.
+Instead of requiring a statement on the item of a constrained statement it requires it on the value.
+The pre-processing changes to extracting all value statements along the constrained statements and ordering them.
+Inequality need to be computed between all properties and all values on value statements.
+The conjunctions remain the same and the final violation rule is
+\begin{equation*}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+&\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \\
+&\text{tripleEDB}(O, V, P, X) \wedge \text{last}(O, V) \wedge \\
+&\text{require}(O, \text{requiredPropertyConstant})
+\end{split}
+\end{equation*}
+
 \subsection{Single best value (Q52060874)}

### 2018-07-31 thesis/paper/paper.tex
Added inverse

@@
 The predicate is defined as require\([2]\) denoting the statement and the required property.
-Without separators the conjunction is \(t = \text{tripleEDB}(S, I, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)\).\\
-
-If a set of allowed values \(A\) has been specified, a second set of require-rules have to be introduced with \(t = \text{tripleEDB}(S, I, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)\).
+Without separators the conjunction is \(t_1 = \text{tripleEDB}(S, I, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)\).\\
+
+If a set of allowed values \(A\) has been specified, a second set of require-rules have to be introduced with \(t_2 = \text{tripleEDB}(S, I, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)\).
 Note that the first set of rules is still required, since the preceding statements do not need to have the required property.\\
@@
 &\text{require}(O, \text{requiredPropertyConstant})
+\end{split}
+\end{equation*}
+
+\subsection{Inverse (Q21510855)}
+The inverse constraint is similar to the value requires statement constraint in requiring a statement on a constrained value with a specific property.
+However, instead of putting no constraint or a list of fixed values on the required statements value it set a relative requirement: the constrained item.
+The pre-processing and inequalities thus remain the same, but the conjunctions change to
+\begin{equation*}
+t_1 = \text{tripleEDB}(S, I, P, V) \wedge \text{unequal}(\text{inversePropertyConstant}, P)
+\end{equation*}
+\begin{equation*}
+\begin{split}
+t_2 = &\text{tripleEDB}(O, R, \text{propertyConstant}, X) \wedge \\
+&\text{tripleEDB}(S, I, \text{inversePropertyConstant}, V) \wedge \\
+&\text{unequal}(R, V)
 \end{split}

### 2018-08-01 thesis/paper/paper.tex
Added symmetric
Added multi-value

@@
 
+\subsection{Symmetric (Q21510862)}
+The symmetric constraint can be translated exactly as the inverse constraint if the inverse property where set to the same as the constrained property.
+
+\subsection{Multi-value (Q21510857)}
+The multi-value constraint requires a limited count of constrained statements.
+Alternatively, it could be said that if the item requires a second statement, the constraint is violated.
+Pre-processing extracts all items statements and inequality between all item statement properties as usual and the conjunction is \(t = \text{tripleEDB}(S, I, P, V) \wedge \text{unequal}(\text{propertyConstant}, P)\).
+The predicate require\_second\([2]\) is introduced with the same meaning as require except that it will be applied after a required property has been found:
+\begin{equation*}
+\text{require\_second}(S, \text{propertyConstant}) \leftarrow \text{next}(Q, R) \wedge \text{next}(R, S) \wedge \text{require}(Q) \wedge t
+\end{equation*}
+\begin{equation*}
+\text{require\_second}(S, \text{propertyConstant}) \leftarrow \text{next}(R, S) \wedge \text{require\_second}(R) \wedge t
+\end{equation*}
+A violations would be any last statement requiring a second constrained statement.
+
 \subsection{Single best value (Q52060874)}

### 2018-08-01 thesis/paper/paper.tex
Modded single value

@@
 If no separators are specified, two triple form a violation of this constraint if they a) belong to the same item, b) have the same constrained property, and c) have the same value.
-If separators are specified, a violation as above can be ignored if the statements each have at least one qualifier with the same separator as predicate and different values.\\
+If separators are specified, a violation as above can be ignored if the statements each have at least one qualifier with the same separator as predicate and different values or an unequal number of separator qualifiers.\\
 In the case without separators the modelling idea would be to find two triples tripleEDB with the same item and constrained property and test if their id is unequal.
-If separators were specified a violation would be the two triples and all their separator qualifiers either equal or non-existent.
+If separators were specified a violation would be the two triples and all their separator qualifiers equal or non-existent.
 

### 2018-08-01 thesis/paper/paper.tex
Added single value

@@
 
+\subsection{Single value (Q19474404)}
+The single value constraint can be divided into two different cases, depending on if the set of separators \(S\) is empty or not.
+Pre-processing needs to extract all item statements and their qualifiers and establish inequality between all statement IDs and all qualifier values.
+
+\subsubsection{No separators}
+If the separator set is empty, the violation simply occurs if there are two constrained statements on the same item. 
+\begin{equation}\label{eq_no_separators_triple}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+&\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \\
+&\text{tripleEDB}(O, I, \text{propertyConstant}, X) \wedge \\
+&\text{unequal}(S, O)
+\end{split}
+\end{equation}
+For qualifiers and reference there is no ID whose inequality could be checked.
+However, since the statement ID and the property of a possible conflict are always the same, inequality can be tested on the values.
+\begin{equation*}
+\begin{split}
+\text{violation\_qualifier}&(S, \text{propertyConstant}, V) \leftarrow \\
+&\text{qualifierEDB}(S, \text{propertyConstant}, V) \wedge \\
+&\text{qualifierEDB}(S, \text{propertyConstant}, X) \wedge \\
+&\text{unequal}(V, X)
+\end{split}
+\end{equation*}
+\begin{equation*}
+\begin{split}
+\text{violation\_reference}&(S, \text{propertyConstant}, V) \leftarrow \\
+&\text{referenceEDB}(S, \text{propertyConstant}, V) \wedge \\
+&\text{referenceEDB}(S, \text{propertyConstant}, X) \wedge \\
+&\text{unequal}(V, X)
+\end{split}
+\end{equation*}
+\todo{Check if all reference-rules look exactly like the qualifier rules and cut them out if it is so.}
+
+\subsubsection{With separators}
+To find all violations under separators a number of things have to be considered.
+Only triples are eligible, since qualifiers cannot have further qualifier, and the order of qualifiers on constrained statements needs to be pre-processed as well since qualifier non-existence is relevant.
+The predicate require\([4]\) is introduced which notes the statement ID, property, and value of a qualifier as well as the separator.
+The pattern needs to be applied for every separator, meaning the conjunction is: \(t = \text{qualifierEDB}(S, P, V) \wedge \text{unequal}(P, \text{requiredPropertyConstant})\) for each separator in \(S\).\\
+
+With qualifier non-existence established the violation can be found.
+Two statements form a violation if they are distinct as per equation \ref{eq_no_separators_triple} and additionally fulfil the condition that all qualifiers are either equal or non-existent.
+\(\forall X \subset S, Y = S \setminus X\) the following rule has to be added.
+Note that \(qualifierEDB(S, {X}, [X])\) denotes one atom for each element \(x \in X\) as introduced in section \ref{subsec_allowed_units} with \([X]\) representing a variable which is distinct for each atom unless they have the same \(x\).
+With the predicate does\_not\_have\([2]\) linking a statement ID to the separator it does not have via \(\text{does\_not\_have}(S, R) \leftarrow \text{last\_qualifier}(S, P, V) \wedge \text{require}(S, P, V, R)\) for brevity the rules are
+\begin{equation*}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+&\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \\
+&\text{tripleEDB}(O, I, \text{propertyConstant}, W) \wedge \\
+&\text{unequal}(S, O) \wedge \\
+&\text{qualifierEDB}(S, \{X\}, [X]) \wedge \\
+&\text{qualifierEDB}(O, \{X\}, [X]) \wedge \\
+&\text{does\_not\_have}(S, \{Y\}) \wedge \\
+&\text{does\_not\_have}(O, \{Y\})
+\end{split}
+\end{equation*}
+
 \subsection{Single best value (Q52060874)}

### 2018-08-01 thesis/paper/paper.tex
Separated constraint labels by chapter

@@
 
-\subsection{Item requires statement (Q21503247)}\label{subsec_item_requires_statement}
+\subsection{Item requires statement (Q21503247)}\label{subsec_2_item_requires_statement}
 The item requires statement constraint requires an additional statement on an item with a statement using the constrained property, possibly also using a specific value from the list of allowed values.
@@
 
-\subsection{Value requires statement (Q21510864)}\label{subsec_value_requires_statement}
+\subsection{Value requires statement (Q21510864)}\label{subsec_2_value_requires_statement}
 The value requires statement constraint requires an additional statement on the value of constrained statement, possibly also using a specific value from the list of allowed values.
-Since the constraint works very similar to item requires statement (see \ref{subsec_item_requires_statement}) --- requiring a statement on the value of the constrained statement instead of on the item --- it is not elaborated on.
-
-\subsection{Inverse (Q21510855)}\label{subsec_inverse}
+Since the constraint works very similar to item requires statement (see \ref{subsec_2_item_requires_statement}) --- requiring a statement on the value of the constrained statement instead of on the item --- it is not elaborated on.
+
+\subsection{Inverse (Q21510855)}\label{subsec_2_inverse}
 The inverse constraint requires an additional statement on the value of a constrained statement with a specific property and the constrained item as value.
-It is very similar to value requires statement (see \ref{subsec_value_requires_statement}), although it does not specify allowed values, since the required statement has one specific value: the constrained item.
+It is very similar to value requires statement (see \ref{subsec_2_value_requires_statement}), although it does not specify allowed values, since the required statement has one specific value: the constrained item.
 While this means it could not be modelled exactly as a value requires statement constraint, it would only require very small rule changes and will thus not be elaborated on.
@@
 The symmetric constraint requires an additional statement on the value of a constrained statement with the constrained property and the constrained item as value.
-It is thus equal to an inverse constraint (see \ref{subsec_inverse}) with the same property as constrained property and required property and will not be elaborated on.
+It is thus equal to an inverse constraint (see \ref{subsec_2_inverse}) with the same property as constrained property and required property and will not be elaborated on.
 
@@
 
-\subsection{Type (Q21503250)}\label{subsec_type}
+\subsection{Type (Q21503250)}\label{subsec_2_type}
 The type constraint requires the existence of either one of a specific statement (instanceOf <class>) or a chain of statements using subclassOf to connect directly (subclassOf* <class>)
@@
 
-If the relation is instanceOf, the constraint can be evaluated like multiple item requires statement constraints (see \ref{subsec_item_requires_statement}) by setting the required property to instanceOf (P31) and, for each created item requires statement constraint, the allowed value to one of the classes.
+If the relation is instanceOf, the constraint can be evaluated like multiple item requires statement constraints (see \ref{subsec_2_item_requires_statement}) by setting the required property to instanceOf (P31) and, for each created item requires statement constraint, the allowed value to one of the classes.
 \\
@@
 \subsection{Value type (Q21510865)}
-The value type constraint works like the type constraint (see \ref{subsec_type}) except the statement chains are required on the value of the constrained statement instead of the constrained item.
+The value type constraint works like the type constraint (see \ref{subsec_2_type}) except the statement chains are required on the value of the constrained statement instead of the constrained item.
 It will thus not be expanded upon.
@@
 
-\subsection{Allowed units (Q21514353)}\label{subsec_allowed_units}
+\subsection{Allowed units (Q21514353)}\label{subsec_3_allowed_units}
 The allowed units constraint needs to determine inequality between all allowed units and each unit on a constrained statement.
@@
 During pre-processing all qualifiers on constrained statements will be extracted and the inequality of their properties and the allowed qualifiers established.
-We will use the notation for set inequality as introduced in section \ref{subsec_allowed_units}.
+We will use the notation for set inequality as introduced in section \ref{subsec_3_allowed_units}.
 With \(A\) the set of allowed qualifiers the rule is:
@@
 \(\forall X \subset S, Y = S \setminus X\) the following rule has to be added.
-Note that \(qualifierEDB(S, {X}, [X])\) denotes one atom for each element \(x \in X\) as introduced in section \ref{subsec_allowed_units} with \([X]\) representing a variable which is distinct for each atom unless they have the same \(x\).
+Note that \(qualifierEDB(S, {X}, [X])\) denotes one atom for each element \(x \in X\) as introduced in section \ref{subsec_3_allowed_units} with \([X]\) representing a variable which is distinct for each atom unless they have the same \(x\).
 With the predicate does\_not\_have\([2]\) linking a statement ID to the separator it does not have via \(\text{does\_not\_have}(S, R) \leftarrow \text{last\_qualifier}(S, P, V) \wedge \text{require}(S, P, V, R)\) for brevity the rules are

### 2018-08-01 thesis/paper/paper.tex
Added mandatory qualifier

@@
 
-\subsubsection{With separators}
+\subsubsection{With separators}\label{subsubsec_with_separators}
 To find all violations under separators a number of things have to be considered.
@@
 
+\subsection{Mandatory qualifier (Q21510856)}
+The mandatory qualifier constraint needs to determine qualifier non-existence on constrained statements to find violations, which is similar to section \ref{subsubsec_with_separators}.
+All constrained statements and their qualifiers have to be extracted and the order of the qualifiers pre-processed.
+Inequality must be established between all qualifier properties and the conjunction is \(t = \text{qualifierEDB}(S, P, V) \wedge \text{unequal}(P, \text{requiredPropertyConstant})\).
+With the usual require-pattern for qualifiers and the does\_not\_have rule added for brevity the violations would be
+\begin{equation*}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+&\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \text{does\_not\_have}(S, \text{requiredPropertyConstant})
+\end{split}
+\end{equation*}
+
 \subsection{Single best value (Q52060874)}

### 2018-08-01 thesis/paper/paper.tex
Added single best value

@@
 
-\subsection{Single value (Q19474404)}
+\subsection{Single value (Q19474404)}\label{subsec_3_single_value}
 The single value constraint can be divided into two different cases, depending on if the set of separators \(S\) is empty or not.
@@
 \subsection{Single best value (Q52060874)}
-The single best value constraint limits the number of constrained statements with rank preferred on the same item to one.
-This means during pre-processing the rank information per triple must be extracted and attached to the statement with the new predicate rank\([2]\) which links the statement ID and the rank.
+The single best value constraint, like the single value constraint (section \ref{subsec_3_single_value}), differentiates between the set of separators \(S\) being empty or not.
+Since the constraint depends on the rank of statements this needs to be extracted during pre-processing and attached to the statement ID with the new predicate rank\([2]\).
 Qualifiers and references are not affected since they do not have ranks.
-Inequality is established between all constrained statement IDs.
-For the simple case without separators the rule is:
-\todo[inline]{Well, this isn't done.}
+For this constraint all item statements, their rank and their qualifiers need to be extracted and inequality established between the statement IDs and the qualifier values.\\
+
+\subsubsection{No separators}
+If the separator set is empty, the violation simply occurs if there are two constrained statements on the same item with the preferred rank.
+Thus the equation \ref{eq_no_separators_triple} is expanded to
+\begin{equation}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+&\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \\
+&\text{rank}(S, \text{preferredRankConstant}) \wedge \\
+&\text{tripleEDB}(O, I, \text{propertyConstant}, X) \wedge \\
+&\text{rank}(O, \text{preferredRankConstant}) \wedge \\
+&\text{unequal}(S, O)
+\end{split}
+\end{equation}
+
+\subsubsection{With separators}
+As in the case above, the qualifier non-existence and violation rules are very similar to single value. The only change is the addition of the rank requirements as shown below.
+\begin{equation*}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+&\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \\
+&\text{rank}(S, \text{preferredRankConstant}) \wedge \\
+&\text{tripleEDB}(O, I, \text{propertyConstant}, W) \wedge \\
+&\text{rank}(O, \text{preferredRankConstant}) \wedge \\
+&\text{unequal}(S, O) \wedge \\
+&\text{qualifierEDB}(S, \{X\}, [X]) \wedge \\
+&\text{qualifierEDB}(O, \{X\}, [X]) \wedge \\
+&\text{does\_not\_have}(S, \{Y\}) \wedge \\
+&\text{does\_not\_have}(O, \{Y\})
+\end{split}
+\end{equation*}
 

### 2018-08-01 thesis/paper/paper.tex
Added comments

@@
 \usepackage{adjustbox}
+
+\setcounter{tocdepth}{1}
 
@@
 If separators were specified a violation would be the two triples and all their separator qualifiers equal or non-existent.
+\todo{Since I found the details via looking at the source code, should/can I cite that?}
 
@@
 
+\subsection{Type (Q21503250)}
+\todo[inline]{Well.
+I wouldn't know.
+I'd have to expand the require pattern to search up the subclass tree and if require arrives at the root the class is not there.
+But a) its not a tree, so that does not work and b) there is no single root.\\
+Alternatively I could compute each items class via transitivity aaaand....
+well.
+Violation if items classes unequal required class, but how to write that?
+Well, maybe I'll get a brilliant idea.
+Or I'll stop being stupid in case this does not require a brilliant idea.}
+
+\subsection{Value type (Q21510865)}
+\todo[]{as above, so below. no wait...}
+
 \chapter{Evaluation/Comparison}

### 2018-08-02 thesis/paper/paper.tex
Changed require-pattern to two specific require-predicates

@@
 Using this pre-processed data the approach is to apply a require-predicate to each statement or qualifier.
-They will be introduced in the respective constraint sections since they may need to carry different information.
-For the introduction the predicate require\([1]\) simply denotes that the statement requires something.
-The qualifier variant works analogous.\\
-
-The pattern consists of two rules, both utilising the same conjunction \(t = \text{tripleEDB}(S, I, P, V) \wedge C_1 ..
-C_m\) with arbitrary atoms \(C_n\) which is used to test if the triple does not fulfil the requirement.
+This is require\([2]\) for statements, connecting the statement ID with a required term \(t\) and require\([4]\) taking the statement ID, property, and value of a qualifier and linking this with the term \(t\).\\
+
+For statements the pattern consists of two rules, both utilising the same conjunction \(c = \text{tripleEDB}(S, I, P, V) \wedge C_1 \wedge ..
+\wedge C_m\) with arbitrary atoms \(C_n\) which is used to test if the triple does not fulfil the requirement.
 Then rule one
 \begin{equation*}
-\text{require}(S) \leftarrow \text{first}(S) \wedge t
+\text{require}(S, t) \leftarrow \text{first}(S) \wedge c
 \end{equation*}
@@
 \begin{equation*}
-\text{require}(O) \leftarrow \text{next}(S, O) \wedge \text{require}(S) \wedge t
-\end{equation*}
-propagates this down the statement chain. If the last statement of an item or the last qualifier of a statement has require the violation occurred.\\
-
-In the following constraints only the conjunction \(t\) and adjustments to this pattern will be described. The require-propagation is regarded as given.
+\text{require}(O, t) \leftarrow \text{next}(S, O) \wedge \text{require}(S, t) \wedge c
+\end{equation*}
+propagates this down the statement chain.
+
+For qualifiers the conjunction has to be adapted to \(c_q = \text{qualifierEDB}(S, P, V) \wedge C_1 \wedge .. \wedge C_m \) and the rules are
+\begin{equation*}
+\text{require}(S, P, V, t) \leftarrow \text{first}(S, P, V) \wedge c
+\end{equation*}
+\begin{equation*}
+\text{require}(O, X, C, t) \leftarrow \text{next}(S, P, V, O, X, C) \wedge \text{require}(S, P, V, t) \wedge c
+\end{equation*}
+
+If the last statement of an item or the last qualifier of a statement has require the violation occurred.
+In the following constraints only the conjunction \(c\) and adjustments to this pattern will be described.
+The require-propagation is regarded as given.
 
@@
 It occurs in two varieties, depending on if allowed values have been specified or not.
-The predicate is defined as require\([2]\) denoting the statement and the required property.
-Without separators the conjunction is \(t_1 = \text{tripleEDB}(S, I, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)\).\\
-
-If a set of allowed values \(A\) has been specified, a second set of require-rules have to be introduced with \(t_2 = \text{tripleEDB}(S, I, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)\).
+The required term \(t\) is the  required property.
+Without separators the conjunction is \(c_1 = \text{tripleEDB}(S, I, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)\).\\
+
+If a set of allowed values \(A\) has been specified, a second set of require-rules have to be introduced with \(c_2 = \text{tripleEDB}(S, I, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)\).
 Note that the first set of rules is still required, since the preceding statements do not need to have the required property.\\
@@
 \begin{equation*}
-t_1 = \text{tripleEDB}(S, I, P, V) \wedge \text{unequal}(\text{inversePropertyConstant}, P)
-\end{equation*}
-\begin{equation*}
-\begin{split}
-t_2 = &\text{tripleEDB}(O, R, \text{propertyConstant}, X) \wedge \\
+c_1 = \text{tripleEDB}(S, I, P, V) \wedge \text{unequal}(\text{inversePropertyConstant}, P)
+\end{equation*}
+\begin{equation*}
+\begin{split}
+c_2 = &\text{tripleEDB}(O, R, \text{propertyConstant}, X) \wedge \\
 &\text{tripleEDB}(S, I, \text{inversePropertyConstant}, V) \wedge \\
@@
 Alternatively, it could be said that if the item requires a second statement, the constraint is violated.
-Pre-processing extracts all items statements and inequality between all item statement properties as usual and the conjunction is \(t = \text{tripleEDB}(S, I, P, V) \wedge \text{unequal}(\text{propertyConstant}, P)\).
+Pre-processing extracts all items statements and inequality between all item statement properties as usual and the conjunction is \(c = \text{tripleEDB}(S, I, P, V) \wedge \text{unequal}(\text{propertyConstant}, P)\).
 The predicate require\_second\([2]\) is introduced with the same meaning as require except that it will be applied after a required property has been found:
@@
 Only triples are eligible, since qualifiers cannot have further qualifier, and the order of qualifiers on constrained statements needs to be pre-processed as well since qualifier non-existence is relevant.
-The predicate require\([4]\) is introduced which notes the statement ID, property, and value of a qualifier as well as the separator.
-The pattern needs to be applied for every separator, meaning the conjunction is: \(t = \text{qualifierEDB}(S, P, V) \wedge \text{unequal}(P, \text{requiredPropertyConstant})\) for each separator in \(S\).\\
+The pattern needs to be applied for every separator, meaning the conjunction is: \(c_q = \text{qualifierEDB}(S, P, V) \wedge \text{unequal}(P, \text{separatorConstant})\) for each separator in \(S\).\\
 
@@
 All constrained statements and their qualifiers have to be extracted and the order of the qualifiers pre-processed.
-Inequality must be established between all qualifier properties and the conjunction is \(t = \text{qualifierEDB}(S, P, V) \wedge \text{unequal}(P, \text{requiredPropertyConstant})\).
+Inequality must be established between all qualifier properties and the conjunction is \(c_q = \text{qualifierEDB}(S, P, V) \wedge \text{unequal}(P, \text{requiredPropertyConstant})\).
 With the usual require-pattern for qualifiers and the does\_not\_have rule added for brevity the violations would be

### 2018-08-03 src/main/java/main/Main.java
Added multiple checkers to main

@@ -25,9 +25,15 @@ import org.wikidata.wdtk.dumpfiles.EntityTimerProcessor;
 import org.wikidata.wdtk.dumpfiles.MwDumpFile;
 import org.wikidata.wdtk.dumpfiles.MwLocalDumpFile;
 
+import impl.CC.AllowedEntityTypesCC;
+import impl.CC.ConflictsWithCC;
 import impl.CC.ConstraintChecker;
 import impl.CC.DistinctValuesCC;
 import impl.CC.ItemRequiresStatementCC;
+import impl.CC.NoneOfCC;
+import impl.CC.ScopeCC;
+import impl.CC.SingleValueCC;
+import impl.CC.TypeCC;
 
 /**
  * @author adrian
@@ -81,7 +87,13 @@ public class Main {
 		
 
 		List<ConstraintChecker> checkers = new ArrayList<ConstraintChecker>();
-		checkers.add(new ItemRequiresStatementCC());
+		//checkers.add(new ItemRequiresStatementCC());
+		//checkers.add(new AllowedEntityTypesCC());
+		checkers.add(new ConflictsWithCC());
+		checkers.add(new DistinctValuesCC());
+		checkers.add(new NoneOfCC());
+		checkers.add(new ScopeCC());
+		checkers.add(new SingleValueCC());
 		try {
 			for (ConstraintChecker constraintChecker : checkers) {
 				constraintChecker.init();

### 2018-08-03 src/main/java/utility/OutputValueVisitor.java
Workaround for null-pointer-exception for QuantityValue

@@ -37,7 +37,19 @@ public class OutputValueVisitor implements ValueVisitor<String> {
 
 	@Override
 	public String visit(QuantityValue value) {
-		String result = value.getNumericValue().toString() + " Lower: " + value.getLowerBound().toString() + " Upper: " + value.getUpperBound().toString() + value.getUnit();
+		String numericValue = "NaN";
+		if (value.getNumericValue() != null)
+			numericValue = value.getNumericValue().toString();
+		String lowerBound = "NaN";
+		if (value.getLowerBound() != null)
+			lowerBound = value.getLowerBound().toString();
+		String upperBound = "NaN";
+		if (value.getUpperBound() != null)
+			upperBound = value.getUpperBound().toString();
+		String unit = "None";
+		if (value.getUnit() != null)
+			unit = value.getUnit().toString();
+		String result = numericValue + " Lower: " + lowerBound + " Upper: " + upperBound + " Unit: " + unit;
 		return result;
 	}
 

### 2018-08-03 thesis/paper/paper.tex
Corrected require-pattern

@@
 
-For statements the pattern consists of two rules, both utilising the same conjunction \(c = \text{tripleEDB}(S, I, P, V) \wedge C_1 \wedge ..
-\wedge C_m\) with arbitrary atoms \(C_n\) which is used to test if the triple does not fulfil the requirement.
+For statements the pattern consists of two rules, both utilising the same conjunction \(c = C_1 \wedge ..
+\wedge C_m\) with arbitrary atoms \(C_n\) which is used to test if the statement does not fulfil the requirement.
 Then rule one
 \begin{equation*}
-\text{require}(S, t) \leftarrow \text{first}(S) \wedge c
+\text{require}(S, t) \leftarrow \text{first}(S, I) \wedge c
 \end{equation*}
@@
 
-For qualifiers the conjunction has to be adapted to \(c_q = \text{qualifierEDB}(S, P, V) \wedge C_1 \wedge .. \wedge C_m \) and the rules are
+For qualifiers the rules are
 \begin{equation*}
@@
 Only triples are eligible, since qualifiers cannot have further qualifier, and the order of qualifiers on constrained statements needs to be pre-processed as well since qualifier non-existence is relevant.
-The pattern needs to be applied for every separator, meaning the conjunction is: \(c_q = \text{qualifierEDB}(S, P, V) \wedge \text{unequal}(P, \text{separatorConstant})\) for each separator in \(S\).\\
+The pattern needs to be applied for every separator, meaning the conjunction is: \(c = \text{qualifierEDB}(S, P, V) \wedge \text{unequal}(P, \text{separatorConstant})\) for each separator in \(S\).\\
 
@@
 All constrained statements and their qualifiers have to be extracted and the order of the qualifiers pre-processed.
-Inequality must be established between all qualifier properties and the conjunction is \(c_q = \text{qualifierEDB}(S, P, V) \wedge \text{unequal}(P, \text{requiredPropertyConstant})\).
+Inequality must be established between all qualifier properties and the conjunction is \(c = \text{qualifierEDB}(S, P, V) \wedge \text{unequal}(P, \text{requiredPropertyConstant})\).
 With the usual require-pattern for qualifiers and the does\_not\_have rule added for brevity the violations would be

### 2018-08-07 src/main/java/impl/CC/AllowedEntityTypesCC.java
FIX: intialise sparql result container before use

@@ -40,6 +40,11 @@ public class AllowedEntityTypesCC extends ConstraintChecker {
 		super("Q52004125");
 		tripleSet = new AllowedEntityTypesTS(allowedEntityTypes.keySet());
 	}
+	
+	@Override
+	void initDataField() {
+		allowedEntityTypes = new HashMap<String, HashSet<String>>();
+	}
 
 	@Override
 	protected Set<String> qualifiers() {
@@ -53,7 +58,6 @@ public class AllowedEntityTypesCC extends ConstraintChecker {
 
 	@Override
 	protected void process(QuerySolution solution) {
-		allowedEntityTypes = new HashMap<String, HashSet<String>>();
 		String property = Utility.addBaseURI(solution.get("item").asResource().getLocalName());
 		
 		if (!allowedEntityTypes.containsKey(property))

### 2018-08-07 src/main/java/impl/CC/AllowedUnitsCC.java

@@ -42,6 +42,11 @@ public class AllowedUnitsCC extends ConstraintChecker {
 	}
 
 	@Override
+	void initDataField() {
+		allowedUnits = new HashMap<String, HashSet<String>>();
+	}
+	
+	@Override
 	protected Set<String> qualifiers() {
 		return asSet();
 	}
@@ -53,7 +58,6 @@ public class AllowedUnitsCC extends ConstraintChecker {
 
 	@Override
 	protected void process(QuerySolution solution) {
-		allowedUnits = new HashMap<String, HashSet<String>>();
 		String property = Utility.addBaseURI(solution.get("item").asResource().getLocalName());
 		
 		if (!allowedUnits.containsKey(property))

### 2018-08-07 src/main/java/impl/CC/ConflictsWithCC.java

@@ -37,6 +37,11 @@ public class ConflictsWithCC extends ConstraintChecker {
 		super("Q21502838");
 		tripleSet = new ConflictsWithTS(configuration.keySet());
 	}
+	
+	@Override
+	void initDataField() {
+		configuration = new HashMap<String, HashMap<String, HashSet<String>>>();
+	}
 
 	@Override
 	protected Set<String> qualifiers() {
@@ -50,7 +55,6 @@ public class ConflictsWithCC extends ConstraintChecker {
 
 	@Override
 	protected void process(QuerySolution solution) {
-		configuration = new HashMap<String, HashMap<String, HashSet<String>>>();
 		String property = Utility.addBaseURI(solution.get("item").asResource().getLocalName());
 		
 		if (!configuration.containsKey(property))

### 2018-08-07 src/main/java/impl/CC/ConstraintChecker.java

@@ -59,8 +59,11 @@ public abstract class ConstraintChecker {
 		constraint = constraint_;
 		internalError = "INTERNAL_ERROR for constraint " + constraint + ".";
 		reasoner.setAlgorithm(Algorithm.RESTRICTED_CHASE);
+		initDataField();
 		init();
 	}
+	
+	abstract void initDataField();
 
 	public void init() throws IOException {
 		Set<String> qualifiers = qualifiers();
@@ -107,7 +110,7 @@ public abstract class ConstraintChecker {
 		while (results.hasNext()) {
 			QuerySolution solution = results.next();
 			//String property = solution.get("item").asResource().getLocalName();
-			//if (!(property.equals("P6"))) 
+			//if (!(property.equals("P103"))) 
 			//	continue;
 
 			process(solution);

### 2018-08-07 src/main/java/impl/CC/DistinctValuesCC.java

@@ -2,6 +2,7 @@ package impl.CC;
 
 import java.io.IOException;
 import java.util.ArrayList;
+import java.util.HashMap;
 import java.util.HashSet;
 import java.util.List;
 import java.util.Set;
@@ -38,13 +39,17 @@ public class DistinctValuesCC extends ConstraintChecker {
 	}
 
 	@Override
+	void initDataField() {
+		properties = new HashSet<String>();
+	}
+	
+	@Override
 	protected Set<String> concatQualifiers() {
 		return asSet();
 	}
 
 	@Override
 	protected void process(QuerySolution solution) {
-		properties = new HashSet<String>();
 		String property = Utility.addBaseURI(solution.get("item").asResource().getLocalName());
 		properties.add(property);
 	}

### 2018-08-07 src/main/java/impl/CC/NoneOfCC.java

@@ -40,6 +40,11 @@ public class NoneOfCC extends ConstraintChecker {
 	}
 
 	@Override
+	void initDataField() {
+		notAllowedValues = new HashMap<String, HashSet<String>>();
+	}
+	
+	@Override
 	protected Set<String> qualifiers() {
 		return asSet();
 	}
@@ -51,7 +56,6 @@ public class NoneOfCC extends ConstraintChecker {
 
 	@Override
 	protected void process(QuerySolution solution) {
-		notAllowedValues = new HashMap<String, HashSet<String>>();
 		String property = Utility.addBaseURI(solution.get("item").asResource().getLocalName());
 		
 		if (!notAllowedValues.containsKey(property))

### 2018-08-07 src/main/java/impl/CC/OneOfCC.java

@@ -41,6 +41,11 @@ public class OneOfCC extends ConstraintChecker {
 	}
 
 	@Override
+	void initDataField() {
+		allowedValues = new HashMap<String, HashSet<String>>();
+	}
+	
+	@Override
 	protected Set<String> qualifiers() {
 		return asSet();
 	}
@@ -52,7 +57,6 @@ public class OneOfCC extends ConstraintChecker {
 
 	@Override
 	protected void process(QuerySolution solution) {
-		allowedValues = new HashMap<String, HashSet<String>>();
 		String property = Utility.addBaseURI(solution.get("item").asResource().getLocalName());
 		
 		if (!allowedValues.containsKey(property))

### 2018-08-07 src/main/java/impl/CC/OneOfQualifierValueCC.java

@@ -44,6 +44,11 @@ public class OneOfQualifierValueCC extends ConstraintChecker {
 	}
 
 	@Override
+	void initDataField() {
+		qualifiersAndValues = new HashMap<String, Map<String, Set<String>>>();
+	}
+	
+	@Override
 	protected Set<String> qualifiers() {
 		return asSet(QUALIFIER_PROPERTY);
 	}
@@ -55,7 +60,6 @@ public class OneOfQualifierValueCC extends ConstraintChecker {
 
 	@Override
 	protected void process(QuerySolution solution) {
-		qualifiersAndValues = new HashMap<String, Map<String, Set<String>>>();
 		String property = Utility.addBaseURI(solution.get("item").asResource().getLocalName());
 		
 		if (!qualifiersAndValues.containsKey(property))

### 2018-08-07 src/main/java/impl/CC/ScopeCC.java

@@ -43,6 +43,11 @@ public class ScopeCC extends ConstraintChecker {
 		tripleSet = new ScopeTS(propertiesAndScopes.keySet());
 	}
 
+	@Override
+	void initDataField() {
+		propertiesAndScopes = new HashMap<String, HashSet<String>>();
+	}
+	
 	protected Set<String> qualifiers() {
 		return asSet();
 	}
@@ -52,7 +57,6 @@ public class ScopeCC extends ConstraintChecker {
 	}
 	
 	protected void process(QuerySolution solution) {
-		propertiesAndScopes = new HashMap<String, HashSet<String>>();
 		String property = Utility.addBaseURI(solution.get("item").asResource().getLocalName());
 		HashSet<String> qualifiers = new HashSet<String>();
 		

### 2018-08-07 src/main/java/impl/CC/SingleValueCC.java

@@ -29,6 +29,11 @@ public class SingleValueCC extends ConstraintChecker {
 	}
 
 	@Override
+	void initDataField() {
+		propertiesAndSeparators = new HashMap<String, HashSet<String>>();
+	}
+	
+	@Override
 	protected Set<String> qualifiers() {
 		return asSet();
 	}
@@ -40,7 +45,6 @@ public class SingleValueCC extends ConstraintChecker {
 
 	@Override
 	protected void process(QuerySolution solution) {
-		propertiesAndSeparators = new HashMap<String, HashSet<String>>();
 		String property = solution.get("item").asResource().getLocalName();
 		
 		if (!propertiesAndSeparators.containsKey(property))

### 2018-08-07 src/main/java/main/Main.java

@@ -103,8 +103,8 @@ public class Main {
 			//checkers.add(new AllowedUnitsCC());
 			//checkers.add(new AllowedQualifiersCC());
 			//checkers.add(new OneOfCC());
-			checkers.add(new OneOfQualifierValueCC());
-			//checkers.add(new ItemRequiresStatementCC());
+			//checkers.add(new OneOfQualifierValueCC());
+			checkers.add(new ItemRequiresStatementCC());
 			//checkers.add(new SingleValueCC());
 		} catch (IOException e) {
 			logger.error("Could not open a file, see the error message for details.", e);

### 2018-08-07 src/main/java/impl/CC/ItemRequiresStatementCC.java
Changed Item requires statement

@@ -12,9 +12,34 @@ import org.apache.jena.query.QuerySolution;
 import org.apache.jena.rdf.model.Literal;
 import org.apache.jena.rdf.model.RDFNode;
 import org.apache.log4j.Logger;
+import org.semanticweb.vlog4j.core.model.api.Atom;
+import org.semanticweb.vlog4j.core.model.api.Constant;
+import org.semanticweb.vlog4j.core.model.api.Term;
+import org.semanticweb.vlog4j.core.model.implementation.Expressions;
+import org.semanticweb.vlog4j.core.reasoner.DataSource;
+import org.semanticweb.vlog4j.core.reasoner.exceptions.ReasonerStateException;
+import org.semanticweb.vlog4j.core.reasoner.implementation.CsvFileDataSource;
 
 import impl.PCC.ItemRequiresStatementPCC;
 import impl.PCC.PropertyConstraintChecker;
+import impl.TS.ItemRequiresStatementTS;
+import utility.InequalityHelper;
+import utility.StatementNonExistenceHelper;
+import utility.Utility;
+
+import static utility.SC.violation_triple_query;
+
+import static utility.SC.first;
+import static utility.SC.next;
+import static utility.SC.last;
+import static utility.SC.tripleEDB;
+
+import static utility.SC.s;
+import static utility.SC.i;
+import static utility.SC.p;
+import static utility.SC.v;
+
+import static utility.SC.require;
 
 public class ItemRequiresStatementCC extends ConstraintChecker {
 	
@@ -23,12 +48,20 @@ public class ItemRequiresStatementCC extends ConstraintChecker {
 	public static final String REQUIRED_PROPERTY = "P2306";
 	public static final String ALLOWED_VALUE = "P2305";
 	
-	Map<String, HashMap<String, HashSet<String>>> configuration;
+	Map<String, Map<String, Set<String>>> configuration;
+	
+	final ItemRequiresStatementTS tripleSet;
 
-	public ItemRequiresStatementCC() {
+	public ItemRequiresStatementCC() throws IOException {
 		super("Q21503247");
+		tripleSet = new ItemRequiresStatementTS(configuration.keySet());
 	}
 
+	@Override
+	void initDataField() {
+		configuration = new HashMap<String, Map<String, Set<String>>>();
+	}
+	
 	@Override
 	protected Set<String> qualifiers() {
 		return asSet(REQUIRED_PROPERTY);
[…]

### 2018-08-07 src/main/java/impl/PCC/ItemRequiresStatementPCC.java

@@ -2,7 +2,6 @@ package impl.PCC;
 
 import java.io.IOException;
 import java.util.ArrayList;
-import java.util.Arrays;
 import java.util.HashSet;
 import java.util.List;
 import java.util.Map;
@@ -13,105 +12,61 @@ import org.semanticweb.vlog4j.core.model.api.Constant;
 import org.semanticweb.vlog4j.core.model.api.Rule;
 import org.semanticweb.vlog4j.core.model.api.Term;
 import org.semanticweb.vlog4j.core.model.implementation.Expressions;
-import org.semanticweb.vlog4j.core.reasoner.DataSource;
-import org.semanticweb.vlog4j.core.reasoner.exceptions.ReasonerStateException;
-import org.semanticweb.vlog4j.core.reasoner.implementation.CsvFileDataSource;
 
-import impl.TS.ItemRequiresStatementTS;
-import impl.TS.TripleSet;
 import utility.InequalityHelper;
-import utility.PrepareQueriesException;
 import utility.StatementNonExistenceHelper;
 import utility.Utility;
 
+import static utility.SC.tripleEDB;
+import static utility.SC.v;
+import static utility.SC.last;
+import static utility.SC.require;
+import static utility.SC.s;
+import static utility.SC.o;
+import static utility.SC.i;
+import static utility.SC.p;
+import static utility.SC.x;
+
 public class ItemRequiresStatementPCC extends PropertyConstraintChecker {
 	
-	final Map<String, HashSet<String>> configuration;
-	
-	final ItemRequiresStatementTS tripleSet;
+	final Map<String, Set<String>> configuration;
 
-	public ItemRequiresStatementPCC(String property_, Map<String, HashSet<String>> configuration_) throws IOException {
+	public ItemRequiresStatementPCC(String property_, Map<String, Set<String>> configuration_) throws IOException {
 		super(property_);
 		configuration = configuration_;
-		tripleSet = new ItemRequiresStatementTS(property);
 	}
 
 	@Override
-	public String violations() throws IOException {
-		if (!tripleSet.notEmpty())
-			return "";
-		
-		// Loading EDB-facts
-		try {
-			loadTripleSets(tripleSet);
-			if (tripleSet.firstNotEmpty()) {
-				final DataSource firstEDBPath = new CsvFileDataSource(tripleSet.getFirstFile());
[…]

### 2018-08-07 src/main/java/impl/TS/ItemRequiresStatementTS.java

@@ -17,15 +17,17 @@ public class ItemRequiresStatementTS extends TripleSet {
 	Set<String> allProperties = new HashSet<String>();
 	Set<String> allValues = new HashSet<String>();
 
-	TripleSetFile first = new TripleSetFile(getTripleSetType(), property + "first");
-	TripleSetFile next = new TripleSetFile(getTripleSetType(), property + "next");
-	TripleSetFile last = new TripleSetFile(getTripleSetType(), property + "last");
+	TripleSetFile first = new TripleSetFile(getTripleSetType(), "first");
+	TripleSetFile next = new TripleSetFile(getTripleSetType(), "next");
+	TripleSetFile last = new TripleSetFile(getTripleSetType(), "last");
 	
 	String lastID;
 	String lastSubject;
+	
+	Set<String> properties;
 
-	public ItemRequiresStatementTS(String property_) throws IOException {
-		super(property_);
+	public ItemRequiresStatementTS(Set<String> properties_) throws IOException {
+		properties = properties_;
 	}
 	
 	@Override
@@ -76,7 +78,7 @@ public class ItemRequiresStatementTS extends TripleSet {
 			if (foundProperty)
 				break;
 			String predicate = sg.getProperty().getIri();
-			if (predicate.endsWith(property)) {
+			if (properties.contains(predicate)) {
 				foundProperty = true;
 				break;
 			}
@@ -130,11 +132,10 @@ public class ItemRequiresStatementTS extends TripleSet {
 		super.close();
 		first.close();
 		next.close();
-		if (!last.isClosed()) {
+		if (!last.isClosed() && lastID != null) {
 			last.write(lastID, lastSubject);
-			last.close();
 		}
-			
+		last.close();
 	}
 
 	@Override

### 2018-08-07 src/main/java/utility/StatementNonExistenceHelper.java

@@ -27,19 +27,13 @@ import static utility.SC.v;
 import static utility.SC.c;
 
 
-public class StatementNonExistenceHelper {
+public class StatementNonExistenceHelper {	
 	
-	static Reasoner reasoner;	
-	
-	public static void setOrReset(Reasoner reasoner_) {
-		reasoner = reasoner_;
+	public static List<Rule> initRequireTriple(Term requiredTerm, Atom...conjunctionAtoms) {
+		return initRequireTriple(requiredTerm, Arrays.asList(conjunctionAtoms));
 	}
 	
-	public static void initRequireTriple(Term requiredTerm, Atom...conjunctionAtoms) throws ReasonerStateException {
-		initRequireTriple(requiredTerm, Arrays.asList(conjunctionAtoms));
-	}
-	
-	public static void initRequireTriple(Term requiredTerm, List<Atom> conjunctionAtoms) throws ReasonerStateException {
+	public static List<Rule> initRequireTriple(Term requiredTerm, List<Atom> conjunctionAtoms) {
 		// require(S, requiredTerm)
 		Atom require_Sr = Expressions.makeAtom(require, s, requiredTerm);
 		
@@ -54,28 +48,28 @@ public class StatementNonExistenceHelper {
 		// require(S, requiredTerm) :- first(S, I), conjunctionAtoms
 		Rule firstRequire = Expressions.makeRule(require_Sr, toArray(firstConjunctionAtoms));
 		
+		// next(O, S)
+		Atom next_OS = Expressions.makeAtom(next, o, s);
+		
 		// require(O, requiredTerm)
 		Atom require_Or = Expressions.makeAtom(require, o, requiredTerm);
 		
-		// next(S, O)
-		Atom next_SO = Expressions.makeAtom(next, s, o);
-		
 		List<Atom> nextConjunctionAtoms = new ArrayList<Atom>();
-		nextConjunctionAtoms.add(next_SO);
-		nextConjunctionAtoms.add(require_Sr);
+		nextConjunctionAtoms.add(next_OS);
+		nextConjunctionAtoms.add(require_Or);
 		nextConjunctionAtoms.addAll(conjunctionAtoms);
 		
-		// require(O, requiredTerm) :- next(S, O), require(S, requiredTerm), conjunctionAtoms
-		Rule nextRequire = Expressions.makeRule(require_Or, toArray(nextConjunctionAtoms));
[…]

### 2018-08-08 src/main/java/main/Main.java
FIX: Typo

@@ -125,7 +125,7 @@ public class Main {
 				case "allowedunits":
 					checkers.add(new AllowedUnitsCC());
 					break;
-				case "allwoedqualifiers":
+				case "allowedqualifiers":
 					checkers.add(new AllowedQualifiersCC());
 					break;
 				case "oneof":

### 2018-08-08 thesis/paper/paper.tex
Fixed variabels in require pattern to allow for conjunction-reuse
FIX: Multi-Value rules

@@
 \begin{equation*}
-\text{require}(O, t) \leftarrow \text{next}(S, O) \wedge \text{require}(S, t) \wedge c
+\text{require}(S, t) \leftarrow \text{next}(O, S) \wedge \text{require}(O, t) \wedge c
 \end{equation*}
@@
 \begin{equation*}
-\text{require}(O, X, C, t) \leftarrow \text{next}(S, P, V, O, X, C) \wedge \text{require}(S, P, V, t) \wedge c
+\text{require}(S, P, V, t) \leftarrow \text{next}(O, X, C, S, P, V) \wedge \text{require}(O, X, C, t) \wedge c
 \end{equation*}
@@
 \begin{equation*}
-\text{require\_second}(S, \text{propertyConstant}) \leftarrow \text{next}(Q, R) \wedge \text{next}(R, S) \wedge \text{require}(Q) \wedge t
-\end{equation*}
-\begin{equation*}
-\text{require\_second}(S, \text{propertyConstant}) \leftarrow \text{next}(R, S) \wedge \text{require\_second}(R) \wedge t
-\end{equation*}
-A violations would be any last statement requiring a second constrained statement.
+\begin{split}
+\text{require\_second}&(S, \text{propertyConstant}) \leftarrow \\
+&\text{next}(Q, R) \wedge \text{next}(R, S) \wedge \text{require}(Q, \text{propertyConstant}) \wedge t \wedge \\
+&\text{tripleEDB}(R, I, \text{propertyConstant}, C)
+\end{split}
+\end{equation*}
+\begin{equation*}
+\text{require\_second}(S, \text{propertyConstant}) \leftarrow \text{next}(R, S) \wedge \text{require\_second}(R, \text{propertyConstant}) \wedge t
+\end{equation*}
+A violations would be any last statement requiring a second constrained statement.\\
 

### 2018-08-09 src/main/java/impl/CC/AllowedQualifiersCC.java
Implemented different inequality modes

@@ -88,7 +88,7 @@ public class AllowedQualifiersCC extends ConstraintChecker {
 		for (Set<String> qualifierSet : allowedQualifiers.values()) {
 			qualifiers.addAll(qualifierSet);
 		}
-		InequalityHelper.addUnequalConstantsToReasoner(qualifiers);
+		InequalityHelper.establishInequality(qualifiers);
 	}
 
 	@Override

### 2018-08-09 src/main/java/impl/CC/AllowedUnitsCC.java

@@ -93,7 +93,7 @@ public class AllowedUnitsCC extends ConstraintChecker {
 		for (Set<String> unitsSet : allowedUnits.values()) {
 			units.addAll(unitsSet);
 		}
-		InequalityHelper.addUnequalConstantsToReasoner(units);
+		InequalityHelper.establishInequality(units);
 	}
 
 	@Override

### 2018-08-09 src/main/java/impl/CC/ConstraintChecker.java

@@ -8,6 +8,7 @@ import java.util.ArrayList;
 import java.util.Arrays;
 import java.util.HashSet;
 import java.util.List;
+import java.util.Map;
 import java.util.Set;
 
 import org.apache.jena.query.Query;
@@ -19,9 +20,12 @@ import org.apache.jena.query.ResultSet;
 import org.apache.jena.riot.web.HttpOp;
 import org.apache.log4j.Logger;
 import org.semanticweb.vlog4j.core.model.api.Atom;
+import org.semanticweb.vlog4j.core.model.api.Conjunction;
 import org.semanticweb.vlog4j.core.model.api.QueryResult;
 import org.semanticweb.vlog4j.core.model.api.Rule;
 import org.semanticweb.vlog4j.core.model.api.Term;
+import org.semanticweb.vlog4j.core.model.api.Variable;
+import org.semanticweb.vlog4j.core.model.implementation.Expressions;
 import org.semanticweb.vlog4j.core.reasoner.Algorithm;
 import org.semanticweb.vlog4j.core.reasoner.DataSource;
 import org.semanticweb.vlog4j.core.reasoner.Reasoner;
@@ -34,8 +38,12 @@ import org.semanticweb.vlog4j.core.reasoner.implementation.QueryResultIterator;
 import impl.PCC.PropertyConstraintChecker;
 import impl.TS.TripleSet;
 import main.Main;
+import utility.InequalityHelper;
 import utility.PrepareQueriesException;
 import utility.SC;
+import utility.Utility;
+
+import static utility.SC.require_inequality;
 
 /**
  * @author adrian
@@ -134,12 +142,53 @@ public abstract class ConstraintChecker {
 		//for (Rule rule : rulesToAdd) {
 		//	System.out.println(rule);
 		//}
+		if (InequalityHelper.mode.equals(InequalityHelper.Mode.DEMANDED)) {
+			rulesToAdd.addAll(addRequireInequality(rulesToAdd));
+		}
+		 
 		try {
 			prepareAndExecuteQueries(rulesToAdd, queries());
 		} catch (PrepareQueriesException e) {
 			resultString += e.getMessage();
 		}
 		delete();
+	}
+	
+	List<Rule> addRequireInequality(List<Rule> rules) {
+		List<Rule> result = new ArrayList<Rule>();
+		for (Rule rule : rules) {
+			List<Atom> unequalAtoms = new ArrayList<Atom>();
+			List<Atom> otherAtoms = new ArrayList<Atom>();
[…]

### 2018-08-09 src/main/java/impl/CC/DistinctValuesCC.java

@@ -62,7 +62,7 @@ public class DistinctValuesCC extends ConstraintChecker {
 	void prepareFacts() throws ReasonerStateException, IOException {
 		loadTripleSets(tripleSet);
 		InequalityHelper.setOrReset(reasoner);
-		InequalityHelper.addUnequalConstantsToReasoner(tripleSet.getStatements());
+		InequalityHelper.establishInequality(tripleSet.getStatements());
 	}
 
 	@Override

### 2018-08-09 src/main/java/impl/CC/ItemRequiresStatementCC.java

@@ -124,8 +124,8 @@ public class ItemRequiresStatementCC extends ConstraintChecker {
 				values.addAll(set);
 			}
 		}
-		InequalityHelper.addUnequalConstantsToReasoner(properties);
-		InequalityHelper.addUnequalConstantsToReasoner(values);
+		InequalityHelper.establishInequality(properties);
+		InequalityHelper.establishInequality(values);
 		
 	}
 

### 2018-08-09 src/main/java/impl/CC/MandatoryQualifierCC.java

@@ -105,7 +105,7 @@ public class MandatoryQualifierCC extends ConstraintChecker {
 		for (Set<String> propertySet : propertiesAndQualifiers.values()) {
 			qualifierProperties.addAll(propertySet);
 		}
-		InequalityHelper.addUnequalConstantsToReasoner(qualifierProperties);
+		InequalityHelper.establishInequality(qualifierProperties);
 	}
 
 	@Override

### 2018-08-09 src/main/java/impl/CC/MultiValueCC.java

@@ -91,7 +91,7 @@ public class MultiValueCC extends ConstraintChecker {
 		// Establishing inequality
 		InequalityHelper.setOrReset(reasoner);
 
-		InequalityHelper.addUnequalConstantsToReasoner(tripleSet.getStatementProperties());
+		InequalityHelper.establishInequality(tripleSet.getStatementProperties());
 
 	}
 

### 2018-08-09 src/main/java/impl/CC/OneOfCC.java

@@ -89,7 +89,7 @@ public class OneOfCC extends ConstraintChecker {
 		for (Set<String> valuesSet : allowedValues.values()) {
 			values.addAll(valuesSet);
 		}
-		InequalityHelper.addUnequalConstantsToReasoner(values);
+		InequalityHelper.establishInequality(values);
 	}
 
 	@Override

### 2018-08-09 src/main/java/impl/CC/OneOfQualifierValueCC.java

@@ -97,7 +97,7 @@ public class OneOfQualifierValueCC extends ConstraintChecker {
 				values.addAll(entry2.getValue());
 			}
 		}
-		InequalityHelper.addUnequalConstantsToReasoner(values);
+		InequalityHelper.establishInequality(values);
 	}
 
 	@Override

### 2018-08-09 src/main/java/impl/CC/SingleValueCC.java

@@ -102,12 +102,12 @@ public class SingleValueCC extends ConstraintChecker {
 		}
 	
 		InequalityHelper.setOrReset(reasoner);
-		InequalityHelper.addUnequalConstantsToReasoner(tripleSet.getStatementIDs());
+		InequalityHelper.establishInequality(tripleSet.getStatementIDs());
 		Set<String> values = tripleSet.getQualifierValues();
 		for (Set<String> valuesSet : propertiesAndSeparators.values()) {
 			values.addAll(valuesSet);
 		}
-		InequalityHelper.addUnequalConstantsToReasoner(values);
+		InequalityHelper.establishInequality(values);
 	}
 
 	@Override

### 2018-08-09 src/main/java/impl/PCC/AllowedQualifiersPCC.java

@@ -17,6 +17,7 @@ import org.semanticweb.vlog4j.core.model.api.Rule;
 import org.semanticweb.vlog4j.core.model.implementation.Expressions;
 
 import utility.InequalityHelper;
+import utility.Utility;
 
 public class AllowedQualifiersPCC extends PropertyConstraintChecker {
 	
@@ -48,7 +49,7 @@ public class AllowedQualifiersPCC extends PropertyConstraintChecker {
 		}
 		
 		// violation_triple(S, I, propertyConstant, V) :- tripleEDB(S, I, propertyConstant, V), qualifierEDB(S, Q, O), unequal({A}, Q)
-		Rule violation = Expressions.makeRule(violation_triple_SIpV, toArray(violation_conjunction));
+		Rule violation = Expressions.makeRule(violation_triple_SIpV, Utility.toArray(violation_conjunction));
 		rules.add(violation);
 
 		return rules;

### 2018-08-09 src/main/java/impl/PCC/AllowedUnitsPCC.java

@@ -16,6 +16,7 @@ import org.semanticweb.vlog4j.core.model.api.Rule;
 import org.semanticweb.vlog4j.core.model.implementation.Expressions;
 
 import utility.InequalityHelper;
+import utility.Utility;
 
 public class AllowedUnitsPCC extends PropertyConstraintChecker {
 	
@@ -50,7 +51,7 @@ public class AllowedUnitsPCC extends PropertyConstraintChecker {
 		violation_triple_conjunction.addAll(unequal_conjunction);
 		
 		// violation_triple(S, I, propertyConstant, V) :-  tripleEDB(S, I, propertyConstant, V), unit(V, U), unequal({A}, U)
-		Rule tripleViolation = Expressions.makeRule(violation_triple_SIpV, toArray(violation_triple_conjunction));
+		Rule tripleViolation = Expressions.makeRule(violation_triple_SIpV, Utility.toArray(violation_triple_conjunction));
 		rules.add(tripleViolation);
 	
 		List<Atom> violation_qualifier_conjunction = new ArrayList<Atom>();
@@ -58,7 +59,7 @@ public class AllowedUnitsPCC extends PropertyConstraintChecker {
 		violation_qualifier_conjunction.addAll(unequal_conjunction);
 		
 		// violation_qualifier(S, propertyConstant, V) :- qualifierEDB(S, propertyConstant, V), unit(V, U), unequal({A}, U)
-		Rule qualifierViolation = Expressions.makeRule(violation_qualifier_SpV, toArray(violation_qualifier_conjunction));
+		Rule qualifierViolation = Expressions.makeRule(violation_qualifier_SpV, Utility.toArray(violation_qualifier_conjunction));
 		rules.add(qualifierViolation);
 		
 		List<Atom> violation_reference_conjunction = new ArrayList<Atom>();
@@ -66,7 +67,7 @@ public class AllowedUnitsPCC extends PropertyConstraintChecker {
 		violation_reference_conjunction.addAll(unequal_conjunction);
 		
 		// violation_reference(S, propertyConstant, V) :- referenceEDB(S, propertyConstant, V), unit(V, U), unequal({A}, U)
-		Rule referenceViolation = Expressions.makeRule(violation_reference_SpV, toArray(violation_reference_conjunction));
+		Rule referenceViolation = Expressions.makeRule(violation_reference_SpV, Utility.toArray(violation_reference_conjunction));
 		rules.add(referenceViolation);
 		
 		return rules;

### 2018-08-09 src/main/java/impl/PCC/OneOfPCC.java

@@ -14,6 +14,7 @@ import org.semanticweb.vlog4j.core.model.api.Rule;
 import org.semanticweb.vlog4j.core.model.implementation.Expressions;
 
 import utility.InequalityHelper;
+import utility.Utility;
 
 public class OneOfPCC extends PropertyConstraintChecker {
 	
@@ -44,21 +45,21 @@ public class OneOfPCC extends PropertyConstraintChecker {
 		violation_triple_conjunction.addAll(unequal_conjunction);
 
 		// violation_triple(S, I, propertyConstant, V) :- tripleEDB(S, I, propertyConstant, V), unequal({A}, V)
-		Rule violationTriple = Expressions.makeRule(violation_triple_SIpV, toArray(violation_triple_conjunction));
+		Rule violationTriple = Expressions.makeRule(violation_triple_SIpV, Utility.toArray(violation_triple_conjunction));
 		
 		List<Atom> violation_qualifier_conjunction = new ArrayList<Atom>();
 		violation_qualifier_conjunction.add(qualifierEDB_SpV);
 		violation_qualifier_conjunction.addAll(unequal_conjunction);
 		
 		// violation_qualifier(S, propertyConstant, V) :- qualifierEDB(S, propertyConstant, V), unequal({A}, V)
-		Rule violationQualifier = Expressions.makeRule(violation_qualifier_SpV, toArray(violation_qualifier_conjunction));
+		Rule violationQualifier = Expressions.makeRule(violation_qualifier_SpV, Utility.toArray(violation_qualifier_conjunction));
 		
 		List<Atom> violation_reference_conjunction = new ArrayList<Atom>();
 		violation_reference_conjunction.add(referenceEDB_SpV);
 		violation_reference_conjunction.addAll(unequal_conjunction);
 		
 		// violation_reference(S, propertyConstant, V) :- referenceEDB(S, propertyConstant, V), unequal({A}, V)
-		Rule violationReference = Expressions.makeRule(violation_qualifier_SpV, toArray(violation_reference_conjunction));
+		Rule violationReference = Expressions.makeRule(violation_qualifier_SpV, Utility.toArray(violation_reference_conjunction));
 		
 		rules.add(violationTriple);
 		rules.add(violationQualifier);

### 2018-08-09 src/main/java/impl/PCC/OneOfQualifierValuePCC.java

@@ -51,7 +51,7 @@ public class OneOfQualifierValuePCC extends PropertyConstraintChecker {
 			}
 			
 			// violation_triple(S, I, propertyConstant, V) :- tripleEDB(S, I, propertyConstant, V), qualifierEDB(S, qualifierPropertyConstant, O), unequal({A}, O)
-			Rule violation = Expressions.makeRule(violation_triple_SIpV, toArray(violation_conjunction));
+			Rule violation = Expressions.makeRule(violation_triple_SIpV, Utility.toArray(violation_conjunction));
 			rules.add(violation);
 		}
 

### 2018-08-09 src/main/java/impl/PCC/PropertyConstraintChecker.java

@@ -65,9 +65,5 @@ public abstract class PropertyConstraintChecker {
 		referenceEDB_SpV = Expressions.makeAtom(referenceEDB, s, propertyConstant, v);
 	}
 	
-	protected static Atom[] toArray(List<Atom> list) {
-		return list.toArray(new Atom[list.size()]);
-	}
-	
 	public abstract List<Rule> rules();
 }

### 2018-08-09 src/main/java/impl/PCC/SingleValuePCC.java

@@ -33,6 +33,7 @@ import com.google.common.collect.Sets;
 
 import utility.InequalityHelper;
 import utility.StatementNonExistenceHelper;
+import utility.Utility;
 
 public class SingleValuePCC extends PropertyConstraintChecker {
 	
@@ -156,7 +157,7 @@ public class SingleValuePCC extends PropertyConstraintChecker {
 				//	qualifierEDB(O, {X}, [X]),
 				//	does_not_have(S, {Y}),
 				//	does_not_have(O, {Y})
-				Rule violation = Expressions.makeRule(violation_triple_SIpV, toArray(conjunction));
+				Rule violation = Expressions.makeRule(violation_triple_SIpV, Utility.toArray(conjunction));
 				if (hasNot.size() > 0)
 					continue;
 				rules.add(violation);

### 2018-08-09 src/main/java/main/Main.java

@@ -39,6 +39,7 @@ import impl.CC.OneOfCC;
 import impl.CC.OneOfQualifierValueCC;
 import impl.CC.ScopeCC;
 import impl.CC.SingleValueCC;
+import utility.InequalityHelper;
 
 /**
  * @author adrian
@@ -70,6 +71,7 @@ public class Main {
 		options.addOption("c", "constraints", true, "The constraint to check.");
 		options.addOption("n", "noviolations", false, "Do not compute violations.");
 		options.addOption("s", "stringResults", false, "Output violations as string.");
+		options.addOption("i", "inequalityMode", true, "Choose the inequality mode. Default is encoded.");
 		
 		CommandLineParser parser = new DefaultParser();
 	    CommandLine cmd;
@@ -92,6 +94,26 @@ public class Main {
 	    	System.out.println("Please specify the constraints.");
 		    formatter.printHelp("help", options);
 		    return;
+	    }
+	    
+	    if (!cmd.hasOption("inequalityMode"))
+	    	InequalityHelper.mode = InequalityHelper.Mode.ENCODED;
+	    else {
+	    	String inequalityMode = cmd.getOptionValue("inequalityMode").toLowerCase();
+	    	switch (inequalityMode) {
+			case "naive":
+				InequalityHelper.mode = InequalityHelper.Mode.NAIVE;
+				break;
+			case "encoded":
+				InequalityHelper.mode = InequalityHelper.Mode.ENCODED;
+				break;
+			case "demanded":
+				InequalityHelper.mode = InequalityHelper.Mode.DEMANDED;
+				break;
+			default:
+				System.out.println("Inequality mode " + inequalityMode + " is unknown.");
+				return;
+			}
 	    }
 	    	
 	    

### 2018-08-09 src/main/java/utility/InequalityHelper.java

@@ -14,7 +14,15 @@ import org.semanticweb.vlog4j.core.model.implementation.Expressions;
 import org.semanticweb.vlog4j.core.reasoner.Reasoner;
 import org.semanticweb.vlog4j.core.reasoner.exceptions.ReasonerStateException;
 
+import static utility.SC.require_inequality;
+
 public class InequalityHelper {
+	
+	public enum Mode {
+		NAIVE,
+		ENCODED,
+		DEMANDED
+	}
 	
 	final static String NONE = "none";
 	
@@ -41,6 +49,9 @@ public class InequalityHelper {
 	// unequal(CON1, CON2)
 	final static Atom unequal_CC = Expressions.makeAtom(unequal, con1, con2);
 	
+	// require_inequality(CON1, CON2)
+	final static Atom require_inequality_CC = Expressions.makeAtom(require_inequality, con1, con2);
+	
 	// unequal(X, Y)
 	final static Atom unequal_XY = Expressions.makeAtom(unequal, x, y);
 	
@@ -52,21 +63,54 @@ public class InequalityHelper {
 	
 	static Reasoner reasoner;
 	
+	public static Mode mode = Mode.ENCODED;
+	
 	public static void setOrReset(Reasoner reasoner_) {
 		reasoner = reasoner_;
 		ITH_LETTER.clear();
 		ith_letter.clear();
 	}
 	
-	public static void addUnequalConstantsToReasoner(String...unequalConstants) throws ReasonerStateException {
+	public static void establishInequality(String...unequalConstants) throws ReasonerStateException {
 		Set<String> unequalConstantsSet = new HashSet<String>();
 		for (int i = 0; i < unequalConstants.length; i++) {
 			unequalConstantsSet.add(unequalConstants[i]);
 		}
-		addUnequalConstantsToReasoner(unequalConstantsSet);
+		establishInequality(unequalConstantsSet);
 	}
 	
-	public static void addUnequalConstantsToReasoner(Set<String> unequalConstants) throws ReasonerStateException {
+	public static void establishInequality(Set<String> unequalConstants) throws ReasonerStateException {
+		reasoner.addRules(unequal_IDB_EDB, inverse);
+		
+		switch (mode) {
+		case NAIVE:
+			naive(unequalConstants);
+			break;
+		case ENCODED:
+			encoded(unequalConstants, false);
+			break;
+		case DEMANDED:
[…]

### 2018-08-09 src/main/java/utility/SC.java

@@ -15,6 +15,8 @@ public class SC {
 	public final static String VIOLATION_QUALIFIER = "violation_qualifier";
 	public final static String VIOLATION_REFERENCE = "violation_reference";
 	
+	public final static String REQUIRE_INEQUALITY = "require_inequality";
+	
 	public final static String S = "s";
 	public final static String O = "o";
 	public final static String I = "i";
@@ -47,9 +49,15 @@ public class SC {
 	
 	public final static String DOES_NOT_HAVE = "does_not_have";
 	
+	public final static Predicate tripleEDB = Expressions.makePredicate(TRIPLE, 4);
+	public final static Predicate qualifierEDB = Expressions.makePredicate(QUALIFIER, 3);
+	public final static Predicate referenceEDB = Expressions.makePredicate(REFERENCE, 3);
+	
 	public final static Predicate violation_triple = Expressions.makePredicate(VIOLATION_TRIPLE, 4);
 	public final static Predicate violation_qualifier = Expressions.makePredicate(VIOLATION_QUALIFIER, 3);
 	public final static Predicate violation_reference = Expressions.makePredicate(VIOLATION_REFERENCE, 3);
+	
+	public final static Predicate require_inequality = Expressions.makePredicate(REQUIRE_INEQUALITY, 2);
 	
 	public final static Variable s = Expressions.makeVariable(S);
 	public final static Variable o = Expressions.makeVariable(O);
@@ -62,10 +70,6 @@ public class SC {
 	public final static Variable u = Expressions.makeVariable(U);
 	public final static Variable w = Expressions.makeVariable(W);
 	public final static Variable r = Expressions.makeVariable(R);
-	
-	public final static Predicate tripleEDB = Expressions.makePredicate(TRIPLE, 4);
-	public final static Predicate qualifierEDB = Expressions.makePredicate(QUALIFIER, 3);
-	public final static Predicate referenceEDB = Expressions.makePredicate(REFERENCE, 3);
 	
 	public final static Atom violation_triple_query = Expressions.makeAtom(violation_triple, s, i, p, v);
 	public final static Atom violation_qualifier_query = Expressions.makeAtom(violation_qualifier, s, p, v);

### 2018-08-09 src/main/java/utility/Utility.java

@@ -1,5 +1,8 @@
 package utility;
 
+import java.util.List;
+
+import org.semanticweb.vlog4j.core.model.api.Atom;
 import org.semanticweb.vlog4j.core.model.api.Constant;
 import org.semanticweb.vlog4j.core.model.implementation.Expressions;
 
@@ -16,5 +19,9 @@ public class Utility {
 			return string;
 		else
 			return BASE_URI + string;
+	}
+	
+	public static Atom[] toArray(List<Atom> list) {
+		return list.toArray(new Atom[list.size()]);
 	}
 }

### 2018-08-09 thesis/paper/paper.tex
Modded demand-driven

@@
 To save space and avoid redundancy the following rules will not explicitly state the inequality requirement predicate.
-Instead, assume that for every premise with the form \(B_1 \wedge \ldots\wedge B_m \wedge I\) where \(B_1 ..
-B_m\) are arbitrary atoms, \(I\) is an atom of the form \(\text{unequal}(t_1, t_2)\) with \(t_1,t_2\) terms and all of \(t_1,t_2\) that are variables appear in at least one of \(B_1 \ldots B_m\), the rule
-\begin{equation*}
-\text{req\_inequality}(t_1, t_2)
+Instead, assume that for every premise with the form \(B_1 \wedge \ldots\wedge B_m \wedge I_1 \wedge I_n\) where \(B_1 ..
+B_m\) are arbitrary atoms, \(I_1 ..
+I_n\) are atoms of the form \(\text{unequal}(t_{n_1}, t_{n_2})\) with \(t_{n_1}, t_{n_2}\) terms and all of \(t_{n_1}, t_{n_2}\) that are variables appear in at least one of \(B_1 \ldots B_m\), the rules
+\begin{equation*}
+\text{req\_inequality}(t_{n_1}, t_{n_2})
  \leftarrow B_1 \wedge \ldots \wedge B_m
-\end{equation*} is introduced.\\
+\end{equation*} for all \(n\) are introduced.\\
 

### 2018-08-19 src/main/java/impl/PCC/ItemRequiresStatementPCC.java
Adapted require predicates to use two terms
Needed to differentiate between requirements from different sources.

@@ -48,7 +48,7 @@ public class ItemRequiresStatementPCC extends PropertyConstraintChecker {
 			// unequal(requiredPropertyConstant, P)
 			Atom unequal_rP = Expressions.makeAtom(InequalityHelper.unequal, requiredPropertyConstant, p);
 		
-			rules.addAll(StatementNonExistenceHelper.initRequireTriple(requiredPropertyConstant, tripleEDB_SIPV, unequal_rP));
+			rules.addAll(StatementNonExistenceHelper.initRequireTriple(propertyConstant, requiredPropertyConstant, tripleEDB_SIPV, unequal_rP));
 			
 			Set<String> allowedValues = entry.getValue();				
 			if (allowedValues.size() != 0) {
@@ -63,7 +63,7 @@ public class ItemRequiresStatementPCC extends PropertyConstraintChecker {
 					conjunction.add(Expressions.makeAtom(InequalityHelper.unequal, allowedValueConstant, v));
 				}
 				
-				rules.addAll(StatementNonExistenceHelper.initRequireTriple(requiredPropertyConstant, conjunction));
+				rules.addAll(StatementNonExistenceHelper.initRequireTriple(propertyConstant, requiredPropertyConstant, conjunction));
 			}
 		}
 		
@@ -76,14 +76,14 @@ public class ItemRequiresStatementPCC extends PropertyConstraintChecker {
 			// last(O, I)
 			Atom last_OI = Expressions.makeAtom(last, o, i);
 			
-			// require(O, requiredPropertyConstant)
-			Atom require_Or = Expressions.makeAtom(require, o, requiredPropertyConstant);
+			// require(O, propertyConstant, requiredPropertyConstant)
+			Atom require_Opr = Expressions.makeAtom(require, o, propertyConstant, requiredPropertyConstant);
 			
 			// violation_triple(S, I, propertyConstant, V) :-
 			//	tripleEDB(S, I, propertyConstant, V),
 			//	tripleEDB(O, I, P, X), last(O, I),
 			//	require(O, requiredPropertyConstant)
-			Rule violation = Expressions.makeRule(violation_triple_SIpV, tripleEDB_SIpV, tripleEDB_OIPX, last_OI, require_Or);
+			Rule violation = Expressions.makeRule(violation_triple_SIpV, tripleEDB_SIpV, tripleEDB_OIPX, last_OI, require_Opr);
 			rules.add(violation);
 		}
 		

### 2018-08-19 src/main/java/impl/PCC/MultiValuePCC.java

@@ -45,10 +45,10 @@ public class MultiValuePCC extends PropertyConstraintChecker {
 		// unequal(propertyConstant, P)
 		Atom unequal_pP = Expressions.makeAtom(InequalityHelper.unequal, propertyConstant, p);
 		
-		rules.addAll(StatementNonExistenceHelper.initRequireTriple(propertyConstant, tripleEDB_SIPV, unequal_pP));
+		rules.addAll(StatementNonExistenceHelper.initRequireTriple(propertyConstant, propertyConstant, tripleEDB_SIPV, unequal_pP));
 		
-		// require_second(S, propertyConstant)
-		Atom require_second_Sp = Expressions.makeAtom(require_second, s, propertyConstant);
+		// require_second(S, propertyConstant, propertyConstant)
+		Atom require_second_Spp = Expressions.makeAtom(require_second, s, propertyConstant, propertyConstant);
 		
 		// next(Q, R)
 		Atom next_QR = Expressions.makeAtom(next, q, r);
@@ -56,30 +56,30 @@ public class MultiValuePCC extends PropertyConstraintChecker {
 		// next(R, S)
 		Atom next_RS = Expressions.makeAtom(next, r, s);
 		
-		// require(Q, propertyConstant)
-		Atom require_Qp = Expressions.makeAtom(require, q, propertyConstant);
+		// require(Q, propertyConstant, propertyConstant)
+		Atom require_Qpp = Expressions.makeAtom(require, q, propertyConstant, propertyConstant);
 		
 		// tripleEDB(R, I, propertyConstant, C)
 		Atom tripleEDB_RIpC = Expressions.makeAtom(tripleEDB, r, i, propertyConstant, c);
 		
-		// require_second(S, propertyConstant) :-
+		// require_second(S, propertyConstant, propertyConstant) :-
 		//	next(Q, R), next(R, S),
-		//	require(Q, propertyConstant),
+		//	require(Q, propertyConstant, propertyConstant),
 		//	tripleEDB(R, I, propertyConstant, C),
 		//	tripleEDB(S, I, P, V),
 		//	unequal(propertyConstant, P)
-		Rule require1 = Expressions.makeRule(require_second_Sp, next_QR, next_RS, require_Qp, tripleEDB_RIpC, tripleEDB_SIPV, unequal_pP);
+		Rule require1 = Expressions.makeRule(require_second_Spp, next_QR, next_RS, require_Qpp, tripleEDB_RIpC, tripleEDB_SIPV, unequal_pP);
[…]

### 2018-08-19 src/main/java/impl/PCC/ValueRequiresStatementPCC.java

@@ -48,7 +48,7 @@ public class ValueRequiresStatementPCC extends PropertyConstraintChecker {
 			// unequal(requiredPropertyConstant, P)
 			Atom unequal_rP = Expressions.makeAtom(InequalityHelper.unequal, requiredPropertyConstant, p);
 		
-			rules.addAll(StatementNonExistenceHelper.initRequireTriple(requiredPropertyConstant, tripleEDB_SIPV, unequal_rP));
+			rules.addAll(StatementNonExistenceHelper.initRequireTriple(propertyConstant, requiredPropertyConstant, tripleEDB_SIPV, unequal_rP));
 			
 			Set<String> allowedValues = entry.getValue();				
 			if (allowedValues.size() != 0) {
@@ -63,7 +63,7 @@ public class ValueRequiresStatementPCC extends PropertyConstraintChecker {
 					conjunction.add(Expressions.makeAtom(InequalityHelper.unequal, allowedValueConstant, v));
 				}
 				
-				rules.addAll(StatementNonExistenceHelper.initRequireTriple(requiredPropertyConstant, conjunction));
+				rules.addAll(StatementNonExistenceHelper.initRequireTriple(propertyConstant, requiredPropertyConstant, conjunction));
 			}
 		}
 		
@@ -76,14 +76,14 @@ public class ValueRequiresStatementPCC extends PropertyConstraintChecker {
 			// last(O, V)
 			Atom last_OV = Expressions.makeAtom(last, o, v);
 			
-			// require(O, requiredPropertyConstant)
-			Atom require_Or = Expressions.makeAtom(require, o, requiredPropertyConstant);
+			// require(O, propertyConstant, requiredPropertyConstant)
+			Atom require_Orr = Expressions.makeAtom(require, o, propertyConstant, requiredPropertyConstant);
 			
 			// violation_triple(S, I, propertyConstant, V) :-
 			//	tripleEDB(S, I, propertyConstant, V),
 			//	tripleEDB(O, V, P, X), last(O, V),
-			//	require(O, requiredPropertyConstant)
-			Rule violation = Expressions.makeRule(violation_triple_SIpV, tripleEDB_SIpV, tripleEDB_OVPX, last_OV, require_Or);
+			//	require(O, propertyConstant, requiredPropertyConstant)
+			Rule violation = Expressions.makeRule(violation_triple_SIpV, tripleEDB_SIpV, tripleEDB_OVPX, last_OV, require_Orr);
 			rules.add(violation);
 		}
 		

### 2018-08-19 src/main/java/utility/SC.java

@@ -75,8 +75,8 @@ public class SC {
 	public final static Atom violation_qualifier_query = Expressions.makeAtom(violation_qualifier, s, p, v);
 	public final static Atom violation_reference_query = Expressions.makeAtom(violation_reference, s, p, v);
 	
-	public static final Predicate require = Expressions.makePredicate(REQUIRE, 2);
-	public static final Predicate require_second = Expressions.makePredicate(REQUIRE_SECOND, 2);
+	public static final Predicate require = Expressions.makePredicate(REQUIRE, 3);
+	public static final Predicate require_second = Expressions.makePredicate(REQUIRE_SECOND, 3);
 	
 	public static final Predicate first = Expressions.makePredicate(FIRST, 2);
 	public static final Predicate next = Expressions.makePredicate(NEXT, 2);

### 2018-08-19 src/main/java/utility/StatementNonExistenceHelper.java

@@ -26,13 +26,13 @@ import org.semanticweb.vlog4j.core.model.implementation.Expressions;
 
 public class StatementNonExistenceHelper {	
 	
-	public static List<Rule> initRequireTriple(Term requiredTerm, Atom...conjunctionAtoms) {
-		return initRequireTriple(requiredTerm, Arrays.asList(conjunctionAtoms));
+	public static List<Rule> initRequireTriple(Term requiringTerm, Term requiredTerm, Atom...conjunctionAtoms) {
+		return initRequireTriple(requiringTerm, requiredTerm, Arrays.asList(conjunctionAtoms));
 	}
 	
-	public static List<Rule> initRequireTriple(Term requiredTerm, List<Atom> conjunctionAtoms) {
-		// require(S, requiredTerm)
-		Atom require_Sr = Expressions.makeAtom(require, s, requiredTerm);
+	public static List<Rule> initRequireTriple(Term requiringTerm, Term requiredTerm, List<Atom> conjunctionAtoms) {
+		// require(S, requiringTerm, requiredTerm)
+		Atom require_Srr = Expressions.makeAtom(require, s, requiringTerm, requiredTerm);
 		
 		// first(S, I)
 		Atom first_SI = Expressions.makeAtom(first, s, i);
@@ -42,22 +42,22 @@ public class StatementNonExistenceHelper {
 		firstConjunctionAtoms.add(first_SI);
 		firstConjunctionAtoms.addAll(conjunctionAtoms);
 		
-		// require(S, requiredTerm) :- first(S, I), conjunctionAtoms
-		Rule firstRequire = Expressions.makeRule(require_Sr, toArray(firstConjunctionAtoms));
+		// require(S, requiringTerm, requiredTerm) :- first(S, I), conjunctionAtoms
+		Rule firstRequire = Expressions.makeRule(require_Srr, toArray(firstConjunctionAtoms));
 		
 		// next(O, S)
 		Atom next_OS = Expressions.makeAtom(next, o, s);
 		
-		// require(O, requiredTerm)
-		Atom require_Or = Expressions.makeAtom(require, o, requiredTerm);
+		// require(O, requiringTerm, requiredTerm)
+		Atom require_Orr = Expressions.makeAtom(require, o, requiringTerm, requiredTerm);
 		
 		List<Atom> nextConjunctionAtoms = new ArrayList<Atom>();
 		nextConjunctionAtoms.add(next_OS);
[…]

### 2018-08-19 thesis/paper/paper.tex
Adapted rules in paper for two required terms

@@
 Using this pre-processed data the approach is to apply a require-predicate to each statement or qualifier.
-This is require\([2]\) for statements, connecting the statement ID with a required term \(t\) and require\([4]\) taking the statement ID, property, and value of a qualifier and linking this with the term \(t\).\\
+This is require\([3]\) for statements, connecting the statement ID with the constrained property \(p\) and the required term \(t\) and require\([5]\) taking the statement ID, property, and value of a qualifier and linking this with \(p\) and \(t\).\\
 
@@
 \begin{equation*}
-\text{require}(S, t) \leftarrow \text{first}(S, I) \wedge c
+\text{require}(S, p, t) \leftarrow \text{first}(S, I) \wedge c
 \end{equation*}
@@
 \begin{equation*}
-\text{require}(S, t) \leftarrow \text{next}(O, S) \wedge \text{require}(O, t) \wedge c
+\text{require}(S, p, t) \leftarrow \text{next}(O, S) \wedge \text{require}(O, p, t) \wedge c
 \end{equation*}
@@
 \begin{equation*}
-\text{require}(S, P, V, t) \leftarrow \text{first}(S, P, V) \wedge c
-\end{equation*}
-\begin{equation*}
-\text{require}(S, P, V, t) \leftarrow \text{next}(O, X, C, S, P, V) \wedge \text{require}(O, X, C, t) \wedge c
+\text{require}(S, P, V, p, t) \leftarrow \text{first}(S, P, V) \wedge c
+\end{equation*}
+\begin{equation*}
+\text{require}(S, P, V, p, t) \leftarrow \text{next}(O, X, C, S, P, V) \wedge \text{require}(O, X, C, p, t) \wedge c
 \end{equation*}
@@
 The required term \(t\) is the  required property.
-Without separators the conjunction is \(c_1 = \text{tripleEDB}(S, I, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)\).\\
-
-If a set of allowed values \(A\) has been specified, a second set of require-rules have to be introduced with \(c_2 = \text{tripleEDB}(S, I, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)\).
-Note that the first set of rules is still required, since the preceding statements do not need to have the required property.\\
-
-In both cases violations can be found with
+Without separators the conjunction is
+\begin{equation*}
+\begin{split}
+c_1 = &\text{tripleEDB}(O, I, \text{propertyConstant}, X) \wedge \\
+&\text{tripleEDB}(S, I, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
+\end{split}
+\end{equation*} \(\)
+If a set of allowed values \(A\) has been specified, a second set of require-rules have to be introduced with
+\begin{equation*}
+\begin{split}
+c_2 = &\text{tripleEDB}(O, I, \text{propertyConstant}, X) \wedge \\
+&\text{tripleEDB}(S, I, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
+\end{split}
+\end{equation*}
+Note that the first set of rules is still required, since the preceding statements do not need to have the required property. In both cases violations can be found with
 \begin{equation*}
@@
 &\text{tripleEDB}(O, I, P, X) \wedge \text{last}(O, I) \wedge \\
-&\text{require}(O, \text{requiredPropertyConstant})
+&\text{require}(O, \text{propertyConstant}, \text{requiredPropertyConstant})
 \end{split}
@@
 &\text{tripleEDB}(O, V, P, X) \wedge \text{last}(O, V) \wedge \\
-&\text{require}(O, \text{requiredPropertyConstant})
+&\text{require}(O, text{propertyConstant}, \text{requiredPropertyConstant})
 \end{split}
@@
 The inverse constraint is similar to the value requires statement constraint in requiring a statement on a constrained value with a specific property.
-However, instead of putting no constraint or a list of fixed values on the required statements value it set a relative requirement: the constrained item.
+However, instead of putting no constraint or a list of fixed values on the required statements value it set a relative requirement: the constrained item, meaning \(t = R\).
 The pre-processing and inequalities thus remain the same, but the conjunctions change to
 \begin{equation*}
-c_1 = \text{tripleEDB}(S, I, P, V) \wedge \text{unequal}(\text{inversePropertyConstant}, P)
-\end{equation*}
-\begin{equation*}
-\begin{split}
-c_2 = &\text{tripleEDB}(O, R, \text{propertyConstant}, X) \wedge \\
+\begin{split}
+c_1 = &\text{tripleEDB}(O, R, \text{propertyConstant}, I) \wedge \\
+&\text{tripleEDB}(S, I, P, V) \wedge \\
+&\text{unequal}(\text{inversePropertyConstant}, P)
+\end{split}
+\end{equation*}
+\begin{equation*}
+\begin{split}
+c_2 = &\text{tripleEDB}(O, R, \text{propertyConstant}, I) \wedge \\
 &\text{tripleEDB}(S, I, \text{inversePropertyConstant}, V) \wedge \\

### 2018-08-19 src/main/java/impl/CC/SymmetricCC.java
Added symmetric

@@ -1,0 +1,95 @@
+package impl.CC;
+
+import static utility.SC.first;
+import static utility.SC.last;
+import static utility.SC.next;
+import static utility.SC.o;
+import static utility.SC.p;
+import static utility.SC.r;
+import static utility.SC.require;
+import static utility.SC.violation_triple_query;
+
+import java.io.IOException;
+import java.util.ArrayList;
+import java.util.HashSet;
+import java.util.List;
+import java.util.Set;
+
+import org.apache.jena.query.QuerySolution;
+import org.apache.log4j.Logger;
+import org.semanticweb.vlog4j.core.model.api.Atom;
+import org.semanticweb.vlog4j.core.model.implementation.Expressions;
+import org.semanticweb.vlog4j.core.reasoner.DataSource;
+import org.semanticweb.vlog4j.core.reasoner.exceptions.ReasonerStateException;
+import org.semanticweb.vlog4j.core.reasoner.implementation.CsvFileDataSource;
+
+import impl.PCC.InversePCC;
+import impl.PCC.PropertyConstraintChecker;
+import main.Main;
+import utility.InequalityHelper;
+import utility.Utility;
+
+public class SymmetricCC extends ConstraintChecker {
+	
+	final static Logger logger = Logger.getLogger(SymmetricCC.class);
+	
+	Set<String> configuration;
+
+	public SymmetricCC() throws IOException {
+		super("Q21510862");
+	}
+
+	@Override
+	void initDataField() {
+		configuration = new HashSet<String>();
+	}
+	
+	@Override
+	protected Set<Atom> queries() {
+		return asSet(violation_triple_query);
+	}
+	
+	@Override
+	protected Set<String> qualifiers() {
+		return asSet();
+	}
+
+	@Override
+	protected Set<String> concatQualifiers() {
+		return asSet();
+	}
+
+	@Override
+	protected void process(QuerySolution solution) {
+		String property = Utility.addBaseURI(solution.get("item").asResource().getLocalName());
+		
+		configuration.add(property);
+	}
+
+	@Override
+	void prepareFacts() throws ReasonerStateException, IOException {
+		final DataSource firstEDBPath = new CsvFileDataSource(Main.tripleSet.getFirstFile());
+		reasoner.addFactsFromDataSource(first, firstEDBPath);
+
+		final DataSource nextEDBPath = new CsvFileDataSource(Main.tripleSet.getNextFile());
[…]

### 2018-08-19 src/main/java/main/Main.java

@@ -40,6 +40,7 @@ import impl.CC.OneOfCC;
 import impl.CC.OneOfQualifierValueCC;
 import impl.CC.ScopeCC;
 import impl.CC.SingleValueCC;
+import impl.CC.SymmetricCC;
 import impl.CC.ValueRequiresStatementCC;
 import impl.TS.TripleSet;
 import utility.InequalityHelper;
@@ -179,6 +180,9 @@ public class Main {
 					break;
 				case "inverse":
 					checkers.add(new InverseCC());
+					break;
+				case "symmetric":
+					checkers.add(new SymmetricCC());
 					break;
 				case "multivalue":
 					checkers.add(new MultiValueCC());

### 2018-08-19 thesis/paper/paper.tex
Modified variable names

@@
 \begin{split}
-c_1 = &\text{tripleEDB}(O, I, \text{propertyConstant}, X) \wedge \\
+c_1 = &\text{tripleEDB}(Q, I, \text{propertyConstant}, X) \wedge \\
 &\text{tripleEDB}(S, I, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
@@
 \begin{split}
-c_2 = &\text{tripleEDB}(O, I, \text{propertyConstant}, X) \wedge \\
+c_2 = &\text{tripleEDB}(Q, I, \text{propertyConstant}, X) \wedge \\
 &\text{tripleEDB}(S, I, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
@@
 \begin{split}
-c_1 = &\text{tripleEDB}(O, R, \text{propertyConstant}, I) \wedge \\
+c_1 = &\text{tripleEDB}(Q, R, \text{propertyConstant}, I) \wedge \\
 &\text{tripleEDB}(S, I, P, V) \wedge \\
@@
 \begin{split}
-c_2 = &\text{tripleEDB}(O, R, \text{propertyConstant}, I) \wedge \\
+c_2 = &\text{tripleEDB}(Q, R, \text{propertyConstant}, I) \wedge \\
 &\text{tripleEDB}(S, I, \text{inversePropertyConstant}, V) \wedge \\

### 2018-08-24 src/main/java/impl/CC/ConstraintChecker.java
Inequality with longer letter predicates

@@ -53,7 +53,7 @@ public abstract class ConstraintChecker {
 	
 	String constraint;
 	
-	protected final Reasoner reasoner = Reasoner.getInstance();
+	protected final Reasoner reasoner;
 	
 	protected final String internalError;
 	
@@ -66,6 +66,7 @@ public abstract class ConstraintChecker {
 	public ConstraintChecker(String constraint_) throws IOException {
 		constraint = constraint_;
 		internalError = "INTERNAL_ERROR for constraint " + constraint + ".";
+		reasoner = Reasoner.getInstance();
 		reasoner.setAlgorithm(Algorithm.RESTRICTED_CHASE);
 		initDataField();
 		init();
@@ -129,6 +130,7 @@ public abstract class ConstraintChecker {
 	}
 	
 	public void violations() throws ReasonerStateException, IOException {
+		InequalityHelper.setOrReset(reasoner);
 		loadTripleSet();
 		prepareFacts();
 		List<Rule> rulesToAdd = new ArrayList<Rule>();
@@ -197,6 +199,13 @@ public abstract class ConstraintChecker {
 	
 	protected void prepareAndExecuteQueries(List<Rule> rules, Set<Atom> queries) throws IOException, PrepareQueriesException {
 		try {
+			InequalityHelper.load();
+		} catch (ReasonerStateException e) {
+			logger.error("Trying to load in the wrong state for constraint " + constraint + ".", e);
+			throw new PrepareQueriesException(internalError);
+		}
+		
+		try {
 			reasoner.addRules(rules);
 		} catch (ReasonerStateException e) {
 			logger.error("Trying to add rules in the wrong state for constraint " + constraint + ".", e);
@@ -238,6 +247,8 @@ public abstract class ConstraintChecker {
 				throw new PrepareQueriesException(internalError);
 			}
 		}
+
+		InequalityHelper.delete();
 	}
 	
 	public int getResultSize() {

### 2018-08-24 src/main/java/impl/CC/SymmetricCC.java

@@ -68,8 +68,6 @@ public class SymmetricCC extends ConstraintChecker {
 		Main.tripleSet.loadLastFile(reasoner);
 
 		// Establishing inequality
-		InequalityHelper.setOrReset(reasoner);
-
 		InequalityHelper.establishInequality(Main.tripleSet.getTripleFile(), 2, configuration);
 		InequalityHelper.establishInequality(Main.tripleSet.getTripleFile(), 1, Main.tripleSet.getTripleFile(), 3);
 	}

### 2018-08-24 src/main/java/impl/TS/TripleSetFile.java

@@ -7,6 +7,7 @@ import java.io.FileOutputStream;
 import java.io.IOException;
 import java.io.OutputStreamWriter;
 import java.util.Arrays;
+import java.util.List;
 import java.util.zip.GZIPOutputStream;
 
 import org.apache.commons.csv.CSVFormat;
@@ -55,13 +56,17 @@ public class TripleSetFile {
 	}
 	
 	public void write(String...strings) {
+		write(Arrays.asList(strings));
+	}
+	
+	public void write(List<String> strings) {
 		if (writer == null) {
 			logger.warn("Trying to write to file " + fileName + " but extraction has not been activated.");
 			return;
 		}
 			
 		try {
-			writer.printRecord(Arrays.asList(strings));
+			writer.printRecord(strings);
 			writer.flush();
 		} catch (IOException e) {
 			logger.error("Could not write line to file " + tripleSetFileGz.getAbsolutePath(), e);

### 2018-08-24 src/main/java/utility/InequalityHelper.java

@@ -2,7 +2,6 @@ package utility;
 
 import java.io.File;
 import java.io.FileInputStream;
-import java.io.FileNotFoundException;
 import java.io.IOException;
 import java.io.InputStreamReader;
 import java.util.ArrayList;
@@ -18,10 +17,13 @@ import org.semanticweb.vlog4j.core.model.api.Atom;
 import org.semanticweb.vlog4j.core.model.api.Constant;
 import org.semanticweb.vlog4j.core.model.api.Predicate;
 import org.semanticweb.vlog4j.core.model.api.Rule;
+import org.semanticweb.vlog4j.core.model.api.Term;
 import org.semanticweb.vlog4j.core.model.api.Variable;
 import org.semanticweb.vlog4j.core.model.implementation.Expressions;
 import org.semanticweb.vlog4j.core.reasoner.Reasoner;
 import org.semanticweb.vlog4j.core.reasoner.exceptions.ReasonerStateException;
+
+import impl.TS.TripleSetFile;
 
 import static utility.SC.require_inequality;
 
@@ -33,6 +35,8 @@ public class InequalityHelper {
 		DEMANDED
 	}
 	
+	final static int chunkSize = 7;
+	
 	final static String NONE = "none";
 	
 	final static String X = "x";
@@ -43,23 +47,17 @@ public class InequalityHelper {
 	
 	final static String UNEQUAL_EDB = "unequal_EDB";
 	final static String UNEQUAL = "unequal";
-	final static List<String> ITH_LETTER = new ArrayList<String>();
 	
 	final static Variable x = Expressions.makeVariable(X);
 	final static Variable y = Expressions.makeVariable(Y);
 	
-	final static Variable con1 = Expressions.makeVariable(CON1);
-	final static Variable con2 = Expressions.makeVariable(CON2);
-	
 	final static Predicate unequal_EDB = Expressions.makePredicate(UNEQUAL_EDB, 2);
 	public final static Predicate unequal = Expressions.makePredicate(UNEQUAL, 2);
-	final static List<Predicate> ith_letter = new ArrayList<Predicate>();
 	
-	// unequal(CON1, CON2)
-	final static Atom unequal_CC = Expressions.makeAtom(unequal, con1, con2);
+	final static List<TripleSetFile> files = new ArrayList<>();
 	
-	// require_inequality(CON1, CON2)
[…]

### 2018-08-24 thesis/paper/paper.tex
Inequalities with longer letter predicates paper

@@
 Then \(l_{\max} = \max\limits_{\forall e \in E}(len(e))\) is the length of the longest entity string.
-Based on this a group of predicates letter\_iEDB[2] is introduced where \(i \in \{0 \ldots l_{\max}-1\}\) and letter\_iEBD denotes the ith letter of the entity string.\\
-
-For each entity string \(e = c_0 \ldots c_{l - 1}\) each predicate letter\_iEDB is created as \(\text{letter\_iEDB}(e, f(e, i))\), where
+Based on this a group of predicates letter\_i\_EDB[8] is introduced where \(0 \equiv i \mod 7\), \(0 \leq i < l_{max} \).
+
+For each entity string \(e = c_0 \ldots c_{l - 1}\) each predicate letter\_i\_EDB is created as \(\text{letter\_i\_EDB}(e, f(e, i), f(e, i+1), \ldots, f(e, i+6)\), where
 \begin{equation*}
@@
 
-Now that inequality between all constitution characters and the link between entities and their characters are introduced, the inequality between two entities can be derived by creating the following rules for all predicates letter\_i:
-\begin{equation*}
-\text{unequal}(X, Y) \leftarrow \text{letter\_i}(X, A) \wedge \text{letter\_i}(Y, B) \wedge \text{unequal}(A, B)
+Now that inequality between all constitution characters and the link between entities and their characters are introduced, the inequality between two entities can be derived by creating the following rules for all predicates letter\_i (with \text{letters} short for \(\text{letter\_i}(X, A_i, \ldots, A_{i+14}) \wedge\text{letter\_i}(Y, B_i, \ldots, B_{i+6})\)):
+\begin{equation*}
+\text{unequal}(X, Y) \leftarrow \text{letters} \wedge \text{unequal}(A_i, B_i)
+\end{equation*}
+\begin{equation*}
+\text{unequal}(X, Y) \leftarrow \text{letters} \wedge \text{unequal}(A_{i+6}, B_{i+6})
 \end{equation*}

### 2018-08-25 src/main/java/impl/CC/ConstraintChecker.java
Removed unused imports

@@ -26,7 +26,6 @@ import org.semanticweb.vlog4j.core.model.api.Term;
 import org.semanticweb.vlog4j.core.model.api.Variable;
 import org.semanticweb.vlog4j.core.model.implementation.Expressions;
 import org.semanticweb.vlog4j.core.reasoner.Algorithm;
-import org.semanticweb.vlog4j.core.reasoner.DataSource;
 import org.semanticweb.vlog4j.core.reasoner.Reasoner;
 import org.semanticweb.vlog4j.core.reasoner.exceptions.EdbIdbSeparationException;
 import org.semanticweb.vlog4j.core.reasoner.exceptions.IncompatiblePredicateArityException;
@@ -35,10 +34,8 @@ import org.semanticweb.vlog4j.core.reasoner.implementation.QueryResultIterator;
 
 import impl.PCC.PropertyConstraintChecker;
 import main.Main;
-import utility.CsvGzFileDataSource;
 import utility.InequalityHelper;
 import utility.PrepareQueriesException;
-import utility.SC;
 import utility.Utility;
 
 import static utility.SC.require_inequality;
@@ -119,7 +116,7 @@ public abstract class ConstraintChecker {
 		while (results.hasNext()) {
 			QuerySolution solution = results.next();
 			//String property = solution.get("item").asResource().getLocalName();
-			//if (!(property.equals("P209"))) 
+			//if (!(property.equals("P103"))) 
 			//	continue;
 
 			process(solution);

### 2018-08-25 src/main/java/impl/PCC/ItemRequiresStatementPCC.java
Added "demand" rule to item requires statement

@@ -4,6 +4,7 @@ import static utility.SC.i;
 import static utility.SC.last;
 import static utility.SC.o;
 import static utility.SC.p;
+import static utility.SC.q;
 import static utility.SC.require;
 import static utility.SC.s;
 import static utility.SC.tripleEDB;
@@ -41,6 +42,9 @@ public class ItemRequiresStatementPCC extends PropertyConstraintChecker {
 		
 		for (Map.Entry<String, Set<String>> entry : configuration.entrySet()) {
 			Term requiredPropertyConstant = Utility.makeConstant(entry.getKey());
+			
+			// tripleEDB(Q, I, propertyConstant, X)
+			Atom tripleEDB_QIpX = Expressions.makeAtom(tripleEDB, q, i, propertyConstant, x);
 		
 			// tripleEDB(S, I, P, V)
 			Atom tripleEDB_SIPV = Expressions.makeAtom(tripleEDB, s, i, p, v);
@@ -48,7 +52,7 @@ public class ItemRequiresStatementPCC extends PropertyConstraintChecker {
 			// unequal(requiredPropertyConstant, P)
 			Atom unequal_rP = Expressions.makeAtom(InequalityHelper.unequal, requiredPropertyConstant, p);
 		
-			rules.addAll(StatementNonExistenceHelper.initRequireTriple(propertyConstant, requiredPropertyConstant, tripleEDB_SIPV, unequal_rP));
+			rules.addAll(StatementNonExistenceHelper.initRequireTriple(propertyConstant, requiredPropertyConstant, tripleEDB_QIpX, tripleEDB_SIPV, unequal_rP));
 			
 			Set<String> allowedValues = entry.getValue();				
 			if (allowedValues.size() != 0) {
@@ -56,6 +60,7 @@ public class ItemRequiresStatementPCC extends PropertyConstraintChecker {
 				Atom tripleEDB_SIrV = Expressions.makeAtom(tripleEDB, s, i, requiredPropertyConstant, v);
 				
 				List<Atom> conjunction = new ArrayList<Atom>();
+				conjunction.add(tripleEDB_QIpX);
 				conjunction.add(tripleEDB_SIrV);
 				
 				for (String allowedValue : allowedValues) {

### 2018-08-25 src/main/java/utility/StatementNonExistenceHelper.java

@@ -39,8 +39,8 @@ public class StatementNonExistenceHelper {
 		
 		List<Atom> firstConjunctionAtoms = new ArrayList<Atom>();
 		
-		firstConjunctionAtoms.add(first_SI);
 		firstConjunctionAtoms.addAll(conjunctionAtoms);
+		firstConjunctionAtoms.add(first_SI);
 		
 		// require(S, requiringTerm, requiredTerm) :- first(S, I), conjunctionAtoms
 		Rule firstRequire = Expressions.makeRule(require_Srr, toArray(firstConjunctionAtoms));
@@ -52,9 +52,9 @@ public class StatementNonExistenceHelper {
 		Atom require_Orr = Expressions.makeAtom(require, o, requiringTerm, requiredTerm);
 		
 		List<Atom> nextConjunctionAtoms = new ArrayList<Atom>();
+		nextConjunctionAtoms.addAll(conjunctionAtoms);
 		nextConjunctionAtoms.add(next_OS);
 		nextConjunctionAtoms.add(require_Orr);
-		nextConjunctionAtoms.addAll(conjunctionAtoms);
 		
 		// require(S, requiringTerm, requiredTerm) :- next(O, S), require(O, requiringTerm, requiredTerm), conjunctionAtoms
 		Rule nextRequire = Expressions.makeRule(require_Srr, toArray(nextConjunctionAtoms));

### 2018-08-25 thesis/paper/paper.tex
FIX: Typo

@@
 &\text{tripleEDB}(O, V, P, X) \wedge \text{last}(O, V) \wedge \\
-&\text{require}(O, text{propertyConstant}, \text{requiredPropertyConstant})
+&\text{require}(O, \text{propertyConstant}, \text{requiredPropertyConstant})
 \end{split}

### 2018-08-25 src/main/java/impl/PCC/ValueRequiresStatementPCC.java
Added "demand" rule to value requires statement

@@ -4,6 +4,8 @@ import static utility.SC.i;
 import static utility.SC.last;
 import static utility.SC.o;
 import static utility.SC.p;
+import static utility.SC.q;
+import static utility.SC.r;
 import static utility.SC.require;
 import static utility.SC.s;
 import static utility.SC.tripleEDB;
@@ -39,16 +41,19 @@ public class ValueRequiresStatementPCC extends PropertyConstraintChecker {
 	public List<Rule> rules() {
 		List<Rule> rules = new ArrayList<Rule>();
 		
+		// tripleEDB(Q, R, propertyConstant, I)
+		Atom tripleEDB_QRpI = Expressions.makeAtom(tripleEDB, q, r, propertyConstant, i);
+		
+		// tripleEDB(S, I, P, V)
+		Atom tripleEDB_SIPV = Expressions.makeAtom(tripleEDB, s, i, p, v);
+		
 		for (Map.Entry<String, Set<String>> entry : configuration.entrySet()) {
 			Term requiredPropertyConstant = Utility.makeConstant(entry.getKey());
-		
-			// tripleEDB(S, I, P, V)
-			Atom tripleEDB_SIPV = Expressions.makeAtom(tripleEDB, s, i, p, v);
 		
 			// unequal(requiredPropertyConstant, P)
 			Atom unequal_rP = Expressions.makeAtom(InequalityHelper.unequal, requiredPropertyConstant, p);
 		
-			rules.addAll(StatementNonExistenceHelper.initRequireTriple(propertyConstant, requiredPropertyConstant, tripleEDB_SIPV, unequal_rP));
+			rules.addAll(StatementNonExistenceHelper.initRequireTriple(propertyConstant, requiredPropertyConstant, tripleEDB_QRpI, tripleEDB_SIPV, unequal_rP));
 			
 			Set<String> allowedValues = entry.getValue();				
 			if (allowedValues.size() != 0) {
@@ -56,6 +61,7 @@ public class ValueRequiresStatementPCC extends PropertyConstraintChecker {
 				Atom tripleEDB_SIrV = Expressions.makeAtom(tripleEDB, s, i, requiredPropertyConstant, v);
 				
 				List<Atom> conjunction = new ArrayList<Atom>();
+				conjunction.add(tripleEDB_QRpI);
 				conjunction.add(tripleEDB_SIrV);
 				
 				for (String allowedValue : allowedValues) {

### 2018-08-25 thesis/paper/paper.tex
Added "demand" rule to item requires statement in paper

@@
 Inequality need to be computed between all properties and all values on value statements.
-The conjunctions remain the same and the final violation rule is
+The conjunctions change in the first triple and are:
+\begin{equation*}
+\begin{split}
+c_1 = &\text{tripleEDB}(Q, R, \text{propertyConstant}, I) \wedge \\
+&\text{tripleEDB}(S, I, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
+\end{split}
+\end{equation*} \(\)
+\begin{equation*}
+\begin{split}
+c_2 = &\text{tripleEDB}(Q, R, \text{propertyConstant}, I) \wedge \\
+&\text{tripleEDB}(S, I, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
+\end{split}
+\end{equation*}
+The final violation rule would then be:
 \begin{equation*}

### 2018-08-25 thesis/paper/paper.tex
Added "demand" rule to multi value

@@
 Alternatively, it could be said that if the item requires a second statement, the constraint is violated.
-Pre-processing extracts all items statements and inequality between all item statement properties as usual and the conjunction is \(c = \text{tripleEDB}(S, I, P, V) \wedge \text{unequal}(\text{propertyConstant}, P)\).
+Pre-processing extracts all items statements and inequality between all item statement properties as usual and the conjunction is
+\begin{equation*}
+\begin{split}
+c = &\text{tripleEDB}(Q, I, \text{propertyConstant}, X) \wedge \\
+&\text{tripleEDB}(S, I, P, V) \wedge \text{unequal}(\text{propertyConstant}, P)
+\end{split}
+\end{equation*}
 The predicate require\_second\([2]\) is introduced with the same meaning as require except that it will be applied after a required property has been found:
@@
 \begin{split}
-\text{require\_second}&(S, \text{propertyConstant}) \leftarrow \\
-&\text{next}(Q, R) \wedge \text{next}(R, S) \wedge \text{require}(Q, \text{propertyConstant}) \wedge t \wedge \\
-&\text{tripleEDB}(R, I, \text{propertyConstant}, C)
-\end{split}
-\end{equation*}
-\begin{equation*}
-\text{require\_second}(S, \text{propertyConstant}) \leftarrow \text{next}(R, S) \wedge \text{require\_second}(R, \text{propertyConstant}) \wedge t
+\text{require\_second}&(S, \text{propertyConstant}, \text{propertyConstant}) \leftarrow \\
+&\text{next}(R, S) \wedge \text{tripleEDB}(R, I, \text{propertyConstant}, C) \wedge \\
+&\text{next}(Q, R) \wedge \text{require}(Q, \text{propertyConstant}, \text{propertyConstant})  \wedge c
+\end{split}
+\end{equation*}
+\begin{equation*}
+\begin{split}
+\text{require\_second}&(S, \text{propertyConstant}, \text{propertyConstant}) \leftarrow \text{next}(R, S) \wedge \\
+&\text{require\_second}(R, \text{propertyConstant}, \text{propertyConstant}) \wedge c
+\end{split}
 \end{equation*}

### 2018-08-25 src/main/java/impl/PCC/MultiValuePCC.java

@@ -11,6 +11,7 @@ import static utility.SC.r;
 import static utility.SC.require;
 import static utility.SC.require_second;
 import static utility.SC.s;
+import static utility.SC.t;
 import static utility.SC.tripleEDB;
 import static utility.SC.v;
 import static utility.SC.x;
@@ -39,47 +40,53 @@ public class MultiValuePCC extends PropertyConstraintChecker {
 	public List<Rule> rules() {
 		List<Rule> rules = new ArrayList<Rule>();
 		
+		// tripleEDB(Q, I, propertyConstant, X)
+		Atom tripleEDB_QIpX = Expressions.makeAtom(tripleEDB, q, i, propertyConstant, x);
+		
 		// tripleEDB(S, I, P, V)
 		Atom tripleEDB_SIPV = Expressions.makeAtom(tripleEDB, s, i, p, v);
 		
 		// unequal(propertyConstant, P)
 		Atom unequal_pP = Expressions.makeAtom(InequalityHelper.unequal, propertyConstant, p);
 		
-		rules.addAll(StatementNonExistenceHelper.initRequireTriple(propertyConstant, propertyConstant, tripleEDB_SIPV, unequal_pP));
+		rules.addAll(StatementNonExistenceHelper.initRequireTriple(propertyConstant, propertyConstant, tripleEDB_QIpX, tripleEDB_SIPV, unequal_pP));
 		
 		// require_second(S, propertyConstant, propertyConstant)
 		Atom require_second_Spp = Expressions.makeAtom(require_second, s, propertyConstant, propertyConstant);
 		
-		// next(Q, R)
-		Atom next_QR = Expressions.makeAtom(next, q, r);
+		// next(T, R)
+		Atom next_TR = Expressions.makeAtom(next, t, r);
+		
+		// require(T, propertyConstant, propertyConstant)
+		Atom require_Tpp = Expressions.makeAtom(require, t, propertyConstant, propertyConstant);
 		
 		// next(R, S)
 		Atom next_RS = Expressions.makeAtom(next, r, s);
-		
-		// require(Q, propertyConstant, propertyConstant)
-		Atom require_Qpp = Expressions.makeAtom(require, q, propertyConstant, propertyConstant);
-		
+
 		// tripleEDB(R, I, propertyConstant, C)
 		Atom tripleEDB_RIpC = Expressions.makeAtom(tripleEDB, r, i, propertyConstant, c);
 		
[…]

### 2018-08-25 src/main/java/utility/SC.java

@@ -18,6 +18,7 @@ public class SC {
 	public final static String REQUIRE_INEQUALITY = "require_inequality";
 	
 	public final static String S = "s";
+	public final static String T = "t";
 	public final static String O = "o";
 	public final static String I = "i";
 	public final static String X = "x";
@@ -60,6 +61,7 @@ public class SC {
 	public final static Predicate require_inequality = Expressions.makePredicate(REQUIRE_INEQUALITY, 2);
 	
 	public final static Variable s = Expressions.makeVariable(S);
+	public final static Variable t = Expressions.makeVariable(T);
 	public final static Variable o = Expressions.makeVariable(O);
 	public final static Variable i = Expressions.makeVariable(I);
 	public final static Variable x = Expressions.makeVariable(X);

### 2018-08-25 thesis/paper/paper.tex
Added "demand" rule to mandatory qualifier

@@
 All constrained statements and their qualifiers have to be extracted and the order of the qualifiers pre-processed.
-Inequality must be established between all qualifier properties and the conjunction is \(c = \text{qualifierEDB}(S, P, V) \wedge \text{unequal}(P, \text{requiredPropertyConstant})\).
+Inequality must be established between all qualifier properties and the conjunction is
+\begin{equation*}
+\begin{split}
+c = &\text{tripleEDB}(S, I, \text{propertyConstant}, C) \wedge \\
+&\text{qualifierEDB}(S, P, V) \wedge \text{unequal}(P, \text{requiredPropertyConstant})
+\end{split}
+\end{equation*}
 With the usual require-pattern for qualifiers and the does\_not\_have rule added for brevity the violations would be

### 2018-08-25 src/main/java/impl/PCC/MandatoryQualifierPCC.java

@@ -1,12 +1,14 @@
 package impl.PCC;
 
 import static utility.SC.c;
+import static utility.SC.i;
 import static utility.SC.last_qualifier;
 import static utility.SC.p;
 import static utility.SC.q;
 import static utility.SC.qualifierEDB;
 import static utility.SC.require_qualifier;
 import static utility.SC.s;
+import static utility.SC.tripleEDB;
 import static utility.SC.v;
 
 import java.io.IOException;
@@ -38,6 +40,9 @@ public class MandatoryQualifierPCC extends PropertyConstraintChecker {
 	public List<Rule> rules() {
 		List<Rule> rules = new ArrayList<Rule>();
 		
+		// tripleEDB(S, I, propertyConstant, C)
+		Atom tripleEDB_SIpC = Expressions.makeAtom(tripleEDB, s, i, propertyConstant, c);
+		
 		// qualifierEDB(S, P, V)
 		Atom qualifierEDB_SPV = Expressions.makeAtom(qualifierEDB, s, p, v);
 		
@@ -47,7 +52,7 @@ public class MandatoryQualifierPCC extends PropertyConstraintChecker {
 			// unequal(P, requiredPropertyConstant)
 			Atom unequal_Pr = Expressions.makeAtom(InequalityHelper.unequal, p, requiredPropertyConstant);
 			
-			rules.addAll(StatementNonExistenceHelper.initRequireQualifier(requiredPropertyConstant, qualifierEDB_SPV, unequal_Pr));
+			rules.addAll(StatementNonExistenceHelper.initRequireQualifier(requiredPropertyConstant, tripleEDB_SIpC, qualifierEDB_SPV, unequal_Pr));
 			
 			// last_qualifier(S, Q, C)
 			Atom last_qualifier_SPV = Expressions.makeAtom(last_qualifier, s, q, c);

### 2018-08-26 src/main/java/impl/CC/ConstraintChecker.java
Easier testing clause

@@ -116,10 +116,8 @@ public abstract class ConstraintChecker {
 		while (results.hasNext()) {
 			QuerySolution solution = results.next();
 			//String property = solution.get("item").asResource().getLocalName();
-			//if (!(property.equals("P103"))) 
-			//	continue;
-
-			process(solution);
+			//if (property.equals("P19") || property.equals("P209")) 
+				process(solution);
 		}       
 
 		qexec.close();

### 2018-08-26 thesis/paper/paper.tex
Changed single value to use less variables

@@
 Only triples are eligible, since qualifiers cannot have further qualifier, and the order of qualifiers on constrained statements needs to be pre-processed as well since qualifier non-existence is relevant.
-The pattern needs to be applied for every separator, meaning the conjunction is: \(c = \text{qualifierEDB}(S, P, V) \wedge \text{unequal}(P, \text{separatorConstant})\) for each separator in \(S\).\\
+The pattern needs to be applied for every separator, meaning the conjunction is
+\begin{equation*}
+c = \text{tripleEDB}(S, I, \text{propertyConstant}, C) \wedge \text{qualifierEDB}(S, P, V) \wedge \text{unequal}(P, \text{separatorConstant})
+\end{equation*}
+for each separator in \(S\).\\
 
@@
 \(\forall X \subset S, Y = S \setminus X\) the following rule has to be added.
-Note that \(qualifierEDB(S, {X}, [X])\) denotes one atom for each element \(x \in X\) as introduced in section \ref{subsec_3_allowed_units} with \([X]\) representing a variable which is distinct for each atom unless they have the same \(x\).
-With the predicate does\_not\_have\([2]\) linking a statement ID to the separator it does not have via \(\text{does\_not\_have}(S, R) \leftarrow \text{last\_qualifier}(S, P, V) \wedge \text{require}(S, P, V, R)\) for brevity the rules are
+Note that \(\text{predicate}(\{X\})\) denotes one atom for each element \(x \in X\) as introduced in section \ref{subsec_3_allowed_units}.
+With the predicate does\_not\_have\([2]\) linking a statement ID to the separator it does not have via
+\begin{equation*}
+\begin{split}
+\text{does\_not\_have}(S, R) \leftarrow &\text{tripleEDB}(S, I, \text{propertyConstant}, C) \wedge \\
+&\text{last\_qualifier}(S, P, V) \wedge \text{require}(S, P, V, \text{propertyConstant}, R)
+\end{split}
+\end{equation*}
+and the predicate has\_same[2] linking two statement IDs to the separator for which the values are the same via
+\begin{equation*}
+\begin{split}
+\text{has\_same}(S, O, Q) \leftarrow &\text{tripleEDB}(S, I, \text{propertyConstant}, C) \wedge \\
+&\text{tripleEDB}(O, I, \text{propertyConstant}, X) \wedge \\
+&\text{qualifierEDB}(S, Q, V) \wedge \text{qualifierEDB}(O, Q, V)
+\end{split}
+\end{equation*}
+ for brevity the rules are
 \begin{equation*}
@@
 &\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \\
-&\text{tripleEDB}(O, I, \text{propertyConstant}, W) \wedge \\
+&\text{tripleEDB}(O, I, \text{propertyConstant}, C) \wedge \\
 &\text{unequal}(S, O) \wedge \\
-&\text{qualifierEDB}(S, \{X\}, [X]) \wedge \\
-&\text{qualifierEDB}(O, \{X\}, [X]) \wedge \\
+&\text{has\_same}(S, O, \{X\}) \wedge \\
 &\text{does\_not\_have}(S, \{Y\}) \wedge \\

### 2018-08-26 src/main/java/impl/PCC/MandatoryQualifierPCC.java
Changed require qualifier to need requiring term

@@ -52,19 +52,19 @@ public class MandatoryQualifierPCC extends PropertyConstraintChecker {
 			// unequal(P, requiredPropertyConstant)
 			Atom unequal_Pr = Expressions.makeAtom(InequalityHelper.unequal, p, requiredPropertyConstant);
 			
-			rules.addAll(StatementNonExistenceHelper.initRequireQualifier(requiredPropertyConstant, tripleEDB_SIpC, qualifierEDB_SPV, unequal_Pr));
+			rules.addAll(StatementNonExistenceHelper.initRequireQualifier(propertyConstant, requiredPropertyConstant, tripleEDB_SIpC, qualifierEDB_SPV, unequal_Pr));
 			
 			// last_qualifier(S, Q, C)
 			Atom last_qualifier_SPV = Expressions.makeAtom(last_qualifier, s, q, c);
 			
-			// require_qualifier(S, Q, C, requiredPropertyConstant)
-			Atom require_qualifier_SQCr = Expressions.makeAtom(require_qualifier, s, q, c, requiredPropertyConstant);
+			// require_qualifier(S, Q, C, propertyConstant, requiredPropertyConstant)
+			Atom require_qualifier_SQCpr = Expressions.makeAtom(require_qualifier, s, q, c, propertyConstant, requiredPropertyConstant);
 			
 			// violation_triple(S, I, propertyConstant, V) :-
 			//	tripleEDB(S, I, propertyConstant, V),
 			//	last_qualifier(S, Q, C),
 			//	require_qualifier(S, Q, C, requiredPropertyConstant)
-			Rule violation = Expressions.makeRule(violation_triple_SIpV, tripleEDB_SIpV, last_qualifier_SPV, require_qualifier_SQCr);
+			Rule violation = Expressions.makeRule(violation_triple_SIpV, tripleEDB_SIpV, last_qualifier_SPV, require_qualifier_SQCpr);
 			rules.add(violation);
 		}
 		

### 2018-08-26 src/main/java/utility/SC.java

@@ -49,6 +49,7 @@ public class SC {
 	public final static String UNIT = "unit";
 	
 	public final static String DOES_NOT_HAVE = "does_not_have";
+	public final static String HAS_SAME = "has_same";
 	
 	public final static Predicate tripleEDB = Expressions.makePredicate(TRIPLE, 4);
 	public final static Predicate qualifierEDB = Expressions.makePredicate(QUALIFIER, 3);
@@ -84,7 +85,7 @@ public class SC {
 	public static final Predicate next = Expressions.makePredicate(NEXT, 2);
 	public static final Predicate last = Expressions.makePredicate(LAST, 2);
 	
-	public static final Predicate require_qualifier = Expressions.makePredicate(REQUIRE_QUALIFIER, 4);
+	public static final Predicate require_qualifier = Expressions.makePredicate(REQUIRE_QUALIFIER, 5);
 	
 	public static final Predicate first_qualifier = Expressions.makePredicate(FIRST_QUALIFIER, 3);
 	public static final Predicate next_qualifier = Expressions.makePredicate(NEXT_QUALIFIER, 6);
@@ -96,4 +97,5 @@ public class SC {
 	public final static Predicate unit = Expressions.makePredicate(UNIT, 2);
 	
 	public final static Predicate does_not_have = Expressions.makePredicate(DOES_NOT_HAVE, 2);
+	public final static Predicate has_same = Expressions.makePredicate(HAS_SAME, 3);
 }

### 2018-08-26 src/main/java/utility/StatementNonExistenceHelper.java

@@ -62,14 +62,14 @@ public class StatementNonExistenceHelper {
 		return Arrays.asList(firstRequire, nextRequire);
 	}
 	
-	public static List<Rule> initRequireQualifier(Term requiredTerm, Atom...conjunctionAtoms) {
-		return initRequireQualifier(requiredTerm, Arrays.asList(conjunctionAtoms));
+	public static List<Rule> initRequireQualifier(Term requiringTerm, Term requiredTerm, Atom...conjunctionAtoms) {
+		return initRequireQualifier(requiringTerm, requiredTerm, Arrays.asList(conjunctionAtoms));
 	}
 	
-	public static List<Rule> initRequireQualifier(Term requiredTerm, List<Atom> conjunctionAtoms) {
+	public static List<Rule> initRequireQualifier(Term requiringTerm, Term requiredTerm, List<Atom> conjunctionAtoms) {
 		
-		// require_qualifier(S, P, V, requiredTerm)
-		Atom require_qualifier_SPVr = Expressions.makeAtom(require_qualifier, s, p, v, requiredTerm);
+		// require_qualifier(S, P, V, requiringTerm, requiredTerm)
+		Atom require_qualifier_SPVrr = Expressions.makeAtom(require_qualifier, s, p, v, requiringTerm, requiredTerm);
 		
 		// first_qualifier(S, P, V)
 		Atom first_qualifier_SPV = Expressions.makeAtom(first_qualifier, s, p, v);
@@ -78,22 +78,22 @@ public class StatementNonExistenceHelper {
 		firstConjunctionAtoms.add(first_qualifier_SPV);
 		firstConjunctionAtoms.addAll(conjunctionAtoms);
 		
-		// require_qualifier(S, P, V, requiredTerm) :- first_qualifier(S, P, V), conjunctionAtoms
-		Rule firstRequire = Expressions.makeRule(require_qualifier_SPVr, toArray(firstConjunctionAtoms));
+		// require_qualifier(S, P, V, requiringTerm, requiredTerm) :- first_qualifier(S, P, V), conjunctionAtoms
+		Rule firstRequire = Expressions.makeRule(require_qualifier_SPVrr, toArray(firstConjunctionAtoms));
 
 		// next_qualifier(O, X, C, S, P, V)
 		Atom next_qualifier_OXCSPV = Expressions.makeAtom(next_qualifier, o, x, c, s, p, v);
 		
[…]

### 2018-08-26 src/main/java/impl/PCC/SingleValuePCC.java
Changed single value to use less variables

@@ -2,10 +2,12 @@ package impl.PCC;
 
 import static utility.SC.c;
 import static utility.SC.does_not_have;
+import static utility.SC.has_same;
 import static utility.SC.i;
 import static utility.SC.last_qualifier;
 import static utility.SC.o;
 import static utility.SC.p;
+import static utility.SC.q;
 import static utility.SC.qualifierEDB;
 import static utility.SC.r;
 import static utility.SC.referenceEDB;
@@ -50,27 +52,30 @@ public class SingleValuePCC extends PropertyConstraintChecker {
 	public List<Rule> rules() { 	
 		List<Rule> rules = new ArrayList<Rule>();
 		
-		// tripleEDB(O, I, propertyConstant, C)
-		Atom tripleEDB_OIpC = Expressions.makeAtom(tripleEDB, o, i, propertyConstant, c);
+		// tripleEDB(O, I, propertyConstant, X)
+		Atom tripleEDB_OIpX = Expressions.makeAtom(tripleEDB, o, i, propertyConstant, x);
 		
 		// unequal(S, O)
 		Atom unequal_SO = Expressions.makeAtom(InequalityHelper.unequal, s, o);
 		
+		// qualifierEDB(S, propertyConstant, V)
+		Atom qualifierEDB_SpV = Expressions.makeAtom(qualifierEDB, s, propertyConstant, v);
+		
 		// qualifierEDB(S, propertyConstant, X)
-		Atom qualifierEDB_SpX = Expressions.makeAtom(qualifierEDB, o, propertyConstant, x);
+		Atom qualifierEDB_SpX = Expressions.makeAtom(qualifierEDB, s, propertyConstant, x);
 		
 		// unequal(V, X)
 		Atom unequal_VX = Expressions.makeAtom(InequalityHelper.unequal, v, x);
 		
 		// referenceEDB(S, propertyConstant, X)
 		Atom referenceEDB_SpX = Expressions.makeAtom(referenceEDB, s, propertyConstant, x);
-
+		
 		if (separators.size() == 0) {
 			// violation_triple(S, I, propertyConstant, V) :-
 			//	tripleEDB(S, I, propertyConstant, V),
-			//	tripleEDB(O, I, propertyConstant, C),
+			//	tripleEDB(O, I, propertyConstant, X),
 			//	unequal (S, O)
[…]

### 2018-08-26 src/main/java/impl/CC/SingleValueCC.java
Added missing queries to single value

@@ -1,9 +1,8 @@
 package impl.CC;
 
-import static utility.SC.first_qualifier;
-import static utility.SC.last_qualifier;
-import static utility.SC.next_qualifier;
 import static utility.SC.violation_triple_query;
+import static utility.SC.violation_qualifier_query;
+import static utility.SC.violation_reference_query;
 
 import java.io.IOException;
 import java.util.ArrayList;
@@ -18,13 +17,11 @@ import org.apache.jena.rdf.model.Literal;
 import org.apache.jena.rdf.model.RDFNode;
 import org.apache.log4j.Logger;
 import org.semanticweb.vlog4j.core.model.api.Atom;
-import org.semanticweb.vlog4j.core.reasoner.DataSource;
 import org.semanticweb.vlog4j.core.reasoner.exceptions.ReasonerStateException;
 
 import impl.PCC.PropertyConstraintChecker;
 import impl.PCC.SingleValuePCC;
 import main.Main;
-import utility.CsvGzFileDataSource;
 import utility.InequalityHelper;
 import utility.Utility;
 
@@ -57,7 +54,7 @@ public class SingleValueCC extends ConstraintChecker {
 	
 	@Override
 	protected Set<Atom> queries() {
-		return asSet(violation_triple_query);
+		return asSet(violation_triple_query, violation_qualifier_query, violation_reference_query);
 		//return asSet(Expressions.makeAtom(tripleEDB, s, i, p, v));
 	}
 	

### 2018-08-26 src/main/java/impl/TS/TripleSet.java
Added rank extraction

@@ -16,6 +16,7 @@ import org.wikidata.wdtk.datamodel.interfaces.SnakGroup;
 import org.wikidata.wdtk.datamodel.interfaces.Statement;
 import org.wikidata.wdtk.datamodel.interfaces.StatementDocument;
 import org.wikidata.wdtk.datamodel.interfaces.StatementGroup;
+import org.wikidata.wdtk.datamodel.interfaces.StatementRank;
 import org.wikidata.wdtk.datamodel.interfaces.Value;
 
 import main.Main;
@@ -35,6 +36,7 @@ public class TripleSet implements EntityDocumentProcessor {
 	TripleSetFile propertiesFile;
 	
 	TripleSetFile unitsFile;
+	TripleSetFile ranksFile;
 	
 	TripleSetFile firstFile;
 	TripleSetFile nextFile;
@@ -61,6 +63,7 @@ public class TripleSet implements EntityDocumentProcessor {
 		propertiesFile = new TripleSetFile("properties", SC.property);
 		
 		unitsFile = new TripleSetFile("units", SC.unit);
+		ranksFile = new TripleSetFile("ranks", SC.rank);
 		
 		firstFile = new TripleSetFile("first", SC.first);
 		nextFile = new TripleSetFile("next", SC.next);
@@ -97,6 +100,10 @@ public class TripleSet implements EntityDocumentProcessor {
 		return unitsFile.getFile();
 	}
 	
+	public File getRanksFile() throws IOException {
+		return ranksFile.getFile();
+	}
+	
 	public void loadTripleFile(Reasoner reasoner) throws IOException, ReasonerStateException {
 		tripleFile.loadFile(reasoner);
 	}
@@ -119,6 +126,10 @@ public class TripleSet implements EntityDocumentProcessor {
 	
 	public void loadUnitsFile(Reasoner reasoner) throws IOException, ReasonerStateException {
 		unitsFile.loadFile(reasoner);
+	}
+	
+	public void loadRanksFile(Reasoner reasoner) throws IOException, ReasonerStateException {
+		ranksFile.loadFile(reasoner);
 	}
 	
 	public void loadFirstFile(Reasoner reasoner) throws IOException, ReasonerStateException {
@@ -193,6 +204,10 @@ public class TripleSet implements EntityDocumentProcessor {
 		}
 	}
 	
+	void processRank(String id, StatementRank rank) {
+		ranksFile.write(id, rank.toString());
+	}
+	
 	public void processStatementDocument(StatementDocument statementDocument) {
[…]

### 2018-08-26 src/main/java/utility/SC.java

@@ -47,6 +47,7 @@ public class SC {
 	public final static String PROPERTY = "property";
 	
 	public final static String UNIT = "unit";
+	public final static String RANK = "rank";
 	
 	public final static String DOES_NOT_HAVE = "does_not_have";
 	public final static String HAS_SAME = "has_same";
@@ -95,6 +96,7 @@ public class SC {
 	public final static Predicate property = Expressions.makePredicate(PROPERTY, 1);
 	
 	public final static Predicate unit = Expressions.makePredicate(UNIT, 2);
+	public final static Predicate rank = Expressions.makePredicate(RANK, 2);
 	
 	public final static Predicate does_not_have = Expressions.makePredicate(DOES_NOT_HAVE, 2);
 	public final static Predicate has_same = Expressions.makePredicate(HAS_SAME, 3);

### 2018-08-26 thesis/paper/paper.tex
Added single best value

@@
 \hline
+Qualifiers & \makecell{separator (P4155) -- 0..* \\ qualifiers that must be different for exception} \\
+\hline
 Violation & \makecell{items with two constrained statements with rank preferred} \\
@@
 Thus the equation \ref{eq_no_separators_triple} is expanded to
-\begin{equation}
+\begin{equation*}
 \begin{split}
@@
 \end{split}
-\end{equation}
+\end{equation*}
 
@@
 &\text{unequal}(S, O) \wedge \\
-&\text{qualifierEDB}(S, \{X\}, [X]) \wedge \\
-&\text{qualifierEDB}(O, \{X\}, [X]) \wedge \\
+&\text{has\_same}(S, O, \{X\}) \wedge \\
 &\text{does\_not\_have}(S, \{Y\}) \wedge \\

### 2018-08-26 src/main/java/impl/CC/SingleBestValueCC.java

@@ -1,0 +1,107 @@
+package impl.CC;
+
+import static utility.SC.violation_triple_query;
+
+import java.io.IOException;
+import java.util.ArrayList;
+import java.util.HashMap;
+import java.util.HashSet;
+import java.util.List;
+import java.util.Map;
+import java.util.Set;
+
+import org.apache.jena.query.QuerySolution;
+import org.apache.jena.rdf.model.Literal;
+import org.apache.jena.rdf.model.RDFNode;
+import org.apache.log4j.Logger;
+import org.semanticweb.vlog4j.core.model.api.Atom;
+import org.semanticweb.vlog4j.core.reasoner.DataSource;
+import org.semanticweb.vlog4j.core.reasoner.exceptions.ReasonerStateException;
+
+import impl.PCC.PropertyConstraintChecker;
+import impl.PCC.SingleBestValuePCC;
+import impl.PCC.SingleValuePCC;
+import main.Main;
+import utility.CsvGzFileDataSource;
+import utility.InequalityHelper;
+import utility.Utility;
+
+public class SingleBestValueCC extends ConstraintChecker {
+	
+	final static Logger logger = Logger.getLogger(SingleBestValueCC.class);
+	
+	public static final String SEPARATOR = "P4155";
+	
+	Map<String, Set<String>> propertiesAndSeparators;
+
+	public SingleBestValueCC() throws IOException {
+		super("Q52060874");
+	}
+
+	@Override
+	void initDataField() {
+		propertiesAndSeparators = new HashMap<String, Set<String>>();
+	}
+	
+	@Override
+	protected Set<String> qualifiers() {
+		return asSet();
+	}
+
+	@Override
+	protected Set<String> concatQualifiers() {
+		return asSet(SEPARATOR);
+	}
+	
+	@Override
+	protected Set<Atom> queries() {
+		return asSet(violation_triple_query);
+	}
+	
+	@Override
+	protected void process(QuerySolution solution) {
+		String property = Utility.addBaseURI(solution.get("item").asResource().getLocalName());
+		
+		if (!propertiesAndSeparators.containsKey(property))
+			propertiesAndSeparators.put(property, new HashSet<String>());
+
+		RDFNode node = solution.get(SEPARATOR);
+		if (node.isLiteral()) {
+			Literal literal = node.asLiteral();
+			String content = literal.getString();
+			if (content.equals(""))
+				return;
[…]

### 2018-08-26 src/main/java/impl/PCC/SingleBestValuePCC.java

@@ -1,0 +1,172 @@
+package impl.PCC;
+
+import static utility.SC.c;
+import static utility.SC.does_not_have;
+import static utility.SC.has_same;
+import static utility.SC.i;
+import static utility.SC.last_qualifier;
+import static utility.SC.o;
+import static utility.SC.p;
+import static utility.SC.q;
+import static utility.SC.qualifierEDB;
+import static utility.SC.r;
+import static utility.SC.rank;
+import static utility.SC.require_qualifier;
+import static utility.SC.s;
+import static utility.SC.tripleEDB;
+import static utility.SC.v;
+import static utility.SC.x;
+
+import java.io.IOException;
+import java.util.ArrayList;
+import java.util.List;
+import java.util.Set;
+
+import org.apache.log4j.Logger;
+import org.semanticweb.vlog4j.core.model.api.Atom;
+import org.semanticweb.vlog4j.core.model.api.Constant;
+import org.semanticweb.vlog4j.core.model.api.Rule;
+import org.semanticweb.vlog4j.core.model.implementation.Expressions;
+import org.wikidata.wdtk.datamodel.interfaces.StatementRank;
+
+import com.google.common.collect.Sets;
+
+import utility.InequalityHelper;
+import utility.StatementNonExistenceHelper;
+
+public class SingleBestValuePCC extends PropertyConstraintChecker {
+	
+	final static Logger logger = Logger.getLogger(SingleBestValuePCC.class);
+	
+	final Set<String> separators;
+
+	public SingleBestValuePCC(String property_, Set<String> separators_) throws IOException {
+		super(property_);
+		separators = separators_;
+	}
+
+	@Override
+	public List<Rule> rules() { 	
+		List<Rule> rules = new ArrayList<Rule>();
+		
+		Constant preferredRankConstant = Expressions.makeConstant(StatementRank.PREFERRED.toString());
+		
+		// rank(S, preferredRankConstant)
+		Atom rank_Sp = Expressions.makeAtom(rank, s, preferredRankConstant);
+		
+		// tripleEDB(O, I, propertyConstant, X)
+		Atom tripleEDB_OIpX = Expressions.makeAtom(tripleEDB, o, i, propertyConstant, x);
+		
[…]

### 2018-08-26 src/main/java/main/Main.java

@@ -42,6 +42,7 @@ import impl.CC.NoneOfCC;
 import impl.CC.OneOfCC;
 import impl.CC.OneOfQualifierValueCC;
 import impl.CC.ScopeCC;
+import impl.CC.SingleBestValueCC;
 import impl.CC.SingleValueCC;
 import impl.CC.SymmetricCC;
 import impl.CC.ValueRequiresStatementCC;
@@ -195,6 +196,9 @@ public class Main {
 					break;
 				case "mandatoryqualifier":
 					checkers.add(new MandatoryQualifierCC());
+					break;
+				case "singlebestvalue":
+					checkers.add(new SingleBestValueCC());
 					break;
 				default:
 					System.out.println("Constraint " + constraintName + " is unknown.");

### 2018-08-27 thesis/paper/paper.tex
Incorporated notes from Markus

@@
 \begin{document}
-\chapter{Introduction/Preliminaries}
+\chapter{Introduction}
 Motivation
-\newpage
+
+\chapter{Preliminaries}
+
 \section{Wikidata}
@@
 The basic unit of information is a triple subject-predicate-object stating an item (subject) has a property (predicate) which is either another item or a value of some kind (object).
-These store the relations of entities with each other or with specific data.
+Triples store the relations of entities with each other or with specific data.
 An entity is either an item, which represents a topic, concept or real world entity, or a property, which denotes a relation an item has to another or with a value.
-They have numeric identifiers that are prefixed with Q and P respectively.
-A value is one of several data types noted in table \ref{tab_datatypes}, although it is also possible to use none or some as values for statements that have no value (akin to negation) or have a currently unknown value.
+They have numeric identifiers that are prefixed with Q and P and will be referred to as Q- and P-IDs, respectively.
+A value is one of several data types shown in Table \ref{tab_datatypes}, although it is also possible to use \emph{none} or \emph{some} as values for statements that have no value (akin to negation) or have a currently unknown value.
 Every property determines the data type it accepts as an object.
@@
 \begin{table}[H]
-\caption{wikidata data types}\label{tab_datatypes}
-\begin{tabularx}{\textwidth}{lllX}
+\caption{Wikidata data types}\label{tab_datatypes}
+\begin{tabularx}{\textwidth}{llp{8.5cm}}
 Data type & Fields & Description \\
 \hline
-string & string (string) & simple string literal \\
-globecoordinate & \makecell{latitude (float) \\ longitude (float) \\ precision (float) \\ globe(URI)} & \makecell{coordinates on a given celestial body \\ or geographic standard} \\
-quantity & \makecell {amount (decimal) \\ upperBound (decimal) \\ lowerBound (decimal) \\ unit (URI or 1)} & \makecell{quantity of a specified unit (or none) \\ with uncertainty} \\
-time & \makecell {time (string) \\ calendarmodel (URI) \\ precision (integer)} & \makecell {time as a string according to calendarmodel \\ (julian or gregorian),\\ with precision mapping to orders of magnitude \\ (e.g..
-10 years, months)} \\
-\end{tabularx}
-\end{table}
+commons media & file (string) & name of a file in the "File" namespace on Wikimedia Commons \\
+globecoordinate & \makecell{latitude (float) \\ longitude (float) \\ precision (float) \\ globe (URI)} & coordinates on a given celestial body or geographic standard \\
+item & item (Q-ID) & the Q-ID of an item on Wikidata \\
+property & property (P-ID) & the P-ID of an item on Wikidata \\
+string & string (string) & simple string literal that does not need to be translated into different languages \\
+monolingual text & string (string) & simple string literal that is written in one specific language and not to be translated \\
+quantity & \makecell {amount (decimal) \\ upperBound (decimal) \\ lowerBound (decimal) \\ unit (URI or 1)} & quantity of a specified unit (or none) with uncertainty \\
+time & \makecell {time (string) \\ calendarmodel (URI) \\ precision (integer)} & time as a string according to calendarmodel (julian or gregorian), with precision mapping to orders of magnitude (e.g..
+10 years, months) \\
+URL & url(URL) & a general URL for an external resource \\
+mathematical expression & formula (string) & string that is formatted to display as a formula \\
+external identifier & id (string) & string representing an identifier of an external system \\
+geographic shape & map data (string) & name of a file in the "Data" namespace on Wikimedia Commons \\
+tabular data & tabular data (string) & name of a file in the "Data" namespace on Wikimedia Commons \\
+\end{tabularx}
+\end{table}
+\todo[inline]{For globe(URI): JSON data model states "typically refer to a data item on wikidata.org", which I would read as can be any URI}
 
 While this is similar to RDF and Wikidata has in fact been exported as RDF \cite{EGKMV2014}, the data structure is more complex than simple triples.
-It allows, to store additional information about statements.
+It allows to store additional information about statements.
 This is done via a statement ID, which is assigned to every triple storing information about an item.
-Every statement can be references by qualifiers and references which contain a predicate and an object, but have no item as subject but instead the statement ID.
+Every statement can be referenced by qualifiers, which consist of the corresponding statement id, a predicate and an object.
 Qualifiers provide contextual information for statements.
@@
 determination method for measurements or the elevation above sea level for the highest point.
-References are information that supports the claim by citing the source and possibly additional information like when it was retrieved or when the source was published.
+Additionally, references can be added to statements.
+References support the claim by citing the source and possibly additional information, e.g.
+when it was retrieved or when the source was published as property-object pairs.
 
@@
 \section{Datalog}
-Datalog is a logic programming language.
-It is fully declarative and uses existential rules to derive additional knowledge from given facts.
-The most basic unit of information in datalog is a \emph{constant}, which will be denoted as someConstant.
-In the context of Wikidata any item or property, e.g.
+Datalog is a logic query language.
+It is fully declarative and uses rules to derive additional knowledge from given facts.
+The most basic domain element in datalog is a \emph{constant}, which will be denoted as \emph{someConstant}.
+In the context of Wikidata, any item or property, e.g.
 Earth (Q2) or head of government (P6), could be represented as a constant.
-Relations in datalog are expressed as \emph{predicates} with a specified arity \(n\), written as predicate\([n]\).
+Relations in datalog are expressed using \emph{predicates} with a specified arity \(n\), written as predicate\([n]\).
 To stay close to Wikidata in our examples let us take the predicate statement\([3]\) as an example.
 This would be used to express the \emph{item}-\emph{property}-\emph{value} relation with each position representing one part of the triple.\\
-Additionally, datalog also knows \emph{variables}, which will be capitalized like \(X\).
+Additionally, datalog also provides \emph{variables}, which will be capitalized like \(X\).
 Constants and variables are collectively known as \emph{terms} and are, together with predicates, used to form \emph{atoms}.
 An atom has the form predicate\((t_1, \ldots , t_n)\) with \(n\) as the arity of the predicate and \(t_1,\ldots,t_n\) terms.\\
-Atoms are the building blocks of \emph{existential rules}, which are used to derive information in datalog.
-A rule is premise that implies a conclusion, and both are conjunctions of atoms.
-Denote \(P_n\) as the premise atoms, \(C_n\) as the conclusion atoms.
-The variables used are divided into three disjunct sets \(V_p\), \(V_c\), and \(V_r\) which are the variables occurring only in premise atoms, only in conclusion atoms, and occurring in both respectively.
+Atoms are the building blocks of \emph{rules}, which are used to derive information in datalog.
+A rule is a premise that implies a conclusion, and both are conjunctions of atoms.
+Let \(P_n\) denote  premise atoms and \(C_n\) the conclusion atoms.
+The variables used are divided into three mutually disjoint sets \(V_p\), \(V_c\), and \(V_r\) which are the variables occurring only in premise atoms, only in conclusion atoms, and occurring in both respectively.
 The general from of an existential rule is (with \(k \in \mathbb{N}^+\), \(l \in \mathbb{N}\))
@@
 \(V_c\) are called existential variables.
-If \(V_c = \emptyset\) the rule is called a \emph{datalog rule} and if \(l = 0\) is called a \emph{fact}.
+If \(V_c = \emptyset\) the rule is called a \emph{datalog rule} and if \(l = 0\) it is called a \emph{fact}.
 Since \(\forall p \in V_p, r \in V_r\) can be taken as given for all rules, it will be omitted in this work.
 
-Within the rule predicates we can differentiate between two different schemas: the \emph{extensional schema} and the \emph{intensional schema}, called EDB and IDB \todo{Shouldn't this be called \\EDP/IDP?} (extensional/intensional database predicate).
-An EDB may only occur in the premise of a rule, meaning it can never be derived.
-All others (predicates that appear in the conclusion at least once) are IDB.
-
-For the following example taken from Wikidata descriptive names will be used instead of the IDs to increase readability.
+Within the rule predicates we can differentiate between two different schemas: the \emph{extensional schema} and the \emph{intensional schema}, called EDB and IDB (extensional/intensional database).
+An EDB predicate may only occur in the premise of a rule, meaning it can never be derived.
+All predicates that appear in the conclusion at least once are IDB predicates.
+
+For the following example, taken from Wikidata, descriptive names will be used instead of the IDs to increase readability.
 The individual entities and properties are introduced with their ID beforehand so the example can be retraced.
-The blue whales (Q42196) parent taxon (P171) is balaenoptera (Q133320), which in turn has the parent taxon rorqual (Q232829).
+The species \emph{blue whale} (Q42196) has the \emph{parent taxon} (P171) \emph{balaenoptera} (Q133320), which in turn has the parent taxon \emph{rorqual} (Q232829).
 The predicates used are tripleEDB\([3]\) as the input and isInTaxonIDB\([2]\), which will be derived.
-The basic facts are:\\
+The basic facts are:
 \[\text{tripleEDB}(\text{blue whale}, \text{parent taxon}, \text{balaenoptera})\]
@@
 To model the basic representation of triples, qualifiers and references three predicates are introduced.\\
-tripleEDB\([4]\) represents a single statement consisting of the statement ID, the item ID, the property ID and the value as either and item ID or a string-representation of the data value.\\
+tripleEDB\([4]\) represents a single statement consisting of the statement ID, the item ID, the property ID and the value as either an item ID or a string-representation of the data value.\\
 qualifierEDB\([3]\) and referenceEDB\([3]\) represent the qualifiers and references on a given statement ID with their respective property id and the value as usual.\\

### 2018-08-27 tools/summarizeOutput.py
Added script to parse the output

@@ -1,0 +1,65 @@
+import csv
+import re
+
+constraintIDLabel = "ConstraintID"
+violationsLabel = "Violations"
+totalTimeLabel = "Total time"
+runtimeMaterializationLabel = "Runtime materialization"
+iterationsLabel = "Iterations"
+derivationsLabel = "Derivations"
+rulesLabel = "Rules"
+demandRulesLabel = "Demand rules"
+totalRulesLabel = "Total rules"
+
+with open("log") as log, open("summary.tsv", "w") as summary:
+    writer = csv.DictWriter(summary, [constraintIDLabel, violationsLabel, totalTimeLabel, runtimeMaterializationLabel, iterationsLabel, derivationsLabel, rulesLabel, demandRulesLabel, totalRulesLabel], delimiter="\t")
+    writer.writeheader()
+
+    blocks = list()
+    block = list()
+
+    rules = re.compile("Created ([0-9]+) rules.")
+    demandRules = re.compile("Created ([0-9]+) additional demand-rules.")
+    totalRules = re.compile("Added ([0-9]+) rules total.")
+    iterations = re.compile("Iterations=([0-9]+)")
+    runtime = re.compile("Runtime materialization = ([0-9]+),([0-9]+) milliseconds")
+    derivations = re.compile("Total # derivations: ([0-9]+)")
+    time = re.compile("Total time elapsed: ([0-9]+)ms")
+    constraint = re.compile("Constraint: (Q[0-9]+), violations: ([0-9]+)")
+
+    stats = dict()
+
+    for line in log:
+        rulesResult = rules.search(line)
+        if rulesResult:
+            stats[rulesLabel] = rulesResult.group(1)
+
+        demandRulesResult = demandRules.search(line)
+        if demandRulesResult:
+            stats[demandRulesLabel] = demandRulesResult.group(1)
+
+        totalRulesResult = totalRules.search(line)
+        if totalRulesResult:
+            stats[totalRulesLabel] = totalRulesResult.group(1)
+
+        iterationsResult = iterations.search(line)
+        if iterationsResult:
+            stats[iterationsLabel] = iterationsResult.group(1)
+
+        runtimeResult = runtime.search(line)
+        if runtimeResult:
+            stats[runtimeMaterializationLabel] = runtimeResult.group(1)
+
+        derivationsResult = derivations.search(line)
+        if derivationsResult:
+            stats[derivationsLabel] = derivationsResult.group(1)
+
+        timeResult = time.search(line)
[…]

### 2018-08-27 thesis/paper/paper.tex
FIX: Missing referenceEDB-hash

@@
 tripleEDB\([4]\) represents a single statement consisting of the statement ID, the item ID, the property ID and the value as either an item ID or a string-representation of the data value.\\
-qualifierEDB\([3]\) and referenceEDB\([3]\) represent the qualifiers and references on a given statement ID with their respective property id and the value as usual.\\
+qualifierEDB\([3]\) represent the qualifiers on a given statement ID with their respective property id and the value as usual.
+ReferenceEDB\([4]\) represents the statement ID, the reference hash, property id and the value.\\
 
@@
 \[\text{violation\_qualifier}(S, \text{propertyConstant}, V) \leftarrow \text{qualifierEDB}(S, \text{propertyConstant}, V)\]
-\[\text{violation\_reference}(S, \text{propertyConstant}, V) \leftarrow \text{referenceEDB}(S, \text{propertyConstant}, V)\]
+\[\text{violation\_reference}(S, \text{propertyConstant}, V) \leftarrow \text{referenceEDB}(S, H, \text{propertyConstant}, V)\]
 depending on if as main value, as qualifier, or as reference was not set in the qualifiers.\\
@@
 \text{violation\_reference}&(S, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
-&\text{referenceEDB}(S, \text{propertyConstant}, \text{forbiddenValueConstant})
+&\text{referenceEDB}(S, H, \text{propertyConstant}, \text{forbiddenValueConstant})
 \end{split}
@@
 \text{violation\_reference}&(S, I \text{propertyConstant}, V) \leftarrow \\
-&\text{referenceEDB}(S, \text{propertyConstant}, V) \wedge \\
+&\text{referenceEDB}(S, H, \text{propertyConstant}, V) \wedge \\
 &\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
@@
 \text{violation\_reference}&(S, \text{propertyConstant}, V) \leftarrow \\
-&\text{referenceEDB}(S, \text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
+&\text{referenceEDB}(S, H, \text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
 \end{split}
@@
 \text{violation\_reference}&(S, \text{propertyConstant}, V) \leftarrow \\
-&\text{referenceEDB}(S, \text{propertyConstant}, V) \wedge \\
-&\text{referenceEDB}(S, \text{propertyConstant}, X) \wedge \\
+&\text{referenceEDB}(S, H, \text{propertyConstant}, V) \wedge \\
+&\text{referenceEDB}(S, J, \text{propertyConstant}, X) \wedge \\
 &\text{unequal}(V, X)

### 2018-08-27 src/main/java/impl/PCC/AllowedUnitsPCC.java

@@ -63,11 +63,11 @@ public class AllowedUnitsPCC extends PropertyConstraintChecker {
 		rules.add(qualifierViolation);
 		
 		List<Atom> violation_reference_conjunction = new ArrayList<Atom>();
-		violation_reference_conjunction.add(qualifierEDB_SpV);
+		violation_reference_conjunction.add(referenceEDB_SHpV);
 		violation_reference_conjunction.addAll(unequal_conjunction);
 		
-		// violation_reference(S, propertyConstant, V) :- referenceEDB(S, propertyConstant, V), unit(V, U), unequal({A}, U)
-		Rule referenceViolation = Expressions.makeRule(violation_reference_SpV, Utility.toArray(violation_reference_conjunction));
+		// violation_reference(S, H, propertyConstant, V) :- referenceEDB(S, H, propertyConstant, V), unit(V, U), unequal({A}, U)
+		Rule referenceViolation = Expressions.makeRule(violation_reference_SHpV, Utility.toArray(violation_reference_conjunction));
 		rules.add(referenceViolation);
 		
 		return rules;

### 2018-08-27 src/main/java/impl/PCC/NoneOfPCC.java

@@ -1,5 +1,6 @@
 package impl.PCC;
 
+import static utility.SC.h;
 import static utility.SC.i;
 import static utility.SC.qualifierEDB;
 import static utility.SC.referenceEDB;
@@ -64,14 +65,14 @@ public class NoneOfPCC extends PropertyConstraintChecker {
 			
 			rules.add(conflict_qualifier);
 			
-			// violation_reference(S, propertyConstant, confValueConstant)
-			Atom violation_reference_Spc = Expressions.makeAtom(violation_reference, s, propertyConstant, confValueConstant);
+			// violation_reference(S, H, propertyConstant, confValueConstant)
+			Atom violation_reference_SHpc = Expressions.makeAtom(violation_reference, s, h, propertyConstant, confValueConstant);
 			
-			// referenceEDB(S, propertyConstant, confValueConstant)
-			Atom referenceEDB_Spc = Expressions.makeAtom(referenceEDB, s, propertyConstant, confValueConstant);
+			// referenceEDB(S, H, propertyConstant, confValueConstant)
+			Atom referenceEDB_SHpc = Expressions.makeAtom(referenceEDB, s, h, propertyConstant, confValueConstant);
 			
-			// violation_reference(S, propertyConstant, confValueConstant) :- referenceEDB(S, propertyConstant, confValueConstant)
-			Rule conflict_reference = Expressions.makeRule(violation_reference_Spc, referenceEDB_Spc);
+			// violation_reference(S, H, propertyConstant, confValueConstant) :- referenceEDB(S, H, propertyConstant, confValueConstant)
+			Rule conflict_reference = Expressions.makeRule(violation_reference_SHpc, referenceEDB_SHpc);
 			
 			rules.add(conflict_reference);
 		}

### 2018-08-27 src/main/java/impl/PCC/OneOfPCC.java

@@ -55,11 +55,11 @@ public class OneOfPCC extends PropertyConstraintChecker {
 		Rule violationQualifier = Expressions.makeRule(violation_qualifier_SpV, Utility.toArray(violation_qualifier_conjunction));
 		
 		List<Atom> violation_reference_conjunction = new ArrayList<Atom>();
-		violation_reference_conjunction.add(referenceEDB_SpV);
+		violation_reference_conjunction.add(referenceEDB_SHpV);
 		violation_reference_conjunction.addAll(unequal_conjunction);
 		
-		// violation_reference(S, propertyConstant, V) :- referenceEDB(S, propertyConstant, V), unequal({A}, V)
-		Rule violationReference = Expressions.makeRule(violation_qualifier_SpV, Utility.toArray(violation_reference_conjunction));
+		// violation_reference(S, H, propertyConstant, V) :- referenceEDB(S, H, propertyConstant, V), unequal({A}, V)
+		Rule violationReference = Expressions.makeRule(violation_reference_SHpV, Utility.toArray(violation_reference_conjunction));
 		
 		rules.add(violationTriple);
 		rules.add(violationQualifier);

### 2018-08-27 src/main/java/impl/PCC/PropertyConstraintChecker.java

@@ -3,6 +3,7 @@
  */
 package impl.PCC;
 
+import static utility.SC.h;
 import static utility.SC.i;
 import static utility.SC.qualifierEDB;
 import static utility.SC.referenceEDB;
@@ -36,15 +37,15 @@ public abstract class PropertyConstraintChecker {
 	protected final Atom violation_triple_SIpV;
 	// violation_qualifier(S, propertyConstant, V)
 	protected final Atom violation_qualifier_SpV;
-	// violation_reference(S, propertyConstant, V)
-	protected final Atom violation_reference_SpV;
+	// violation_reference(S, H, propertyConstant, V)
+	protected final Atom violation_reference_SHpV;
 	
 	// tripleEDB(S, I, propertyConstant, V)
 	protected final Atom tripleEDB_SIpV;
 	// qualifierEDB(S, propertyConstant, V)
 	protected final Atom qualifierEDB_SpV;
-	// referenceEDB(S, propertyConstant, V)
-	protected final Atom referenceEDB_SpV;
+	// referenceEDB(S, H, propertyConstant, V)
+	protected final Atom referenceEDB_SHpV;
 	
 	protected final Constant propertyConstant;
 	
@@ -58,11 +59,11 @@ public abstract class PropertyConstraintChecker {
 		
 		violation_triple_SIpV = Expressions.makeAtom(violation_triple, s, i, propertyConstant, v);
 		violation_qualifier_SpV = Expressions.makeAtom(violation_qualifier, s, propertyConstant, v);
-		violation_reference_SpV = Expressions.makeAtom(violation_reference, s, propertyConstant, v);
+		violation_reference_SHpV = Expressions.makeAtom(violation_reference, s, h, propertyConstant, v);
 		
 		tripleEDB_SIpV = Expressions.makeAtom(tripleEDB, s, i, propertyConstant, v);
 		qualifierEDB_SpV = Expressions.makeAtom(qualifierEDB, s, propertyConstant, v);
-		referenceEDB_SpV = Expressions.makeAtom(referenceEDB, s, propertyConstant, v);
+		referenceEDB_SHpV = Expressions.makeAtom(referenceEDB, s, h, propertyConstant, v);
 	}
 	
 	public abstract List<Rule> rules();

### 2018-08-27 src/main/java/impl/PCC/ScopePCC.java

@@ -41,8 +41,8 @@ public class ScopePCC extends PropertyConstraintChecker {
 		Rule notTriple = Expressions.makeRule(violation_triple_SIpV, tripleEDB_SIpV);
 		// violation_qualfier(S, propertyConstant, V) :- qualifierEDB(S, propertyConstant, V)
 		Rule notQualifier = Expressions.makeRule(violation_qualifier_SpV, qualifierEDB_SpV);
-		// violation_reference(S, propertyConstant, V) :- referenceEDB(S, propertyConstant, V)
-		Rule notReference = Expressions.makeRule(violation_reference_SpV, referenceEDB_SpV); 
+		// violation_reference(S, H, propertyConstant, V) :- referenceEDB(S, H, propertyConstant, V)
+		Rule notReference = Expressions.makeRule(violation_reference_SHpV, referenceEDB_SHpV); 
 		
 		rules.add(notTriple);
 		rules.add(notQualifier);

### 2018-08-27 src/main/java/impl/PCC/SingleValuePCC.java

@@ -3,6 +3,7 @@ package impl.PCC;
 import static utility.SC.c;
 import static utility.SC.does_not_have;
 import static utility.SC.has_same;
+import static utility.SC.g;
 import static utility.SC.i;
 import static utility.SC.last_qualifier;
 import static utility.SC.o;
@@ -63,8 +64,8 @@ public class SingleValuePCC extends PropertyConstraintChecker {
 		// unequal(V, X)
 		Atom unequal_VX = Expressions.makeAtom(InequalityHelper.unequal, v, x);
 		
-		// referenceEDB(S, propertyConstant, X)
-		Atom referenceEDB_SpX = Expressions.makeAtom(referenceEDB, s, propertyConstant, x);
+		// referenceEDB(S, G, propertyConstant, X)
+		Atom referenceEDB_SGpX = Expressions.makeAtom(referenceEDB, s, g, propertyConstant, x);
 		
 		if (separators.size() == 0) {
 			// violation_triple(S, I, propertyConstant, V) :-
@@ -81,11 +82,11 @@ public class SingleValuePCC extends PropertyConstraintChecker {
 			Rule violationQualifier = Expressions.makeRule(violation_qualifier_SpV, qualifierEDB_SpV, qualifierEDB_SpX, unequal_VX);
 			rules.add(violationQualifier);
 			
-			// violation_reference(S, propertyConstant, V) :-
-			//	referenceEDB(S, propertyConstant, V),
-			//	referenceEDB(S, propertyConstant, X),
+			// violation_reference(S, H, propertyConstant, V) :-
+			//	referenceEDB(S, H, propertyConstant, V),
+			//	referenceEDB(S, G, propertyConstant, X),
 			// 	unequal(V, X)
-			Rule violationReference = Expressions.makeRule(violation_reference_SpV, referenceEDB_SpV, referenceEDB_SpX, unequal_VX);
+			Rule violationReference = Expressions.makeRule(violation_reference_SHpV, referenceEDB_SHpV, referenceEDB_SGpX, unequal_VX);
 			rules.add(violationReference);
 		} else {
 			// tripleEDB(S, I, propertyConstant, C)

### 2018-08-27 src/main/java/impl/TS/TripleSet.java

@@ -236,6 +236,8 @@ public class TripleSet implements EntityDocumentProcessor {
 						processUnit(qualifier_object, qualifier_value);
 					}
 				}
+				// TODO this is not extracting the actual hash, but it does not seem to be available
+				int semiHash = 0;
 				for (Reference reference : statement.getReferences()) {
 					for (SnakGroup rGroup : reference.getSnakGroups()) {
 						String reference_predicate = rGroup.getProperty().getIri();
@@ -245,10 +247,11 @@ public class TripleSet implements EntityDocumentProcessor {
 							if (reference_value != null) {
 								reference_object = reference_value.accept(new OutputValueVisitor());
 							}
-							referenceFile.write(id, reference_predicate, reference_object);
+							referenceFile.write(id, String.valueOf(semiHash), reference_predicate, reference_object);
 							processUnit(reference_object, reference_value);
 						}
 					}
+					semiHash++;
 				}
 			}
 		}

### 2018-08-27 src/main/java/utility/SC.java

@@ -17,6 +17,8 @@ public class SC {
 	
 	public final static String REQUIRE_INEQUALITY = "require_inequality";
 	
+	public final static String H = "h";
+	public final static String G = "G";
 	public final static String S = "s";
 	public final static String T = "t";
 	public final static String O = "o";
@@ -54,14 +56,17 @@ public class SC {
 	
 	public final static Predicate tripleEDB = Expressions.makePredicate(TRIPLE, 4);
 	public final static Predicate qualifierEDB = Expressions.makePredicate(QUALIFIER, 3);
-	public final static Predicate referenceEDB = Expressions.makePredicate(REFERENCE, 3);
+	public final static Predicate referenceEDB = Expressions.makePredicate(REFERENCE, 4);
 	
 	public final static Predicate violation_triple = Expressions.makePredicate(VIOLATION_TRIPLE, 4);
 	public final static Predicate violation_qualifier = Expressions.makePredicate(VIOLATION_QUALIFIER, 3);
-	public final static Predicate violation_reference = Expressions.makePredicate(VIOLATION_REFERENCE, 3);
+	public final static Predicate violation_reference = Expressions.makePredicate(VIOLATION_REFERENCE, 4);
 	
 	public final static Predicate require_inequality = Expressions.makePredicate(REQUIRE_INEQUALITY, 2);
 	
+	
+	public final static Variable h = Expressions.makeVariable(H);
+	public final static Variable g = Expressions.makeVariable(G);
 	public final static Variable s = Expressions.makeVariable(S);
 	public final static Variable t = Expressions.makeVariable(T);
 	public final static Variable o = Expressions.makeVariable(O);
@@ -77,7 +82,7 @@ public class SC {
 	
 	public final static Atom violation_triple_query = Expressions.makeAtom(violation_triple, s, i, p, v);
 	public final static Atom violation_qualifier_query = Expressions.makeAtom(violation_qualifier, s, p, v);
-	public final static Atom violation_reference_query = Expressions.makeAtom(violation_reference, s, p, v);
+	public final static Atom violation_reference_query = Expressions.makeAtom(violation_reference, s, h, p, v);
 	
 	public static final Predicate require = Expressions.makePredicate(REQUIRE, 3);
 	public static final Predicate require_second = Expressions.makePredicate(REQUIRE_SECOND, 3);

### 2018-08-27 thesis/paper/paper.tex
Renamed tripleEDB to statementEDB in paper

@@
 To model the basic representation of triples, qualifiers and references three predicates are introduced.\\
-tripleEDB\([4]\) represents a single statement consisting of the statement ID, the item ID, the property ID and the value as either an item ID or a string-representation of the data value.\\
+statementEDB\([4]\) represents a single statement consisting of the statement ID, the item ID, the property ID and the value as either an item ID or a string-representation of the data value.\\
 qualifierEDB\([3]\) represent the qualifiers on a given statement ID with their respective property id and the value as usual.
@@
 
-Since only three possibilities exist, this is equivalent to forbidding all other possibilities and the rules are simple mapping all of tripleEDB, qualifierEDB and referenceEDB with the constrained property that are not allowed.
+Since only three possibilities exist, this is equivalent to forbidding all other possibilities and the rules are simple mapping all of statementEDB, qualifierEDB and referenceEDB with the constrained property that are not allowed.
 
@@
 \end{table}
-The rule modelling can simply be done by choosing all tripleEDB with the constrained property and then all tripleEDB on the same item with the conflicting property, specifying the values as well if necessary.
+The rule modelling can simply be done by choosing all statementEDB with the constrained property and then all statementEDB on the same item with the conflicting property, specifying the values as well if necessary.
 
@@
 
-Since there are only three options (and only two in use), violations can be found without inequality by finding the tripleEDB with the constrained property whose subject is of the type that is not allowed, meaning either wikibase item or wikibase property.
+Since there are only three options (and only two in use), violations can be found without inequality by finding the statementEDB with the constrained property whose subject is of the type that is not allowed, meaning either wikibase item or wikibase property.
 
@@
 Modelling this constraint is very straightforward, since each forbidden value forms a property-value pair with the constrained statement that may not exists.
-Thus a violations is a tripleEDB with the constrained statement and a forbidden value.
+Thus a violations is a statementEDB with the constrained statement and a forbidden value.
 
@@
 \end{table}
-To model the distinct values constraint it is necessary to find triples tripleEDB with the constrained property and the same value whose statement id is unequal.
+To model the distinct values constraint it is necessary to find statements statementEDB with the constrained property and the same value whose statement id is unequal.
 In theory the constraint can also be applied to qualifiers and references by setting the respective value on a constraint scope (P4680) qualifier.
@@
 Modelling the constraint requires a predicate associating the value or the statement with its unit and inequality between every unit and all allowed units.
-Then a violation can be found by finding all tripleEDB, qualifierEDB and referenceEDB with the constrained property and a unit unequal to all allowed units.
+Then a violation can be found by finding all statementEDB, qualifierEDB and referenceEDB with the constrained property and a unit unequal to all allowed units.
 Theoretically some of these can be excluded by using the constraint scope (P4680) qualifier on the constraint but there is no current case where this is done.
@@
 To represent a violation of this constraint in datalog inequality between all specified allowed qualifiers and the properties of the qualifiers on constrained statements is required.
-A tripleEDB with the constrained property and a qualifierEDB on that statement with a property different from all allowed qualifiers would violate the constraint.
+A statementEDB with the constrained property and a qualifierEDB on that statement with a property different from all allowed qualifiers would violate the constraint.
 
@@
 
-Modelling the one-of constraint requires a rule finding a tripleEDB with the constrained property whose value is unequal to all allowed values, meaning the inequality between each value of constrained statements and all allowed values must be given.
+Modelling the one-of constraint requires a rule finding a statementEDB with the constrained property whose value is unequal to all allowed values, meaning the inequality between each value of constrained statements and all allowed values must be given.
 
@@
 The constraint can be modelled if the inequality between a value of a qualifier property with all qualifier values is given.
-The violations could then be found by taking a tripleEDB with the constrained property and a qualifier with the qualifier property on the former statement whose value is unequal to all qualifier values.
+The violations could then be found by taking a statementEDB with the constrained property and a qualifier with the qualifier property on the former statement whose value is unequal to all qualifier values.
 
@@
 An item violates the constraint if it a) has a statement using constrained property as predicate, b1) has no statement using the required property as predicate, or (if allowed values are specified) b2) has only statement using required property as predicate that do not use one of the allowed values as value.\\
-While condition a) can easily be tested by a tripleEDB-atom with the constrained property, condition b1) and b2) pose the question if an item does not have a statement.
+While condition a) can easily be tested by a statementEDB-atom with the constrained property, condition b1) and b2) pose the question if an item does not have a statement.
 This is addressed in section [?].
@@
 If separators are specified, a violation as above can be ignored if the statements each have at least one qualifier with the same separator as predicate and different values or an unequal number of separator qualifiers.\\
-In the case without separators the modelling idea would be to find two triples tripleEDB with the same item and constrained property and test if their id is unequal.
+In the case without separators the modelling idea would be to find two triples statementEDB with the same item and constrained property and test if their id is unequal.
 If separators were specified a violation would be the two triples and all their separator qualifiers equal or non-existent.
@@
 
-To find violations of this constraint one would find a tripleEDB with the constrained property, but where no qualifierEDB for the statement and with the required property exists.
+To find violations of this constraint one would find a statementEDB with the constrained property, but where no qualifierEDB for the statement and with the required property exists.
 
@@
 With the exception of Property scope all constraints work on triples.
-The atom that will thus always be included is \(\text{tripleEDB}(X, Y, \text{propertyConstant}, Z)\).\\
+The atom that will thus always be included is \(\text{statementEDB}(X, Y, \text{propertyConstant}, Z)\).\\
 
@@
 A violation for a property scope constraint is
-\[\text{violation\_triple}(S, I, \text{propertyConstant}, V) \leftarrow \text{tripleEDB}(S, I, \text{propertyConstant}, V)\]
+\[\text{violation\_triple}(S, I, \text{propertyConstant}, V) \leftarrow \text{statementEDB}(S, I, \text{propertyConstant}, V)\]
 \[\text{violation\_qualifier}(S, \text{propertyConstant}, V) \leftarrow \text{qualifierEDB}(S, \text{propertyConstant}, V)\]
@@
 \text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow\\
-&\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge\\
-&\text{tripleEDB}(O, I, \text{conflictingPropertyConstant}, C)
+&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge\\
+&\text{statementEDB}(O, I, \text{conflictingPropertyConstant}, C)
 \end{split}
@@
 \text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow\\
-&\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge\\
-&\text{tripleEDB}(O, I, \text{conflictingPropertyConstant}, \text{conflictingValueConstant})
+&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge\\
+&\text{statementEDB}(O, I, \text{conflictingPropertyConstant}, \text{conflictingValueConstant})
 \end{split}
@@
 \text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \text{item}(I)
+&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \text{item}(I)
 \end{split}
@@
 \text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \text{property}(I)
+&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \text{property}(I)
 \end{split}
@@
 \text{violation\_triple}&(S, I, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
-&\text{tripleEDB}(S, I, \text{propertyConstant}, \text{forbiddenValueConstant})
+&\text{statementEDB}(S, I, \text{propertyConstant}, \text{forbiddenValueConstant})
 \end{split}
@@
 \text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \\
-&\text{tripleEDB}(O, X, \text{propertyConstant}, V) \wedge \\
+&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \\
+&\text{statementEDB}(O, X, \text{propertyConstant}, V) \wedge \\
 &\text{unequal}(S, O)
@@
 \begin{equation*}
-\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \text{unit}(V, U)
+\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \text{unit}(V, U)
 \end{equation*}
@@
 \text{violation\_triple}&(S, I \text{propertyConstant}, V) \leftarrow \\
-&\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \\
+&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \\
 &\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
@@
 \text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \\
+&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \\
 &\text{qualifierEDB}(S, Q, O) \wedge \text{unequal}(\{A\}, Q)
@@
 \text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
+&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
 \end{split}
@@
 \text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \\
+&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \\
 &\text{qualifierEDB}(S, \text{qualifierProperty}, O) \wedge \text{unequal}(\{A\}, O)
@@
 \begin{split}
-c_1 = &\text{tripleEDB}(Q, I, \text{propertyConstant}, X) \wedge \\
-&\text{tripleEDB}(S, I, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
+c_1 = &\text{statementEDB}(Q, I, \text{propertyConstant}, X) \wedge \\
+&\text{statementEDB}(S, I, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
 \end{split}
@@
 \begin{split}
-c_2 = &\text{tripleEDB}(Q, I, \text{propertyConstant}, X) \wedge \\
-&\text{tripleEDB}(S, I, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
+c_2 = &\text{statementEDB}(Q, I, \text{propertyConstant}, X) \wedge \\
+&\text{statementEDB}(S, I, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
 \end{split}
@@
 \text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \\
-&\text{tripleEDB}(O, I, P, X) \wedge \text{last}(O, I) \wedge \\
+&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \\
+&\text{statementEDB}(O, I, P, X) \wedge \text{last}(O, I) \wedge \\
 &\text{require}(O, \text{propertyConstant}, \text{requiredPropertyConstant})
@@
 \begin{split}
-c_1 = &\text{tripleEDB}(Q, R, \text{propertyConstant}, I) \wedge \\
-&\text{tripleEDB}(S, I, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
+c_1 = &\text{statementEDB}(Q, R, \text{propertyConstant}, I) \wedge \\
+&\text{statementEDB}(S, I, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
 \end{split}
@@
 \begin{split}
-c_2 = &\text{tripleEDB}(Q, R, \text{propertyConstant}, I) \wedge \\
-&\text{tripleEDB}(S, I, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
+c_2 = &\text{statementEDB}(Q, R, \text{propertyConstant}, I) \wedge \\
+&\text{statementEDB}(S, I, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
 \end{split}
@@
 \text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \\
-&\text{tripleEDB}(O, V, P, X) \wedge \text{last}(O, V) \wedge \\
+&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \\
+&\text{statementEDB}(O, V, P, X) \wedge \text{last}(O, V) \wedge \\
 &\text{require}(O, \text{propertyConstant}, \text{requiredPropertyConstant})
@@
 \begin{split}
-c_1 = &\text{tripleEDB}(Q, R, \text{propertyConstant}, I) \wedge \\
-&\text{tripleEDB}(S, I, P, V) \wedge \\
+c_1 = &\text{statementEDB}(Q, R, \text{propertyConstant}, I) \wedge \\
+&\text{statementEDB}(S, I, P, V) \wedge \\
 &\text{unequal}(\text{inversePropertyConstant}, P)
@@
 \begin{split}
-c_2 = &\text{tripleEDB}(Q, R, \text{propertyConstant}, I) \wedge \\
-&\text{tripleEDB}(S, I, \text{inversePropertyConstant}, V) \wedge \\
+c_2 = &\text{statementEDB}(Q, R, \text{propertyConstant}, I) \wedge \\
+&\text{statementEDB}(S, I, \text{inversePropertyConstant}, V) \wedge \\
 &\text{unequal}(R, V)
@@
 \begin{split}
-c = &\text{tripleEDB}(Q, I, \text{propertyConstant}, X) \wedge \\
-&\text{tripleEDB}(S, I, P, V) \wedge \text{unequal}(\text{propertyConstant}, P)
+c = &\text{statementEDB}(Q, I, \text{propertyConstant}, X) \wedge \\
+&\text{statementEDB}(S, I, P, V) \wedge \text{unequal}(\text{propertyConstant}, P)
 \end{split}
@@
 \text{require\_second}&(S, \text{propertyConstant}, \text{propertyConstant}) \leftarrow \\
-&\text{next}(R, S) \wedge \text{tripleEDB}(R, I, \text{propertyConstant}, C) \wedge \\
+&\text{next}(R, S) \wedge \text{statementEDB}(R, I, \text{propertyConstant}, C) \wedge \\
 &\text{next}(Q, R) \wedge \text{require}(Q, \text{propertyConstant}, \text{propertyConstant})  \wedge c
@@
 \text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \\
-&\text{tripleEDB}(O, I, \text{propertyConstant}, X) \wedge \\
+&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \\
+&\text{statementEDB}(O, I, \text{propertyConstant}, X) \wedge \\
 &\text{unequal}(S, O)
@@
 \begin{equation*}
-c = \text{tripleEDB}(S, I, \text{propertyConstant}, C) \wedge \text{qualifierEDB}(S, P, V) \wedge \text{unequal}(P, \text{separatorConstant})
+c = \text{statementEDB}(S, I, \text{propertyConstant}, C) \wedge \text{qualifierEDB}(S, P, V) \wedge \text{unequal}(P, \text{separatorConstant})
 \end{equation*}
@@
 \begin{split}
-\text{does\_not\_have}(S, R) \leftarrow &\text{tripleEDB}(S, I, \text{propertyConstant}, C) \wedge \\
+\text{does\_not\_have}(S, R) \leftarrow &\text{statementEDB}(S, I, \text{propertyConstant}, C) \wedge \\
 &\text{last\_qualifier}(S, P, V) \wedge \text{require}(S, P, V, \text{propertyConstant}, R)
@@
 \begin{split}
-\text{has\_same}(S, O, Q) \leftarrow &\text{tripleEDB}(S, I, \text{propertyConstant}, C) \wedge \\
-&\text{tripleEDB}(O, I, \text{propertyConstant}, X) \wedge \\
+\text{has\_same}(S, O, Q) \leftarrow &\text{statementEDB}(S, I, \text{propertyConstant}, C) \wedge \\
+&\text{statementEDB}(O, I, \text{propertyConstant}, X) \wedge \\
 &\text{qualifierEDB}(S, Q, V) \wedge \text{qualifierEDB}(O, Q, V)
@@
 \text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \\
-&\text{tripleEDB}(O, I, \text{propertyConstant}, C) \wedge \\
+&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \\
+&\text{statementEDB}(O, I, \text{propertyConstant}, C) \wedge \\
 &\text{unequal}(S, O) \wedge \\
@@
 \begin{split}
-c = &\text{tripleEDB}(S, I, \text{propertyConstant}, C) \wedge \\
+c = &\text{statementEDB}(S, I, \text{propertyConstant}, C) \wedge \\
 &\text{qualifierEDB}(S, P, V) \wedge \text{unequal}(P, \text{requiredPropertyConstant})
@@
 \text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \text{does\_not\_have}(S, \text{requiredPropertyConstant})
+&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \text{does\_not\_have}(S, \text{requiredPropertyConstant})
 \end{split}
@@
 \text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \\
+&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \\
 &\text{rank}(S, \text{preferredRankConstant}) \wedge \\
-&\text{tripleEDB}(O, I, \text{propertyConstant}, X) \wedge \\
+&\text{statementEDB}(O, I, \text{propertyConstant}, X) \wedge \\
 &\text{rank}(O, \text{preferredRankConstant}) \wedge \\
@@
 \text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{tripleEDB}(S, I, \text{propertyConstant}, V) \wedge \\
+&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \\
 &\text{rank}(S, \text{preferredRankConstant}) \wedge \\
-&\text{tripleEDB}(O, I, \text{propertyConstant}, W) \wedge \\
+&\text{statementEDB}(O, I, \text{propertyConstant}, W) \wedge \\
 &\text{rank}(O, \text{preferredRankConstant}) \wedge \\

### 2018-08-29 thesis/paper/paper.tex
Incorporated criticism

@@
 
-\setcounter{tocdepth}{1}
+\setcounter{tocdepth}{2}
 
@@
 In the context of Wikidata, any item or property, e.g.
-Earth (Q2) or head of government (P6), could be represented as a constant.
+\emph{Earth} (Q2) or \emph{head of government} (P6), could be represented as a constant.
 Relations in datalog are expressed using \emph{predicates} with a specified arity \(n\), written as predicate\([n]\).
-To stay close to Wikidata in our examples let us take the predicate statement\([3]\) as an example.
+To stay close to Wikidata in our examples let us take the predicate triple\([3]\) as an example.
 This would be used to express the \emph{item}-\emph{property}-\emph{value} relation with each position representing one part of the triple.\\
 Additionally, datalog also provides \emph{variables}, which will be capitalized like \(X\).
+The sets of predicates {\bf P}, constants {\bf C} and variables {\bf V} are mutually disjoint.
 Constants and variables are collectively known as \emph{terms} and are, together with predicates, used to form \emph{atoms}.
 An atom has the form predicate\((t_1, \ldots , t_n)\) with \(n\) as the arity of the predicate and \(t_1,\ldots,t_n\) terms.\\
-Atoms are the building blocks of \emph{rules}, which are used to derive information in datalog.
-A rule is a premise that implies a conclusion, and both are conjunctions of atoms.
+Atoms are the building blocks of \emph{rules}, which are used to derive information in datalog. A rule is a premise that implies a conclusion, and both are conjunctions of atoms.
 Let \(P_n\) denote  premise atoms and \(C_n\) the conclusion atoms.
-The variables used are divided into three mutually disjoint sets \(V_p\), \(V_c\), and \(V_r\) which are the variables occurring only in premise atoms, only in conclusion atoms, and occurring in both respectively.
+The variables used in one rule are divided into three mutually disjoint sets \(V_p\), \(V_c\), and \(V_r\) which are the variables occurring only in its premise atoms, only in its conclusion atoms, and occurring in both, respectively.
 The general from of an existential rule is (with \(k \in \mathbb{N}^+\), \(l \in \mathbb{N}\))
-\begin{equation*}
-\forall p \in V_p, r \in V_r : \exists c \in V_c: C_1 \ldots C_k \leftarrow P_1 \ldots P_l
-\end{equation*}
+\begin{equation}
+\forall V_p \forall V_r.(\exists V_c.C_1 \wedge \ldots \wedge C_k \leftarrow P_1 \wedge \ldots \wedge P_l)
+\end{equation}
 
@@
 If \(V_c = \emptyset\) the rule is called a \emph{datalog rule} and if \(l = 0\) it is called a \emph{fact}.
-Since \(\forall p \in V_p, r \in V_r\) can be taken as given for all rules, it will be omitted in this work.
+If a fact contains no variables but only constants it is a \emph{ground rule}.
+Since all \(V_p\) and \(V_r\) are  quantified, the preceding universal quantifiers \(\forall V_p \forall V_r\) will be omitted for the rules presented in this work.
 
@@
 \todo[inline]{Should I explain the chase? And in what detail...}
-\section{Representation}
+\section{Representation of Wikidata as Facts}
 To model the basic representation of triples, qualifiers and references three predicates are introduced.\\
@@
 \begin{table}[H]
-\caption{abbreviations}
+\caption{abbreviations}\label{tab_abbreviations}
 \begin{tabular}{ll}
@@
 constrained reference & reference with the constrained property as predicate \\
-constrained item & item with a constrained statement \\
-item statement & statement on the same item of a constrained statement \\
+constrained entity & entity with a constrained statement \\
+entity statement & statement on the same entity of a constrained statement \\
 value statement & statement on the value of a constrained statement
@@
 
-To improve the data quality, help editors avoid common mistakes and clarify the usage of properties, Wikidata allows constraints to be defined on properties.
-They are however not firm rules and exceptions can and should be made if necessary.
+To improve the data quality, to help editors avoid common mistakes and to clarify the usage of properties, Wikidata allows constraints to be defined on properties.
+They are, however, not firm rules and exceptions can and should be made if necessary.
 A property constraint says something about the way a property may or may not be used.
 This can mean simple restrictions on the statement a property is part of, e.g.
-only certain values are allowed or more complex requirements on connected parts of the data graph, e.g.
-the value of a statement with this property should have certain statements.
-Further down you can find table \ref{tab_property_constraints}, which list the existing property constraints.
-Afterwards the constraints will be explained in detail and either a basic approach for their modelling or reasoning why it is not possible or sensible to model it given.\\
-
-To apply a constraint to a property a statement <property> <P2302> <constraint> is added to the properties item page, where <property> is the property that should be constrained, <P2302> (property constraint) denotes that the property has a constraint and <constraint> is the specific constraint.
+that only certain values are allowed, or more complex requirements on connected parts of the data graph, e.g.
+that the value of a statement with this property should have certain statements.
+In Table \ref{tab_property_constraints} you can find the existing property constraints.
+Afterwards the constraints will be explained in detail and a basic approach for their modelling given.
+For constraint that cannot be modelled, reasoning for this assessment is included.
+
+To apply a constraint to a property, a statement <property> <P2302> <constraint> is added to the properties item page, where <property> is the property that should be constrained, <P2302> (property constraint) denotes that the property has a constraint and <constraint> is the specific constraint.
 Note that a constrained statement means a statement whose predicate is a property that has a specific constraint.
-A simple example for this would be distinct values (Q21502410) stating that values of the property must be unique across all constrained statements.
+A simple example for this would be \emph{distinct values} (Q21502410) stating that values of the property must be unique across all constrained statements.
 For this case no additional information regarding the constraint is necessary and violation could be found by surveying all constrained statements and comparing their values.\\
@@
 These qualifiers are either optional or mandatory and may be applied once or multiple times.
-For example, the none of constraint (Q52558054) states that values for a specific property may not be in a set of disallowed values.
-These are specified as qualifiers using item of property constraint (P2305) on the constraint statement and require at least one such statement.\\
+For example, the \emph{none of} constraint (Q52558054) states that values for the specific property may not be in a given set of disallowed values.
+This set is specified using qualifiers with property \emph{item of} (P2305), which list all the disallowed values.
+At least one such qualifier must be present for the statement.\\
 More complex dependencies are also possible.
-A constraint can depend on all constrained statements of a specific item (single value constraint [Q19474404]), on all statements of the value of a constrained statement (value requires statement constraint [Q21510864]) or in the most complex case on all statements using the subclassOf-predicate (type constraint [Q21503250]).
-
-Table \ref{tab_property_constraints} shows all currently used property constrains in Wikidata.
-It is sorted by difficulty, which is an estimate of the necessary pre-processing and rule complexity based on required features.
-These required features are conditions that cannot natively be expressed in datalog and would require some kind of pre-processing to be made available.
-Difficulty 0 does not need any pre-processing except for loading the data into datalog as ground rules.
-1 requires inequality, which can be easily solved with brute force but will be improved in chapter \ref{cha_rules}.
-2 needs negation in a form that can be solved without pre-processing everything and 3 in a form that would require all logic as part of pre-processing.
-Other features needed for category three are regular expressions, which could theoretically be translated but would require large amounts of pre-processing, value comparison, general negation and finally web requests to test if a link is valid.
+A constraint can depend on all constrained statements of a specific item (\emph{single value} constraint [Q19474404]), on all statements of the item specified as the value of a constrained statement (\emph{value requires statement} constraint [Q21510864]) or on all statements using the subclassOf-predicate (\emph{type} constraint [Q21503250]).
+
+Table \ref{tab_property_constraints} shows all constraints that are used as of September 2018.
+It is roughly sorted by how difficult it is to model the given constraint using logical rules, based on the required expressive features and pre-processing.
+These required features are conditions that cannot natively be expressed in datalog and would require some kind of pre-processing of the dataset.
+Four constraints do not need any pre-processing except for loading the data into datalog as ground rules.
+Six can be solved if constant inequality is defined.
+This can be easily solved with brute force but will be improved in section \ref{sec_inequality}.
+A further seven require negation of a specific fact which can be achieved with reasonable effort.
+The last ten \todo{maybe type can be done in some way} are not deemed translatable because they require a feature that cannot be reasonably emulated in datalog.
+The feature mostly needed is a general form of negation that would allow to test for the non-existence of a derived fact.
+Other features needed for these constraints are regular expressions, which could theoretically be translated but would require large amounts of pre-processing, value comparison, and finally web requests to test if a link is valid.
 
@@
 \caption{property constraints}\label{tab_property_constraints}
-\begin{tabularx}{\textheight}{rrllllX}
-Difficulty & \makecell{\# constrained\\ properties} & Identifier & Name & Required Features \\
-\hline
-0 & 1410 & Q53869507 & Scope &  \\
-0 & 781 & Q21502838 & Conflicts with &  \\
-0 & 27 & Q52004125 & Allowed entity types &  \\
-0 & 17 & Q52558054 & None of &  \\
-1 & 2843 & Q21502410 & Distinct value & Inequality \\
-1 & 343 & Q21514353 & Allowed units & Inequality \\
-1 & 303 & Q21510851 & Allowed qualifiers & Inequality \\
-1 & 116 & Q21510859 & One of & Inequality \\
-1 & 3 & Q52060874 & Single best value & Inequality \\
-1 & 1 & Q52712340 & One-of qualifier value & Inequality \\
-2 & 4265 & Q21503247 & Item requires statement & Negation (statement non-existence) \\
-2 & 244 & Q21510864 & Value requires statement & Negation (statement non-existence) \\
-2 & 119 & Q21510855 & Inverse & Negation (statement non-existence) \\
-2 & 33 & Q21510862 & Symmetric & Negation (statement non-existence) \\
-2 & 13 & Q21510857 & Multi-value & Negation, Count (up to two) \\
-2 & 2909 & Q19474404 & Single value & Inequality, partially negation (qualifier non-existence) \\
-2 & 203 & Q21510856 & Mandatory qualifier & Negation (qualifier non-existence) \\
-3 & 2613 & Q21503250 & Type & Negation (derived fact negation) \\
-3 & 707 & Q21510865 & Value type & Negation (derived fact negation) \\
-3 & 3021 & Q21502404 & Format & Regular expressions \\
-3 & 188 & Q21510860 & Range & Value comparison \\
-3 & 106 & Q52848401 & Integer & Regular expressions \\
-3 & 65 & Q51723761 & No bounds & Negation \\
-3 & 53 & Q21510852 & Commons link & Web request \\
-3 & 10 & Q54554025 & Link & Negation \\
-3 & 6 & Q21510854 & Difference within range & Value comparison \\
-3 & 1 & Q25796498 & Contemporary constraint & Negation \\
+\begin{tabularx}{\textheight}{rlll}
+\makecell{\# constrained\\ properties} & Identifier & Name & Required Features \\
+\hline
+1410 & Q53869507 & Scope &  \\
+781 & Q21502838 & Conflicts with &  \\
+27 & Q52004125 & Allowed entity types &  \\
+17 & Q52558054 & None of &  \\
+\hline
+2843 & Q21502410 & Distinct value & Inequality \\
+343 & Q21514353 & Allowed units & Inequality \\
+303 & Q21510851 & Allowed qualifiers & Inequality \\
+116 & Q21510859 & One of & Inequality \\
+3 & Q52060874 & Single best value & Inequality \\
+1 & Q52712340 & One-of qualifier value & Inequality \\
+\hline
+4265 & Q21503247 & Item requires statement & Negation (statement non-existence) \\
+244 & Q21510864 & Value requires statement & Negation (statement non-existence) \\
+119 & Q21510855 & Inverse & Negation (statement non-existence) \\
+33 & Q21510862 & Symmetric & Negation (statement non-existence) \\
+13 & Q21510857 & Multi-value & Negation, Count (up to two) \\
+2909 & Q19474404 & Single value & Inequality, Negation (qualifier non-existence) \\
+203 & Q21510856 & Mandatory qualifier & Negation (qualifier non-existence) \\
+\hline
+2613 & Q21503250 & Type & Negation (derived fact negation) \\
+707 & Q21510865 & Value type & Negation (derived fact negation) \\
+3021 & Q21502404 & Format & Regular expressions \\
+188 & Q21510860 & Range & Value comparison \\
+106 & Q52848401 & Integer & Regular expressions \\
+65 & Q51723761 & No bounds & Negation \\
+53 & Q21510852 & Commons link & Web request \\
+10 & Q54554025 & Link & Negation \\
+6 & Q21510854 & Difference within range & Value comparison \\
+1 & Q25796498 & Contemporary constraint & Negation \\
 \end{tabularx}
@@
 They will be presented roughly in descending order of the number of properties with this constraint, although constraints with similar function will be explained directly after one another.
-Also, all constraints that were considered untranslatable are listed at the and, each with a short explanation as to why they were placed in that group.\\
-
-Every constraint will start with a table showing basic properties of the constraint.
-This includes the requirement, relevant qualifiers, possible violations of the constraint and the data the constraint depends on.
-The requirement states which part of the data graph in relation to the constrained statement are restricted or need to exist.
+Also, all constraints that were considered untranslatable are listed at the and, each with a short explanation as to why they were placed in that group.
+
+Every section describing a constraint will start with a table showing basic properties of the constraint.
+These include the required data, relevant qualifiers, and possible violations of the constraint.
+The required data states which part of the data in relation to the constrained statement are restricted or need to exist.
 The qualifiers list their cardinality, meaning if they are required and if they may be used multiple times, and a short description.
 The violations list the conflicting (part of a) statement in relation to the constrained statement.
-The dependencies list their relation to the constrained statement.
-Table \ref{tab_example_constraint} shows the formatting as an example.
-
-\begin{table}[H]
-\caption{example constraint (Q31415926)}\label{tab_example_constraint}
-\begin{tabularx}{\textwidth}{ ll X}
-\hline
-Requirement & requirement \\
-\hline
-Qualifiers & \makecell{example qualifier (P1234) -- 1..* \\ description \\ another qualifier (P4321) -- 0..1 \\ different description} \\
-\hline
-Violation & \makecell{one way to violate the constraint \\ another violation possibility} \\
-\hline
-Dependency & statements, qualifiers or references in relation to the constrained statement\\
-\hline
-Rule features & features needed (missing if no features needed) \\
-\hline
-\end{tabularx}
-\end{table}
+The first constraint, \emph{property scope}, will be used as an example to explain the table in detail.
 
 \subsection{Property scope (Q53869507)}
-The property scope constraint specifies if a property may be used in a statement, qualifier or reference as predicate.
+\subsubsection{Constraint}
+The \emph{property scope} constraint specifies if a property may be used as predicate in a statement, qualifier or reference.
 \begin{table}[H]
@@
 \hline
-Requirement & constrained statement \\
-\hline
-Qualifiers & \makecell{property scope (P5314) -- 1..3 \\ must be as main value (Q54828448), as qualifier (Q54828449) or as reference (Q54828450) \\ specifies allowed use in statements, qualifiers or references} \\
-\hline
-Violation & constrained statement in a position that is not allowed \\
-\hline
-Dependency & constrained statement, qualifier or reference \\
-\hline
-\end{tabularx}
-\end{table}
-
-Since only three possibilities exist, this is equivalent to forbidding all other possibilities and the rules are simple mapping all of statementEDB, qualifierEDB and referenceEDB with the constrained property that are not allowed.
-
-\subsection{Conflicts-with (Q21502838)}
-The conflicts with constraint forbids either certain statements or statements with a specific property on constrained items.
+Required data & constrained statement, qualifier and reference \\
+\hline
+Qualifiers & \makecell{\emph{property scope} (P5314) -- 1..3 \\ must be \emph{as main value} (Q54828448), \emph{as qualifier} (Q54828449) \\ or \emph{as reference} (Q54828450) \\ specifies allowed use in statements, qualifiers or references} \\
+\hline
+Violation & constrained property in a position that is not allowed \\
+\hline
+\end{tabularx}
+\end{table}
+
+The required data specifies which data, in relation to the constrained statement, needs to be evaluated to conclude if the constraint was violated.
+The used abbreviations are listed in Table \ref{tab_abbreviations}.
+In this case this is simply each statement, qualifier or reference whose predicate is the constrained property.\\
+The \emph{property scope} constraint requires the scope, meaning statement, qualifier or reference, to be set via qualifier \emph{property scope} (P5314).
+The cardinality is 1..3 because it is mandatory and has only three possible values.\\
+It is violated if a statement, qualifier or reference has a predicate with the \emph{property scope} constraint that does not allow this scope.\\
+
+\subsubsection{Rules}
+The property scope constraint requires no context information to be violated; the triple, qualifier, or reference with the constrained property is enough.
+This means violations can be found using these rules:
+\begin{align*}
+\text{violation\_triple}(S, I, \text{propertyConstant}, V) &\leftarrow \text{statementEDB}(S, I, \text{propertyConstant}, V)\\
+\text{violation\_qualifier}(S, \text{propertyConstant}, V) &\leftarrow \text{qualifierEDB}(S, \text{propertyConstant}, V)\\
+\text{violation\_reference}(S, \text{propertyConstant}, V) &\leftarrow \text{referenceEDB}(S, H, \text{propertyConstant}, V)
+\end{align*}
+Which of these rules need to be used depends on the values of the \emph{property scope} qualifier.
+If a concrete constrained on a property does not have \emph{as main value} as a value of the qualifier, the first rule is used.
+The case for \emph{as qualifier} and \emph{as reference} is analogue with the second and third rule, respectively.
+
+\subsection{Conflicts with (Q21502838)}
+\subsubsection{Constraint}
+The \emph{conflicts with} constraint forbids statements on constrained entities whose predicate is a specified \emph{conflicting property}.
+If \emph{conflicting values} have been specified as well, the constraint is weakened and only violated by statements if in addition to the predicate being a \emph{conflicting property}, the object is one of the \emph{conflicting values}.
 \begin{table}[H]
@@
 \hline
-Requirement & item statement \\
-\hline
-Qualifiers & \makecell{conflicting property (P2306) -- 1 \\ property no item statement may have (with conflicting values if specified) \\
-conflicting values (P2305) -- 0..* \\ values that may not appear in item statements with the conflicting property} \\
-\hline
-Violation & \makecell{item statement with the conflicting property \\ item statement with the conflicting property and one of the conflicting values} \\
-\hline
-Dependency & all item statements \\
-\hline
-\end{tabularx}
-\end{table}
-The rule modelling can simply be done by choosing all statementEDB with the constrained property and then all statementEDB on the same item with the conflicting property, specifying the values as well if necessary.
+Required data & entity statements \\
+\hline
+Qualifiers & \makecell{\emph{conflicting property} (P2306) -- 1 \\ property that may not be predicate of entity statements \\
+\emph{conflicting values} (P2305) -- 0..* \\ values that may not be objects of entity statements with the \emph{conflicting property}} \\
+\hline
+Violation & \makecell{entity statement with the conflicting property \\ entity statement with the conflicting property and one of the conflicting values} \\
+\hline
+\end{tabularx}
+\end{table}
+
+\subsubsection{Rules}
+The conflicts-with constraint is violated if another statement on the same item either has a conflicting property or a conflicting property and a conflicting value depending on specified qualifiers.
+It depends on a conjunction (\(O\) denoting another statement ID, \(C\) another value) and a violation for a conflicts-with constraint is thus either
+\begin{equation*}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow\\
+&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge{}\\
+&\text{statementEDB}(O, I, \text{conflictingPropertyConstant}, C)
+\end{split}
+\end{equation*}
+for constraints with no conflicting values specified or one
+\begin{equation*}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow\\
+&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge{}\\
+&\text{statementEDB}(O, I, \text{conflictingPropertyConstant}, \text{conflictingValueConstant})
+\end{split}
+\end{equation*}
+for each conflicting value if specified.\\
 
@@
 \hline
-Requirement & constrained statement \\
+Required data & constrained statement \\
 \hline
@@
 \hline
-Requirement & constrained statement value \\
+Required data & constrained statement value \\
 \hline
@@
 \hline
-Requirement & constrained statement unit \\
+Required data & constrained statement unit \\
 \hline
@@
 \hline
-Requirement & qualifier on constrained statement \\
+Required data & qualifier on constrained statement \\
 \hline
@@
 \hline
-Requirement & constrained statement qualifier value \\
+Required data & constrained statement qualifier value \\
 \hline
@@
 \hline
-Requirement & item statement \\
+Required data & item statement \\
 \hline
@@
 \hline
-Requirement & item statement rank \\
+Required data & item statement rank \\
 \hline
@@
 \hline
-Requirement & chain of statements \\
+Required data & chain of statements \\
 \hline
@@
 
-The second step is the most complex: Depending on the constrained statement and the requirements, has a violation occurred?
+The second step is the most complex: Depending on the constrained statement and the Required data, has a violation occurred?
 For the easiest category of constraints this step does not entail any additional work since the existence of the dependent structure is itself a violation.
@@
 where \(S\) specifies the statement ID, \(I\) the item, \(P\) the property, and \(V\) the value. The term propertyConstant will be used for the constant of the constrained property.
-
-\subsection{Property scope (Q53869507)}
-The property scope constraint requires no context information to be violated; the triple, qualifier, or reference with the constrained property is enough.
-A violation for a property scope constraint is
-\[\text{violation\_triple}(S, I, \text{propertyConstant}, V) \leftarrow \text{statementEDB}(S, I, \text{propertyConstant}, V)\]
-\[\text{violation\_qualifier}(S, \text{propertyConstant}, V) \leftarrow \text{qualifierEDB}(S, \text{propertyConstant}, V)\]
-\[\text{violation\_reference}(S, \text{propertyConstant}, V) \leftarrow \text{referenceEDB}(S, H, \text{propertyConstant}, V)\]
-depending on if as main value, as qualifier, or as reference was not set in the qualifiers.\\
-
-There is no further information that can be materialized regarding this violations.
-
-\subsection{Conflicts-with (Q21502838)}
-The conflicts-with constraint is violated if another statement on the same item either has a conflicting property or a conflicting property and a conflicting value depending on specified qualifiers.
-It depends on a conjunction (\(O\) denoting another statement ID, \(C\) another value) and a violation for a conflicts-with constraint is thus either
-\begin{equation*}
-\begin{split}
-\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow\\
-&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge\\
-&\text{statementEDB}(O, I, \text{conflictingPropertyConstant}, C)
-\end{split}
-\end{equation*}
-for constraints with no conflicting values specified or one
-\begin{equation*}
-\begin{split}
-\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow\\
-&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge\\
-&\text{statementEDB}(O, I, \text{conflictingPropertyConstant}, \text{conflictingValueConstant})
-\end{split}
-\end{equation*}
-for each conflicting value if specified.\\
 

### 2018-08-29 thesis/paper/paper.tex
Corrected wedges at line end

@@
 \text{violation}&(X, Y, \text{PROPERTY}, Z) \leftarrow \\
-&\text{constrained}(\text{PROPERTY}, \text{constrainedConstant}) \wedge\\
+&\text{constrained}(\text{PROPERTY}, \text{constrainedConstant}) \wedge{} \\
 &\text{rules determining the violation}
@@
 \text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \\
-&\text{statementEDB}(O, X, \text{propertyConstant}, V) \wedge \\
+&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
+&\text{statementEDB}(O, X, \text{propertyConstant}, V) \wedge{} \\
 &\text{unequal}(S, O)
@@
 \text{violation\_triple}&(S, I \text{propertyConstant}, V) \leftarrow \\
-&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \\
+&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
 &\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
@@
 \text{violation\_qualifier}&(S, I \text{propertyConstant}, V) \leftarrow \\
-&\text{qualifierEDB}(S, \text{propertyConstant}, V) \wedge \\
+&\text{qualifierEDB}(S, \text{propertyConstant}, V) \wedge{} \\
 &\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
@@
 \text{violation\_reference}&(S, I \text{propertyConstant}, V) \leftarrow \\
-&\text{referenceEDB}(S, H, \text{propertyConstant}, V) \wedge \\
+&\text{referenceEDB}(S, H, \text{propertyConstant}, V) \wedge{} \\
 &\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
@@
 \text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \\
+&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
 &\text{qualifierEDB}(S, Q, O) \wedge \text{unequal}(\{A\}, Q)
@@
 \text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \\
+&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
 &\text{qualifierEDB}(S, \text{qualifierProperty}, O) \wedge \text{unequal}(\{A\}, O)
@@
 \begin{split}
-c_1 = &\text{statementEDB}(Q, I, \text{propertyConstant}, X) \wedge \\
+c_1 = &\text{statementEDB}(Q, I, \text{propertyConstant}, X) \wedge{} \\
 &\text{statementEDB}(S, I, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
@@
 \begin{split}
-c_2 = &\text{statementEDB}(Q, I, \text{propertyConstant}, X) \wedge \\
+c_2 = &\text{statementEDB}(Q, I, \text{propertyConstant}, X) \wedge{} \\
 &\text{statementEDB}(S, I, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
@@
 \text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \\
-&\text{statementEDB}(O, I, P, X) \wedge \text{last}(O, I) \wedge \\
+&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
+&\text{statementEDB}(O, I, P, X) \wedge \text{last}(O, I) \wedge{} \\
 &\text{require}(O, \text{propertyConstant}, \text{requiredPropertyConstant})
@@
 \begin{split}
-c_1 = &\text{statementEDB}(Q, R, \text{propertyConstant}, I) \wedge \\
+c_1 = &\text{statementEDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
 &\text{statementEDB}(S, I, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
@@
 \begin{split}
-c_2 = &\text{statementEDB}(Q, R, \text{propertyConstant}, I) \wedge \\
+c_2 = &\text{statementEDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
 &\text{statementEDB}(S, I, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
@@
 \text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \\
-&\text{statementEDB}(O, V, P, X) \wedge \text{last}(O, V) \wedge \\
+&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
+&\text{statementEDB}(O, V, P, X) \wedge \text{last}(O, V) \wedge{} \\
 &\text{require}(O, \text{propertyConstant}, \text{requiredPropertyConstant})
@@
 \begin{split}
-c_1 = &\text{statementEDB}(Q, R, \text{propertyConstant}, I) \wedge \\
-&\text{statementEDB}(S, I, P, V) \wedge \\
+c_1 = &\text{statementEDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
+&\text{statementEDB}(S, I, P, V) \wedge{} \\
 &\text{unequal}(\text{inversePropertyConstant}, P)
@@
 \begin{split}
-c_2 = &\text{statementEDB}(Q, R, \text{propertyConstant}, I) \wedge \\
-&\text{statementEDB}(S, I, \text{inversePropertyConstant}, V) \wedge \\
+c_2 = &\text{statementEDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
+&\text{statementEDB}(S, I, \text{inversePropertyConstant}, V) \wedge{} \\
 &\text{unequal}(R, V)
@@
 \begin{split}
-c = &\text{statementEDB}(Q, I, \text{propertyConstant}, X) \wedge \\
+c = &\text{statementEDB}(Q, I, \text{propertyConstant}, X) \wedge{} \\
 &\text{statementEDB}(S, I, P, V) \wedge \text{unequal}(\text{propertyConstant}, P)
@@
 \text{require\_second}&(S, \text{propertyConstant}, \text{propertyConstant}) \leftarrow \\
-&\text{next}(R, S) \wedge \text{statementEDB}(R, I, \text{propertyConstant}, C) \wedge \\
+&\text{next}(R, S) \wedge \text{statementEDB}(R, I, \text{propertyConstant}, C) \wedge{} \\
 &\text{next}(Q, R) \wedge \text{require}(Q, \text{propertyConstant}, \text{propertyConstant})  \wedge c
@@
 \begin{split}
-\text{require\_second}&(S, \text{propertyConstant}, \text{propertyConstant}) \leftarrow \text{next}(R, S) \wedge \\
+\text{require\_second}&(S, \text{propertyConstant}, \text{propertyConstant}) \leftarrow \text{next}(R, S) \wedge{} \\
 &\text{require\_second}(R, \text{propertyConstant}, \text{propertyConstant}) \wedge c
@@
 \text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \\
-&\text{statementEDB}(O, I, \text{propertyConstant}, X) \wedge \\
+&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
+&\text{statementEDB}(O, I, \text{propertyConstant}, X) \wedge{} \\
 &\text{unequal}(S, O)
@@
 \text{violation\_qualifier}&(S, \text{propertyConstant}, V) \leftarrow \\
-&\text{qualifierEDB}(S, \text{propertyConstant}, V) \wedge \\
-&\text{qualifierEDB}(S, \text{propertyConstant}, X) \wedge \\
+&\text{qualifierEDB}(S, \text{propertyConstant}, V) \wedge{} \\
+&\text{qualifierEDB}(S, \text{propertyConstant}, X) \wedge{} \\
 &\text{unequal}(V, X)
@@
 \text{violation\_reference}&(S, \text{propertyConstant}, V) \leftarrow \\
-&\text{referenceEDB}(S, H, \text{propertyConstant}, V) \wedge \\
-&\text{referenceEDB}(S, J, \text{propertyConstant}, X) \wedge \\
+&\text{referenceEDB}(S, H, \text{propertyConstant}, V) \wedge{} \\
+&\text{referenceEDB}(S, J, \text{propertyConstant}, X) \wedge{} \\
 &\text{unequal}(V, X)
@@
 \begin{split}
-\text{does\_not\_have}(S, R) \leftarrow &\text{statementEDB}(S, I, \text{propertyConstant}, C) \wedge \\
+\text{does\_not\_have}(S, R) \leftarrow &\text{statementEDB}(S, I, \text{propertyConstant}, C) \wedge{} \\
 &\text{last\_qualifier}(S, P, V) \wedge \text{require}(S, P, V, \text{propertyConstant}, R)
@@
 \begin{split}
-\text{has\_same}(S, O, Q) \leftarrow &\text{statementEDB}(S, I, \text{propertyConstant}, C) \wedge \\
-&\text{statementEDB}(O, I, \text{propertyConstant}, X) \wedge \\
+\text{has\_same}(S, O, Q) \leftarrow &\text{statementEDB}(S, I, \text{propertyConstant}, C) \wedge{} \\
+&\text{statementEDB}(O, I, \text{propertyConstant}, X) \wedge{} \\
 &\text{qualifierEDB}(S, Q, V) \wedge \text{qualifierEDB}(O, Q, V)
@@
 \text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \\
-&\text{statementEDB}(O, I, \text{propertyConstant}, C) \wedge \\
-&\text{unequal}(S, O) \wedge \\
-&\text{has\_same}(S, O, \{X\}) \wedge \\
-&\text{does\_not\_have}(S, \{Y\}) \wedge \\
+&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
+&\text{statementEDB}(O, I, \text{propertyConstant}, C) \wedge{} \\
+&\text{unequal}(S, O) \wedge{} \\
+&\text{has\_same}(S, O, \{X\}) \wedge{} \\
+&\text{does\_not\_have}(S, \{Y\}) \wedge{} \\
 &\text{does\_not\_have}(O, \{Y\})
@@
 \begin{split}
-c = &\text{statementEDB}(S, I, \text{propertyConstant}, C) \wedge \\
+c = &\text{statementEDB}(S, I, \text{propertyConstant}, C) \wedge{} \\
 &\text{qualifierEDB}(S, P, V) \wedge \text{unequal}(P, \text{requiredPropertyConstant})
@@
 \text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \\
-&\text{rank}(S, \text{preferredRankConstant}) \wedge \\
-&\text{statementEDB}(O, I, \text{propertyConstant}, X) \wedge \\
-&\text{rank}(O, \text{preferredRankConstant}) \wedge \\
+&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
+&\text{rank}(S, \text{preferredRankConstant}) \wedge{} \\
+&\text{statementEDB}(O, I, \text{propertyConstant}, X) \wedge{} \\
+&\text{rank}(O, \text{preferredRankConstant}) \wedge{} \\
 &\text{unequal}(S, O)
@@
 \text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \\
-&\text{rank}(S, \text{preferredRankConstant}) \wedge \\
-&\text{statementEDB}(O, I, \text{propertyConstant}, W) \wedge \\
-&\text{rank}(O, \text{preferredRankConstant}) \wedge \\
-&\text{unequal}(S, O) \wedge \\
-&\text{has\_same}(S, O, \{X\}) \wedge \\
-&\text{does\_not\_have}(S, \{Y\}) \wedge \\
+&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
+&\text{rank}(S, \text{preferredRankConstant}) \wedge{} \\
+&\text{statementEDB}(O, I, \text{propertyConstant}, W) \wedge{} \\
+&\text{rank}(O, \text{preferredRankConstant}) \wedge{} \\
+&\text{unequal}(S, O) \wedge{} \\
+&\text{has\_same}(S, O, \{X\}) \wedge{} \\
+&\text{does\_not\_have}(S, \{Y\}) \wedge{} \\
 &\text{does\_not\_have}(O, \{Y\})

### 2018-08-29 thesis/paper/paper.tex
Improved allowed entity types

@@
 They have numeric identifiers that are prefixed with Q and P and will be referred to as Q- and P-IDs, respectively.
+In the future there will be a third entity type, lexeme.
+This is already incorporated into some constraints, but as of September 2018 they are not deployed.
 A value is one of several data types shown in Table \ref{tab_datatypes}, although it is also possible to use \emph{none} or \emph{some} as values for statements that have no value (akin to negation) or have a currently unknown value.
@@
 \section{Representation of Wikidata as Facts}
-To model the basic representation of triples, qualifiers and references three predicates are introduced.\\
-statementEDB\([4]\) represents a single statement consisting of the statement ID, the item ID, the property ID and the value as either an item ID or a string-representation of the data value.\\
-qualifierEDB\([3]\) represent the qualifiers on a given statement ID with their respective property id and the value as usual.
-ReferenceEDB\([4]\) represents the statement ID, the reference hash, property id and the value.\\
+To model the representation of Wikidata the following predicates are introduced:\\
+statementEDB\([4]\) represents a single statement consisting of the statement ID, the entity ID, the property ID and the value as either an entity ID or a string-representation of the data value.\\
+qualifierEDB\([3]\) represent the qualifiers on a given statement ID with their respective property id and the value as usual. \\
+referenceEDB\([4]\) represents the statement ID, the reference hash, property id and the value.\\
+itemEDB\([1]\) and propertyEDB\([1]\) specify if an entity ID is an item or a property.\\
 
@@
 \subsection{Allowed entity types (Q52004125)}
-The allowed entity types constraint limits if a property may be used on an item, a property or a lexeme, although lexemes are not yet deployed.
+\subsubsection{Constraint}
+The \emph{allowed entity types} constraint limits if a property may be used in a statement whose subject is an item, a property or a lexeme, although as of September 2018 lexemes are not deployed.
 
@@
 \hline
-Required data & constrained statement \\
-\hline
-Qualifiers & \makecell{allowed entity type (P2305) -- 1..3 \\ must be wikibase item (Q29934200), wikibase property (Q29934218) \\ or lexeme (Q51885771); \\ specifies the type the property may be used on} \\
+Required data & constrained statements and entity types\\
+\hline
+Qualifiers & \makecell{allowed entity type (P2305) -- 1..3 \\ must be \emph{Wikibase item} (Q29934200), \emph{Wikibase property} (Q29934218) \\ or lexeme (Q51885771); \\ specifies the entity type the property may be used on} \\
 \hline
@@
 \hline
-Dependency & constrained statement and entity type of the subject \\
-\hline
-\end{tabularx}
-\end{table}
-
-Since there are only three options (and only two in use), violations can be found without inequality by finding the statementEDB with the constrained property whose subject is of the type that is not allowed, meaning either wikibase item or wikibase property.
+\end{tabularx}
+\end{table}
+
+Since there are only three options (and only two in use), violations can be found without inequality by finding the statementEDB with the constrained property whose subject is of the type that is not allowed, meaning either \emph{Wikibase item} or \emph{Wikibase property}.
+
+\subsubsection{Rules}
+The allowed entity types constraint is violated by a statement with the wrong entity type.
+Constraints regarding the lexeme entity type can be ignored since they are currently not used.
+The Violations depend on a simple conjunction:
+\begin{equation*}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \text{item}(I)
+\end{split}
+\end{equation*}
+\begin{equation*}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \text{property}(I)
+\end{split}
+\end{equation*}
+The first rule are created for all \emph{allowed entity type} constraints whose \emph{allowed entity type} does not contain \emph{Wikibase item}.
+The second rule works analogue for \emph{Wikibase property}.
 
@@
 where \(S\) specifies the statement ID, \(I\) the item, \(P\) the property, and \(V\) the value. The term propertyConstant will be used for the constant of the constrained property.
-
-\subsection{Allowed entity types (Q52004125)}
-The allowed entity types constraint is violated by a statement with the wrong entity type.
-This means that during extraction additional data has to be readied.
-To encode this two new predicates are introduced, item\([1]\) and property\([1]\).
-All items and properties with a constrained statement will then be sorted into one of these two categories.
-Constraints regarding the lexeme entity type can be ignored since they are currently not deployed.
-The violations the depend on a simple conjunction and follow directly as
-\begin{equation*}
-\begin{split}
-\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \text{item}(I)
-\end{split}
-\end{equation*}
-\begin{equation*}
-\begin{split}
-\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \text{property}(I)
-\end{split}
-\end{equation*}
 

### 2018-08-29 thesis/paper/paper.tex
Improved none of

@@
 \subsection{None of (Q52558054)}
+\subsubsection{Constraint}
 The none of constraint restricts the values that the property may refer to by listing a set of values as forbidden.
@@
 \hline
-Required data & constrained statement \\
-\hline
-Qualifiers & \makecell{forbidden value (P2305) -- 1..* \\ values that may not appear in a constrained property} \\
-\hline
-Violation & \makecell{constrained statement with a forbidden value} \\
-\hline
-Dependency & constrained statement \\
-\hline
-\end{tabularx}
-\end{table}
-
-Modelling this constraint is very straightforward, since each forbidden value forms a property-value pair with the constrained statement that may not exists.
-Thus a violations is a statementEDB with the constrained statement and a forbidden value.
+Required data & constrained statements, qualifiers and references \\
+\hline
+Qualifiers & \makecell{\emph{forbidden value} (P2305) -- 1..* \\ values that may not be object of a constraint statement, qualifier or reference} \\
+\hline
+Violation & \makecell{constrained statement, qualifier or reference with a forbidden value} \\
+\hline
+\end{tabularx}
+\end{table}
+
+\subsubsection{Rules}
+The none of constraint simply forbids certain values for properties, thus effectively prohibiting certain triples, qualifiers and references.
+The rules, depending on nothing but specific database entries, are
+\begin{equation*}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
+&\text{statementEDB}(S, I, \text{propertyConstant}, \text{forbiddenValueConstant})
+\end{split}
+\end{equation*}
+\begin{equation*}
+\begin{split}
+\text{violation\_qualifier}&(S, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
+&\text{qualifierEDB}(S, \text{propertyConstant}, \text{forbiddenValueConstant})
+\end{split}
+\end{equation*}
+\begin{equation*}
+\begin{split}
+\text{violation\_reference}&(S, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
+&\text{referenceEDB}(S, H, \text{propertyConstant}, \text{forbiddenValueConstant})
+\end{split}
+\end{equation*}
+Given a specified constraint, each of the three rules is created once for each \emph{forbidden value}.
 
@@
 where \(S\) specifies the statement ID, \(I\) the item, \(P\) the property, and \(V\) the value. The term propertyConstant will be used for the constant of the constrained property.
-
-\subsection{None of (Q52558054)}
-The none of constraint simply forbids certain values for properties, thus effectively prohibiting certain triples, qualifiers and references.
-The violations, depending on nothing but specific database entries, are
-\begin{equation*}
-\begin{split}
-\text{violation\_triple}&(S, I, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
-&\text{statementEDB}(S, I, \text{propertyConstant}, \text{forbiddenValueConstant})
-\end{split}
-\end{equation*}
-\begin{equation*}
-\begin{split}
-\text{violation\_qualifier}&(S, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
-&\text{qualifierEDB}(S, \text{propertyConstant}, \text{forbiddenValueConstant})
-\end{split}
-\end{equation*}
-\begin{equation*}
-\begin{split}
-\text{violation\_reference}&(S, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
-&\text{referenceEDB}(S, H, \text{propertyConstant}, \text{forbiddenValueConstant})
-\end{split}
-\end{equation*}
 

### 2018-08-29 thesis/paper/paper.tex
Improved rules intro

@@
 \chapter{Rules}\label{cha_rules}
-The previous chapter has show that most constraints cannot be modelled without at least some amount of preprocessing.
-Only Scope, Conflicts with, Allowed entity types and None of can be modelled directly.
-This is due to the fact that Conflicts with and None of exclude a limited and specified set of statements from validity and these cases can thus be directly takes as violations.
-Scope and Allowed entity types work very similar, since there are only three respectively two possible values and violations are simply the options that have not been allowed.
+The previous chapter has shown that most constraints cannot be modelled without at least some amount of preprocessing.
+Only \emph{scope}, \emph{conflicts with}, \emph{allowed entity types} and \emph{none of} can be modelled directly.
+This is due to the fact that \emph{conflicts with} and \emph{none of} exclude a limited and specified set of statements from validity and these cases can thus be directly taken as violations.
+\emph{scope} and \emph{allowed entity types} work very similar, since there are only three respectively two possible values and violations are simply the options that have not been allowed.
 The rules for these constraints are described in section \ref{sec_no_preprocessing}.\\
@@
 A problem that occurs multiple times is the need to determine inequality between constants.
-Distinct value, Allowed units, Allowed qualifiers, One of, Single best value and One-of qualifier value each require a rule to determine if a value, unit or property is not equal to another.
+\emph{distinct value}, \emph{allowed units}, \emph{allowed qualifiers}, \emph{one of}, \emph{single best value} and \emph{one-of qualifier value} each require a rule to determine if a value, unit or property is not equal to another.
 To this end a certain amount of preprocessing is necessary, most notably the extraction of all constant pairs whose inequality has to be determined for finding the violations.
-To avoid putting and undue amount of processing outside of datalog this extraction will not check more than that the constant is in the right position relative to the constrained statement, e.g.
-all values on statements constrained by Distinct value or all qualifiers on statements constrained by Allowed qualifiers.
+To avoid putting and undue amount of processing outside of datalog this extraction will not check more than that the constant is in the right position relative to the constrained statement, e.g., all values on statements constrained by \emph{distinct value} or all qualifiers on statements constrained by \emph{allowed qualifiers}.
 From this set inequality will be generated, different versions of which will be introduced in section \ref{sec_inequality} and evaluated with regards to their performance in section \ref{sec_inequality_evaluation}.\\
@@
 However, not all constraints require a general negation.
-Item requires statement, Value requires statement, Inverse and Symmetric are all violated if a statement does not exist on a specific item.
+\emph{item requires statement}, \emph{value requires statement}, \emph{inverse }and \emph{symmetric} are all violated if a statement does not exist on a specific item.
 This problem can be solved with limited amounts of preprocessing by leveraging a preprocessed statement order per item and 'iterating' through the statements until the required conditions are met.\\
-A similar approach can be taken to find items with exactly one statement matching certain criteria, required for 	Multi-value.
+A similar approach can be taken to find items with exactly one statement matching certain criteria, required for 	\emph{multi-value}.
 The idea is to ask the question if a second statement exists, then evaluating it as above.

### 2018-08-30 thesis/paper/paper.tex
Improved distinct values

@@
 \subsection{Distinct values (Q21502410)}
-The distinct values constraint forbids the existence of two constrained statements with the same value.
+\subsubsection{Constraint}
+The \emph{distinct values} constraint forbids the existence of two constrained statements with the same object.
 \begin{table}[H]
@@
 \hline
-Required data & constrained statement value \\
+Required data & constrained statements \\
 \hline
@@
 \hline
-Dependency & all constrained statements\\
-\hline
 Rule features & inequality \\
@@
 \end{table}
-To model the distinct values constraint it is necessary to find statements statementEDB with the constrained property and the same value whose statement id is unequal.
-In theory the constraint can also be applied to qualifiers and references by setting the respective value on a constraint scope (P4680) qualifier.
-However, there is currently no such case and will therefore not be regarded.
+The constraint can be expanded to cover qualifiers and references as well by setting a \emph{constraint scope} (P4680) qualifier with the respective value.
+However, as of September 2018 there is no constraint making use of this.
+
+\subsubsection{Rules}
+The \emph{distinct values} constraint requires two distinct constrained statements with the same value. 
+Two constrained statements with the same value can be found via a simple conjunction of two statementEDB atoms with the propertyConstant as predicate and a variable \(V\) as value.
+Since this would mark every constrained statement as a violations because it can fulfil both atoms at the same time, inequality of the statement IDs \(S\) and \(O\) must be tested as well.t
+\begin{equation*}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
+&\text{statementEDB}(O, X, \text{propertyConstant}, V) \wedge{} \\
+&\text{unequal}(S, O)
+\end{split}
+\end{equation*}
+The rules for qualifiers and reference are analogue.
 
@@
 \end{equation*}
-
-\todo{could use this to restrict the character inequality too but do we gain much?}
-
-\subsection{Distinct values (Q21502410)}
-The distinct values constraints requires two distinct constrained statements with the same value. 
-While finding constrained statements with the same value is straightforward, violations only occur if the statements are actually different, i.e. have an unequal statement id.
-\begin{equation*}
-\begin{split}
-\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
-&\text{statementEDB}(O, X, \text{propertyConstant}, V) \wedge{} \\
-&\text{unequal}(S, O)
-\end{split}
-\end{equation*}
-While it is possible to set this constraint on qualifiers or references, this is currently not done and is thus not evaluated here.
 

### 2018-08-30 src/main/java/impl/CC/ConstraintChecker.java
FIX: Inequality mode getter needed

@@ -138,7 +138,7 @@ public abstract class ConstraintChecker {
 			rulesToAdd.addAll(propertyConstraintChecker.rules());
 		}
 		logger.info("Created " + rulesToAdd.size() + " rules.");
-		if (InequalityHelper.mode.equals(InequalityHelper.Mode.DEMANDED)) {
+		if (InequalityHelper.getMode().equals(InequalityHelper.Mode.DEMANDED)) {
 			List<Rule> demandRules = addRequireInequality(rulesToAdd);
 			rulesToAdd.addAll(demandRules);
 			logger.info("Created " + demandRules.size() + " additional demand-rules.");

### 2018-08-30 src/main/java/utility/InequalityHelper.java

@@ -118,6 +118,10 @@ public class InequalityHelper {
 		}
 	}
 	
+	public static Mode getMode() {
+		return mode;
+	}
+	
 	public static InequalityHelper getInequalityHelper(ConstraintChecker checker) {
 		String folder = normalize(checker.getConstraint());
 		if (!helpers.containsKey(checker))

### 2018-08-31 thesis/paper/paper.tex
Improved allowed units

@@
 \chapter{Preliminaries}
-
 \section{Wikidata}
@@
 itemEDB\([1]\) and propertyEDB\([1]\) specify if an entity ID is an item or a property.\\
+unitEDB\([2]\) connects a value to its unit.\\
 
@@
 
-\subsection{Allowed units (Q21514353)}
-The allowed units constraint specifies that a constrained statement may only have a unit that is in a set of specific units.
+\subsection{Allowed units (Q21514353)}\label{subsec_3_allowed_units}
+\subsubsection{Constraint}
+The \emph{allowed units} constraint specifies that the value of a constrained statement may only have a unit that is in a set of specific units.
 
@@
 \hline
-Required data & constrained statement unit \\
-\hline
-Qualifiers & \makecell{allowed unit (P2305) -- 1..* \\ units that the statement may have (or none for dimensionless units)} \\
+Required data & constrained statements \\
+\hline
+Qualifiers & \makecell{\emph{allowed unit} (P2305) -- 1..* \\ units that are allowed on values of constrained statements (\emph{none} for dimensionless units)} \\
 \hline
@@
 \hline
-Dependency & constrained statement and its unit \\
-\hline
 Rule features & inequality \\
@@
 
-Modelling the constraint requires a predicate associating the value or the statement with its unit and inequality between every unit and all allowed units.
-Then a violation can be found by finding all statementEDB, qualifierEDB and referenceEDB with the constrained property and a unit unequal to all allowed units.
-Theoretically some of these can be excluded by using the constraint scope (P4680) qualifier on the constraint but there is no current case where this is done.
+By setting the \emph{constraint scope} (P4680) qualifier on the constraint it can be limited to only affect a subset of statement, qualifier, and reference.
+As of September 2018 this is not done on any constraint.
+
+\subsubsection{Rules}
+To handle this constraint inequality between all units need to be established.
+Then a violation can be identified by finding all statementEDB, qualifierEDB and referenceEDB with the constrained property and a unit unequal to all allowed units.
+The allowed units constraint needs to determine inequality between all allowed units and each unit on a constrained statement.
+To this end the units of a triple, qualifier or reference needs to be extracted during pre-processing and their inequality established.
+Quantities are rendered as string during pre-processing, specifying their value, bounds and unit.
+
+To determine if a violation occurred, the unit of the constrained statement needs to be found. This is done via the unit-predicate as below
+\begin{equation*}
+\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \text{unitEDB}(V, U)
+\end{equation*}
+with qualifiers and references treated analogue.
+
+The complete rules require multiple unequal-atoms for all allowed units.
+Given a set \(A\) of constants, in this case allowed units, \(\text{unequal}(\{A\}, t)\) with an arbitrary term \(t\) is short for \(\bigwedge_{a \in A} \text{unequal}(a, t)\)
+
+Violation are thus
+\begin{equation*}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
+&\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
+\end{split}
+\end{equation*}
+\begin{equation*}
+\begin{split}
+\text{violation\_qualifier}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+&\text{qualifierEDB}(S, \text{propertyConstant}, V) \wedge{} \\
+&\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
+\end{split}
+\end{equation*}
+\begin{equation*}
+\begin{split}
+\text{violation\_reference}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+&\text{referenceEDB}(S, H, \text{propertyConstant}, V) \wedge{} \\
+&\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
+\end{split}
+\end{equation*}
 
@@
 \text{unequal}(X, Y) \leftarrow \text{req\_inequality}(X, Y) \wedge \text{letter\_i}(X, A) \wedge \text{letter\_i}(Y, B) \wedge \text{unequal}(A, B)
-\end{equation*}
-
-\subsection{Allowed units (Q21514353)}\label{subsec_3_allowed_units}
-The allowed units constraint needs to determine inequality between all allowed units and each unit on a constrained statement.
-To this end the units of a triple, qualifier or reference needs to be extracted during pre-processing and their inequality established.
-Quantities are rendered as string during pre-processing, specifying their value, bounds and unit.
-The new predicate unit\([2]\) will associate a value with its unit.\\
-
-To determine if a violation occurred the unit to the constrained statement needs to be found. This is done via the unit-predicate as below
-\begin{equation*}
-\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \text{unit}(V, U)
-\end{equation*}
-with qualifiers and references working on the same lines.
-
-The full rules require multiple unequal-atoms for all allowed units.
-Given a set \(A\) of constants, in this case allowed units, \(\text{unequal}(\{A\}, t)\) is short for the conjunction of the set \(\{\text{unequal}(a, t) | \forall a \in A\}\), with \(t\) being an arbitrary term.
-
-Violation are thus
-\begin{equation*}
-\begin{split}
-\text{violation\_triple}&(S, I \text{propertyConstant}, V) \leftarrow \\
-&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
-&\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
-\end{split}
-\end{equation*}
-\begin{equation*}
-\begin{split}
-\text{violation\_qualifier}&(S, I \text{propertyConstant}, V) \leftarrow \\
-&\text{qualifierEDB}(S, \text{propertyConstant}, V) \wedge{} \\
-&\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
-\end{split}
-\end{equation*}
-\begin{equation*}
-\begin{split}
-\text{violation\_reference}&(S, I \text{propertyConstant}, V) \leftarrow \\
-&\text{referenceEDB}(S, H, \text{propertyConstant}, V) \wedge{} \\
-&\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
-\end{split}
 \end{equation*}

### 2018-08-31 thesis/paper/paper.tex
Improved allowed qualifiers

@@
 \subsection{Allowed qualifiers (Q21510851)}
-The allowed qualifiers constraint specifies that only a limited set of properties may be used as qualifiers on a constrained statement.
+\subsubsection{Constraint}
+The \emph{allowed qualifiers} constraint specifies that only a limited set of properties may be used as predicates of qualifiers on a constrained statement.
 
@@
 \hline
-Required data & qualifier on constrained statement \\
-\hline
-Qualifiers & \makecell{	allowed qualifier (P2306) -- 1..* \\ qualifiers that the statement may have} \\
+Required data & qualifiers on constrained statements \\
+\hline
+Qualifiers & \makecell{\emph{allowed qualifier} (P2306) -- 1..* \\ properties that may appear in predicates of qualifiers on constrained statements} \\
 \hline
@@
 \hline
-Dependency & constrained statement and its qualifiers \\
-\hline
 Rule features & inequality \\
@@
 
-To represent a violation of this constraint in datalog inequality between all specified allowed qualifiers and the properties of the qualifiers on constrained statements is required.
-A statementEDB with the constrained property and a qualifierEDB on that statement with a property different from all allowed qualifiers would violate the constraint.
+\subsubsection{Rules}
+To represent a violation of the \emph{allowed qualifiers} constraint in datalog inequality between all specified allowed qualifiers and the properties of qualifiers is required.
+The constraint is violated if a qualifier on a constrained statement exists that is unequal to all allowed qualifiers.
+The notation for set inequality as introduced in section \ref{subsec_3_allowed_units} will be used.
+With \(A\) the set of allowed qualifiers the rule is:
+\begin{equation*}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
+&\text{qualifierEDB}(S, Q, O) \wedge \text{unequal}(\{A\}, Q)
+\end{split}
+\end{equation*}
 
@@
 \text{unequal}(X, Y) \leftarrow \text{req\_inequality}(X, Y) \wedge \text{letter\_i}(X, A) \wedge \text{letter\_i}(Y, B) \wedge \text{unequal}(A, B)
-\end{equation*}
-
-\subsection{Allowed qualifiers (Q21510851)}
-The allowed qualifiers constraint is violated if a qualifier on a constrained statement exists that is different from all allowed qualifiers.
-During pre-processing all qualifiers on constrained statements will be extracted and the inequality of their properties and the allowed qualifiers established.
-We will use the notation for set inequality as introduced in section \ref{subsec_3_allowed_units}.
-With \(A\) the set of allowed qualifiers the rule is:
-\begin{equation*}
-\begin{split}
-\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
-&\text{qualifierEDB}(S, Q, O) \wedge \text{unequal}(\{A\}, Q)
-\end{split}
 \end{equation*}

### 2018-08-31 thesis/paper/paper.tex
Improved one-of

@@
 \subsection{One-of (Q21510859)}
-The one-of constraint limits the values a property may use to a set of given values.
+\subsubsection{Constraint}
+The \emph{one-of} constraint limits the values of a constraint statement, qualifier or reference to a set of given values.
 
@@
 \hline
-Required data & constrained statement \\
-\hline
-Qualifiers & \makecell{allowed values (P2305) -- 1..* \\ values that may be used in the constrained statement} \\
+Required data & constrained statements, qualifiers and references \\
+\hline
+Qualifiers & \makecell{\emph{allowed values} (P2305) -- 1..* \\ values that may be used as objects of a constrained statement} \\
 \hline
@@
 \hline
-Dependency & constrained statement \\
-\hline
 Rule features & inequality \\
@@
 
-Modelling the one-of constraint requires a rule finding a statementEDB with the constrained property whose value is unequal to all allowed values, meaning the inequality between each value of constrained statements and all allowed values must be given.
+\subsubsection{Rules}
+The one-of constraint fits in the pattern established by the previous two constraints and disallows certain property-value combinations in statements, qualifiers and references.
+Inequality is established between all \emph{allowed values} and the values of statements, qualifiers and references, respectively.
+With \(A\) the set of allowed values the rules for triple, qualifier and reference are:
+\begin{align*}
+\text{violation\_triple}(S, I, \text{propertyConstant}, V) &\leftarrow \\
+\text{statementEDB}(S, I, &\text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V) \\
+\text{violation\_qualifier}(S, \text{propertyConstant}, V) &\leftarrow \\
+\text{qualifierEDB}(S, &\text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V) \\
+\text{violation\_reference}(S, \text{propertyConstant}, V) &\leftarrow \\
+\text{referenceEDB}(S, H, &\text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
+\end{align*}
 
@@
 \text{unequal}(X, Y) \leftarrow \text{req\_inequality}(X, Y) \wedge \text{letter\_i}(X, A) \wedge \text{letter\_i}(Y, B) \wedge \text{unequal}(A, B)
-\end{equation*}
-
-\subsection{One-of (Q21510859)}
-The one-of constraint fits in the pattern established by the previous two constraints and disallows certain property-value combinations.
-Inequality is pre-processed between all values of constrained statements.
-With \(A\) the set of allowed values the rules for triple, qualifier and reference are respectively:
-\begin{equation*}
-\begin{split}
-\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
-\end{split}
-\end{equation*}
-\begin{equation*}
-\begin{split}
-\text{violation\_qualifier}&(S, \text{propertyConstant}, V) \leftarrow \\
-&\text{qualifierEDB}(S, \text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
-\end{split}
-\end{equation*}
-\begin{equation*}
-\begin{split}
-\text{violation\_reference}&(S, \text{propertyConstant}, V) \leftarrow \\
-&\text{referenceEDB}(S, H, \text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
-\end{split}
 \end{equation*}

### 2018-08-31 thesis/paper/paper.tex
Improved one-of qualifier value

@@
 \subsection{One-of qualifier value (Q52712340)}
-The one-of qualifier value constraint restricts which values may be used on a qualifier with a given property on a constrained statement.
+\subsubsection{Constraint}
+The \emph{one-of qualifier value} constraint restricts which \emph{qualifier values} may be used on a qualifier with a given \emph{qualifier property} on a constrained statement.
 
@@
 \hline
-Required data & constrained statement qualifier value \\
-\hline
-Qualifiers: & \makecell{qualifier property (P2306) -- 1 \\ property of the restricted qualifier\\
-qualifier values (P2305) -- 1..* \\ values the qualifiers may take}\\
-\hline
-Violation & \makecell{constrained statement with a qualifier with the qualifier property and a value unequal to all qualifier values} \\
-\hline
-Dependency & constrained item and qualifiers with the qualifier property \\
+Required data & constrained statements and their qualifiers \\
+\hline
+Qualifiers: & \makecell{\emph{qualifier property} (P2306) -- 1 \\ property that, if used as predicate, constrains a qualifier \\
+\emph{qualifier values} (P2305) -- 1..* \\ values that are allowed for qualifiers on constrained statements \\ with the \emph{qualifier property} as predicate}\\
+\hline
+Violation & \makecell{constrained statement with a qualifier with the qualifier property \\ and a value unequal to all qualifier values} \\
 \hline
@@
 
-The constraint can be modelled if the inequality between a value of a qualifier property with all qualifier values is given.
-The violations could then be found by taking a statementEDB with the constrained property and a qualifier with the qualifier property on the former statement whose value is unequal to all qualifier values.
+\subsubsection{Rules}
+The one-of qualifier value constraint is violated if a constrained statement has a qualifier with a value unequal to all allowed values.
+Inequality needs to be established between all values of qualifiers and all allowed values.
+With the set of allowed Values \(A\) the rules are
+\begin{equation*}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
+&\text{qualifierEDB}(S, \text{qualifierProperty}, O) \wedge \text{unequal}(\{A\}, O)
+\end{split}
+\end{equation*}
 
@@
 \text{unequal}(X, Y) \leftarrow \text{req\_inequality}(X, Y) \wedge \text{letter\_i}(X, A) \wedge \text{letter\_i}(Y, B) \wedge \text{unequal}(A, B)
-\end{equation*}
-
-\subsection{One-of qualifier value (Q52712340)}
-The one-of qualifier value constraint is violated if a constrained statement has a qualifier with a value unequal to all allowed values.
-Inequality needs to be pre-processed between all values of qualifiers on constrained statements.
-With the set of allowed Values \(A\) the rules are
-\begin{equation*}
-\begin{split}
-\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
-&\text{qualifierEDB}(S, \text{qualifierProperty}, O) \wedge \text{unequal}(\{A\}, O)
-\end{split}
 \end{equation*}

### 2018-08-31 thesis/paper/paper.tex
Restructured paper

@@
 The species \emph{blue whale} (Q42196) has the \emph{parent taxon} (P171) \emph{balaenoptera} (Q133320), which in turn has the parent taxon \emph{rorqual} (Q232829).
-The predicates used are tripleEDB\([3]\) as the input and isInTaxonIDB\([2]\), which will be derived.
+The predicates used are triple\_EDB\([3]\) as the input and isInTaxon\_IDB\([2]\), which will be derived.
 The basic facts are:
-\[\text{tripleEDB}(\text{blue whale}, \text{parent taxon}, \text{balaenoptera})\]
-\[\text{tripleEDB}(\text{balaenoptera}, \text{parent taxon}, \text{rorqual})\]
+\[\text{triple\_EDB}(\text{blue whale}, \text{parent taxon}, \text{balaenoptera})\]
+\[\text{triple\_EDB}(\text{balaenoptera}, \text{parent taxon}, \text{rorqual})\]
 From this we can draw conclusions. The first step is to represent the already existing facts as IDB:
-\[\text{isInTaxonIDB}(X, Y) \leftarrow \text{tripleEDB}(X, \text{parent taxon}, Y)\]
+\[\text{isInTaxon\_IDB}(X, Y) \leftarrow \text{triple\_EDB}(X, \text{parent taxon}, Y)\]
 Finally the transitive closure of this relation can be derived with
-\[\text{isInTaxonIDB}(X, Y) \leftarrow \text{isInTaxonIDB}(X, Z) \wedge \text{isInTaxonIDB}(Z, Y)\]
-
-\todo[inline]{Should I explain the chase? And in what detail...}
-\section{Representation of Wikidata as Facts}
-To model the representation of Wikidata the following predicates are introduced:\\
-statementEDB\([4]\) represents a single statement consisting of the statement ID, the entity ID, the property ID and the value as either an entity ID or a string-representation of the data value.\\
-qualifierEDB\([3]\) represent the qualifiers on a given statement ID with their respective property id and the value as usual. \\
-referenceEDB\([4]\) represents the statement ID, the reference hash, property id and the value.\\
-itemEDB\([1]\) and propertyEDB\([1]\) specify if an entity ID is an item or a property.\\
-unitEDB\([2]\) connects a value to its unit.\\
-
-Additional predicates modelling different pre-processing steps are explained in chapter \ref{cha_rules}.
-\section{Terminology}
-\begin{table}[H]
-\caption{abbreviations}\label{tab_abbreviations}
-\begin{tabular}{ll}
-Shorthand & Description \\
-\hline
-constrained statement & statement with the constrained property as predicate \\
-constrained qualifier & qualifier with the constrained property as predicate \\
-constrained reference & reference with the constrained property as predicate \\
-constrained entity & entity with a constrained statement \\
-entity statement & statement on the same entity of a constrained statement \\
-value statement & statement on the value of a constrained statement
-\end{tabular}
-\end{table}
+\[\text{isInTaxon\_IDB}(X, Y) \leftarrow \text{isInTaxon\_IDB}(X, Z) \wedge \text{isInTaxon\_IDB}(Z, Y)\]
+
 \chapter{Property Constraints}
@@
 
-\section{Wikidata constraints explained}
-This section will explain all interesting, meaning translatable or partially translatable, currently used property constraints on Wikidata.
-They will be presented roughly in descending order of the number of properties with this constraint, although constraints with similar function will be explained directly after one another.
-Also, all constraints that were considered untranslatable are listed at the and, each with a short explanation as to why they were placed in that group.
-
-Every section describing a constraint will start with a table showing basic properties of the constraint.
-These include the required data, relevant qualifiers, and possible violations of the constraint.
-The required data states which part of the data in relation to the constrained statement are restricted or need to exist.
-The qualifiers list their cardinality, meaning if they are required and if they may be used multiple times, and a short description.
-The violations list the conflicting (part of a) statement in relation to the constrained statement.
-The first constraint, \emph{property scope}, will be used as an example to explain the table in detail.
-
-\subsection{Property scope (Q53869507)}
-\subsubsection{Constraint}
-The \emph{property scope} constraint specifies if a property may be used as predicate in a statement, qualifier or reference.
-\begin{table}[H]
-\caption{property scope constraint (Q53869507)}
-\begin{tabularx}{\textwidth}{ ll X}
-\hline
-Required data & constrained statement, qualifier and reference \\
-\hline
-Qualifiers & \makecell{\emph{property scope} (P5314) -- 1..3 \\ must be \emph{as main value} (Q54828448), \emph{as qualifier} (Q54828449) \\ or \emph{as reference} (Q54828450) \\ specifies allowed use in statements, qualifiers or references} \\
-\hline
-Violation & constrained property in a position that is not allowed \\
-\hline
-\end{tabularx}
-\end{table}
-
-The required data specifies which data, in relation to the constrained statement, needs to be evaluated to conclude if the constraint was violated.
-The used abbreviations are listed in Table \ref{tab_abbreviations}.
-In this case this is simply each statement, qualifier or reference whose predicate is the constrained property.\\
-The \emph{property scope} constraint requires the scope, meaning statement, qualifier or reference, to be set via qualifier \emph{property scope} (P5314).
-The cardinality is 1..3 because it is mandatory and has only three possible values.\\
-It is violated if a statement, qualifier or reference has a predicate with the \emph{property scope} constraint that does not allow this scope.\\
-
-\subsubsection{Rules}
-The property scope constraint requires no context information to be violated; the triple, qualifier, or reference with the constrained property is enough.
-This means violations can be found using these rules:
-\begin{align*}
-\text{violation\_triple}(S, I, \text{propertyConstant}, V) &\leftarrow \text{statementEDB}(S, I, \text{propertyConstant}, V)\\
-\text{violation\_qualifier}(S, \text{propertyConstant}, V) &\leftarrow \text{qualifierEDB}(S, \text{propertyConstant}, V)\\
-\text{violation\_reference}(S, \text{propertyConstant}, V) &\leftarrow \text{referenceEDB}(S, H, \text{propertyConstant}, V)
-\end{align*}
-Which of these rules need to be used depends on the values of the \emph{property scope} qualifier.
-If a concrete constrained on a property does not have \emph{as main value} as a value of the qualifier, the first rule is used.
-The case for \emph{as qualifier} and \emph{as reference} is analogue with the second and third rule, respectively.
-
-\subsection{Conflicts with (Q21502838)}
-\subsubsection{Constraint}
-The \emph{conflicts with} constraint forbids statements on constrained entities whose predicate is a specified \emph{conflicting property}.
-If \emph{conflicting values} have been specified as well, the constraint is weakened and only violated by statements if in addition to the predicate being a \emph{conflicting property}, the object is one of the \emph{conflicting values}.
-\begin{table}[H]
-\caption{Conflicts-with constraint (Q21502838)}
-\begin{tabularx}{\textwidth}{ ll X}
-\hline
-Required data & entity statements \\
-\hline
-Qualifiers & \makecell{\emph{conflicting property} (P2306) -- 1 \\ property that may not be predicate of entity statements \\
-\emph{conflicting values} (P2305) -- 0..* \\ values that may not be objects of entity statements with the \emph{conflicting property}} \\
-\hline
-Violation & \makecell{entity statement with the conflicting property \\ entity statement with the conflicting property and one of the conflicting values} \\
-\hline
-\end{tabularx}
-\end{table}
-
-\subsubsection{Rules}
-The conflicts-with constraint is violated if another statement on the same item either has a conflicting property or a conflicting property and a conflicting value depending on specified qualifiers.
-It depends on a conjunction (\(O\) denoting another statement ID, \(C\) another value) and a violation for a conflicts-with constraint is thus either
-\begin{equation*}
-\begin{split}
-\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow\\
-&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge{}\\
-&\text{statementEDB}(O, I, \text{conflictingPropertyConstant}, C)
-\end{split}
-\end{equation*}
-for constraints with no conflicting values specified or one
-\begin{equation*}
-\begin{split}
-\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow\\
-&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge{}\\
-&\text{statementEDB}(O, I, \text{conflictingPropertyConstant}, \text{conflictingValueConstant})
-\end{split}
-\end{equation*}
-for each conflicting value if specified.\\
-
-\subsection{Allowed entity types (Q52004125)}
-\subsubsection{Constraint}
-The \emph{allowed entity types} constraint limits if a property may be used in a statement whose subject is an item, a property or a lexeme, although as of September 2018 lexemes are not deployed.
-
-\begin{table}[H]
-\caption{Allowed entity types constraint (Q52004125)}
-\begin{tabularx}{\textwidth}{ ll X}
-\hline
-Required data & constrained statements and entity types\\
-\hline
-Qualifiers & \makecell{allowed entity type (P2305) -- 1..3 \\ must be \emph{Wikibase item} (Q29934200), \emph{Wikibase property} (Q29934218) \\ or lexeme (Q51885771); \\ specifies the entity type the property may be used on} \\
-\hline
-Violation & \makecell{constrained statement on an entity of a type unequal to all allowed types} \\
-\hline
-\end{tabularx}
-\end{table}
-
-Since there are only three options (and only two in use), violations can be found without inequality by finding the statementEDB with the constrained property whose subject is of the type that is not allowed, meaning either \emph{Wikibase item} or \emph{Wikibase property}.
-
-\subsubsection{Rules}
-The allowed entity types constraint is violated by a statement with the wrong entity type.
-Constraints regarding the lexeme entity type can be ignored since they are currently not used.
-The Violations depend on a simple conjunction:
-\begin{equation*}
-\begin{split}
-\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \text{item}(I)
-\end{split}
-\end{equation*}
-\begin{equation*}
-\begin{split}
-\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \text{property}(I)
-\end{split}
-\end{equation*}
-The first rule are created for all \emph{allowed entity type} constraints whose \emph{allowed entity type} does not contain \emph{Wikibase item}.
-The second rule works analogue for \emph{Wikibase property}.
-
-\subsection{None of (Q52558054)}
-\subsubsection{Constraint}
-The none of constraint restricts the values that the property may refer to by listing a set of values as forbidden.
-
-\begin{table}[H]
-\caption{None of constraint (Q52558054)}
-\begin{tabularx}{\textwidth}{ ll X}
-\hline
-Required data & constrained statements, qualifiers and references \\
-\hline
-Qualifiers & \makecell{\emph{forbidden value} (P2305) -- 1..* \\ values that may not be object of a constraint statement, qualifier or reference} \\
-\hline
-Violation & \makecell{constrained statement, qualifier or reference with a forbidden value} \\
-\hline
-\end{tabularx}
-\end{table}
-
-\subsubsection{Rules}
-The none of constraint simply forbids certain values for properties, thus effectively prohibiting certain triples, qualifiers and references.
-The rules, depending on nothing but specific database entries, are
-\begin{equation*}
-\begin{split}
-\text{violation\_triple}&(S, I, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
-&\text{statementEDB}(S, I, \text{propertyConstant}, \text{forbiddenValueConstant})
-\end{split}
-\end{equation*}
-\begin{equation*}
-\begin{split}
-\text{violation\_qualifier}&(S, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
-&\text{qualifierEDB}(S, \text{propertyConstant}, \text{forbiddenValueConstant})
-\end{split}
-\end{equation*}
-\begin{equation*}
-\begin{split}
-\text{violation\_reference}&(S, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
-&\text{referenceEDB}(S, H, \text{propertyConstant}, \text{forbiddenValueConstant})
-\end{split}
-\end{equation*}
-Given a specified constraint, each of the three rules is created once for each \emph{forbidden value}.
-
-\subsection{Distinct values (Q21502410)}
-\subsubsection{Constraint}
-The \emph{distinct values} constraint forbids the existence of two constrained statements with the same object.
-\begin{table}[H]
-\caption{distinct values constraint (Q21502410)}
-\begin{tabularx}{\textwidth}{ ll X}
-\hline
-Required data & constrained statements \\
-\hline
-Violations & two constrained statements with same value \\
-\hline
-Rule features & inequality \\
-\hline
-\end{tabularx}
-\end{table}
-The constraint can be expanded to cover qualifiers and references as well by setting a \emph{constraint scope} (P4680) qualifier with the respective value.
-However, as of September 2018 there is no constraint making use of this.
-
-\subsubsection{Rules}
-The \emph{distinct values} constraint requires two distinct constrained statements with the same value. 
-Two constrained statements with the same value can be found via a simple conjunction of two statementEDB atoms with the propertyConstant as predicate and a variable \(V\) as value.
-Since this would mark every constrained statement as a violations because it can fulfil both atoms at the same time, inequality of the statement IDs \(S\) and \(O\) must be tested as well.t
-\begin{equation*}
-\begin{split}
-\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
-&\text{statementEDB}(O, X, \text{propertyConstant}, V) \wedge{} \\
-&\text{unequal}(S, O)
-\end{split}
-\end{equation*}
-The rules for qualifiers and reference are analogue.
-
-\subsection{Allowed units (Q21514353)}\label{subsec_3_allowed_units}
-\subsubsection{Constraint}
-The \emph{allowed units} constraint specifies that the value of a constrained statement may only have a unit that is in a set of specific units.
-
-\begin{table}[H]
-\caption{Allowed units constraint (Q21514353)}
-\begin{tabularx}{\textwidth}{ ll X}
-\hline
-Required data & constrained statements \\
-\hline
-Qualifiers & \makecell{\emph{allowed unit} (P2305) -- 1..* \\ units that are allowed on values of constrained statements (\emph{none} for dimensionless units)} \\
-\hline
-Violation & \makecell{constrained statement with a unit unequal to all allowed units} \\
-\hline
-Rule features & inequality \\
-\hline
-\end{tabularx}
-\end{table}
-
-By setting the \emph{constraint scope} (P4680) qualifier on the constraint it can be limited to only affect a subset of statement, qualifier, and reference.
-As of September 2018 this is not done on any constraint.
-
-\subsubsection{Rules}
-To handle this constraint inequality between all units need to be established.
-Then a violation can be identified by finding all statementEDB, qualifierEDB and referenceEDB with the constrained property and a unit unequal to all allowed units.
-The allowed units constraint needs to determine inequality between all allowed units and each unit on a constrained statement.
-To this end the units of a triple, qualifier or reference needs to be extracted during pre-processing and their inequality established.
-Quantities are rendered as string during pre-processing, specifying their value, bounds and unit.
-
-To determine if a violation occurred, the unit of the constrained statement needs to be found. This is done via the unit-predicate as below
-\begin{equation*}
-\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \text{unitEDB}(V, U)
-\end{equation*}
-with qualifiers and references treated analogue.
-
-The complete rules require multiple unequal-atoms for all allowed units.
-Given a set \(A\) of constants, in this case allowed units, \(\text{unequal}(\{A\}, t)\) with an arbitrary term \(t\) is short for \(\bigwedge_{a \in A} \text{unequal}(a, t)\)
-
-Violation are thus
-\begin{equation*}
-\begin{split}
-\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
-&\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
-\end{split}
-\end{equation*}
-\begin{equation*}
-\begin{split}
-\text{violation\_qualifier}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{qualifierEDB}(S, \text{propertyConstant}, V) \wedge{} \\
-&\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
-\end{split}
-\end{equation*}
-\begin{equation*}
-\begin{split}
-\text{violation\_reference}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{referenceEDB}(S, H, \text{propertyConstant}, V) \wedge{} \\
-&\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
-\end{split}
-\end{equation*}
-
-\subsection{Allowed qualifiers (Q21510851)}
-\subsubsection{Constraint}
-The \emph{allowed qualifiers} constraint specifies that only a limited set of properties may be used as predicates of qualifiers on a constrained statement.
-
-\begin{table}[H]
-\caption{Allowed qualifiers constraint (Q21510851)}
-\begin{tabularx}{\textwidth}{ ll X}
-\hline
-Required data & qualifiers on constrained statements \\
-\hline
-Qualifiers & \makecell{\emph{allowed qualifier} (P2306) -- 1..* \\ properties that may appear in predicates of qualifiers on constrained statements} \\
-\hline
-Violation & \makecell{constrained statement with a qualifier unequal to all allowed qualifiers} \\
-\hline
-Rule features & inequality \\
-\hline
-\end{tabularx}
-\end{table}
-
-\subsubsection{Rules}
-To represent a violation of the \emph{allowed qualifiers} constraint in datalog inequality between all specified allowed qualifiers and the properties of qualifiers is required.
-The constraint is violated if a qualifier on a constrained statement exists that is unequal to all allowed qualifiers.
-The notation for set inequality as introduced in section \ref{subsec_3_allowed_units} will be used.
-With \(A\) the set of allowed qualifiers the rule is:
-\begin{equation*}
-\begin{split}
-\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
-&\text{qualifierEDB}(S, Q, O) \wedge \text{unequal}(\{A\}, Q)
-\end{split}
-\end{equation*}
-
-\subsection{One-of (Q21510859)}
-\subsubsection{Constraint}
-The \emph{one-of} constraint limits the values of a constraint statement, qualifier or reference to a set of given values.
-
-\begin{table}[H]
-\caption{One-of constraint (Q21510859)}
-\begin{tabularx}{\textwidth}{ ll X}
-\hline
-Required data & constrained statements, qualifiers and references \\
-\hline
-Qualifiers & \makecell{\emph{allowed values} (P2305) -- 1..* \\ values that may be used as objects of a constrained statement} \\
-\hline
-Violation & \makecell{constrained statement with a value unequal to all allowed values} \\
-\hline
-Rule features & inequality \\
-\hline
-\end{tabularx}
-\end{table}
-
-\subsubsection{Rules}
-The one-of constraint fits in the pattern established by the previous two constraints and disallows certain property-value combinations in statements, qualifiers and references.
-Inequality is established between all \emph{allowed values} and the values of statements, qualifiers and references, respectively.
-With \(A\) the set of allowed values the rules for triple, qualifier and reference are:
-\begin{align*}
-\text{violation\_triple}(S, I, \text{propertyConstant}, V) &\leftarrow \\
-\text{statementEDB}(S, I, &\text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V) \\
-\text{violation\_qualifier}(S, \text{propertyConstant}, V) &\leftarrow \\
-\text{qualifierEDB}(S, &\text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V) \\
-\text{violation\_reference}(S, \text{propertyConstant}, V) &\leftarrow \\
-\text{referenceEDB}(S, H, &\text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
-\end{align*}
-
-\subsection{One-of qualifier value (Q52712340)}
-\subsubsection{Constraint}
-The \emph{one-of qualifier value} constraint restricts which \emph{qualifier values} may be used on a qualifier with a given \emph{qualifier property} on a constrained statement.
-
-\begin{table}[H]
-\caption{One-of qualifier value constraint (Q52712340)}
-\begin{tabularx}{\textwidth}{ ll X}
-\hline
-Required data & constrained statements and their qualifiers \\
-\hline
-Qualifiers: & \makecell{\emph{qualifier property} (P2306) -- 1 \\ property that, if used as predicate, constrains a qualifier \\
-\emph{qualifier values} (P2305) -- 1..* \\ values that are allowed for qualifiers on constrained statements \\ with the \emph{qualifier property} as predicate}\\
-\hline
-Violation & \makecell{constrained statement with a qualifier with the qualifier property \\ and a value unequal to all qualifier values} \\
-\hline
-Rule features & inequality \\
-\hline
-\end{tabularx}
-\end{table}
-
-\subsubsection{Rules}
-The one-of qualifier value constraint is violated if a constrained statement has a qualifier with a value unequal to all allowed values.
-Inequality needs to be established between all values of qualifiers and all allowed values.
-With the set of allowed Values \(A\) the rules are
-\begin{equation*}
-\begin{split}
-\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
-&\text{qualifierEDB}(S, \text{qualifierProperty}, O) \wedge \text{unequal}(\{A\}, O)
-\end{split}
-\end{equation*}
-
-\subsection{Item requires statement (Q21503247)}\label{subsec_2_item_requires_statement}
-The item requires statement constraint requires an additional statement on an item with a statement using the constrained property, possibly also using a specific value from the list of allowed values.
-\begin{table}[H]
-\caption{item requires statement constraint (Q21503247)}
-\begin{tabularx}{\textwidth}{ ll X}
-\hline
-Required data & item statement \\
-\hline
-Qualifiers & \makecell{required property (P2306) -- 1 \\ property the item statement must have \\ allowed value (P2305) -- 0..* \\ values the item statement may have} \\
-\hline
-Violation & \makecell{no item statement with required property \\ all item statements with required property have no allowed value} \\
-\hline
-Dependency &  all item statements\\
-\hline
-Rule features & negation \\
-\hline
-\end{tabularx}
-\end{table}
-
-An item violates the constraint if it a) has a statement using constrained property as predicate, b1) has no statement using the required property as predicate, or (if allowed values are specified) b2) has only statement using required property as predicate that do not use one of the allowed values as value.\\
-While condition a) can easily be tested by a statementEDB-atom with the constrained property, condition b1) and b2) pose the question if an item does not have a statement.
-This is addressed in section [?].
-
-\subsection{Value requires statement (Q21510864)}\label{subsec_2_value_requires_statement}
-The value requires statement constraint requires an additional statement on the value of constrained statement, possibly also using a specific value from the list of allowed values.
-Since the constraint works very similar to item requires statement (see \ref{subsec_2_item_requires_statement}) --- requiring a statement on the value of the constrained statement instead of on the item --- it is not elaborated on.
-
-\subsection{Inverse (Q21510855)}\label{subsec_2_inverse}
-The inverse constraint requires an additional statement on the value of a constrained statement with a specific property and the constrained item as value.
-It is very similar to value requires statement (see \ref{subsec_2_value_requires_statement}), although it does not specify allowed values, since the required statement has one specific value: the constrained item.
-While this means it could not be modelled exactly as a value requires statement constraint, it would only require very small rule changes and will thus not be elaborated on.
-
-\subsection{Symmetric (Q21510862)}
-The symmetric constraint requires an additional statement on the value of a constrained statement with the constrained property and the constrained item as value.
-It is thus equal to an inverse constraint (see \ref{subsec_2_inverse}) with the same property as constrained property and required property and will not be elaborated on.
-
-\subsection{Multi-value (Q21510857)}
-The multi-value constraint specifies that the constrained property should appear in multiple statements of an item or none.
-
-\begin{table}[H]
-\caption{Multi-value constraint (Q21510857)}
-\begin{tabularx}{\textwidth}{ ll X}
-\hline
-Required data & item statement \\
-\hline
-Violation & \makecell{constrained item with exactly one constrained statement} \\
-\hline
-Dependency & all item statements \\
-\hline
-Rule features & count \\
-\hline
-\end{tabularx}
-\end{table}
-
-To model this constraint a count of statements using the constrained property per item would be needed.
-Then a violation would be any constrained item with a count of one for the constrained property.
-Since there is no need for arbitrarily large counts, but only for the difference between one and two, this could be solved explicitly.
-
-\subsection{Single value (Q19474404)}
-The single value constraint constrains the values a property may take on a single item.
-A property marked with the single value constraint can have properties marked as separators, which define exceptions under specific circumstances.
-\begin{table}[H]
-\caption{single value constraint (Q19474404)}
-\begin{tabularx}{\textwidth}{ ll X}
-\hline
-Required data & constrained statement value \\
-\hline
-Qualifiers & \makecell{separator (P4155) -- 0..* \\ qualifiers that must be different for exception} \\
-\hline
-Violation & \makecell{two constrained statements with same item and value \\ (exception: statements with separators as qualifiers with different values)} \\
-\hline
-Dependency & \makecell{without separators: all constrained statements on one item \\ with separators: additionally all qualifiers with separators on those statements }\\
-\hline
-Rule features & inequality, partially negation \\
-\hline
-\end{tabularx}
-\end{table}
-
-If no separators are specified, two triple form a violation of this constraint if they a) belong to the same item, b) have the same constrained property, and c) have the same value.
-If separators are specified, a violation as above can be ignored if the statements each have at least one qualifier with the same separator as predicate and different values or an unequal number of separator qualifiers.\\
-In the case without separators the modelling idea would be to find two triples statementEDB with the same item and constrained property and test if their id is unequal.
-If separators were specified a violation would be the two triples and all their separator qualifiers equal or non-existent.
-\todo{Since I found the details via looking at the source code, should/can I cite that?}
-
-\subsection{Mandatory qualifier (Q21510856)}
-The mandatory qualifier constraint requires that a constrained statement has a qualifier with a specific property.
-
-\begin{table}[H]
-\caption{Mandatory qualifier constraint (Q21510856)}
-\begin{tabularx}{\textwidth}{ ll X}
-\hline
-Required data & qualifier on constrained statement \\
-\hline
-Qualifiers & \makecell{required property (P2306) -- 1 \\ qualifier that the statement must have} \\
-\hline
-Violation & \makecell{constrained statement without a qualifier with the required property} \\
-\hline
-Dependency & constrained statement and its qualifiers \\
-\hline
-Rule features & negation \\
-\hline
-\end{tabularx}
-\end{table}
-
-To find violations of this constraint one would find a statementEDB with the constrained property, but where no qualifierEDB for the statement and with the required property exists.
-
-\subsection{Single best value (Q52060874)}
-The single best value constraint requires that only one constrained statement per item has the rank preferred.
-
-\begin{table}[H]
-\caption{Single best value constraint (Q52060874)}
-\begin{tabularx}{\textwidth}{ ll X}
-\hline
-Required data & item statement rank \\
-\hline
-Qualifiers & \makecell{separator (P4155) -- 0..* \\ qualifiers that must be different for exception} \\
-\hline
-Violation & \makecell{items with two constrained statements with rank preferred} \\
-\hline
-Dependency & all item statements and their rank \\
-\hline
-Rule features & inequality \\
-\hline
-\end{tabularx}
-\end{table}
-
-Modelling the constraint means a count of all constrained statements with rank preferred is necessary, with a violation constituted if an item has a count of two or higher.
-Since it is only needed to differentiate between one and many, the rules could be written explicitly.
-
-\subsection{Type (Q21503250)}\label{subsec_2_type}
-The type constraint requires the existence of either one of a specific statement (instanceOf <class>) or a chain of statements using subclassOf to connect directly (subclassOf* <class>)
-or via instanceOf (instanceOf/subclassOf* <class>) for all classes on the constrained item.
-\begin{table}[H]
-\caption{type constraint (Q21503250)}
-\begin{tabularx}{\textwidth}{ ll X}
-\hline
-Required data & chain of statements \\
-\hline
-Qualifiers & \makecell{relation (P2309) -- 1 \\ must be instance of (Q21503252), subclass of (Q21514624), \\ or instance or subclass of (Q30208840); \\ specifies the relation the constrained item must have to the classes\\
-class (P2308) -- 1..* \\ entities that must be reached using the relation} \\
-\hline
-Violation & constrained item without relation to a class \\
-\hline
-Dependency & \makecell{statements of constrained items using instanceOf (P31) or subclassOf(P279) \\ all statements using subclassOf (P279) }\\
-\hline
-Rule features & inequality, negation \\
-\hline
-\end{tabularx}
-\end{table}
-
-If the relation is instanceOf, the constraint can be evaluated like multiple item requires statement constraints (see \ref{subsec_2_item_requires_statement}) by setting the required property to instanceOf (P31) and, for each created item requires statement constraint, the allowed value to one of the classes.
-\\
-For the relations involving subclasses all transitive subclass relationships would be concluded by setting a subclass relationship if a third item can be found that is subclass of one and subclassed by the other item.
-This would be extended with one step of instanceOf.
-A violation would then be any constrained item that is not subclass of any of the classes.
-
-\subsection{Value type (Q21510865)}
-The value type constraint works like the type constraint (see \ref{subsec_2_type}) except the statement chains are required on the value of the constrained statement instead of the constrained item.
-It will thus not be expanded upon.
-
-\section{Reasoning for unconsidered constraints}
-Multiple constraints have not been described above for different reasons.
-The most prominent of these is the format constraint (Q21502404) which specifies a regular expressions the values of the constrained property must adhere to.
-Evaluating these expressions would be theoretically possible but \\
-a) primarily results in a large amount of preprocessing and\\
-b) does not solve the question of negation, since violations would be values that do not conform to their required regular expression. \\
-
-The integer constraint (Q52848401), which requires values of a constrained property be integers (notably a quantity without decimal places), could easily be seen as a specific format constraint.
-While precomputing if a value is an integer or not could be done, it would mean that no actually relevant rules would be needed for the evaluation in datalog.
-\\
-
-Another problem that could not be sensibly solved is the comparison of values, needed for range (Q21510860), difference within range (Q21510854) and contemporary constraint (Q25796498).
-Range restricts the values of a constrained property within a certain range, difference within range specifies that two item statements should have values within a certain range and contemporary constraint specifies that subject and object of a constrained statement have to have coexisted at some point in time.
-While it would be possible to extract all values that might need to be compared and precompute the comparison, it would make most of the work precomputing and leave little to be done in datalog.\\
-
-Link constraint (Q54554025), requiring that constraint statements must have one or more references, requires a negation that cannot be concluded except by explicitly precomputing it, which makes the evaluation unnecessary.
-\\
-A similar situation is the no bounds constraint (Q51723761), which requires that quantity values have neither upper nor lower bound.
-While it would be possible to precompute if a statement has no upper or lower bound, this would shift basically all evaluation logic into precomputing.
-
-And finally commons link constraint (Q21510852) requires that values for a property should be valid links to to Wikimedia Commons.
-This validity cannot be testes from within datalog since there is no way to query the web for information.
-
-\chapter{Rules}\label{cha_rules}
-The previous chapter has shown that most constraints cannot be modelled without at least some amount of preprocessing.
+\section{Rules}
+Table \ref{tab_property_constraints} shows that most constraints cannot be modelled without at least some amount of preprocessing.
 Only \emph{scope}, \emph{conflicts with}, \emph{allowed entity types} and \emph{none of} can be modelled directly.
@@
 \emph{scope} and \emph{allowed entity types} work very similar, since there are only three respectively two possible values and violations are simply the options that have not been allowed.
-The rules for these constraints are described in section \ref{sec_no_preprocessing}.\\
+The rules for these constraints are described in section \ref{sec_simple_constraints}.\\
 
@@
 
-\section{Task}
+\subsection{Representation of Wikidata as Facts}
+To represent the content of Wikidata as facts the following predicates are introduced:
+\begin{table}[H]
+\caption{EDB predicates}\label{tab_predicates}
+\begin{tabular}{llp{8.5cm}}
+Predicate & Positions & Description \\
+\hline
+statement\_EDB & \makecell{statementID \\ entityID \\ propertyID \\ value (entityID or string)} & represents basic statements \\
+\hline
+qualifier\_EDB & \makecell{statementID \\ propertyID \\ value} & represents qualifiers on statements \\
+\hline
+reference\_EDB & \makecell{statementID \\ reference hash \\ propertyID \\ value} & represents references on statements; reference hash groups the property-value pairs belonging to one reference \\
+\hline
+item\_EDB & entityID & marks an entity as an item \\
+property\_EDB & entityID & marks an entity as a property \\
+\hline
+unit\_EDB & \makecell{value \\ unitID} & denotes which unit a value has \\
+\hline
+rank\_EDB & \makecell{statementID \\ rank} & denotes the rank a statement has
+\end{tabular}
+\end{table}
+
+\subsection{Terminology}
+\begin{table}[H]
+\caption{abbreviations}\label{tab_abbreviations}
+\begin{tabular}{ll}
+Shorthand & Description \\
+\hline
+constrained statement & statement with the constrained property as predicate \\
+constrained qualifier & qualifier with the constrained property as predicate \\
+constrained reference & reference with the constrained property as predicate \\
+constrained entity & entity with a constrained statement \\
+entity statement & statement on the same entity of a constrained statement \\
+value statement & statement on the value of a constrained statement
+\end{tabular}
+\end{table}
+
+\subsection{Task}
 Each property constraint will be modelled and evaluated as one set of rules per property per constraint, meaning each violating statement will be of a form similar to
-\begin{equation}
-\begin{split}
-\text{violation}(X, Y, \text{propertyConstant}, Z) \leftarrow \text{atoms determining the violation}
-\end{split}
-\end{equation}
-where propertyConstant denotes the constrained property. While it would be possible to load all data at once and restrict the property constant with a separate predicate akin to
-\begin{equation}
-\begin{split}
-\text{violation}&(X, Y, \text{PROPERTY}, Z) \leftarrow \\
-&\text{constrained}(\text{PROPERTY}, \text{constrainedConstant}) \wedge{} \\
-&\text{rules determining the violation}
-\end{split}
-\end{equation}
-this would increase the amount of facts loaded significantly while almost not reducing preprocessing.\\
+\begin{align*}
+&\text{violation\_statement}(S, E, \text{propertyConstant}, V) &\leftarrow \text{atoms determining the violation} \\
+&\text{violation\_qualifier}(S, \text{propertyConstant}, V) &\leftarrow \text{atoms determining the violation} \\
+&\text{violation\_reference}(S, H, \text{propertyConstant}, V) &\leftarrow \text{atoms determining the violation}
+\end{align*}
+where propertyConstant denotes the constrained property. The variables \(S\), \(I\), \(V\), and \(H\) are used for statement ID, entity ID, value and reference hash, respectively.
 
 The task for each constraint is thus: Given the constraint and a constrained property, potentially with qualifiers, which rule finds all violations of the rule in a given set of triples, qualifiers and references if applicable?
-The necessary preprocessing and filtering will be listed and the amount of rules necessary to model the constrained depending on the number of qualifiers given.
-
-\subsection{General approach}
-The goal outlined above is to identify violations of constrained properties, meaning a triple, qualifier, or reference with the constrained property will always be part of the rule body.
-With the exception of Property scope all constraints work on triples.
-The atom that will thus always be included is \(\text{statementEDB}(X, Y, \text{propertyConstant}, Z)\).\\
-
-Every constraint violation depends on some constant in relation to the constrained statement.
-These requirements have been roughly identified in the previous chapter and will now be specified as rules.
-This step can include expanding the pre-processing to load more data such as units for quantities.\\
-
-The second step is the most complex: Depending on the constrained statement and the Required data, has a violation occurred?
-For the easiest category of constraints this step does not entail any additional work since the existence of the dependent structure is itself a violation.
-However, most constraints require at least evaluation of inequality between allowed values and existing values (\ref{sec_inequality}) or the non-existence of certain facts (\ref{sec_statement_non-existence}).
-Both processes will be explained in their respective sections.
-
-The final step of translation is not necessary to identify violations per se, but provides additional information as to where exactly a violation has occurred.
-\todo[inline]{lets see how much we can actually do here}
-
-\section{No additional pre-processing}\label{sec_no_preprocessing}
-Four constraints do not require any additional features and can be easily translated, since they are violated if specific statements exists.
-A violation of these statements is thus always a triple, qualifier, or reference and violations will be expressed as the result of a query atom
-\[\text{violation\_triple}(S, I, P, V)\]
-\[\text{violation\_qualifier}(S, P, V)\]
-\[\text{violation\_reference}(S, P, V)\]
-where \(S\) specifies the statement ID, \(I\) the item, \(P\) the property, and \(V\) the value. The term propertyConstant will be used for the constant of the constrained property.
-
-\section{Inequality}\label{sec_inequality}
-All rules except the four covered in the previous section require inequality in some way or another.
-This means it is necessary to extract additional information during pre-processing, namely the constants for whom inequality needs to be stated.
+The given rule set should return all violations if queries with the following atoms:
+\begin{align*}
+&\text{violation\_statement}(S, E, P, V) \\
+&\text{violation\_qualifier}(S, P, V) \\
+&\text{violation\_reference}(S, H, P, V)
+\end{align*}
+
+\chapter{Wikidata constraints explained}
+This section will explain all interesting, meaning translatable or partially translatable, currently used property constraints on Wikidata.
+They will be presented roughly in descending order of the number of properties with this constraint, although constraints with similar function will be explained directly after one another.
+Additionally, all constraints that were considered untranslatable are listed at the and, each with a short explanation as to why they were placed in that group.
+
+Every section describing a constraint will consist of two subsections: the constraint explanation and the rules.
+
+{\bf Constraint:} This subsection starts with a table showing basic properties of the constraint.
+These include the required data, relevant qualifiers, and possible violations of the constraint.
+The required data states which part of the data in relation to the constrained statement are restricted or need to exist.
+The qualifiers list their cardinality, meaning if they are required and if they may be used multiple times, and a short description.
+The violations list the conflicting (part of a) statement in relation to the constrained statement.
+The first constraint, \emph{property scope}, will be used as an example to explain the table in detail.
+
+{\bf Rules:} Most constraints require more data than just the statements, qualifiers and references.
+This can include additional data from Wikidata to be provided as facts using the introduced predicates (see Table \ref{tab_predicates}), inequalities to be established as a pre-processing step (see Section \ref{sec_inequality}) or preparations to determine specific kinds of negation (see Section \ref{sec_statement_non-existence}).
+Afterwards, the rules necessary to determine violations are described.
+
+\section{Simple Constraints}\label{sec_simple_constraints}
+This section contains constraints that do not require any additional features and can be translated into rules directly based on the facts introduced from Wikidata.
+
+\subsection{Scope (Q53869507)}
+\paragraph{Constraint}
+The \emph{property scope} constraint specifies if a property may be used as predicate in a statement, qualifier or reference.
+\begin{table}[H]
+\caption{property scope constraint (Q53869507)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Required data & constrained statement, qualifier and reference \\
+\hline
+Qualifiers & \makecell{\emph{property scope} (P5314) -- 1..3 \\ must be \emph{as main value} (Q54828448), \emph{as qualifier} (Q54828449) \\ or \emph{as reference} (Q54828450) \\ specifies allowed use in statements, qualifiers or references} \\
+\hline
+Violation & constrained property in a position that is not allowed \\
+\hline
+\end{tabularx}
+\end{table}
+
+The required data specifies which data, in relation to the constrained statement, needs to be evaluated to conclude if the constraint was violated.
+The used abbreviations are listed in Table \ref{tab_abbreviations}.
+In this case this is simply each statement, qualifier or reference whose predicate is the constrained property.\\
+The \emph{property scope} constraint requires the scope, meaning statement, qualifier or reference, to be set via qualifier \emph{property scope} (P5314).
+The cardinality is 1..3 because it is mandatory and has only three possible values.\\
+It is violated if a statement, qualifier or reference has a predicate with the \emph{property scope} constraint that does not allow this scope.\\
+
+\paragraph{Rules}
+The property scope constraint requires no context information to be violated; the triple, qualifier, or reference with the constrained property is enough.
+This means violations can be found using these rules:
+\begin{align*}
+\text{violation\_triple}(S, I, \text{propertyConstant}, V) &\leftarrow \text{statement\_EDB}(S, I, \text{propertyConstant}, V)\\
+\text{violation\_qualifier}(S, \text{propertyConstant}, V) &\leftarrow \text{qualifier\_EDB}(S, \text{propertyConstant}, V)\\
+\text{violation\_reference}(S, \text{propertyConstant}, V) &\leftarrow \text{reference\_EDB}(S, H, \text{propertyConstant}, V)
+\end{align*}
+Which of these rules need to be used depends on the values of the \emph{property scope} qualifier.
+If a concrete constrained on a property does not have \emph{as main value} as a value of the qualifier, the first rule is used.
+The case for \emph{as qualifier} and \emph{as reference} is analogue with the second and third rule, respectively.
+
+\subsection{Conflicts with (Q21502838)}
+\paragraph{Constraint}
+The \emph{conflicts with} constraint forbids statements on constrained entities whose predicate is a specified \emph{conflicting property}.
+If \emph{conflicting values} have been specified as well, the constraint is weakened and only violated by statements if in addition to the predicate being a \emph{conflicting property}, the object is one of the \emph{conflicting values}.
+\begin{table}[H]
+\caption{Conflicts-with constraint (Q21502838)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Required data & entity statements \\
+\hline
+Qualifiers & \makecell{\emph{conflicting property} (P2306) -- 1 \\ property that may not be predicate of entity statements \\
+\emph{conflicting values} (P2305) -- 0..* \\ values that may not be objects of entity statements with the \emph{conflicting property}} \\
+\hline
+Violation & \makecell{entity statement with the conflicting property \\ entity statement with the conflicting property and one of the conflicting values} \\
+\hline
+\end{tabularx}
+\end{table}
+
+\paragraph{Rules}
+The conflicts-with constraint is violated if another statement on the same item either has a conflicting property or a conflicting property and a conflicting value depending on specified qualifiers.
+It depends on a conjunction (\(O\) denoting another statement ID, \(C\) another value) and a violation for a conflicts-with constraint is thus either
+\begin{equation*}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow\\
+&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{}\\
+&\text{statement\_EDB}(O, I, \text{conflictingPropertyConstant}, C)
+\end{split}
+\end{equation*}
+for constraints with no conflicting values specified or one
+\begin{equation*}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow\\
+&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{}\\
+&\text{statement\_EDB}(O, I, \text{conflictingPropertyConstant}, \text{conflictingValueConstant})
+\end{split}
+\end{equation*}
+for each conflicting value if specified.\\
+
+\subsection{Allowed entity types (Q52004125)}
+\paragraph{Constraint}
+The \emph{allowed entity types} constraint limits if a property may be used in a statement whose subject is an item, a property or a lexeme, although as of September 2018 lexemes are not deployed.
+
+\begin{table}[H]
+\caption{Allowed entity types constraint (Q52004125)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Required data & constrained statements and entity types\\
+\hline
+Qualifiers & \makecell{allowed entity type (P2305) -- 1..3 \\ must be \emph{Wikibase item} (Q29934200), \emph{Wikibase property} (Q29934218) \\ or lexeme (Q51885771); \\ specifies the entity type the property may be used on} \\
+\hline
+Violation & \makecell{constrained statement on an entity of a type unequal to all allowed types} \\
+\hline
+\end{tabularx}
+\end{table}
+
+Since there are only three options (and only two in use), violations can be found without inequality by finding the statement\_EDB with the constrained property whose subject is of the type that is not allowed, meaning either \emph{Wikibase item} or \emph{Wikibase property}.
+
+\paragraph{Rules}
+The allowed entity types constraint is violated by a statement with the wrong entity type.
+Constraints regarding the lexeme entity type can be ignored since they are currently not used.
+The Violations depend on a simple conjunction:
+\begin{equation*}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge \text{item}(I)
+\end{split}
+\end{equation*}
+\begin{equation*}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge \text{property}(I)
+\end{split}
+\end{equation*}
+The first rule are created for all \emph{allowed entity type} constraints whose \emph{allowed entity type} does not contain \emph{Wikibase item}.
+The second rule works analogue for \emph{Wikibase property}.
+
+\subsection{None of (Q52558054)}
+\paragraph{Constraint}
+The none of constraint restricts the values that the property may refer to by listing a set of values as forbidden.
+
+\begin{table}[H]
+\caption{None of constraint (Q52558054)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Required data & constrained statements, qualifiers and references \\
+\hline
+Qualifiers & \makecell{\emph{forbidden value} (P2305) -- 1..* \\ values that may not be object of a constraint statement, qualifier or reference} \\
+\hline
+Violation & \makecell{constrained statement, qualifier or reference with a forbidden value} \\
+\hline
+\end{tabularx}
+\end{table}
+
+\paragraph{Rules}
+The none of constraint simply forbids certain values for properties, thus effectively prohibiting certain triples, qualifiers and references.
+The rules, depending on nothing but specific database entries, are
+\begin{equation*}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
+&\text{statement\_EDB}(S, I, \text{propertyConstant}, \text{forbiddenValueConstant})
+\end{split}
+\end{equation*}
+\begin{equation*}
+\begin{split}
+\text{violation\_qualifier}&(S, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
+&\text{qualifier\_EDB}(S, \text{propertyConstant}, \text{forbiddenValueConstant})
+\end{split}
+\end{equation*}
+\begin{equation*}
+\begin{split}
+\text{violation\_reference}&(S, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
+&\text{reference\_EDB}(S, H, \text{propertyConstant}, \text{forbiddenValueConstant})
+\end{split}
+\end{equation*}
+Given a specified constraint, each of the three rules is created once for each \emph{forbidden value}.
+
+\section{Constraints using Inequality}\label{sec_constraints_using_inequality}
+All translatable constraints except the four covered in the previous section require inequality in some way or another.
+This means it is necessary to extract additional information during pre-processing, namely the constants for which inequality needs to be stated.
 Each constraint will list the possible unequal entities in relation to the constrained statement.
@@
 In every approach it will be necessary to load some basic inequalities into the database.
-To this end the predicate unequalEDB\([2]\) is introduced.
+To this end the predicate unequal\_EDB\([2]\) is introduced.
 Since the goal is to keep the number of applications of this predicate small, an IDB predicate unequal\([2]\) has to be introduced.
-This leads to the linear rule
-\begin{equation*}
-\text{unequal}(X, Y) \leftarrow \text{unequalEDB}(X, Y)
-\end{equation*}
-
-Given a set of entities \(E\), the task is to generate rules so that \(\forall x, y \in E, x \ne y: \text{unequal}(x, y)\).
-In doing so, dependent on the number of entities \(n = |E|\), the sum of the number of introduced EDB facts \(f(n)\) and the number of introduced rules \(r(n)\) should be kept as small as possible.
-Note that \(n > 1\) is presumed.
-The effects will be evaluated in section \ref{sec_inequality_evaluation}.
-
-\todo[inline]{How are the three numbers weighted? Or do I determine that in evaluation?}
-\todo[inline]{Is the number of materialised facts even relevant? Maybe look at rule applications.}
-%, and the number of materialised facts \(m(n)\)
+This leads to the rule
+\begin{equation*}
+\text{unequal}(X, Y) \leftarrow \text{unequal\_EDB}(X, Y)
+\end{equation*}
+
+Given a set of entities \(E\), the task is to generate rules so that for all \( x, y \in E: x \ne y \Rightarrow \text{unequal}(x, y)\).
+
 \subsubsection{Naive}\label{subsubsec_naive}
-The naive approach would be to to simply add all possible combinations:
-\begin{equation*}
-\forall x, y \in E, x \ne y \text{ add } \text{unequalEDB}(x, y)
-\end{equation*}.
+The naive approach would be to simply add all possible combinations:
+\begin{equation*}
+\text{for all } x, y \in E, x \ne y \text{ add } \text{unequal\_EDB}(x, y)
+\end{equation*}
 This can be easily improved using the inverse and adding the following rule:
@@
 \end{equation}
-The it is sufficient to add each combinations once if disregarding their order:
-\begin{equation*}
-\forall x, y \in E, x < y \text{ add } \text{unequalEDB}(x, y)
+Then it is sufficient to add each combination once. To achieve this assume an arbitrary total order \(<\) has been established on the set \(E\):
+\begin{equation*}
+\text{for all } x, y \in E, x < y \text{ add } \text{unequal\_EDB}(x, y)
 \end{equation*}.
-
-The naive approach with inverse would result in \(f(n) = \binom{n}{2}\) EDB facts and \(r(n) = 1\) additional rule.
-%and \(m(n) = \binom{n}{2}\) materialized facts since equation \ref{equ_inverse} would be applied exactly once to every EDB fact.
 
@@
 
-\section{Statement non-existence}\label{sec_statement_non-existence}
-Multiple statements are violated if specific statements do not exist in relation to constrained statements.
+\subsection{Distinct values (Q21502410)}
+\paragraph{Constraint}
+The \emph{distinct values} constraint forbids the existence of two constrained statements with the same object.
+\begin{table}[H]
+\caption{distinct values constraint (Q21502410)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Required data & constrained statements \\
+\hline
+Violations & two constrained statements with same value \\
+\hline
+Rule features & inequality \\
+\hline
+\end{tabularx}
+\end{table}
+The constraint can be expanded to cover qualifiers and references as well by setting a \emph{constraint scope} (P4680) qualifier with the respective value.
+However, as of September 2018 there is no constraint making use of this.
+
+\paragraph{Rules}
+The \emph{distinct values} constraint requires two distinct constrained statements with the same value. 
+Two constrained statements with the same value can be found via a simple conjunction of two statement\_EDB atoms with the propertyConstant as predicate and a variable \(V\) as value.
+Since this would mark every constrained statement as a violations because it can fulfil both atoms at the same time, inequality of the statement IDs \(S\) and \(O\) must be tested as well.t
+\begin{equation*}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
+&\text{statement\_EDB}(O, X, \text{propertyConstant}, V) \wedge{} \\
+&\text{unequal}(S, O)
+\end{split}
+\end{equation*}
+The rules for qualifiers and reference are analogue.
+
+\subsection{Allowed units (Q21514353)}\label{subsec_3_allowed_units}
+\paragraph{Constraint}
+The \emph{allowed units} constraint specifies that the value of a constrained statement may only have a unit that is in a set of specific units.
+
+\begin{table}[H]
+\caption{Allowed units constraint (Q21514353)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Required data & constrained statements \\
+\hline
+Qualifiers & \makecell{\emph{allowed unit} (P2305) -- 1..* \\ units that are allowed on values of constrained statements (\emph{none} for dimensionless units)} \\
+\hline
+Violation & \makecell{constrained statement with a unit unequal to all allowed units} \\
+\hline
+Rule features & inequality \\
+\hline
+\end{tabularx}
+\end{table}
+
+By setting the \emph{constraint scope} (P4680) qualifier on the constraint it can be limited to only affect a subset of statement, qualifier, and reference.
+As of September 2018 this is not done on any constraint.
+
+\paragraph{Rules}
+To handle this constraint inequality between all units need to be established.
+Then a violation can be identified by finding all statement\_EDB, qualifier\_EDB and reference\_EDB with the constrained property and a unit unequal to all allowed units.
+The allowed units constraint needs to determine inequality between all allowed units and each unit on a constrained statement.
+To this end the units of a triple, qualifier or reference needs to be extracted during pre-processing and their inequality established.
+Quantities are rendered as string during pre-processing, specifying their value, bounds and unit.
+
+To determine if a violation occurred, the unit of the constrained statement needs to be found. This is done via the unit-predicate as below
+\begin{equation*}
+\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge \text{unit\_EDB}(V, U)
+\end{equation*}
+with qualifiers and references treated analogue.
+
+The complete rules require multiple unequal-atoms for all allowed units.
+Given a set \(A\) of constants, in this case allowed units, \(\text{unequal}(\{A\}, t)\) with an arbitrary term \(t\) is short for \(\bigwedge_{a \in A} \text{unequal}(a, t)\)
+
+Violation are thus
+\begin{equation*}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
+&\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
+\end{split}
+\end{equation*}
+\begin{equation*}
+\begin{split}
+\text{violation\_qualifier}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+&\text{qualifier\_EDB}(S, \text{propertyConstant}, V) \wedge{} \\
+&\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
+\end{split}
+\end{equation*}
+\begin{equation*}
+\begin{split}
+\text{violation\_reference}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+&\text{reference\_EDB}(S, H, \text{propertyConstant}, V) \wedge{} \\
+&\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
+\end{split}
+\end{equation*}
+
+\subsection{Allowed qualifiers (Q21510851)}
+\paragraph{Constraint}
+The \emph{allowed qualifiers} constraint specifies that only a limited set of properties may be used as predicates of qualifiers on a constrained statement.
+
+\begin{table}[H]
+\caption{Allowed qualifiers constraint (Q21510851)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Required data & qualifiers on constrained statements \\
+\hline
+Qualifiers & \makecell{\emph{allowed qualifier} (P2306) -- 1..* \\ properties that may appear in predicates of qualifiers on constrained statements} \\
+\hline
+Violation & \makecell{constrained statement with a qualifier unequal to all allowed qualifiers} \\
+\hline
+Rule features & inequality \\
+\hline
+\end{tabularx}
+\end{table}
+
+\paragraph{Rules}
+To represent a violation of the \emph{allowed qualifiers} constraint in datalog inequality between all specified allowed qualifiers and the properties of qualifiers is required.
+The constraint is violated if a qualifier on a constrained statement exists that is unequal to all allowed qualifiers.
+The notation for set inequality as introduced in section \ref{subsec_3_allowed_units} will be used.
+With \(A\) the set of allowed qualifiers the rule is:
+\begin{equation*}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
+&\text{qualifier\_EDB}(S, Q, O) \wedge \text{unequal}(\{A\}, Q)
+\end{split}
+\end{equation*}
+
+\subsection{One-of (Q21510859)}
+\paragraph{Constraint}
+The \emph{one-of} constraint limits the values of a constraint statement, qualifier or reference to a set of given values.
+
+\begin{table}[H]
+\caption{One-of constraint (Q21510859)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Required data & constrained statements, qualifiers and references \\
+\hline
+Qualifiers & \makecell{\emph{allowed values} (P2305) -- 1..* \\ values that may be used as objects of a constrained statement} \\
+\hline
+Violation & \makecell{constrained statement with a value unequal to all allowed values} \\
+\hline
+Rule features & inequality \\
+\hline
+\end{tabularx}
+\end{table}
+
+\paragraph{Rules}
+The one-of constraint fits in the pattern established by the previous two constraints and disallows certain property-value combinations in statements, qualifiers and references.
+Inequality is established between all \emph{allowed values} and the values of statements, qualifiers and references, respectively.
+With \(A\) the set of allowed values the rules for triple, qualifier and reference are:
+\begin{align*}
+\text{violation\_triple}(S, I, \text{propertyConstant}, V) &\leftarrow \\
+\text{statement\_EDB}(S, I, &\text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V) \\
+\text{violation\_qualifier}(S, \text{propertyConstant}, V) &\leftarrow \\
+\text{qualifier\_EDB}(S, &\text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V) \\
+\text{violation\_reference}(S, \text{propertyConstant}, V) &\leftarrow \\
+\text{reference\_EDB}(S, H, &\text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
+\end{align*}
+
+\subsection{One-of qualifier value (Q52712340)}
+\paragraph{Constraint}
+The \emph{one-of qualifier value} constraint restricts which \emph{qualifier values} may be used on a qualifier with a given \emph{qualifier property} on a constrained statement.
+
+\begin{table}[H]
+\caption{One-of qualifier value constraint (Q52712340)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Required data & constrained statements and their qualifiers \\
+\hline
+Qualifiers: & \makecell{\emph{qualifier property} (P2306) -- 1 \\ property that, if used as predicate, constrains a qualifier \\
+\emph{qualifier values} (P2305) -- 1..* \\ values that are allowed for qualifiers on constrained statements \\ with the \emph{qualifier property} as predicate}\\
+\hline
+Violation & \makecell{constrained statement with a qualifier with the qualifier property \\ and a value unequal to all qualifier values} \\
+\hline
+Rule features & inequality \\
+\hline
+\end{tabularx}
+\end{table}
+
+\paragraph{Rules}
+The one-of qualifier value constraint is violated if a constrained statement has a qualifier with a value unequal to all allowed values.
+Inequality needs to be established between all values of qualifiers and all allowed values.
+With the set of allowed Values \(A\) the rules are
+\begin{equation*}
+\begin{split}
+\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
+&\text{qualifier\_EDB}(S, \text{qualifierProperty}, O) \wedge \text{unequal}(\{A\}, O)
+\end{split}
+\end{equation*}
+
+\section{Constraints using Statement non-Existence}
+Several constraints are violated if specific statements do not exist in relation to constrained statements.
 It is not possible to simply negate said statements and pre-processing every required statement would defeat the purpose of implementing them in datalog.
@@
 The require-propagation is regarded as given.
+
+\subsection{Item requires statement (Q21503247)}\label{subsec_2_item_requires_statement}
+The item requires statement constraint requires an additional statement on an item with a statement using the constrained property, possibly also using a specific value from the list of allowed values.
+\begin{table}[H]
+\caption{item requires statement constraint (Q21503247)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Required data & item statement \\
+\hline
+Qualifiers & \makecell{required property (P2306) -- 1 \\ property the item statement must have \\ allowed value (P2305) -- 0..* \\ values the item statement may have} \\
+\hline
+Violation & \makecell{no item statement with required property \\ all item statements with required property have no allowed value} \\
+\hline
+Dependency &  all item statements\\
+\hline
+Rule features & negation \\
+\hline
+\end{tabularx}
+\end{table}
+
+An item violates the constraint if it a) has a statement using constrained property as predicate, b1) has no statement using the required property as predicate, or (if allowed values are specified) b2) has only statement using required property as predicate that do not use one of the allowed values as value.\\
+While condition a) can easily be tested by a statement\_EDB-atom with the constrained property, condition b1) and b2) pose the question if an item does not have a statement.
+This is addressed in section [?].
+
+\subsection{Value requires statement (Q21510864)}\label{subsec_2_value_requires_statement}
+The value requires statement constraint requires an additional statement on the value of constrained statement, possibly also using a specific value from the list of allowed values.
+Since the constraint works very similar to item requires statement (see \ref{subsec_2_item_requires_statement}) --- requiring a statement on the value of the constrained statement instead of on the item --- it is not elaborated on.
+
+\subsection{Inverse (Q21510855)}\label{subsec_2_inverse}
+The inverse constraint requires an additional statement on the value of a constrained statement with a specific property and the constrained item as value.
+It is very similar to value requires statement (see \ref{subsec_2_value_requires_statement}), although it does not specify allowed values, since the required statement has one specific value: the constrained item.
+While this means it could not be modelled exactly as a value requires statement constraint, it would only require very small rule changes and will thus not be elaborated on.
+
+\subsection{Symmetric (Q21510862)}
+The symmetric constraint requires an additional statement on the value of a constrained statement with the constrained property and the constrained item as value.
+It is thus equal to an inverse constraint (see \ref{subsec_2_inverse}) with the same property as constrained property and required property and will not be elaborated on.
+
+\subsection{Multi-value (Q21510857)}
+The multi-value constraint specifies that the constrained property should appear in multiple statements of an item or none.
+
+\begin{table}[H]
+\caption{Multi-value constraint (Q21510857)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Required data & item statement \\
+\hline
+Violation & \makecell{constrained item with exactly one constrained statement} \\
+\hline
+Dependency & all item statements \\
+\hline
+Rule features & count \\
+\hline
+\end{tabularx}
+\end{table}
+
+To model this constraint a count of statements using the constrained property per item would be needed.
+Then a violation would be any constrained item with a count of one for the constrained property.
+Since there is no need for arbitrarily large counts, but only for the difference between one and two, this could be solved explicitly.
+
+\subsection{Single value (Q19474404)}
+The single value constraint constrains the values a property may take on a single item.
+A property marked with the single value constraint can have properties marked as separators, which define exceptions under specific circumstances.
+\begin{table}[H]
+\caption{single value constraint (Q19474404)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Required data & constrained statement value \\
+\hline
+Qualifiers & \makecell{separator (P4155) -- 0..* \\ qualifiers that must be different for exception} \\
+\hline
+Violation & \makecell{two constrained statements with same item and value \\ (exception: statements with separators as qualifiers with different values)} \\
+\hline
+Dependency & \makecell{without separators: all constrained statements on one item \\ with separators: additionally all qualifiers with separators on those statements }\\
+\hline
+Rule features & inequality, partially negation \\
+\hline
+\end{tabularx}
+\end{table}
+
+If no separators are specified, two triple form a violation of this constraint if they a) belong to the same item, b) have the same constrained property, and c) have the same value.
+If separators are specified, a violation as above can be ignored if the statements each have at least one qualifier with the same separator as predicate and different values or an unequal number of separator qualifiers.\\
+In the case without separators the modelling idea would be to find two triples statement\_EDB with the same item and constrained property and test if their id is unequal.
+If separators were specified a violation would be the two triples and all their separator qualifiers equal or non-existent.
+\todo{Since I found the details via looking at the source code, should/can I cite that?}
+
+\subsection{Mandatory qualifier (Q21510856)}
+The mandatory qualifier constraint requires that a constrained statement has a qualifier with a specific property.
+
+\begin{table}[H]
+\caption{Mandatory qualifier constraint (Q21510856)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Required data & qualifier on constrained statement \\
+\hline
+Qualifiers & \makecell{required property (P2306) -- 1 \\ qualifier that the statement must have} \\
+\hline
+Violation & \makecell{constrained statement without a qualifier with the required property} \\
+\hline
+Dependency & constrained statement and its qualifiers \\
+\hline
+Rule features & negation \\
+\hline
+\end{tabularx}
+\end{table}
+
+To find violations of this constraint one would find a statement\_EDB with the constrained property, but where no qualifier\_EDB for the statement and with the required property exists.
+
+\subsection{Single best value (Q52060874)}
+The single best value constraint requires that only one constrained statement per item has the rank preferred.
+
+\begin{table}[H]
+\caption{Single best value constraint (Q52060874)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Required data & item statement rank \\
+\hline
+Qualifiers & \makecell{separator (P4155) -- 0..* \\ qualifiers that must be different for exception} \\
+\hline
+Violation & \makecell{items with two constrained statements with rank preferred} \\
+\hline
+Dependency & all item statements and their rank \\
+\hline
+Rule features & inequality \\
+\hline
+\end{tabularx}
+\end{table}
+
+Modelling the constraint means a count of all constrained statements with rank preferred is necessary, with a violation constituted if an item has a count of two or higher.
+Since it is only needed to differentiate between one and many, the rules could be written explicitly.
+
+\subsection{Type (Q21503250)}\label{subsec_2_type}
+The type constraint requires the existence of either one of a specific statement (instanceOf <class>) or a chain of statements using subclassOf to connect directly (subclassOf* <class>)
+or via instanceOf (instanceOf/subclassOf* <class>) for all classes on the constrained item.
+\begin{table}[H]
+\caption{type constraint (Q21503250)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Required data & chain of statements \\
+\hline
+Qualifiers & \makecell{relation (P2309) -- 1 \\ must be instance of (Q21503252), subclass of (Q21514624), \\ or instance or subclass of (Q30208840); \\ specifies the relation the constrained item must have to the classes\\
+class (P2308) -- 1..* \\ entities that must be reached using the relation} \\
+\hline
+Violation & constrained item without relation to a class \\
+\hline
+Dependency & \makecell{statements of constrained items using instanceOf (P31) or subclassOf(P279) \\ all statements using subclassOf (P279) }\\
+\hline
+Rule features & inequality, negation \\
+\hline
+\end{tabularx}
+\end{table}
+
+If the relation is instanceOf, the constraint can be evaluated like multiple item requires statement constraints (see \ref{subsec_2_item_requires_statement}) by setting the required property to instanceOf (P31) and, for each created item requires statement constraint, the allowed value to one of the classes.
+\\
+For the relations involving subclasses all transitive subclass relationships would be concluded by setting a subclass relationship if a third item can be found that is subclass of one and subclassed by the other item.
+This would be extended with one step of instanceOf.
+A violation would then be any constrained item that is not subclass of any of the classes.
+
+\subsection{Value type (Q21510865)}
+The value type constraint works like the type constraint (see \ref{subsec_2_type}) except the statement chains are required on the value of the constrained statement instead of the constrained item.
+It will thus not be expanded upon.
+
+\section{Reasoning for unconsidered constraints}
+Multiple constraints have not been described above for different reasons.
+The most prominent of these is the format constraint (Q21502404) which specifies a regular expressions the values of the constrained property must adhere to.
+Evaluating these expressions would be theoretically possible but \\
+a) primarily results in a large amount of preprocessing and\\
+b) does not solve the question of negation, since violations would be values that do not conform to their required regular expression. \\
+
+The integer constraint (Q52848401), which requires values of a constrained property be integers (notably a quantity without decimal places), could easily be seen as a specific format constraint.
+While precomputing if a value is an integer or not could be done, it would mean that no actually relevant rules would be needed for the evaluation in datalog.
+\\
+
+Another problem that could not be sensibly solved is the comparison of values, needed for range (Q21510860), difference within range (Q21510854) and contemporary constraint (Q25796498).
+Range restricts the values of a constrained property within a certain range, difference within range specifies that two item statements should have values within a certain range and contemporary constraint specifies that subject and object of a constrained statement have to have coexisted at some point in time.
+While it would be possible to extract all values that might need to be compared and precompute the comparison, it would make most of the work precomputing and leave little to be done in datalog.\\
+
+Link constraint (Q54554025), requiring that constraint statements must have one or more references, requires a negation that cannot be concluded except by explicitly precomputing it, which makes the evaluation unnecessary.
+\\
+A similar situation is the no bounds constraint (Q51723761), which requires that quantity values have neither upper nor lower bound.
+While it would be possible to precompute if a statement has no upper or lower bound, this would shift basically all evaluation logic into precomputing.
+
+And finally commons link constraint (Q21510852) requires that values for a property should be valid links to to Wikimedia Commons.
+This validity cannot be testes from within datalog since there is no way to query the web for information.
+
+\section{Statement non-existence}\label{sec_statement_non-existence}
 
@@
 \begin{split}
-c_1 = &\text{statementEDB}(Q, I, \text{propertyConstant}, X) \wedge{} \\
-&\text{statementEDB}(S, I, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
+c_1 = &\text{statement\_EDB}(Q, I, \text{propertyConstant}, X) \wedge{} \\
+&\text{statement\_EDB}(S, I, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
 \end{split}
@@
 \begin{split}
-c_2 = &\text{statementEDB}(Q, I, \text{propertyConstant}, X) \wedge{} \\
-&\text{statementEDB}(S, I, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
+c_2 = &\text{statement\_EDB}(Q, I, \text{propertyConstant}, X) \wedge{} \\
+&\text{statement\_EDB}(S, I, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
 \end{split}
@@
 \text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
-&\text{statementEDB}(O, I, P, X) \wedge \text{last}(O, I) \wedge{} \\
+&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
+&\text{statement\_EDB}(O, I, P, X) \wedge \text{last}(O, I) \wedge{} \\
 &\text{require}(O, \text{propertyConstant}, \text{requiredPropertyConstant})
@@
 \begin{split}
-c_1 = &\text{statementEDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
-&\text{statementEDB}(S, I, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
+c_1 = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
+&\text{statement\_EDB}(S, I, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
 \end{split}
@@
 \begin{split}
-c_2 = &\text{statementEDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
-&\text{statementEDB}(S, I, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
+c_2 = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
+&\text{statement\_EDB}(S, I, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
 \end{split}
@@
 \text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
-&\text{statementEDB}(O, V, P, X) \wedge \text{last}(O, V) \wedge{} \\
+&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
+&\text{statement\_EDB}(O, V, P, X) \wedge \text{last}(O, V) \wedge{} \\
 &\text{require}(O, \text{propertyConstant}, \text{requiredPropertyConstant})
@@
 \begin{split}
-c_1 = &\text{statementEDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
-&\text{statementEDB}(S, I, P, V) \wedge{} \\
+c_1 = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
+&\text{statement\_EDB}(S, I, P, V) \wedge{} \\
 &\text{unequal}(\text{inversePropertyConstant}, P)
@@
 \begin{split}
-c_2 = &\text{statementEDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
-&\text{statementEDB}(S, I, \text{inversePropertyConstant}, V) \wedge{} \\
+c_2 = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
+&\text{statement\_EDB}(S, I, \text{inversePropertyConstant}, V) \wedge{} \\
 &\text{unequal}(R, V)
@@
 \begin{split}
-c = &\text{statementEDB}(Q, I, \text{propertyConstant}, X) \wedge{} \\
-&\text{statementEDB}(S, I, P, V) \wedge \text{unequal}(\text{propertyConstant}, P)
+c = &\text{statement\_EDB}(Q, I, \text{propertyConstant}, X) \wedge{} \\
+&\text{statement\_EDB}(S, I, P, V) \wedge \text{unequal}(\text{propertyConstant}, P)
 \end{split}
@@
 \text{require\_second}&(S, \text{propertyConstant}, \text{propertyConstant}) \leftarrow \\
-&\text{next}(R, S) \wedge \text{statementEDB}(R, I, \text{propertyConstant}, C) \wedge{} \\
+&\text{next}(R, S) \wedge \text{statement\_EDB}(R, I, \text{propertyConstant}, C) \wedge{} \\
 &\text{next}(Q, R) \wedge \text{require}(Q, \text{propertyConstant}, \text{propertyConstant})  \wedge c
@@
 \text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
-&\text{statementEDB}(O, I, \text{propertyConstant}, X) \wedge{} \\
+&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
+&\text{statement\_EDB}(O, I, \text{propertyConstant}, X) \wedge{} \\
 &\text{unequal}(S, O)
@@
 \text{violation\_qualifier}&(S, \text{propertyConstant}, V) \leftarrow \\
-&\text{qualifierEDB}(S, \text{propertyConstant}, V) \wedge{} \\
-&\text{qualifierEDB}(S, \text{propertyConstant}, X) \wedge{} \\
+&\text{qualifier\_EDB}(S, \text{propertyConstant}, V) \wedge{} \\
+&\text{qualifier\_EDB}(S, \text{propertyConstant}, X) \wedge{} \\
 &\text{unequal}(V, X)
@@
 \text{violation\_reference}&(S, \text{propertyConstant}, V) \leftarrow \\
-&\text{referenceEDB}(S, H, \text{propertyConstant}, V) \wedge{} \\
-&\text{referenceEDB}(S, J, \text{propertyConstant}, X) \wedge{} \\
+&\text{reference\_EDB}(S, H, \text{propertyConstant}, V) \wedge{} \\
+&\text{reference\_EDB}(S, J, \text{propertyConstant}, X) \wedge{} \\
 &\text{unequal}(V, X)
@@
 \begin{equation*}
-c = \text{statementEDB}(S, I, \text{propertyConstant}, C) \wedge \text{qualifierEDB}(S, P, V) \wedge \text{unequal}(P, \text{separatorConstant})
+c = \text{statement\_EDB}(S, I, \text{propertyConstant}, C) \wedge \text{qualifier\_EDB}(S, P, V) \wedge \text{unequal}(P, \text{separatorConstant})
 \end{equation*}
@@
 \begin{split}
-\text{does\_not\_have}(S, R) \leftarrow &\text{statementEDB}(S, I, \text{propertyConstant}, C) \wedge{} \\
+\text{does\_not\_have}(S, R) \leftarrow &\text{statement\_EDB}(S, I, \text{propertyConstant}, C) \wedge{} \\
 &\text{last\_qualifier}(S, P, V) \wedge \text{require}(S, P, V, \text{propertyConstant}, R)
@@
 \begin{split}
-\text{has\_same}(S, O, Q) \leftarrow &\text{statementEDB}(S, I, \text{propertyConstant}, C) \wedge{} \\
-&\text{statementEDB}(O, I, \text{propertyConstant}, X) \wedge{} \\
-&\text{qualifierEDB}(S, Q, V) \wedge \text{qualifierEDB}(O, Q, V)
+\text{has\_same}(S, O, Q) \leftarrow &\text{statement\_EDB}(S, I, \text{propertyConstant}, C) \wedge{} \\
+&\text{statement\_EDB}(O, I, \text{propertyConstant}, X) \wedge{} \\
+&\text{qualifier\_EDB}(S, Q, V) \wedge \text{qualifier\_EDB}(O, Q, V)
 \end{split}
@@
 \text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
-&\text{statementEDB}(O, I, \text{propertyConstant}, C) \wedge{} \\
+&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
+&\text{statement\_EDB}(O, I, \text{propertyConstant}, C) \wedge{} \\
 &\text{unequal}(S, O) \wedge{} \\
@@
 \begin{split}
-c = &\text{statementEDB}(S, I, \text{propertyConstant}, C) \wedge{} \\
-&\text{qualifierEDB}(S, P, V) \wedge \text{unequal}(P, \text{requiredPropertyConstant})
+c = &\text{statement\_EDB}(S, I, \text{propertyConstant}, C) \wedge{} \\
+&\text{qualifier\_EDB}(S, P, V) \wedge \text{unequal}(P, \text{requiredPropertyConstant})
 \end{split}
@@
 \text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge \text{does\_not\_have}(S, \text{requiredPropertyConstant})
+&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge \text{does\_not\_have}(S, \text{requiredPropertyConstant})
 \end{split}
@@
 \text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
-&\text{rank}(S, \text{preferredRankConstant}) \wedge{} \\
-&\text{statementEDB}(O, I, \text{propertyConstant}, X) \wedge{} \\
-&\text{rank}(O, \text{preferredRankConstant}) \wedge{} \\
+&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
+&&\text{rank\_EDB}(S, \text{preferredRankConstant}) \wedge{} \\
+&\text{statement\_EDB}(O, I, \text{propertyConstant}, X) \wedge{} \\
+&&\text{rank\_EDB}(O, \text{preferredRankConstant}) \wedge{} \\
 &\text{unequal}(S, O)
@@
 \text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statementEDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
-&\text{rank}(S, \text{preferredRankConstant}) \wedge{} \\
-&\text{statementEDB}(O, I, \text{propertyConstant}, W) \wedge{} \\
-&\text{rank}(O, \text{preferredRankConstant}) \wedge{} \\
+&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
+&&\text{rank\_EDB}(S, \text{preferredRankConstant}) \wedge{} \\
+&\text{statement\_EDB}(O, I, \text{propertyConstant}, W) \wedge{} \\
+&&\text{rank\_EDB}(O, \text{preferredRankConstant}) \wedge{} \\
 &\text{unequal}(S, O) \wedge{} \\

### 2018-08-31 thesis/paper/paper.tex
Improved inequality description

@@
 
-Given a set of entities \(E\), the task is to generate rules so that for all \( x, y \in E: x \ne y \Rightarrow \text{unequal}(x, y)\).
+Given a set of unequal constants \(U\), the task is to generate rules so that for all \( x, y \in U: x \ne y \Rightarrow \text{unequal}(x, y)\).
 
@@
 \begin{equation*}
-\text{for all } x, y \in E, x \ne y \text{ add } \text{unequal\_EDB}(x, y)
+\text{for all } x, y \in U, x \ne y \text{ add } \text{unequal\_EDB}(x, y)
 \end{equation*}
@@
 \begin{equation*}
-\text{for all } x, y \in E, x < y \text{ add } \text{unequal\_EDB}(x, y)
-\end{equation*}.
-
-\subsubsection{Encoding}\label{subsubsec_encoding}
-One option to determine inequality would be to utilise that all entities are strings of single characters.
-The idea is to determine the set of characters used in all entities \(C\), set them as unequal as outlined above in section \ref{subsubsec_naive} and associate each entity with their characters.\\
-
-An entity \(e \in E\) is always a string and can be written as \(c_0 \ldots c_{l - 1}\) with \(l = len(e)\) being the length of the string.
-Denote the set \(C = \{c | \exists e \in E: c \in e\}\)\todo{\(c \in e\) does not seem correct} as the set of characters occurring in entities.
-Then \(l_{\max} = \max\limits_{\forall e \in E}(len(e))\) is the length of the longest entity string.
-Based on this a group of predicates letter\_i\_EDB[8] is introduced where \(0 \equiv i \mod 7\), \(0 \leq i < l_{max} \).
-
-For each entity string \(e = c_0 \ldots c_{l - 1}\) each predicate letter\_i\_EDB is created as \(\text{letter\_i\_EDB}(e, f(e, i), f(e, i+1), \ldots, f(e, i+6)\), where
+\text{for all } x, y \in U, x < y \text{ add } \text{unequal\_EDB}(x, y)
+\end{equation*}
+
+\subsubsection{Encoding constant strings to establish inequality}\label{subsubsec_encoding}
+One way of determining inequality would be to take advantage of the fact that inequality is established in a set of unequal constants \(U\), which can be represented as a string.
+The idea is to determine the set of characters used in \(U\), set these characters as unequal as outlined in section \ref{subsubsec_encoding} and associate each unequal constant with their characters.\\
+
+An unequal constant \(u \in U\) can be represented a string and be written as \(c_0 \ldots c_{l - 1}\) with \(l = len(u)\) being the length of the string.
+Denote the set \(C\) as the set of characters occurring in unequal strings.
+Then \(l_{\max} = \max\limits_{\forall u \in U}(len(u))\) is the length of the longest entity string that occurs in the given data.
+Based on this a group of predicates letter\(_i\)\_EDB[8] is introduced where \(0 \equiv i \mod 7\), \(0 \leq i < l_{max} \).
+
+For each unequal constant string \(u = c_0 \ldots c_{l - 1}\) and \(0 \equiv i \mod 7, 0 \leq i < len(u)\) an atom using predicate letter\(_i\)\_EDB is created as 
+\begin{equation*}
+\text{letter}_i\text{\_EDB}(e, f(e, i), f(e, i+1), \ldots, f(e, i+6)
+\end{equation*}where
 \begin{equation*}
@@
       c_i & i < len(e) \\
-      \text{NONE} & i \geq len(e)
+      \text{none} & i \geq len(e)
 \end{cases}
@@
 
-Now that inequality between all constitution characters and the link between entities and their characters are introduced, the inequality between two entities can be derived by creating the following rules for all predicates letter\_i (with \text{letters} short for \(\text{letter\_i}(X, A_i, \ldots, A_{i+14}) \wedge\text{letter\_i}(Y, B_i, \ldots, B_{i+6})\)):
-\begin{equation*}
-\text{unequal}(X, Y) \leftarrow \text{letters} \wedge \text{unequal}(A_i, B_i)
-\end{equation*}
-\begin{equation*}
-\text{unequal}(X, Y) \leftarrow \text{letters} \wedge \text{unequal}(A_{i+6}, B_{i+6})
-\end{equation*}
-
-With this approach the statistics depend on the number of distinct characters \(n_c = |C|\).
-The number of introduced EDB facts is \(f(n, n_c) = \binom{n_c}{2} + n*l_{\max}\) and the number of introduced rules \(r(n, n_c) = 1 + l_{\max}\).
-The effectiveness of this approach thus depends on the ratio between \(n\) and \(n_c\).
+Now that inequality between all characters and the link between entities and their characters are introduced, the inequality between two entities can be derived by creating the following rules for all predicates letter\(_i\)\_EDB (with \text{letters} short for \(\text{letter}_i\text{\_EDB}(X, A_i, \ldots, A_{i+6}) \wedge\text{letter}_i\text{\_EDB}(Y, B_i, \ldots, B_{i+6})\), \(i \leq j < i + 7\)):
+\begin{equation*}
+\text{unequal}(X, Y) \leftarrow \text{letters} \wedge \text{unequal}(A_j, B_j)
+\end{equation*}
+Note that the facts fulfilling \(\text{unequal}(A_j, B_j)\) are created by using the approach outlined in Section \ref{subsubsec_naive} for the set of characters \(C\) in all unequal constants \(U\).
 
 \subsubsection{Demand-driven materialisation}
-The approaches outlined above will always result in all inequalities between the members of \(E\) being materialised.
+The approaches outlined above will always result in all inequalities between the members of \(U\) being materialised.
 However, the purpose of determining inequality is to assist in finding violations of specific constraints.
@@
 To save space and avoid redundancy the following rules will not explicitly state the inequality requirement predicate.
-Instead, assume that for every premise with the form \(B_1 \wedge \ldots\wedge B_m \wedge I_1 \wedge I_n\) where \(B_1 ..
+Instead, assume that for every premise of the form \(B_1 \wedge \ldots\wedge B_m \wedge I_1 \wedge I_n\) where \(B_1 ..
 B_m\) are arbitrary atoms, \(I_1 ..
-I_n\) are atoms of the form \(\text{unequal}(t_{n_1}, t_{n_2})\) with \(t_{n_1}, t_{n_2}\) terms and all of \(t_{n_1}, t_{n_2}\) that are variables appear in at least one of \(B_1 \ldots B_m\), the rules
+I_n\) are atoms of the form \(\text{unequal}(t_{n_1}, t_{n_2})\) with \(t_{n_1}, t_{n_2}\) terms and all variables amount \(t_{n_1}\) and \(t_{n_2}\) appear in at least one of \(B_1 \ldots B_m\), the rules
 \begin{equation*}
@@
 
-The rules outlined in section \ref{subsubsec_encoding} are changed to utilise this new requirement.
-\begin{equation*}
-\text{unequal}(X, Y) \leftarrow \text{req\_inequality}(X, Y) \wedge \text{letter\_i}(X, A) \wedge \text{letter\_i}(Y, B) \wedge \text{unequal}(A, B)
+The rules outlined in Section \ref{subsubsec_encoding} are changed to utilise this new requirement.
+\begin{equation*}
+\text{unequal}(X, Y) \leftarrow \text{req\_inequality}	(X, Y) \wedge \text{letters} \wedge \text{unequal}(A_j, B_j)
 \end{equation*}

### 2018-08-31 thesis/paper/paper.tex
Improved require pattern

@@
 
-If qualifier non-existence should be tested the predicates have to be adapted, since they do not have a unique identifier.
-Their order has to be established using the entire set of statement ID, property and value.
+If qualifier non-existence should be tested the predicates have to be adapted, since qualifiers do not have a unique identifier.
+To assign them a unique identity, statement ID, property and value have to be used.
 For this the predicates first\_qualifier\([3]\) and last\_qualifier\([3]\) taking the statement ID, the property and the value and next\_qualifier\([6]\) with ID, property and value for the current and next qualifier are introduced.\\
 
-Using this pre-processed data the approach is to apply a require-predicate to each statement or qualifier.
+Using this pre-processed data, the approach is to apply a require-predicate to each statement or qualifier.
 This is require\([3]\) for statements, connecting the statement ID with the constrained property \(p\) and the required term \(t\) and require\([5]\) taking the statement ID, property, and value of a qualifier and linking this with \(p\) and \(t\).\\
 
-For statements the pattern consists of two rules, both utilising the same conjunction \(c = C_1 \wedge ..
-\wedge C_m\) with arbitrary atoms \(C_n\) which is used to test if the statement does not fulfil the requirement.
+For statements the pattern consists of two rules, both utilising the same conjunction \(c = C_1 \wedge \ldots \wedge C_m\) with atoms \(C_i\) depending on the constraint which is used to test if the statement does not fulfil the requirement.
 Then rule one

### 2018-09-01 thesis/paper/paper.tex
Improved item requires statement

@@
 \usepackage{adjustbox}
+\usepackage[shortlabels]{enumitem}
 
@@
 \subsection{Item requires statement (Q21503247)}\label{subsec_2_item_requires_statement}
-The item requires statement constraint requires an additional statement on an item with a statement using the constrained property, possibly also using a specific value from the list of allowed values.
+\paragraph{Constraint}
+The \emph{item requires statement} constraint requires an additional statement on an item with a statement using the constrained property.
+It can be even more restrictive by also requiring the value of said additional statement to be in a set of \emph{allowed values}.
 \begin{table}[H]
@@
 \hline
-Required data & item statement \\
-\hline
-Qualifiers & \makecell{required property (P2306) -- 1 \\ property the item statement must have \\ allowed value (P2305) -- 0..* \\ values the item statement may have} \\
+Required data & item statements \\
+\hline
+Qualifiers & \makecell{\emph{required property} (P2306) -- 1 \\ property that must be predicate of an item statement \\
+\emph{allowed value} (P2305) -- 0..* \\ value that must be object of the item statement} \\
 \hline
@@
 \hline
-Dependency &  all item statements\\
-\hline
 Rule features & negation \\
@@
 
-An item violates the constraint if it a) has a statement using constrained property as predicate, b1) has no statement using the required property as predicate, or (if allowed values are specified) b2) has only statement using required property as predicate that do not use one of the allowed values as value.\\
-While condition a) can easily be tested by a statement\_EDB-atom with the constrained property, condition b1) and b2) pose the question if an item does not have a statement.
-This is addressed in section [?].
+\paragraph{Rules}
+The \emph{item requires statement} constraint requires inequality between all properties of statements and the required property as well as between all statement values and the allowed values.
+An item violates the constraint if it a) has a statement using constrained property as predicate, b1) has no statement using the \emph{required property} as predicate, or (if \emph{allowed values} are specified) b2) has only statement using the \emph{required property} as predicate that do not use one of the \emph{allowed values} as value.
+The required term \(t\) is the \emph{required property}.
+Without separators the conjunction is
+\begin{equation*}
+\begin{split}
+c_1 = &\text{statement\_EDB}(Q, I, \text{propertyConstant}, X) \wedge{} \\
+&\text{statement\_EDB}(S, I, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
+\end{split}
+\end{equation*} \(\)
+If a set of \emph{allowed values} \(A\) has been specified, a second set of require-rules have to be introduced with
+\begin{equation*}
+\begin{split}
+c_2 = &\text{statement\_EDB}(Q, I, \text{propertyConstant}, X) \wedge{} \\
+&\text{statement\_EDB}(S, I, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
+\end{split}
+\end{equation*}
+Note that the first set of rules is still required, since the preceding statements do not need to have the \emph{required property}. In both cases violations can be found with
+\begin{equation*}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
+&\text{statement\_EDB}(O, I, P, X) \wedge \text{last}(O, I) \wedge{} \\
+&\text{require}(O, \text{propertyConstant}, \text{requiredPropertyConstant})
+\end{split}
+\end{equation*}
 
 \subsection{Value requires statement (Q21510864)}\label{subsec_2_value_requires_statement}
+\paragraph{Constraint}
 The value requires statement constraint requires an additional statement on the value of constrained statement, possibly also using a specific value from the list of allowed values.
@@
 
+\paragraph{Rules}
+The value requires statement constraint is syntactically very similar to the previous item requires statement constraint.
+Instead of requiring a statement on the item of a constrained statement it requires it on the value.
+The pre-processing changes to extracting all value statements along the constrained statements and ordering them.
+Inequality need to be computed between all properties and all values on value statements.
+The conjunctions change in the first triple and are:
+\begin{equation*}
+\begin{split}
+c_1 = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
+&\text{statement\_EDB}(S, I, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
+\end{split}
+\end{equation*} \(\)
+\begin{equation*}
+\begin{split}
+c_2 = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
+&\text{statement\_EDB}(S, I, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
+\end{split}
+\end{equation*}
+The final violation rule would then be:
+\begin{equation*}
+\begin{split}
+\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
+&\text{statement\_EDB}(O, V, P, X) \wedge \text{last}(O, V) \wedge{} \\
+&\text{require}(O, \text{propertyConstant}, \text{requiredPropertyConstant})
+\end{split}
+\end{equation*}
+
 \subsection{Inverse (Q21510855)}\label{subsec_2_inverse}
+\paragraph{Constraint}
 The inverse constraint requires an additional statement on the value of a constrained statement with a specific property and the constrained item as value.
@@
 
+\paragraph{Rules}
+The inverse constraint is similar to the value requires statement constraint in requiring a statement on a constrained value with a specific property.
+However, instead of putting no constraint or a list of fixed values on the required statements value it set a relative requirement: the constrained item, meaning \(t = R\).
+The pre-processing and inequalities thus remain the same, but the conjunctions change to
+\begin{equation*}
+\begin{split}
+c_1 = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
+&\text{statement\_EDB}(S, I, P, V) \wedge{} \\
+&\text{unequal}(\text{inversePropertyConstant}, P)
+\end{split}
+\end{equation*}
+\begin{equation*}
+\begin{split}
+c_2 = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
+&\text{statement\_EDB}(S, I, \text{inversePropertyConstant}, V) \wedge{} \\
+&\text{unequal}(R, V)
+\end{split}
+\end{equation*}
+
 \subsection{Symmetric (Q21510862)}
@@
 
+The symmetric constraint can be translated exactly as the inverse constraint if the inverse property where set to the same as the constrained property.
+
 \subsection{Multi-value (Q21510857)}
+\paragraph{Constraint}
 The multi-value constraint specifies that the constrained property should appear in multiple statements of an item or none.
@@
 
+\paragraph{Rules}
+The multi-value constraint requires a limited count of constrained statements.
+Alternatively, it could be said that if the item requires a second statement, the constraint is violated.
+Pre-processing extracts all items statements and inequality between all item statement properties as usual and the conjunction is
+\begin{equation*}
+\begin{split}
+c = &\text{statement\_EDB}(Q, I, \text{propertyConstant}, X) \wedge{} \\
+&\text{statement\_EDB}(S, I, P, V) \wedge \text{unequal}(\text{propertyConstant}, P)
+\end{split}
+\end{equation*}
+The predicate require\_second\([2]\) is introduced with the same meaning as require except that it will be applied after a required property has been found:
+\begin{equation*}
+\begin{split}
+\text{require\_second}&(S, \text{propertyConstant}, \text{propertyConstant}) \leftarrow \\
+&\text{next}(R, S) \wedge \text{statement\_EDB}(R, I, \text{propertyConstant}, C) \wedge{} \\
+&\text{next}(Q, R) \wedge \text{require}(Q, \text{propertyConstant}, \text{propertyConstant})  \wedge c
+\end{split}
+\end{equation*}
+\begin{equation*}
+\begin{split}
+\text{require\_second}&(S, \text{propertyConstant}, \text{propertyConstant}) \leftarrow \text{next}(R, S) \wedge{} \\
+&\text{require\_second}(R, \text{propertyConstant}, \text{propertyConstant}) \wedge c
+\end{split}
+\end{equation*}
+A violations would be any last statement requiring a second constrained statement.\\
+
 \subsection{Single value (Q19474404)}
+\paragraph{Constraint}
 The single value constraint constrains the values a property may take on a single item.
@@
 
-\subsection{Mandatory qualifier (Q21510856)}
-The mandatory qualifier constraint requires that a constrained statement has a qualifier with a specific property.
-
-\begin{table}[H]
-\caption{Mandatory qualifier constraint (Q21510856)}
-\begin{tabularx}{\textwidth}{ ll X}
-\hline
-Required data & qualifier on constrained statement \\
-\hline
-Qualifiers & \makecell{required property (P2306) -- 1 \\ qualifier that the statement must have} \\
-\hline
-Violation & \makecell{constrained statement without a qualifier with the required property} \\
-\hline
-Dependency & constrained statement and its qualifiers \\
-\hline
-Rule features & negation \\
-\hline
-\end{tabularx}
-\end{table}
-
-To find violations of this constraint one would find a statement\_EDB with the constrained property, but where no qualifier\_EDB for the statement and with the required property exists.
-
-\subsection{Single best value (Q52060874)}
-The single best value constraint requires that only one constrained statement per item has the rank preferred.
-
-\begin{table}[H]
-\caption{Single best value constraint (Q52060874)}
-\begin{tabularx}{\textwidth}{ ll X}
-\hline
-Required data & item statement rank \\
-\hline
-Qualifiers & \makecell{separator (P4155) -- 0..* \\ qualifiers that must be different for exception} \\
-\hline
-Violation & \makecell{items with two constrained statements with rank preferred} \\
-\hline
-Dependency & all item statements and their rank \\
-\hline
-Rule features & inequality \\
-\hline
-\end{tabularx}
-\end{table}
-
-Modelling the constraint means a count of all constrained statements with rank preferred is necessary, with a violation constituted if an item has a count of two or higher.
-Since it is only needed to differentiate between one and many, the rules could be written explicitly.
-
-\subsection{Type (Q21503250)}\label{subsec_2_type}
-The type constraint requires the existence of either one of a specific statement (instanceOf <class>) or a chain of statements using subclassOf to connect directly (subclassOf* <class>)
-or via instanceOf (instanceOf/subclassOf* <class>) for all classes on the constrained item.
-\begin{table}[H]
-\caption{type constraint (Q21503250)}
-\begin{tabularx}{\textwidth}{ ll X}
-\hline
-Required data & chain of statements \\
-\hline
-Qualifiers & \makecell{relation (P2309) -- 1 \\ must be instance of (Q21503252), subclass of (Q21514624), \\ or instance or subclass of (Q30208840); \\ specifies the relation the constrained item must have to the classes\\
-class (P2308) -- 1..* \\ entities that must be reached using the relation} \\
-\hline
-Violation & constrained item without relation to a class \\
-\hline
-Dependency & \makecell{statements of constrained items using instanceOf (P31) or subclassOf(P279) \\ all statements using subclassOf (P279) }\\
-\hline
-Rule features & inequality, negation \\
-\hline
-\end{tabularx}
-\end{table}
-
-If the relation is instanceOf, the constraint can be evaluated like multiple item requires statement constraints (see \ref{subsec_2_item_requires_statement}) by setting the required property to instanceOf (P31) and, for each created item requires statement constraint, the allowed value to one of the classes.
-\\
-For the relations involving subclasses all transitive subclass relationships would be concluded by setting a subclass relationship if a third item can be found that is subclass of one and subclassed by the other item.
-This would be extended with one step of instanceOf.
-A violation would then be any constrained item that is not subclass of any of the classes.
-
-\subsection{Value type (Q21510865)}
-The value type constraint works like the type constraint (see \ref{subsec_2_type}) except the statement chains are required on the value of the constrained statement instead of the constrained item.
-It will thus not be expanded upon.
-
-\section{Reasoning for unconsidered constraints}
-Multiple constraints have not been described above for different reasons.
-The most prominent of these is the format constraint (Q21502404) which specifies a regular expressions the values of the constrained property must adhere to.
-Evaluating these expressions would be theoretically possible but \\
-a) primarily results in a large amount of preprocessing and\\
-b) does not solve the question of negation, since violations would be values that do not conform to their required regular expression. \\
-
-The integer constraint (Q52848401), which requires values of a constrained property be integers (notably a quantity without decimal places), could easily be seen as a specific format constraint.
-While precomputing if a value is an integer or not could be done, it would mean that no actually relevant rules would be needed for the evaluation in datalog.
-\\
-
-Another problem that could not be sensibly solved is the comparison of values, needed for range (Q21510860), difference within range (Q21510854) and contemporary constraint (Q25796498).
-Range restricts the values of a constrained property within a certain range, difference within range specifies that two item statements should have values within a certain range and contemporary constraint specifies that subject and object of a constrained statement have to have coexisted at some point in time.
-While it would be possible to extract all values that might need to be compared and precompute the comparison, it would make most of the work precomputing and leave little to be done in datalog.\\
-
-Link constraint (Q54554025), requiring that constraint statements must have one or more references, requires a negation that cannot be concluded except by explicitly precomputing it, which makes the evaluation unnecessary.
-\\
-A similar situation is the no bounds constraint (Q51723761), which requires that quantity values have neither upper nor lower bound.
-While it would be possible to precompute if a statement has no upper or lower bound, this would shift basically all evaluation logic into precomputing.
-
-And finally commons link constraint (Q21510852) requires that values for a property should be valid links to to Wikimedia Commons.
-This validity cannot be testes from within datalog since there is no way to query the web for information.
-
-\section{Statement non-existence}\label{sec_statement_non-existence}
-
-\subsection{Item requires statement (Q21503247)}
-The item requires statement constraint requires the pre-processing of all item statements.
-Additionally, inequality between all properties of item statements and the required property as well as between all values and the allowed values has to be established.
-It occurs in two varieties, depending on if allowed values have been specified or not.
-The required term \(t\) is the  required property.
-Without separators the conjunction is
-\begin{equation*}
-\begin{split}
-c_1 = &\text{statement\_EDB}(Q, I, \text{propertyConstant}, X) \wedge{} \\
-&\text{statement\_EDB}(S, I, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
-\end{split}
-\end{equation*} \(\)
-If a set of allowed values \(A\) has been specified, a second set of require-rules have to be introduced with
-\begin{equation*}
-\begin{split}
-c_2 = &\text{statement\_EDB}(Q, I, \text{propertyConstant}, X) \wedge{} \\
-&\text{statement\_EDB}(S, I, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
-\end{split}
-\end{equation*}
-Note that the first set of rules is still required, since the preceding statements do not need to have the required property. In both cases violations can be found with
-\begin{equation*}
-\begin{split}
-\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
-&\text{statement\_EDB}(O, I, P, X) \wedge \text{last}(O, I) \wedge{} \\
-&\text{require}(O, \text{propertyConstant}, \text{requiredPropertyConstant})
-\end{split}
-\end{equation*}
-
-\subsection{Value requires statement (Q21510864)}
-The value requires statement constraint is syntactically very similar to the previous item requires statement constraint.
-Instead of requiring a statement on the item of a constrained statement it requires it on the value.
-The pre-processing changes to extracting all value statements along the constrained statements and ordering them.
-Inequality need to be computed between all properties and all values on value statements.
-The conjunctions change in the first triple and are:
-\begin{equation*}
-\begin{split}
-c_1 = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
-&\text{statement\_EDB}(S, I, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
-\end{split}
-\end{equation*} \(\)
-\begin{equation*}
-\begin{split}
-c_2 = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
-&\text{statement\_EDB}(S, I, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
-\end{split}
-\end{equation*}
-The final violation rule would then be:
-\begin{equation*}
-\begin{split}
-\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
-&\text{statement\_EDB}(O, V, P, X) \wedge \text{last}(O, V) \wedge{} \\
-&\text{require}(O, \text{propertyConstant}, \text{requiredPropertyConstant})
-\end{split}
-\end{equation*}
-
-\subsection{Inverse (Q21510855)}
-The inverse constraint is similar to the value requires statement constraint in requiring a statement on a constrained value with a specific property.
-However, instead of putting no constraint or a list of fixed values on the required statements value it set a relative requirement: the constrained item, meaning \(t = R\).
-The pre-processing and inequalities thus remain the same, but the conjunctions change to
-\begin{equation*}
-\begin{split}
-c_1 = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
-&\text{statement\_EDB}(S, I, P, V) \wedge{} \\
-&\text{unequal}(\text{inversePropertyConstant}, P)
-\end{split}
-\end{equation*}
-\begin{equation*}
-\begin{split}
-c_2 = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
-&\text{statement\_EDB}(S, I, \text{inversePropertyConstant}, V) \wedge{} \\
-&\text{unequal}(R, V)
-\end{split}
-\end{equation*}
-
-\subsection{Symmetric (Q21510862)}
-The symmetric constraint can be translated exactly as the inverse constraint if the inverse property where set to the same as the constrained property.
-
-\subsection{Multi-value (Q21510857)}
-The multi-value constraint requires a limited count of constrained statements.
-Alternatively, it could be said that if the item requires a second statement, the constraint is violated.
-Pre-processing extracts all items statements and inequality between all item statement properties as usual and the conjunction is
-\begin{equation*}
-\begin{split}
-c = &\text{statement\_EDB}(Q, I, \text{propertyConstant}, X) \wedge{} \\
-&\text{statement\_EDB}(S, I, P, V) \wedge \text{unequal}(\text{propertyConstant}, P)
-\end{split}
-\end{equation*}
-The predicate require\_second\([2]\) is introduced with the same meaning as require except that it will be applied after a required property has been found:
-\begin{equation*}
-\begin{split}
-\text{require\_second}&(S, \text{propertyConstant}, \text{propertyConstant}) \leftarrow \\
-&\text{next}(R, S) \wedge \text{statement\_EDB}(R, I, \text{propertyConstant}, C) \wedge{} \\
-&\text{next}(Q, R) \wedge \text{require}(Q, \text{propertyConstant}, \text{propertyConstant})  \wedge c
-\end{split}
-\end{equation*}
-\begin{equation*}
-\begin{split}
-\text{require\_second}&(S, \text{propertyConstant}, \text{propertyConstant}) \leftarrow \text{next}(R, S) \wedge{} \\
-&\text{require\_second}(R, \text{propertyConstant}, \text{propertyConstant}) \wedge c
-\end{split}
-\end{equation*}
-A violations would be any last statement requiring a second constrained statement.\\
-
-\subsection{Single value (Q19474404)}\label{subsec_3_single_value}
+\paragraph{Rules}
 The single value constraint can be divided into two different cases, depending on if the set of separators \(S\) is empty or not.
@@
 \subsection{Mandatory qualifier (Q21510856)}
+\paragraph{Constraint}
+The mandatory qualifier constraint requires that a constrained statement has a qualifier with a specific property.
+
+\begin{table}[H]
+\caption{Mandatory qualifier constraint (Q21510856)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Required data & qualifier on constrained statement \\
+\hline
+Qualifiers & \makecell{required property (P2306) -- 1 \\ qualifier that the statement must have} \\
+\hline
+Violation & \makecell{constrained statement without a qualifier with the required property} \\
+\hline
+Dependency & constrained statement and its qualifiers \\
+\hline
+Rule features & negation \\
+\hline
+\end{tabularx}
+\end{table}
+
+To find violations of this constraint one would find a statement\_EDB with the constrained property, but where no qualifier\_EDB for the statement and with the required property exists.
+
+\paragraph{Rules}
 The mandatory qualifier constraint needs to determine qualifier non-existence on constrained statements to find violations, which is similar to section \ref{subsubsec_with_separators}.
@@
 \subsection{Single best value (Q52060874)}
+\paragraph{Constraint}
+The single best value constraint requires that only one constrained statement per item has the rank preferred.
+
+\begin{table}[H]
+\caption{Single best value constraint (Q52060874)}
+\begin{tabularx}{\textwidth}{ ll X}
+\hline
+Required data & item statement rank \\
+\hline
+Qualifiers & \makecell{separator (P4155) -- 0..* \\ qualifiers that must be different for exception} \\
+\hline
+Violation & \makecell{items with two constrained statements with rank preferred} \\
+\hline
+Dependency & all item statements and their rank \\
+\hline
+Rule features & inequality \\
+\hline
+\end{tabularx}
+\end{table}
+
+Modelling the constraint means a count of all constrained statements with rank preferred is necessary, with a violation constituted if an item has a count of two or higher.
+Since it is only needed to differentiate between one and many, the rules could be written explicitly.
+
+\paragraph{Rules}
 The single best value constraint, like the single value constraint (section \ref{subsec_3_single_value}), differentiates between the set of separators \(S\) being empty or not.
@@
 
-\subsection{Type (Q21503250)}
-\todo[inline]{Well.
-I wouldn't know.
-I'd have to expand the require pattern to search up the subclass tree and if require arrives at the root the class is not there.
-But a) its not a tree, so that does not work and b) there is no single root.\\
-Alternatively I could compute each items class via transitivity aaaand....
-well.
-Violation if items classes unequal required class, but how to write that?
-Well, maybe I'll get a brilliant idea.
-Or I'll stop being stupid in case this does not require a brilliant idea.}
-
-\subsection{Value type (Q21510865)}
-\todo[]{as above, so below. no wait...}
+\section{Reasoning for unconsidered constraints}
+Multiple constraints have not been described above for different reasons.
+The most prominent of these is the \emph{format} constraint (Q21502404) which specifies a regular expressions the values of the constrained property must adhere to.
+Evaluating these expressions would be theoretically possible but \\
+\begin{enumerate}[a)]
+\item primarily results in a large amount of preprocessing and
+\item does not solve the question of negation, since violations would be values that do not conform to their required regular expression
+\end{enumerate}
+
+The two type constraints \emph{type} (Q21503250) and \emph{value type} (Q21510865) require an entity to have a certain type.
+This is defined transitively via the properties \emph{instance of} (P31) and \emph{subclass of} (P279), which have to form a statement chain to the entity given as the type.
+Evaluating these constraints would require pre-processing to compute all types of an entity transitively, add these as statements to the data and establish their order to allow testing for the non-existence of a type.
+Since the constraint could easily be checked when computing the types of an entity (is the required type in the types it has), this does not make sense in datalog.
+
+The \emph{integer} constraint (Q52848401), which requires values of a constrained property be integers (notably a quantity without decimal places), could easily be seen as a specific format constraint.
+While precomputing if a value is an integer or not could be done, it would mean that no actually relevant rules would be needed for the evaluation in datalog.
+
+Another problem that could not be sensibly solved is the comparison of values, needed for \emph{range} (Q21510860), \emph{difference within range} (Q21510854) and \emph{contemporary} constraint (Q25796498).
+\emph{Range} restricts the values of a constrained property within a certain range, \emph{difference within range} specifies that the values of two item statements should be within a certain range and \emph{contemporary} constraint specifies that subject and object of a constrained statement have to have coexisted at some point in time.
+While it would be possible to extract all values that might need to be compared and precompute the comparison, it would make most of the work precomputing and leave little to be done in datalog.
+
+\emph{Link} constraint (Q54554025), stating that constraint statements must have one or more references, requires a negation that cannot be concluded except by explicitly precomputing it, which makes the evaluation unnecessary.
+\\
+A similar situation is the \emph{no bounds} constraint (Q51723761), which requires that quantity values have neither upper nor lower bound.
+While it would be possible to precompute if a statement has no upper or lower bound, this would shift basically all evaluation logic into precomputing.
+
+And finally \emph{commons link} constraint (Q21510852) requires that values for a property should be valid links to to Wikimedia Commons.
+This validity cannot be tested from within datalog since there is no way to query the web for information.
 

### 2018-09-01 thesis/paper/paper.tex
Improved value requires statement

@@
 \begin{table}[H]
-\caption{item requires statement constraint (Q21503247)}
+\caption{item requires statement constraint (Q21503247)}\label{tab_item_requires_statements}
 \begin{tabularx}{\textwidth}{ ll X}
@@
 \paragraph{Rules}
-The \emph{item requires statement} constraint requires inequality between all properties of statements and the required property as well as between all statement values and the allowed values.
+The \emph{item requires statement} constraint requires inequality between all properties of statements and the \emph{required property} as well as between all statement values and the \emph{allowed values}.
+The order of the item statements needs to be established.
 An item violates the constraint if it a) has a statement using constrained property as predicate, b1) has no statement using the \emph{required property} as predicate, or (if \emph{allowed values} are specified) b2) has only statement using the \emph{required property} as predicate that do not use one of the \emph{allowed values} as value.
@@
 \paragraph{Constraint}
-The value requires statement constraint requires an additional statement on the value of constrained statement, possibly also using a specific value from the list of allowed values.
-Since the constraint works very similar to item requires statement (see \ref{subsec_2_item_requires_statement}) --- requiring a statement on the value of the constrained statement instead of on the item --- it is not elaborated on.
-
-\paragraph{Rules}
-The value requires statement constraint is syntactically very similar to the previous item requires statement constraint.
+The \emph{value requires statement} constraint requires an additional statement on the value of constrained statement, possibly also using a specific value from the list of allowed values.
+Since the constraint works very similar to item requires statement (see \ref{subsec_2_item_requires_statement}) --- requiring a statement on the value of the constrained statement instead of on the item --- Table \ref{tab_item_requires_statements} is not repeated.
+
+\paragraph{Rules}
+The \emph{value requires statement} constraint is syntactically very similar to the previous \emph{item requires statement} constraint.
 Instead of requiring a statement on the item of a constrained statement it requires it on the value.
-The pre-processing changes to extracting all value statements along the constrained statements and ordering them.
-Inequality need to be computed between all properties and all values on value statements.
-The conjunctions change in the first triple and are:
+Inequality need to be computed between all properties and all values and the item statement order established.
+The conjunctions change in the first statement and are:
 \begin{equation*}

### 2018-09-01 thesis/paper/paper.tex
Improved inverse

@@
 \paragraph{Constraint}
-The inverse constraint requires an additional statement on the value of a constrained statement with a specific property and the constrained item as value.
-It is very similar to value requires statement (see \ref{subsec_2_value_requires_statement}), although it does not specify allowed values, since the required statement has one specific value: the constrained item.
-While this means it could not be modelled exactly as a value requires statement constraint, it would only require very small rule changes and will thus not be elaborated on.
-
-\paragraph{Rules}
-The inverse constraint is similar to the value requires statement constraint in requiring a statement on a constrained value with a specific property.
-However, instead of putting no constraint or a list of fixed values on the required statements value it set a relative requirement: the constrained item, meaning \(t = R\).
-The pre-processing and inequalities thus remain the same, but the conjunctions change to
+The \emph{inverse} constraint requires an additional statement on the value of a constrained statement with a \emph{required property} and the constrained item as value.
+It is very similar to \emph{value requires statement} (see \ref{subsec_2_value_requires_statement}), although it does not specify allowed values, since the required statement needs to have one specific value: the item of the constrained statement.
+
+\paragraph{Rules}
+The \emph{inverse} constraint can be modelled in almost the same way as the \emph{value requires statement} constraint.
+The only change is that instead of requiring inequality to all \emph{allowed values} in conjunction \(c_2\) the statement requires this constraint if its value is not the subject \(R\) of the original constrained statement, meaning \(t = R\).
+Inequalities have to be established between all properties and the \emph{inverse} property as well as all subjects and values of statements.
 \begin{equation*}
@@
 \end{equation*}
+The final violation rule remains the same.
 

### 2018-09-01 thesis/paper/paper.tex
Improved symmetric

@@
 \subsection{Symmetric (Q21510862)}
-The symmetric constraint requires an additional statement on the value of a constrained statement with the constrained property and the constrained item as value.
-It is thus equal to an inverse constraint (see \ref{subsec_2_inverse}) with the same property as constrained property and required property and will not be elaborated on.
-
-The symmetric constraint can be translated exactly as the inverse constraint if the inverse property where set to the same as the constrained property.
+The \emph{symmetric} constraint requires an additional statement on the value of a constrained statement with the constrained property and the constrained item as value.
+It is thus equal to an \emph{inverse} constraint (see \ref{subsec_2_inverse}) with the same property as constrained property and inverse property.
+It can be translated exactly as the inverse constraint if the inverse property where set to the same as the constrained property.
 

### 2018-09-01 thesis/paper/paper.tex
Improved multi-value

@@
 \paragraph{Constraint}
-The multi-value constraint specifies that the constrained property should appear in multiple statements of an item or none.
+The \emph{multi-value} constraint specifies that the constrained property should appear in multiple statements of an item or none.
 
@@
 \hline
-Required data & item statement \\
+Required data & item statements \\
 \hline
@@
 
-To model this constraint a count of statements using the constrained property per item would be needed.
-Then a violation would be any constrained item with a count of one for the constrained property.
-Since there is no need for arbitrarily large counts, but only for the difference between one and two, this could be solved explicitly.
-
-\paragraph{Rules}
-The multi-value constraint requires a limited count of constrained statements.
+\paragraph{Rules}
+The \emph{multi-value} constraint requires a limited count of constrained statements.
 Alternatively, it could be said that if the item requires a second statement, the constraint is violated.
-Pre-processing extracts all items statements and inequality between all item statement properties as usual and the conjunction is
+Inequality is established between all properties of statements and the conjunction is:
 \begin{equation*}

### 2018-09-01 thesis/paper/paper.tex
Improved single value

@@
 \paragraph{Constraint}
-The single value constraint constrains the values a property may take on a single item.
-A property marked with the single value constraint can have properties marked as separators, which define exceptions under specific circumstances.
+The \emph{single value} constraint states that the constrained property may only be used once per item.
+A constrained statement marked with the single value constraint can have \emph{separators}, which define exceptions under specific circumstances.
 \begin{table}[H]
@@
 \hline
-Required data & constrained statement value \\
-\hline
-Qualifiers & \makecell{separator (P4155) -- 0..* \\ qualifiers that must be different for exception} \\
+Required data & constrained statements and their qualifiers, constrained qualifiers and references \\
+\hline
+Qualifiers & \makecell{\emph{separator} (P4155) -- 0..* \\ qualifiers that, when used on constrained statements, can warrant an exception} \\
 \hline
@@
 \hline
-Dependency & \makecell{without separators: all constrained statements on one item \\ with separators: additionally all qualifiers with separators on those statements }\\
-\hline
 Rule features & inequality, partially negation \\
@@
 
-If no separators are specified, two triple form a violation of this constraint if they a) belong to the same item, b) have the same constrained property, and c) have the same value.
-If separators are specified, a violation as above can be ignored if the statements each have at least one qualifier with the same separator as predicate and different values or an unequal number of separator qualifiers.\\
-In the case without separators the modelling idea would be to find two triples statement\_EDB with the same item and constrained property and test if their id is unequal.
-If separators were specified a violation would be the two triples and all their separator qualifiers equal or non-existent.
-\todo{Since I found the details via looking at the source code, should/can I cite that?}
+If no separators are specified, two statements form a violation of this constraint if they a) belong to the same item, b) have the same constrained property, and c) have the same value.
+If separators are specified, a violation as above can be ignored if the statements each have at least one qualifier with the same separator as predicate and different values or an unequal number of separator qualifiers.
 
@@
 \end{equation}
-For qualifiers and reference there is no ID whose inequality could be checked.
+For qualifiers and references there is no ID whose inequality could be checked.
 However, since the statement ID and the property of a possible conflict are always the same, inequality can be tested on the values.
@@
 \begin{split}
-\text{violation\_reference}&(S, \text{propertyConstant}, V) \leftarrow \\
+\text{violation\_reference}&(S, H, \text{propertyConstant}, V) \leftarrow \\
 &\text{reference\_EDB}(S, H, \text{propertyConstant}, V) \wedge{} \\
@@
 \end{equation*}
-\todo{Check if all reference-rules look exactly like the qualifier rules and cut them out if it is so.}
 
 \subsubsection{With separators}\label{subsubsec_with_separators}
-To find all violations under separators a number of things have to be considered.
-Only triples are eligible, since qualifiers cannot have further qualifier, and the order of qualifiers on constrained statements needs to be pre-processed as well since qualifier non-existence is relevant.
-The pattern needs to be applied for every separator, meaning the conjunction is
+To find all violations under \emph{separators} a number of things have to be considered.
+Only statements are eligible, since qualifiers cannot have further qualifier.
+The order of qualifiers on constrained statements needs to be provided as well since qualifier non-existence is relevant.
+The pattern needs to be applied for every \emph{separator}, meaning the conjunction is
 \begin{equation*}
@@
 With qualifier non-existence established the violation can be found.
-Two statements form a violation if they are distinct as per equation \ref{eq_no_separators_triple} and additionally fulfil the condition that all qualifiers are either equal or non-existent.
-\(\forall X \subset S, Y = S \setminus X\) the following rule has to be added.
-Note that \(\text{predicate}(\{X\})\) denotes one atom for each element \(x \in X\) as introduced in section \ref{subsec_3_allowed_units}.
-With the predicate does\_not\_have\([2]\) linking a statement ID to the separator it does not have via
-\begin{equation*}
-\begin{split}
-\text{does\_not\_have}(S, R) \leftarrow &\text{statement\_EDB}(S, I, \text{propertyConstant}, C) \wedge{} \\
-&\text{last\_qualifier}(S, P, V) \wedge \text{require}(S, P, V, \text{propertyConstant}, R)
-\end{split}
-\end{equation*}
-and the predicate has\_same[2] linking two statement IDs to the separator for which the values are the same via
-\begin{equation*}
-\begin{split}
-\text{has\_same}(S, O, Q) \leftarrow &\text{statement\_EDB}(S, I, \text{propertyConstant}, C) \wedge{} \\
-&\text{statement\_EDB}(O, I, \text{propertyConstant}, X) \wedge{} \\
-&\text{qualifier\_EDB}(S, Q, V) \wedge \text{qualifier\_EDB}(O, Q, V)
-\end{split}
-\end{equation*}
- for brevity the rules are
+Two statements form a violation if they are distinct as per equation \ref{eq_no_separators_triple} and additionally fulfil the condition that all values of \emph{separator} qualifiers are equal or no such \emph{separator} qualifier exist.
+To determine this a number of rules have to be introduced which are based on two sets the \emph{separators} are split into: \(X\) contains the \emph{separators} that exist as qualifiers and \(Y\) the \emph{separators} that do not.
+The following rule is then created for every possible split of the \emph{separators} into two such sets.
+Note that \(\text{predicate}(\{X\})\) denotes \(\bigwedge_{x \in X} \text{predicate}(x)\) as introduced in section \ref{subsec_3_allowed_units}.
 \begin{equation*}
@@
 \end{equation*}
+The predicate does\_not\_have\([2]\) linking a statement ID to the separator it does not have via 
+\begin{equation*}
+\begin{split}
+\text{does\_not\_have}(S, R) \leftarrow &\text{statement\_EDB}(S, I, \text{propertyConstant}, C) \wedge{} \\
+&\text{last\_qualifier}(S, P, V) \wedge \text{require}(S, P, V, \text{propertyConstant}, R)
+\end{split}
+\end{equation*}
+and the predicate has\_same[2] linking two statement IDs to the separator for which the values are the same via
+\begin{equation*}
+\begin{split}
+\text{has\_same}(S, O, Q) \leftarrow &\text{statement\_EDB}(S, I, \text{propertyConstant}, C) \wedge{} \\
+&\text{statement\_EDB}(O, I, \text{propertyConstant}, X) \wedge{} \\
+&\text{qualifier\_EDB}(S, Q, V) \wedge \text{qualifier\_EDB}(O, Q, V)
+\end{split}
+\end{equation*}
+are introduced for brevity.
 

### 2018-09-01 thesis/paper/paper.tex
Improved mandatory qualifier

@@
 This is done via a statement ID, which is assigned to every triple storing information about an item.
-Every statement can be referenced by qualifiers, which consist of the corresponding statement id, a predicate and an object.
+Statements have one of three ranks, \emph{normal}, \emph{preferred} and \emph{deprecated} which are used to differentiate similar statements on the same item.
+For example, the \emph{preferred} rank would be given to the current \emph{head of state} (P35) of a country, while all statements denoting former \emph{heads of state} have the rank \emph{normal}.
+\emph{Deprecated} is used on statements that where once considered correct but are not anymore, e.g., \emph{Chelsea Manning} (Q298423) having the \emph{sex or gender} (P21) \emph{male} (Q6581097).
+
+Every statement can have added qualifiers, which consist of the corresponding statement id, a predicate and an object.
 Qualifiers provide contextual information for statements.
 The most common context is the time frame in which the statement was accurate, e.g.
-stating the start and end times of a head of state, but other information is possible, e.g.
+stating the \emph{start time} (P580) and \emph{end time} (P582) of the term of a head of state, but other information is possible, e.g.
 determination method for measurements or the elevation above sea level for the highest point.
 Additionally, references can be added to statements.
-References support the claim by citing the source and possibly additional information, e.g.
-when it was retrieved or when the source was published as property-object pairs.
+
+References support the claim by citing the source and additional information, e.g.
+when it was retrieved or when the source was published as property-value pairs.
+Each references consists from multiple such pairs, and statements can have multiple references.
 
 Other information regarding items like labels and descriptions in different languages as well as aliases and site links to other Wikimedia projects are also part of the data model but are not subject to constraints and are thus not explored further.
+
 \section{Datalog}
@@
 \paragraph{Constraint}
-The mandatory qualifier constraint requires that a constrained statement has a qualifier with a specific property.
+The \emph{mandatory qualifier} constraint requires that a constrained statement has a qualifier with a \emph{required property}.
 
@@
 \hline
-Required data & qualifier on constrained statement \\
-\hline
-Qualifiers & \makecell{required property (P2306) -- 1 \\ qualifier that the statement must have} \\
+Required data & constrained statements and their qualifiers \\
+\hline
+Qualifiers & \makecell{\emph{required property} (P2306) -- 1 \\ property that must appear as predicate of a qualifier on the constrained statement} \\
 \hline
@@
 \hline
-Dependency & constrained statement and its qualifiers \\
-\hline
 Rule features & negation \\
@@
 
-To find violations of this constraint one would find a statement\_EDB with the constrained property, but where no qualifier\_EDB for the statement and with the required property exists.
-
-\paragraph{Rules}
-The mandatory qualifier constraint needs to determine qualifier non-existence on constrained statements to find violations, which is similar to section \ref{subsubsec_with_separators}.
-All constrained statements and their qualifiers have to be extracted and the order of the qualifiers pre-processed.
-Inequality must be established between all qualifier properties and the conjunction is
+\paragraph{Rules}
+The \emph{mandatory qualifier} constraint needs to determine qualifier non-existence on constrained statements to find violations, which is similar to section \ref{subsubsec_with_separators}.
+The qualifier order needs to be provided and inequality must be established between all qualifier properties.
+The conjunction for the require-pattern is:
 \begin{equation*}
@@
 \end{equation*}
-With the usual require-pattern for qualifiers and the does\_not\_have rule added for brevity the violations would be
+With the usual require-pattern for qualifiers and the does\_not\_have rule as introduced in section \ref{subsubsec_with_separators} added for brevity the violations would be found with:
 \begin{equation*}
@@
 \text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge \text{does\_not\_have}(S, \text{requiredPropertyConstant})
+&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge \\
+&\text{does\_not\_have}(S, \text{requiredPropertyConstant})
 \end{split}

### 2018-09-01 thesis/paper/paper.tex
Improved single best value

@@
 \paragraph{Constraint}
-The single best value constraint requires that only one constrained statement per item has the rank preferred.
+The \emph{single best value} constraint requires that only one constrained statement per item has the rank preferred.
 
@@
 \hline
-Required data & item statement rank \\
-\hline
-Qualifiers & \makecell{separator (P4155) -- 0..* \\ qualifiers that must be different for exception} \\
-\hline
-Violation & \makecell{items with two constrained statements with rank preferred} \\
-\hline
-Dependency & all item statements and their rank \\
-\hline
-Rule features & inequality \\
-\hline
-\end{tabularx}
-\end{table}
-
-Modelling the constraint means a count of all constrained statements with rank preferred is necessary, with a violation constituted if an item has a count of two or higher.
-Since it is only needed to differentiate between one and many, the rules could be written explicitly.
-
-\paragraph{Rules}
-The single best value constraint, like the single value constraint (section \ref{subsec_3_single_value}), differentiates between the set of separators \(S\) being empty or not.
-Since the constraint depends on the rank of statements this needs to be extracted during pre-processing and attached to the statement ID with the new predicate rank\([2]\).
+Required data & item statements and ranks \\
+\hline
+Qualifiers & \makecell{\emph{separator} (P4155) -- 0..* \\ qualifiers that, when used on constrained statements, can warrant an exception } \\
+\hline
+Violation & \makecell{items with two constrained statements with rank preferred \\ (exception: statements with separators as qualifiers with different values)} \\
+\hline
+Rule features & inequality, partially negation \\
+\hline
+\end{tabularx}
+\end{table}
+
+\paragraph{Rules}
+The \emph{single best value} constraint, like the \emph{single value} constraint (section \ref{subsec_3_single_value}), differentiates between the set of \emph{separators} \(S\) being empty or not.
+Since the constraint depends on the rank of statements this needs to be provided.
 Qualifiers and references are not affected since they do not have ranks.
-For this constraint all item statements, their rank and their qualifiers need to be extracted and inequality established between the statement IDs and the qualifier values.\\
+For this constraint all inequality needs to be established between the statement IDs and the qualifier values as well as the order of the qualifiers given.\\
 
 \subsubsection{No separators}
-If the separator set is empty, the violation simply occurs if there are two constrained statements on the same item with the preferred rank.
-Thus the equation \ref{eq_no_separators_triple} is expanded to
+If the \emph{separator} set is empty, the violation simply occurs if there are two constrained statements on the same item with the preferred rank.
+Thus the rule \eqref{eq_no_separators_triple} is expanded to
 \begin{equation*}
@@
 &\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
-&&\text{rank\_EDB}(S, \text{preferredRankConstant}) \wedge{} \\
+&\text{rank\_EDB}(S, \text{preferredRankConstant}) \wedge{} \\
 &\text{statement\_EDB}(O, I, \text{propertyConstant}, X) \wedge{} \\
-&&\text{rank\_EDB}(O, \text{preferredRankConstant}) \wedge{} \\
+&\text{rank\_EDB}(O, \text{preferredRankConstant}) \wedge{} \\
 &\text{unequal}(S, O)
@@
 \subsubsection{With separators}
-As in the case above, the qualifier non-existence and violation rules are very similar to single value. The only change is the addition of the rank requirements as shown below.
+As in the case above, the qualifier non-existence and violation rules are very similar to the \emph{single value} constraint. The only change is the addition of the rank requirements as shown below.
 \begin{equation*}
@@
 &\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
-&&\text{rank\_EDB}(S, \text{preferredRankConstant}) \wedge{} \\
+&\text{rank\_EDB}(S, \text{preferredRankConstant}) \wedge{} \\
 &\text{statement\_EDB}(O, I, \text{propertyConstant}, W) \wedge{} \\
-&&\text{rank\_EDB}(O, \text{preferredRankConstant}) \wedge{} \\
+&\text{rank\_EDB}(O, \text{preferredRankConstant}) \wedge{} \\
 &\text{unequal}(S, O) \wedge{} \\

### 2018-09-01 thesis/paper/paper.tex
Minor improvements

@@
 This can mean simple restrictions on the statement a property is part of, e.g.
-that only certain values are allowed, or more complex requirements on connected parts of the data graph, e.g.
+that only certain values are allowed, or more complex requirements on connected parts of the data, e.g.
 that the value of a statement with this property should have certain statements.
-In Table \ref{tab_property_constraints} you can find the existing property constraints.
-Afterwards the constraints will be explained in detail and a basic approach for their modelling given.
+In Table \ref{tab_property_constraints} you can find the existing property constraints as of September 2018.
+Afterwards the constraints will be explained in detail and their modelling given.
 For constraint that cannot be modelled, reasoning for this assessment is included.
@@
 A further seven require negation of a specific fact which can be achieved with reasonable effort.
-The last ten \todo{maybe type can be done in some way} are not deemed translatable because they require a feature that cannot be reasonably emulated in datalog.
+The last ten are not deemed translatable because they require a feature that cannot be reasonably emulated in datalog.
 The feature mostly needed is a general form of negation that would allow to test for the non-existence of a derived fact.

### 2018-09-01 thesis/paper/paper.tex
Renamed Link to Citation Needed

@@
 53 & Q21510852 & Commons link & Web request \\
-10 & Q54554025 & Link & Negation \\
+10 & Q54554025 & Citation needed & Negation \\
 6 & Q21510854 & Difference within range & Value comparison \\
@@
 
-\emph{Link} constraint (Q54554025), stating that constraint statements must have one or more references, requires a negation that cannot be concluded except by explicitly precomputing it, which makes the evaluation unnecessary.
+\emph{Citation needed} constraint (Q54554025), stating that constraint statements must have one or more references, requires a negation that cannot be concluded except by explicitly precomputing it, which makes the evaluation unnecessary.
 \\

### 2018-09-02 thesis/paper/paper.tex
Described software

@@
 \usepackage[shortlabels]{enumitem}
+\usepackage{graphicx}
 
@@
 
-\subsection{Representation of Wikidata as Facts}
+\subsection{Representation of Wikidata as Facts}\label{subsec_representation}
 To represent the content of Wikidata as facts the following predicates are introduced:
@@
 
-\subsection{Task}
+\subsection{Task}\label{subsec_task}
 Each property constraint will be modelled and evaluated as one set of rules per property per constraint, meaning each violating statement will be of a form similar to
@@
 
-\chapter{Wikidata constraints explained}
+\chapter{Wikidata constraints explained}\label{cha_wikidata_constraints_explained}
 This section will explain all interesting, meaning translatable or partially translatable, currently used property constraints on Wikidata.
@@
 
- \section{Constraints using Statement non-Existence}
+ \section{Constraints using Statement non-Existence}\label{sec_constraints_using_statement_non}
 Several constraints are violated if specific statements do not exist in relation to constrained statements.
@@
 
-\chapter{Evaluation/Comparison}
+\chapter{Evaluation}
+
+To evaluate the rules described in chapter \ref{cha_wikidata_constraints_explained} they have been implemented for VLog4J \cite{vlog4j}.
+VLog4J is a java library base on VLog \cite{vlog}, a Datalog engine utilising a column-based memory layout to allow processing of large datasets on commodity hardware.
+It was first published in 2016\cite{UJK2016} but has been actively developed since then.
+
+The entire content of Wikidata is available in several formats: the aforementioned RDF, JSON and XML. To evaluate the violations, Wikidata needs to be provided in form of atoms as described in 
+section \ref{subsec_representation}. For this purpose the Wikidata Toolkit (WDTK)\cite{wdtk} was used. It provides an iterator over the JSON dumps, allowing the processing of all entities in Wikidata.
+
+The evaluation aims to document how suitable Datalog and VLog in particular are to extract violations from Wikidata.
+The performance impact of the additionally introduced features, inequality (section \ref{sec_constraints_using_inequality}) and statement non-existence (section \ref{sec_constraints_using_statement_non}) is evaluated.
+
+\section{Software structure}
+The evaluation software\cite{wcd} was written in Java and utilises WDTK and VLog4J to evaluate the violation rules on the content of Wikidata.
+Figure \ref{fig_pipeline} illustrates the steps undertaken to evaluate the constraints.
+
+The first step is to transform the content of the Wikidata JSON-Dump, read via WDKT, into facts readable by VLog4J.
+These take the form of comma-separated value files which created for each predicate in section \ref{subsec_representation}, denoted as "Facts" in the diagram.
+
+Based on this set the Inequaliy Helper creates all files necessary to established inequality as described in section \ref{sec_constraints_using_inequality}, based on which approach was chosen.
+
+The rules determining violations are created by the Constraint Checker, which uses the Wikidata Query Service\footnote{\url{query.wikidata.org}} to fetch the current constraint configurations.
+Based on these configurations the rules are created as described in chapter \ref{cha_wikidata_constraints_explained}.
+
+Computing all violations based on the provided (inequality)-facts and violations rules is done per constraint.
+All facts pertaining to the constraint as well as the rules are loaded into VLog, which creates and in-memory database from the provided .csv-files.
+Once this is done, the reasoning engine materializes all possible facts.
+Afterwards, all violations are extracted by querying the database for the queries described in section \ref{subsec_task}.
+
+\begin{figure}
+\includegraphics[width=\linewidth]{images/Pipeline.pdf}
+\caption{Evaluation pipeline}\label{fig_pipeline}
+\end{figure}
 

### 2018-09-04 thesis/paper/paper.tex
Small changes
Continued evaluation

@@
 
-\subsubsection{Demand-driven materialisation}
+\subsubsection{Demand-driven materialisation}\label{subsubsec_demand-driven_materialisation}
 The approaches outlined above will always result in all inequalities between the members of \(U\) being materialised.
@@
 Two constrained statements with the same value can be found via a simple conjunction of two statement\_EDB atoms with the propertyConstant as predicate and a variable \(V\) as value.
-Since this would mark every constrained statement as a violations because it can fulfil both atoms at the same time, inequality of the statement IDs \(S\) and \(O\) must be tested as well.t
+Since this would mark every constrained statement as a violations because it can fulfil both atoms at the same time, inequality of the statement IDs \(S\) and \(O\) must be tested as well.
 \begin{equation*}
@@
 
-The first step is to transform the content of the Wikidata JSON-Dump, read via WDKT, into facts readable by VLog4J.
+At the beginning the content of the Wikidata JSON-Dump, read via WDKT, is transformed into facts readable by VLog4J.
 These take the form of comma-separated value files which created for each predicate in section \ref{subsec_representation}, denoted as "Facts" in the diagram.
@@
 
-\section{Inequality evaluation}\label{sec_inequality_evaluation}
+\section{Results}
+
+The experiments where carried out on a server with a 3.50GHz Intel Xeon CPU, 1 TB SSD and 32 GB assigned RAM running Debian 4.9.
+Multiple sets of sizes ranging from one thousand to one million entities where extracted from the Wikidata JSON-Dump and all constraints evaluated against them.
+
+\subsection{Statistics}
+Several statistics about the executed runs are reported.
+
+{\bf Total Time} is the execution time taken for evaluating one constraint.
+This does not include the data extraction from JSON, nor the generation of the files necessary to establish inequality.
+It does include the time taken to generate the rules as well as to load, reason and query with VLog.
+
+{\bf Runtime materialization} is the time taken by the VLog engine to compute all facts that can be derived from the given input.
+
+{\bf Derivations} counts how many new facts were created during materialization. This will include all found violations but can be more in cases where additional facts were created as helpers.
+
+{\bf Total Rules} states how many rules where created to establish the violations.
+They are split into two subfields: {\bf Demand rules}, which are additional rules created to satisfy the requirements outlined in section \ref{subsubsec_demand-driven_materialisation}, and {\bf Base Rules}, which are the rules created without accounting for the demand-driven scheme.
+
+\subsection{Overview}
+
+Most constraints could be evaluated in under two hours on even the largest dataset.
+The exceptions are \emph{distinct values}, \emph{item requires statement} and \emph{single value}.
+In parts this is expected behaviour, since these three constraints are applied to an order of magnitude more properties than all other properties with the exception of \emph{scope}, which is described in very simple rules.
+However, as seen in Figure \ref{fig_runtime}, the runtimes of these three constraints are multiple orders of magnitude higher than those of other constraints.
+The following sections will thus focus on the comparison of these three constraints to identify bottlenecks and reasons for this huge difference in performance.
+
+\begin{figure}
+\includegraphics[width=\linewidth]{images/runtime100,000.pdf}
+\caption{Runtime materialization for 100,000 entities}\label{fig_runtime}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\linewidth]{images/rules.pdf}
+\caption{Number of rules and demanded rules}\label{fig_rules}
+\end{figure}
+
+\subsection{Inequality evaluation}\label{sec_inequality_evaluation}
+Inequality computation was implemented in three variants outlined in section \ref{sec_constraints_using_inequality}: naive, encoded and demanded.
+Naive and encoded could not terminate for any dataset other than the smallest with 1,000 entities.
+Naive failed because the disk space needed for the precomputed inequality facts was not available.
+Encoded, while improving on the disk space needed, still.
+The runtimes for this dataset, presented in Figure \ref{fig_inequality}, shows a clear improvement when using the demanded approach.
+This shows that the overhead in rules introduced to enable the demanded approach, seen in Figure \ref{fig_rules}, is worth the improvements gained through it.
+
+\begin{figure}
+\includegraphics[width=\linewidth]{images/inequality1000.pdf}
+\caption{Runtime materialization for 1,000 entities (encoded and demanded)}\label{fig_inequality}
+\end{figure}
+
 \end{document}

### 2018-09-04 thesis/paper/paper.tex
Added methodology to constraints explained

@@
 
+The constraints are described based on the Wikidata Property constraints portal\cite{constraintsPortal}, which links to a help page for every constraint.
+These help pages describe the constraint, list the possible or required qualifiers and show the intended usage with an example.
+In some cases the documentation was unclear and inquiries had to be made on the discussion pages of their respective constraints.
+
 Every section describing a constraint will consist of two subsections: the constraint explanation and the rules.

### 2018-09-05 thesis/paper/paper.tex
Continued evaluation

@@
 \begin{align*}
-\text{violation\_triple}(S, I, \text{propertyConstant}, V) &\leftarrow \text{statement\_EDB}(S, I, \text{propertyConstant}, V)\\
+\text{violation\_statement}(S, I, \text{propertyConstant}, V) &\leftarrow \text{statement\_EDB}(S, I, \text{propertyConstant}, V)\\
 \text{violation\_qualifier}(S, \text{propertyConstant}, V) &\leftarrow \text{qualifier\_EDB}(S, \text{propertyConstant}, V)\\
@@
 \begin{split}
-\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow\\
+\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow\\
 &\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{}\\
@@
 \begin{split}
-\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
 &\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge \text{item}(I)
@@
 \begin{split}
-\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
 &\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge \text{property}(I)
@@
 \begin{split}
-\text{violation\_triple}&(S, I, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
+\text{violation\_statement}&(S, I, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
 &\text{statement\_EDB}(S, I, \text{propertyConstant}, \text{forbiddenValueConstant})
@@
 \begin{split}
-\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
 &\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
@@
 \begin{align*}
-\text{violation\_triple}(S, I, \text{propertyConstant}, V) &\leftarrow \\
+\text{violation\_statement}(S, I, \text{propertyConstant}, V) &\leftarrow \\
 \text{statement\_EDB}(S, I, &\text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V) \\
@@
 \begin{split}
-\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
 &\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge \\
@@
 
-{\bf Total Time} is the execution time taken for evaluating one constraint.
-This does not include the data extraction from JSON, nor the generation of the files necessary to establish inequality.
-It does include the time taken to generate the rules as well as to load, reason and query with VLog.
-
 {\bf Runtime materialization} is the time taken by the VLog engine to compute all facts that can be derived from the given input.
 
-{\bf Derivations} counts how many new facts were created during materialization. This will include all found violations but can be more in cases where additional facts were created as helpers.
+{\bf Additional Derivations} counts how many new facts were created during materialization, excluding the violations themselves.
 
@@
 
-Most constraints could be evaluated in under two hours on even the largest dataset.
+Because of a bug in VLog\footnote{\url{https://github.com/knowsys/vlog4j/issues/43}}, the largest dataset that could be completely evaluated contained 100,000 entities.
+With datasets of larger size, more than half of the constraints were affected by this bug, meaning the analysis could not be conducted.
+All further statistics describe the results on the 100,000 entity set unless specifically noted otherwise.
+Since inequality could only be computed in demanded mode (see section  \ref{sec_inequality_evaluation}), all statistics will show the results attained in this mode.
+
+Most constraints could be evaluated in under two hours.
 The exceptions are \emph{distinct values}, \emph{item requires statement} and \emph{single value}.
@@
 The following sections will thus focus on the comparison of these three constraints to identify bottlenecks and reasons for this huge difference in performance.
+They will be marked with a border in every figure.
 
@@
 
+\subsection{Facts}
+Figure \ref{fig_facts} shows the number of facts loaded into VLog per constraint.
+Base facts denote those directly extracted from the JSON-Dump and corresponding to one of the predicates listed in section \ref{subsec_representation}.
+Inequality are the letter\(_i\) facts created to establish inequality as described in section \ref{sec_constraints_using_inequality}.
+Note that in the demanded (and encoded) mode, every entry for which inequality should be established is usually split into multiple letter\(_i\) facts based on its length, which is why the number of inequality facts can be up to ten times the number of base facts.
+Require facts list the order of statements or qualifiers as described in section \ref{sec_constraints_using_statement_non}.
+
+\begin{figure}
+\includegraphics[width=\linewidth]{images/facts100,000.pdf}
+\caption{Number of input facts per constraint}\label{fig_facts}
+\end{figure}
+
+The long runtimes of \emph{distinct values}, \emph{item requires statement} and \emph{single value} cannot be explained through the size of the input data.
+While all of them have more than 10 million input facts, this is also the case for other constraints that do not take nearly as much runtime.
+\emph{One-of}, \emph{symmetric} and \emph{inverse} have even more input facts than the three long-runtime constraints.
+The reason for the performance difference must be found elsewhere.
+
 \subsection{Inequality evaluation}\label{sec_inequality_evaluation}
@@
 Naive failed because the disk space needed for the precomputed inequality facts was not available.
-Encoded, while improving on the disk space needed, still.
+Encoded, while improving on the disk space needed, still failed due to out of memory errors.
+This behaviour is expected, since for an input of size \(n\) \(n^2\) inequalities have to be materialized.
 The runtimes for this dataset, presented in Figure \ref{fig_inequality}, shows a clear improvement when using the demanded approach.
-This shows that the overhead in rules introduced to enable the demanded approach, seen in Figure \ref{fig_rules}, is worth the improvements gained through it.
+This shows that the overhead in rules introduced to enable the demanded approach, seen in Figure \ref{fig_rules}, is worth the improvements in performance gained through it.
 
@@
 
+\subsection{Additional Derivations}
+One possible bottleneck are the additional derivations necessary to identify violations.
+This includes the inequalities, demands for inequalities, requirements to determine statement non-existence as well as other helper rules introduced in chapter \ref{cha_wikidata_constraints_explained}.
+Note that the four simple constraints described in section \ref{sec_simple_constraints} do not generate any additional facts and are excluded from the chart.
+Figure \ref{fig_add_derivations} shows that the three performance-heavy constraints do generate by far the most additional derivations.
+The constraint with the smallest number of additional derivations, \emph{distinct values}, is almost an order of magnitude higher than the next highest, \emph{one-of}.
+
+\begin{figure}
+\includegraphics[width=\linewidth]{images/additionalDerivations100,000.pdf}
+\caption{Additional derivations for 100,000 entities}\label{fig_add_derivations}
+\end{figure}
+
+\chapter{Conclusion}
+
 \end{document}

### 2018-09-06 thesis/paper/paper.tex
More or less finished evaluation

@@
 \matno{3694323}
-\betreuer{Prof. Dr. Markus Krötzsch}
+\betreuer{Prof. Dr. Markus Kr{\"{o}}tzsch}
 \bibfiles{literatur}
@@
 \usepackage{graphicx}
+\usepackage{amsthm}
 
 \setcounter{tocdepth}{2}
+
+\newtheorem{example}{Example}
 
@@
 \end{table}
-\todo[inline]{For globe(URI): JSON data model states "typically refer to a data item on wikidata.org", which I would read as can be any URI}
 
@@
 Other information regarding items like labels and descriptions in different languages as well as aliases and site links to other Wikimedia projects are also part of the data model but are not subject to constraints and are thus not explored further.
+
+\begin{example}
+Figure \ref{fig_example} is an excerpt from the entity page of the clipper \emph{Cutty Sark} (Q171255) \footnote{\url{https://www.wikidata.org/wiki/Q171255}}.
+An entity page lists all statements with the entity as the subject.
+It then groups all statements by their property, in this case \emph{length} (P2043).
+For the length of the Cutty Sark there is only one value, 85.4 metres, but this could include multiple values, e.g., if there were different ways to determine the length.
+This statement is further qualified with \emph{criterion used} (P1013) as the qualifier property, which specifies how the value was determined.
+In this case the qualifier value \emph{length overall} (Q2358152) was taken.
+The statement is backed up by one reference citing the property \emph{reference URL} (P854) pointing to the reference value, which is an URL confirming the length of 85.4 metres.
+
+\begin{figure}[H]
+\includegraphics[width=\linewidth]{images/cutty_sark_length.png}
+\caption{Statement with qualifier and reference about the length of the Cutty Sark}\label{fig_example}
+\end{figure}
+\end{example}
 
@@
 
+\begin{example}
 For the following example, taken from Wikidata, descriptive names will be used instead of the IDs to increase readability.
@@
 \[\text{isInTaxon\_IDB}(X, Y) \leftarrow \text{isInTaxon\_IDB}(X, Z) \wedge \text{isInTaxon\_IDB}(Z, Y)\]
+\end{example}
 
@@
 
-\subsection{Distinct values (Q21502410)}
+\subsection{Distinct values (Q21502410)}\label{subsec_distinct_values}
 \paragraph{Constraint}
@@
 
-\subsection{Item requires statement (Q21503247)}\label{subsec_2_item_requires_statement}
+\subsection{Item requires statement (Q21503247)}\label{subsec_item_requires_statement}
 \paragraph{Constraint}
@@
 
-\subsection{Value requires statement (Q21510864)}\label{subsec_2_value_requires_statement}
+\subsection{Value requires statement (Q21510864)}\label{subsec_value_requires_statement}
 \paragraph{Constraint}
 The \emph{value requires statement} constraint requires an additional statement on the value of constrained statement, possibly also using a specific value from the list of allowed values.
-Since the constraint works very similar to item requires statement (see \ref{subsec_2_item_requires_statement}) --- requiring a statement on the value of the constrained statement instead of on the item --- Table \ref{tab_item_requires_statements} is not repeated.
+Since the constraint works very similar to item requires statement (see \ref{subsec_item_requires_statement}) --- requiring a statement on the value of the constrained statement instead of on the item --- Table \ref{tab_item_requires_statements} is not repeated.
 
@@
 The \emph{inverse} constraint requires an additional statement on the value of a constrained statement with a \emph{required property} and the constrained item as value.
-It is very similar to \emph{value requires statement} (see \ref{subsec_2_value_requires_statement}), although it does not specify allowed values, since the required statement needs to have one specific value: the item of the constrained statement.
+It is very similar to \emph{value requires statement} (see \ref{subsec_value_requires_statement}), although it does not specify allowed values, since the required statement needs to have one specific value: the item of the constrained statement.
 
@@
 
-\subsection{Single value (Q19474404)}
+\subsection{Single value (Q19474404)}\label{subsec_single_value}
 \paragraph{Constraint}
@@
 
-\subsubsection{No separators}
+\subsubsection{No separators}\label{subsubsec_single_value_no_separators}
 If the separator set is empty, the violation simply occurs if there are two constrained statements on the same item. 
@@
 
-\subsubsection{With separators}\label{subsubsec_with_separators}
+\subsubsection{With separators}\label{subsubsec_single_value_with_separators}
 To find all violations under \emph{separators} a number of things have to be considered.
@@
 \paragraph{Rules}
-The \emph{mandatory qualifier} constraint needs to determine qualifier non-existence on constrained statements to find violations, which is similar to section \ref{subsubsec_with_separators}.
+The \emph{mandatory qualifier} constraint needs to determine qualifier non-existence on constrained statements to find violations, which is similar to section \ref{subsubsec_single_value_with_separators}.
 The qualifier order needs to be provided and inequality must be established between all qualifier properties.
@@
 \end{equation*}
-With the usual require-pattern for qualifiers and the does\_not\_have rule as introduced in section \ref{subsubsec_with_separators} added for brevity the violations would be found with:
+With the usual require-pattern for qualifiers and the does\_not\_have rule as introduced in section \ref{subsubsec_single_value_with_separators} added for brevity the violations would be found with:
 \begin{equation*}
@@
 
-\subsection{Single best value (Q52060874)}
+\subsection{Single best value (Q52060874)}\label{subsec_single_best_value}
 \paragraph{Constraint}
@@
 
-\section{Results}
+\section{Results}\label{sec_results}
 
@@
 
-Most constraints could be evaluated in under two hours.
-The exceptions are \emph{distinct values}, \emph{item requires statement} and \emph{single value}.
+Most constraints could be evaluated in under a minute.
+The exceptions are \emph{distinct values} (1.22 hours), \emph{item requires statement} (2.38 hours) and \emph{single value} (4.22 hours).
 In parts this is expected behaviour, since these three constraints are applied to an order of magnitude more properties than all other properties with the exception of \emph{scope}, which is described in very simple rules.
-However, as seen in Figure \ref{fig_runtime}, the runtimes of these three constraints are multiple orders of magnitude higher than those of other constraints.
+However, also seen in Figure \ref{fig_runtime}, the runtimes of these three constraints are multiple orders of magnitude higher than those of other constraints.
 The following sections will thus focus on the comparison of these three constraints to identify bottlenecks and reasons for this huge difference in performance.
@@
 
+\section{Findings}
+Most constraints could be evaluated within a very short timeframe (< 1 minute).
+While the dataset does not include the entire database of Wikidata, which with 50m entities is about 500 times bigger, the runtimes for the entire dataset should still be within a reasonable frame.
+The exception are the three constraints focused on in section \ref{sec_results}.
+As shown in that section, the main difference between \emph{distinct values}, \emph{item requires statement}, and \emph{single value} and other constraints is
+\begin{enumerate}[a)]
+\item a larger amount of constrained properties (with a corresponding larger amount of rules) and
+\item a larger amount of additional derivations
+\end{enumerate}
+The rules for these three constraints will now be analysed and possible reasons for this huge difference in performance outlined.
+
+\subsection{Distinct values}
+While the rules for the \emph{distinct values} constraint (see section \ref{subsec_distinct_values}) are relatively simple in writing, they do require a large amount of processing.
+Distinct values requires the comparison of every constrained statement with every other constrained statement, meaning its evaluation would scale badly due to the quadratic growth of statements to be compared.
+Additionally, the number of inequalities that have to be evaluated according to the demanded approach (see section \ref{sec_constraints_using_statement_non}) grows accordingly.
+Due to the nature of the constraint -- requiring distinct values for a property across all of Wikidata -- this is in all likelihood unavoidable.
+
+\subsection{Item requires statement}\label{subsec_findings_item_requires_statement}
+The \emph{item requires statement} constraint (see section \ref{subsec_item_requires_statement}) creates additional facts about every statement on an entity with a constrained statement (if the constraint was violated on that entity).
+Of interest to note is that the \emph{value requires statement} constraint (see section \ref{subsec_value_requires_statement}), which works almost identically, does not show a performance nearly as bad as \emph{item requires statement}.
+\emph{Value requires statement} has, however, only 244 constrained properties, while \emph{item requires statement} has 4265.
+This shows that the require approach scales badly with increasing number of require rules.
+
+\subsection{Single value}
+The \emph{single value} constraint (see section \ref{subsec_single_value}) has to be analysed with regards to the two cases: without separators (section \ref{subsubsec_single_value_no_separators}) and with separators (section \ref{subsubsec_single_value_with_separators}.
+
+\paragraph{No separators}
+If no separators were given, the rules simply compare every statement on an entity with a constrained statement to said constrained statement.
+Since entity size does not increase with dataset size, this should not be the reason for \emph{single values} bad performance.
+
+\paragraph{With separators}
+The \emph{single value} constraint, like the \emph{item requires statement} constraint, uses a require approach to determine non-existence.
+For \emph{single value} this is qualifier non-existence on constrained statements.
+As seen in section \ref{subsec_findings_item_requires_statement}, this could be a reason for the bad performance of the analysis for this constraint.
+Another source of processing time for \emph{single value} are the has\_same facts, which need to compare all constrained statements on the same item and all their qualifiers.
+Of interest to note is that the \emph{single best value} constraint (see section \ref{subsec_single_best_value}) works in almost the same way but does not suffer from the same performance.
+Again, the number of constrained properties is significantly lower for \emph{single best value} (203) compared to \emph{single value} (2909).
+
+\subsection{Possible adaptations}
+Since the problematic performance of \emph{item requires statement} and \emph{single value} are not visible in \emph{value requires statement} and \emph{single best value}, which are very similar in their rule modelling, a possible adaptation would be to not evaluate all constrained properties for these constraints in bulk but split them into smaller sets, evaluating each set individually to combat the problematic scaling of the require pattern.
+The precomputed datasets could be kept, meaning only the reasoning step would have to be repeated.
+
 \chapter{Conclusion}

### 2018-09-09 thesis/paper/paper.tex
Added examples and removed unneeded rule parts

@@
 \chapter{Introduction}
-Motivation
 
@@
 
+\begin{example}
+The \emph{reference URL} (P854) property is restricted to the \emph{property scope} (P5314) \emph{as reference} (Q54828450), because it is to be used to mark a URL that substantiates a statement.
+\end{example}
+
 \paragraph{Rules}
@@
 
+\begin{example}
+The \emph{author} (P50) constraint has (among others) the \emph{conflicting property} (P2306) \emph{instance of} (P31) with the \emph{conflicting value} (P2305) \emph{Wikimedia disambiguation page} (Q4167410).
+It is set because editors might easily confuse a work of art with a specific title, e.g., \emph{Tempest} (Q86440) by \emph{author} \emph{William Shakespear} (Q692) with the Wikimedia disambiguation page \emph{Tempest} (Q1589529).
+\end{example}
+
 \paragraph{Rules}
@@
 
+\begin{example}
+The \emph{property constraint} (P2302) property has the \emph{allowed entity type} (P2305) \emph{Wikibase property} (Q29934218) because non-properties cannot be subject to a property constraint.
+\end{example}
+
 \paragraph{Rules}
@@
 \end{table}
+
+\begin{example}
+The \emph{license} (P31) property, denoting the license under which a work is released, has the \emph{forbidden value} (P2305) \emph{proprietary software} (Q218616), since it is often confused with \emph{proprietary license} (Q3238057).
+\end{example}
 
@@
 
+\begin{example}
+The \emph{ISBN-13} (P212) property has a \emph{distinct values} constraint since an ISBN is supposed to be a unique identifier for a book and should not be shared between two items.
+\end{example}
+
 \paragraph{Rules}
@@
 As of September 2018 this is not done on any constraint.
+
+\begin{example}
+The \emph{length} (P2043) property has (among others) the \emph{allowed unit} (P2305) \emph{metre} (Q11573).
+This helps avoid confusing it for \emph{duration} (P2047), which denotes the length of time of an event.
+\end{example}
 
@@
 
+\begin{example}
+The \emph{head of state} (P35) property has (among others) the \emph{allowed qualifiers} (P2306) \emph{start time} (P580) and \emph{end time} (P582) properties, used to describe the time frame the person held that office.
+\end{example}
+
 \paragraph{Rules}
@@
 \end{table}
+
+\begin{example}
+The \emph{wheelchair accessibility} (P2846) property, which describes the accessibility of an location or event, has the three \emph{allowed values} (P2305) \emph{wheelchair accessible} (Q24192067), \emph{wheelchair accessible with help} (Q24192068), and \emph{wheelchair inaccessible} (Q24192069).
+\end{example}
 
@@
 
+\begin{example}
+The \emph{flag bearer} (P3022) property, naming the person carrying a flag at a ceremony, has the \emph{qualifier property} (P2306) \emph{of} (P642) with the \emph{qualifier values} (P2305) \emph{opening ceremony} (Q3010369) and \emph{closing ceremony} (Q13454063).
+\end{example}
+
 \paragraph{Rules}
@@
 
+\begin{example}
+The \emph{professorship} (P803) property requires an additional statement with the \emph{required property} (P2306) \emph{occupation} (P106) and the \emph{allowed value} (P2305) \emph{university teacher} (Q1622272).
+\end{example}
+
 \paragraph{Rules}
@@
 \text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
-&\text{statement\_EDB}(O, I, P, X) \wedge \text{last}(O, I) \wedge{} \\
+&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \text{last}(O, I) \wedge{} \\
 &\text{require}(O, \text{propertyConstant}, \text{requiredPropertyConstant})
@@
 Since the constraint works very similar to item requires statement (see \ref{subsec_item_requires_statement}) --- requiring a statement on the value of the constrained statement instead of on the item --- Table \ref{tab_item_requires_statements} is not repeated.
+
+\begin{example}
+The \emph{list of monuments} (P1456) property requires its value to have the \emph{required property} (P2306) \emph{is a list of} (P360).
+\end{example}
 
@@
 \paragraph{Constraint}
-The \emph{inverse} constraint requires an additional statement on the value of a constrained statement with a \emph{required property} and the constrained item as value.
+The \emph{inverse} constraint requires an additional statement on the value of a constrained statement with an \emph{inverse property} and the constrained item as value.
 It is very similar to \emph{value requires statement} (see \ref{subsec_value_requires_statement}), although it does not specify allowed values, since the required statement needs to have one specific value: the item of the constrained statement.
+
+\begin{example}
+The \emph{mother} (P25) property requires values to have a statement with the \emph{inverse property} (P2306) \emph{child} (P40) and the original item as value.
+\end{example}
 
@@
 
+\begin{example}
+The \emph{sibling} (P3373) property requires a statement on the value with the same property and the original item as the value.
+\end{example}
+
 \subsection{Multi-value (Q21510857)}
@@
 \end{table}
+
+\begin{example}
+The \emph{cast member} (P161) property requires an item to have multiple statements using this property, since most plays or films have multiple cast members.
+This is an example for a constraint from which exception can and should be made and is only meant to point to possibly missing data.
+\end{example}
 
@@
 If separators are specified, a violation as above can be ignored if the statements each have at least one qualifier with the same separator as predicate and different values or an unequal number of separator qualifiers.
+
+\begin{example}
+The \emph{capital} (P36) property should only have one single value per item.
+If conflicts exist, it is not a violation of this constraint if one of the values for the \emph{separators} (P4155) \emph{start time} (P580), \emph{point in time} (P585), \emph{end time} (P582), and \emph{determination method} (P459) are unequal.
+\end{example}
 
@@
 
+\begin{example}
+The \emph{head of state} (P35) property has the \emph{required property} (P2306) \emph{start time} (P580), meaning every \emph{head of state} must have assumed that position at some time.
+\end{example}
+
 \paragraph{Rules}
@@
 \end{table}
+
+\begin{example}
+The \emph{property proposal discussion} (P3254) property, denoting the URL of a web page where the creation of a Wikidata property was discussed should only have one best value.
+Since a discussion might have been moved to a different page at some point, the best value should be the page the discussion was last moved to.
+\end{example}
 

### 2018-09-09 src/main/java/impl/PCC/ItemRequiresStatementPCC.java

@@ -75,9 +75,6 @@ public class ItemRequiresStatementPCC extends PropertyConstraintChecker {
 		for (String requiredProperty : configuration.keySet()) {
 			Term requiredPropertyConstant = Utility.makeConstant(requiredProperty);
 			
-			// statementEDB(O, I, P, X)
-			Atom statementEDB_OIPX = Expressions.makeAtom(statementEDB, o, i, p, x);
-			
 			// last(O, I)
 			Atom last_OI = Expressions.makeAtom(last, o, i);
 			
@@ -86,9 +83,9 @@ public class ItemRequiresStatementPCC extends PropertyConstraintChecker {
 			
 			// violation_statement(S, I, propertyConstant, V) :-
 			//	statementEDB(S, I, propertyConstant, V),
-			//	statementEDB(O, I, P, X), last(O, I),
+			//	last(O, I),
 			//	require(O, requiredPropertyConstant)
-			Rule violation = Expressions.makeRule(violation_statement_SIpV, statementEDB_SIpV, statementEDB_OIPX, last_OI, require_Opr);
+			Rule violation = Expressions.makeRule(violation_statement_SIpV, statementEDB_SIpV, last_OI, require_Opr);
 			rules.add(violation);
 		}
 		

### 2018-09-09 thesis/paper/paper.tex
Modified paper with rule part removal

@@
 \text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
-&\text{statement\_EDB}(O, V, P, X) \wedge \text{last}(O, V) \wedge{} \\
+&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \text{last}(O, V) \wedge{} \\
 &\text{require}(O, \text{propertyConstant}, \text{requiredPropertyConstant})

### 2018-09-18 thesis/paper/paper.tex
Added introduction

@@
 \chapter{Introduction}
-
-\chapter{Preliminaries}
-\section{Wikidata}
+Wikidata, the knowledge base sister project of Wikipedia, has grown to be one of the most used open data resources available on the web.
+Wikidata is closely integrated with Wikipedia, where it links pages across languages and provides language-independent content (such as birth dates or population figures).
+Additionally, several services outside of the Wikimedia make use of Wikidata.
+Probably the best-known example is the knowledge graph by Google, shown in information boxes to the right or top of search results on their main page.
+These facts are sourced from a variety of databases, Wikidata among them.
+Openstreetmap links objects with a physical presence to its Wikidata pages (and vice versa).
+\todo{more examples, if they can be found}
+
+As of September 2018, Wikidata contains more than 50 million pages with over 500 million statements.
+More than 34.000 contributors worked on Wikidata in August 2018 alone, and more than 18.4 million facts where edited or created.
+Multiple open data collections are linked through Wikidata, such as VIAF, DOI, or BBC Programmes.
+More on the structure of Wikidata is explained in section \ref{sec_wikidata}.
+
+Because of Wikidatas widespread usage and frequent editing, it is necessary to ensure the correctness and quality of the data.
+Mechanisms to support data quality are, among others, referencing the sources of statements, cross-checking by the community, and constraints.
+
+Constraints limit how the content of Wikidata may related to itself.
+Some facts can be in conflict with others, meaning one of the two is in all likelihood wrong.
+Other facts can require certain facts to be stated in the database as well, thus pointing to data that may be missing.
+Constraints are used to ensure data correctness as well as clarify the data structure in cases where it might be unclear how facts should be described.
+A detailed explanation of these constrains can be found
+in chapter \ref{cha_property_constraints}.
+
+Since Wikidata is a very large and constantly changing resource, it is necessary to find ways to quickly and efficiently determine violations of these constraints.
+One solution offering itself are rule engines that allow conclusions about a given set of facts (the data in Wikidata) using specific rules (the constraints).
+To lay the foundation for this work, chapter \ref{cha_preliminaries} will explain the basics of Wikidata and Datalog.
+The property constrains are explained in general in chapter \ref{cha_property_constraints} and each individually in chapter \ref{cha_wikidata_constraints_explained}.
+In this work a Datalog-engine, VLog, was chosen because its column-based storage mechanism should be particularly suited to Wikidata.
+The evaluation of this implementation is found in chaper \ref{cha_evaluation}.
+
+\chapter{Preliminaries}\label{cha_preliminaries}
+\section{Wikidata}\label{sec_wikidata}
 Wikidata is the knowledge base sister project of Wikipedia.
@@
 
-\chapter{Property Constraints}
+\chapter{Property Constraints}\label{cha_property_constraints}
 
@@
 
-\chapter{Evaluation}
+\chapter{Evaluation}\label{cha_evaluation}
 
@@
 
+\subsection{Adaptations to VLog}
+
+\subsubsection{Sorting facts for column-based storage}
+VLog uses a column-based storage format where the first column is fully sorted and all following columns are sorted w.r.t.
+the preceding column.
+These columns are then run-length encoded.
+To minimize memory requirements and maximize access speed, predicates should be sorted in such a way that each entry has a smaller number of unique constants than the next entry.
+
+\begin{example}
+The predicate \emph{statement\_EDB} lists the \emph{statement ID}, \emph{entity ID}, \emph{property ID} and \emph{value}.
+To optimize this for VLog, a new order must be created.
+\emph{property ID} should be the first entry, because there are only \todo{number of properties and link} different properties on Wikidata as of September 2018.
+The \emph{statement ID} needs to be moved to the last position since it is unique for all statements, meaning for any data source with \(n\) statements there would be \(n\) unique \emph{statement IDs}.
+Finally \emph{entity ID} should be before \emph{value}, because every \emph{entity ID} could also be a \emph{value}, but value can take additional string values.
+
+The final, optimized order is then: \emph{property ID}, \emph{entity ID}, \emph{value}, \emph{statement ID}.
+\end{example}
+
+This order is, however, not easily readable and, if taken as presented, would likely cause implementation errors because the different positions are confused for each other.
+To prevent this the datasets imported from Wikidata and created during preprocessing as well as the rules outlined in chapter \ref{cha_wikidata_constraints_explained} are sorted afterwards.
+Thus the rules can be kept in a readable form while still optimizing for VLog.
+
+\subsubsection{Sorting conjunctions for left-to-right evaluation}
+Conjunctions in VLog are evaluated from left to right, meaning the engine computes the matches of the first atom and, based on those findings, proceeds with the next atom.
+The rules as presented are not optimal regarding this process.
+Ideally, atoms should be sorted by the size of the intermediate results they incur.
+
+\subsubsection{Results}
+
 \chapter{Conclusion}

### 2018-09-18 thesis/paper/paper.tex
Captialised chapters and sections

@@
 Multiple open data collections are linked through Wikidata, such as VIAF, DOI, or BBC Programmes.
-More on the structure of Wikidata is explained in section \ref{sec_wikidata}.
+More on the structure of Wikidata is explained in Section \ref{sec_wikidata}.
 
@@
 A detailed explanation of these constrains can be found
-in chapter \ref{cha_property_constraints}.
+in Chapter \ref{cha_property_constraints}.
 
@@
 One solution offering itself are rule engines that allow conclusions about a given set of facts (the data in Wikidata) using specific rules (the constraints).
-To lay the foundation for this work, chapter \ref{cha_preliminaries} will explain the basics of Wikidata and Datalog.
-The property constrains are explained in general in chapter \ref{cha_property_constraints} and each individually in chapter \ref{cha_wikidata_constraints_explained}.
-In this work a Datalog-engine, VLog, was chosen because its column-based storage mechanism should be particularly suited to Wikidata.
+To lay the foundation for this work, Chapter \ref{cha_preliminaries} will explain the basics of Wikidata and Datalog.
+The property constrains are explained in general in Chapter \ref{cha_property_constraints} and each individually in Chapter \ref{cha_wikidata_constraints_explained}.
+In this work a Datalog-engine, VLog, was chosen because its column-based storage mechanism seems particularly suited to Wikidata.
 The evaluation of this implementation is found in chaper \ref{cha_evaluation}.
@@
 Six can be solved if constant inequality is defined.
-This can be easily solved with brute force but will be improved in section \ref{sec_inequality}.
+This can be easily solved with brute force but will be improved in Section \ref{sec_inequality}.
 A further seven require negation of a specific fact which can be achieved with reasonable effort.
@@
 \emph{scope} and \emph{allowed entity types} work very similar, since there are only three respectively two possible values and violations are simply the options that have not been allowed.
-The rules for these constraints are described in section \ref{sec_simple_constraints}.\\
+The rules for these constraints are described in Section \ref{sec_simple_constraints}.\\
 
@@
 To avoid putting and undue amount of processing outside of datalog this extraction will not check more than that the constant is in the right position relative to the constrained statement, e.g., all values on statements constrained by \emph{distinct value} or all qualifiers on statements constrained by \emph{allowed qualifiers}.
-From this set inequality will be generated, different versions of which will be introduced in section \ref{sec_inequality} and evaluated with regards to their performance in section \ref{sec_inequality_evaluation}.\\
+From this set inequality will be generated, different versions of which will be introduced in Section \ref{sec_inequality} and evaluated with regards to their performance in Section \ref{sec_inequality_evaluation}.\\
 
@@
 The idea is to ask the question if a second statement exists, then evaluating it as above.
-Both methods will be explained in section \ref{sec_statement_non-existence}.
+Both methods will be explained in Section \ref{sec_statement_non-existence}.
 
@@
 \chapter{Wikidata constraints explained}\label{cha_wikidata_constraints_explained}
-This section will explain all interesting, meaning translatable or partially translatable, currently used property constraints on Wikidata.
+This Section will explain all interesting, meaning translatable or partially translatable, currently used property constraints on Wikidata.
 They will be presented roughly in descending order of the number of properties with this constraint, although constraints with similar function will be explained directly after one another.
@@
 
-Every section describing a constraint will consist of two subsections: the constraint explanation and the rules.
+Every Section describing a constraint will consist of two subsections: the constraint explanation and the rules.
 
@@
 \section{Simple Constraints}\label{sec_simple_constraints}
-This section contains constraints that do not require any additional features and can be translated into rules directly based on the facts introduced from Wikidata.
+This Section contains constraints that do not require any additional features and can be translated into rules directly based on the facts introduced from Wikidata.
 
@@
 \section{Constraints using Inequality}\label{sec_constraints_using_inequality}
-All translatable constraints except the four covered in the previous section require inequality in some way or another.
+All translatable constraints except the four covered in the previous Section require inequality in some way or another.
 This means it is necessary to extract additional information during pre-processing, namely the constants for which inequality needs to be stated.
@@
 One way of determining inequality would be to take advantage of the fact that inequality is established in a set of unequal constants \(U\), which can be represented as a string.
-The idea is to determine the set of characters used in \(U\), set these characters as unequal as outlined in section \ref{subsubsec_encoding} and associate each unequal constant with their characters.\\
+The idea is to determine the set of characters used in \(U\), set these characters as unequal as outlined in Section \ref{subsubsec_encoding} and associate each unequal constant with their characters.\\
 
@@
 The constraint is violated if a qualifier on a constrained statement exists that is unequal to all allowed qualifiers.
-The notation for set inequality as introduced in section \ref{subsec_3_allowed_units} will be used.
+The notation for set inequality as introduced in Section \ref{subsec_3_allowed_units} will be used.
 With \(A\) the set of allowed qualifiers the rule is:
@@
 The following rule is then created for every possible split of the \emph{separators} into two such sets.
-Note that \(\text{predicate}(\{X\})\) denotes \(\bigwedge_{x \in X} \text{predicate}(x)\) as introduced in section \ref{subsec_3_allowed_units}.
+Note that \(\text{predicate}(\{X\})\) denotes \(\bigwedge_{x \in X} \text{predicate}(x)\) as introduced in Section \ref{subsec_3_allowed_units}.
 \begin{equation*}
@@
 \paragraph{Rules}
-The \emph{mandatory qualifier} constraint needs to determine qualifier non-existence on constrained statements to find violations, which is similar to section \ref{subsubsec_single_value_with_separators}.
+The \emph{mandatory qualifier} constraint needs to determine qualifier non-existence on constrained statements to find violations, which is similar to Section \ref{subsubsec_single_value_with_separators}.
 The qualifier order needs to be provided and inequality must be established between all qualifier properties.
@@
 \end{equation*}
-With the usual require-pattern for qualifiers and the does\_not\_have rule as introduced in section \ref{subsubsec_single_value_with_separators} added for brevity the violations would be found with:
+With the usual require-pattern for qualifiers and the does\_not\_have rule as introduced in Section \ref{subsubsec_single_value_with_separators} added for brevity the violations would be found with:
 \begin{equation*}
@@
 \paragraph{Rules}
-The \emph{single best value} constraint, like the \emph{single value} constraint (section \ref{subsec_3_single_value}), differentiates between the set of \emph{separators} \(S\) being empty or not.
+The \emph{single best value} constraint, like the \emph{single value} constraint (Section \ref{subsec_3_single_value}), differentiates between the set of \emph{separators} \(S\) being empty or not.
 Since the constraint depends on the rank of statements this needs to be provided.
@@
 
-To evaluate the rules described in chapter \ref{cha_wikidata_constraints_explained} they have been implemented for VLog4J \cite{vlog4j}.
+To evaluate the rules described in Chapter \ref{cha_wikidata_constraints_explained} they have been implemented for VLog4J \cite{vlog4j}.
 VLog4J is a java library base on VLog \cite{vlog}, a Datalog engine utilising a column-based memory layout to allow processing of large datasets on commodity hardware.
@@
 The entire content of Wikidata is available in several formats: the aforementioned RDF, JSON and XML. To evaluate the violations, Wikidata needs to be provided in form of atoms as described in 
-section \ref{subsec_representation}. For this purpose the Wikidata Toolkit (WDTK)\cite{wdtk} was used. It provides an iterator over the JSON dumps, allowing the processing of all entities in Wikidata.
+Section \ref{subsec_representation}. For this purpose the Wikidata Toolkit (WDTK)\cite{wdtk} was used. It provides an iterator over the JSON dumps, allowing the processing of all entities in Wikidata.
 
 The evaluation aims to document how suitable Datalog and VLog in particular are to extract violations from Wikidata.
-The performance impact of the additionally introduced features, inequality (section \ref{sec_constraints_using_inequality}) and statement non-existence (section \ref{sec_constraints_using_statement_non}) is evaluated.
+The performance impact of the additionally introduced features, inequality (Section \ref{sec_constraints_using_inequality}) and statement non-existence (Section \ref{sec_constraints_using_statement_non}) is evaluated.
 
@@
 At the beginning the content of the Wikidata JSON-Dump, read via WDKT, is transformed into facts readable by VLog4J.
-These take the form of comma-separated value files which created for each predicate in section \ref{subsec_representation}, denoted as "Facts" in the diagram.
-
-Based on this set the Inequaliy Helper creates all files necessary to established inequality as described in section \ref{sec_constraints_using_inequality}, based on which approach was chosen.
+These take the form of comma-separated value files which created for each predicate in Section \ref{subsec_representation}, denoted as "Facts" in the diagram.
+
+Based on this set the Inequaliy Helper creates all files necessary to established inequality as described in Section \ref{sec_constraints_using_inequality}, based on which approach was chosen.
 
 The rules determining violations are created by the Constraint Checker, which uses the Wikidata Query Service\footnote{\url{query.wikidata.org}} to fetch the current constraint configurations.
-Based on these configurations the rules are created as described in chapter \ref{cha_wikidata_constraints_explained}.
+Based on these configurations the rules are created as described in Chapter \ref{cha_wikidata_constraints_explained}.
 
@@
 Once this is done, the reasoning engine materializes all possible facts.
-Afterwards, all violations are extracted by querying the database for the queries described in section \ref{subsec_task}.
+Afterwards, all violations are extracted by querying the database for the queries described in Section \ref{subsec_task}.
 
@@
 {\bf Total Rules} states how many rules where created to establish the violations.
-They are split into two subfields: {\bf Demand rules}, which are additional rules created to satisfy the requirements outlined in section \ref{subsubsec_demand-driven_materialisation}, and {\bf Base Rules}, which are the rules created without accounting for the demand-driven scheme.
+They are split into two subfields: {\bf Demand rules}, which are additional rules created to satisfy the requirements outlined in Section \ref{subsubsec_demand-driven_materialisation}, and {\bf Base Rules}, which are the rules created without accounting for the demand-driven scheme.
 
@@
 All further statistics describe the results on the 100,000 entity set unless specifically noted otherwise.
-Since inequality could only be computed in demanded mode (see section  \ref{sec_inequality_evaluation}), all statistics will show the results attained in this mode.
+Since inequality could only be computed in demanded mode (see Section  \ref{sec_inequality_evaluation}), all statistics will show the results attained in this mode.
 
@@
 However, also seen in Figure \ref{fig_runtime}, the runtimes of these three constraints are multiple orders of magnitude higher than those of other constraints.
-The following sections will thus focus on the comparison of these three constraints to identify bottlenecks and reasons for this huge difference in performance.
+The following Sections will thus focus on the comparison of these three constraints to identify bottlenecks and reasons for this huge difference in performance.
 They will be marked with a border in every figure.
@@
 Figure \ref{fig_facts} shows the number of facts loaded into VLog per constraint.
-Base facts denote those directly extracted from the JSON-Dump and corresponding to one of the predicates listed in section \ref{subsec_representation}.
-Inequality are the letter\(_i\) facts created to establish inequality as described in section \ref{sec_constraints_using_inequality}.
+Base facts denote those directly extracted from the JSON-Dump and corresponding to one of the predicates listed in Section \ref{subsec_representation}.
+Inequality are the letter\(_i\) facts created to establish inequality as described in Section \ref{sec_constraints_using_inequality}.
 Note that in the demanded (and encoded) mode, every entry for which inequality should be established is usually split into multiple letter\(_i\) facts based on its length, which is why the number of inequality facts can be up to ten times the number of base facts.
-Require facts list the order of statements or qualifiers as described in section \ref{sec_constraints_using_statement_non}.
+Require facts list the order of statements or qualifiers as described in Section \ref{sec_constraints_using_statement_non}.
 
@@
 \subsection{Inequality evaluation}\label{sec_inequality_evaluation}
-Inequality computation was implemented in three variants outlined in section \ref{sec_constraints_using_inequality}: naive, encoded and demanded.
+Inequality computation was implemented in three variants outlined in Section \ref{sec_constraints_using_inequality}: naive, encoded and demanded.
 Naive and encoded could not terminate for any dataset other than the smallest with 1,000 entities.
@@
 One possible bottleneck are the additional derivations necessary to identify violations.
-This includes the inequalities, demands for inequalities, requirements to determine statement non-existence as well as other helper rules introduced in chapter \ref{cha_wikidata_constraints_explained}.
-Note that the four simple constraints described in section \ref{sec_simple_constraints} do not generate any additional facts and are excluded from the chart.
+This includes the inequalities, demands for inequalities, requirements to determine statement non-existence as well as other helper rules introduced in Chapter \ref{cha_wikidata_constraints_explained}.
+Note that the four simple constraints described in Section \ref{sec_simple_constraints} do not generate any additional facts and are excluded from the chart.
 Figure \ref{fig_add_derivations} shows that the three performance-heavy constraints do generate by far the most additional derivations.
@@
 While the dataset does not include the entire database of Wikidata, which with 50m entities is about 500 times bigger, the runtimes for the entire dataset should still be within a reasonable frame.
-The exception are the three constraints focused on in section \ref{sec_results}.
-As shown in that section, the main difference between \emph{distinct values}, \emph{item requires statement}, and \emph{single value} and other constraints is
+The exception are the three constraints focused on in Section \ref{sec_results}.
+As shown in that Section, the main difference between \emph{distinct values}, \emph{item requires statement}, and \emph{single value} and other constraints is
 \begin{enumerate}[a)]
@@
 \subsection{Distinct values}
-While the rules for the \emph{distinct values} constraint (see section \ref{subsec_distinct_values}) are relatively simple in writing, they do require a large amount of processing.
+While the rules for the \emph{distinct values} constraint (see Section \ref{subsec_distinct_values}) are relatively simple in writing, they do require a large amount of processing.
 Distinct values requires the comparison of every constrained statement with every other constrained statement, meaning its evaluation would scale badly due to the quadratic growth of statements to be compared.
-Additionally, the number of inequalities that have to be evaluated according to the demanded approach (see section \ref{sec_constraints_using_statement_non}) grows accordingly.
+Additionally, the number of inequalities that have to be evaluated according to the demanded approach (see Section \ref{sec_constraints_using_statement_non}) grows accordingly.
 Due to the nature of the constraint -- requiring distinct values for a property across all of Wikidata -- this is in all likelihood unavoidable.
@@
 \subsection{Item requires statement}\label{subsec_findings_item_requires_statement}
-The \emph{item requires statement} constraint (see section \ref{subsec_item_requires_statement}) creates additional facts about every statement on an entity with a constrained statement (if the constraint was violated on that entity).
-Of interest to note is that the \emph{value requires statement} constraint (see section \ref{subsec_value_requires_statement}), which works almost identically, does not show a performance nearly as bad as \emph{item requires statement}.
+The \emph{item requires statement} constraint (see Section \ref{subsec_item_requires_statement}) creates additional facts about every statement on an entity with a constrained statement (if the constraint was violated on that entity).
+Of interest to note is that the \emph{value requires statement} constraint (see Section \ref{subsec_value_requires_statement}), which works almost identically, does not show a performance nearly as bad as \emph{item requires statement}.
 \emph{Value requires statement} has, however, only 244 constrained properties, while \emph{item requires statement} has 4265.
@@
 \subsection{Single value}
-The \emph{single value} constraint (see section \ref{subsec_single_value}) has to be analysed with regards to the two cases: without separators (section \ref{subsubsec_single_value_no_separators}) and with separators (section \ref{subsubsec_single_value_with_separators}.
+The \emph{single value} constraint (see Section \ref{subsec_single_value}) has to be analysed with regards to the two cases: without separators (Section \ref{subsubsec_single_value_no_separators}) and with separators (Section \ref{subsubsec_single_value_with_separators}.
 
@@
 For \emph{single value} this is qualifier non-existence on constrained statements.
-As seen in section \ref{subsec_findings_item_requires_statement}, this could be a reason for the bad performance of the analysis for this constraint.
+As seen in Section \ref{subsec_findings_item_requires_statement}, this could be a reason for the bad performance of the analysis for this constraint.
 Another source of processing time for \emph{single value} are the has\_same facts, which need to compare all constrained statements on the same item and all their qualifiers.
-Of interest to note is that the \emph{single best value} constraint (see section \ref{subsec_single_best_value}) works in almost the same way but does not suffer from the same performance.
+Of interest to note is that the \emph{single best value} constraint (see Section \ref{subsec_single_best_value}) works in almost the same way but does not suffer from the same performance.
 Again, the number of constrained properties is significantly lower for \emph{single best value} (203) compared to \emph{single value} (2909).
@@
 This order is, however, not easily readable and, if taken as presented, would likely cause implementation errors because the different positions are confused for each other.
-To prevent this the datasets imported from Wikidata and created during preprocessing as well as the rules outlined in chapter \ref{cha_wikidata_constraints_explained} are sorted afterwards.
+To prevent this the datasets imported from Wikidata and created during preprocessing as well as the rules outlined in Chapter \ref{cha_wikidata_constraints_explained} are sorted afterwards.
 Thus the rules can be kept in a readable form while still optimizing for VLog.

### 2018-09-18 src/main/java/main/Main.java
Added correct extraction reuse

@@ -246,7 +246,14 @@ public class Main {
 				logger.error("Error closing the data set files.", e);
 				return;
 			}
-		}	
+
+			try {
+				InequalityHelper.close();
+			} catch (IOException e) {
+				logger.error("Error closing the inequality set files.", e);
+				return;
+			}
+		}
 
 		if (!cmd.hasOption("noviolations")) {
 			try {

### 2018-09-18 src/main/java/utility/InequalityHelper.java

@@ -211,7 +211,13 @@ public class InequalityHelper {
 		}
 		
 		return atoms;
-	}	
+	}
+	
+	public static void close() throws IOException {
+		for (InequalityHandler handler : helpers.values()) {
+			handler.close();
+		}
+	}
 	
 	public static class InequalityHandler {
 		final String FOLDER;
@@ -247,12 +253,6 @@ public class InequalityHelper {
 		}
 
 		void close() throws IOException {
-			for (DataSetFile file : files) {
-				file.close();
-			}
-		}
-
-		void load(Reasoner reasoner) throws ReasonerStateException, IOException {
 			encodedAdditional();
 			
 			File uniqueCharacters = new File(DataSetFile.BASE_LOCATION + FOLDER + UNIQUE_CHARACTERS);
@@ -263,7 +263,13 @@ public class InequalityHelper {
 				writer.write(character + "\n");
 			}		
 			writer.close();
-			
+
+			for (DataSetFile file : files) {
+				file.close();
+			}
+		}
+
+		void load(Reasoner reasoner) throws ReasonerStateException, IOException {			
 			commonLoad(reasoner);
 		}
 		

### 2018-09-18 thesis/paper/paper.tex
Typos

@@
 This is done via a statement ID, which is assigned to every triple storing information about an item.
-Statements have one of three ranks, \emph{normal}, \emph{preferred} and \emph{deprecated} which are used to differentiate similar statements on the same item.
+Statements have one of three ranks, \emph{normal}, \emph{preferred}, and \emph{deprecated} which are used to differentiate similar statements on the same item.
 For example, the \emph{preferred} rank would be given to the current \emph{head of state} (P35) of a country, while all statements denoting former \emph{heads of state} have the rank \emph{normal}.
@@
 
-Every statement can have added qualifiers, which consist of the corresponding statement id, a predicate and an object.
+Every statement can have added qualifiers, which consist of the corresponding statement id, a predicate, and an object.
 Qualifiers provide contextual information for statements.
-The most common context is the time frame in which the statement was accurate, e.g.
-stating the \emph{start time} (P580) and \emph{end time} (P582) of the term of a head of state, but other information is possible, e.g.
+The most common context is the time frame in which the statement was accurate, e.g., stating the \emph{start time} (P580) and \emph{end time} (P582) of the term of a head of state, but other information is possible, e.g.
 determination method for measurements or the elevation above sea level for the highest point.
@@
 when it was retrieved or when the source was published as property-value pairs.
-Each references consists from multiple such pairs, and statements can have multiple references.
+Each reference consists of multiple such pairs, and statements can have multiple references.
 
@@
 Relations in datalog are expressed using \emph{predicates} with a specified arity \(n\), written as predicate\([n]\).
-To stay close to Wikidata in our examples let us take the predicate triple\([3]\) as an example.
+To stay close to Wikidata in our examples, let us take the predicate triple\([3]\) as an example.
 This would be used to express the \emph{item}-\emph{property}-\emph{value} relation with each position representing one part of the triple.\\
-Additionally, datalog also provides \emph{variables}, which will be capitalized like \(X\).
+Additionally, datalog also provides \emph{variables}, which will be capitalized, like \(X\).
 The sets of predicates {\bf P}, constants {\bf C} and variables {\bf V} are mutually disjoint.
@@
 If \(V_c = \emptyset\) the rule is called a \emph{datalog rule} and if \(l = 0\) it is called a \emph{fact}.
-If a fact contains no variables but only constants it is a \emph{ground rule}.
+If a fact contains no variables but only constants, it is a \emph{ground rule}.
 Since all \(V_p\) and \(V_r\) are  quantified, the preceding universal quantifiers \(\forall V_p \forall V_r\) will be omitted for the rules presented in this work.
@@
 For the following example, taken from Wikidata, descriptive names will be used instead of the IDs to increase readability.
-The individual entities and properties are introduced with their ID beforehand so the example can be retraced.
+The individual entities and properties are introduced with their ID beforehand, so the example can be retraced.
 The species \emph{blue whale} (Q42196) has the \emph{parent taxon} (P171) \emph{balaenoptera} (Q133320), which in turn has the parent taxon \emph{rorqual} (Q232829).
@@
 
-To improve the data quality, to help editors avoid common mistakes and to clarify the usage of properties, Wikidata allows constraints to be defined on properties.
-They are, however, not firm rules and exceptions can and should be made if necessary.
+To improve the data quality, to help editors avoid common mistakes, and to clarify the usage of properties, Wikidata allows constraints to be defined on properties.
+They are, however, not firm rules, and exceptions can and should be made if necessary.
 A property constraint says something about the way a property may or may not be used.
@@
 In Table \ref{tab_property_constraints} you can find the existing property constraints as of September 2018.
-Afterwards the constraints will be explained in detail and their modelling given.
+Afterwards, the constraints will be explained in detail and their modelling given.
 For constraint that cannot be modelled, reasoning for this assessment is included.
@@
 Note that a constrained statement means a statement whose predicate is a property that has a specific constraint.
-A simple example for this would be \emph{distinct values} (Q21502410) stating that values of the property must be unique across all constrained statements.
+A simple example for this would be \emph{distinct values} (Q21502410), stating that values of the property must be unique across all constrained statements.
 For this case no additional information regarding the constraint is necessary and violation could be found by surveying all constrained statements and comparing their values.\\
 Constraints can be further specified by using qualifiers on the constraint statement.
-These qualifiers are either optional or mandatory and may be applied once or multiple times.
+These qualifiers are either optional or mandatory, and may be applied once or multiple times.
 For example, the \emph{none of} constraint (Q52558054) states that values for the specific property may not be in a given set of disallowed values.
@@
 More complex dependencies are also possible.
-A constraint can depend on all constrained statements of a specific item (\emph{single value} constraint [Q19474404]), on all statements of the item specified as the value of a constrained statement (\emph{value requires statement} constraint [Q21510864]) or on all statements using the subclassOf-predicate (\emph{type} constraint [Q21503250]).
+A constraint can depend on all constrained statements of a specific item (\emph{single value} constraint [Q19474404]), on all statements of the item specified as the value of a constrained statement (\emph{value requires statement} constraint [Q21510864]), or on all statements using the subclass of-predicate (\emph{type} constraint [Q21503250]).
 
@@
 Six can be solved if constant inequality is defined.
-This can be easily solved with brute force but will be improved in Section \ref{sec_inequality}.
+This can be easily solved with brute force but will be improved in Section \ref{sec_constraints_using_inequality}.
 A further seven require negation of a specific fact which can be achieved with reasonable effort.
@@
 The feature mostly needed is a general form of negation that would allow to test for the non-existence of a derived fact.
-Other features needed for these constraints are regular expressions, which could theoretically be translated but would require large amounts of pre-processing, value comparison, and finally web requests to test if a link is valid.
+Other features needed for these constraints are regular expressions, which could theoretically be translated, but would require large amounts of pre-processing, value comparison, and finally web requests to test if a link is valid.
 
@@
 A problem that occurs multiple times is the need to determine inequality between constants.
-\emph{distinct value}, \emph{allowed units}, \emph{allowed qualifiers}, \emph{one of}, \emph{single best value} and \emph{one-of qualifier value} each require a rule to determine if a value, unit or property is not equal to another.
+\emph{distinct value}, \emph{allowed units}, \emph{allowed qualifiers}, \emph{one of}, \emph{single best value} and \emph{one-of qualifier value} each require a rule to determine if a value, unit, or property is not equal to another.
 To this end a certain amount of preprocessing is necessary, most notably the extraction of all constant pairs whose inequality has to be determined for finding the violations.
-To avoid putting and undue amount of processing outside of datalog this extraction will not check more than that the constant is in the right position relative to the constrained statement, e.g., all values on statements constrained by \emph{distinct value} or all qualifiers on statements constrained by \emph{allowed qualifiers}.
-From this set inequality will be generated, different versions of which will be introduced in Section \ref{sec_inequality} and evaluated with regards to their performance in Section \ref{sec_inequality_evaluation}.\\
+To avoid putting and undue amount of processing outside of datalog, this extraction will not check more than that the constant is in the right position relative to the constrained statement, e.g., all values on statements constrained by \emph{distinct value} or all qualifiers on statements constrained by \emph{allowed qualifiers}.
+From this set inequality will be generated, different versions of which will be introduced in Section \ref{sec_inequality} and evaluated with regards to their performance in Section \ref{sec_constraints_using_inequality}.\\
 
@@
 The idea is to ask the question if a second statement exists, then evaluating it as above.
-Both methods will be explained in Section \ref{sec_statement_non-existence}.
+Both methods will be explained in Section \ref{sec_constraints_using_statement_non}.
 
@@
 To evaluate the rules described in Chapter \ref{cha_wikidata_constraints_explained} they have been implemented for VLog4J \cite{vlog4j}.
-VLog4J is a java library base on VLog \cite{vlog}, a Datalog engine utilising a column-based memory layout to allow processing of large datasets on commodity hardware.
+VLog4J is a Java library based on VLog \cite{vlog}, a Datalog engine utilising a column-based memory layout to allow processing of large datasets on commodity hardware.
 It was first published in 2016\cite{UJK2016} but has been actively developed since then.
 
-The entire content of Wikidata is available in several formats: the aforementioned RDF, JSON and XML. To evaluate the violations, Wikidata needs to be provided in form of atoms as described in 
+The entire content of Wikidata is available in several formats: RDF (as mentioned), JSON and XML. To evaluate the violations, Wikidata needs to be provided in form of atoms as described in 
 Section \ref{subsec_representation}. For this purpose the Wikidata Toolkit (WDTK)\cite{wdtk} was used. It provides an iterator over the JSON dumps, allowing the processing of all entities in Wikidata.
@@
 
-Based on this set the Inequaliy Helper creates all files necessary to established inequality as described in Section \ref{sec_constraints_using_inequality}, based on which approach was chosen.
-
-The rules determining violations are created by the Constraint Checker, which uses the Wikidata Query Service\footnote{\url{query.wikidata.org}} to fetch the current constraint configurations.
+At the same time the Inequaliy Helper, receiving the Facts from WDTK, creates all files necessary to establish inequality as described in Section \ref{sec_constraints_using_inequality}, based on which approach was chosen.
+
+The rules determining violations are created by the Constraint Checker, which uses the Wikidata Query Service (WDQS)\footnote{\url{query.wikidata.org}} to fetch the current constraint configurations, which are the set constraints and their qualifiers as described in Chapter \ref{cha_property_constraints}.
 Based on these configurations the rules are created as described in Chapter \ref{cha_wikidata_constraints_explained}.
 
-Computing all violations based on the provided (inequality)-facts and violations rules is done per constraint.
+Computing all violations based on the provided (inequality)-facts and violations rules is done per constraint type.
 All facts pertaining to the constraint as well as the rules are loaded into VLog, which creates and in-memory database from the provided .csv-files.
@@
 The experiments where carried out on a server with a 3.50GHz Intel Xeon CPU, 1 TB SSD and 32 GB assigned RAM running Debian 4.9.
-Multiple sets of sizes ranging from one thousand to one million entities where extracted from the Wikidata JSON-Dump and all constraints evaluated against them.
+Multiple sets of sizes ranging from one thousand to one million entities where randomly chosen from the Wikidata JSON-Dump and all constraints evaluated against them.
 
@@
 All further statistics describe the results on the 100,000 entity set unless specifically noted otherwise.
-Since inequality could only be computed in demanded mode (see Section  \ref{sec_inequality_evaluation}), all statistics will show the results attained in this mode.
+Since inequality could only be computed in on-demand mode (see Section  \ref{sec_inequality_evaluation}), all statistics will show the results attained in this mode.
 
@@
 \includegraphics[width=\linewidth]{images/rules.pdf}
-\caption{Number of rules and demanded rules}\label{fig_rules}
+\caption{Number of rules and on-demand rules}\label{fig_rules}
 \end{figure}
@@
 Inequality are the letter\(_i\) facts created to establish inequality as described in Section \ref{sec_constraints_using_inequality}.
-Note that in the demanded (and encoded) mode, every entry for which inequality should be established is usually split into multiple letter\(_i\) facts based on its length, which is why the number of inequality facts can be up to ten times the number of base facts.
+Note that in the on-demand (and encoded) mode, every entry for which inequality should be established is usually split into multiple letter\(_i\) facts based on its length, which is why the number of inequality facts can be up to ten times the number of base facts.
 Require facts list the order of statements or qualifiers as described in Section \ref{sec_constraints_using_statement_non}.
@@
 \subsection{Inequality evaluation}\label{sec_inequality_evaluation}
-Inequality computation was implemented in three variants outlined in Section \ref{sec_constraints_using_inequality}: naive, encoded and demanded.
+Inequality computation was implemented in three variants outlined in Section \ref{sec_constraints_using_inequality}: naive, encoded and on-demand.
 Naive and encoded could not terminate for any dataset other than the smallest with 1,000 entities.
@@
 Encoded, while improving on the disk space needed, still failed due to out of memory errors.
-This behaviour is expected, since for an input of size \(n\) \(n^2\) inequalities have to be materialized.
-The runtimes for this dataset, presented in Figure \ref{fig_inequality}, shows a clear improvement when using the demanded approach.
-This shows that the overhead in rules introduced to enable the demanded approach, seen in Figure \ref{fig_rules}, is worth the improvements in performance gained through it.
+This behaviour is expected, since for an input of size \(n\), the number of inequalities that have to be materialized is \(n^2\).
+The runtimes for this dataset, presented in Figure \ref{fig_inequality}, shows a clear improvement when using the on-demand approach.
+This shows that the overhead in rules introduced to enable the on-demand approach, seen in Figure \ref{fig_rules}, is worth the improvements in performance gained through it.
 
@@
 \includegraphics[width=\linewidth]{images/inequality1000.pdf}
-\caption{Runtime materialization for 1,000 entities (encoded and demanded)}\label{fig_inequality}
+\caption{Runtime materialization for 1,000 entities (encoded and on-demand)}\label{fig_inequality}
 \end{figure}
@@
 Distinct values requires the comparison of every constrained statement with every other constrained statement, meaning its evaluation would scale badly due to the quadratic growth of statements to be compared.
-Additionally, the number of inequalities that have to be evaluated according to the demanded approach (see Section \ref{sec_constraints_using_statement_non}) grows accordingly.
+Additionally, the number of inequalities that have to be evaluated according to the on-demand approach (see Section \ref{sec_constraints_using_statement_non}) grows accordingly.
 Due to the nature of the constraint -- requiring distinct values for a property across all of Wikidata -- this is in all likelihood unavoidable.

### 2018-09-19 thesis/paper/paper.tex
Typos

@@
 Constraints are used to ensure data correctness as well as clarify the data structure in cases where it might be unclear how facts should be described.
-A detailed explanation of these constrains can be found
-in Chapter \ref{cha_property_constraints}.
+A detailed explanation of these constrains can be found in Chapter \ref{cha_property_constraints}.
 
@@
 \subsection{Representation of Wikidata as Facts}\label{subsec_representation}
-To represent the content of Wikidata as facts the following predicates are introduced:
+To represent the content of Wikidata as facts, the following predicates are introduced:
 \begin{table}[H]
@@
 The task for each constraint is thus: Given the constraint and a constrained property, potentially with qualifiers, which rule finds all violations of the rule in a given set of triples, qualifiers and references if applicable?
-The given rule set should return all violations if queries with the following atoms:
+The given rule set should return all violations if queried with the following atoms:
 \begin{align*}
@@
 
-{\bf Rules:} Most constraints require more data than just the statements, qualifiers and references.
-This can include additional data from Wikidata to be provided as facts using the introduced predicates (see Table \ref{tab_predicates}), inequalities to be established as a pre-processing step (see Section \ref{sec_inequality}) or preparations to determine specific kinds of negation (see Section \ref{sec_statement_non-existence}).
+{\bf Rules:} Most constraints require more data than just the statements, qualifiers, and references.
+This can include additional data from Wikidata to be provided as facts using the introduced predicates (see Table \ref{tab_predicates}), inequalities to be established as a pre-processing step (see Section \ref{sec_constraints_using_inequality}), or preparations to determine specific kinds of negation (see Section \ref{sec_constraints_using_statement_non}).
 Afterwards, the rules necessary to determine violations are described.
@@
 \paragraph{Constraint}
-The \emph{property scope} constraint specifies if a property may be used as predicate in a statement, qualifier or reference.
+The \emph{property scope} constraint specifies if a property may be used as predicate in a statement, qualifier, or reference.
 \begin{table}[H]
@@
 \hline
-Qualifiers & \makecell{\emph{property scope} (P5314) -- 1..3 \\ must be \emph{as main value} (Q54828448), \emph{as qualifier} (Q54828449) \\ or \emph{as reference} (Q54828450) \\ specifies allowed use in statements, qualifiers or references} \\
+Qualifiers & \makecell{\emph{property scope} (P5314) -- 1..3 \\ must be \emph{as main value} (Q54828448), \emph{as qualifier} (Q54828449), \\ or \emph{as reference} (Q54828450) \\ specifies allowed use in statements, qualifiers, or references} \\
 \hline
@@
 The used abbreviations are listed in Table \ref{tab_abbreviations}.
-In this case this is simply each statement, qualifier or reference whose predicate is the constrained property.\\
-The \emph{property scope} constraint requires the scope, meaning statement, qualifier or reference, to be set via qualifier \emph{property scope} (P5314).
+In this case, this is simply each statement, qualifier, or reference whose predicate is the constrained property.\\
+The \emph{property scope} constraint requires the scope, meaning statement, qualifier, or reference, to be set via qualifier \emph{property scope} (P5314).
 The cardinality is 1..3 because it is mandatory and has only three possible values.\\
-It is violated if a statement, qualifier or reference has a predicate with the \emph{property scope} constraint that does not allow this scope.\\
+It is violated if a statement, qualifier, or reference has a predicate with the \emph{property scope} constraint that does not allow this scope.\\
 
@@
 If a concrete constrained on a property does not have \emph{as main value} as a value of the qualifier, the first rule is used.
-The case for \emph{as qualifier} and \emph{as reference} is analogue with the second and third rule, respectively.
+The case for \emph{as qualifier} and \emph{as reference} is analogous with the second and third rule, respectively.
 
@@
 \paragraph{Constraint}
-The \emph{allowed entity types} constraint limits if a property may be used in a statement whose subject is an item, a property or a lexeme, although as of September 2018 lexemes are not deployed.
+The \emph{allowed entity types} constraint limits if a property may be used in a statement whose subject is an item, a property, or a lexeme, although as of September 2018 lexemes are not deployed.
 
@@
 \hline
-Qualifiers & \makecell{allowed entity type (P2305) -- 1..3 \\ must be \emph{Wikibase item} (Q29934200), \emph{Wikibase property} (Q29934218) \\ or lexeme (Q51885771); \\ specifies the entity type the property may be used on} \\
+Qualifiers & \makecell{allowed entity type (P2305) -- 1..3 \\ must be \emph{Wikibase item} (Q29934200), \emph{Wikibase property} (Q29934218), \\ or lexeme (Q51885771); \\ specifies the entity type the property may be used on} \\
 \hline
@@
 The allowed entity types constraint is violated by a statement with the wrong entity type.
-Constraints regarding the lexeme entity type can be ignored since they are currently not used.
-The Violations depend on a simple conjunction:
+Constraints regarding the lexeme entity type can be ignored since they are not currently used.
+The violations depend on a simple conjunction:
 \begin{equation*}
@@
 The first rule are created for all \emph{allowed entity type} constraints whose \emph{allowed entity type} does not contain \emph{Wikibase item}.
-The second rule works analogue for \emph{Wikibase property}.
+The second rule works analogous for \emph{Wikibase property}.
 
@@
 \hline
-Qualifiers & \makecell{\emph{forbidden value} (P2305) -- 1..* \\ values that may not be object of a constraint statement, qualifier or reference} \\
-\hline
-Violation & \makecell{constrained statement, qualifier or reference with a forbidden value} \\
+Qualifiers & \makecell{\emph{forbidden value} (P2305) -- 1..* \\ values that may not be object of a constraint statement, qualifier, or reference} \\
+\hline
+Violation & \makecell{constrained statement, qualifier, or reference with a forbidden value} \\
 \hline
@@
 \paragraph{Rules}
-The none of constraint simply forbids certain values for properties, thus effectively prohibiting certain triples, qualifiers and references.
+The none of constraint simply forbids certain values for properties, thus effectively prohibiting certain statements, qualifiers, and references.
 The rules, depending on nothing but specific database entries, are
@@
 Each constraint will list the possible unequal entities in relation to the constrained statement.
-Multiple ways of implementing inequality while keeping the number of rules small will be discussed.
+Multiple ways of implementing inequality will be discussed.
 Afterwards all constraints that can be translated using inequality will be explained.
@@
 
-To save space and avoid redundancy the following rules will not explicitly state the inequality requirement predicate.
+To save space and avoid redundancy, the following rules will not explicitly state the inequality requirement predicate.
 Instead, assume that for every premise of the form \(B_1 \wedge \ldots\wedge B_m \wedge I_1 \wedge I_n\) where \(B_1 ..
@@
 Two constrained statements with the same value can be found via a simple conjunction of two statement\_EDB atoms with the propertyConstant as predicate and a variable \(V\) as value.
-Since this would mark every constrained statement as a violations because it can fulfil both atoms at the same time, inequality of the statement IDs \(S\) and \(O\) must be tested as well.
+Since this would mark every constrained statement as a violation because it can fulfil both atoms at the same time, inequality of the statement IDs \(S\) and \(O\) must be tested as well.
 \begin{equation*}
@@
 \end{equation*}
-The rules for qualifiers and reference are analogue.
+The rules for qualifiers and reference are analogous.
 
@@
 
-By setting the \emph{constraint scope} (P4680) qualifier on the constraint it can be limited to only affect a subset of statement, qualifier, and reference.
+By setting the \emph{constraint scope} (P4680) qualifier on the constraint, it can be limited to only affect a subset of statement, qualifier, and reference.
 As of September 2018 this is not done on any constraint.
@@
 \paragraph{Rules}
-To handle this constraint inequality between all units need to be established.
-Then a violation can be identified by finding all statement\_EDB, qualifier\_EDB and reference\_EDB with the constrained property and a unit unequal to all allowed units.
+To handle this constraint, inequality between all units need to be established.
+Then a violation can be identified by finding all statement\_EDB, qualifier\_EDB, and reference\_EDB with the constrained property and a unit unequal to all allowed units.
 The allowed units constraint needs to determine inequality between all allowed units and each unit on a constrained statement.
-To this end the units of a triple, qualifier or reference needs to be extracted during pre-processing and their inequality established.
-Quantities are rendered as string during pre-processing, specifying their value, bounds and unit.
+To this end the units of a statement, qualifier, or reference need to be extracted during pre-processing and their inequality established.
+Quantities are rendered as string during pre-processing, specifying their value, bounds, and unit.
 
@@
 \end{equation*}
-with qualifiers and references treated analogue.
+with qualifiers and references treated analogous.
 
@@
 
-Violation are thus
+Violations are thus
 \begin{equation*}
@@
 \paragraph{Rules}
-To represent a violation of the \emph{allowed qualifiers} constraint in datalog inequality between all specified allowed qualifiers and the properties of qualifiers is required.
+To represent a violation of the \emph{allowed qualifiers} constraint in datalog, inequality between all specified allowed qualifiers and the properties of qualifiers is required.
 The constraint is violated if a qualifier on a constrained statement exists that is unequal to all allowed qualifiers.
@@
 \paragraph{Constraint}
-The \emph{one-of} constraint limits the values of a constraint statement, qualifier or reference to a set of given values.
+The \emph{one-of} constraint limits the values of a constraint statement, qualifier, or reference to a set of given values.
 
@@
 \hline
-Required data & constrained statements, qualifiers and references \\
+Required data & constrained statements, qualifiers, and references \\
 \hline
@@
 \paragraph{Rules}
-The one-of constraint fits in the pattern established by the previous two constraints and disallows certain property-value combinations in statements, qualifiers and references.
-Inequality is established between all \emph{allowed values} and the values of statements, qualifiers and references, respectively.
-With \(A\) the set of allowed values the rules for triple, qualifier and reference are:
+The one-of constraint fits the pattern established by the previous two constraints and disallows certain property-value combinations in statements, qualifiers, and references.
+Inequality is established between all \emph{allowed values} and the values of statements, qualifiers, and references, respectively.
+With \(A\) the set of allowed values the rules for triple, qualifier, and reference are:
 \begin{align*}
@@
 
-If qualifier non-existence should be tested the predicates have to be adapted, since qualifiers do not have a unique identifier.
-To assign them a unique identity, statement ID, property and value have to be used.
-For this the predicates first\_qualifier\([3]\) and last\_qualifier\([3]\) taking the statement ID, the property and the value and next\_qualifier\([6]\) with ID, property and value for the current and next qualifier are introduced.\\
+If qualifier non-existence should be tested, the predicates have to be adapted since qualifiers do not have a unique identifier.
+To assign them a unique identity, statement ID, property, and value have to be used.
+For this the predicates first\_qualifier\([3]\) and last\_qualifier\([3]\) taking the statement ID, the property, and the value, as well as next\_qualifier\([6]\) with ID, property, and value for the current and next qualifier are introduced.\\
 
@@
 
-For statements the pattern consists of two rules, both utilising the same conjunction \(c = C_1 \wedge \ldots \wedge C_m\) with atoms \(C_i\) depending on the constraint which is used to test if the statement does not fulfil the requirement.
+For statements the pattern consists of two rules, both utilising the same conjunction \(c = C_1 \wedge \ldots \wedge C_m\) with atoms \(C_i\) depending on the constraint.
+This conjunction is used to test if the statement does not fulfil the requirement.
 Then rule one
@@
 
-If the last statement of an item or the last qualifier of a statement has require the violation occurred.
+If the last statement of an item or the last qualifier of a statement has the require-predicate the violation occurred.
 In the following constraints only the conjunction \(c\) and adjustments to this pattern will be described.
@@
 The order of the item statements needs to be established.
-An item violates the constraint if it a) has a statement using constrained property as predicate, b1) has no statement using the \emph{required property} as predicate, or (if \emph{allowed values} are specified) b2) has only statement using the \emph{required property} as predicate that do not use one of the \emph{allowed values} as value.
+An item violates the constraint if it a) has a statement using the constrained property as predicate, b1) has no statements using the \emph{required property} as predicate, or (if \emph{allowed values} are specified) b2) has only statement using the \emph{required property} as predicate that do not use one of the \emph{allowed values} as value.
 The required term \(t\) is the \emph{required property}.
@@
 \paragraph{Constraint}
-The \emph{value requires statement} constraint requires an additional statement on the value of constrained statement, possibly also using a specific value from the list of allowed values.
+The \emph{value requires statement} constraint requires an additional statement on the value of the constrained statement, possibly also using a specific value from the list of allowed values.
 Since the constraint works very similar to item requires statement (see \ref{subsec_item_requires_statement}) --- requiring a statement on the value of the constrained statement instead of on the item --- Table \ref{tab_item_requires_statements} is not repeated.
@@
 The \emph{value requires statement} constraint is syntactically very similar to the previous \emph{item requires statement} constraint.
-Instead of requiring a statement on the item of a constrained statement it requires it on the value.
-Inequality need to be computed between all properties and all values and the item statement order established.
+Instead of requiring a statement on the item of a constrained statement, it requires it on the value.
+Inequality needs to be computed between all properties and all values and the item statement order established.
 The conjunctions change in the first statement and are:
@@
 It is thus equal to an \emph{inverse} constraint (see \ref{subsec_2_inverse}) with the same property as constrained property and inverse property.
-It can be translated exactly as the inverse constraint if the inverse property where set to the same as the constrained property.
+It can be translated exactly as the inverse constraint if the inverse property were set to the same as the constrained property.
 
@@
 The \emph{cast member} (P161) property requires an item to have multiple statements using this property, since most plays or films have multiple cast members.
-This is an example for a constraint from which exception can and should be made and is only meant to point to possibly missing data.
+This is an example for a constraint from which exceptions can and should be made and is only meant to point to possibly missing data.
 \end{example}
@@
 \end{equation*}
-A violations would be any last statement requiring a second constrained statement.\\
+A violation would be any last statement requiring a second constrained statement.\\
 
@@
 \end{equation}
-For qualifiers and references there is no ID whose inequality could be checked.
+For qualifiers and references, there is no ID whose inequality could be checked.
 However, since the statement ID and the property of a possible conflict are always the same, inequality can be tested on the values.
@@
 \subsubsection{With separators}\label{subsubsec_single_value_with_separators}
-To find all violations under \emph{separators} a number of things have to be considered.
-Only statements are eligible, since qualifiers cannot have further qualifier.
+To find all violations under \emph{separators}, a number of things have to be considered.
+Only statements are eligible, since qualifiers cannot have further qualifiers.
 The order of qualifiers on constrained statements needs to be provided as well since qualifier non-existence is relevant.
@@
 \begin{equation*}
-c = \text{statement\_EDB}(S, I, \text{propertyConstant}, C) \wedge \text{qualifier\_EDB}(S, P, V) \wedge \text{unequal}(P, \text{separatorConstant})
+\begin{split}
+c = \text{statement\_EDB}&(S, I, \text{propertyConstant}, C) \wedge{} \\
+\text{qualifier\_EDB}&(S, P, V) \wedge{} \\
+\text{unequal}&(P, \text{separatorConstant})
+\end{split}
 \end{equation*}
@@
 
-With qualifier non-existence established the violation can be found.
+With qualifier non-existence established, the violation can be found.
 Two statements form a violation if they are distinct as per equation \ref{eq_no_separators_triple} and additionally fulfil the condition that all values of \emph{separator} qualifiers are equal or no such \emph{separator} qualifier exist.
-To determine this a number of rules have to be introduced which are based on two sets the \emph{separators} are split into: \(X\) contains the \emph{separators} that exist as qualifiers and \(Y\) the \emph{separators} that do not.
+To determine this, a number of rules have to be introduced which are based on two sets the \emph{separators} are split into: \(X\) contains the \emph{separators} that exist as qualifiers and \(Y\) the \emph{separators} that do not.
 The following rule is then created for every possible split of the \emph{separators} into two such sets.
@@
 \end{equation*}
-With the usual require-pattern for qualifiers and the does\_not\_have rule as introduced in Section \ref{subsubsec_single_value_with_separators} added for brevity the violations would be found with:
+With the usual require-pattern for qualifiers and the does\_not\_have rule as introduced in Section \ref{subsubsec_single_value_with_separators} added for brevity, the violations would be found with:
 \begin{equation*}
@@
 \paragraph{Rules}
-The \emph{single best value} constraint, like the \emph{single value} constraint (Section \ref{subsec_3_single_value}), differentiates between the set of \emph{separators} \(S\) being empty or not.
-Since the constraint depends on the rank of statements this needs to be provided.
+The \emph{single best value} constraint, like the \emph{single value} constraint (Section \ref{subsec_single_value}), differentiates between the set of \emph{separators} \(S\) being empty or not.
+Since the constraint depends on the rank of statements, this needs to be provided.
 Qualifiers and references are not affected since they do not have ranks.
@@
 Multiple constraints have not been described above for different reasons.
-The most prominent of these is the \emph{format} constraint (Q21502404) which specifies a regular expressions the values of the constrained property must adhere to.
+The most prominent of these is the \emph{format} constraint (Q21502404) which specifies a regular expression the values of the constrained property must adhere to.
 Evaluating these expressions would be theoretically possible but \\
@@
 
-To evaluate the rules described in Chapter \ref{cha_wikidata_constraints_explained} they have been implemented for VLog4J \cite{vlog4j}.
+To evaluate the rules described in Chapter \ref{cha_wikidata_constraints_explained}, they have been implemented for VLog4J \cite{vlog4j}.
 VLog4J is a Java library based on VLog \cite{vlog}, a Datalog engine utilising a column-based memory layout to allow processing of large datasets on commodity hardware.
@@
 
-The entire content of Wikidata is available in several formats: RDF (as mentioned), JSON and XML. To evaluate the violations, Wikidata needs to be provided in form of atoms as described in 
+The entire content of Wikidata is available in several formats: RDF (as mentioned), JSON, and XML. To evaluate the violations, Wikidata needs to be provided in form of atoms as described in 
 Section \ref{subsec_representation}. For this purpose the Wikidata Toolkit (WDTK)\cite{wdtk} was used. It provides an iterator over the JSON dumps, allowing the processing of all entities in Wikidata.
@@
 The evaluation aims to document how suitable Datalog and VLog in particular are to extract violations from Wikidata.
-The performance impact of the additionally introduced features, inequality (Section \ref{sec_constraints_using_inequality}) and statement non-existence (Section \ref{sec_constraints_using_statement_non}) is evaluated.
+The performance impact of the additionally introduced features, inequality (Section \ref{sec_constraints_using_inequality}) and statement non-existence (Section \ref{sec_constraints_using_statement_non}), is evaluated.
 
@@
 At the beginning the content of the Wikidata JSON-Dump, read via WDKT, is transformed into facts readable by VLog4J.
-These take the form of comma-separated value files which created for each predicate in Section \ref{subsec_representation}, denoted as "Facts" in the diagram.
+These take the form of comma-separated value files, which are created for each predicate in Section \ref{subsec_representation}, denoted as "Facts" in the diagram.
 
@@
 Computing all violations based on the provided (inequality)-facts and violations rules is done per constraint type.
-All facts pertaining to the constraint as well as the rules are loaded into VLog, which creates and in-memory database from the provided .csv-files.
+All facts pertaining to the constraint, as well as the rules, are loaded into VLog, which creates an in-memory database from the provided .csv-files.
 Once this is done, the reasoning engine materializes all possible facts.
@@
 
-The experiments where carried out on a server with a 3.50GHz Intel Xeon CPU, 1 TB SSD and 32 GB assigned RAM running Debian 4.9.
+The experiments where carried out on a server with a 3.50GHz Intel Xeon CPU, 1 TB SSD, and 32 GB assigned RAM running Debian 4.9.
 Multiple sets of sizes ranging from one thousand to one million entities where randomly chosen from the Wikidata JSON-Dump and all constraints evaluated against them.
@@
 
-{\bf Total Rules} states how many rules where created to establish the violations.
+{\bf Total Rules} states how many rules were created to establish the violations.
 They are split into two subfields: {\bf Demand rules}, which are additional rules created to satisfy the requirements outlined in Section \ref{subsubsec_demand-driven_materialisation}, and {\bf Base Rules}, which are the rules created without accounting for the demand-driven scheme.
@@
 The exceptions are \emph{distinct values} (1.22 hours), \emph{item requires statement} (2.38 hours) and \emph{single value} (4.22 hours).
-In parts this is expected behaviour, since these three constraints are applied to an order of magnitude more properties than all other properties with the exception of \emph{scope}, which is described in very simple rules.
+In part this is expected behaviour, since these three constraints are applied to an order of magnitude more properties than all other properties with the exception of \emph{scope}, which is described in very simple rules.
 However, also seen in Figure \ref{fig_runtime}, the runtimes of these three constraints are multiple orders of magnitude higher than those of other constraints.
@@
 
-The long runtimes of \emph{distinct values}, \emph{item requires statement} and \emph{single value} cannot be explained through the size of the input data.
+The long runtimes of \emph{distinct values}, \emph{item requires statement}, and \emph{single value} cannot be explained through the size of the input data.
 While all of them have more than 10 million input facts, this is also the case for other constraints that do not take nearly as much runtime.
-\emph{One-of}, \emph{symmetric} and \emph{inverse} have even more input facts than the three long-runtime constraints.
+\emph{One-of}, \emph{symmetric}, and \emph{inverse} have even more input facts than the three long-runtime constraints.
 The reason for the performance difference must be found elsewhere.
@@
 \subsection{Inequality evaluation}\label{sec_inequality_evaluation}
-Inequality computation was implemented in three variants outlined in Section \ref{sec_constraints_using_inequality}: naive, encoded and on-demand.
+Inequality computation was implemented in three variants outlined in Section \ref{sec_constraints_using_inequality}: naive, encoded, and on-demand.
 Naive and encoded could not terminate for any dataset other than the smallest with 1,000 entities.
@@
 This behaviour is expected, since for an input of size \(n\), the number of inequalities that have to be materialized is \(n^2\).
-The runtimes for this dataset, presented in Figure \ref{fig_inequality}, shows a clear improvement when using the on-demand approach.
+The runtimes for this dataset, presented in Figure \ref{fig_inequality}, show a clear improvement when using the on-demand approach.
 This shows that the overhead in rules introduced to enable the on-demand approach, seen in Figure \ref{fig_rules}, is worth the improvements in performance gained through it.

### 2018-09-19 thesis/paper/paper.tex
Changed single value

@@
 The order of qualifiers on constrained statements needs to be provided as well since qualifier non-existence is relevant.
-The pattern needs to be applied for every \emph{separator}, meaning the conjunction is
-\begin{equation*}
-\begin{split}
-c = \text{statement\_EDB}&(S, I, \text{propertyConstant}, C) \wedge{} \\
-\text{qualifier\_EDB}&(S, P, V) \wedge{} \\
-\text{unequal}&(P, \text{separatorConstant})
-\end{split}
-\end{equation*}
-for each separator in \(S\).\\
+The require qualifier pattern needs to be applied to the qualifiers of constrained statements.
+The conjunction c is created for every \emph{separator} and the pattern applied.
+\begin{equation*}
+\begin{split}
+c = &\text{statement\_EDB}(S, I, \text{propertyConstant}, C) \wedge{} \\
+&\text{qualifier\_EDB}(S, P, V) \wedge{} \\
+&\text{unequal}(P, \text{separatorConstant})
+\end{split}
+\end{equation*}
 
@@
 &\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
-&\text{statement\_EDB}(O, I, \text{propertyConstant}, C) \wedge{} \\
+&\text{statement\_EDB}(O, I, \text{propertyConstant}, X) \wedge{} \\
 &\text{unequal}(S, O) \wedge{} \\
-&\text{has\_same}(S, O, \{X\}) \wedge{} \\
-&\text{does\_not\_have}(S, \{Y\}) \wedge{} \\
-&\text{does\_not\_have}(O, \{Y\})
-\end{split}
-\end{equation*}
-The predicate does\_not\_have\([2]\) linking a statement ID to the separator it does not have via 
-\begin{equation*}
-\begin{split}
-\text{does\_not\_have}(S, R) \leftarrow &\text{statement\_EDB}(S, I, \text{propertyConstant}, C) \wedge{} \\
-&\text{last\_qualifier}(S, P, V) \wedge \text{require}(S, P, V, \text{propertyConstant}, R)
-\end{split}
-\end{equation*}
-and the predicate has\_same[2] linking two statement IDs to the separator for which the values are the same via
-\begin{equation*}
-\begin{split}
-\text{has\_same}(S, O, Q) \leftarrow &\text{statement\_EDB}(S, I, \text{propertyConstant}, C) \wedge{} \\
-&\text{statement\_EDB}(O, I, \text{propertyConstant}, X) \wedge{} \\
-&\text{qualifier\_EDB}(S, Q, V) \wedge \text{qualifier\_EDB}(O, Q, V)
-\end{split}
-\end{equation*}
-are introduced for brevity.
+&\text{same\_or\_non\_existent}(S, O, \{Q\})
+\end{split}
+\end{equation*}
+\begin{equation*}
+\begin{split}
+\text{same\_or\_non\_existent}&(S, O, Q) \leftarrow \\
+&\text{statement\_EDB}(S, I, \text{propertyConstant}, X) \wedge{} \\
+&\text{statement\_EDB}(O, I, \text{propertyConstant}, Y) \wedge{} \\
+&\text{qualifier\_EDB}(S, Q, V) \wedge{} \\
+&\text{qualifier\_EDB}(O, Q, V)
+\end{split}
+\end{equation*}
+\begin{equation*}
+\begin{split}
+\text{same\_or\_non\_existent}&(S, O, Q) \leftarrow{} \\
+&\text{statement\_EDB}(S, I, \text{propertyConstant}, X) \wedge{} \\
+&\text{statement\_EDB}(O, I, \text{propertyConstant}, Y) \wedge{} \\
+&\text{last\_qualifier}(S, P, V) \wedge{} \\
+&\text{require}(S, P, V, \text{propertyConstant}, Q) \wedge{} \\
+&\text{last\_qualifier}(O, R, C) \wedge{} \\
+&\text{require}(O, R, C, \text{propertyConstant}, Q)
+\end{split}
+\end{equation*}
 

### 2018-09-19 src/main/java/impl/CC/ConstraintChecker.java

@@ -138,7 +138,7 @@ public abstract class ConstraintChecker {
 		prepareFacts();
 		logger.info("Loaded basic data sets.");
 		
-		List<Rule> rulesToAdd = new ArrayList<>();
+		Set<Rule> rulesToAdd = new HashSet<>();
 		InequalityHelper.load(this);
 		logger.info("Loaded inequalities.");
 		
@@ -147,7 +147,7 @@ public abstract class ConstraintChecker {
 		}
 		logger.info("Created " + rulesToAdd.size() + " rules.");
 		if (InequalityHelper.getMode().equals(InequalityHelper.Mode.ON_DEMAND)) {
-			List<Rule> demandRules = addRequireInequality(rulesToAdd);
+			Set<Rule> demandRules = addRequireInequality(rulesToAdd);
 			rulesToAdd.addAll(demandRules);
 			logger.info("Created " + demandRules.size() + " additional on-demand-rules.");
 		}
@@ -160,8 +160,8 @@ public abstract class ConstraintChecker {
 		reasoner.close();
 	}
 	
-	List<Rule> addRequireInequality(List<Rule> rules) {
-		List<Rule> result = new ArrayList<Rule>();
+	Set<Rule> addRequireInequality(Set<Rule> rules) {
+		Set<Rule> result = new HashSet<Rule>();
 		for (Rule rule : rules) {
 			List<Atom> unequalAtoms = new ArrayList<Atom>();
 			List<Atom> otherAtoms = new ArrayList<Atom>();
@@ -197,9 +197,9 @@ public abstract class ConstraintChecker {
 		return result;
 	}
 	
-	protected void prepareAndExecuteQueries(List<Rule> rules, Set<Atom> queries) throws IOException, PrepareQueriesException {
+	protected void prepareAndExecuteQueries(Set<Rule> rules, Set<Atom> queries) throws IOException, PrepareQueriesException {
 		try {
-			reasoner.addRules(rules);
+			reasoner.addRules(rules.toArray(new Rule[rules.size()]));
 		} catch (ReasonerStateException e) {
 			logger.error("Trying to add rules in the wrong state for constraint " + constraint + ".", e);
 			throw new PrepareQueriesException(internalError);

### 2018-09-19 src/main/java/impl/PCC/AllowedEntityTypesPCC.java

@@ -5,6 +5,7 @@ import static utility.SC.item;
 
 import java.io.IOException;
 import java.util.ArrayList;
+import java.util.HashSet;
 import java.util.List;
 import java.util.Set;
 
@@ -30,8 +31,8 @@ public class AllowedEntityTypesPCC extends PropertyConstraintChecker {
 	}
 
 	@Override
-	public List<Rule> rules() {
-		List<Rule> rules = new ArrayList<Rule>();
+	public Set<Rule> rules() {
+		Set<Rule> rules = new HashSet<Rule>();
 		
 		// item(I)
 		Atom item_I = Expressions.makeAtom(item, i);

### 2018-09-19 src/main/java/impl/PCC/AllowedQualifiersPCC.java

@@ -7,6 +7,7 @@ import static utility.SC.s;
 
 import java.io.IOException;
 import java.util.ArrayList;
+import java.util.HashSet;
 import java.util.List;
 import java.util.Set;
 
@@ -31,8 +32,8 @@ public class AllowedQualifiersPCC extends PropertyConstraintChecker {
 	}
 
 	@Override
-	public List<Rule> rules() { 	
-		List<Rule> rules = new ArrayList<Rule>();
+	public Set<Rule> rules() {
+		Set<Rule> rules = new HashSet<Rule>();
 		
 		// qualifierEDB(S, Q, O)
 		Atom qualfierEDB_SQO = Expressions.makeAtom(qualifierEDB, s, q, o);

### 2018-09-19 src/main/java/impl/PCC/AllowedUnitsPCC.java

@@ -8,6 +8,7 @@ import java.io.IOException;
 import java.util.ArrayList;
 import java.util.HashSet;
 import java.util.List;
+import java.util.Set;
 
 import org.apache.log4j.Logger;
 import org.semanticweb.vlog4j.core.model.api.Atom;
@@ -30,8 +31,8 @@ public class AllowedUnitsPCC extends PropertyConstraintChecker {
 	}
 
 	@Override
-	public List<Rule> rules() {
-		List<Rule> rules = new ArrayList<Rule>();
+	public Set<Rule> rules() {
+		Set<Rule> rules = new HashSet<Rule>();
 
 		List<Atom> unequal_conjunction = new ArrayList<Atom>();
 		// unit(V, U)

### 2018-09-19 src/main/java/impl/PCC/ConflictsWithPCC.java

@@ -10,6 +10,7 @@ import java.util.ArrayList;
 import java.util.HashSet;
 import java.util.List;
 import java.util.Map;
+import java.util.Set;
 
 import org.apache.log4j.Logger;
 import org.semanticweb.vlog4j.core.model.api.Atom;
@@ -31,8 +32,8 @@ public class ConflictsWithPCC extends PropertyConstraintChecker {
 	}
 
 	@Override
-	public List<Rule> rules() {	
-		List<Rule> rules = new ArrayList<Rule>();
+	public Set<Rule> rules() {
+		Set<Rule> rules = new HashSet<Rule>();
 		
 		for (Map.Entry<String, HashSet<String>> entry : conflicts.entrySet()) {
 			String confProperty = entry.getKey();

### 2018-09-19 src/main/java/impl/PCC/DistinctValuesPCC.java

@@ -8,7 +8,9 @@ import static utility.SC.x;
 
 import java.io.IOException;
 import java.util.ArrayList;
+import java.util.HashSet;
 import java.util.List;
+import java.util.Set;
 
 import org.apache.log4j.Logger;
 import org.semanticweb.vlog4j.core.model.api.Atom;
@@ -26,8 +28,8 @@ public class DistinctValuesPCC extends PropertyConstraintChecker {
 	}
 
 	@Override
-	public List<Rule> rules() {
-		List<Rule> rules = new ArrayList<Rule>();
+	public Set<Rule> rules() {
+		Set<Rule> rules = new HashSet<Rule>();
 		
 		// statementEDB(O, X, propertyConstant, V)
 		Atom statementEDB_OXpV = Expressions.makeAtom(statementEDB, o, x, propertyConstant, v);

### 2018-09-19 src/main/java/impl/PCC/InversePCC.java

@@ -14,6 +14,7 @@ import static utility.SC.x;
 
 import java.io.IOException;
 import java.util.ArrayList;
+import java.util.HashSet;
 import java.util.List;
 import java.util.Set;
 
@@ -36,8 +37,8 @@ public class InversePCC extends PropertyConstraintChecker {
 	}
 
 	@Override
-	public List<Rule> rules() {		
-		List<Rule> rules = new ArrayList<Rule>();
+	public Set<Rule> rules() {
+		Set<Rule> rules = new HashSet<Rule>();
 		
 		// statementEDB(Q, R, propertyConstant, I)
 		Atom statementEDB_QRpI = Expressions.makeAtom(statementEDB, q, r, propertyConstant, i);

### 2018-09-19 src/main/java/impl/PCC/ItemRequiresStatementPCC.java

@@ -13,6 +13,7 @@ import static utility.SC.x;
 
 import java.io.IOException;
 import java.util.ArrayList;
+import java.util.HashSet;
 import java.util.List;
 import java.util.Map;
 import java.util.Set;
@@ -37,8 +38,8 @@ public class ItemRequiresStatementPCC extends PropertyConstraintChecker {
 	}
 
 	@Override
-	public List<Rule> rules() {		
-		List<Rule> rules = new ArrayList<Rule>();
+	public Set<Rule> rules() {
+		Set<Rule> rules = new HashSet<Rule>();
 
 		// statementEDB(Q, I, propertyConstant, X)
 		Atom statementEDB_QIpX = Expressions.makeAtom(statementEDB, q, i, propertyConstant, x);

### 2018-09-19 src/main/java/impl/PCC/MandatoryQualifierPCC.java

@@ -13,6 +13,7 @@ import static utility.SC.v;
 
 import java.io.IOException;
 import java.util.ArrayList;
+import java.util.HashSet;
 import java.util.List;
 import java.util.Set;
 
@@ -37,8 +38,8 @@ public class MandatoryQualifierPCC extends PropertyConstraintChecker {
 	}
 
 	@Override
-	public List<Rule> rules() {
-		List<Rule> rules = new ArrayList<Rule>();
+	public Set<Rule> rules() {
+		Set<Rule> rules = new HashSet<Rule>();
 		
 		// statementEDB(S, I, propertyConstant, C)
 		Atom statementEDB_SIpC = Expressions.makeAtom(statementEDB, s, i, propertyConstant, c);

### 2018-09-19 src/main/java/impl/PCC/MultiValuePCC.java

@@ -18,7 +18,9 @@ import static utility.SC.x;
 
 import java.io.IOException;
 import java.util.ArrayList;
+import java.util.HashSet;
 import java.util.List;
+import java.util.Set;
 
 import org.apache.log4j.Logger;
 import org.semanticweb.vlog4j.core.model.api.Atom;
@@ -37,8 +39,8 @@ public class MultiValuePCC extends PropertyConstraintChecker {
 	}
 
 	@Override
-	public List<Rule> rules() {
-		List<Rule> rules = new ArrayList<Rule>();
+	public Set<Rule> rules() {
+		Set<Rule> rules = new HashSet<Rule>();
 		
 		// statementEDB(Q, I, propertyConstant, X)
 		Atom statementEDB_QIpX = Expressions.makeAtom(statementEDB, q, i, propertyConstant, x);

### 2018-09-19 src/main/java/impl/PCC/NoneOfPCC.java

@@ -12,6 +12,7 @@ import static utility.SC.violation_statement;
 
 import java.io.IOException;
 import java.util.ArrayList;
+import java.util.HashSet;
 import java.util.List;
 import java.util.Set;
 
@@ -35,8 +36,8 @@ public class NoneOfPCC extends PropertyConstraintChecker {
 	}
 
 	@Override
-	public List<Rule> rules() {
-		List<Rule> rules = new ArrayList<Rule>();
+	public Set<Rule> rules() {
+		Set<Rule> rules = new HashSet<Rule>();
 		
 		
 		for (String notAllowedValue : forbiddenValues) {

### 2018-09-19 src/main/java/impl/PCC/OneOfPCC.java

@@ -4,6 +4,7 @@ import static utility.SC.v;
 
 import java.io.IOException;
 import java.util.ArrayList;
+import java.util.HashSet;
 import java.util.List;
 import java.util.Set;
 
@@ -28,8 +29,8 @@ public class OneOfPCC extends PropertyConstraintChecker {
 	}
 
 	@Override
-	public List<Rule> rules() {	    	
-		List<Rule> rules = new ArrayList<Rule>();
+	public Set<Rule> rules() {
+		Set<Rule> rules = new HashSet<Rule>();
 		
 		List<Atom> unequal_conjunction = new ArrayList<Atom>();
 		for (String allowedValue : allowedValues) {

### 2018-09-19 src/main/java/impl/PCC/OneOfQualifierValuePCC.java

@@ -6,6 +6,7 @@ import static utility.SC.s;
 
 import java.io.IOException;
 import java.util.ArrayList;
+import java.util.HashSet;
 import java.util.List;
 import java.util.Map;
 import java.util.Set;
@@ -31,8 +32,8 @@ public class OneOfQualifierValuePCC extends PropertyConstraintChecker {
 	}
 
 	@Override
-	public List<Rule> rules() {  	
-		List<Rule> rules = new ArrayList<Rule>();
+	public Set<Rule> rules() {
+		Set<Rule> rules = new HashSet<Rule>();
 		
 		for (Map.Entry<String, Set<String>> entry : qualifiersAndValues.entrySet()) {
 			Constant qualifierPropertyConstant = Utility.makeConstant(entry.getKey());

### 2018-09-19 src/main/java/impl/PCC/PropertyConstraintChecker.java

@@ -18,6 +18,7 @@ import java.io.IOException;
 import java.util.ArrayList;
 import java.util.Arrays;
 import java.util.List;
+import java.util.Set;
 
 import org.apache.log4j.Logger;
 import org.semanticweb.vlog4j.core.model.api.Atom;
@@ -68,5 +69,5 @@ public abstract class PropertyConstraintChecker {
 		referenceEDB_SHpV = Expressions.makeAtom(referenceEDB, s, h, propertyConstant, v);
 	}
 	
-	public abstract List<Rule> rules();
+	public abstract Set<Rule> rules();
 }

### 2018-09-19 src/main/java/impl/PCC/ScopePCC.java

@@ -9,6 +9,7 @@ import static utility.SC.violation_statement_query;
 
 import java.io.IOException;
 import java.util.ArrayList;
+import java.util.HashSet;
 import java.util.List;
 import java.util.Set;
 
@@ -34,8 +35,8 @@ public class ScopePCC extends PropertyConstraintChecker {
 		allowedScopes = qualifiers_;
 	}
 	
-	public List<Rule> rules() {		
-		List<Rule> rules = new ArrayList<Rule>();
+	public Set<Rule> rules() {
+		Set<Rule> rules = new HashSet<Rule>();
 
 		// violation_statement(S, I, propertyConstant, V) :- statementEDB(S, I, propertyConstant, V)
 		Rule notStatement = Expressions.makeRule(violation_statement_SIpV, statementEDB_SIpV);

### 2018-09-19 src/main/java/impl/PCC/SingleBestValuePCC.java

@@ -1,8 +1,7 @@
 package impl.PCC;
 
 import static utility.SC.c;
-import static utility.SC.does_not_have;
-import static utility.SC.has_same;
+import static utility.SC.g;
 import static utility.SC.i;
 import static utility.SC.last_qualifier;
 import static utility.SC.o;
@@ -11,14 +10,18 @@ import static utility.SC.q;
 import static utility.SC.qualifierEDB;
 import static utility.SC.r;
 import static utility.SC.rank;
+import static utility.SC.referenceEDB;
 import static utility.SC.require_qualifier;
 import static utility.SC.s;
+import static utility.SC.same_or_non_existent;
 import static utility.SC.statementEDB;
 import static utility.SC.v;
 import static utility.SC.x;
+import static utility.SC.y;
 
 import java.io.IOException;
 import java.util.ArrayList;
+import java.util.HashSet;
 import java.util.List;
 import java.util.Set;
 
@@ -28,8 +31,6 @@ import org.semanticweb.vlog4j.core.model.api.Constant;
 import org.semanticweb.vlog4j.core.model.api.Rule;
 import org.semanticweb.vlog4j.core.model.implementation.Expressions;
 import org.wikidata.wdtk.datamodel.interfaces.StatementRank;
-
-import com.google.common.collect.Sets;
 
 import utility.InequalityHelper;
 import utility.StatementNonExistenceHelper;
@@ -46,33 +47,57 @@ public class SingleBestValuePCC extends PropertyConstraintChecker {
 	}
 
 	@Override
-	public List<Rule> rules() { 	
-		List<Rule> rules = new ArrayList<Rule>();
+	public Set<Rule> rules() {
+		Set<Rule> rules = new HashSet<Rule>();
 		
 		Constant preferredRankConstant = Expressions.makeConstant(StatementRank.PREFERRED.toString());
 		
 		// rank(S, preferredRankConstant)
 		Atom rank_Sp = Expressions.makeAtom(rank, s, preferredRankConstant);
 		
-		// statementEDB(O, I, propertyConstant, X)
-		Atom statementEDB_OIpX = Expressions.makeAtom(statementEDB, o, i, propertyConstant, x);
-		
 		// rank(O, preferredRankConstant)
[…]

### 2018-09-19 src/main/java/impl/PCC/SingleValuePCC.java

@@ -1,8 +1,6 @@
 package impl.PCC;
 
 import static utility.SC.c;
-import static utility.SC.does_not_have;
-import static utility.SC.has_same;
 import static utility.SC.g;
 import static utility.SC.i;
 import static utility.SC.last_qualifier;
@@ -14,12 +12,15 @@ import static utility.SC.r;
 import static utility.SC.referenceEDB;
 import static utility.SC.require_qualifier;
 import static utility.SC.s;
+import static utility.SC.same_or_non_existent;
 import static utility.SC.statementEDB;
 import static utility.SC.v;
 import static utility.SC.x;
+import static utility.SC.y;
 
 import java.io.IOException;
 import java.util.ArrayList;
+import java.util.HashSet;
 import java.util.List;
 import java.util.Set;
 
@@ -28,8 +29,6 @@ import org.semanticweb.vlog4j.core.model.api.Atom;
 import org.semanticweb.vlog4j.core.model.api.Constant;
 import org.semanticweb.vlog4j.core.model.api.Rule;
 import org.semanticweb.vlog4j.core.model.implementation.Expressions;
-
-import com.google.common.collect.Sets;
 
 import utility.InequalityHelper;
 import utility.StatementNonExistenceHelper;
@@ -46,8 +45,8 @@ public class SingleValuePCC extends PropertyConstraintChecker {
 	}
 
 	@Override
-	public List<Rule> rules() { 	
-		List<Rule> rules = new ArrayList<Rule>();
+	public Set<Rule> rules() {
+		Set<Rule> rules = new HashSet<Rule>();
 		
 		// statementEDB(O, I, propertyConstant, X)
 		Atom statementEDB_OIpX = Expressions.makeAtom(statementEDB, o, i, propertyConstant, x);
@@ -88,7 +87,7 @@ public class SingleValuePCC extends PropertyConstraintChecker {
 			// 	unequal(V, X)
 			Rule violationReference = Expressions.makeRule(violation_reference_SHpV, referenceEDB_SHpV, referenceEDB_SGpX, unequal_VX);
 			rules.add(violationReference);
-		} else {
+		} else {			
 			// statementEDB(S, I, propertyConstant, C)
 			Atom statementEDB_SIpC = Expressions.makeAtom(statementEDB, s, i, propertyConstant, c);
 			
[…]

### 2018-09-19 src/main/java/impl/PCC/ValueRequiresStatementPCC.java

@@ -14,6 +14,7 @@ import static utility.SC.x;
 
 import java.io.IOException;
 import java.util.ArrayList;
+import java.util.HashSet;
 import java.util.List;
 import java.util.Map;
 import java.util.Set;
@@ -38,8 +39,8 @@ public class ValueRequiresStatementPCC extends PropertyConstraintChecker {
 	}
 
 	@Override
-	public List<Rule> rules() {
-		List<Rule> rules = new ArrayList<Rule>();
+	public Set<Rule> rules() {
+		Set<Rule> rules = new HashSet<Rule>();
 		
 		// statementEDB(Q, R, propertyConstant, I)
 		Atom statementEDB_QRpI = Expressions.makeAtom(statementEDB, q, r, propertyConstant, i);

### 2018-09-19 src/main/java/utility/SC.java

@@ -16,6 +16,7 @@ public class SC {
 	public final static String O = "o";
 	public final static String I = "i";
 	public final static String X = "x";
+	public final static String Y = "y";
 	public final static String P = "p";
 	public final static String Q = "q";
 	public final static String V = "v";
@@ -53,8 +54,7 @@ public class SC {
 	public final static String UNIT = "unit";
 	public final static String RANK = "rank";
 	
-	public final static String DOES_NOT_HAVE = "does_not_have";
-	public final static String HAS_SAME = "has_same";	
+	public final static String SAME_OR_NON_EXISTENT = "same_or_non_existent";
 	
 	public final static Variable h = Expressions.makeVariable(H);
 	public final static Variable g = Expressions.makeVariable(G);
@@ -63,6 +63,7 @@ public class SC {
 	public final static Variable o = Expressions.makeVariable(O);
 	public final static Variable i = Expressions.makeVariable(I);
 	public final static Variable x = Expressions.makeVariable(X);
+	public final static Variable y = Expressions.makeVariable(Y);
 	public final static Variable p = Expressions.makeVariable(P);
 	public final static Variable q = Expressions.makeVariable(Q);
 	public final static Variable v = Expressions.makeVariable(V);
@@ -120,10 +121,8 @@ public class SC {
 	// id, rank -> rank, id
 	public final static Predicate rank = new PositionPredicate(RANK, 2, 1, 0);
 	
-	// id, property -> property, id
-	public final static Predicate does_not_have = new PositionPredicate(DOES_NOT_HAVE, 2, 1, 0);
-	// id1, id2, property -> property, id1, id2
-	public final static Predicate has_same = new PositionPredicate(HAS_SAME, 3, 1, 2, 0);
+	// id1, id2, qualifier -> qualifier, id1, id2
+	public final static Predicate same_or_non_existent = new PositionPredicate(SAME_OR_NON_EXISTENT, 3, 1, 2, 0);
 	
 	public final static Atom violation_statement_query = Expressions.makeAtom(violation_statement, s, i, p, v);
 	public final static Atom violation_qualifier_query = Expressions.makeAtom(violation_qualifier, s, p, v);

### 2018-09-20 thesis/paper/paper.tex
Started correcting formatting

@@
 \setcounter{tocdepth}{2}
+\expandafter\def\expandafter\normalsize\expandafter{%
+    \normalsize
+    \setlength\abovedisplayskip{3pt}
+    \setlength\belowdisplayskip{3pt}
+    \setlength\abovedisplayshortskip{3pt}
+    \setlength\belowdisplayshortskip{3pt}
+}
 
@@
 It depends on a conjunction (\(O\) denoting another statement ID, \(C\) another value) and a violation for a conflicts-with constraint is thus either
-\begin{equation*}
+\begin{equation}
 \begin{split}
@@
 \end{split}
-\end{equation*}
+\end{equation}
 for constraints with no conflicting values specified or one
-\begin{equation*}
+\begin{equation}
 \begin{split}
@@
 \end{split}
-\end{equation*}
+\end{equation}
 for each conflicting value if specified.\\
@@
 The violations depend on a simple conjunction:
-\begin{equation*}
+\begin{equation}
 \begin{split}
@@
 \end{split}
-\end{equation*}
-\begin{equation*}
+\end{equation}
+\begin{equation}
 \begin{split}
@@
 \end{split}
-\end{equation*}
+\end{equation}
 The first rule are created for all \emph{allowed entity type} constraints whose \emph{allowed entity type} does not contain \emph{Wikibase item}.
@@
 The rules, depending on nothing but specific database entries, are
-\begin{equation*}
+\begin{equation}
 \begin{split}
@@
 \end{split}
-\end{equation*}
+\end{equation}
 Given a specified constraint, each of the three rules is created once for each \emph{forbidden value}.
@@
 This leads to the rule
-\begin{equation*}
+\begin{equation}
 \text{unequal}(X, Y) \leftarrow \text{unequal\_EDB}(X, Y)
-\end{equation*}
+\end{equation}
 
@@
 The naive approach would be to simply add all possible combinations:
-\begin{equation*}
+\begin{equation}
 \text{for all } x, y \in U, x \ne y \text{ add } \text{unequal\_EDB}(x, y)
-\end{equation*}
+\end{equation}
 This can be easily improved using the inverse and adding the following rule:
@@
 Then it is sufficient to add each combination once. To achieve this assume an arbitrary total order \(<\) has been established on the set \(E\):
-\begin{equation*}
+\begin{equation}
 \text{for all } x, y \in U, x < y \text{ add } \text{unequal\_EDB}(x, y)
-\end{equation*}
+\end{equation}
 
@@
 For each unequal constant string \(u = c_0 \ldots c_{l - 1}\) and \(0 \equiv i \mod 7, 0 \leq i < len(u)\) an atom using predicate letter\(_i\)\_EDB is created as 
-\begin{equation*}
+\begin{equation}
 \text{letter}_i\text{\_EDB}(e, f(e, i), f(e, i+1), \ldots, f(e, i+6)
-\end{equation*}where
-\begin{equation*}
+\end{equation}where
+\begin{equation}
 f(e, i) = 
@@
 \end{cases}
-\end{equation*}
+\end{equation}
 
 Now that inequality between all characters and the link between entities and their characters are introduced, the inequality between two entities can be derived by creating the following rules for all predicates letter\(_i\)\_EDB (with \text{letters} short for \(\text{letter}_i\text{\_EDB}(X, A_i, \ldots, A_{i+6}) \wedge\text{letter}_i\text{\_EDB}(Y, B_i, \ldots, B_{i+6})\), \(i \leq j < i + 7\)):
-\begin{equation*}
+\begin{equation}
 \text{unequal}(X, Y) \leftarrow \text{letters} \wedge \text{unequal}(A_j, B_j)
-\end{equation*}
+\end{equation}
 Note that the facts fulfilling \(\text{unequal}(A_j, B_j)\) are created by using the approach outlined in Section \ref{subsubsec_naive} for the set of characters \(C\) in all unequal constants \(U\).
@@
 I_n\) are atoms of the form \(\text{unequal}(t_{n_1}, t_{n_2})\) with \(t_{n_1}, t_{n_2}\) terms and all variables amount \(t_{n_1}\) and \(t_{n_2}\) appear in at least one of \(B_1 \ldots B_m\), the rules
-\begin{equation*}
+\begin{equation}
 \text{req\_inequality}(t_{n_1}, t_{n_2})
  \leftarrow B_1 \wedge \ldots \wedge B_m
-\end{equation*} for all \(n\) are introduced.\\
+\end{equation} for all \(n\) are introduced.\\
 
 The rules outlined in Section \ref{subsubsec_encoding} are changed to utilise this new requirement.
-\begin{equation*}
+\begin{equation}
 \text{unequal}(X, Y) \leftarrow \text{req\_inequality}	(X, Y) \wedge \text{letters} \wedge \text{unequal}(A_j, B_j)
-\end{equation*}
+\end{equation}
 
@@
 Since this would mark every constrained statement as a violation because it can fulfil both atoms at the same time, inequality of the statement IDs \(S\) and \(O\) must be tested as well.
-\begin{equation*}
+\begin{equation}
 \begin{split}
@@
 \end{split}
-\end{equation*}
+\end{equation}
 The rules for qualifiers and reference are analogous.
@@
 To determine if a violation occurred, the unit of the constrained statement needs to be found. This is done via the unit-predicate as below
-\begin{equation*}
+\begin{equation}
 \text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge \text{unit\_EDB}(V, U)
-\end{equation*}
+\end{equation}
 with qualifiers and references treated analogous.
@@
 Violations are thus
-\begin{equation*}
+\begin{equation}
 \begin{split}
@@
 \end{split}
-\end{equation*}
+\end{equation}
 
@@
 With \(A\) the set of allowed qualifiers the rule is:
-\begin{equation*}
+\begin{equation}
 \begin{split}
@@
 With the set of allowed Values \(A\) the rules are
-\begin{equation*}
+\begin{equation}
 \begin{split}
@@
 Using this pre-processed data, the approach is to apply a require-predicate to each statement or qualifier.
-This is require\([3]\) for statements, connecting the statement ID with the constrained property \(p\) and the required term \(t\) and require\([5]\) taking the statement ID, property, and value of a qualifier and linking this with \(p\) and \(t\).\\
-
-For statements the pattern consists of two rules, both utilising the same conjunction \(c = C_1 \wedge \ldots \wedge C_m\) with atoms \(C_i\) depending on the constraint.
+This is require\([2]\) for statements, connecting the statement ID with an identifier unique to the declared constraint and require\([4]\) taking the statement ID, property, and value of a qualifier and linking this with the unique identifier.\\
+
+\todo{Move qualifier require to single value}
+
+For statements the pattern consists of two rules, both utilising the same conjunction \(c\) whose atoms depend on the constraint type.
 This conjunction is used to test if the statement does not fulfil the requirement.
-Then rule one
-\begin{equation*}
-\text{require}(S, p, t) \leftarrow \text{first}(S, I) \wedge c
-\end{equation*}
-states that the statement requires the condition and
-\begin{equation*}
-\text{require}(S, p, t) \leftarrow \text{next}(O, S) \wedge \text{require}(O, p, t) \wedge c
-\end{equation*}
-propagates this down the statement chain.
-
-For qualifiers the rules are
-\begin{equation*}
-\text{require}(S, P, V, p, t) \leftarrow \text{first}(S, P, V) \wedge c
-\end{equation*}
-\begin{equation*}
-\text{require}(S, P, V, p, t) \leftarrow \text{next}(O, X, C, S, P, V) \wedge \text{require}(O, X, C, p, t) \wedge c
-\end{equation*}
+Then rule one states that the statement requires the condition and rule two propagates this down the statement chain.
+For qualifiers the rules three and four apply.
+\begin{align}
+&\text{require}(S, \text{id}) \leftarrow \text{first}(S, I) \wedge c \\
+&\text{require}(S, \text{id}) \leftarrow \text{next}(O, S) \wedge \text{require}(O, \text{id}) \wedge c \\
+&\text{require}(S, P, V, \text{id}) \leftarrow \text{first}(S, P, V) \wedge c \\
+&\text{require}(S, P, V, \text{id}) \leftarrow \text{next}(O, X, C, S, P, V) \wedge \text{require}(O, X, C, \text{id}) \wedge c
+\end{align}
 
@@
 An item violates the constraint if it a) has a statement using the constrained property as predicate, b1) has no statements using the \emph{required property} as predicate, or (if \emph{allowed values} are specified) b2) has only statement using the \emph{required property} as predicate that do not use one of the \emph{allowed values} as value.
-The required term \(t\) is the \emph{required property}.
-Without separators the conjunction is
-\begin{equation*}
+If no \emph{allowed values} have been specified for the declared constraint the conjunction for the require pattern is [one].
+If a set of \emph{allowed values} has been specified, a second set of require-rules have to be introduced with conjunction [two].
+Note that the require pattern using conjunction [one] is still required, since the preceding statements do not need to have the \emph{required property}.
+In both cases violations can be found with rule [three]
+
+\begin{align}
 \begin{split}
@@
 &\text{statement\_EDB}(S, I, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
-\end{split}
-\end{equation*} \(\)
-If a set of \emph{allowed values} \(A\) has been specified, a second set of require-rules have to be introduced with
-\begin{equation*}
+\end{split} \\
 \begin{split}
@@
 \end{split}
-\end{equation*}
-Note that the first set of rules is still required, since the preceding statements do not need to have the \emph{required property}. In both cases violations can be found with
-\begin{equation*}
+\end{align}
+\begin{align}
 \begin{split}
@@
 \end{split}
-\end{equation*}
+\end{align}
 
@@
 The conjunctions change in the first statement and are:
-\begin{equation*}
+\begin{equation}
 \begin{split}
@@
 \end{split}
-\end{equation*} \(\)
-\begin{equation*}
+\end{equation} \(\)
+\begin{equation}
 \begin{split}
@@
 \end{split}
-\end{equation*}
+\end{equation}
 The final violation rule would then be:
-\begin{equation*}
+\begin{equation}
 \begin{split}
@@
 Inequalities have to be established between all properties and the \emph{inverse} property as well as all subjects and values of statements.
-\begin{equation*}
+\begin{equation}
 \begin{split}
@@
 \end{split}
-\end{equation*}
+\end{equation}
 The final violation rule remains the same.
@@
 Inequality is established between all properties of statements and the conjunction is:
-\begin{equation*}
+\begin{equation}
 \begin{split}
@@
 \end{split}
-\end{equation*}
+\end{equation}
 The predicate require\_second\([2]\) is introduced with the same meaning as require except that it will be applied after a required property has been found:
-\begin{equation*}
+\begin{equation}
 \begin{split}
@@
 \end{split}
-\end{equation*}
+\end{equation}
 A violation would be any last statement requiring a second constrained statement.\\
@@
 However, since the statement ID and the property of a possible conflict are always the same, inequality can be tested on the values.
-\begin{equation*}
+\begin{equation}
 \begin{split}
@@
 The conjunction c is created for every \emph{separator} and the pattern applied.
-\begin{equation*}
+\begin{equation}
 \begin{split}
@@
 Note that \(\text{predicate}(\{X\})\) denotes \(\bigwedge_{x \in X} \text{predicate}(x)\) as introduced in Section \ref{subsec_3_allowed_units}.
-\begin{equation*}
+\begin{equation}
 \begin{split}
@@
 The conjunction for the require-pattern is:
-\begin{equation*}
+\begin{equation}
 \begin{split}
@@
 \end{split}
-\end{equation*}
+\end{equation}
 With the usual require-pattern for qualifiers and the does\_not\_have rule as introduced in Section \ref{subsubsec_single_value_with_separators} added for brevity, the violations would be found with:
-\begin{equation*}
+\begin{equation}
 \begin{split}
@@
 Thus the rule \eqref{eq_no_separators_triple} is expanded to
-\begin{equation*}
+\begin{equation}
 \begin{split}
@@
 As in the case above, the qualifier non-existence and violation rules are very similar to the \emph{single value} constraint. The only change is the addition of the rank requirements as shown below.
-\begin{equation*}
+\begin{equation}
 \begin{split}

### 2018-09-20 thesis/paper/paper.tex
Formatted formulas

@@
 
-\subsection{Scope (Q53869507)}
+\subsection{Property scope (Q53869507)}
 \paragraph{Constraint}
@@
 \paragraph{Rules}
-The property scope constraint requires no context information to be violated; the triple, qualifier, or reference with the constrained property is enough.
+The \emph{property scope} constraint requires no context information to be violated; the triple, qualifier, or reference with the constrained property is enough.
 This means violations can be found using these rules:
-\begin{align*}
-\text{violation\_statement}(S, I, \text{propertyConstant}, V) &\leftarrow \text{statement\_EDB}(S, I, \text{propertyConstant}, V)\\
-\text{violation\_qualifier}(S, \text{propertyConstant}, V) &\leftarrow \text{qualifier\_EDB}(S, \text{propertyConstant}, V)\\
-\text{violation\_reference}(S, \text{propertyConstant}, V) &\leftarrow \text{reference\_EDB}(S, H, \text{propertyConstant}, V)
-\end{align*}
+\begin{align}
+\begin{split}
+\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+&\text{statement\_EDB}(S, I, \text{propertyConstant}, V)\\
+\end{split}\\
+\begin{split}
+\text{violation\_qualifier}&(S, \text{propertyConstant}, V) \leftarrow \\
+&\text{qualifier\_EDB}(S, \text{propertyConstant}, V)\\
+\end{split}\\
+\begin{split}
+\text{violation\_reference}&(S, \text{propertyConstant}, V) \leftarrow \\
+&\text{reference\_EDB}(S, H, \text{propertyConstant}, V)\\
+\end{split}
+\end{align}
 Which of these rules need to be used depends on the values of the \emph{property scope} qualifier.
@@
 The conflicts-with constraint is violated if another statement on the same item either has a conflicting property or a conflicting property and a conflicting value depending on specified qualifiers.
-It depends on a conjunction (\(O\) denoting another statement ID, \(C\) another value) and a violation for a conflicts-with constraint is thus either
-\begin{equation}
+It depends on a conjunction (\(O\) denoting another statement ID, \(C\) another value) and a violation for a conflicts-with constraint is thus either rule one for constraints with no conflicting values specified or one instance of rule two for each conflicting value if specified.
+\begin{align}
 \begin{split}
@@
 &\text{statement\_EDB}(O, I, \text{conflictingPropertyConstant}, C)
-\end{split}
-\end{equation}
-for constraints with no conflicting values specified or one
-\begin{equation}
+\end{split}\\
 \begin{split}
@@
 \end{split}
-\end{equation}
-for each conflicting value if specified.\\
+\end{align}
 
@@
 The violations depend on a simple conjunction:
-\begin{equation}
+\begin{align}
 \begin{split}
@@
 &\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge \text{item}(I)
-\end{split}
-\end{equation}
-\begin{equation}
+\end{split}\\
 \begin{split}
@@
 \end{split}
-\end{equation}
-The first rule are created for all \emph{allowed entity type} constraints whose \emph{allowed entity type} does not contain \emph{Wikibase item}.
+\end{align}
+The first rule is created for all \emph{allowed entity type} constraints whose \emph{allowed entity type} does not contain \emph{Wikibase item}.
 The second rule works analogous for \emph{Wikibase property}.
@@
 The rules, depending on nothing but specific database entries, are
-\begin{equation}
+\begin{align}
 \begin{split}
@@
 &\text{statement\_EDB}(S, I, \text{propertyConstant}, \text{forbiddenValueConstant})
-\end{split}
-\end{equation}
-\begin{equation}
+\end{split}\\
 \begin{split}
@@
 &\text{qualifier\_EDB}(S, \text{propertyConstant}, \text{forbiddenValueConstant})
-\end{split}
-\end{equation}
-\begin{equation}
+\end{split}\\
 \begin{split}
@@
 \end{split}
-\end{equation}
+\end{align}
 Given a specified constraint, each of the three rules is created once for each \emph{forbidden value}.
@@
 This leads to the rule
-\begin{equation}
+\begin{align}
 \text{unequal}(X, Y) \leftarrow \text{unequal\_EDB}(X, Y)
-\end{equation}
+\end{align}
 
@@
 The naive approach would be to simply add all possible combinations:
-\begin{equation}
+\begin{align}
 \text{for all } x, y \in U, x \ne y \text{ add } \text{unequal\_EDB}(x, y)
-\end{equation}
+\end{align}
 This can be easily improved using the inverse and adding the following rule:
-\begin{equation}\label{equ_inverse}
+\begin{align}\label{equ_inverse}
 \text{unequal}(X, Y) \leftarrow \text{unequal}(Y, X)
-\end{equation}
+\end{align}
 Then it is sufficient to add each combination once. To achieve this assume an arbitrary total order \(<\) has been established on the set \(E\):
-\begin{equation}
+\begin{align}
 \text{for all } x, y \in U, x < y \text{ add } \text{unequal\_EDB}(x, y)
-\end{equation}
+\end{align}
 
@@
 For each unequal constant string \(u = c_0 \ldots c_{l - 1}\) and \(0 \equiv i \mod 7, 0 \leq i < len(u)\) an atom using predicate letter\(_i\)\_EDB is created as 
-\begin{equation}
+\begin{align}
+\begin{split}
 \text{letter}_i\text{\_EDB}(e, f(e, i), f(e, i+1), \ldots, f(e, i+6)
-\end{equation}where
-\begin{equation}
+\end{split}
+\begin{split}
 f(e, i) = 
@@
 \end{cases}
-\end{equation}
+\end{split}
+\end{align}
 
 Now that inequality between all characters and the link between entities and their characters are introduced, the inequality between two entities can be derived by creating the following rules for all predicates letter\(_i\)\_EDB (with \text{letters} short for \(\text{letter}_i\text{\_EDB}(X, A_i, \ldots, A_{i+6}) \wedge\text{letter}_i\text{\_EDB}(Y, B_i, \ldots, B_{i+6})\), \(i \leq j < i + 7\)):
-\begin{equation}
+\begin{align}
 \text{unequal}(X, Y) \leftarrow \text{letters} \wedge \text{unequal}(A_j, B_j)
-\end{equation}
+\end{align}
 Note that the facts fulfilling \(\text{unequal}(A_j, B_j)\) are created by using the approach outlined in Section \ref{subsubsec_naive} for the set of characters \(C\) in all unequal constants \(U\).
@@
 B_m\) are arbitrary atoms, \(I_1 ..
-I_n\) are atoms of the form \(\text{unequal}(t_{n_1}, t_{n_2})\) with \(t_{n_1}, t_{n_2}\) terms and all variables amount \(t_{n_1}\) and \(t_{n_2}\) appear in at least one of \(B_1 \ldots B_m\), the rules
-\begin{equation}
+I_n\) are atoms of the form \(\text{unequal}(t_{n_1}, t_{n_2})\) with \(t_{n_1}, t_{n_2}\) terms and all variables amount \(t_{n_1}\) and \(t_{n_2}\) appear in at least one of \(B_1 \ldots B_m\), rule one is introduced for all \(n\).
+The rules outlined in Section \ref{subsubsec_encoding} are changed to utilise this new requirement as shown in rule two.
+\begin{align}
+\begin{split}
 \text{req\_inequality}(t_{n_1}, t_{n_2})
  \leftarrow B_1 \wedge \ldots \wedge B_m
-\end{equation} for all \(n\) are introduced.\\
-
-The rules outlined in Section \ref{subsubsec_encoding} are changed to utilise this new requirement.
-\begin{equation}
+\end{split}\\
+\begin{split}
 \text{unequal}(X, Y) \leftarrow \text{req\_inequality}	(X, Y) \wedge \text{letters} \wedge \text{unequal}(A_j, B_j)
-\end{equation}
+\end{split}
+\end{align}
 
@@
 Since this would mark every constrained statement as a violation because it can fulfil both atoms at the same time, inequality of the statement IDs \(S\) and \(O\) must be tested as well.
-\begin{equation}
+The rules for qualifiers and reference are analogous.
+\begin{align}
 \begin{split}
@@
 \end{split}
-\end{equation}
-The rules for qualifiers and reference are analogous.
+\end{align}
 
@@
 
-To determine if a violation occurred, the unit of the constrained statement needs to be found. This is done via the unit-predicate as below
-\begin{equation}
-\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge \text{unit\_EDB}(V, U)
-\end{equation}
-with qualifiers and references treated analogous.
-
-The complete rules require multiple unequal-atoms for all allowed units.
-Given a set \(A\) of constants, in this case allowed units, \(\text{unequal}(\{A\}, t)\) with an arbitrary term \(t\) is short for \(\bigwedge_{a \in A} \text{unequal}(a, t)\)
-
-Violations are thus
-\begin{equation}
+The rules require multiple unequal-atoms for all allowed units in the set \(A\) of allowed units. Violations are thus
+\begin{align}
 \begin{split}
@@
 &\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
-&\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
-\end{split}
-\end{equation}
-\begin{equation}
+&\text{unit}(V, U)  {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, U)
+\end{split}\\
 \begin{split}
@@
 &\text{qualifier\_EDB}(S, \text{propertyConstant}, V) \wedge{} \\
-&\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
-\end{split}
-\end{equation}
-\begin{equation}
+&\text{unit}(V, U) {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, U)
+\end{split}\\
 \begin{split}
@@
 &\text{reference\_EDB}(S, H, \text{propertyConstant}, V) \wedge{} \\
-&\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
-\end{split}
-\end{equation}
+&\text{unit}(V, U) {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, U)
+\end{split}
+\end{align}
 
@@
 With \(A\) the set of allowed qualifiers the rule is:
-\begin{equation}
+\begin{align}
 \begin{split}
@@
 &\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
-&\text{qualifier\_EDB}(S, Q, O) \wedge \text{unequal}(\{A\}, Q)
-\end{split}
-\end{equation}
+&\text{qualifier\_EDB}(S, Q, O) {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, Q)
+\end{split}
+\end{align}
 
@@
 With \(A\) the set of allowed values the rules for triple, qualifier, and reference are:
-\begin{align*}
-\text{violation\_statement}(S, I, \text{propertyConstant}, V) &\leftarrow \\
-\text{statement\_EDB}(S, I, &\text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V) \\
-\text{violation\_qualifier}(S, \text{propertyConstant}, V) &\leftarrow \\
-\text{qualifier\_EDB}(S, &\text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V) \\
-\text{violation\_reference}(S, \text{propertyConstant}, V) &\leftarrow \\
-\text{reference\_EDB}(S, H, &\text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
-\end{align*}
+\begin{align}
+\begin{split}
+\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, V)
+\end{split}\\
+\begin{split}
+\text{violation\_qualifier}&(S, \text{propertyConstant}, V) \leftarrow \\
+&\text{qualifier\_EDB}(S, \text{propertyConstant}, V) {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, V)
+\end{split}\\
+\begin{split}
+\text{violation\_reference}&(S, \text{propertyConstant}, V) \leftarrow \\
+&\text{reference\_EDB}(S, H, \text{propertyConstant}, V) {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, V)
+\end{split}
+\end{align}
 
@@
 With the set of allowed Values \(A\) the rules are
-\begin{equation}
+\begin{align}
 \begin{split}
@@
 &\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
-&\text{qualifier\_EDB}(S, \text{qualifierProperty}, O) \wedge \text{unequal}(\{A\}, O)
-\end{split}
-\end{equation}
+&\text{qualifier\_EDB}(S, \text{qualifierProperty}, O) {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, O)
+\end{split}
+\end{align}
 
@@
 
-\todo{Move qualifier require to single value}
-
 For statements the pattern consists of two rules, both utilising the same conjunction \(c\) whose atoms depend on the constraint type.
@@
 An item violates the constraint if it a) has a statement using the constrained property as predicate, b1) has no statements using the \emph{required property} as predicate, or (if \emph{allowed values} are specified) b2) has only statement using the \emph{required property} as predicate that do not use one of the \emph{allowed values} as value.
-If no \emph{allowed values} have been specified for the declared constraint the conjunction for the require pattern is [one].
-If a set of \emph{allowed values} has been specified, a second set of require-rules have to be introduced with conjunction [two].
-Note that the require pattern using conjunction [one] is still required, since the preceding statements do not need to have the \emph{required property}.
-In both cases violations can be found with rule [three]
+If no \emph{allowed values} have been specified for the declared constraint the conjunction for the require pattern is one.
+If a set of \emph{allowed values} has been specified, a second set of require-rules have to be introduced with conjunction two.
+Note that the require pattern using conjunction one is still required, since the preceding statements do not need to have the \emph{required property}.
+In both cases violations can be found with rule three
 
@@
 Inequality needs to be computed between all properties and all values and the item statement order established.
-The conjunctions change in the first statement and are:
-\begin{equation}
+The conjunctions change in the first statement and rule one and two.
+The final violation rule would be rule three:
+\begin{align}
 \begin{split}
@@
 &\text{statement\_EDB}(S, I, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
-\end{split}
-\end{equation} \(\)
-\begin{equation}
+\end{split}\\
 \begin{split}
@@
 \end{split}
-\end{equation}
-The final violation rule would then be:
-\begin{equation}
+\end{align}
+
+\begin{align}
 \begin{split}
@@
 \end{split}
-\end{equation}
+\end{align}
 
@@
 Inequalities have to be established between all properties and the \emph{inverse} property as well as all subjects and values of statements.
-\begin{equation}
+The final violation rule remains the same.
+\begin{align}
 \begin{split}
@@
 &\text{unequal}(\text{inversePropertyConstant}, P)
-\end{split}
-\end{equation}
-\begin{equation}
+\end{split}\\
 \begin{split}
@@
 \end{split}
-\end{equation}
-The final violation rule remains the same.
+\end{align}
 
@@
 \end{equation}
-The predicate require\_second\([2]\) is introduced with the same meaning as require except that it will be applied after a required property has been found:
-\begin{equation}
+The predicate require\_second\([2]\) is introduced with the same meaning as require except that it will be applied after a required property has been found.
+A violation would be any last statement requiring a second constrained statement, analogous to the violation rule for \emph{item requires statement} (see Section \ref{subsec_item_requires_statement}).
+\begin{align}
 \begin{split}
@@
 &\text{next}(Q, R) \wedge \text{require}(Q, \text{propertyConstant}, \text{propertyConstant})  \wedge c
-\end{split}
-\end{equation}
-\begin{equation}
+\end{split}\\
 \begin{split}
@@
 \end{split}
-\end{equation}
-A violation would be any last statement requiring a second constrained statement.\\
+\end{align}
 
@@
 If the separator set is empty, the violation simply occurs if there are two constrained statements on the same item.
-\begin{equation}\label{eq_no_separators_triple}
-\begin{split}
+
+For qualifiers and references, there is no ID whose inequality could be checked.
+However, since the statement ID and the property of a possible conflict are always the same, inequality can be tested on the values.
+\begin{align}
+\begin{split}\label{eq_no_separators_triple}
 \text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
@@
 &\text{unequal}(S, O)
-\end{split}
-\end{equation}
-For qualifiers and references, there is no ID whose inequality could be checked.
-However, since the statement ID and the property of a possible conflict are always the same, inequality can be tested on the values.
-\begin{equation}
+\end{split}\\
 \begin{split}
@@
 &\text{unequal}(V, X)
-\end{split}
-\end{equation}
-\begin{equation}
+\end{split}\\
 \begin{split}
@@
 Note that \(\text{predicate}(\{X\})\) denotes \(\bigwedge_{x \in X} \text{predicate}(x)\) as introduced in Section \ref{subsec_3_allowed_units}.
-\begin{equation}
+\begin{align}
 \begin{split}
@@
 &\text{same\_or\_non\_existent}(S, O, \{Q\})
-\end{split}
-\end{equation}
-\begin{equation}
+\end{split}\\
 \begin{split}
@@
 &\text{qualifier\_EDB}(O, Q, V)
-\end{split}
-\end{equation}
-\begin{equation}
+\end{split}\\
 \begin{split}
@@
 The qualifier order needs to be provided and inequality must be established between all qualifier properties.
-The conjunction for the require-pattern is:
-\begin{equation}
+The conjunction for the require-pattern is rule one.
+With the usual require-pattern for qualifiers and the does\_not\_have rule as introduced in Section \ref{subsubsec_single_value_with_separators} added for brevity, the violations would be found with rule two.
+\begin{align}
 \begin{split}
@@
 &\text{qualifier\_EDB}(S, P, V) \wedge \text{unequal}(P, \text{requiredPropertyConstant})
-\end{split}
-\end{equation}
-With the usual require-pattern for qualifiers and the does\_not\_have rule as introduced in Section \ref{subsubsec_single_value_with_separators} added for brevity, the violations would be found with:
-\begin{equation}
+\end{split}\\
 \begin{split}
@@
 Thus the rule \eqref{eq_no_separators_triple} is expanded to
-\begin{equation}
+\begin{align}
 \begin{split}
@@
 As in the case above, the qualifier non-existence and violation rules are very similar to the \emph{single value} constraint. The only change is the addition of the rank requirements as shown below.
-\begin{equation}
+\begin{align}
 \begin{split}
@@
 The exceptions are \emph{distinct values} (1.22 hours), \emph{item requires statement} (2.38 hours) and \emph{single value} (4.22 hours).
-In part this is expected behaviour, since these three constraints are applied to an order of magnitude more properties than all other properties with the exception of \emph{scope}, which is described in very simple rules.
+In part this is expected behaviour, since these three constraints are applied to an order of magnitude more properties than all other properties with the exception of \emph{property scope}, which is described in very simple rules.
 However, also seen in Figure \ref{fig_runtime}, the runtimes of these three constraints are multiple orders of magnitude higher than those of other constraints.

### 2018-09-20 thesis/paper/paper.tex
Changed I for item to E for entity

@@
 \begin{split}
-\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, I, \text{propertyConstant}, V)\\
+\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
+&\text{statement\_EDB}(S, E, \text{propertyConstant}, V)\\
 \end{split}\\
@@
 \begin{split}
-\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow\\
-&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{}\\
-&\text{statement\_EDB}(O, I, \text{conflictingPropertyConstant}, C)
-\end{split}\\
-\begin{split}
-\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow\\
-&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{}\\
-&\text{statement\_EDB}(O, I, \text{conflictingPropertyConstant}, \text{conflictingValueConstant})
+\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow\\
+&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{}\\
+&\text{statement\_EDB}(O, E, \text{conflictingPropertyConstant}, C)
+\end{split}\\
+\begin{split}
+\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow\\
+&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{}\\
+&\text{statement\_EDB}(O, E, \text{conflictingPropertyConstant}, \text{conflictingValueConstant})
 \end{split}
@@
 \begin{split}
-\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge \text{item}(I)
-\end{split}\\
-\begin{split}
-\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge \text{property}(I)
+\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
+&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge \text{item}(I)
+\end{split}\\
+\begin{split}
+\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
+&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge \text{property}(I)
 \end{split}
@@
 \begin{split}
-\text{violation\_statement}&(S, I, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
-&\text{statement\_EDB}(S, I, \text{propertyConstant}, \text{forbiddenValueConstant})
+\text{violation\_statement}&(S, E, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
+&\text{statement\_EDB}(S, E, \text{propertyConstant}, \text{forbiddenValueConstant})
 \end{split}\\
@@
 \begin{split}
-\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
+\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
+&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{} \\
 &\text{statement\_EDB}(O, X, \text{propertyConstant}, V) \wedge{} \\
@@
 \begin{split}
-\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
+\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
+&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{} \\
 &\text{unit}(V, U)  {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, U)
@@
 \begin{split}
-\text{violation\_qualifier}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+\text{violation\_qualifier}&(S, \text{propertyConstant}, V) \leftarrow \\
 &\text{qualifier\_EDB}(S, \text{propertyConstant}, V) \wedge{} \\
@@
 \begin{split}
-\text{violation\_reference}&(S, I, \text{propertyConstant}, V) \leftarrow \\
+\text{violation\_reference}&(S, H, \text{propertyConstant}, V) \leftarrow \\
 &\text{reference\_EDB}(S, H, \text{propertyConstant}, V) \wedge{} \\
@@
 \begin{split}
-\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
+\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
+&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{} \\
 &\text{qualifier\_EDB}(S, Q, O) {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, Q)
@@
 \begin{split}
-\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, V)
+\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
+&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, V)
 \end{split}\\
@@
 \begin{split}
-\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
+\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
+&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{} \\
 &\text{qualifier\_EDB}(S, \text{qualifierProperty}, O) {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, O)
@@
 \begin{split}
-c_1 = &\text{statement\_EDB}(Q, I, \text{propertyConstant}, X) \wedge{} \\
-&\text{statement\_EDB}(S, I, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
+c_1 = &\text{statement\_EDB}(Q, E, \text{propertyConstant}, X) \wedge{} \\
+&\text{statement\_EDB}(S, E, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
 \end{split} \\
 \begin{split}
-c_2 = &\text{statement\_EDB}(Q, I, \text{propertyConstant}, X) \wedge{} \\
-&\text{statement\_EDB}(S, I, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
-\end{split}
-\end{align}
-\begin{align}
-\begin{split}
-\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \text{last}(O, I) \wedge{} \\
+c_2 = &\text{statement\_EDB}(Q, E, \text{propertyConstant}, X) \wedge{} \\
+&\text{statement\_EDB}(S, E, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
+\end{split}
+\end{align}
+\begin{align}
+\begin{split}
+\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
+&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{} \text{last}(O, I) \wedge{} \\
 &\text{require}(O, \text{propertyConstant}, \text{requiredPropertyConstant})
@@
 c_1 = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
-&\text{statement\_EDB}(S, I, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
+&\text{statement\_EDB}(S, E, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
 \end{split}\\
@@
 c_2 = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
-&\text{statement\_EDB}(S, I, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
-\end{split}
-\end{align}
-
-\begin{align}
-\begin{split}
-\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \text{last}(O, V) \wedge{} \\
+&\text{statement\_EDB}(S, E, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
+\end{split}
+\end{align}
+
+\begin{align}
+\begin{split}
+\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
+&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{} \text{last}(O, V) \wedge{} \\
 &\text{require}(O, \text{propertyConstant}, \text{requiredPropertyConstant})
@@
 c_1 = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
-&\text{statement\_EDB}(S, I, P, V) \wedge{} \\
+&\text{statement\_EDB}(S, E, P, V) \wedge{} \\
 &\text{unequal}(\text{inversePropertyConstant}, P)
@@
 c_2 = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
-&\text{statement\_EDB}(S, I, \text{inversePropertyConstant}, V) \wedge{} \\
+&\text{statement\_EDB}(S, E, \text{inversePropertyConstant}, V) \wedge{} \\
 &\text{unequal}(R, V)
@@
 \begin{split}
-c = &\text{statement\_EDB}(Q, I, \text{propertyConstant}, X) \wedge{} \\
-&\text{statement\_EDB}(S, I, P, V) \wedge \text{unequal}(\text{propertyConstant}, P)
+c = &\text{statement\_EDB}(Q, E, \text{propertyConstant}, X) \wedge{} \\
+&\text{statement\_EDB}(S, E, P, V) \wedge \text{unequal}(\text{propertyConstant}, P)
 \end{split}
@@
 \text{require\_second}&(S, \text{propertyConstant}, \text{propertyConstant}) \leftarrow \\
-&\text{next}(R, S) \wedge \text{statement\_EDB}(R, I, \text{propertyConstant}, C) \wedge{} \\
+&\text{next}(R, S) \wedge \text{statement\_EDB}(R, E, \text{propertyConstant}, C) \wedge{} \\
 &\text{next}(Q, R) \wedge \text{require}(Q, \text{propertyConstant}, \text{propertyConstant})  \wedge c
@@
 \begin{split}\label{eq_no_separators_triple}
-\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
-&\text{statement\_EDB}(O, I, \text{propertyConstant}, X) \wedge{} \\
+\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
+&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{} \\
+&\text{statement\_EDB}(O, E, \text{propertyConstant}, X) \wedge{} \\
 &\text{unequal}(S, O)
@@
 \begin{split}
-c = &\text{statement\_EDB}(S, I, \text{propertyConstant}, C) \wedge{} \\
+c = &\text{statement\_EDB}(S, E, \text{propertyConstant}, C) \wedge{} \\
 &\text{qualifier\_EDB}(S, P, V) \wedge{} \\
@@
 \begin{split}
-\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
-&\text{statement\_EDB}(O, I, \text{propertyConstant}, X) \wedge{} \\
+\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
+&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{} \\
+&\text{statement\_EDB}(O, E, \text{propertyConstant}, X) \wedge{} \\
 &\text{unequal}(S, O) \wedge{} \\
@@
 \text{same\_or\_non\_existent}&(S, O, Q) \leftarrow \\
-&\text{statement\_EDB}(S, I, \text{propertyConstant}, X) \wedge{} \\
-&\text{statement\_EDB}(O, I, \text{propertyConstant}, Y) \wedge{} \\
+&\text{statement\_EDB}(S, E, \text{propertyConstant}, X) \wedge{} \\
+&\text{statement\_EDB}(O, E, \text{propertyConstant}, Y) \wedge{} \\
 &\text{qualifier\_EDB}(S, Q, V) \wedge{} \\
@@
 \text{same\_or\_non\_existent}&(S, O, Q) \leftarrow{} \\
-&\text{statement\_EDB}(S, I, \text{propertyConstant}, X) \wedge{} \\
-&\text{statement\_EDB}(O, I, \text{propertyConstant}, Y) \wedge{} \\
+&\text{statement\_EDB}(S, E, \text{propertyConstant}, X) \wedge{} \\
+&\text{statement\_EDB}(O, E, \text{propertyConstant}, Y) \wedge{} \\
 &\text{last\_qualifier}(S, P, V) \wedge{} \\
@@
 \begin{split}
-c = &\text{statement\_EDB}(S, I, \text{propertyConstant}, C) \wedge{} \\
+c = &\text{statement\_EDB}(S, E, \text{propertyConstant}, C) \wedge{} \\
 &\text{qualifier\_EDB}(S, P, V) \wedge \text{unequal}(P, \text{requiredPropertyConstant})
@@
 \begin{split}
-\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge \\
+\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
+&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge \\
 &\text{does\_not\_have}(S, \text{requiredPropertyConstant})
@@
 \begin{split}
-\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
+\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
+&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{} \\
 &\text{rank\_EDB}(S, \text{preferredRankConstant}) \wedge{} \\
-&\text{statement\_EDB}(O, I, \text{propertyConstant}, X) \wedge{} \\
+&\text{statement\_EDB}(O, E, \text{propertyConstant}, X) \wedge{} \\
 &\text{rank\_EDB}(O, \text{preferredRankConstant}) \wedge{} \\
@@
 \begin{split}
-\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
+\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
+&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{} \\
 &\text{rank\_EDB}(S, \text{preferredRankConstant}) \wedge{} \\
-&\text{statement\_EDB}(O, I, \text{propertyConstant}, W) \wedge{} \\
+&\text{statement\_EDB}(O, E, \text{propertyConstant}, W) \wedge{} \\
 &\text{rank\_EDB}(O, \text{preferredRankConstant}) \wedge{} \\

### 2018-09-20 thesis/paper/paper.tex
Renamed conjunction, removed existential

@@
 Let \(P_n\) denote  premise atoms and \(C_n\) the conclusion atoms.
-The variables used in one rule are divided into three mutually disjoint sets \(V_p\), \(V_c\), and \(V_r\) which are the variables occurring only in its premise atoms, only in its conclusion atoms, and occurring in both, respectively.
-The general from of an existential rule is (with \(k \in \mathbb{N}^+\), \(l \in \mathbb{N}\))
+The variables used in one rule are divided into two mutually disjoint sets \(V_p\) and \(V_r\) which are the variables occurring only in its premise atoms and occurring in conclusion atoms and premise atoms, respectively.
+The general from of a rule is (with \(k \in \mathbb{N}^+\), \(l \in \mathbb{N}\))
 \begin{equation}
-\forall V_p \forall V_r.(\exists V_c.C_1 \wedge \ldots \wedge C_k \leftarrow P_1 \wedge \ldots \wedge P_l)
+\forall V_p \forall V_r.(C_1 \wedge \ldots \wedge C_k \leftarrow P_1 \wedge \ldots \wedge P_l)
 \end{equation}
 
-\(V_c\) are called existential variables.
 If \(V_c = \emptyset\) the rule is called a \emph{datalog rule} and if \(l = 0\) it is called a \emph{fact}.
@@
 
-For statements the pattern consists of two rules, both utilising the same conjunction \(c\) whose atoms depend on the constraint type.
+For statements the pattern consists of two rules, both utilising the same conjunction \(\text{doesNotFulfil}\) whose atoms depend on the constraint type.
 This conjunction is used to test if the statement does not fulfil the requirement.
@@
 \begin{align}
-&\text{require}(S, \text{id}) \leftarrow \text{first}(S, I) \wedge c \\
-&\text{require}(S, \text{id}) \leftarrow \text{next}(O, S) \wedge \text{require}(O, \text{id}) \wedge c \\
-&\text{require}(S, P, V, \text{id}) \leftarrow \text{first}(S, P, V) \wedge c \\
-&\text{require}(S, P, V, \text{id}) \leftarrow \text{next}(O, X, C, S, P, V) \wedge \text{require}(O, X, C, \text{id}) \wedge c
+&\text{require}(S, \text{id}) \leftarrow \text{first}(S, I) \wedge \text{doesNotFulfil} \\
+&\text{require}(S, \text{id}) \leftarrow \text{next}(O, S) \wedge \text{require}(O, \text{id}) \wedge \text{doesNotFulfil} \\
+&\text{require}(S, P, V, \text{id}) \leftarrow \text{first}(S, P, V) \wedge \text{doesNotFulfil} \\
+&\text{require}(S, P, V, \text{id}) \leftarrow \text{next}(O, X, C, S, P, V) \wedge \text{require}(O, X, C, \text{id}) \wedge \text{doesNotFulfil}
 \end{align}
@@
 \begin{split}
-c_1 = &\text{statement\_EDB}(Q, E, \text{propertyConstant}, X) \wedge{} \\
+c = &\text{statement\_EDB}(Q, E, \text{propertyConstant}, X) \wedge{} \\
 &\text{statement\_EDB}(S, E, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
@@
 \begin{split}
-c_2 = &\text{statement\_EDB}(Q, E, \text{propertyConstant}, X) \wedge{} \\
+c = &\text{statement\_EDB}(Q, E, \text{propertyConstant}, X) \wedge{} \\
 &\text{statement\_EDB}(S, E, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
@@
 \begin{split}
-c_1 = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
+c = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
 &\text{statement\_EDB}(S, E, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
@@
 \begin{split}
-c_2 = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
+c = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
 &\text{statement\_EDB}(S, E, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
@@
 \begin{split}
-c_1 = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
+c = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
 &\text{statement\_EDB}(S, E, P, V) \wedge{} \\
@@
 \begin{split}
-c_2 = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
+c = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
 &\text{statement\_EDB}(S, E, \text{inversePropertyConstant}, V) \wedge{} \\
@@
 The require qualifier pattern needs to be applied to the qualifiers of constrained statements.
-The conjunction c is created for every \emph{separator} and the pattern applied.
+The conjunction \(\text{doesNotFulfil}\) is created for every \emph{separator} and the pattern applied.
 \begin{equation}

### 2018-09-20 thesis/paper/paper.tex
Fixed more multiple-atoms writing

@@
 c = &\text{statement\_EDB}(Q, E, \text{propertyConstant}, X) \wedge{} \\
-&\text{statement\_EDB}(S, E, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
+&\text{statement\_EDB}(S, E, \text{requiredPropertyConstant}, V) \wedge {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, V)
 \end{split}
@@
 c = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
-&\text{statement\_EDB}(S, E, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
+&\text{statement\_EDB}(S, E, \text{requiredPropertyConstant}, V) \wedge {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, V)
 \end{split}

### 2018-09-20 thesis/paper/paper.tex
Typos

@@
 At the beginning the content of the Wikidata JSON-Dump, read via WDKT, is transformed into facts readable by VLog4J.
-These take the form of comma-separated value files, which are created for each predicate in Section \ref{subsec_representation}, denoted as "Facts" in the diagram.
+These take the form of comma-separated value files, one for each predicate in Section \ref{subsec_representation}, denoted as "Facts" in the diagram.
 
 At the same time the Inequaliy Helper, receiving the Facts from WDTK, creates all files necessary to establish inequality as described in Section \ref{sec_constraints_using_inequality}, based on which approach was chosen.
+These files are constraint type specific, whereas the "Facts" are not.
 
@@
 All facts pertaining to the constraint, as well as the rules, are loaded into VLog, which creates an in-memory database from the provided .csv-files.
-Once this is done, the reasoning engine materializes all possible facts.
+For example, Once this is done, the reasoning engine materializes all possible facts.
 Afterwards, all violations are extracted by querying the database for the queries described in Section \ref{subsec_task}.
@@
 
-\subsection{Possible adaptations}
-Since the problematic performance of \emph{item requires statement} and \emph{single value} are not visible in \emph{value requires statement} and \emph{single best value}, which are very similar in their rule modelling, a possible adaptation would be to not evaluate all constrained properties for these constraints in bulk but split them into smaller sets, evaluating each set individually to combat the problematic scaling of the require pattern.
-The precomputed datasets could be kept, meaning only the reasoning step would have to be repeated.
-
-\subsection{Adaptations to VLog}
-
-\subsubsection{Sorting facts for column-based storage}
+\subsection{Adaptations to VLog - Sorting facts for column-based storage}
 VLog uses a column-based storage format where the first column is fully sorted and all following columns are sorted w.r.t.
@@
 
-\subsubsection{Sorting conjunctions for left-to-right evaluation}
-Conjunctions in VLog are evaluated from left to right, meaning the engine computes the matches of the first atom and, based on those findings, proceeds with the next atom.
-The rules as presented are not optimal regarding this process.
-Ideally, atoms should be sorted by the size of the intermediate results they incur.
-
 \subsubsection{Results}

### 2018-09-20 thesis/paper/paper.tex
Changed emphasis to sc

@@
 This is already incorporated into some constraints, but as of September 2018 they are not deployed.
-A value is one of several data types shown in Table \ref{tab_datatypes}, although it is also possible to use \emph{none} or \emph{some} as values for statements that have no value (akin to negation) or have a currently unknown value.
+A value is one of several data types shown in Table \ref{tab_datatypes}, although it is also possible to use \textsc{none} or \textsc{some} as values for statements that have no value (akin to negation) or have a currently unknown value.
 Every property determines the data type it accepts as an object.
@@
 This is done via a statement ID, which is assigned to every triple storing information about an item.
-Statements have one of three ranks, \emph{normal}, \emph{preferred}, and \emph{deprecated} which are used to differentiate similar statements on the same item.
-For example, the \emph{preferred} rank would be given to the current \emph{head of state} (P35) of a country, while all statements denoting former \emph{heads of state} have the rank \emph{normal}.
-\emph{Deprecated} is used on statements that where once considered correct but are not anymore, e.g., \emph{Chelsea Manning} (Q298423) having the \emph{sex or gender} (P21) \emph{male} (Q6581097).
+Statements have one of three ranks, \textsc{normal}, \textsc{preferred}, and \textsc{deprecated} which are used to differentiate similar statements on the same item.
+For example, the \textsc{preferred} rank would be given to the current \textsc{head of state} (P35) of a country, while all statements denoting former \textsc{heads of state} have the rank \textsc{normal}.
+\textsc{Deprecated} is used on statements that where once considered correct but are not anymore, e.g., \textsc{Chelsea Manning} (Q298423) having the \textsc{sex or gender} (P21) \textsc{male} (Q6581097).
 
@@
 Qualifiers provide contextual information for statements.
-The most common context is the time frame in which the statement was accurate, e.g., stating the \emph{start time} (P580) and \emph{end time} (P582) of the term of a head of state, but other information is possible, e.g.
+The most common context is the time frame in which the statement was accurate, e.g., stating the \textsc{start time} (P580) and \textsc{end time} (P582) of the term of a head of state, but other information is possible, e.g.
 determination method for measurements or the elevation above sea level for the highest point.
@@
 \begin{example}
-Figure \ref{fig_example} is an excerpt from the entity page of the clipper \emph{Cutty Sark} (Q171255) \footnote{\url{https://www.wikidata.org/wiki/Q171255}}.
+Figure \ref{fig_example} is an excerpt from the entity page of the clipper \textsc{Cutty Sark} (Q171255) \footnote{\url{https://www.wikidata.org/wiki/Q171255}}.
 An entity page lists all statements with the entity as the subject.
-It then groups all statements by their property, in this case \emph{length} (P2043).
+It then groups all statements by their property, in this case \textsc{length} (P2043).
 For the length of the Cutty Sark there is only one value, 85.4 metres, but this could include multiple values, e.g., if there were different ways to determine the length.
-This statement is further qualified with \emph{criterion used} (P1013) as the qualifier property, which specifies how the value was determined.
-In this case the qualifier value \emph{length overall} (Q2358152) was taken.
-The statement is backed up by one reference citing the property \emph{reference URL} (P854) pointing to the reference value, which is an URL confirming the length of 85.4 metres.
+This statement is further qualified with \textsc{criterion used} (P1013) as the qualifier property, which specifies how the value was determined.
+In this case the qualifier value \textsc{length overall} (Q2358152) was taken.
+The statement is backed up by one reference citing the property \textsc{reference URL} (P854) pointing to the reference value, which is an URL confirming the length of 85.4 metres.
 
@@
 It is fully declarative and uses rules to derive additional knowledge from given facts.
-The most basic domain element in datalog is a \emph{constant}, which will be denoted as \emph{someConstant}.
+The most basic domain element in datalog is a \textsc{constant}, which will be denoted as \textsc{someConstant}.
 In the context of Wikidata, any item or property, e.g.
-\emph{Earth} (Q2) or \emph{head of government} (P6), could be represented as a constant.
-Relations in datalog are expressed using \emph{predicates} with a specified arity \(n\), written as predicate\([n]\).
+\textsc{Earth} (Q2) or \textsc{head of government} (P6), could be represented as a constant.
+Relations in datalog are expressed using \textsc{predicates} with a specified arity \(n\), written as predicate\([n]\).
 To stay close to Wikidata in our examples, let us take the predicate triple\([3]\) as an example.
-This would be used to express the \emph{item}-\emph{property}-\emph{value} relation with each position representing one part of the triple.\\
-Additionally, datalog also provides \emph{variables}, which will be capitalized, like \(X\).
+This would be used to express the \textsc{item}-\textsc{property}-\textsc{value} relation with each position representing one part of the triple.\\
+Additionally, datalog also provides \textsc{variables}, which will be capitalized, like \(X\).
 The sets of predicates {\bf P}, constants {\bf C} and variables {\bf V} are mutually disjoint.
-Constants and variables are collectively known as \emph{terms} and are, together with predicates, used to form \emph{atoms}.
+Constants and variables are collectively known as \textsc{terms} and are, together with predicates, used to form \textsc{atoms}.
 An atom has the form predicate\((t_1, \ldots , t_n)\) with \(n\) as the arity of the predicate and \(t_1,\ldots,t_n\) terms.\\
-Atoms are the building blocks of \emph{rules}, which are used to derive information in datalog. A rule is a premise that implies a conclusion, and both are conjunctions of atoms.
+Atoms are the building blocks of \textsc{rules}, which are used to derive information in datalog. A rule is a premise that implies a conclusion, and both are conjunctions of atoms.
 Let \(P_n\) denote  premise atoms and \(C_n\) the conclusion atoms.
@@
 
-If \(V_c = \emptyset\) the rule is called a \emph{datalog rule} and if \(l = 0\) it is called a \emph{fact}.
-If a fact contains no variables but only constants, it is a \emph{ground rule}.
+If \(V_c = \emptyset\) the rule is called a \textsc{datalog rule} and if \(l = 0\) it is called a \textsc{fact}.
+If a fact contains no variables but only constants, it is a \textsc{ground rule}.
 Since all \(V_p\) and \(V_r\) are  quantified, the preceding universal quantifiers \(\forall V_p \forall V_r\) will be omitted for the rules presented in this work.
 
-Within the rule predicates we can differentiate between two different schemas: the \emph{extensional schema} and the \emph{intensional schema}, called EDB and IDB (extensional/intensional database).
-An EDB predicate may only occur in the premise of a rule, meaning it can never be derived.
-All predicates that appear in the conclusion at least once are IDB predicates.
+Within the rule predicates we can differentiate between two different schemas: the \textsc{extensional schema} and the \textsc{intensional schema}, called \textsc{edb} and \textsc{idb} (extensional/intensional database).
+An \textsc{edb} predicate may only occur in the premise of a rule, meaning it can never be derived.
+All predicates that appear in the conclusion at least once are \textsc{idb} predicates.
 
@@
 The individual entities and properties are introduced with their ID beforehand, so the example can be retraced.
-The species \emph{blue whale} (Q42196) has the \emph{parent taxon} (P171) \emph{balaenoptera} (Q133320), which in turn has the parent taxon \emph{rorqual} (Q232829).
-The predicates used are triple\_EDB\([3]\) as the input and isInTaxon\_IDB\([2]\), which will be derived.
+The species \textsc{blue whale} (Q42196) has the \textsc{parent taxon} (P171) \textsc{balaenoptera} (Q133320), which in turn has the parent taxon \textsc{rorqual} (Q232829).
+The predicates used are triple\_edb\([3]\) as the input and isInTaxon\_idb\([2]\), which will be derived.
 The basic facts are:
-\[\text{triple\_EDB}(\text{blue whale}, \text{parent taxon}, \text{balaenoptera})\]
-\[\text{triple\_EDB}(\text{balaenoptera}, \text{parent taxon}, \text{rorqual})\]
-From this we can draw conclusions. The first step is to represent the already existing facts as IDB:
-\[\text{isInTaxon\_IDB}(X, Y) \leftarrow \text{triple\_EDB}(X, \text{parent taxon}, Y)\]
+\[\textsc{triple\_edb}(\textsc{blue whale}, \textsc{parent taxon}, \textsc{balaenoptera})\]
+\[\textsc{triple\_edb}(\textsc{balaenoptera}, \textsc{parent taxon}, \textsc{rorqual})\]
+From this we can draw conclusions. The first step is to represent the already existing facts as idb:
+\[\textsc{isInTaxon\_idb}(X, Y) \leftarrow \textsc{triple\_edb}(X, \textsc{parent taxon}, Y)\]
 Finally the transitive closure of this relation can be derived with
-\[\text{isInTaxon\_IDB}(X, Y) \leftarrow \text{isInTaxon\_IDB}(X, Z) \wedge \text{isInTaxon\_IDB}(Z, Y)\]
+\[\textsc{isInTaxon\_idb}(X, Y) \leftarrow \textsc{isInTaxon\_idb}(X, Z) \wedge \textsc{isInTaxon\_idb}(Z, Y)\]
 \end{example}
@@
 Note that a constrained statement means a statement whose predicate is a property that has a specific constraint.
-A simple example for this would be \emph{distinct values} (Q21502410), stating that values of the property must be unique across all constrained statements.
+A simple example for this would be \textsc{distinct values} (Q21502410), stating that values of the property must be unique across all constrained statements.
 For this case no additional information regarding the constraint is necessary and violation could be found by surveying all constrained statements and comparing their values.\\
@@
 These qualifiers are either optional or mandatory, and may be applied once or multiple times.
-For example, the \emph{none of} constraint (Q52558054) states that values for the specific property may not be in a given set of disallowed values.
-This set is specified using qualifiers with property \emph{item of} (P2305), which list all the disallowed values.
+For example, the \textsc{none of} constraint (Q52558054) states that values for the specific property may not be in a given set of disallowed values.
+This set is specified using qualifiers with property \textsc{item of} (P2305), which list all the disallowed values.
 At least one such qualifier must be present for the statement.\\
 More complex dependencies are also possible.
-A constraint can depend on all constrained statements of a specific item (\emph{single value} constraint [Q19474404]), on all statements of the item specified as the value of a constrained statement (\emph{value requires statement} constraint [Q21510864]), or on all statements using the subclass of-predicate (\emph{type} constraint [Q21503250]).
+A constraint can depend on all constrained statements of a specific item (\textsc{single value} constraint [Q19474404]), on all statements of the item specified as the value of a constrained statement (\textsc{value requires statement} constraint [Q21510864]), or on all statements using the subclass of-predicate (\textsc{type} constraint [Q21503250]).
 
@@
 Table \ref{tab_property_constraints} shows that most constraints cannot be modelled without at least some amount of preprocessing.
-Only \emph{scope}, \emph{conflicts with}, \emph{allowed entity types} and \emph{none of} can be modelled directly.
-This is due to the fact that \emph{conflicts with} and \emph{none of} exclude a limited and specified set of statements from validity and these cases can thus be directly taken as violations.
-\emph{scope} and \emph{allowed entity types} work very similar, since there are only three respectively two possible values and violations are simply the options that have not been allowed.
+Only \textsc{scope}, \textsc{conflicts with}, \textsc{allowed entity types} and \textsc{none of} can be modelled directly.
+This is due to the fact that \textsc{conflicts with} and \textsc{none of} exclude a limited and specified set of statements from validity and these cases can thus be directly taken as violations.
+\textsc{scope} and \textsc{allowed entity types} work very similar, since there are only three respectively two possible values and violations are simply the options that have not been allowed.
 The rules for these constraints are described in Section \ref{sec_simple_constraints}.\\
@@
 A problem that occurs multiple times is the need to determine inequality between constants.
-\emph{distinct value}, \emph{allowed units}, \emph{allowed qualifiers}, \emph{one of}, \emph{single best value} and \emph{one-of qualifier value} each require a rule to determine if a value, unit, or property is not equal to another.
+\textsc{distinct value}, \textsc{allowed units}, \textsc{allowed qualifiers}, \textsc{one of}, \textsc{single best value} and \textsc{one-of qualifier value} each require a rule to determine if a value, unit, or property is not equal to another.
 To this end a certain amount of preprocessing is necessary, most notably the extraction of all constant pairs whose inequality has to be determined for finding the violations.
-To avoid putting and undue amount of processing outside of datalog, this extraction will not check more than that the constant is in the right position relative to the constrained statement, e.g., all values on statements constrained by \emph{distinct value} or all qualifiers on statements constrained by \emph{allowed qualifiers}.
+To avoid putting and undue amount of processing outside of datalog, this extraction will not check more than that the constant is in the right position relative to the constrained statement, e.g., all values on statements constrained by \textsc{distinct value} or all qualifiers on statements constrained by \textsc{allowed qualifiers}.
 From this set inequality will be generated, different versions of which will be introduced in Section \ref{sec_inequality} and evaluated with regards to their performance in Section \ref{sec_constraints_using_inequality}.\\
@@
 However, not all constraints require a general negation.
-\emph{item requires statement}, \emph{value requires statement}, \emph{inverse }and \emph{symmetric} are all violated if a statement does not exist on a specific item.
+\textsc{item requires statement}, \textsc{value requires statement}, \textsc{inverse }and \textsc{symmetric} are all violated if a statement does not exist on a specific item.
 This problem can be solved with limited amounts of preprocessing by leveraging a preprocessed statement order per item and 'iterating' through the statements until the required conditions are met.\\
-A similar approach can be taken to find items with exactly one statement matching certain criteria, required for 	\emph{multi-value}.
+A similar approach can be taken to find items with exactly one statement matching certain criteria, required for 	\textsc{multi-value}.
 The idea is to ask the question if a second statement exists, then evaluating it as above.
@@
 \begin{table}[H]
-\caption{EDB predicates}\label{tab_predicates}
+\caption{\textsc{edb} predicates}\label{tab_predicates}
 \begin{tabular}{llp{8.5cm}}
@@
 \hline
-statement\_EDB & \makecell{statementID \\ entityID \\ propertyID \\ value (entityID or string)} & represents basic statements \\
-\hline
-qualifier\_EDB & \makecell{statementID \\ propertyID \\ value} & represents qualifiers on statements \\
-\hline
-reference\_EDB & \makecell{statementID \\ reference hash \\ propertyID \\ value} & represents references on statements; reference hash groups the property-value pairs belonging to one reference \\
-\hline
-item\_EDB & entityID & marks an entity as an item \\
-property\_EDB & entityID & marks an entity as a property \\
-\hline
-unit\_EDB & \makecell{value \\ unitID} & denotes which unit a value has \\
-\hline
-rank\_EDB & \makecell{statementID \\ rank} & denotes the rank a statement has
+\textsc{statement\_edb} & \makecell{statementID \\ entityID \\ propertyID \\ value (entityID or string)} & represents basic statements \\
+\hline
+\textsc{qualifier\_edb} & \makecell{statementID \\ propertyID \\ value} & represents qualifiers on statements \\
+\hline
+\textsc{reference\_edb} & \makecell{statementID \\ reference hash \\ propertyID \\ value} & represents references on statements; reference hash groups the property-value pairs belonging to one reference \\
+\hline
+\textsc{item\_edb} & entityID & marks an entity as an item \\
+\textsc{property\_edb} & entityID & marks an entity as a property \\
+\hline
+\textsc{unit\_edb} & \makecell{value \\ unitID} & denotes which unit a value has \\
+\hline
+\textsc{rank\_edb} & \makecell{statementID \\ rank} & denotes the rank a statement has
 \end{tabular}
@@
 \begin{align*}
-&\text{violation\_statement}(S, E, \text{propertyConstant}, V) &\leftarrow \text{atoms determining the violation} \\
-&\text{violation\_qualifier}(S, \text{propertyConstant}, V) &\leftarrow \text{atoms determining the violation} \\
-&\text{violation\_reference}(S, H, \text{propertyConstant}, V) &\leftarrow \text{atoms determining the violation}
+&\textsc{violation\_statement}(S, E, \textsc{propertyConstant}, V) &\leftarrow \textsc{atoms determining the violation} \\
+&\textsc{violation\_qualifier}(S, \textsc{propertyConstant}, V) &\leftarrow \textsc{atoms determining the violation} \\
+&\textsc{violation\_reference}(S, H, \textsc{propertyConstant}, V) &\leftarrow \textsc{atoms determining the violation}
 \end{align*}
@@
 \begin{align*}
-&\text{violation\_statement}(S, E, P, V) \\
-&\text{violation\_qualifier}(S, P, V) \\
-&\text{violation\_reference}(S, H, P, V)
+&\textsc{violation\_statement}(S, E, P, V) \\
+&\textsc{violation\_qualifier}(S, P, V) \\
+&\textsc{violation\_reference}(S, H, P, V)
 \end{align*}
@@
 The violations list the conflicting (part of a) statement in relation to the constrained statement.
-The first constraint, \emph{property scope}, will be used as an example to explain the table in detail.
+The first constraint, \textsc{property scope}, will be used as an example to explain the table in detail.
 
@@
 \paragraph{Constraint}
-The \emph{property scope} constraint specifies if a property may be used as predicate in a statement, qualifier, or reference.
+The \textsc{property scope} constraint specifies if a property may be used as predicate in a statement, qualifier, or reference.
 \begin{table}[H]
@@
 \hline
-Qualifiers & \makecell{\emph{property scope} (P5314) -- 1..3 \\ must be \emph{as main value} (Q54828448), \emph{as qualifier} (Q54828449), \\ or \emph{as reference} (Q54828450) \\ specifies allowed use in statements, qualifiers, or references} \\
+Qualifiers & \makecell{\textsc{property scope} (P5314) -- 1..3 \\ must be \textsc{as main value} (Q54828448), \textsc{as qualifier} (Q54828449), \\ or \textsc{as reference} (Q54828450) \\ specifies allowed use in statements, qualifiers, or references} \\
 \hline
@@
 In this case, this is simply each statement, qualifier, or reference whose predicate is the constrained property.\\
-The \emph{property scope} constraint requires the scope, meaning statement, qualifier, or reference, to be set via qualifier \emph{property scope} (P5314).
+The \textsc{property scope} constraint requires the scope, meaning statement, qualifier, or reference, to be set via qualifier \textsc{property scope} (P5314).
 The cardinality is 1..3 because it is mandatory and has only three possible values.\\
-It is violated if a statement, qualifier, or reference has a predicate with the \emph{property scope} constraint that does not allow this scope.\\
-
-\begin{example}
-The \emph{reference URL} (P854) property is restricted to the \emph{property scope} (P5314) \emph{as reference} (Q54828450), because it is to be used to mark a URL that substantiates a statement.
-\end{example}
-
-\paragraph{Rules}
-The \emph{property scope} constraint requires no context information to be violated; the triple, qualifier, or reference with the constrained property is enough.
+It is violated if a statement, qualifier, or reference has a predicate with the \textsc{property scope} constraint that does not allow this scope.\\
+
+\begin{example}
+The \textsc{reference URL} (P854) property is restricted to the \textsc{property scope} (P5314) \textsc{as reference} (Q54828450), because it is to be used to mark a URL that substantiates a statement.
+\end{example}
+
+\paragraph{Rules}
+The \textsc{property scope} constraint requires no context information to be violated; the triple, qualifier, or reference with the constrained property is enough.
 This means violations can be found using these rules:
@@
 \begin{split}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, V)\\
-\end{split}\\
-\begin{split}
-\text{violation\_qualifier}&(S, \text{propertyConstant}, V) \leftarrow \\
-&\text{qualifier\_EDB}(S, \text{propertyConstant}, V)\\
-\end{split}\\
-\begin{split}
-\text{violation\_reference}&(S, \text{propertyConstant}, V) \leftarrow \\
-&\text{reference\_EDB}(S, H, \text{propertyConstant}, V)\\
-\end{split}
-\end{align}
-Which of these rules need to be used depends on the values of the \emph{property scope} qualifier.
-If a concrete constrained on a property does not have \emph{as main value} as a value of the qualifier, the first rule is used.
-The case for \emph{as qualifier} and \emph{as reference} is analogous with the second and third rule, respectively.
+\textsc{violation\_statement}&(S, E, \textsc{propertyConstant}, V) \leftarrow \\
+&\textsc{statement\_edb}(S, E, \textsc{propertyConstant}, V)\\
+\end{split}\\
+\begin{split}
+\textsc{violation\_qualifier}&(S, \textsc{propertyConstant}, V) \leftarrow \\
+&\textsc{qualifier\_edb}(S, \textsc{propertyConstant}, V)\\
+\end{split}\\
+\begin{split}
+\textsc{violation\_reference}&(S, \textsc{propertyConstant}, V) \leftarrow \\
+&\textsc{reference\_edb}(S, H, \textsc{propertyConstant}, V)\\
+\end{split}
+\end{align}
+Which of these rules need to be used depends on the values of the \textsc{property scope} qualifier.
+If a concrete constrained on a property does not have \textsc{as main value} as a value of the qualifier, the first rule is used.
+The case for \textsc{as qualifier} and \textsc{as reference} is analogous with the second and third rule, respectively.
 
@@
 \paragraph{Constraint}
-The \emph{conflicts with} constraint forbids statements on constrained entities whose predicate is a specified \emph{conflicting property}.
-If \emph{conflicting values} have been specified as well, the constraint is weakened and only violated by statements if in addition to the predicate being a \emph{conflicting property}, the object is one of the \emph{conflicting values}.
+The \textsc{conflicts with} constraint forbids statements on constrained entities whose predicate is a specified \textsc{conflicting property}.
+If \textsc{conflicting values} have been specified as well, the constraint is weakened and only violated by statements if in addition to the predicate being a \textsc{conflicting property}, the object is one of the \textsc{conflicting values}.
 \begin{table}[H]
@@
 \hline
-Qualifiers & \makecell{\emph{conflicting property} (P2306) -- 1 \\ property that may not be predicate of entity statements \\
-\emph{conflicting values} (P2305) -- 0..* \\ values that may not be objects of entity statements with the \emph{conflicting property}} \\
+Qualifiers & \makecell{\textsc{conflicting property} (P2306) -- 1 \\ property that may not be predicate of entity statements \\
+\textsc{conflicting values} (P2305) -- 0..* \\ values that may not be objects of entity statements with the \textsc{conflicting property}} \\
 \hline
@@
 \begin{example}
-The \emph{author} (P50) constraint has (among others) the \emph{conflicting property} (P2306) \emph{instance of} (P31) with the \emph{conflicting value} (P2305) \emph{Wikimedia disambiguation page} (Q4167410).
-It is set because editors might easily confuse a work of art with a specific title, e.g., \emph{Tempest} (Q86440) by \emph{author} \emph{William Shakespear} (Q692) with the Wikimedia disambiguation page \emph{Tempest} (Q1589529).
+The \textsc{author} (P50) constraint has (among others) the \textsc{conflicting property} (P2306) \textsc{instance of} (P31) with the \textsc{conflicting value} (P2305) \textsc{Wikimedia disambiguation page} (Q4167410).
+It is set because editors might easily confuse a work of art with a specific title, e.g., \textsc{Tempest} (Q86440) by \textsc{author} \textsc{William Shakespear} (Q692) with the Wikimedia disambiguation page \textsc{Tempest} (Q1589529).
 \end{example}
@@
 \begin{split}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow\\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{}\\
-&\text{statement\_EDB}(O, E, \text{conflictingPropertyConstant}, C)
-\end{split}\\
-\begin{split}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow\\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{}\\
-&\text{statement\_EDB}(O, E, \text{conflictingPropertyConstant}, \text{conflictingValueConstant})
+\textsc{violation\_statement}&(S, E, \textsc{propertyConstant}, V) \leftarrow\\
+&\textsc{statement\_edb}(S, E, \textsc{propertyConstant}, V) \wedge{}\\
+&\textsc{statement\_edb}(O, E, \textsc{conflictingPropertyConstant}, C)
+\end{split}\\
+\begin{split}
+\textsc{violation\_statement}&(S, E, \textsc{propertyConstant}, V) \leftarrow\\
+&\textsc{statement\_edb}(S, E, \textsc{propertyConstant}, V) \wedge{}\\
+&\textsc{statement\_edb}(O, E, \textsc{conflictingPropertyConstant}, \textsc{conflictingValueConstant})
 \end{split}
@@
 \paragraph{Constraint}
-The \emph{allowed entity types} constraint limits if a property may be used in a statement whose subject is an item, a property, or a lexeme, although as of September 2018 lexemes are not deployed.
+The \textsc{allowed entity types} constraint limits if a property may be used in a statement whose subject is an item, a property, or a lexeme, although as of September 2018 lexemes are not deployed.
 
@@
 \hline
-Qualifiers & \makecell{allowed entity type (P2305) -- 1..3 \\ must be \emph{Wikibase item} (Q29934200), \emph{Wikibase property} (Q29934218), \\ or lexeme (Q51885771); \\ specifies the entity type the property may be used on} \\
+Qualifiers & \makecell{allowed entity type (P2305) -- 1..3 \\ must be \textsc{Wikibase item} (Q29934200), \textsc{Wikibase property} (Q29934218), \\ or lexeme (Q51885771); \\ specifies the entity type the property may be used on} \\
 \hline
@@
 
-Since there are only three options (and only two in use), violations can be found without inequality by finding the statement\_EDB with the constrained property whose subject is of the type that is not allowed, meaning either \emph{Wikibase item} or \emph{Wikibase property}.
-
-\begin{example}
-The \emph{property constraint} (P2302) property has the \emph{allowed entity type} (P2305) \emph{Wikibase property} (Q29934218) because non-properties cannot be subject to a property constraint.
+Since there are only three options (and only two in use), violations can be found without inequality by finding the \textsc{statement\_edb} with the constrained property whose subject is of the type that is not allowed, meaning either \textsc{Wikibase item} or \textsc{Wikibase property}.
+
+\begin{example}
+The \textsc{property constraint} (P2302) property has the \textsc{allowed entity type} (P2305) \textsc{Wikibase property} (Q29934218) because non-properties cannot be subject to a property constraint.
 \end{example}
@@
 \begin{split}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge \text{item}(I)
-\end{split}\\
-\begin{split}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge \text{property}(I)
-\end{split}
-\end{align}
-The first rule is created for all \emph{allowed entity type} constraints whose \emph{allowed entity type} does not contain \emph{Wikibase item}.
-The second rule works analogous for \emph{Wikibase property}.
+\textsc{violation\_statement}&(S, E, \textsc{propertyConstant}, V) \leftarrow \\
+&\textsc{statement\_edb}(S, E, \textsc{propertyConstant}, V) \wedge \textsc{item}(I)
+\end{split}\\
+\begin{split}
+\textsc{violation\_statement}&(S, E, \textsc{propertyConstant}, V) \leftarrow \\
+&\textsc{statement\_edb}(S, E, \textsc{propertyConstant}, V) \wedge \textsc{property}(I)
+\end{split}
+\end{align}
+The first rule is created for all \textsc{allowed entity type} constraints whose \textsc{allowed entity type} does not contain \textsc{Wikibase item}.
+The second rule works analogous for \textsc{Wikibase property}.
 
@@
 \hline
-Qualifiers & \makecell{\emph{forbidden value} (P2305) -- 1..* \\ values that may not be object of a constraint statement, qualifier, or reference} \\
+Qualifiers & \makecell{\textsc{forbidden value} (P2305) -- 1..* \\ values that may not be object of a constraint statement, qualifier, or reference} \\
 \hline
@@
 \begin{example}
-The \emph{license} (P31) property, denoting the license under which a work is released, has the \emph{forbidden value} (P2305) \emph{proprietary software} (Q218616), since it is often confused with \emph{proprietary license} (Q3238057).
+The \textsc{license} (P31) property, denoting the license under which a work is released, has the \textsc{forbidden value} (P2305) \textsc{proprietary software} (Q218616), since it is often confused with \textsc{proprietary license} (Q3238057).
 \end{example}
@@
 \begin{split}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, \text{forbiddenValueConstant})
-\end{split}\\
-\begin{split}
-\text{violation\_qualifier}&(S, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
-&\text{qualifier\_EDB}(S, \text{propertyConstant}, \text{forbiddenValueConstant})
-\end{split}\\
-\begin{split}
-\text{violation\_reference}&(S, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
-&\text{reference\_EDB}(S, H, \text{propertyConstant}, \text{forbiddenValueConstant})
-\end{split}
-\end{align}
-Given a specified constraint, each of the three rules is created once for each \emph{forbidden value}.
+\textsc{violation\_statement}&(S, E, \textsc{propertyConstant}, \textsc{forbiddenValueConstant}) \leftarrow \\
+&\textsc{statement\_edb}(S, E, \textsc{propertyConstant}, \textsc{forbiddenValueConstant})
+\end{split}\\
+\begin{split}
+\textsc{violation\_qualifier}&(S, \textsc{propertyConstant}, \textsc{forbiddenValueConstant}) \leftarrow \\
+&\textsc{qualifier\_edb}(S, \textsc{propertyConstant}, \textsc{forbiddenValueConstant})
+\end{split}\\
+\begin{split}
+\textsc{violation\_reference}&(S, \textsc{propertyConstant}, \textsc{forbiddenValueConstant}) \leftarrow \\
+&\textsc{reference\_edb}(S, H, \textsc{propertyConstant}, \textsc{forbiddenValueConstant})
+\end{split}
+\end{align}
+Given a specified constraint, each of the three rules is created once for each \textsc{forbidden value}.
 
@@
 In every approach it will be necessary to load some basic inequalities into the database.
-To this end the predicate unequal\_EDB\([2]\) is introduced.
-Since the goal is to keep the number of applications of this predicate small, an IDB predicate unequal\([2]\) has to be introduced.
+To this end the predicate unequal\_edb\([2]\) is introduced.
+Since the goal is to keep the number of applications of this predicate small, an \textsc{idb} predicate unequal\([2]\) has to be introduced.
 This leads to the rule
 \begin{align}
-\text{unequal}(X, Y) \leftarrow \text{unequal\_EDB}(X, Y)
-\end{align}
-
-Given a set of unequal constants \(U\), the task is to generate rules so that for all \( x, y \in U: x \ne y \Rightarrow \text{unequal}(x, y)\).
+\textsc{unequal}(X, Y) \leftarrow \textsc{unequal\_edb}(X, Y)
+\end{align}
+
+Given a set of unequal constants \(U\), the task is to generate rules so that for all \( x, y \in U: x \ne y \Rightarrow \textsc{unequal}(x, y)\).
 
@@
 \begin{align}
-\text{for all } x, y \in U, x \ne y \text{ add } \text{unequal\_EDB}(x, y)
+\textsc{for all } x, y \in U, x \ne y \textsc{ add } \textsc{unequal\_edb}(x, y)
 \end{align}
@@
 \begin{align}\label{equ_inverse}
-\text{unequal}(X, Y) \leftarrow \text{unequal}(Y, X)
+\textsc{unequal}(X, Y) \leftarrow \textsc{unequal}(Y, X)
 \end{align}
@@
 \begin{align}
-\text{for all } x, y \in U, x < y \text{ add } \text{unequal\_EDB}(x, y)
+\textsc{for all } x, y \in U, x < y \textsc{ add } \textsc{unequal\_edb}(x, y)
 \end{align}
@@
 Then \(l_{\max} = \max\limits_{\forall u \in U}(len(u))\) is the length of the longest entity string that occurs in the given data.
-Based on this a group of predicates letter\(_i\)\_EDB[8] is introduced where \(0 \equiv i \mod 7\), \(0 \leq i < l_{max} \).
-
-For each unequal constant string \(u = c_0 \ldots c_{l - 1}\) and \(0 \equiv i \mod 7, 0 \leq i < len(u)\) an atom using predicate letter\(_i\)\_EDB is created as 
-\begin{align}
-\begin{split}
-\text{letter}_i\text{\_EDB}(e, f(e, i), f(e, i+1), \ldots, f(e, i+6)
+Based on this a group of predicates letter\(_i\)\_edb[8] is introduced where \(0 \equiv i \mod 7\), \(0 \leq i < l_{max} \).
+
+For each unequal constant string \(u = c_0 \ldots c_{l - 1}\) and \(0 \equiv i \mod 7, 0 \leq i < len(u)\) an atom using predicate \textsc{letter\(_i\)\_edb} is created as 
+\begin{align}
+\begin{split}
+\textsc{letter}_i\textsc{\_edb}(e, f(e, i), f(e, i+1), \ldots, f(e, i+6)
 \end{split}
@@
       c_i & i < len(e) \\
-      \text{none} & i \geq len(e)
+      \textsc{none} & i \geq len(e)
 \end{cases}
@@
 
-Now that inequality between all characters and the link between entities and their characters are introduced, the inequality between two entities can be derived by creating the following rules for all predicates letter\(_i\)\_EDB (with \text{letters} short for \(\text{letter}_i\text{\_EDB}(X, A_i, \ldots, A_{i+6}) \wedge\text{letter}_i\text{\_EDB}(Y, B_i, \ldots, B_{i+6})\), \(i \leq j < i + 7\)):
-\begin{align}
-\text{unequal}(X, Y) \leftarrow \text{letters} \wedge \text{unequal}(A_j, B_j)
-\end{align}
-Note that the facts fulfilling \(\text{unequal}(A_j, B_j)\) are created by using the approach outlined in Section \ref{subsubsec_naive} for the set of characters \(C\) in all unequal constants \(U\).
+Now that inequality between all characters and the link between entities and their characters are introduced, the inequality between two entities can be derived by creating the following rules for all predicates \textsc{letter\(_i\)\_edb} (with \textsc{letters} short for \(\textsc{letter}_i\textsc{\_edb}(X, A_i, \ldots, A_{i+6}) \wedge\textsc{letter}_i\textsc{\_edb}(Y, B_i, \ldots, B_{i+6})\), \(i \leq j < i + 7\)):
+\begin{align}
+\textsc{unequal}(X, Y) \leftarrow \textsc{letters} \wedge \textsc{unequal}(A_j, B_j)
+\end{align}
+Note that the facts fulfilling \(\textsc{unequal}(A_j, B_j)\) are created by using the approach outlined in Section \ref{subsubsec_naive} for the set of characters \(C\) in all unequal constants \(U\).
 
@@
 B_m\) are arbitrary atoms, \(I_1 ..
-I_n\) are atoms of the form \(\text{unequal}(t_{n_1}, t_{n_2})\) with \(t_{n_1}, t_{n_2}\) terms and all variables amount \(t_{n_1}\) and \(t_{n_2}\) appear in at least one of \(B_1 \ldots B_m\), rule one is introduced for all \(n\).
+I_n\) are atoms of the form \(\textsc{unequal}(t_{n_1}, t_{n_2})\) with \(t_{n_1}, t_{n_2}\) terms and all variables amount \(t_{n_1}\) and \(t_{n_2}\) appear in at least one of \(B_1 \ldots B_m\), rule one is introduced for all \(n\).
 The rules outlined in Section \ref{subsubsec_encoding} are changed to utilise this new requirement as shown in rule two.
@@
 \begin{split}
-\text{req\_inequality}(t_{n_1}, t_{n_2})
+\textsc{req\_inequality}(t_{n_1}, t_{n_2})
  \leftarrow B_1 \wedge \ldots \wedge B_m
@@
 \begin{split}
-\text{unequal}(X, Y) \leftarrow \text{req\_inequality}	(X, Y) \wedge \text{letters} \wedge \text{unequal}(A_j, B_j)
+\textsc{unequal}(X, Y) \leftarrow \textsc{req\_inequality}	(X, Y) \wedge \textsc{letters} \wedge \textsc{unequal}(A_j, B_j)
 \end{split}
@@
 \paragraph{Constraint}
-The \emph{distinct values} constraint forbids the existence of two constrained statements with the same object.
+The \textsc{distinct values} constraint forbids the existence of two constrained statements with the same object.
 \begin{table}[H]
@@
 \end{table}
-The constraint can be expanded to cover qualifiers and references as well by setting a \emph{constraint scope} (P4680) qualifier with the respective value.
+The constraint can be expanded to cover qualifiers and references as well by setting a \textsc{constraint scope} (P4680) qualifier with the respective value.
 However, as of September 2018 there is no constraint making use of this.
@@
 \begin{example}
-The \emph{ISBN-13} (P212) property has a \emph{distinct values} constraint since an ISBN is supposed to be a unique identifier for a book and should not be shared between two items.
-\end{example}
-
-\paragraph{Rules}
-The \emph{distinct values} constraint requires two distinct constrained statements with the same value. 
-Two constrained statements with the same value can be found via a simple conjunction of two statement\_EDB atoms with the propertyConstant as predicate and a variable \(V\) as value.
+The \textsc{ISBN-13} (P212) property has a \textsc{distinct values} constraint since an ISBN is supposed to be a unique identifier for a book and should not be shared between two items.
+\end{example}
+
+\paragraph{Rules}
+The \textsc{distinct values} constraint requires two distinct constrained statements with the same value. 
+Two constrained statements with the same value can be found via a simple conjunction of two \textsc{statement\_edb} atoms with the propertyConstant as predicate and a variable \(V\) as value.
 Since this would mark every constrained statement as a violation because it can fulfil both atoms at the same time, inequality of the statement IDs \(S\) and \(O\) must be tested as well.
@@
 \begin{split}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{} \\
-&\text{statement\_EDB}(O, X, \text{propertyConstant}, V) \wedge{} \\
-&\text{unequal}(S, O)
+\textsc{violation\_statement}&(S, E, \textsc{propertyConstant}, V) \leftarrow \\
+&\textsc{statement\_edb}(S, E, \textsc{propertyConstant}, V) \wedge{} \\
+&\textsc{statement\_edb}(O, X, \textsc{propertyConstant}, V) \wedge{} \\
+&\textsc{unequal}(S, O)
 \end{split}
@@
 \paragraph{Constraint}
-The \emph{allowed units} constraint specifies that the value of a constrained statement may only have a unit that is in a set of specific units.
+The \textsc{allowed units} constraint specifies that the value of a constrained statement may only have a unit that is in a set of specific units.
 
@@
 \hline
-Qualifiers & \makecell{\emph{allowed unit} (P2305) -- 1..* \\ units that are allowed on values of constrained statements (\emph{none} for dimensionless units)} \\
+Qualifiers & \makecell{\textsc{allowed unit} (P2305) -- 1..* \\ units that are allowed on values of constrained statements (\textsc{none} for dimensionless units)} \\
 \hline
@@
 
-By setting the \emph{constraint scope} (P4680) qualifier on the constraint, it can be limited to only affect a subset of statement, qualifier, and reference.
+By setting the \textsc{constraint scope} (P4680) qualifier on the constraint, it can be limited to only affect a subset of statement, qualifier, and reference.
 As of September 2018 this is not done on any constraint.
@@
 \begin{example}
-The \emph{length} (P2043) property has (among others) the \emph{allowed unit} (P2305) \emph{metre} (Q11573).
-This helps avoid confusing it for \emph{duration} (P2047), which denotes the length of time of an event.
+The \textsc{length} (P2043) property has (among others) the \textsc{allowed unit} (P2305) \textsc{metre} (Q11573).
+This helps avoid confusing it for \textsc{duration} (P2047), which denotes the length of time of an event.
 \end{example}
@@
 To handle this constraint, inequality between all units need to be established.
-Then a violation can be identified by finding all statement\_EDB, qualifier\_EDB, and reference\_EDB with the constrained property and a unit unequal to all allowed units.
+Then a violation can be identified by finding all \textsc{statement\_edb}, \textsc{qualifier\_edb}, and \textsc{reference\_edb} with the constrained property and a unit unequal to all allowed units.
 The allowed units constraint needs to determine inequality between all allowed units and each unit on a constrained statement.
@@
 \begin{split}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{} \\
-&\text{unit}(V, U)  {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, U)
-\end{split}\\
-\begin{split}
-\text{violation\_qualifier}&(S, \text{propertyConstant}, V) \leftarrow \\
-&\text{qualifier\_EDB}(S, \text{propertyConstant}, V) \wedge{} \\
-&\text{unit}(V, U) {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, U)
-\end{split}\\
-\begin{split}
-\text{violation\_reference}&(S, H, \text{propertyConstant}, V) \leftarrow \\
-&\text{reference\_EDB}(S, H, \text{propertyConstant}, V) \wedge{} \\
-&\text{unit}(V, U) {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, U)
+\textsc{violation\_statement}&(S, E, \textsc{propertyConstant}, V) \leftarrow \\
+&\textsc{statement\_edb}(S, E, \textsc{propertyConstant}, V) \wedge{} \\
+&\textsc{unit}(V, U)  {\textstyle\bigwedge}_{a \in A} \textsc{unequal}(a, U)
+\end{split}\\
+\begin{split}
+\textsc{violation\_qualifier}&(S, \textsc{propertyConstant}, V) \leftarrow \\
+&\textsc{qualifier\_edb}(S, \textsc{propertyConstant}, V) \wedge{} \\
+&\textsc{unit}(V, U) {\textstyle\bigwedge}_{a \in A} \textsc{unequal}(a, U)
+\end{split}\\
+\begin{split}
+\textsc{violation\_reference}&(S, H, \textsc{propertyConstant}, V) \leftarrow \\
+&\textsc{reference\_edb}(S, H, \textsc{propertyConstant}, V) \wedge{} \\
+&\textsc{unit}(V, U) {\textstyle\bigwedge}_{a \in A} \textsc{unequal}(a, U)
 \end{split}
@@
 \paragraph{Constraint}
-The \emph{allowed qualifiers} constraint specifies that only a limited set of properties may be used as predicates of qualifiers on a constrained statement.
+The \textsc{allowed qualifiers} constraint specifies that only a limited set of properties may be used as predicates of qualifiers on a constrained statement.
 
@@
 \hline
-Qualifiers & \makecell{\emph{allowed qualifier} (P2306) -- 1..* \\ properties that may appear in predicates of qualifiers on constrained statements} \\
+Qualifiers & \makecell{\textsc{allowed qualifier} (P2306) -- 1..* \\ properties that may appear in predicates of qualifiers on constrained statements} \\
 \hline
@@
 \begin{example}
-The \emph{head of state} (P35) property has (among others) the \emph{allowed qualifiers} (P2306) \emph{start time} (P580) and \emph{end time} (P582) properties, used to describe the time frame the person held that office.
-\end{example}
-
-\paragraph{Rules}
-To represent a violation of the \emph{allowed qualifiers} constraint in datalog, inequality between all specified allowed qualifiers and the properties of qualifiers is required.
+The \textsc{head of state} (P35) property has (among others) the \textsc{allowed qualifiers} (P2306) \textsc{start time} (P580) and \textsc{end time} (P582) properties, used to describe the time frame the person held that office.
+\end{example}
+
+\paragraph{Rules}
+To represent a violation of the \textsc{allowed qualifiers} constraint in datalog, inequality between all specified allowed qualifiers and the properties of qualifiers is required.
 The constraint is violated if a qualifier on a constrained statement exists that is unequal to all allowed qualifiers.
@@
 \begin{split}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{} \\
-&\text{qualifier\_EDB}(S, Q, O) {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, Q)
+\textsc{violation\_statement}&(S, E, \textsc{propertyConstant}, V) \leftarrow \\
+&\textsc{statement\_edb}(S, E, \textsc{propertyConstant}, V) \wedge{} \\
+&\textsc{qualifier\_edb}(S, Q, O) {\textstyle\bigwedge}_{a \in A} \textsc{unequal}(a, Q)
 \end{split}
@@
 \paragraph{Constraint}
-The \emph{one-of} constraint limits the values of a constraint statement, qualifier, or reference to a set of given values.
+The \textsc{one-of} constraint limits the values of a constraint statement, qualifier, or reference to a set of given values.
 
@@
 \hline
-Qualifiers & \makecell{\emph{allowed values} (P2305) -- 1..* \\ values that may be used as objects of a constrained statement} \\
+Qualifiers & \makecell{\textsc{allowed values} (P2305) -- 1..* \\ values that may be used as objects of a constrained statement} \\
 \hline
@@
 \begin{example}
-The \emph{wheelchair accessibility} (P2846) property, which describes the accessibility of an location or event, has the three \emph{allowed values} (P2305) \emph{wheelchair accessible} (Q24192067), \emph{wheelchair accessible with help} (Q24192068), and \emph{wheelchair inaccessible} (Q24192069).
+The \textsc{wheelchair accessibility} (P2846) property, which describes the accessibility of an location or event, has the three \textsc{allowed values} (P2305) \textsc{wheelchair accessible} (Q24192067), \textsc{wheelchair accessible with help} (Q24192068), and \textsc{wheelchair inaccessible} (Q24192069).
 \end{example}
@@
 The one-of constraint fits the pattern established by the previous two constraints and disallows certain property-value combinations in statements, qualifiers, and references.
-Inequality is established between all \emph{allowed values} and the values of statements, qualifiers, and references, respectively.
+Inequality is established between all \textsc{allowed values} and the values of statements, qualifiers, and references, respectively.
 With \(A\) the set of allowed values the rules for triple, qualifier, and reference are:
@@
 \begin{split}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, V)
-\end{split}\\
-\begin{split}
-\text{violation\_qualifier}&(S, \text{propertyConstant}, V) \leftarrow \\
-&\text{qualifier\_EDB}(S, \text{propertyConstant}, V) {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, V)
-\end{split}\\
-\begin{split}
-\text{violation\_reference}&(S, \text{propertyConstant}, V) \leftarrow \\
-&\text{reference\_EDB}(S, H, \text{propertyConstant}, V) {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, V)
+\textsc{violation\_statement}&(S, E, \textsc{propertyConstant}, V) \leftarrow \\
+&\textsc{statement\_edb}(S, E, \textsc{propertyConstant}, V) {\textstyle\bigwedge}_{a \in A} \textsc{unequal}(a, V)
+\end{split}\\
+\begin{split}
+\textsc{violation\_qualifier}&(S, \textsc{propertyConstant}, V) \leftarrow \\
+&\textsc{qualifier\_edb}(S, \textsc{propertyConstant}, V) {\textstyle\bigwedge}_{a \in A} \textsc{unequal}(a, V)
+\end{split}\\
+\begin{split}
+\textsc{violation\_reference}&(S, \textsc{propertyConstant}, V) \leftarrow \\
+&\textsc{reference\_edb}(S, H, \textsc{propertyConstant}, V) {\textstyle\bigwedge}_{a \in A} \textsc{unequal}(a, V)
 \end{split}
@@
 \paragraph{Constraint}
-The \emph{one-of qualifier value} constraint restricts which \emph{qualifier values} may be used on a qualifier with a given \emph{qualifier property} on a constrained statement.
+The \textsc{one-of qualifier value} constraint restricts which \textsc{qualifier values} may be used on a qualifier with a given \textsc{qualifier property} on a constrained statement.
 
@@
 \hline
-Qualifiers: & \makecell{\emph{qualifier property} (P2306) -- 1 \\ property that, if used as predicate, constrains a qualifier \\
-\emph{qualifier values} (P2305) -- 1..* \\ values that are allowed for qualifiers on constrained statements \\ with the \emph{qualifier property} as predicate}\\
+Qualifiers: & \makecell{\textsc{qualifier property} (P2306) -- 1 \\ property that, if used as predicate, constrains a qualifier \\
+\textsc{qualifier values} (P2305) -- 1..* \\ values that are allowed for qualifiers on constrained statements \\ with the \textsc{qualifier property} as predicate}\\
 \hline
@@
 \begin{example}
-The \emph{flag bearer} (P3022) property, naming the person carrying a flag at a ceremony, has the \emph{qualifier property} (P2306) \emph{of} (P642) with the \emph{qualifier values} (P2305) \emph{opening ceremony} (Q3010369) and \emph{closing ceremony} (Q13454063).
+The \textsc{flag bearer} (P3022) property, naming the person carrying a flag at a ceremony, has the \textsc{qualifier property} (P2306) \textsc{of} (P642) with the \textsc{qualifier values} (P2305) \textsc{opening ceremony} (Q3010369) and \textsc{closing ceremony} (Q13454063).
 \end{example}
@@
 \begin{split}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{} \\
-&\text{qualifier\_EDB}(S, \text{qualifierProperty}, O) {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, O)
+\textsc{violation\_statement}&(S, E, \textsc{propertyConstant}, V) \leftarrow \\
+&\textsc{statement\_edb}(S, E, \textsc{propertyConstant}, V) \wedge{} \\
+&\textsc{qualifier\_edb}(S, \textsc{qualifierProperty}, O) {\textstyle\bigwedge}_{a \in A} \textsc{unequal}(a, O)
 \end{split}
@@
 
-For statements the pattern consists of two rules, both utilising the same conjunction \(\text{doesNotFulfil}\) whose atoms depend on the constraint type.
+For statements the pattern consists of two rules, both utilising the same conjunction \(\textsc{doesNotFulfil}\) whose atoms depend on the constraint type.
 This conjunction is used to test if the statement does not fulfil the requirement.
@@
 \begin{align}
-&\text{require}(S, \text{id}) \leftarrow \text{first}(S, I) \wedge \text{doesNotFulfil} \\
-&\text{require}(S, \text{id}) \leftarrow \text{next}(O, S) \wedge \text{require}(O, \text{id}) \wedge \text{doesNotFulfil} \\
-&\text{require}(S, P, V, \text{id}) \leftarrow \text{first}(S, P, V) \wedge \text{doesNotFulfil} \\
-&\text{require}(S, P, V, \text{id}) \leftarrow \text{next}(O, X, C, S, P, V) \wedge \text{require}(O, X, C, \text{id}) \wedge \text{doesNotFulfil}
+&\textsc{require}(S, \textsc{id}) \leftarrow \textsc{first}(S, I) \wedge \textsc{doesNotFulfil} \\
+&\textsc{require}(S, \textsc{id}) \leftarrow \textsc{next}(O, S) \wedge \textsc{require}(O, \textsc{id}) \wedge \textsc{doesNotFulfil} \\
+&\textsc{require}(S, P, V, \textsc{id}) \leftarrow \textsc{first}(S, P, V) \wedge \textsc{doesNotFulfil} \\
+&\textsc{require}(S, P, V, \textsc{id}) \leftarrow \textsc{next}(O, X, C, S, P, V) \wedge \textsc{require}(O, X, C, \textsc{id}) \wedge \textsc{doesNotFulfil}
 \end{align}
@@
 \paragraph{Constraint}
-The \emph{item requires statement} constraint requires an additional statement on an item with a statement using the constrained property.
-It can be even more restrictive by also requiring the value of said additional statement to be in a set of \emph{allowed values}.
+The \textsc{item requires statement} constraint requires an additional statement on an item with a statement using the constrained property.
+It can be even more restrictive by also requiring the value of said additional statement to be in a set of \textsc{allowed values}.
 \begin{table}[H]
@@
 \hline
-Qualifiers & \makecell{\emph{required property} (P2306) -- 1 \\ property that must be predicate of an item statement \\
-\emph{allowed value} (P2305) -- 0..* \\ value that must be object of the item statement} \\
+Qualifiers & \makecell{\textsc{required property} (P2306) -- 1 \\ property that must be predicate of an item statement \\
+\textsc{allowed value} (P2305) -- 0..* \\ value that must be object of the item statement} \\
 \hline
@@
 \begin{example}
-The \emph{professorship} (P803) property requires an additional statement with the \emph{required property} (P2306) \emph{occupation} (P106) and the \emph{allowed value} (P2305) \emph{university teacher} (Q1622272).
-\end{example}
-
-\paragraph{Rules}
-The \emph{item requires statement} constraint requires inequality between all properties of statements and the \emph{required property} as well as between all statement values and the \emph{allowed values}.
+The \textsc{professorship} (P803) property requires an additional statement with the \textsc{required property} (P2306) \textsc{occupation} (P106) and the \textsc{allowed value} (P2305) \textsc{university teacher} (Q1622272).
+\end{example}
+
+\paragraph{Rules}
+The \textsc{item requires statement} constraint requires inequality between all properties of statements and the \textsc{required property} as well as between all statement values and the \textsc{allowed values}.
 The order of the item statements needs to be established.
-An item violates the constraint if it a) has a statement using the constrained property as predicate, b1) has no statements using the \emph{required property} as predicate, or (if \emph{allowed values} are specified) b2) has only statement using the \emph{required property} as predicate that do not use one of the \emph{allowed values} as value.
-If no \emph{allowed values} have been specified for the declared constraint the conjunction for the require pattern is one.
-If a set of \emph{allowed values} has been specified, a second set of require-rules have to be introduced with conjunction two.
-Note that the require pattern using conjunction one is still required, since the preceding statements do not need to have the \emph{required property}.
+An item violates the constraint if it a) has a statement using the constrained property as predicate, b1) has no statements using the \textsc{required property} as predicate, or (if \textsc{allowed values} are specified) b2) has only statement using the \textsc{required property} as predicate that do not use one of the \textsc{allowed values} as value.
+If no \textsc{allowed values} have been specified for the declared constraint the conjunction for the require pattern is one.
+If a set of \textsc{allowed values} has been specified, a second set of require-rules have to be introduced with conjunction two.
+Note that the require pattern using conjunction one is still required, since the preceding statements do not need to have the \textsc{required property}.
 In both cases violations can be found with rule three
@@
 \begin{split}
-c = &\text{statement\_EDB}(Q, E, \text{propertyConstant}, X) \wedge{} \\
-&\text{statement\_EDB}(S, E, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
+c = &\textsc{statement\_edb}(Q, E, \textsc{propertyConstant}, X) \wedge{} \\
+&\textsc{statement\_edb}(S, E, P, V) \wedge \textsc{unequal}(\textsc{requiredPropertyConstant}, P)
 \end{split} \\
 \begin{split}
-c = &\text{statement\_EDB}(Q, E, \text{propertyConstant}, X) \wedge{} \\
-&\text{statement\_EDB}(S, E, \text{requiredPropertyConstant}, V) \wedge {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, V)
-\end{split}
-\end{align}
-\begin{align}
-\begin{split}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{} \text{last}(O, I) \wedge{} \\
-&\text{require}(O, \text{propertyConstant}, \text{requiredPropertyConstant})
+c = &\textsc{statement\_edb}(Q, E, \textsc{propertyConstant}, X) \wedge{} \\
+&\textsc{statement\_edb}(S, E, \textsc{requiredPropertyConstant}, V) \wedge {\textstyle\bigwedge}_{a \in A} \textsc{unequal}(a, V)
+\end{split}
+\end{align}
+\begin{align}
+\begin{split}
+\textsc{violation\_statement}&(S, E, \textsc{propertyConstant}, V) \leftarrow \\
+&\textsc{statement\_edb}(S, E, \textsc{propertyConstant}, V) \wedge{} \textsc{last}(O, I) \wedge{} \\
+&\textsc{require}(O, \textsc{propertyConstant}, \textsc{requiredPropertyConstant})
 \end{split}
@@
 \paragraph{Constraint}
-The \emph{value requires statement} constraint requires an additional statement on the value of the constrained statement, possibly also using a specific value from the list of allowed values.
+The \textsc{value requires statement} constraint requires an additional statement on the value of the constrained statement, possibly also using a specific value from the list of allowed values.
 Since the constraint works very similar to item requires statement (see \ref{subsec_item_requires_statement}) --- requiring a statement on the value of the constrained statement instead of on the item --- Table \ref{tab_item_requires_statements} is not repeated.
@@
 \begin{example}
-The \emph{list of monuments} (P1456) property requires its value to have the \emph{required property} (P2306) \emph{is a list of} (P360).
-\end{example}
-
-\paragraph{Rules}
-The \emph{value requires statement} constraint is syntactically very similar to the previous \emph{item requires statement} constraint.
+The \textsc{list of monuments} (P1456) property requires its value to have the \textsc{required property} (P2306) \textsc{is a list of} (P360).
+\end{example}
+
+\paragraph{Rules}
+The \textsc{value requires statement} constraint is syntactically very similar to the previous \textsc{item requires statement} constraint.
 Instead of requiring a statement on the item of a constrained statement, it requires it on the value.
@@
 \begin{split}
-c = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
-&\text{statement\_EDB}(S, E, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
-\end{split}\\
-\begin{split}
-c = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
-&\text{statement\_EDB}(S, E, \text{requiredPropertyConstant}, V) \wedge {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, V)
-\end{split}
-\end{align}
-
-\begin{align}
-\begin{split}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{} \text{last}(O, V) \wedge{} \\
-&\text{require}(O, \text{propertyConstant}, \text{requiredPropertyConstant})
+c = &\textsc{statement\_edb}(Q, R, \textsc{propertyConstant}, I) \wedge{} \\
+&\textsc{statement\_edb}(S, E, P, V) \wedge \textsc{unequal}(\textsc{requiredPropertyConstant}, P)
+\end{split}\\
+\begin{split}
+c = &\textsc{statement\_edb}(Q, R, \textsc{propertyConstant}, I) \wedge{} \\
+&\textsc{statement\_edb}(S, E, \textsc{requiredPropertyConstant}, V) \wedge {\textstyle\bigwedge}_{a \in A} \textsc{unequal}(a, V)
+\end{split}
+\end{align}
+
+\begin{align}
+\begin{split}
+\textsc{violation\_statement}&(S, E, \textsc{propertyConstant}, V) \leftarrow \\
+&\textsc{statement\_edb}(S, E, \textsc{propertyConstant}, V) \wedge{} \textsc{last}(O, V) \wedge{} \\
+&\textsc{require}(O, \textsc{propertyConstant}, \textsc{requiredPropertyConstant})
 \end{split}
@@
 \paragraph{Constraint}
-The \emph{inverse} constraint requires an additional statement on the value of a constrained statement with an \emph{inverse property} and the constrained item as value.
-It is very similar to \emph{value requires statement} (see \ref{subsec_value_requires_statement}), although it does not specify allowed values, since the required statement needs to have one specific value: the item of the constrained statement.
-
-\begin{example}
-The \emph{mother} (P25) property requires values to have a statement with the \emph{inverse property} (P2306) \emph{child} (P40) and the original item as value.
-\end{example}
-
-\paragraph{Rules}
-The \emph{inverse} constraint can be modelled in almost the same way as the \emph{value requires statement} constraint.
-The only change is that instead of requiring inequality to all \emph{allowed values} in conjunction \(c_2\) the statement requires this constraint if its value is not the subject \(R\) of the original constrained statement, meaning \(t = R\).
-Inequalities have to be established between all properties and the \emph{inverse} property as well as all subjects and values of statements.
+The \textsc{inverse} constraint requires an additional statement on the value of a constrained statement with an \textsc{inverse property} and the constrained item as value.
+It is very similar to \textsc{value requires statement} (see \ref{subsec_value_requires_statement}), although it does not specify allowed values, since the required statement needs to have one specific value: the item of the constrained statement.
+
+\begin{example}
+The \textsc{mother} (P25) property requires values to have a statement with the \textsc{inverse property} (P2306) \textsc{child} (P40) and the original item as value.
+\end{example}
+
+\paragraph{Rules}
+The \textsc{inverse} constraint can be modelled in almost the same way as the \textsc{value requires statement} constraint.
+The only change is that instead of requiring inequality to all \textsc{allowed values} in conjunction \(c_2\) the statement requires this constraint if its value is not the subject \(R\) of the original constrained statement, meaning \(t = R\).
+Inequalities have to be established between all properties and the \textsc{inverse} property as well as all subjects and values of statements.
 The final violation rule remains the same.
@@
 \begin{split}
-c = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
-&\text{statement\_EDB}(S, E, P, V) \wedge{} \\
-&\text{unequal}(\text{inversePropertyConstant}, P)
-\end{split}\\
-\begin{split}
-c = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
-&\text{statement\_EDB}(S, E, \text{inversePropertyConstant}, V) \wedge{} \\
-&\text{unequal}(R, V)
+c = &\textsc{statement\_edb}(Q, R, \textsc{propertyConstant}, I) \wedge{} \\
+&\textsc{statement\_edb}(S, E, P, V) \wedge{} \\
+&\textsc{unequal}(\textsc{inversePropertyConstant}, P)
+\end{split}\\
+\begin{split}
+c = &\textsc{statement\_edb}(Q, R, \textsc{propertyConstant}, I) \wedge{} \\
+&\textsc{statement\_edb}(S, E, \textsc{inversePropertyConstant}, V) \wedge{} \\
+&\textsc{unequal}(R, V)
 \end{split}
@@
 \subsection{Symmetric (Q21510862)}
-The \emph{symmetric} constraint requires an additional statement on the value of a constrained statement with the constrained property and the constrained item as value.
-It is thus equal to an \emph{inverse} constraint (see \ref{subsec_2_inverse}) with the same property as constrained property and inverse property.
+The \textsc{symmetric} constraint requires an additional statement on the value of a constrained statement with the constrained property and the constrained item as value.
+It is thus equal to an \textsc{inverse} constraint (see \ref{subsec_2_inverse}) with the same property as constrained property and inverse property.
 It can be translated exactly as the inverse constraint if the inverse property were set to the same as the constrained property.
@@
 \begin{example}
-The \emph{sibling} (P3373) property requires a statement on the value with the same property and the original item as the value.
+The \textsc{sibling} (P3373) property requires a statement on the value with the same property and the original item as the value.
 \end{example}
@@
 \paragraph{Constraint}
-The \emph{multi-value} constraint specifies that the constrained property should appear in multiple statements of an item or none.
+The \textsc{multi-value} constraint specifies that the constrained property should appear in multiple statements of an item or none.
 
@@
 \begin{example}
-The \emph{cast member} (P161) property requires an item to have multiple statements using this property, since most plays or films have multiple cast members.
+The \textsc{cast member} (P161) property requires an item to have multiple statements using this property, since most plays or films have multiple cast members.
 This is an example for a constraint from which exceptions can and should be made and is only meant to point to possibly missing data.
@@
 \paragraph{Rules}
-The \emph{multi-value} constraint requires a limited count of constrained statements.
+The \textsc{multi-value} constraint requires a limited count of constrained statements.
 Alternatively, it could be said that if the item requires a second statement, the constraint is violated.
@@
 \begin{split}
-c = &\text{statement\_EDB}(Q, E, \text{propertyConstant}, X) \wedge{} \\
-&\text{statement\_EDB}(S, E, P, V) \wedge \text{unequal}(\text{propertyConstant}, P)
+c = &\textsc{statement\_edb}(Q, E, \textsc{propertyConstant}, X) \wedge{} \\
+&\textsc{statement\_edb}(S, E, P, V) \wedge \textsc{unequal}(\textsc{propertyConstant}, P)
 \end{split}
@@
 The predicate require\_second\([2]\) is introduced with the same meaning as require except that it will be applied after a required property has been found.
-A violation would be any last statement requiring a second constrained statement, analogous to the violation rule for \emph{item requires statement} (see Section \ref{subsec_item_requires_statement}).
-\begin{align}
-\begin{split}
-\text{require\_second}&(S, \text{propertyConstant}, \text{propertyConstant}) \leftarrow \\
-&\text{next}(R, S) \wedge \text{statement\_EDB}(R, E, \text{propertyConstant}, C) \wedge{} \\
-&\text{next}(Q, R) \wedge \text{require}(Q, \text{propertyConstant}, \text{propertyConstant})  \wedge c
-\end{split}\\
-\begin{split}
-\text{require\_second}&(S, \text{propertyConstant}, \text{propertyConstant}) \leftarrow \text{next}(R, S) \wedge{} \\
-&\text{require\_second}(R, \text{propertyConstant}, \text{propertyConstant}) \wedge c
+A violation would be any last statement requiring a second constrained statement, analogous to the violation rule for \textsc{item requires statement} (see Section \ref{subsec_item_requires_statement}).
+\begin{align}
+\begin{split}
+\textsc{require\_second}&(S, \textsc{propertyConstant}, \textsc{propertyConstant}) \leftarrow \\
+&\textsc{next}(R, S) \wedge \textsc{statement\_edb}(R, E, \textsc{propertyConstant}, C) \wedge{} \\
+&\textsc{next}(Q, R) \wedge \textsc{require}(Q, \textsc{propertyConstant}, \textsc{propertyConstant})  \wedge c
+\end{split}\\
+\begin{split}
+\textsc{require\_second}&(S, \textsc{propertyConstant}, \textsc{propertyConstant}) \leftarrow \textsc{next}(R, S) \wedge{} \\
+&\textsc{require\_second}(R, \textsc{propertyConstant}, \textsc{propertyConstant}) \wedge c
 \end{split}
@@
 \paragraph{Constraint}
-The \emph{single value} constraint states that the constrained property may only be used once per item.
-A constrained statement marked with the single value constraint can have \emph{separators}, which define exceptions under specific circumstances.
+The \textsc{single value} constraint states that the constrained property may only be used once per item.
+A constrained statement marked with the single value constraint can have \textsc{separators}, which define exceptions under specific circumstances.
 \begin{table}[H]
@@
 \hline
-Qualifiers & \makecell{\emph{separator} (P4155) -- 0..* \\ qualifiers that, when used on constrained statements, can warrant an exception} \\
+Qualifiers & \makecell{\textsc{separator} (P4155) -- 0..* \\ qualifiers that, when used on constrained statements, can warrant an exception} \\
 \hline
@@
 \begin{example}
-The \emph{capital} (P36) property should only have one single value per item.
-If conflicts exist, it is not a violation of this constraint if one of the values for the \emph{separators} (P4155) \emph{start time} (P580), \emph{point in time} (P585), \emph{end time} (P582), and \emph{determination method} (P459) are unequal.
+The \textsc{capital} (P36) property should only have one single value per item.
+If conflicts exist, it is not a violation of this constraint if one of the values for the \textsc{separators} (P4155) \textsc{start time} (P580), \textsc{point in time} (P585), \textsc{end time} (P582), and \textsc{determination method} (P459) are unequal.
 \end{example}
@@
 \begin{split}\label{eq_no_separators_triple}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{} \\
-&\text{statement\_EDB}(O, E, \text{propertyConstant}, X) \wedge{} \\
-&\text{unequal}(S, O)
-\end{split}\\
-\begin{split}
-\text{violation\_qualifier}&(S, \text{propertyConstant}, V) \leftarrow \\
-&\text{qualifier\_EDB}(S, \text{propertyConstant}, V) \wedge{} \\
-&\text{qualifier\_EDB}(S, \text{propertyConstant}, X) \wedge{} \\
-&\text{unequal}(V, X)
-\end{split}\\
-\begin{split}
-\text{violation\_reference}&(S, H, \text{propertyConstant}, V) \leftarrow \\
-&\text{reference\_EDB}(S, H, \text{propertyConstant}, V) \wedge{} \\
-&\text{reference\_EDB}(S, J, \text{propertyConstant}, X) \wedge{} \\
-&\text{unequal}(V, X)
+\textsc{violation\_statement}&(S, E, \textsc{propertyConstant}, V) \leftarrow \\
+&\textsc{statement\_edb}(S, E, \textsc{propertyConstant}, V) \wedge{} \\
+&\textsc{statement\_edb}(O, E, \textsc{propertyConstant}, X) \wedge{} \\
+&\textsc{unequal}(S, O)
+\end{split}\\
+\begin{split}
+\textsc{violation\_qualifier}&(S, \textsc{propertyConstant}, V) \leftarrow \\
+&\textsc{qualifier\_edb}(S, \textsc{propertyConstant}, V) \wedge{} \\
+&\textsc{qualifier\_edb}(S, \textsc{propertyConstant}, X) \wedge{} \\
+&\textsc{unequal}(V, X)
+\end{split}\\
+\begin{split}
+\textsc{violation\_reference}&(S, H, \textsc{propertyConstant}, V) \leftarrow \\
+&\textsc{reference\_edb}(S, H, \textsc{propertyConstant}, V) \wedge{} \\
+&\textsc{reference\_edb}(S, J, \textsc{propertyConstant}, X) \wedge{} \\
+&\textsc{unequal}(V, X)
 \end{split}
@@
 \subsubsection{With separators}\label{subsubsec_single_value_with_separators}
-To find all violations under \emph{separators}, a number of things have to be considered.
+To find all violations under \textsc{separators}, a number of things have to be considered.
 Only statements are eligible, since qualifiers cannot have further qualifiers.
@@
 The require qualifier pattern needs to be applied to the qualifiers of constrained statements.
-The conjunction \(\text{doesNotFulfil}\) is created for every \emph{separator} and the pattern applied.
+The conjunction \(\textsc{doesNotFulfil}\) is created for every \textsc{separator} and the pattern applied.
 \begin{equation}
 \begin{split}
-c = &\text{statement\_EDB}(S, E, \text{propertyConstant}, C) \wedge{} \\
-&\text{qualifier\_EDB}(S, P, V) \wedge{} \\
-&\text{unequal}(P, \text{separatorConstant})
+c = &\textsc{statement\_edb}(S, E, \textsc{propertyConstant}, C) \wedge{} \\
+&\textsc{qualifier\_edb}(S, P, V) \wedge{} \\
+&\textsc{unequal}(P, \textsc{separatorConstant})
 \end{split}
@@
 With qualifier non-existence established, the violation can be found.
-Two statements form a violation if they are distinct as per equation \ref{eq_no_separators_triple} and additionally fulfil the condition that all values of \emph{separator} qualifiers are equal or no such \emph{separator} qualifier exist.
-To determine this, a number of rules have to be introduced which are based on two sets the \emph{separators} are split into: \(X\) contains the \emph{separators} that exist as qualifiers and \(Y\) the \emph{separators} that do not.
-The following rule is then created for every possible split of the \emph{separators} into two such sets.
-Note that \(\text{predicate}(\{X\})\) denotes \(\bigwedge_{x \in X} \text{predicate}(x)\) as introduced in Section \ref{subsec_3_allowed_units}.
-\begin{align}
-\begin{split}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{} \\
-&\text{statement\_EDB}(O, E, \text{propertyConstant}, X) \wedge{} \\
-&\text{unequal}(S, O) \wedge{} \\
-&\text{same\_or\_non\_existent}(S, O, \{Q\})
-\end{split}\\
-\begin{split}
-\text{same\_or\_non\_existent}&(S, O, Q) \leftarrow \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, X) \wedge{} \\
-&\text{statement\_EDB}(O, E, \text{propertyConstant}, Y) \wedge{} \\
-&\text{qualifier\_EDB}(S, Q, V) \wedge{} \\
-&\text{qualifier\_EDB}(O, Q, V)
-\end{split}\\
-\begin{split}
-\text{same\_or\_non\_existent}&(S, O, Q) \leftarrow{} \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, X) \wedge{} \\
-&\text{statement\_EDB}(O, E, \text{propertyConstant}, Y) \wedge{} \\
-&\text{last\_qualifier}(S, P, V) \wedge{} \\
-&\text{require}(S, P, V, \text{propertyConstant}, Q) \wedge{} \\
-&\text{last\_qualifier}(O, R, C) \wedge{} \\
-&\text{require}(O, R, C, \text{propertyConstant}, Q)
+Two statements form a violation if they are distinct as per equation \ref{eq_no_separators_triple} and additionally fulfil the condition that all values of \textsc{separator} qualifiers are equal or no such \textsc{separator} qualifier exist.
+To determine this, a number of rules have to be introduced which are based on two sets the \textsc{separators} are split into: \(X\) contains the \textsc{separators} that exist as qualifiers and \(Y\) the \textsc{separators} that do not.
+The following rule is then created for every possible split of the \textsc{separators} into two such sets.
+Note that \(\textsc{predicate}(\{X\})\) denotes \(\bigwedge_{x \in X} \textsc{predicate}(x)\) as introduced in Section \ref{subsec_3_allowed_units}.
+\begin{align}
+\begin{split}
+\textsc{violation\_statement}&(S, E, \textsc{propertyConstant}, V) \leftarrow \\
+&\textsc{statement\_edb}(S, E, \textsc{propertyConstant}, V) \wedge{} \\
+&\textsc{statement\_edb}(O, E, \textsc{propertyConstant}, X) \wedge{} \\
+&\textsc{unequal}(S, O) \wedge{} \\
+&\textsc{same\_or\_non\_existent}(S, O, \{Q\})
+\end{split}\\
+\begin{split}
+\textsc{same\_or\_non\_existent}&(S, O, Q) \leftarrow \\
+&\textsc{statement\_edb}(S, E, \textsc{propertyConstant}, X) \wedge{} \\
+&\textsc{statement\_edb}(O, E, \textsc{propertyConstant}, Y) \wedge{} \\
+&\textsc{qualifier\_edb}(S, Q, V) \wedge{} \\
+&\textsc{qualifier\_edb}(O, Q, V)
+\end{split}\\
+\begin{split}
+\textsc{same\_or\_non\_existent}&(S, O, Q) \leftarrow{} \\
+&\textsc{statement\_edb}(S, E, \textsc{propertyConstant}, X) \wedge{} \\
+&\textsc{statement\_edb}(O, E, \textsc{propertyConstant}, Y) \wedge{} \\
+&\textsc{last\_qualifier}(S, P, V) \wedge{} \\
+&\textsc{require}(S, P, V, \textsc{propertyConstant}, Q) \wedge{} \\
+&\textsc{last\_qualifier}(O, R, C) \wedge{} \\
+&\textsc{require}(O, R, C, \textsc{propertyConstant}, Q)
 \end{split}
@@
 \paragraph{Constraint}
-The \emph{mandatory qualifier} constraint requires that a constrained statement has a qualifier with a \emph{required property}.
+The \textsc{mandatory qualifier} constraint requires that a constrained statement has a qualifier with a \textsc{required property}.
 
@@
 \hline
-Qualifiers & \makecell{\emph{required property} (P2306) -- 1 \\ property that must appear as predicate of a qualifier on the constrained statement} \\
+Qualifiers & \makecell{\textsc{required property} (P2306) -- 1 \\ property that must appear as predicate of a qualifier on the constrained statement} \\
 \hline
@@
 \begin{example}
-The \emph{head of state} (P35) property has the \emph{required property} (P2306) \emph{start time} (P580), meaning every \emph{head of state} must have assumed that position at some time.
-\end{example}
-
-\paragraph{Rules}
-The \emph{mandatory qualifier} constraint needs to determine qualifier non-existence on constrained statements to find violations, which is similar to Section \ref{subsubsec_single_value_with_separators}.
+The \textsc{head of state} (P35) property has the \textsc{required property} (P2306) \textsc{start time} (P580), meaning every \textsc{head of state} must have assumed that position at some time.
+\end{example}
+
+\paragraph{Rules}
+The \textsc{mandatory qualifier} constraint needs to determine qualifier non-existence on constrained statements to find violations, which is similar to Section \ref{subsubsec_single_value_with_separators}.
 The qualifier order needs to be provided and inequality must be established between all qualifier properties.
@@
 \begin{split}
-c = &\text{statement\_EDB}(S, E, \text{propertyConstant}, C) \wedge{} \\
-&\text{qualifier\_EDB}(S, P, V) \wedge \text{unequal}(P, \text{requiredPropertyConstant})
-\end{split}\\
-\begin{split}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge \\
-&\text{does\_not\_have}(S, \text{requiredPropertyConstant})
+c = &\textsc{statement\_edb}(S, E, \textsc{propertyConstant}, C) \wedge{} \\
+&\textsc{qualifier\_edb}(S, P, V) \wedge \textsc{unequal}(P, \textsc{requiredPropertyConstant})
+\end{split}\\
+\begin{split}
+\textsc{violation\_statement}&(S, E, \textsc{propertyConstant}, V) \leftarrow \\
+&\textsc{statement\_edb}(S, E, \textsc{propertyConstant}, V) \wedge \\
+&\textsc{does\_not\_have}(S, \textsc{requiredPropertyConstant})
 \end{split}
@@
 \paragraph{Constraint}
-The \emph{single best value} constraint requires that only one constrained statement per item has the rank preferred.
+The \textsc{single best value} constraint requires that only one constrained statement per item has the rank preferred.
 
@@
 \hline
-Qualifiers & \makecell{\emph{separator} (P4155) -- 0..* \\ qualifiers that, when used on constrained statements, can warrant an exception } \\
+Qualifiers & \makecell{\textsc{separator} (P4155) -- 0..* \\ qualifiers that, when used on constrained statements, can warrant an exception } \\
 \hline
@@
 \begin{example}
-The \emph{property proposal discussion} (P3254) property, denoting the URL of a web page where the creation of a Wikidata property was discussed should only have one best value.
+The \textsc{property proposal discussion} (P3254) property, denoting the URL of a web page where the creation of a Wikidata property was discussed should only have one best value.
 Since a discussion might have been moved to a different page at some point, the best value should be the page the discussion was last moved to.
@@
 \paragraph{Rules}
-The \emph{single best value} constraint, like the \emph{single value} constraint (Section \ref{subsec_single_value}), differentiates between the set of \emph{separators} \(S\) being empty or not.
+The \textsc{single best value} constraint, like the \textsc{single value} constraint (Section \ref{subsec_single_value}), differentiates between the set of \textsc{separators} \(S\) being empty or not.
 Since the constraint depends on the rank of statements, this needs to be provided.
@@
 \subsubsection{No separators}
-If the \emph{separator} set is empty, the violation simply occurs if there are two constrained statements on the same item with the preferred rank.
+If the \textsc{separator} set is empty, the violation simply occurs if there are two constrained statements on the same item with the preferred rank.
 Thus the rule \eqref{eq_no_separators_triple} is expanded to
@@
 \begin{split}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{} \\
-&\text{rank\_EDB}(S, \text{preferredRankConstant}) \wedge{} \\
-&\text{statement\_EDB}(O, E, \text{propertyConstant}, X) \wedge{} \\
-&\text{rank\_EDB}(O, \text{preferredRankConstant}) \wedge{} \\
-&\text{unequal}(S, O)
+\textsc{violation\_statement}&(S, E, \textsc{propertyConstant}, V) \leftarrow \\
+&\textsc{statement\_edb}(S, E, \textsc{propertyConstant}, V) \wedge{} \\
+&\textsc{rank\_edb}(S, \textsc{preferredRankConstant}) \wedge{} \\
+&\textsc{statement\_edb}(O, E, \textsc{propertyConstant}, X) \wedge{} \\
+&\textsc{rank\_edb}(O, \textsc{preferredRankConstant}) \wedge{} \\
+&\textsc{unequal}(S, O)
 \end{split}
@@
 \subsubsection{With separators}
-As in the case above, the qualifier non-existence and violation rules are very similar to the \emph{single value} constraint. The only change is the addition of the rank requirements as shown below.
-\begin{align}
-\begin{split}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{} \\
-&\text{rank\_EDB}(S, \text{preferredRankConstant}) \wedge{} \\
-&\text{statement\_EDB}(O, E, \text{propertyConstant}, W) \wedge{} \\
-&\text{rank\_EDB}(O, \text{preferredRankConstant}) \wedge{} \\
-&\text{unequal}(S, O) \wedge{} \\
-&\text{has\_same}(S, O, \{X\}) \wedge{} \\
-&\text{does\_not\_have}(S, \{Y\}) \wedge{} \\
-&\text{does\_not\_have}(O, \{Y\})
+As in the case above, the qualifier non-existence and violation rules are very similar to the \textsc{single value} constraint. The only change is the addition of the rank requirements as shown below.
+\begin{align}
+\begin{split}
+\textsc{violation\_statement}&(S, E, \textsc{propertyConstant}, V) \leftarrow \\
+&\textsc{statement\_edb}(S, E, \textsc{propertyConstant}, V) \wedge{} \\
+&\textsc{rank\_edb}(S, \textsc{preferredRankConstant}) \wedge{} \\
+&\textsc{statement\_edb}(O, E, \textsc{propertyConstant}, W) \wedge{} \\
+&\textsc{rank\_edb}(O, \textsc{preferredRankConstant}) \wedge{} \\
+&\textsc{unequal}(S, O) \wedge{} \\
+&\textsc{has\_same}(S, O, \{X\}) \wedge{} \\
+&\textsc{does\_not\_have}(S, \{Y\}) \wedge{} \\
+&\textsc{does\_not\_have}(O, \{Y\})
 \end{split}
@@
 Multiple constraints have not been described above for different reasons.
-The most prominent of these is the \emph{format} constraint (Q21502404) which specifies a regular expression the values of the constrained property must adhere to.
+The most prominent of these is the \textsc{format} constraint (Q21502404) which specifies a regular expression the values of the constrained property must adhere to.
 Evaluating these expressions would be theoretically possible but \\
@@
 
-The two type constraints \emph{type} (Q21503250) and \emph{value type} (Q21510865) require an entity to have a certain type.
-This is defined transitively via the properties \emph{instance of} (P31) and \emph{subclass of} (P279), which have to form a statement chain to the entity given as the type.
+The two type constraints \textsc{type} (Q21503250) and \textsc{value type} (Q21510865) require an entity to have a certain type.
+This is defined transitively via the properties \textsc{instance of} (P31) and \textsc{subclass of} (P279), which have to form a statement chain to the entity given as the type.
 Evaluating these constraints would require pre-processing to compute all types of an entity transitively, add these as statements to the data and establish their order to allow testing for the non-existence of a type.
@@
 
-The \emph{integer} constraint (Q52848401), which requires values of a constrained property be integers (notably a quantity without decimal places), could easily be seen as a specific format constraint.
+The \textsc{integer} constraint (Q52848401), which requires values of a constrained property be integers (notably a quantity without decimal places), could easily be seen as a specific format constraint.
 While precomputing if a value is an integer or not could be done, it would mean that no actually relevant rules would be needed for the evaluation in datalog.
 
-Another problem that could not be sensibly solved is the comparison of values, needed for \emph{range} (Q21510860), \emph{difference within range} (Q21510854) and \emph{contemporary} constraint (Q25796498).
-\emph{Range} restricts the values of a constrained property within a certain range, \emph{difference within range} specifies that the values of two item statements should be within a certain range and \emph{contemporary} constraint specifies that subject and object of a constrained statement have to have coexisted at some point in time.
+Another problem that could not be sensibly solved is the comparison of values, needed for \textsc{range} (Q21510860), \textsc{difference within range} (Q21510854) and \textsc{contemporary} constraint (Q25796498).
+\textsc{Range} restricts the values of a constrained property within a certain range, \textsc{difference within range} specifies that the values of two item statements should be within a certain range and \textsc{contemporary} constraint specifies that subject and object of a constrained statement have to have coexisted at some point in time.
 While it would be possible to extract all values that might need to be compared and precompute the comparison, it would make most of the work precomputing and leave little to be done in datalog.
 
-\emph{Citation needed} constraint (Q54554025), stating that constraint statements must have one or more references, requires a negation that cannot be concluded except by explicitly precomputing it, which makes the evaluation unnecessary.
+\textsc{Citation needed} constraint (Q54554025), stating that constraint statements must have one or more references, requires a negation that cannot be concluded except by explicitly precomputing it, which makes the evaluation unnecessary.
 \\
-A similar situation is the \emph{no bounds} constraint (Q51723761), which requires that quantity values have neither upper nor lower bound.
+A similar situation is the \textsc{no bounds} constraint (Q51723761), which requires that quantity values have neither upper nor lower bound.
 While it would be possible to precompute if a statement has no upper or lower bound, this would shift basically all evaluation logic into precomputing.
 
-And finally \emph{commons link} constraint (Q21510852) requires that values for a property should be valid links to to Wikimedia Commons.
+And finally \textsc{commons link} constraint (Q21510852) requires that values for a property should be valid links to to Wikimedia Commons.
 This validity cannot be tested from within datalog since there is no way to query the web for information.
@@
 Most constraints could be evaluated in under a minute.
-The exceptions are \emph{distinct values} (1.22 hours), \emph{item requires statement} (2.38 hours) and \emph{single value} (4.22 hours).
-In part this is expected behaviour, since these three constraints are applied to an order of magnitude more properties than all other properties with the exception of \emph{property scope}, which is described in very simple rules.
+The exceptions are \textsc{distinct values} (1.22 hours), \textsc{item requires statement} (2.38 hours) and \textsc{single value} (4.22 hours).
+In part this is expected behaviour, since these three constraints are applied to an order of magnitude more properties than all other properties with the exception of \textsc{property scope}, which is described in very simple rules.
 However, also seen in Figure \ref{fig_runtime}, the runtimes of these three constraints are multiple orders of magnitude higher than those of other constraints.
@@
 
-The long runtimes of \emph{distinct values}, \emph{item requires statement}, and \emph{single value} cannot be explained through the size of the input data.
+The long runtimes of \textsc{distinct values}, \textsc{item requires statement}, and \textsc{single value} cannot be explained through the size of the input data.
 While all of them have more than 10 million input facts, this is also the case for other constraints that do not take nearly as much runtime.
-\emph{One-of}, \emph{symmetric}, and \emph{inverse} have even more input facts than the three long-runtime constraints.
+\textsc{One-of}, \textsc{symmetric}, and \textsc{inverse} have even more input facts than the three long-runtime constraints.
 The reason for the performance difference must be found elsewhere.
@@
 Figure \ref{fig_add_derivations} shows that the three performance-heavy constraints do generate by far the most additional derivations.
-The constraint with the smallest number of additional derivations, \emph{distinct values}, is almost an order of magnitude higher than the next highest, \emph{one-of}.
+The constraint with the smallest number of additional derivations, \textsc{distinct values}, is almost an order of magnitude higher than the next highest, \textsc{one-of}.
 
@@
 The exception are the three constraints focused on in Section \ref{sec_results}.
-As shown in that Section, the main difference between \emph{distinct values}, \emph{item requires statement}, and \emph{single value} and other constraints is
+As shown in that Section, the main difference between \textsc{distinct values}, \textsc{item requires statement}, and \textsc{single value} and other constraints is
 \begin{enumerate}[a)]
@@
 \subsection{Distinct values}
-While the rules for the \emph{distinct values} constraint (see Section \ref{subsec_distinct_values}) are relatively simple in writing, they do require a large amount of processing.
+While the rules for the \textsc{distinct values} constraint (see Section \ref{subsec_distinct_values}) are relatively simple in writing, they do require a large amount of processing.
 Distinct values requires the comparison of every constrained statement with every other constrained statement, meaning its evaluation would scale badly due to the quadratic growth of statements to be compared.
@@
 \subsection{Item requires statement}\label{subsec_findings_item_requires_statement}
-The \emph{item requires statement} constraint (see Section \ref{subsec_item_requires_statement}) creates additional facts about every statement on an entity with a constrained statement (if the constraint was violated on that entity).
-Of interest to note is that the \emph{value requires statement} constraint (see Section \ref{subsec_value_requires_statement}), which works almost identically, does not show a performance nearly as bad as \emph{item requires statement}.
-\emph{Value requires statement} has, however, only 244 constrained properties, while \emph{item requires statement} has 4265.
+The \textsc{item requires statement} constraint (see Section \ref{subsec_item_requires_statement}) creates additional facts about every statement on an entity with a constrained statement (if the constraint was violated on that entity).
+Of interest to note is that the \textsc{value requires statement} constraint (see Section \ref{subsec_value_requires_statement}), which works almost identically, does not show a performance nearly as bad as \textsc{item requires statement}.
+\textsc{Value requires statement} has, however, only 244 constrained properties, while \textsc{item requires statement} has 4265.
 This shows that the require approach scales badly with increasing number of require rules.
@@
 \subsection{Single value}
-The \emph{single value} constraint (see Section \ref{subsec_single_value}) has to be analysed with regards to the two cases: without separators (Section \ref{subsubsec_single_value_no_separators}) and with separators (Section \ref{subsubsec_single_value_with_separators}.
+The \textsc{single value} constraint (see Section \ref{subsec_single_value}) has to be analysed with regards to the two cases: without separators (Section \ref{subsubsec_single_value_no_separators}) and with separators (Section \ref{subsubsec_single_value_with_separators}.
 
@@
 If no separators were given, the rules simply compare every statement on an entity with a constrained statement to said constrained statement.
-Since entity size does not increase with dataset size, this should not be the reason for \emph{single values} bad performance.
+Since entity size does not increase with dataset size, this should not be the reason for \textsc{single values} bad performance.
 
 \paragraph{With separators}
-The \emph{single value} constraint, like the \emph{item requires statement} constraint, uses a require approach to determine non-existence.
-For \emph{single value} this is qualifier non-existence on constrained statements.
+The \textsc{single value} constraint, like the \textsc{item requires statement} constraint, uses a require approach to determine non-existence.
+For \textsc{single value} this is qualifier non-existence on constrained statements.
 As seen in Section \ref{subsec_findings_item_requires_statement}, this could be a reason for the bad performance of the analysis for this constraint.
-Another source of processing time for \emph{single value} are the has\_same facts, which need to compare all constrained statements on the same item and all their qualifiers.
-Of interest to note is that the \emph{single best value} constraint (see Section \ref{subsec_single_best_value}) works in almost the same way but does not suffer from the same performance.
-Again, the number of constrained properties is significantly lower for \emph{single best value} (203) compared to \emph{single value} (2909).
+Another source of processing time for \textsc{single value} are the has\_same facts, which need to compare all constrained statements on the same item and all their qualifiers.
+Of interest to note is that the \textsc{single best value} constraint (see Section \ref{subsec_single_best_value}) works in almost the same way but does not suffer from the same performance.
+Again, the number of constrained properties is significantly lower for \textsc{single best value} (203) compared to \textsc{single value} (2909).
 
@@
 \begin{example}
-The predicate \emph{statement\_EDB} lists the \emph{statement ID}, \emph{entity ID}, \emph{property ID} and \emph{value}.
+The predicate \textsc{statement\_edb} lists the \textsc{statement ID}, \textsc{entity ID}, \textsc{property ID} and \textsc{value}.
 To optimize this for VLog, a new order must be created.
-\emph{property ID} should be the first entry, because there are only \todo{number of properties and link} different properties on Wikidata as of September 2018.
-The \emph{statement ID} needs to be moved to the last position since it is unique for all statements, meaning for any data source with \(n\) statements there would be \(n\) unique \emph{statement IDs}.
-Finally \emph{entity ID} should be before \emph{value}, because every \emph{entity ID} could also be a \emph{value}, but value can take additional string values.
-
-The final, optimized order is then: \emph{property ID}, \emph{entity ID}, \emph{value}, \emph{statement ID}.
+\textsc{property ID} should be the first entry, because there are only \todo{number of properties and link} different properties on Wikidata as of September 2018.
+The \textsc{statement ID} needs to be moved to the last position since it is unique for all statements, meaning for any data source with \(n\) statements there would be \(n\) unique \textsc{statement IDs}.
+Finally \textsc{entity ID} should be before \textsc{value}, because every \textsc{entity ID} could also be a \textsc{value}, but value can take additional string values.
+
+The final, optimized order is then: \textsc{property ID}, \textsc{entity ID}, \textsc{value}, \textsc{statement ID}.
 \end{example}

### 2018-09-20 thesis/paper/paper.tex
Changed font of predicates and constants

--- 2018-09-20 thesis/paper/paper.tex
@@
 \emph{Earth} (Q2) or \emph{head of government} (P6), could be represented as a constant.
-Relations in datalog are expressed using \emph{predicates} with a specified arity \(n\), written as predicate\([n]\).
-To stay close to Wikidata in our examples, let us take the predicate triple\([3]\) as an example.
+Relations in datalog are expressed using \emph{predicates} with a specified arity \(n\), written as \textsc{predicate}\([n]\).
+To stay close to Wikidata in our examples, let us take the predicate \textsc{triple}\([3]\) as an example.
 This would be used to express the \emph{item}-\emph{property}-\emph{value} relation with each position representing one part of the triple.\\
@@
 Constants and variables are collectively known as \emph{terms} and are, together with predicates, used to form \emph{atoms}.
-An atom has the form predicate\((t_1, \ldots , t_n)\) with \(n\) as the arity of the predicate and \(t_1,\ldots,t_n\) terms.\\
+An atom has the form \textsc{predicate}\((t_1, \ldots , t_n)\) with \(n\) as the arity of the predicate and \(t_1,\ldots,t_n\) terms.\\
 Atoms are the building blocks of \emph{rules}, which are used to derive information in datalog. A rule is a premise that implies a conclusion, and both are conjunctions of atoms.
@@
 
-Within the rule predicates we can differentiate between two different schemas: the \emph{extensional schema} and the \emph{intensional schema}, called EDB and IDB (extensional/intensional database).
-An EDB predicate may only occur in the premise of a rule, meaning it can never be derived.
-All predicates that appear in the conclusion at least once are IDB predicates.
+Within the rule predicates we can differentiate between two different schemas: the \emph{extensional schema} and the \emph{intensional schema}, called \textsc{edb} and \textsc{idb} (extensional/intensional database).
+An \textsc{edb} predicate may only occur in the premise of a rule, meaning it can never be derived.
+All predicates that appear in the conclusion at least once are \textsc{idb} predicates.
 
@@
 The individual entities and properties are introduced with their ID beforehand, so the example can be retraced.
-The species \emph{blue whale} (Q42196) has the \emph{parent taxon} (P171) \emph{balaenoptera} (Q133320), which in turn has the parent taxon \emph{rorqual} (Q232829).
-The predicates used are triple\_EDB\([3]\) as the input and isInTaxon\_IDB\([2]\), which will be derived.
+The species \emph{blueWhale} (Q42196) has the \emph{parentTaxon} (P171) \emph{balaenoptera} (Q133320), which in turn has the parent taxon \emph{rorqual} (Q232829).
+The predicates used are \textsc{triple\_edb}\([3]\) as the input and \textsc{is\_in\_taxon\_idb}\([2]\), which will be derived.
 The basic facts are:
-\[\text{triple\_EDB}(\text{blue whale}, \text{parent taxon}, \text{balaenoptera})\]
-\[\text{triple\_EDB}(\text{balaenoptera}, \text{parent taxon}, \text{rorqual})\]
-From this we can draw conclusions. The first step is to represent the already existing facts as IDB:
-\[\text{isInTaxon\_IDB}(X, Y) \leftarrow \text{triple\_EDB}(X, \text{parent taxon}, Y)\]
+\[\textsc{triple\_edb}(\emph{blueWhale}, \emph{parentTaxon}, \emph{balaenoptera})\]
+\[\textsc{triple\_edb}(\emph{balaenoptera}, \emph{parenTaxon}, \emph{rorqual})\]
+From this we can draw conclusions. The first step is to represent the already existing facts as idb:
+\[\textsc{is\_in\_taxon\_idb}(X, Y) \leftarrow \textsc{triple\_edb}(X, \emph{parentTaxon}, Y)\]
 Finally the transitive closure of this relation can be derived with
-\[\text{isInTaxon\_IDB}(X, Y) \leftarrow \text{isInTaxon\_IDB}(X, Z) \wedge \text{isInTaxon\_IDB}(Z, Y)\]
+\[\textsc{is\_in\_taxon\_idb}(X, Y) \leftarrow \textsc{is\_in\_taxon\_idb}(X, Z) \wedge \textsc{is\_in\_taxon\_idb}(Z, Y)\]
 \end{example}
@@
 \begin{align*}
-&\text{violation\_statement}(S, E, \text{propertyConstant}, V) &\leftarrow \text{atoms determining the violation} \\
-&\text{violation\_qualifier}(S, \text{propertyConstant}, V) &\leftarrow \text{atoms determining the violation} \\
-&\text{violation\_reference}(S, H, \text{propertyConstant}, V) &\leftarrow \text{atoms determining the violation}
+&\textsc{violation\_statement}(S, E, \emph{constrainedProperty}, V) &\leftarrow \emph{atoms determining the violation} \\
+&\textsc{violation\_qualifier}(S, \emph{constrainedProperty}, V) &\leftarrow \emph{atoms determining the violation} \\
+&\textsc{violation\_reference}(S, H, \emph{constrainedProperty}, V) &\leftarrow \emph{atoms determining the violation}
 \end{align*}
@@
 \begin{split}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, V)\\
-\end{split}\\
-\begin{split}
-\text{violation\_qualifier}&(S, \text{propertyConstant}, V) \leftarrow \\
-&\text{qualifier\_EDB}(S, \text{propertyConstant}, V)\\
-\end{split}\\
-\begin{split}
-\text{violation\_reference}&(S, \text{propertyConstant}, V) \leftarrow \\
-&\text{reference\_EDB}(S, H, \text{propertyConstant}, V)\\
+\textsc{violation\_statement}&(S, E, \emph{constrainedProperty}, V) \leftarrow \\
+&\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, V)\\
+\end{split}\\
+\begin{split}
+\textsc{violation\_qualifier}&(S, \emph{constrainedProperty}, V) \leftarrow \\
+&\textsc{qualifier\_edb}(S, \emph{constrainedProperty}, V)\\
+\end{split}\\
+\begin{split}
+\textsc{violation\_reference}&(S, \emph{constrainedProperty}, V) \leftarrow \\
+&\textsc{reference\_edb}(S, H, \emph{constrainedProperty}, V)\\
 \end{split}
@@
 \begin{split}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow\\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{}\\
-&\text{statement\_EDB}(O, E, \text{conflictingPropertyConstant}, C)
-\end{split}\\
-\begin{split}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow\\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{}\\
-&\text{statement\_EDB}(O, E, \text{conflictingPropertyConstant}, \text{conflictingValueConstant})
+\textsc{violation\_statement}&(S, E, \emph{constrainedProperty}, V) \leftarrow\\
+&\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, V) \wedge{}\\
+&\textsc{statement\_edb}(O, E, \emph{conflictingProperty}, C)
+\end{split}\\
+\begin{split}
+\textsc{violation\_statement}&(S, E, \emph{constrainedProperty}, V) \leftarrow\\
+&\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, V) \wedge{}\\
+&\textsc{statement\_edb}(O, E, \emph{conflictingProperty}, \emph{conflictingValue})
 \end{split}
@@
 
-Since there are only three options (and only two in use), violations can be found without inequality by finding the statement\_EDB with the constrained property whose subject is of the type that is not allowed, meaning either \emph{Wikibase item} or \emph{Wikibase property}.
+Since there are only three options (and only two in use), violations can be found without inequality by finding the \textsc{statement\_edb} with the constrained property whose subject is of the type that is not allowed, meaning either \emph{Wikibase item} or \emph{Wikibase property}.
 
@@
 \begin{split}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge \text{item}(I)
-\end{split}\\
-\begin{split}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge \text{property}(I)
+\textsc{violation\_statement}&(S, E, \emph{constrainedProperty}, V) \leftarrow \\
+&\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, V) \wedge \textsc{item\_edb}(I)
+\end{split}\\
+\begin{split}
+\textsc{violation\_statement}&(S, E, \emph{constrainedProperty}, V) \leftarrow \\
+&\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, V) \wedge \textsc{property\_edb}(I)
 \end{split}
@@
 \begin{split}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, \text{forbiddenValueConstant})
-\end{split}\\
-\begin{split}
-\text{violation\_qualifier}&(S, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
-&\text{qualifier\_EDB}(S, \text{propertyConstant}, \text{forbiddenValueConstant})
-\end{split}\\
-\begin{split}
-\text{violation\_reference}&(S, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
-&\text{reference\_EDB}(S, H, \text{propertyConstant}, \text{forbiddenValueConstant})
+\textsc{violation\_statement}&(S, E, \emph{constrainedProperty}, \emph{forbiddenValue}) \leftarrow \\
+&\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, \emph{forbiddenValue})
+\end{split}\\
+\begin{split}
+\textsc{violation\_qualifier}&(S, \emph{constrainedProperty}, \emph{forbiddenValue}) \leftarrow \\
+&\textsc{qualifier\_edb}(S, \emph{constrainedProperty}, \emph{forbiddenValue})
+\end{split}\\
+\begin{split}
+\textsc{violation\_reference}&(S, \emph{constrainedProperty}, \emph{forbiddenValue}) \leftarrow \\
+&\textsc{reference\_edb}(S, H, \emph{constrainedProperty}, \emph{forbiddenValue})
 \end{split}
@@
 In every approach it will be necessary to load some basic inequalities into the database.
-To this end the predicate unequal\_EDB\([2]\) is introduced.
-Since the goal is to keep the number of applications of this predicate small, an IDB predicate unequal\([2]\) has to be introduced.
+To this end the predicate unequal\_edb\([2]\) is introduced.
+Since the goal is to keep the number of applications of this predicate small, an \emph{idb} predicate unequal\([2]\) has to be introduced.
 This leads to the rule
 \begin{align}
-\text{unequal}(X, Y) \leftarrow \text{unequal\_EDB}(X, Y)
-\end{align}
-
-Given a set of unequal constants \(U\), the task is to generate rules so that for all \( x, y \in U: x \ne y \Rightarrow \text{unequal}(x, y)\).
+\textsc{unequal}(X, Y) \leftarrow \emph{unequal\_edb}(X, Y)
+\end{align}
+
+Given a set of unequal constants \(U\), the task is to generate rules so that for all \( x, y \in U: x \ne y \Rightarrow \textsc{unequal}(x, y)\).
 
@@
 \begin{align}
-\text{for all } x, y \in U, x \ne y \text{ add } \text{unequal\_EDB}(x, y)
+\emph{for all } x, y \in U, x \ne y \emph{ add } \emph{unequal\_edb}(x, y)
 \end{align}
@@
 \begin{align}
-\text{for all } x, y \in U, x < y \text{ add } \text{unequal\_EDB}(x, y)
+\emph{for all } x, y \in U, x < y \emph{ add } \emph{unequal\_edb}(x, y)
 \end{align}
@@
 Then \(l_{\max} = \max\limits_{\forall u \in U}(len(u))\) is the length of the longest entity string that occurs in the given data.
-Based on this a group of predicates letter\(_i\)\_EDB[8] is introduced where \(0 \equiv i \mod 7\), \(0 \leq i < l_{max} \).
-
-For each unequal constant string \(u = c_0 \ldots c_{l - 1}\) and \(0 \equiv i \mod 7, 0 \leq i < len(u)\) an atom using predicate letter\(_i\)\_EDB is created as 
-\begin{align}
-\begin{split}
-\text{letter}_i\text{\_EDB}(e, f(e, i), f(e, i+1), \ldots, f(e, i+6)
+Based on this a group of predicates letter\(_i\)\_edb[8] is introduced where \(0 \equiv i \mod 7\), \(0 \leq i < l_{max} \).
+
+For each unequal constant string \(u = c_0 \ldots c_{l - 1}\) and \(0 \equiv i \mod 7, 0 \leq i < len(u)\) an atom using predicate \emph{letter\(_i\)\_edb} is created as 
+\begin{align}
+\begin{split}
+\emph{letter}_i\emph{\_edb}(e, f(e, i), f(e, i+1), \ldots, f(e, i+6)
 \end{split}
@@
       c_i & i < len(e) \\
-      \text{none} & i \geq len(e)
+      \emph{none} & i \geq len(e)
 \end{cases}
@@
 
-Now that inequality between all characters and the link between entities and their characters are introduced, the inequality between two entities can be derived by creating the following rules for all predicates letter\(_i\)\_EDB (with \text{letters} short for \(\text{letter}_i\text{\_EDB}(X, A_i, \ldots, A_{i+6}) \wedge\text{letter}_i\text{\_EDB}(Y, B_i, \ldots, B_{i+6})\), \(i \leq j < i + 7\)):
-\begin{align}
-\text{unequal}(X, Y) \leftarrow \text{letters} \wedge \text{unequal}(A_j, B_j)
-\end{align}
-Note that the facts fulfilling \(\text{unequal}(A_j, B_j)\) are created by using the approach outlined in Section \ref{subsubsec_naive} for the set of characters \(C\) in all unequal constants \(U\).
+Now that inequality between all characters and the link between entities and their characters are introduced, the inequality between two entities can be derived by creating the following rules for all predicates \emph{letter\(_i\)\_edb} (with \emph{letters} short for \(\emph{letter}_i\emph{\_edb}(X, A_i, \ldots, A_{i+6}) \wedge\emph{letter}_i\emph{\_edb}(Y, B_i, \ldots, B_{i+6})\), \(i \leq j < i + 7\)):
+\begin{align}
+\textsc{unequal}(X, Y) \leftarrow \emph{letters} \wedge \textsc{unequal}(A_j, B_j)
+\end{align}
+Note that the facts fulfilling \(\textsc{unequal}(A_j, B_j)\) are created by using the approach outlined in Section \ref{subsubsec_naive} for the set of characters \(C\) in all unequal constants \(U\).
 
@@
 \begin{split}
-\text{unequal}(X, Y) \leftarrow \text{req\_inequality}	(X, Y) \wedge \text{letters} \wedge \text{unequal}(A_j, B_j)
+\textsc{unequal}(X, Y) \leftarrow \textsc{req\_inequality}	(X, Y) \wedge \emph{letters} \wedge \textsc{unequal}(A_j, B_j)
 \end{split}
@@
 The \emph{distinct values} constraint requires two distinct constrained statements with the same value. 
-Two constrained statements with the same value can be found via a simple conjunction of two statement\_EDB atoms with the propertyConstant as predicate and a variable \(V\) as value.
+Two constrained statements with the same value can be found via a simple conjunction of two \textsc{statement\_edb} atoms with the propertyConstant as predicate and a variable \(V\) as value.
 Since this would mark every constrained statement as a violation because it can fulfil both atoms at the same time, inequality of the statement IDs \(S\) and \(O\) must be tested as well.
@@
 \begin{split}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{} \\
-&\text{statement\_EDB}(O, X, \text{propertyConstant}, V) \wedge{} \\
-&\text{unequal}(S, O)
+\textsc{violation\_statement}&(S, E, \emph{constrainedProperty}, V) \leftarrow \\
+&\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, V) \wedge{} \\
+&\textsc{statement\_edb}(O, X, \emph{constrainedProperty}, V) \wedge{} \\
+&\textsc{unequal}(S, O)
 \end{split}
@@
 \begin{split}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{} \\
-&\text{unit}(V, U)  {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, U)
-\end{split}\\
-\begin{split}
-\text{violation\_qualifier}&(S, \text{propertyConstant}, V) \leftarrow \\
-&\text{qualifier\_EDB}(S, \text{propertyConstant}, V) \wedge{} \\
-&\text{unit}(V, U) {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, U)
-\end{split}\\
-\begin{split}
-\text{violation\_reference}&(S, H, \text{propertyConstant}, V) \leftarrow \\
-&\text{reference\_EDB}(S, H, \text{propertyConstant}, V) \wedge{} \\
-&\text{unit}(V, U) {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, U)
+\textsc{violation\_statement}&(S, E, \emph{constrainedProperty}, V) \leftarrow \\
+&\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, V) \wedge{} \\
+&\textsc{unit\_edb}(V, U)  {\textstyle\bigwedge}_{a \in A} \textsc{unequal}(a, U)
+\end{split}\\
+\begin{split}
+\textsc{violation\_qualifier}&(S, \emph{constrainedProperty}, V) \leftarrow \\
+&\textsc{qualifier\_edb}(S, \emph{constrainedProperty}, V) \wedge{} \\
+&\textsc{unit\_edb}(V, U) {\textstyle\bigwedge}_{a \in A} \textsc{unequal}(a, U)
+\end{split}\\
+\begin{split}
+\textsc{violation\_reference}&(S, H, \emph{constrainedProperty}, V) \leftarrow \\
+&\textsc{reference\_edb}(S, H, \emph{constrainedProperty}, V) \wedge{} \\
+&\textsc{unit\_edb}(V, U) {\textstyle\bigwedge}_{a \in A} \textsc{unequal}(a, U)
 \end{split}
@@
 \begin{split}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{} \\
-&\text{qualifier\_EDB}(S, Q, O) {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, Q)
+\textsc{violation\_statement}&(S, E, \emph{constrainedProperty}, V) \leftarrow \\
+&\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, V) \wedge{} \\
+&\textsc{qualifier\_edb}(S, Q, O) {\textstyle\bigwedge}_{a \in A} \textsc{unequal}(a, Q)
 \end{split}
@@
 \begin{split}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, V)
-\end{split}\\
-\begin{split}
-\text{violation\_qualifier}&(S, \text{propertyConstant}, V) \leftarrow \\
-&\text{qualifier\_EDB}(S, \text{propertyConstant}, V) {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, V)
-\end{split}\\
-\begin{split}
-\text{violation\_reference}&(S, \text{propertyConstant}, V) \leftarrow \\
-&\text{reference\_EDB}(S, H, \text{propertyConstant}, V) {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, V)
+\textsc{violation\_statement}&(S, E, \emph{constrainedProperty}, V) \leftarrow \\
+&\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, V) {\textstyle\bigwedge}_{a \in A} \textsc{unequal}(a, V)
+\end{split}\\
+\begin{split}
+\textsc{violation\_qualifier}&(S, \emph{constrainedProperty}, V) \leftarrow \\
+&\textsc{qualifier\_edb}(S, \emph{constrainedProperty}, V) {\textstyle\bigwedge}_{a \in A} \textsc{unequal}(a, V)
+\end{split}\\
+\begin{split}
+\textsc{violation\_reference}&(S, \emph{constrainedProperty}, V) \leftarrow \\
+&\textsc{reference\_edb}(S, H, \emph{constrainedProperty}, V) {\textstyle\bigwedge}_{a \in A} \textsc{unequal}(a, V)
 \end{split}
@@
 \begin{split}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{} \\
-&\text{qualifier\_EDB}(S, \text{qualifierProperty}, O) {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, O)
+\textsc{violation\_statement}&(S, E, \emph{constrainedProperty}, V) \leftarrow \\
+&\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, V) \wedge{} \\
+&\textsc{qualifier\_edb}(S, \emph{qualifierProperty}, O) {\textstyle\bigwedge}_{a \in A} \textsc{unequal}(a, O)
 \end{split}
@@
 However, non-existence can be derived utilising the order of statements.
-Three predicates are introduced: first\([2]\), last\([2]\), and next\([2]\), which denote the first and last statement of an item and the following statement to another using their statement IDs.\\
+Three predicates are introduced: \textsc{first}\([2]\), \textsc{last}\([2]\), and \textsc{next}\([2]\), which denote the first and last statement of an item and the following statement to another using their statement IDs.\\
 
@@
 To assign them a unique identity, statement ID, property, and value have to be used.
-For this the predicates first\_qualifier\([3]\) and last\_qualifier\([3]\) taking the statement ID, the property, and the value, as well as next\_qualifier\([6]\) with ID, property, and value for the current and next qualifier are introduced.\\
-
-Using this pre-processed data, the approach is to apply a require-predicate to each statement or qualifier.
-This is require\([2]\) for statements, connecting the statement ID with an identifier unique to the declared constraint and require\([4]\) taking the statement ID, property, and value of a qualifier and linking this with the unique identifier.\\
-
-For statements the pattern consists of two rules, both utilising the same conjunction \(\text{doesNotFulfil}\) whose atoms depend on the constraint type.
+For this the predicates \textsc{first\_qualifier}\([3]\) and \textsc{last\_qualifier}\([3]\) taking the statement ID, the property, and the value, as well as \textsc{next\_qualifier}\([6]\) with ID, property, and value for the current and next qualifier are introduced.\\
+
+Using this pre-processed data, the approach is to apply a not\_found-predicate to each statement or qualifier.
+This is \textsc{not\_found}\([2]\) for statements, connecting the statement ID with an identifier unique to the declared constraint and \textsc{not\_found}\([4]\) taking the statement ID, property, and value of a qualifier and linking this with the unique identifier.\\
+
+For statements the pattern consists of two rules, both utilising the same conjunction \(\emph{doesNotFulfil}\) whose atoms depend on the constraint type.
 This conjunction is used to test if the statement does not fulfil the requirement.
@@
 \begin{align}
-&\text{require}(S, \text{id}) \leftarrow \text{first}(S, I) \wedge \text{doesNotFulfil} \\
-&\text{require}(S, \text{id}) \leftarrow \text{next}(O, S) \wedge \text{require}(O, \text{id}) \wedge \text{doesNotFulfil} \\
-&\text{require}(S, P, V, \text{id}) \leftarrow \text{first}(S, P, V) \wedge \text{doesNotFulfil} \\
-&\text{require}(S, P, V, \text{id}) \leftarrow \text{next}(O, X, C, S, P, V) \wedge \text{require}(O, X, C, \text{id}) \wedge \text{doesNotFulfil}
+\textsc{not\_found}&(S, \emph{id}) \leftarrow \textsc{first}(S, I) \wedge \emph{doesNotFulfil} \\
+\textsc{not\_found}&(S, \emph{id}) \leftarrow \textsc{next}(O, S) \wedge \textsc{not\_found}(O, \emph{id}) \wedge \emph{doesNotFulfil} \\
+\textsc{not\_found}&(S, P, V, \emph{id}) \leftarrow \textsc{first\_qualifier}(S, P, V) \wedge \emph{doesNotFulfil} \\
+\begin{split}
+\textsc{not\_found}&(S, P, V, \emph{id}) \leftarrow \textsc{next\_qualifier}(O, X, C, S, P, V) \wedge{} \\
+&\textsc{not\_found}(O, X, C, \emph{id}) \wedge \emph{doesNotFulfil}
+\end{split}
 \end{align}
@@
 An item violates the constraint if it a) has a statement using the constrained property as predicate, b1) has no statements using the \emph{required property} as predicate, or (if \emph{allowed values} are specified) b2) has only statement using the \emph{required property} as predicate that do not use one of the \emph{allowed values} as value.
-If no \emph{allowed values} have been specified for the declared constraint the conjunction for the require pattern is one.
+If no \emph{allowed values} have been specified for the declared constraint the conjunction for the not-found-pattern is one.
 If a set of \emph{allowed values} has been specified, a second set of require-rules have to be introduced with conjunction two.
-Note that the require pattern using conjunction one is still required, since the preceding statements do not need to have the \emph{required property}.
+Note that the not-found-pattern using conjunction one is still required, since the preceding statements do not need to have the \emph{required property}.
 In both cases violations can be found with rule three
@@
 \begin{split}
-c = &\text{statement\_EDB}(Q, E, \text{propertyConstant}, X) \wedge{} \\
-&\text{statement\_EDB}(S, E, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
+c = &\textsc{statement\_edb}(Q, E, \emph{constrainedProperty}, X) \wedge{} \\
+&\textsc{statement\_edb}(S, E, P, V) \wedge \textsc{unequal}(\emph{requiredProperty}, P)
 \end{split} \\
 \begin{split}
-c = &\text{statement\_EDB}(Q, E, \text{propertyConstant}, X) \wedge{} \\
-&\text{statement\_EDB}(S, E, \text{requiredPropertyConstant}, V) \wedge {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, V)
-\end{split}
-\end{align}
-\begin{align}
-\begin{split}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{} \text{last}(O, I) \wedge{} \\
-&\text{require}(O, \text{propertyConstant}, \text{requiredPropertyConstant})
+c = &\textsc{statement\_edb}(Q, E, \emph{constrainedProperty}, X) \wedge{} \\
+&\textsc{statement\_edb}(S, E, \emph{requiredProperty}, V) \wedge {\textstyle\bigwedge}_{a \in A} \textsc{unequal}(a, V)
+\end{split}
+\end{align}
+\begin{align}
+\begin{split}
+\textsc{violation\_statement}&(S, E, \emph{constrainedProperty}, V) \leftarrow \\
+&\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, V) \wedge{} \textsc{last}(O, I) \wedge{} \\
+&\textsc{not\_found}(O, \emph{constrainedProperty}, \emph{requiredProperty})
 \end{split}
@@
 \begin{split}
-c = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
-&\text{statement\_EDB}(S, E, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
-\end{split}\\
-\begin{split}
-c = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
-&\text{statement\_EDB}(S, E, \text{requiredPropertyConstant}, V) \wedge {\textstyle\bigwedge}_{a \in A} \text{unequal}(a, V)
-\end{split}
-\end{align}
-
-\begin{align}
-\begin{split}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{} \text{last}(O, V) \wedge{} \\
-&\text{require}(O, \text{propertyConstant}, \text{requiredPropertyConstant})
+c = &\textsc{statement\_edb}(Q, R, \emph{constrainedProperty}, I) \wedge{} \\
+&\textsc{statement\_edb}(S, E, P, V) \wedge \textsc{unequal}(\emph{requiredProperty}, P)
+\end{split}\\
+\begin{split}
+c = &\textsc{statement\_edb}(Q, R, \emph{constrainedProperty}, I) \wedge{} \\
+&\textsc{statement\_edb}(S, E, \emph{requiredProperty}, V) \wedge {\textstyle\bigwedge}_{a \in A} \textsc{unequal}(a, V)
+\end{split}
+\end{align}
+
+\begin{align}
+\begin{split}
+\textsc{violation\_statement}&(S, E, \emph{constrainedProperty}, V) \leftarrow \\
+&\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, V) \wedge{} \textsc{last}(O, V) \wedge{} \\
+&\textsc{not\_found}(O, \emph{constrainedProperty}, \emph{requiredProperty})
 \end{split}
@@
 \begin{split}
-c = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
-&\text{statement\_EDB}(S, E, P, V) \wedge{} \\
-&\text{unequal}(\text{inversePropertyConstant}, P)
-\end{split}\\
-\begin{split}
-c = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
-&\text{statement\_EDB}(S, E, \text{inversePropertyConstant}, V) \wedge{} \\
-&\text{unequal}(R, V)
+c = &\textsc{statement\_edb}(Q, R, \emph{constrainedProperty}, I) \wedge{} \\
+&\textsc{statement\_edb}(S, E, P, V) \wedge{} \\
+&\textsc{unequal}(\emph{inverseProperty}, P)
+\end{split}\\
+\begin{split}
+c = &\textsc{statement\_edb}(Q, R, \emph{constrainedProperty}, I) \wedge{} \\
+&\textsc{statement\_edb}(S, E, \emph{inverseProperty}, V) \wedge{} \\
+&\textsc{unequal}(R, V)
 \end{split}
@@
 \begin{split}
-c = &\text{statement\_EDB}(Q, E, \text{propertyConstant}, X) \wedge{} \\
-&\text{statement\_EDB}(S, E, P, V) \wedge \text{unequal}(\text{propertyConstant}, P)
+c = &\textsc{statement\_edb}(Q, E, \emph{constrainedProperty}, X) \wedge{} \\
+&\textsc{statement\_edb}(S, E, P, V) \wedge \textsc{unequal}(\emph{constrainedProperty}, P)
 \end{split}
 \end{equation}
-The predicate require\_second\([2]\) is introduced with the same meaning as require except that it will be applied after a required property has been found.
+The predicate \textsc{require\_second}\([2]\) is introduced with the same meaning as require except that it will be applied after a required property has been found.
 A violation would be any last statement requiring a second constrained statement, analogous to the violation rule for \emph{item requires statement} (see Section \ref{subsec_item_requires_statement}).
@@
 \begin{split}
-\text{require\_second}&(S, \text{propertyConstant}, \text{propertyConstant}) \leftarrow \\
-&\text{next}(R, S) \wedge \text{statement\_EDB}(R, E, \text{propertyConstant}, C) \wedge{} \\
-&\text{next}(Q, R) \wedge \text{require}(Q, \text{propertyConstant}, \text{propertyConstant})  \wedge c
-\end{split}\\
-\begin{split}
-\text{require\_second}&(S, \text{propertyConstant}, \text{propertyConstant}) \leftarrow \text{next}(R, S) \wedge{} \\
-&\text{require\_second}(R, \text{propertyConstant}, \text{propertyConstant}) \wedge c
+\textsc{require\_second}&(S, \emph{constrainedProperty}, \emph{constrainedProperty}) \leftarrow \\
+&\textsc{next}(R, S) \wedge \textsc{statement\_edb}(R, E, \emph{constrainedProperty}, C) \wedge{} \\
+&\textsc{next}(Q, R) \wedge \textsc{not\_found}(Q, \emph{constrainedProperty}, \emph{constrainedProperty})  \wedge c
+\end{split}\\
+\begin{split}
+\textsc{require\_second}&(S, \emph{constrainedProperty}, \emph{constrainedProperty}) \leftarrow \textsc{next}(R, S) \wedge{} \\
+&\textsc{require\_second}(R, \emph{constrainedProperty}, \emph{constrainedProperty}) \wedge c
 \end{split}
@@
 \begin{split}\label{eq_no_separators_triple}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{} \\
-&\text{statement\_EDB}(O, E, \text{propertyConstant}, X) \wedge{} \\
-&\text{unequal}(S, O)
-\end{split}\\
-\begin{split}
-\text{violation\_qualifier}&(S, \text{propertyConstant}, V) \leftarrow \\
-&\text{qualifier\_EDB}(S, \text{propertyConstant}, V) \wedge{} \\
-&\text{qualifier\_EDB}(S, \text{propertyConstant}, X) \wedge{} \\
-&\text{unequal}(V, X)
-\end{split}\\
-\begin{split}
-\text{violation\_reference}&(S, H, \text{propertyConstant}, V) \leftarrow \\
-&\text{reference\_EDB}(S, H, \text{propertyConstant}, V) \wedge{} \\
-&\text{reference\_EDB}(S, J, \text{propertyConstant}, X) \wedge{} \\
-&\text{unequal}(V, X)
+\textsc{violation\_statement}&(S, E, \emph{constrainedProperty}, V) \leftarrow \\
+&\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, V) \wedge{} \\
+&\textsc{statement\_edb}(O, E, \emph{constrainedProperty}, X) \wedge{} \\
+&\textsc{unequal}(S, O)
+\end{split}\\
+\begin{split}
+\textsc{violation\_qualifier}&(S, \emph{constrainedProperty}, V) \leftarrow \\
+&\textsc{qualifier\_edb}(S, \emph{constrainedProperty}, V) \wedge{} \\
+&\textsc{qualifier\_edb}(S, \emph{constrainedProperty}, X) \wedge{} \\
+&\textsc{unequal}(V, X)
+\end{split}\\
+\begin{split}
+\textsc{violation\_reference}&(S, H, \emph{constrainedProperty}, V) \leftarrow \\
+&\textsc{reference\_edb}(S, H, \emph{constrainedProperty}, V) \wedge{} \\
+&\textsc{reference\_edb}(S, J, \emph{constrainedProperty}, X) \wedge{} \\
+&\textsc{unequal}(V, X)
 \end{split}
@@
 The require qualifier pattern needs to be applied to the qualifiers of constrained statements.
-The conjunction \(\text{doesNotFulfil}\) is created for every \emph{separator} and the pattern applied.
+The conjunction \(\emph{doesNotFulfil}\) is created for every \emph{separator} and the pattern applied.
 \begin{equation}
 \begin{split}
-c = &\text{statement\_EDB}(S, E, \text{propertyConstant}, C) \wedge{} \\
-&\text{qualifier\_EDB}(S, P, V) \wedge{} \\
-&\text{unequal}(P, \text{separatorConstant})
+c = &\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, C) \wedge{} \\
+&\textsc{qualifier\_edb}(S, P, V) \wedge{} \\
+&\textsc{unequal}(P, \emph{separator})
 \end{split}
@@
 The following rule is then created for every possible split of the \emph{separators} into two such sets.
-Note that \(\text{predicate}(\{X\})\) denotes \(\bigwedge_{x \in X} \text{predicate}(x)\) as introduced in Section \ref{subsec_3_allowed_units}.
-\begin{align}
-\begin{split}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{} \\
-&\text{statement\_EDB}(O, E, \text{propertyConstant}, X) \wedge{} \\
-&\text{unequal}(S, O) \wedge{} \\
-&\text{same\_or\_non\_existent}(S, O, \{Q\})
-\end{split}\\
-\begin{split}
-\text{same\_or\_non\_existent}&(S, O, Q) \leftarrow \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, X) \wedge{} \\
-&\text{statement\_EDB}(O, E, \text{propertyConstant}, Y) \wedge{} \\
-&\text{qualifier\_EDB}(S, Q, V) \wedge{} \\
-&\text{qualifier\_EDB}(O, Q, V)
-\end{split}\\
-\begin{split}
-\text{same\_or\_non\_existent}&(S, O, Q) \leftarrow{} \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, X) \wedge{} \\
-&\text{statement\_EDB}(O, E, \text{propertyConstant}, Y) \wedge{} \\
-&\text{last\_qualifier}(S, P, V) \wedge{} \\
-&\text{require}(S, P, V, \text{propertyConstant}, Q) \wedge{} \\
-&\text{last\_qualifier}(O, R, C) \wedge{} \\
-&\text{require}(O, R, C, \text{propertyConstant}, Q)
+\begin{align}
+\begin{split}
+\textsc{violation\_statement}&(S, E, \emph{constrainedProperty}, V) \leftarrow \\
+&\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, V) \wedge{} \\
+&\textsc{statement\_edb}(O, E, \emph{constrainedProperty}, X) \wedge{} \\
+&\textsc{unequal}(S, O) \wedge{} \\
+&\textsc{same\_or\_non\_existent}(S, O, \{Q\})
+\end{split}\\
+\begin{split}
+\textsc{same\_or\_non\_existent}&(S, O, Q) \leftarrow \\
+&\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, X) \wedge{} \\
+&\textsc{statement\_edb}(O, E, \emph{constrainedProperty}, Y) \wedge{} \\
+&\textsc{qualifier\_edb}(S, Q, V) \wedge{} \\
+&\textsc{qualifier\_edb}(O, Q, V)
+\end{split}\\
+\begin{split}
+\textsc{same\_or\_non\_existent}&(S, O, Q) \leftarrow{} \\
+&\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, X) \wedge{} \\
+&\textsc{statement\_edb}(O, E, \emph{constrainedProperty}, Y) \wedge{} \\
+&\textsc{last\_qualifier}(S, P, V) \wedge{} \\
+&\textsc{not\_found}(S, P, V, \emph{constrainedProperty}, Q) \wedge{} \\
+&\textsc{last\_qualifier}(O, R, C) \wedge{} \\
+&\textsc{not\_found}(O, R, C, \emph{constrainedProperty}, Q)
 \end{split}
@@
 The qualifier order needs to be provided and inequality must be established between all qualifier properties.
-The conjunction for the require-pattern is rule one.
-With the usual require-pattern for qualifiers and the does\_not\_have rule as introduced in Section \ref{subsubsec_single_value_with_separators} added for brevity, the violations would be found with rule two.
-\begin{align}
-\begin{split}
-c = &\text{statement\_EDB}(S, E, \text{propertyConstant}, C) \wedge{} \\
-&\text{qualifier\_EDB}(S, P, V) \wedge \text{unequal}(P, \text{requiredPropertyConstant})
-\end{split}\\
-\begin{split}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge \\
-&\text{does\_not\_have}(S, \text{requiredPropertyConstant})
+The conjunction for the not-found-pattern is rule one.
+With the usual not-found-pattern for qualifiers the violations would be found with rule two.
+\begin{align}
+\begin{split}
+c = &\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, C) \wedge{} \\
+&\textsc{qualifier\_edb}(S, P, V) \wedge \textsc{unequal}(P, \emph{requiredProperty})
+\end{split}
+\end{align}
+\begin{align}
+\begin{split}
+\textsc{violation\_statement}&(S, E, \emph{constrainedProperty}, V) \leftarrow \\
+&\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, V) \wedge \\
+&\textsc{last\_qualifier}(S, P, V) \wedge \textsc{not\_found}(S, P, V, \emph{requiredProperty})
 \end{split}
@@
 \begin{split}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{} \\
-&\text{rank\_EDB}(S, \text{preferredRankConstant}) \wedge{} \\
-&\text{statement\_EDB}(O, E, \text{propertyConstant}, X) \wedge{} \\
-&\text{rank\_EDB}(O, \text{preferredRankConstant}) \wedge{} \\
-&\text{unequal}(S, O)
+\textsc{violation\_statement}&(S, E, \emph{constrainedProperty}, V) \leftarrow \\
+&\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, V) \wedge{} \\
+&\textsc{rank\_edb}(S, \emph{preferredRank}) \wedge{} \\
+&\textsc{statement\_edb}(O, E, \emph{constrainedProperty}, X) \wedge{} \\
+&\textsc{rank\_edb}(O, \emph{preferredRank}) \wedge{} \\
+&\textsc{unequal}(S, O)
 \end{split}
@@
 \begin{split}
-\text{violation\_statement}&(S, E, \text{propertyConstant}, V) \leftarrow \\
-&\text{statement\_EDB}(S, E, \text{propertyConstant}, V) \wedge{} \\
-&\text{rank\_EDB}(S, \text{preferredRankConstant}) \wedge{} \\
-&\text{statement\_EDB}(O, E, \text{propertyConstant}, W) \wedge{} \\
-&\text{rank\_EDB}(O, \text{preferredRankConstant}) \wedge{} \\
-&\text{unequal}(S, O) \wedge{} \\
-&\text{has\_same}(S, O, \{X\}) \wedge{} \\
-&\text{does\_not\_have}(S, \{Y\}) \wedge{} \\
-&\text{does\_not\_have}(O, \{Y\})
+\textsc{violation\_statement}&(S, E, \emph{constrainedProperty}, V) \leftarrow \\
+&\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, V) \wedge{} \\
+&\textsc{rank\_edb}(S, \emph{preferredRank}) \wedge{} \\
+&\textsc{statement\_edb}(O, E, \emph{constrainedProperty}, W) \wedge{} \\
+&\textsc{rank\_edb}(O, \emph{preferredRank}) \wedge{} \\
+&\textsc{unequal}(S, O) \wedge{} \\
+&\textsc{has\_same}(S, O, \{X\}) \wedge{} \\
+&\textsc{does\_not\_have}(S, \{Y\}) \wedge{} \\
+&\textsc{does\_not\_have}(O, \{Y\})
 \end{split}
@@
 \begin{example}
-The predicate \emph{statement\_EDB} lists the \emph{statement ID}, \emph{entity ID}, \emph{property ID} and \emph{value}.
+The predicate \textsc{statement\_edb} lists the \emph{statement ID}, \emph{entity ID}, \emph{property ID} and \emph{value}.
 To optimize this for VLog, a new order must be created.

### 2018-09-20 thesis/paper/paper.tex
Corrected does-not-fulfil

@@
 
-For statements the pattern consists of two rules, both utilising the same conjunction \(\emph{doesNotFulfil}\) whose atoms depend on the constraint type.
+For statements the pattern consists of two rules, both utilising the same conjunction \emph{does-not-fulfil} whose atoms depend on the constraint type.
 This conjunction is used to test if the statement does not fulfil the requirement.
@@
 \begin{align}
-\textsc{not\_found}&(S, \emph{id}) \leftarrow \textsc{first}(S, I) \wedge \emph{doesNotFulfil} \\
-\textsc{not\_found}&(S, \emph{id}) \leftarrow \textsc{next}(O, S) \wedge \textsc{not\_found}(O, \emph{id}) \wedge \emph{doesNotFulfil} \\
-\textsc{not\_found}&(S, P, V, \emph{id}) \leftarrow \textsc{first\_qualifier}(S, P, V) \wedge \emph{doesNotFulfil} \\
+\textsc{not\_found}&(S, \emph{id}) \leftarrow \textsc{first}(S, I) \wedge \emph{does-not-fulfil} \\
+\textsc{not\_found}&(S, \emph{id}) \leftarrow \textsc{next}(O, S) \wedge \textsc{not\_found}(O, \emph{id}) \wedge \emph{does-not-fulfil} \\
+\textsc{not\_found}&(S, P, V, \emph{id}) \leftarrow \textsc{first\_qualifier}(S, P, V) \wedge \emph{does-not-fulfil} \\
 \begin{split}
 \textsc{not\_found}&(S, P, V, \emph{id}) \leftarrow \textsc{next\_qualifier}(O, X, C, S, P, V) \wedge{} \\
-&\textsc{not\_found}(O, X, C, \emph{id}) \wedge \emph{doesNotFulfil}
+&\textsc{not\_found}(O, X, C, \emph{id}) \wedge \emph{does-not-fulfil}
 \end{split}
@@
 \begin{split}
-c = &\textsc{statement\_edb}(Q, E, \emph{constrainedProperty}, X) \wedge{} \\
+\emph{does-not-fulfil} = &\textsc{statement\_edb}(Q, E, \emph{constrainedProperty}, X) \wedge{} \\
 &\textsc{statement\_edb}(S, E, P, V) \wedge \textsc{unequal}(\emph{requiredProperty}, P)
@@
 \begin{split}
-c = &\textsc{statement\_edb}(Q, E, \emph{constrainedProperty}, X) \wedge{} \\
+\emph{does-not-fulfil} = &\textsc{statement\_edb}(Q, E, \emph{constrainedProperty}, X) \wedge{} \\
 &\textsc{statement\_edb}(S, E, \emph{requiredProperty}, V) \wedge {\textstyle\bigwedge}_{a \in A} \textsc{unequal}(a, V)
@@
 \textsc{violation\_statement}&(S, E, \emph{constrainedProperty}, V) \leftarrow \\
-&\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, V) \wedge{} \textsc{last}(O, I) \wedge{} \\
+&\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, V) \wedge{} \\
+&\textsc{last}(O, I) \wedge{} \\
 &\textsc{not\_found}(O, \emph{constrainedProperty}, \emph{requiredProperty})
@@
 \begin{split}
-c = &\textsc{statement\_edb}(Q, R, \emph{constrainedProperty}, I) \wedge{} \\
+\emph{does-not-fulfil} = &\textsc{statement\_edb}(Q, R, \emph{constrainedProperty}, I) \wedge{} \\
 &\textsc{statement\_edb}(S, E, P, V) \wedge \textsc{unequal}(\emph{requiredProperty}, P)
@@
 \begin{split}
-c = &\textsc{statement\_edb}(Q, R, \emph{constrainedProperty}, I) \wedge{} \\
+\emph{does-not-fulfil} = &\textsc{statement\_edb}(Q, R, \emph{constrainedProperty}, I) \wedge{} \\
 &\textsc{statement\_edb}(S, E, \emph{requiredProperty}, V) \wedge {\textstyle\bigwedge}_{a \in A} \textsc{unequal}(a, V)
@@
 \end{align}
-
 \begin{align}
@@
 \textsc{violation\_statement}&(S, E, \emph{constrainedProperty}, V) \leftarrow \\
-&\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, V) \wedge{} \textsc{last}(O, V) \wedge{} \\
+&\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, V) \wedge{} \\
+&\textsc{last}(O, V) \wedge{} \\
 &\textsc{not\_found}(O, \emph{constrainedProperty}, \emph{requiredProperty})
@@
 The \emph{inverse} constraint can be modelled in almost the same way as the \emph{value requires statement} constraint.
-The only change is that instead of requiring inequality to all \emph{allowed values} in conjunction \(c_2\) the statement requires this constraint if its value is not the subject \(R\) of the original constrained statement, meaning \(t = R\).
+The only change is that instead of requiring inequality to all \emph{allowed values} in conjunction \(c_2\) the statement requires this constraint if its value is not the subject \(R\) of the original constrained statement.
 Inequalities have to be established between all properties and the \emph{inverse} property as well as all subjects and values of statements.
@@
 \begin{split}
-c = &\textsc{statement\_edb}(Q, R, \emph{constrainedProperty}, I) \wedge{} \\
+\emph{does-not-fulfil} = &\textsc{statement\_edb}(Q, R, \emph{constrainedProperty}, I) \wedge{} \\
 &\textsc{statement\_edb}(S, E, P, V) \wedge{} \\
@@
 \begin{split}
-c = &\textsc{statement\_edb}(Q, R, \emph{constrainedProperty}, I) \wedge{} \\
+\emph{does-not-fulfil} = &\textsc{statement\_edb}(Q, R, \emph{constrainedProperty}, I) \wedge{} \\
 &\textsc{statement\_edb}(S, E, \emph{inverseProperty}, V) \wedge{} \\
@@
 \begin{split}
-c = &\textsc{statement\_edb}(Q, E, \emph{constrainedProperty}, X) \wedge{} \\
+\emph{does-not-fulfil} = &\textsc{statement\_edb}(Q, E, \emph{constrainedProperty}, X) \wedge{} \\
 &\textsc{statement\_edb}(S, E, P, V) \wedge \textsc{unequal}(\emph{constrainedProperty}, P)
@@
 &\textsc{next}(R, S) \wedge \textsc{statement\_edb}(R, E, \emph{constrainedProperty}, C) \wedge{} \\
-&\textsc{next}(Q, R) \wedge \textsc{not\_found}(Q, \emph{constrainedProperty}, \emph{constrainedProperty})  \wedge c
+&\textsc{next}(Q, R) \wedge{} \\
+&\textsc{not\_found}(Q, \emph{constrainedProperty}, \emph{constrainedProperty})  \wedge{} \\
+&\emph{does-not-fulfil}
 \end{split}\\
@@
 \textsc{require\_second}&(S, \emph{constrainedProperty}, \emph{constrainedProperty}) \leftarrow \textsc{next}(R, S) \wedge{} \\
-&\textsc{require\_second}(R, \emph{constrainedProperty}, \emph{constrainedProperty}) \wedge c
+&\textsc{require\_second}(R, \emph{constrainedProperty}, \emph{constrainedProperty}) \wedge{} \\
+&\emph{does-not-fulfil}
 \end{split}
@@
 \begin{split}
-c = &\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, C) \wedge{} \\
+\emph{does-not-fulfil} = &\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, C) \wedge{} \\
 &\textsc{qualifier\_edb}(S, P, V) \wedge{} \\
@@
 With qualifier non-existence established, the violation can be found.
-Two statements form a violation if they are distinct as per equation \ref{eq_no_separators_triple} and additionally fulfil the condition that all values of \emph{separator} qualifiers are equal or no such \emph{separator} qualifier exist.
-To determine this, a number of rules have to be introduced which are based on two sets the \emph{separators} are split into: \(X\) contains the \emph{separators} that exist as qualifiers and \(Y\) the \emph{separators} that do not.
-The following rule is then created for every possible split of the \emph{separators} into two such sets.
+Two statements form a violation if they are distinct as per equation \ref{eq_no_separators_triple} and additionally fulfil the condition that all values of \emph{separator} qualifiers are equal or no such \emph{separator} qualifier exist on both statements.
+To determine this, a number of rules have to be introduced which are based on the \emph{separators}-set \(Q\).
+Rule one is created and uses the auxiliary rules two and three.
 \begin{align}
@@
 &\textsc{unequal}(S, O) \wedge{} \\
-&\textsc{same\_or\_non\_existent}(S, O, \{Q\})
+&{\textstyle\bigwedge}_{q \in Q} \textsc{same\_or\_non\_existent}(S, O, q)
 \end{split}\\
@@
 \begin{split}
-c = &\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, C) \wedge{} \\
+\emph{does-not-fulfil} = &\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, C) \wedge{} \\
 &\textsc{qualifier\_edb}(S, P, V) \wedge \textsc{unequal}(P, \emph{requiredProperty})
@@
 &\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, V) \wedge \\
-&\textsc{last\_qualifier}(S, P, V) \wedge \textsc{not\_found}(S, P, V, \emph{requiredProperty})
+&\textsc{last\_qualifier}(S, P, V) \wedge{} \\
+&\textsc{not\_found}(S, P, V, \emph{requiredProperty})
 \end{split}
@@
 &\textsc{rank\_edb}(S, \emph{preferredRank}) \wedge{} \\
-&\textsc{statement\_edb}(O, E, \emph{constrainedProperty}, W) \wedge{} \\
+&\textsc{statement\_edb}(O, E, \emph{constrainedProperty}, X) \wedge{} \\
 &\textsc{rank\_edb}(O, \emph{preferredRank}) \wedge{} \\
 &\textsc{unequal}(S, O) \wedge{} \\
-&\textsc{has\_same}(S, O, \{X\}) \wedge{} \\
-&\textsc{does\_not\_have}(S, \{Y\}) \wedge{} \\
-&\textsc{does\_not\_have}(O, \{Y\})
+&{\textstyle\bigwedge}_{q \in Q} \textsc{same\_or\_non\_existent}(S, O, q)
 \end{split}

### 2018-09-20 thesis/paper/paper.tex

@@
 However, non-existence can be derived utilising the order of statements.
-Three predicates are introduced: \textsc{first}\([2]\), \textsc{last}\([2]\), and \textsc{next}\([2]\), which denote the first and last statement of an item and the following statement to another using their statement IDs.\\
+Three predicates are introduced: \textsc{first}\([1]\), \textsc{last}\([2]\), and \textsc{next}\([2]\), which denote the first and last statement of an item and the following statement to another using their statement IDs.\\
 
@@
 This conjunction is used to test if the statement does not fulfil the requirement.
+It needs to bind the variable \(S\), denoting the current statement ID, but not bind the variable \(O\), denoting the preceding statement ID.
+Care must be taken that the rules presented below also utilise the variables \(C, P, V,\) and \(X\) as the value of the preceding qualifier, the property of the current qualifier, the value of the current qualifier, and the property of the preceding qualifier, respectively.
 Then rule one states that the statement requires the condition and rule two propagates this down the statement chain.
@@
 \begin{align}
-\textsc{not\_found}&(S, \emph{id}) \leftarrow \textsc{first}(S, I) \wedge \emph{does-not-fulfil} \\
+\textsc{not\_found}&(S, \emph{id}) \leftarrow \textsc{first}(S) \wedge \emph{does-not-fulfil} \\
 \textsc{not\_found}&(S, \emph{id}) \leftarrow \textsc{next}(O, S) \wedge \textsc{not\_found}(O, \emph{id}) \wedge \emph{does-not-fulfil} \\
@@
 \begin{split}
-\textsc{not\_found}&(S, P, V, \emph{id}) \leftarrow \textsc{next\_qualifier}(O, X, C, S, P, V) \wedge{} \\
-&\textsc{not\_found}(O, X, C, \emph{id}) \wedge \emph{does-not-fulfil}
-\end{split}
-\end{align}
-
-If the last statement of an item or the last qualifier of a statement has the require-predicate the violation occurred.
-In the following constraints only the conjunction \(c\) and adjustments to this pattern will be described.
-The require-propagation is regarded as given.
+\textsc{not\_found}&(S, P, V, \emph{id}) \leftarrow \textsc{next\_qualifier}(S, X, C, S, P, V) \wedge{} \\
+&\textsc{not\_found}(S, X, C, \emph{id}) \wedge \emph{does-not-fulfil}
+\end{split}
+\end{align}
+
+If the last statement of an item or the last qualifier of a statement has the \textsc{not\_found}-predicate the violation occurred.
+In the following constraints only the conjunction \emph{does-not-fulfil} and adjustments to this pattern will be described.
+The \textsc{not\_found}-propagation is regarded as given.
+If multiple \emph{does-not-fulfil} conjunction are specified for one constraint type the pattern is applied multiple times.
 

### 2018-09-20 thesis/paper/paper.tex
Changed arity declaration

@@
 It is a public, open database system that, at its core, stores statements about specific items.
-The basic unit of information is a triple subject-predicate-object stating an item (subject) has a property (predicate) which is either another item or a value of some kind (object).
+The basic unit of information is a triple of the form subject-predicate-object stating an item (subject) has a property (predicate) which is either another item or a value of some kind (object).
 Triples store the relations of entities with each other or with specific data.
@@
 They have numeric identifiers that are prefixed with Q and P and will be referred to as Q- and P-IDs, respectively.
-In the future there will be a third entity type, lexeme.
-This is already incorporated into some constraints, but as of September 2018 they are not deployed.
-A value is one of several data types shown in Table \ref{tab_datatypes}, although it is also possible to use \emph{none} or \emph{some} as values for statements that have no value (akin to negation) or have a currently unknown value.
+A third entity type, lexeme, has been deployed since May 2018.
+This is already incorporated into some constraint type definitions, but as of September 2018 there are no constraint declarations using this.
+A value is one of several data types shown in Table \ref{tab_datatypes}, although it is also possible to use \emph{none} or \emph{some} as values for statements that have no value (a kind of negation) or have a currently unknown value.
+Values consist of components, separating different aspects of a value into separate entries.
 Every property determines the data type it accepts as an object.
@@
 \caption{Wikidata data types}\label{tab_datatypes}
-\begin{tabularx}{\textwidth}{llp{8.5cm}}
-Data type & Fields & Description \\
+\begin{tabularx}{\textwidth}{llp{7cm}}
+Data type & Components & Description \\
 \hline
 commons media & file (string) & name of a file in the "File" namespace on Wikimedia Commons \\
-globecoordinate & \makecell{latitude (float) \\ longitude (float) \\ precision (float) \\ globe (URI)} & coordinates on a given celestial body or geographic standard \\
+globecoordinate & \makecell{latitude (float) \\ longitude (float) \\ precision (float) \\ globe (URI)} & coordinates on a given celestial body or in a given geographic standard (specified via the globe URI) \\
 item & item (Q-ID) & the Q-ID of an item on Wikidata \\
 property & property (P-ID) & the P-ID of an item on Wikidata \\
-string & string (string) & simple string literal that does not need to be translated into different languages \\
+string & \makecell{string (string) \\ language (language code)} & simple string literal that does not need to be translated into different languages \\
 monolingual text & string (string) & simple string literal that is written in one specific language and not to be translated \\
 quantity & \makecell {amount (decimal) \\ upperBound (decimal) \\ lowerBound (decimal) \\ unit (URI or 1)} & quantity of a specified unit (or none) with uncertainty \\
-time & \makecell {time (string) \\ calendarmodel (URI) \\ precision (integer)} & time as a string according to calendarmodel (julian or gregorian), with precision mapping to orders of magnitude (e.g..
+time & \makecell {time (string) \\ calendarmodel (URI) \\ precision (integer)} & time as a string according to calendarmodel (Julian or Gregorian), with precision mapping to orders of magnitude (e.g..
 10 years, months) \\
@@
 mathematical expression & formula (string) & string that is formatted to display as a formula \\
-external identifier & id (string) & string representing an identifier of an external system \\
-geographic shape & map data (string) & name of a file in the "Data" namespace on Wikimedia Commons \\
-tabular data & tabular data (string) & name of a file in the "Data" namespace on Wikimedia Commons \\
+external identifier & ID (string) & string representing an identifier of an external system \\
+geographic shape & map data (string) & name of a file in the ``Data'' namespace on Wikimedia Commons \\
+tabular data & tabular data (string) & name of a file in the ``Data'' namespace on Wikimedia Commons \\
 \end{tabularx}
@@
 While this is similar to RDF and Wikidata has in fact been exported as RDF \cite{EGKMV2014}, the data structure is more complex than simple triples.
-It allows to store additional information about statements.
+It allows storing additional information about statements.
 This is done via a statement ID, which is assigned to every triple storing information about an item.
@@
 For example, the \emph{preferred} rank would be given to the current \emph{head of state} (P35) of a country, while all statements denoting former \emph{heads of state} have the rank \emph{normal}.
-\emph{Deprecated} is used on statements that where once considered correct but are not anymore, e.g., \emph{Chelsea Manning} (Q298423) having the \emph{sex or gender} (P21) \emph{male} (Q6581097).
-
-Every statement can have added qualifiers, which consist of the corresponding statement id, a predicate, and an object.
+\emph{Deprecated} is used on statements that are not correct, e.g., \emph{Chelsea Manning} (Q298423) having the \emph{sex or gender} (P21) \emph{male} (Q6581097).
+
+Every statement can have added qualifiers, which consist of a predicate and an object.
 Qualifiers provide contextual information for statements.
@@
 determination method for measurements or the elevation above sea level for the highest point.
+
 Additionally, references can be added to statements.
-
 References support the claim by citing the source and additional information, e.g.
-when it was retrieved or when the source was published as property-value pairs.
+when it was retrieved or when the source was published, as property-value pairs.
 Each reference consists of multiple such pairs, and statements can have multiple references.
@@
 It is fully declarative and uses rules to derive additional knowledge from given facts.
-The most basic domain element in datalog is a \emph{constant}, which will be denoted as \emph{someConstant}.
+The most basic domain element in datalog is a \emph{constant}.
 In the context of Wikidata, any item or property, e.g.
 \emph{Earth} (Q2) or \emph{head of government} (P6), could be represented as a constant.
-Relations in datalog are expressed using \emph{predicates} with a specified arity \(n\), written as \textsc{predicate}\([n]\).
-To stay close to Wikidata in our examples, let us take the predicate \textsc{triple}\([3]\) as an example.
+Relations in datalog are expressed using \emph{predicates} with a specified arity \(n\).
+To stay close to Wikidata in our examples, let us take the predicate \textsc{triple}with arity 3 as an example.
 This would be used to express the \emph{item}-\emph{property}-\emph{value} relation with each position representing one part of the triple.\\
@@
 Constants and variables are collectively known as \emph{terms} and are, together with predicates, used to form \emph{atoms}.
-An atom has the form \textsc{predicate}\((t_1, \ldots , t_n)\) with \(n\) as the arity of the predicate and \(t_1,\ldots,t_n\) terms.\\
-Atoms are the building blocks of \emph{rules}, which are used to derive information in datalog. A rule is a premise that implies a conclusion, and both are conjunctions of atoms.
+An atom has the form \textsc{predicate}\((t_1, \ldots , t_n)\) with \(t_1,\ldots,t_n\) being terms.\\
+Atoms are the building blocks of \emph{rules}, which are used to derive information in datalog. A rule specifies a premise that implies a conclusion, and both are conjunctions of atoms.
 Let \(P_n\) denote  premise atoms and \(C_n\) the conclusion atoms.
@@
 The species \emph{blueWhale} (Q42196) has the \emph{parentTaxon} (P171) \emph{balaenoptera} (Q133320), which in turn has the parent taxon \emph{rorqual} (Q232829).
-The predicates used are \textsc{triple\_edb}\([3]\) as the input and \textsc{is\_in\_taxon\_idb}\([2]\), which will be derived.
+The predicates used are \textsc{triple} with arity 3 as the input and \textsc{is\_in\_taxon}with arity 2, which will be derived.
 The basic facts are:
-\[\textsc{triple\_edb}(\emph{blueWhale}, \emph{parentTaxon}, \emph{balaenoptera})\]
-\[\textsc{triple\_edb}(\emph{balaenoptera}, \emph{parenTaxon}, \emph{rorqual})\]
+\[\textsc{triple}(\emph{blueWhale}, \emph{parentTaxon}, \emph{balaenoptera})\]
+\[\textsc{triple}(\emph{balaenoptera}, \emph{parenTaxon}, \emph{rorqual})\]
 From this we can draw conclusions. The first step is to represent the already existing facts as idb:
-\[\textsc{is\_in\_taxon\_idb}(X, Y) \leftarrow \textsc{triple\_edb}(X, \emph{parentTaxon}, Y)\]
+\[\textsc{is\_in\_taxon\_idb}(X, Y) \leftarrow \textsc{triple}(X, \emph{parentTaxon}, Y)\]
 Finally the transitive closure of this relation can be derived with
@@
 that the value of a statement with this property should have certain statements.
-In Table \ref{tab_property_constraints} you can find the existing property constraints as of September 2018.
-Afterwards, the constraints will be explained in detail and their modelling given.
+Table \ref{tab_property_constraints} contains the existing property constraints as of September 2018.
+The constraints will be explained in detail and their modelling given.
 For constraint that cannot be modelled, reasoning for this assessment is included.
 
-To apply a constraint to a property, a statement <property> <P2302> <constraint> is added to the properties item page, where <property> is the property that should be constrained, <P2302> (property constraint) denotes that the property has a constraint and <constraint> is the specific constraint.
-Note that a constrained statement means a statement whose predicate is a property that has a specific constraint.
+To apply a constraint to a property, a statement <property> <P2302> <constraint> is added to the property's page, where <property> is the property that should be constrained, <P2302> (property constraint) denotes that the property has a constraint and <constraint> is the specific constraint.
+A constrained statement is a statement whose predicate is a property that has a constraint declaration.
 A simple example for this would be \emph{distinct values} (Q21502410), stating that values of the property must be unique across all constrained statements.
-For this case no additional information regarding the constraint is necessary and violation could be found by surveying all constrained statements and comparing their values.\\
+No additional information regarding the constraint is necessary in this case, and violation could be found by surveying all constrained statements and comparing their values.\\
 Constraints can be further specified by using qualifiers on the constraint statement.
@@
 More complex dependencies are also possible.
-A constraint can depend on all constrained statements of a specific item (\emph{single value} constraint [Q19474404]), on all statements of the item specified as the value of a constrained statement (\emph{value requires statement} constraint [Q21510864]), or on all statements using the subclass of-predicate (\emph{type} constraint [Q21503250]).
+A constraint type can depend on all constrained statements of a specific item (\emph{single value} constraint [Q19474404]), on all statements of the item specified as the value of a constrained statement (\emph{value requires statement} constraint [Q21510864]), or on all statements using the subclass of-predicate (\emph{type} constraint [Q21503250]).
 
@@
 These required features are conditions that cannot natively be expressed in datalog and would require some kind of pre-processing of the dataset.
-Four constraints do not need any pre-processing except for loading the data into datalog as ground rules.
+Four constraints do not need any pre-processing except for loading the data into datalog as facts.
 Six can be solved if constant inequality is defined.
@@
 The feature mostly needed is a general form of negation that would allow to test for the non-existence of a derived fact.
-Other features needed for these constraints are regular expressions, which could theoretically be translated, but would require large amounts of pre-processing, value comparison, and finally web requests to test if a link is valid.
+Other features needed for these constraints are value comparison, web requests to test if a link is valid, and regular expressions, which could theoretically be translated, but would require large amounts of pre-processing.
 
@@
 \hline
-\textsc{statement\_edb} & \makecell{statementID \\ entityID \\ propertyID \\ value (entityID or string)} & represents basic statements \\
-\hline
-\textsc{qualifier\_edb} & \makecell{statementID \\ propertyID \\ value} & represents qualifiers on statements \\
-\hline
-\textsc{reference\_edb} & \makecell{statementID \\ reference hash \\ propertyID \\ value} & represents references on statements; reference hash groups the property-value pairs belonging to one reference \\
-\hline
-\textsc{item\_edb} & entityID & marks an entity as an item \\
-\textsc{property\_edb} & entityID & marks an entity as a property \\
-\hline
-\textsc{unit\_edb} & \makecell{value \\ unitID} & denotes which unit a value has \\
-\hline
-\textsc{rank\_edb} & \makecell{statementID \\ rank} & denotes the rank a statement has
+\textsc{statement\_edb} & \makecell{statement ID \\ entity ID \\ property ID \\ value (entity ID or string)} & represents basic statements \\
+\hline
+\textsc{qualifier\_edb} & \makecell{statement ID \\ property ID \\ value} & represents qualifiers on statements \\
+\hline
+\textsc{reference\_edb} & \makecell{statement ID \\ reference hash \\ property ID \\ value} & represents references on statements; reference hash groups the property-value pairs belonging to one reference \\
+\hline
+\textsc{item\_edb} & entity ID & marks an entity as an item \\
+\textsc{property\_edb} & entity ID & marks an entity as a property \\
+\hline
+\textsc{unit\_edb} & \makecell{value \\ unit ID} & denotes which unit a value has \\
+\hline
+\textsc{rank\_edb} & \makecell{statement ID \\ rank} & denotes the rank a statement has
 \end{tabular}
@@
 In every approach it will be necessary to load some basic inequalities into the database.
-To this end the predicate unequal\_edb\([2]\) is introduced.
-Since the goal is to keep the number of applications of this predicate small, an \emph{idb} predicate unequal\([2]\) has to be introduced.
+To this end the predicate unequal\_edbwith arity 2 is introduced.
+Since the goal is to keep the number of applications of this predicate small, an \emph{idb} predicate unequalwith arity 2 has to be introduced.
 This leads to the rule
@@
 This means that only some inequalities -- the ones that actually appear in the dataset -- have to be evaluated.
-The rules can then be limited by introducing a new predicate req\_inequality\([2]\) which will only be derived if a condition occurs in which the inequality must be evaluated.\\
+The rules can then be limited by introducing a new predicate req\_inequalitywith arity 2 which will only be derived if a condition occurs in which the inequality must be evaluated.\\
 
@@
 However, non-existence can be derived utilising the order of statements.
-Three predicates are introduced: \textsc{first}\([1]\), \textsc{last}\([2]\), and \textsc{next}\([2]\), which denote the first and last statement of an item and the following statement to another using their statement IDs.\\
+Three predicates are introduced: \textsc{first}with arity 1, \textsc{last}with arity 2, and \textsc{next}with arity 2, which denote the first and last statement of an item and the following statement to another using their statement IDs.\\
 
@@
 To assign them a unique identity, statement ID, property, and value have to be used.
-For this the predicates \textsc{first\_qualifier}\([3]\) and \textsc{last\_qualifier}\([3]\) taking the statement ID, the property, and the value, as well as \textsc{next\_qualifier}\([6]\) with ID, property, and value for the current and next qualifier are introduced.\\
+For this the predicates \textsc{first\_qualifier} with arity 3 and \textsc{last\_qualifier} with arity 3 taking the statement ID, the property, and the value, as well as \textsc{next\_qualifier}with arity 6 with ID, property, and value for the current and next qualifier are introduced.\\
 
 Using this pre-processed data, the approach is to apply a not\_found-predicate to each statement or qualifier.
-This is \textsc{not\_found}\([2]\) for statements, connecting the statement ID with an identifier unique to the declared constraint and \textsc{not\_found}\([4]\) taking the statement ID, property, and value of a qualifier and linking this with the unique identifier.\\
+This is \textsc{not\_found}with arity 2 for statements, connecting the statement ID with an identifier unique to the declared constraint and \textsc{not\_found}with arity 4 taking the statement ID, property, and value of a qualifier and linking this with the unique identifier.\\
 
@@
 \end{equation}
-The predicate \textsc{require\_second}\([2]\) is introduced with the same meaning as require except that it will be applied after a required property has been found.
+The predicate \textsc{require\_second}with arity 2 is introduced with the same meaning as require except that it will be applied after a required property has been found.
 A violation would be any last statement requiring a second constrained statement, analogous to the violation rule for \emph{item requires statement} (see Section \ref{subsec_item_requires_statement}).

### 2018-09-22 thesis/paper/paper.tex

@@
 that the value of a statement with this property should have certain statements.
-Table \ref{tab_property_constraints} contains the existing property constraints as of September 2018.
-The constraints will be explained in detail and their modelling given.
+Table \ref{tab_property_constraints} contains the existing constraint types as of September 2018.
+The constraint types will be explained in detail and their modelling given.
 For constraint that cannot be modelled, reasoning for this assessment is included.
@@
 Table \ref{tab_property_constraints} shows all constraints that are used as of September 2018.
-It is roughly sorted by how difficult it is to model the given constraint using logical rules, based on the required expressive features and pre-processing.
-These required features are conditions that cannot natively be expressed in datalog and would require some kind of pre-processing of the dataset.
+It contains the name of the constraint and its Q-ID, followed by the number of constraint declarations and the features required to model the constraint type.
+The table is roughly sorted by how difficult it is to model the given constraint using logical rules, based on the required expressive features and pre-processing.
+The horizontal lines group  These required features are conditions that cannot natively be expressed in datalog and would require some kind of pre-processing of the dataset.
 Four constraints do not need any pre-processing except for loading the data into datalog as facts.
@@
 \caption{property constraints}\label{tab_property_constraints}
-\begin{tabularx}{\textheight}{rlll}
-\makecell{\# constrained\\ properties} & Identifier & Name & Required Features \\
-\hline
-1410 & Q53869507 & Scope &  \\
-781 & Q21502838 & Conflicts with &  \\
-27 & Q52004125 & Allowed entity types &  \\
-17 & Q52558054 & None of &  \\
-\hline
-2843 & Q21502410 & Distinct value & Inequality \\
-343 & Q21514353 & Allowed units & Inequality \\
-303 & Q21510851 & Allowed qualifiers & Inequality \\
-116 & Q21510859 & One of & Inequality \\
-3 & Q52060874 & Single best value & Inequality \\
-1 & Q52712340 & One-of qualifier value & Inequality \\
-\hline
-4265 & Q21503247 & Item requires statement & Negation (statement non-existence) \\
-244 & Q21510864 & Value requires statement & Negation (statement non-existence) \\
-119 & Q21510855 & Inverse & Negation (statement non-existence) \\
-33 & Q21510862 & Symmetric & Negation (statement non-existence) \\
-13 & Q21510857 & Multi-value & Negation, Count (up to two) \\
-2909 & Q19474404 & Single value & Inequality, Negation (qualifier non-existence) \\
-203 & Q21510856 & Mandatory qualifier & Negation (qualifier non-existence) \\
-\hline
-2613 & Q21503250 & Type & Negation (derived fact negation) \\
-707 & Q21510865 & Value type & Negation (derived fact negation) \\
-3021 & Q21502404 & Format & Regular expressions \\
-188 & Q21510860 & Range & Value comparison \\
-106 & Q52848401 & Integer & Regular expressions \\
-65 & Q51723761 & No bounds & Negation \\
-53 & Q21510852 & Commons link & Web request \\
-10 & Q54554025 & Citation needed & Negation \\
-6 & Q21510854 & Difference within range & Value comparison \\
-1 & Q25796498 & Contemporary constraint & Negation \\
-\end{tabularx}
+\begin{tabular}{llrp{6.5cm}}
+Name & Identifier & \makecell{\# constrained \\ properties} & Required Features \\
+\hline
+Scope & Q53869507 & 1410 & \\
+Conflicts with & Q21502838 & 781 & \\
+Allowed entity types & Q52004125 & 27 & \\
+None of & Q52558054 & 17 & \\
+\hline
+Distinct value & Q21502410 & 2843 & Inequality \\
+Allowed units & Q21514353 & 343 & Inequality \\
+Allowed qualifiers & Q21510851 & 303 & Inequality \\
+One of & Q21510859 & 116 & Inequality \\
+Single best value & Q52060874 & 3 & Inequality \\
+One-of qualifier value & Q52712340 & 1 & Inequality \\
+\hline
+Item requires statement & Q21503247 & 4265 & Negation (statement non-existence) \\
+Value requires statement & Q21510864 & 244 & Negation (statement non-existence) \\
+Inverse & Q21510855 & 119 & Negation (statement non-existence) \\
+Symmetric & Q21510862 & 33 & Negation (statement non-existence) \\
+Multi-value & Q21510857 & 13 & Negation, Count (up to two)  \\
+Single value & Q19474404 & 2909 & Inequality, Negation (qualifier non-existence) \\
+Mandatory qualifier & Q21510856 & 203 & Negation (qualifier non-existence) \\
+\hline
+Type & Q21503250 & 2613 & Negation (derived fact negation) \\
+Value type & Q21510865 & 707 & Negation (derived fact negation) \\
+Format & Q21502404 & 3021 & Regular expressions \\
+Range & Q21510860 & 188 & Value comparison \\
+Integer & Q52848401 & 106 & Regular expressions \\
+No bounds & Q51723761 & 65 & Negation  \\
+Commons link & Q21510852 & 53 & Web request \\
+Citation needed & Q54554025 & 10 & Negation \\
+Difference within range & Q21510854 & 6 & Value comparison \\
+Contemporary constraint & Q25796498 & 1 & Negation  \\
+\end{tabular}
 \end{table}
@@
 \caption{\textsc{edb} predicates}\label{tab_predicates}
-\begin{tabular}{llp{8.5cm}}
+\begin{tabular}{llp{8cm}}
 Predicate & Positions & Description \\
@@
 These include the required data, relevant qualifiers, and possible violations of the constraint.
-The required data states which part of the data in relation to the constrained statement are restricted or need to exist.
+The required data states which facts need to be loaded.
 The qualifiers list their cardinality, meaning if they are required and if they may be used multiple times, and a short description.
@@
 {\bf Rules:} Most constraints require more data than just the statements, qualifiers, and references.
-This can include additional data from Wikidata to be provided as facts using the introduced predicates (see Table \ref{tab_predicates}), inequalities to be established as a pre-processing step (see Section \ref{sec_constraints_using_inequality}), or preparations to determine specific kinds of negation (see Section \ref{sec_constraints_using_statement_non}).
+This can include inequalities to be established as a pre-processing step (see Section \ref{sec_constraints_using_inequality}) or preparations to determine specific kinds of negation (see Section \ref{sec_constraints_using_statement_non}).
 Afterwards, the rules necessary to determine violations are described.
@@
 \hline
-Required data & constrained statement, qualifier and reference \\
+Required data & statements, qualifiers, and references\\
 \hline
@@
 \hline
-Required data & entity statements \\
+Required data & statements \\
 \hline
@@
 \hline
-Required data & constrained statements and entity types\\
+Required data & statements, entity type (item or property)\\
 \hline
@@
 \hline
-Required data & constrained statements, qualifiers and references \\
+Required data & statements, qualifiers, and references \\
 \hline
@@
 \subsubsection{Encoding constant strings to establish inequality}\label{subsubsec_encoding}
-One way of determining inequality would be to take advantage of the fact that inequality is established in a set of unequal constants \(U\), which can be represented as a string.
-The idea is to determine the set of characters used in \(U\), set these characters as unequal as outlined in Section \ref{subsubsec_encoding} and associate each unequal constant with their characters.\\
-
-An unequal constant \(u \in U\) can be represented a string and be written as \(c_0 \ldots c_{l - 1}\) with \(l = len(u)\) being the length of the string.
-Denote the set \(C\) as the set of characters occurring in unequal strings.
+One way of determining inequality would be to take advantage of the fact that inequality is established in a set of constants \(U\), which can be represented as a string.
+The idea is to determine the set of characters used in \(U\), set these characters as unequal as outlined in Section \ref{subsubsec_encoding} and associate each unequal constant with their characters.
+One possibility would be to create one predicate to associate all characters of a string with said string.
+This predicates arity would likely be very large and thus not efficient to handle.
+If each position was described with its own binary predicate, this could be avoided.
+It would however require the string to be saved to disk repeatedly, once for each position in the string.\\
+
+The approach chosen here thus splits the constant string into chunks of length \(c\) and saves these with one predicate per chunk.
+Varying the chunk length allows balancing the disk space requirements with potential performance increases gained through shorter predicates.
+
+A constant \(u \in U\) can be represented a string and be written as \(c_0 \ldots c_{l - 1}\) with \(l = len(u)\) being the length of the string.
+Denote the set \(C\) as the set of characters occurring in strings.
 Then \(l_{\max} = \max\limits_{\forall u \in U}(len(u))\) is the length of the longest entity string that occurs in the given data.
-Based on this a group of predicates letter\(_i\)\_edb[8] is introduced where \(0 \equiv i \mod 7\), \(0 \leq i < l_{max} \).
-
-For each unequal constant string \(u = c_0 \ldots c_{l - 1}\) and \(0 \equiv i \mod 7, 0 \leq i < len(u)\) an atom using predicate \emph{letter\(_i\)\_edb} is created as 
-\begin{align}
-\begin{split}
-\emph{letter}_i\emph{\_edb}(e, f(e, i), f(e, i+1), \ldots, f(e, i+6)
+Based on this a group of predicates letter\(_i\)\_edb[c+1] is introduced where \(0 \equiv i \mod c\), \(0 \leq i < l_{max} \).
+
+For each constant string \(u = c_0 \ldots c_{l - 1}\) and \(0 \equiv i \mod c, 0 \leq i < len(u)\) an atom using predicate \emph{letter\(_i\)\_edb} is created as 
+\begin{align}
+\begin{split}
+\emph{letter}_i\emph{\_edb}(e, f(e, i), f(e, i+1), \ldots, f(e, i+c-1)
 \end{split}
@@
 
-Now that inequality between all characters and the link between entities and their characters are introduced, the inequality between two entities can be derived by creating the following rules for all predicates \emph{letter\(_i\)\_edb} (with \emph{letters} short for \(\emph{letter}_i\emph{\_edb}(X, A_i, \ldots, A_{i+6}) \wedge\emph{letter}_i\emph{\_edb}(Y, B_i, \ldots, B_{i+6})\), \(i \leq j < i + 7\)):
+Now that inequality between all characters and the link between entities and their characters are introduced, the inequality between two entities can be derived by creating the following rules for all predicates \emph{letter\(_i\)\_edb} (with \emph{letters} short for \(\emph{letter}_i\emph{\_edb}(X, A_i, \ldots, A_{i+6}) \wedge\emph{letter}_i\emph{\_edb}(Y, B_i, \ldots, B_{i+c-1})\), \(i \leq j < i + c\)):
 \begin{align}
@@
 \end{align}
-Note that the facts fulfilling \(\textsc{unequal}(A_j, B_j)\) are created by using the approach outlined in Section \ref{subsubsec_naive} for the set of characters \(C\) in all unequal constants \(U\).
+Note that the facts fulfilling \(\textsc{unequal}(A_j, B_j)\) are created by using the approach outlined in Section \ref{subsubsec_naive} for the set of characters \(C\) in all constants \(U\).
+
+Another possibility to apply this approach would be to encode inequality for the string hashes instead of the strings themselves.
+The string and its hash would then be associated via another predicate.
+Since generating these hashes is no trivial task and the additional step from constant to hash would complicate the rules presented below, this was not done here.
 
@@
 \hline
-Required data & constrained statements \\
+Required data & \makecell{statements; \\
+inequality for statement IDs} \\
 \hline
@@
 \hline
-Required data & constrained statements \\
-\hline
-Qualifiers & \makecell{\emph{allowed unit} (P2305) -- 1..* \\ units that are allowed on values of constrained statements (\emph{none} for dimensionless units)} \\
+Required data & \makecell{statements, qualifiers, and references; \\
+inequality for units} \\
+\hline
+Qualifiers & \makecell{\emph{allowed unit} (P2305) -- 1..* \\ units that are allowed on values of constrained statements \\ (\emph{none} for dimensionless units)} \\
 \hline
@@
 \hline
-Required data & qualifiers on constrained statements \\
+Required data & \makecell{statements and qualifiers; \\
+inequality for qualifier properties} \\
 \hline
@@
 \hline
-Required data & constrained statements, qualifiers, and references \\
+Required data & \makecell{statements, qualifiers, and references; \\
+inequality for values of statements, qualifiers, and references} \\
 \hline
@@
 \textsc{violation\_statement}&(S, E, \emph{constrainedProperty}, V) \leftarrow \\
-&\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, V) {\textstyle\bigwedge}_{a \in A} \textsc{unequal}(a, V)
+&\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, V) \\
+&{\textstyle\bigwedge}_{a \in A} \textsc{unequal}(a, V)
 \end{split}\\
@@
 \textsc{violation\_qualifier}&(S, \emph{constrainedProperty}, V) \leftarrow \\
-&\textsc{qualifier\_edb}(S, \emph{constrainedProperty}, V) {\textstyle\bigwedge}_{a \in A} \textsc{unequal}(a, V)
+&\textsc{qualifier\_edb}(S, \emph{constrainedProperty}, V) \\
+&{\textstyle\bigwedge}_{a \in A} \textsc{unequal}(a, V)
 \end{split}\\
@@
 \textsc{violation\_reference}&(S, \emph{constrainedProperty}, V) \leftarrow \\
-&\textsc{reference\_edb}(S, H, \emph{constrainedProperty}, V) {\textstyle\bigwedge}_{a \in A} \textsc{unequal}(a, V)
+&\textsc{reference\_edb}(S, H, \emph{constrainedProperty}, V) \\
+&{\textstyle\bigwedge}_{a \in A} \textsc{unequal}(a, V)
 \end{split}
@@
 \hline
-Required data & constrained statements and their qualifiers \\
+Required data & \makecell{statements and qualifiers; \\
+inequality for values of qualifiers} \\
 \hline
@@
 
-\section{Constraints using Statement non-Existence}\label{sec_constraints_using_statement_non}
+\section{Constraints Using Statement Non-Existence}\label{sec_constraints_using_statement_non}
 Several constraints are violated if specific statements do not exist in relation to constrained statements.
@@
 \hline
-Required data & item statements \\
+Required data & \makecell{statements; \\
+inequality for statement properties and values; \\
+statement order} \\
 \hline
@@
 \hline
-Required data & item statements \\
+Required data & \makecell{statements; \\
+inequality for statement properties; \\
+statement order} \\
 \hline
@@
 \hline
-Required data & constrained statements and their qualifiers, constrained qualifiers and references \\
+Required data & \makecell{statements, qualifiers, and references; \\
+inequality for statement IDs and qualifier values; \\
+qualifier order} \\
 \hline
@@
 \hline
-Required data & constrained statements and their qualifiers \\
+Required data & \makecell{statements and qualifiers; \\
+inequality between qualifier properties; \\
+qualifier order} \\
 \hline
@@
 \hline
-Required data & item statements and ranks \\
+Required data & \makecell{statements, qualifiers, references, and ranks; \\
+inequality for statement IDs and qualifier values; \\
+qualifier order} \\
 \hline
@@
 These take the form of comma-separated value files, one for each predicate in Section \ref{subsec_representation}, denoted as "Facts" in the diagram.
-
-At the same time the Inequaliy Helper, receiving the Facts from WDTK, creates all files necessary to establish inequality as described in Section \ref{sec_constraints_using_inequality}, based on which approach was chosen.
+The statement and qualifier orders are generated as well.
+These consist of three files each for the first, next, and last predicate.
+
+At the same time the Inequality Helper, receiving the Facts from WDTK, creates all files necessary to establish inequality as described in Section \ref{sec_constraints_using_inequality}, based on which approach was chosen.
 These files are constraint type specific, whereas the "Facts" are not.
 
-The rules determining violations are created by the Constraint Checker, which uses the Wikidata Query Service (WDQS)\footnote{\url{query.wikidata.org}} to fetch the current constraint configurations, which are the set constraints and their qualifiers as described in Chapter \ref{cha_property_constraints}.
-Based on these configurations the rules are created as described in Chapter \ref{cha_wikidata_constraints_explained}.
+The rules determining violations are created by the Constraint Checker, which uses the Wikidata Query Service (WDQS)\footnote{\url{query.wikidata.org}} to fetch the constraint declarations.
+Based on these declarations the rules are created as described in Chapter \ref{cha_wikidata_constraints_explained}.
 
@@
 All facts pertaining to the constraint, as well as the rules, are loaded into VLog, which creates an in-memory database from the provided .csv-files.
-For example, Once this is done, the reasoning engine materializes all possible facts.
+Once this is done, the reasoning engine materializes all possible facts.
 Afterwards, all violations are extracted by querying the database for the queries described in Section \ref{subsec_task}.
@@
 
-{\bf Runtime materialization} is the time taken by the VLog engine to compute all facts that can be derived from the given input.
-
-{\bf Additional Derivations} counts how many new facts were created during materialization, excluding the violations themselves.
+{\bf Total time} is the time taken to evaluate one constraint type on the given data.
+The preparation of the facts from the JSON-Dump is not reported in this figure, since it is constraint type independent and not executed by VLog.
+
+{\bf Loading time} is the time taken to load the data from the .csv-files into memory by VLog. It is expected that this correlates with the number of input facts.
+
+{\bf Reasoning time}  is the time taken to reason on the given facts and materialize all possible conclusions.
+
+{\bf Querying time} is the time taken to query for all violations and count them.
+For this experiment the violations where not written to a file or the console.
+In a production setup the querying time would be higher since it would depend on disk I/O.
 
@@
 Most constraints could be evaluated in under a minute.
-The exceptions are \emph{distinct values} (1.22 hours), \emph{item requires statement} (2.38 hours) and \emph{single value} (4.22 hours).
+The exceptions are \emph{distinct values} (1.26 hours), \emph{item requires statement} (2.27 hours) and \emph{single value} (4.22 hours).
 In part this is expected behaviour, since these three constraints are applied to an order of magnitude more properties than all other properties with the exception of \emph{property scope}, which is described in very simple rules.
@@
 The following Sections will thus focus on the comparison of these three constraints to identify bottlenecks and reasons for this huge difference in performance.
-They will be marked with a border in every figure.
 
 \begin{figure}
-\includegraphics[width=\linewidth]{images/runtime100,000.pdf}
-\caption{Runtime materialization for 100,000 entities}\label{fig_runtime}
+\includegraphics[width=\linewidth]{images/TotalTime.pdf}
+\caption{Total time for each constraint type}\label{fig_runtime}
 \end{figure}
 
+\subsection{Inequality evaluation}\label{sec_inequality_evaluation}
+Inequality computation was implemented in three variants outlined in Section \ref{sec_constraints_using_inequality}: naive, encoded, and on-demand.
+Naive and encoded could not terminate for any dataset other than the smallest with 1,000 entities.
+Naive failed because the disk space needed for the precomputed inequality facts was not available.
+Encoded, while improving on the disk space needed, still failed due to out of memory errors.
+This behaviour is expected, since for an input of size \(n\), the number of inequalities that have to be materialized is \(n^2\).
+The reasoning times for this dataset, presented in Figure \ref{fig_inequality}, show a clear improvement when using the on-demand approach.
+This shows that the overhead in rules introduced to enable the on-demand approach, seen in Figure \ref{fig_rules}, is worth the improvements in performance gained through it.
+
 \begin{figure}
-\includegraphics[width=\linewidth]{images/rules.pdf}
-\caption{Number of rules and on-demand rules}\label{fig_rules}
+\includegraphics[width=\linewidth]{images/inequality1000.pdf}
+\caption{Reasoning time for 1,000 entities (encoded and on-demand)}\label{fig_inequality}
 \end{figure}
+
+\begin{figure}
+\includegraphics[width=\linewidth]{images/Rules.pdf}
+\caption{Rules per constraint type}\label{fig_inequality}
+\end{figure}
+
+\subsection{Runtime}
+Runtime was dominated by reasoning time, which took close to 8 hours for all constraint types combined.
+Data extraction as well as loading time and querying time for all constraint types took less than 5 minutes each.
+
+Figure \ref{fig_loading_time} shows the loading times as bars, aligned with the left axis, and the number of input facts as diamonds, aligned with the right axis.
+As expected, the loading times are directly correlated with the number of input facts.
+
+\begin{figure}
+\includegraphics[width=\linewidth]{images/LoadingTime.pdf}
+\caption{Loading time and input size for each constraint type}\label{fig_loading_time}
+\end{figure}
+
+Figure \ref{fig_querying_time} shows the querying times as bars, aligned with the left axis, and the number of violations as well as total derivations, aligned with the right axis.
+While the querying time roughly aligns with the number of violations, with the constraints after \emph{allowed qualifiers} having higher querying times than expected from the number of violations.
+This is likely because of additional facts for inequality and not-found-pattern that were materialized putting a greater strain on the system.
+
+\begin{figure}
+\includegraphics[width=\linewidth]{images/QueryingTime.pdf}
+\caption{Querying time, violations and derivations for each constraint type}\label{fig_querying_time}
+\end{figure}
+
+\subsubsection{Reasoning time}
 
@@
 \begin{figure}
-\includegraphics[width=\linewidth]{images/facts100,000.pdf}
+\includegraphics[width=\linewidth]{images/Facts.pdf}
 \caption{Number of input facts per constraint}\label{fig_facts}
@@
 
-\subsection{Inequality evaluation}\label{sec_inequality_evaluation}
-Inequality computation was implemented in three variants outlined in Section \ref{sec_constraints_using_inequality}: naive, encoded, and on-demand.
-Naive and encoded could not terminate for any dataset other than the smallest with 1,000 entities.
-Naive failed because the disk space needed for the precomputed inequality facts was not available.
-Encoded, while improving on the disk space needed, still failed due to out of memory errors.
-This behaviour is expected, since for an input of size \(n\), the number of inequalities that have to be materialized is \(n^2\).
-The runtimes for this dataset, presented in Figure \ref{fig_inequality}, show a clear improvement when using the on-demand approach.
-This shows that the overhead in rules introduced to enable the on-demand approach, seen in Figure \ref{fig_rules}, is worth the improvements in performance gained through it.
+\subsection{Derivations}
+One possible bottleneck are the derivations necessary to identify violations.
+This includes the inequalities, demands for inequalities, requirements to determine statement non-existence as well as other helper rules introduced in Chapter \ref{cha_wikidata_constraints_explained}.
+Figure \ref{fig_derivations} shows that the three performance-heavy constraint types do generate by far the most derivations.
+The performance-heavy constraint type with the smallest number of derivations, \emph{distinct values}, is almost an order of magnitude higher than the next highest, \emph{one-of}.
 
 \begin{figure}
-\includegraphics[width=\linewidth]{images/inequality1000.pdf}
-\caption{Runtime materialization for 1,000 entities (encoded and on-demand)}\label{fig_inequality}
-\end{figure}
-
-\subsection{Additional Derivations}
-One possible bottleneck are the additional derivations necessary to identify violations.
-This includes the inequalities, demands for inequalities, requirements to determine statement non-existence as well as other helper rules introduced in Chapter \ref{cha_wikidata_constraints_explained}.
-Note that the four simple constraints described in Section \ref{sec_simple_constraints} do not generate any additional facts and are excluded from the chart.
-Figure \ref{fig_add_derivations} shows that the three performance-heavy constraints do generate by far the most additional derivations.
-The constraint with the smallest number of additional derivations, \emph{distinct values}, is almost an order of magnitude higher than the next highest, \emph{one-of}.
-
-\begin{figure}
-\includegraphics[width=\linewidth]{images/additionalDerivations100,000.pdf}
-\caption{Additional derivations for 100,000 entities}\label{fig_add_derivations}
+\includegraphics[width=\linewidth]{images/Derivations.pdf}
+\caption{Derivations for each constraint type}\label{fig_derivations}
 \end{figure}
@@
 While the rules for the \emph{distinct values} constraint (see Section \ref{subsec_distinct_values}) are relatively simple in writing, they do require a large amount of processing.
-Distinct values requires the comparison of every constrained statement with every other constrained statement, meaning its evaluation would scale badly due to the quadratic growth of statements to be compared.
+Distinct values requires the comparison of every constrained statement with every other constrained statement, meaning its evaluation would scale badly due to the growth of statements to be compared.
 Additionally, the number of inequalities that have to be evaluated according to the on-demand approach (see Section \ref{sec_constraints_using_statement_non}) grows accordingly.
 Due to the nature of the constraint -- requiring distinct values for a property across all of Wikidata -- this is in all likelihood unavoidable.
+Nevertheless, the current approach could be improved upon.
+One possibility is to reorder the \textsc{statement\_edb}-predicate to have the value in the first position.
+This could allow a quicker join of constrained statements with the same value, reducing the reasoning time.
+The downside is having to recreate the \textsc{statement\_edb} facts for this constraint type.
+This is likely to be offset by even small improvements in reasoning time, since the data extraction only took 5 minutes, compared to the 1.26 hours of reasoning time for \emph{distinct values}.
 
@@
 \emph{Value requires statement} has, however, only 244 constrained properties, while \emph{item requires statement} has 4265.
-This shows that the require approach scales badly with increasing number of require rules.
+This shows that the not-found-pattern scales badly with increasing number of require rules.
 
@@
 
-\subsubsection{Results}
+Figure \ref{fig_ps_comparision} shows the difference in runtime between the normal approach and the approach with predicate sorting.
+Predicate sorting does not seem to have any effect on the total runtimes, with differences both up and down and generally very close to the normal approach.
+
+\begin{figure}
+\includegraphics[width=\linewidth]{images/PredicateSortingComparison.pdf}
+\caption{Comparison of total time with and without predicate sorting}\label{fig_ps_comparision}
+\end{figure}
 
